### 2016-09-13 analysis/pheno_predability.Rmd
Display predictive abilities with rmarkdown

---
title: "Evaluation of LOOCV predictive abilities"
output: 
  html_document:
    toc: true
    toc_float: true
    smooth_scroll: true
    toc_depth: 3
    number_sections: true
    theme: lumen
    highlight: tango
    fig_width: 8
    fig_height: 5
---

# Data preparation
Load packages
```{r message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "MASS")
```

Set knitr options
```{r message = FALSE}
knitr::opts_knit$set(root.dir = "~/gamazon/")
```

Load log files and data
```{r Load_Data, cache = TRUE}
run_log <- fread("./data/derived/pred_log.txt")

# LOOCV 
loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
  readRDS(paste0("./data/derived/predictions/", loocv_nms[i], ".RDS"))
})
loocv <- rbindlist(loocv)
```

Change trait names and compute residuals.
```{r Rename_Traits, cache = TRUE}
loocv_clean <- loocv %>%
  mutate(Trait = fct_recode(Trait, 
                            "FAT" = "FETT",
                            "PRO" = "RPR",
                            "DMY" = "GTM",
                            "DMC" = "GTS"),
         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
                                      "XZ")),

         Dent_NA_Fraction = as.factor(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction)),
         Residuals = y - yhat
  )
```

## Predictive ability for non-imputed data
Peak at predictive abilities based on non-imputed predictor information and rank
the results by their predictive ability.
```{r Non_Imputed_Pred_Ability, cache = TRUE}
loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  mutate(Predictor = as.factor(Predictor)) %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  arrange(desc(Pred_Ability))
```

# Plots
## Complete data (no ssBLUP)
### Predicted vs. residuals
```{r Complete_Data_Residuals, cache = TRUE}
full_pred_resid <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
       geom_point(size = 0.5) +
       geom_ref_line(h = 0) +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Predicted") +
       ylab("Residuals") +
       ggtitle(.$Predictor))
invisible(lapply(full_pred_resid$plot, FUN = print))
```

### Predicted vs. Observed
```{r Complete_Data_y_yHat, cache = TRUE}
full_y_yhat <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = y, y = yhat)) +
       geom_point(size = 0.5) +
       geom_abline(slope = 1, linetype = "dashed", color = "blue") +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Predicted") +
       ylab("Residuals") +
       ggtitle(.$Predictor))
invisible(lapply(full_y_yhat$plot, FUN = print))
```

## Imputed data
### Predicted vs. residuals
Generate residual plots for each trait and each combination of missing values 
in Dent (rows) and Flint (columns) parents.
```{r Imputed_Residual, cache = TRUE}
resid_plots <- loocv_clean %>%
  filter(Predictor == "mrna") %>%
  group_by(Trait) %>%
  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
            geom_point(size = 0.5) +
            geom_ref_line(h = 0) +
            facet_grid(Dent_NA_Fraction ~ Flint_NA_Fraction) +
            theme_light() +
            theme(axis.text.x = element_text(angle = 60, 
                                             hjust = 1, vjust = 1)) +
            xlab("Predicted") +
            ylab("Residuals") +
            ggtitle(.$Trait))
invisible(lapply(resid_plots$plot, print))
```

### Predicted vs. observed
Likewise, compare observed with predicted values for combinations that were used
in the previous plot.
```{r Imputed_y_yHat, cache = TRUE}
y_yhat_plots <- loocv_clean %>%
  filter(Predictor == "mrna") %>%
  group_by(Trait) %>%
  do(plot = ggplot(., aes(x = y, y = yhat)) +
            geom_point(size = 0.5) +
            geom_abline(slope = 1, linetype = "dashed", color = "blue") +
            facet_grid(Dent_NA_Fraction ~ Flint_NA_Fraction) +
            theme(axis.text.x = element_text(angle = 60,
                                             hjust = 1, vjust = 1)) +
            xlab("Observed") +
            ylab("Predicted") +
            ggtitle(.$Trait))
invisible(lapply(y_yhat_plots$plot, print))
```

## LOOCV predictive ability (DMY)
### mrna
```{r DMY, cache = TRUE}
dmy <- loocv_clean %>%
  mutate(Dent_NA_Fraction = as.numeric(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.numeric(as.character(Flint_NA_Fraction))) %>%
  filter(Trait == "DMY", Predictor == "mrna") %>%
  droplevels() %>%
  unite(NA_Fraction, Dent_NA_Fraction, Flint_NA_Fraction, sep = "_") %>%
  group_by(Predictor, Trait, NA_Fraction) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  ungroup() %>%
  arrange(desc(Pred_Ability)) %>%
  separate(NA_Fraction,
           into = c("Dent_NA_Fraction", "Flint_NA_Fraction"), 
           sep = "_") %>%
  mutate(Dent_NA_Fraction = as.numeric(Dent_NA_Fraction),
         Flint_NA_Fraction = as.numeric(Flint_NA_Fraction)) %>%
  rowwise() %>%
  mutate(Avg_NA = mean(c(Dent_NA_Fraction, Flint_NA_Fraction)))
```

### Average r as a function of global missingness and missingness in Flint
```{r DMY_Global_NA_Flint_NA, cache = TRUE}
dmy %>%
  mutate(Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction))) %>%
  mutate(Flint_NA_Fraction = fct_relevel(Flint_NA_Fraction, 
                                         "0", "0.1", "0.2", "0.3", "0.4",
                                         "0.5", "0.6")) %>%
  arrange(desc(Avg_NA)) %>%
  ggplot(aes(x = Avg_NA, y = Pred_Ability, color = Flint_NA_Fraction)) +
  geom_point(size = 2) +
  geom_line() +
  geom_smooth(method = "rlm", aes(group = 1)) +
  theme_light() +
  ylab("r")
```

### Average r as a function of missingness in Dent and missingness in Flint
```{r DMY_Dent_NA_Flint_NA, cache = TRUE}
dmy %>%
  mutate(Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction))) %>%
  mutate(Flint_NA_Fraction = fct_relevel(Flint_NA_Fraction, 
                                         "0", "0.1", "0.2", "0.3", "0.4",
                                         "0.5", "0.6")) %>%
  arrange(desc(Dent_NA_Fraction)) %>%
  ggplot(aes(x = Dent_NA_Fraction, y = Pred_Ability, 
             color = Flint_NA_Fraction)) +
  geom_point(size = 2) +
  geom_line() +
  geom_smooth(method = "rlm", aes(group = 1)) +
  theme_light() +
  ylab("r")
```

### Predictive ability as a function of missingness in Dent and missingness in 
### Flint for all traits
```{r All_Traits_Dent_NA_Flint_NA, cache = TRUE} 
loocv_clean %>%
  filter(Dent_NA_Fraction != 0, Flint_NA_Fraction != 0) %>%
  unite(NA_Fraction, Dent_NA_Fraction, Flint_NA_Fraction, sep = "_") %>%
  group_by(Trait, NA_Fraction) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  ungroup() %>%
  separate(NA_Fraction, into = c("Dent_NA_Fraction",
                                 "Flint_NA_Fraction"), sep = "_") %>%
  mutate(Dent_NA_Fraction = as.numeric(Dent_NA_Fraction)) %>%
  ggplot(aes(x = Dent_NA_Fraction, y = Pred_Ability, 
             color = Flint_NA_Fraction)) +
  geom_line() +
  facet_wrap(~ Trait, scales = "free_x") +
  theme_light() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "top") +
  ylab("r")
```

### 2016-09-13 analysis/pheno_predability.R

@@ -1,57 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis")
-run_log <- fread("./data/processed/prediction_log/log_list.txt")
-geno <- readRDS("./data/processed/common_genotypes.RDS")
-dent_imp_nms <- setdiff(geno$Dent$snp, geno$Dent$mrna)
-flint_imp_nms <- setdiff(geno$Flint$snp, geno$Flint$mrna)
-
-# CV1000
-cv1000_nms <- run_log[CV != "LOOCV", Job_ID, ]
-cv1000_lst <- lapply(seq_along(cv1000_nms), FUN = function(i) {
-  readRDS(paste0("./data/processed/predictions/", cv1000_nms[i], ".RDS"))
-})
-cv1000 <- rbindlist(cv1000_lst)
-cv1000_avg <- cv1000[Set == "T0", 
-                     .(Pred_Ability = cor(y, yHat)),
-                     by = .(Imputation, Trait)]
-cv1000_avg[, CV_Mode := "CV1000", ]
-
-
-# LOOCV 
-loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
-loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
-  readRDS(paste0("./data/processed/predictions/", loocv_nms[i], ".RDS"))
-})
-loocv <- rbindlist(loocv)
-loocv_avg <- loocv[, .(Pred_Ability = cor(y, yhat)), 
-                   by = .(Imputation, Trait)]
-loocv_avg[, CV_Mode := "LOOCV", ]
-
-# Concatenation and transformation
-cv_avg <- rbind(loocv_avg, cv1000_avg)
-cv_avg <- cv_avg %>%
-  mutate(Trait = fct_recode(Trait, 
-    "FAT" = "FETT",
-    "PRO" = "RPR",
-    "DMY" = "GTM",
-    "DMC" = "GTS"
-  )) %>%
-  mutate(Trait = fct_relevel(Trait,
-    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "XZ")
-  )) %>%
-  rename(ssBLUP = Imputation)
-
-# Plot
-g1 <- cv_avg %>%
-  ggplot(aes(x = Trait, y = Pred_Ability, fill = ssBLUP)) +
[…]

### 2016-09-14 analysis/prediction.R
Move report on Pred_Ability with NAs in parents

@@ -291,7 +291,8 @@ if (hypred_model %in% caret_mod_nms) {
     if (isTRUE(hypred_model == "gbm")) {
       tune_grid <- expand.grid(interaction.depth = seq(1, to = 7, by = 2),
                                n.trees = seq(100, to = 1000, by = 50),
-                               shrinkage = c(0.01, 0.1))
+                               shrinkage = c(0.01, 0.1),
+                               n.minobsinnode = 10)
       tuned_mod <- train(pred_mat, y,
                          method = "gbm",
                          tuneGrid = tune_grid,

### 2016-09-14 reports/agronomic_data.Rmd
Bundle reports in one gitbook

# Explore phenotypic data
Load packages
```{r packages, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "caret",
               "knitr")
```

Set knitr options
```{r knitr_options, message = FALSE}
knitr::opts_knit$set(root.dir = "~/gamazon/")
```

Load data
```{r Load_Data}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
```

## All hybrids
Prepare agronomic data.
```{r Prepare_agro}
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
```

Evaluate the skewness of agronomic traits.
```{r Full_Agro_Skewness}
all_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2)) %>%
  arrange(abs(Skewness)) %>%
  kable(padding = 0)
```

Evaluate the distribution of agronomic traits via kernel density plots.
```{r Full_Agro_Distribution}
all_pheno %>%
  ggplot(aes(x = Observed)) +
  geom_density() +
  facet_wrap(~ Trait, scales = "free") +
  theme_hc() +
  ylab("Density")
```

## Hybrids of parents with genomic and transcriptomic data 
bla bla

### 2016-09-14 reports/index.Rmd
Bundle reports in one gitbook

---
title: Single-step Hybrid Predition
author: Matthias Westhues, Claas Heuer
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::gitbook: default
documentclass: book
bibliography: ["library.bib", "cheuer.bib"]
biblio-style: apalike
link-citations: yes
---

# Preface {-}
This book covers an extension of the methodology of single-step prediction 
&mdash; which was developed for the imputation of SNP data from pedigree 
information &mdash; to the imputation of quantitative predictor variables using
genomic information.

The companion R-package for this analysis is [sspredr](https://github.com/mwesthues/sspredr).
The development version can be installed from Github:
```{r Install_sspred, eval = FALSE}
devtools::install_github("mwesthues/sspredr")
```

### 2016-09-19 reports/agronomic_data.Rmd
Fix bug in class of resulting data.table

@@
                "viridis", "tidyr", "gridExtra", "modelr", "caret",
-               "knitr")
+               "knitr", "tibble")
 ```
@@
 
-## All hybrids
 Prepare agronomic data.
@@
     ))
+
+# Select hybrids for which both parents have genomic as well as transcriptomic
+# records.
+dent <- genos$Dent$mrna
+flint <- genos$Flint$mrna
+hybrid <- genos$Hybrid %>%
+  as_data_frame() %>%
+  separate(value, into = c("DF", "Dent", "Flint")) %>%
+  mutate(Avail_Dent = ifelse(Dent %in% dent, yes = 1, no = 0),
+         Avail_Flint = ifelse(Flint %in% flint, yes = 1, no = 0),
+         Tested_Parents = Avail_Dent + Avail_Flint) %>%
+  filter(Tested_Parents == 2) %>%
+  unite(Hybrid, DF, Dent, Flint) %>%
+  .$Hybrid
+
+# All hybrids
+all_pheno <- all_pheno %>%
+  mutate(Reduced_Data = "no")
+
+# Only hybrids for which parents have both, genomic and transcriptomic data
+reduced_pheno <- all_pheno %>%
+  mutate(Reduced_Data = ifelse(G %in% hybrid, yes = "yes", no = "no")) %>%
+  filter(Reduced_Data == "yes")
 ```
 
-Evaluate the skewness of agronomic traits.
-```{r Full_Agro_Skewness}
-all_pheno %>%
+## Skewness
+A general **rule of thumb** to consider is that skewed data whose ratio of the 
+highest value to the lowest value is greater than 20 have significant skewness.
+The formula for the skewness statistic is
+$$
+\text{skewness}  = \frac{\sum (x_i - \bar{x})^{3}}{(n - 1)v^{3/2}},
+$$
+where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ is the predictor
+variable, $n$ is the number of values and $\bar{x}$ is the sample mean of the 
+predictor.
+```{r Agro_Skewness}
+all_skewness <- all_pheno %>%
   group_by(Trait) %>%
   summarize(Skewness = e1071::skewness(Observed),
-            Skewness = round(Skewness, digits = 2)) %>%
-  arrange(abs(Skewness)) %>%
+            Skewness = round(Skewness, digits = 2))
+reduced_skewness <- reduced_pheno %>%
+  group_by(Trait) %>%
+  summarize(Skewness = e1071::skewness(Observed),
+            Skewness = round(Skewness, digits = 2))
+full_join(x = all_skewness, y = reduced_skewness,
+          by = "Trait",
+          suffix = c("_All", "_Reduced")) %>%
   kable(padding = 0)
@@
 
-Evaluate the distribution of agronomic traits via kernel density plots.
-```{r Full_Agro_Distribution}
-all_pheno %>%
-  ggplot(aes(x = Observed)) +
-  geom_density() +
+## Distribution
+```{r Full_Agro_Distribution, message = FALSE}
+ggplot(all_pheno, aes(x = Observed)) +
+  geom_histogram(fill = "grey", alpha = 0.5) +
+  geom_histogram(data = reduced_pheno, color = "black") +
   facet_wrap(~ Trait, scales = "free") +
-  theme_hc() +
-  ylab("Density")
+  theme_bw()
 ```
 
-## Hybrids of parents with genomic and transcriptomic data 
-bla bla
+## Correlation
+```{r Full_Agro_Correlation}
+par(mar = c(0, 2, 0, 0),
+    mfrow = c(1, 2))
+all_pheno %>%
+  select(-Reduced_Data) %>%
+  spread(key = Trait, value = Observed) %>%
+  as.data.frame() %>%
+  remove_rownames() %>%
+  column_to_rownames(var = "G") %>%
+  cor() %>%
+  corrplot::corrplot(col = viridis(256),
+                     method = "color",
+                     diag = FALSE,
+                     type = "lower",
+                     order = "hclust",
+                     tl.srt = 45,
+                     addCoef.col = "black",
+                     tl.col = "black",
+                     title = "All hybrids",
+                     mar = c(0, 1, 1, 0))
+reduced_pheno %>%
+  filter(Reduced_Data == "yes") %>%
+  select(-Reduced_Data) %>%
+  spread(key = Trait, value = Observed) %>%
+  as.data.frame() %>%
+  remove_rownames() %>%
+  column_to_rownames(var = "G") %>%
+  cor() %>%
+  corrplot::corrplot(col = viridis(256),
+                     method = "color",
+                     diag = FALSE,
+                     type = "lower",
+                     order = "hclust",
+                     tl.srt = 45,
+                     addCoef.col = "black",
+                     tl.col = "black",
+                     title = "Reduced hybrids",
+                     mar = c(0, 1, 1, 0))
+```

### 2016-09-19 analysis/prediction.R

@@ -30,13 +30,13 @@ if (isTRUE(interactive())) {
   Sys.setenv("ALL_PHENO" = "FALSE")
   Sys.setenv("TRAIT" = "GTM")
   # Number of iterations in BGLR()
-  Sys.setenv("ITER" = "0")
+  Sys.setenv("ITER" = "500")
   # Prediction model in BGLR()
-  Sys.setenv("MODEL" = "pls")
+  Sys.setenv("MODEL" = "BRR")
   # Algorithm to generate variance-covariance matrices.
   Sys.setenv("VCOV" = "RadenII")
-  Sys.setenv("PI" = "0")
-  Sys.setenv("PRIOR_PI_COUNT" = "0")
+  Sys.setenv("PI" = "0.5")
+  Sys.setenv("PRIOR_PI_COUNT" = "10")
   # 'mrna' or 'snp' 
   Sys.setenv("PREDICTOR" = "mrna")
 }
@@ -353,7 +353,8 @@ if (!hypred_model %in% caret_mod_nms) {
            PI = Pi,
            PriorPiCount = PriorPiCount,
            Predictor = predictor,
-           Elapsed_Time = elapsed_time)
+           Elapsed_Time = elapsed_time) %>%
+    as.data.table()
   
   log_file <- unique(res[, .(Job_ID, All_Pheno, Elapsed_Time, Trait, Iter, CV, 
                              Model, PI, PriorPiCount), ])

### 2016-09-19 moab_scripts/prediction.sh

@@ -8,7 +8,7 @@
 #MOAB -l nodes=1:ppn=16
 #
 # Request memory per process
-#MOAB -l pmem=3000mb
+#MOAB -l pmem=3800mb
 #
 # Estimated wallclock time for job
 #MOAB -l walltime=03:00:00:00

### 2016-09-19 reports/agronomic_data.Rmd
Do not display code chunks in report

@@
-# Explore phenotypic data
-Load packages
-```{r packages, message = FALSE}
+# Agronomic Data
+```{r echo = FALSE, message = FALSE}
 if (!require("pacman")) install.packages("pacman")
@@
                "knitr", "tibble")
+
+knit_hooks$set(inline = function(x) {
+  prettyNum(x, big.mark = ",", digits = 2)
+})
+knitr::opts_knit$set(root.dir = "~/gamazon/",
+                     inline = TRUE)
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
+                      warning = FALSE, echo = FALSE)
 ```
 
-Set knitr options
-```{r knitr_options, message = FALSE}
-knitr::opts_knit$set(root.dir = "~/gamazon/")
-```
+```{r echo = FALSE}
+# Load data
+pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
 
-Load data
-```{r Load_Data}
-pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
-```
-
-Prepare agronomic data.
-```{r Prepare_agro}
+# Prepare agronomic data.
 genos <- readRDS("./data/processed/common_genotypes.RDS")
@@
 predictor.
-```{r Agro_Skewness}
+```{r Agro-Skewness, echo = FALSE}
 all_skewness <- all_pheno %>%
@@
           by = "Trait",
-          suffix = c("_All", "_Reduced")) %>%
+          suffix = c(" (Full)", " (Reduced)")) %>%
   kable(padding = 0)
@@
 ## Distribution
-```{r Full_Agro_Distribution, message = FALSE}
+```{r Agro-Distribution, message = FALSE, echo = FALSE, fig.cap = "Histogram of the distribution of seven agronomic traits.
+Black bins show the distribution of the reduced set of hybrids whereas light grey bins show the distribtution of the full set of hybrids."}
 ggplot(all_pheno, aes(x = Observed)) +
@@
 ## Correlation
-```{r Full_Agro_Correlation}
-par(mar = c(0, 2, 0, 0),
+```{r Agro-Correlation, echo = FALSE, fig.cap = "Pairwise correlations among seven agronomic traits for the full set of hybrids and the reduced set of hybrids, respectively."}
+par(mar = c(0, 1, 0, 0),
     mfrow = c(1, 2))
@@
                      diag = FALSE,
-                     type = "lower",
+                     cl.pos = "b",
+                     tl.pos = "d",
+                     tl.srt = 60,
                      order = "hclust",
-                     tl.srt = 45,
                      addCoef.col = "black",
-                     tl.col = "black",
-                     title = "All hybrids",
+                     title = "Full",
                      mar = c(0, 1, 1, 0))
@@
                      diag = FALSE,
-                     type = "lower",
+                     cl.pos = "b",
+                     tl.pos = "d",
+                     tl.srt = 60,
                      order = "hclust",
-                     tl.srt = 45,
                      addCoef.col = "black",
-                     tl.col = "black",
-                     title = "Reduced hybrids",
+                     title = "Reduced",
                      mar = c(0, 1, 1, 0))

### 2016-09-19 reports/data_overview.Rmd
Do not display code chunks in report

# Data Overview
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "VennDiagram", "grid",
               "gridExtra", "purrr", "htmlTable")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
                      warning = FALSE, echo = FALSE)
```

```{r Load_Agro_Data, message = FALSE, echo = FALSE}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
genos <- readRDS("./data/processed/common_genotypes.RDS")

full_hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% full_hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
traits <- all_pheno %>%
  .$Trait %>%
  levels()

# Hybrids of parents with records for both endophenotypes
red_dent <- Reduce(intersect, genos[["Dent"]][c("snp", "mrna")])
red_flint <- Reduce(intersect, genos[["Flint"]][c("snp", "mrna")])
red_hybrid <- genos$Hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% red_dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% red_flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid
full_hybrid <- genos$Hybrid
genos$Hybrid <- NULL
genos$Hybrid$snp <- full_hybrid
genos$Hybrid$mrna <- red_hybrid
```

```{r Feature-Number, echo = FALSE}
snp <- readRDS("./data/processed/snp_mat.RDS")
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
```

Agronomic data for `r length(traits)` traits (including dry-matter yield (DMY))
and `r length(full_hybrid)` hybrids are available 
(Table \@ref(tab:Geno-Number)).
After applying a series of quality checks, `r ncol(snp)` SNPs and 
best linear unbiased estimates of `r ncol(mrna)` were available for further 
analyses.
Both, genomic and transcriptomic records, were collected for inbred parent lines
of `r length(red_hybrid)` hybrids.

```{r Geno-Number, echo = FALSE}
geno_tab_cap <- paste("Number of genotypes for Dent and Flint parent lines and",
                      "their hybrid progeny for which genomic and/or",
                      "transcriptomic data are available.",
                      "All genotypes for which transcriptomic data are", 
                      "available, also have genomic data.",
                      "Agronomic data are available for all 1,521 hybrid",
                      "genotypes.")
# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
# parental Dent and Flint lines, respectively.
genos %>% 
  at_depth(.depth = 2, .f = length) %>%
  bind_rows() %>%
  select(-agro) %>%
  mutate(Type = names(genos)) %>%
  select(Type, snp, mrna) %>%
  rename("Genomic Data" = snp,
         "Transcriptomic Data" = mrna) %>%
  knitr::kable(caption = geno_tab_cap, booktabs = TRUE)
```

### 2016-09-19 reports/index.Rmd
Do not display code chunks in report

@@
 
+Here, we are working with a "full" or "complete" data set &mdash; referring to 
+the maximum number of hybrids for which agronomic and genomic, but not 
+necessarily transcriptomic data are available &mdash; and a "reduced" data set
+with less hybrids but all of their parental inbred lines having complete
+genomic and transcriptomic records.
+
+All our predictive abilities were evaluated using a leave-one-out cross 
+validation (LOOCV) routine.
+
 The companion R-package for this analysis is [sspredr](https://github.com/mwesthues/sspredr).
@@
 ```
+
+## Overview
+Chapter \@ref(introduction) gives a summary of the relevant literature.
+
+Chapter \@ref(data-overview) provides a brief summary of the available data 
+without delving into their properties.
+
+Chapter \@ref(agronomic-data) gives an overview of the distribution of the 
+agronomic traits as well as correlations among them, separately for the full
+set of hybrids as well as the reduced set of hybrids.
+
+Chapter \@ref(prediction-for-reduced-data) deals with the results of hybrid 
+prediction for the reduced set of data.
+
+Chapter \@ref(evaluation-of-loocv-predictive-abilities) examines the influence
+of certain levels of missingness in the transcriptomic data.

### 2016-09-19 reports/intro.Rmd
Do not display code chunks in report

---
output: html_document
---
# Introduction
The integration of different sources of information for the prediction of 
breeding or production values is one of the major challenges in current genomic 
prediction research.
In the early years of genomic prediction, a 2-step procedure was the most 
intuitive choice for establishing a new selection scheme.
The first consists of conventionally estimating breeding values based on 
pedigree relationships and therefore the phenotypes of relatives.
Individuals with accurate breeding values out of the first step (i.e. high 
number of offspring), enter the second step in which SNP genotypes are regressed 
against the conventional breeding values.
Those markers can then be used to predict genomic breeding values for solely 
genotyped and potentially very young individuals.
Considering the two sources of information, pedigree and high density SNP 
genotypes, the process can be described as follows:

### First Step
Denote $\mathbf{y}$ as the vector of measured phenotypes for a normally 
distributed trait for all individuals available in the population.
The relationship, and therefor the strucure of the additive genetic covariance, 
across individuals is based on the numerator relationship matrix $\mathbf{A}$.
In order to obtain breeding values for any individual in the population, we 
consider the following model:

\begin{equation}
	\mathbf{y} = \mathbf{Xb} + \mathbf{Za} + \mathbf{e},
	\label{eq:animal_model}
\end{equation}

where $\mathbf{X}$ is a design matrix for fixed effects, $\mathbf{b}$ a vector 
of fixed effects, $\mathbf{Z}$ is the design matrix for the random animal 
effects, which related observations in $\mathbf{y}$ to elements in $\mathbf{A}$.
$\mathbf{a}$ is the vector of random animal effects (=breeding values) and
$\mathbf{e}$ is a vector of iid residual errors wich expectation $0$ and 
variance $\sigma^2_e$.
The distribution of the random vector $\mathbf{u}$ is 
$MVN(0, \mathbf{A}\sigma^2_a)$.

### Second Step
In the second step, a subset of the individuals for which conventional breeding 
values have been estimated, is selected and enters the reference population for 
training a genomic prediction model.
That selection includes as necessary condition, that the individual in genotyped 
and has $\text{PEV}(u_{\text{candidate}}) \leq \text{threshold}$.
As the breeding values from step one enter the genomic prediction model as 
phenotypes, one has to be careful about distribution of the residual error of 
such a model.
For instance, if the range of the prediction error variances is very broud in 
the reference population, the residual errors will not be identically 
distributed anymore, e.g. they should be weighted differently or estimated 
individually.
One popular approach to account for that problem is *Deregressing* the breeding 
values and weighting the residuals according to the prediction error variances 
[@Garrick2009].
As the breeding values from step one enter the genomic prediction model as 
phenotypes, one has to be careful about distribution of the residual error of 
such a model.
For instance, if the range of the prediction error variances is very broud in 
the reference population, the residual errors will not be identically 
distributed anymore, e.g. they should be weighted differently or estimated 
individually.
One popular approach to account for that problem is *Deregressing* the breeding 
values and weighting the residuals according to the prediction error variances 
[@Garrick2009].

\begin{equation}
	\label{eq:gblup}
	\mathbf{\hat{a}} = \mathbf{1'b} + \mathbf{Mu} + \mathbf{e},
\end{equation}

where $\mathbf{M}$ is the matrix of marker covariates and $\mathbf{u}$ is a 
vector of marker effects with iid normal distribution.

Genomic breeding values for any animal in the population can then be obtained 
by:

\begin{equation}
	\label{eq:gebv}
	\mathbf{\hat{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
\end{equation}

where $\mathbf{p}$ is a row-vector of allele frequencies for all the markers in 
$\mathbf{M}$ in the base population.
The prediction error variance of the vector $\mathbf{\hat{g}}$ is usually 
globally obtained using cross validation.
Alternatively they could be obtained individually by the posterior distribution 
of $(\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}}$ when a Bayesian model is used.

## Single Step Transciptomic prediction of Breeding Values
The breeding values in a mRNA prediction model are in general given by:
\begin{equation}
	\label{eq:mrnaebv}
	\mathbf{\hat{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model 
equaivalent to equation \@ref(eq:gblup) but using MRNA covariates instead of 
marker genotypes.

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in [@fernando_class_2014] and impute missing
mRNA data.
The difference here is, that instead of predicting SNP covariates using 
pedigree relationship, we impute MRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data is available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual:
\begin{align}
	\label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The distribution of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$.

Using those results from [@legarra_relationship_2009] we can model the breeding 
values of the individuals for wich no MRNA information is available as expected 
values from the breeding values in group $2$.
The structure of the residual imputation/prediction error is known and can 
therefore be modelled.
The missing MRNA covariates can be predicted using the expectation of a 
multivariate normal random vector given correlated observations, which related 
to the Best Linear Predictor:

### 2016-09-19 reports/loocv_reduced.Rmd
Do not display code chunks in report

# Prediction For Reduced Data
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "MASS")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE,
                     verbose = FALSE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
                      warning = FALSE, echo = FALSE)
```

```{r Load-LOOCV-Data, cache = TRUE}
# Load log files and data
run_log <- fread("./data/derived/na_fraction_pred_log.txt")

# LOOCV 
loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
  readRDS(paste0("./data/derived/predictions/", loocv_nms[i], ".RDS"))
})
loocv <- rbindlist(loocv)
```

```{r Rename-Traits, cache = TRUE}
loocv_clean <- loocv %>%
  mutate(Trait = fct_recode(Trait, 
                            "FAT" = "FETT",
                            "PRO" = "RPR",
                            "DMY" = "GTM",
                            "DMC" = "GTS",
                            "SUG" = "XZ"),
         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
                                      "SUG")),
         Dent_NA_Fraction = as.factor(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction)),
         Residuals = y - yhat,
         Predictor = fct_recode(Predictor,
                                "mRNA" = "mrna",
                                "SNP" = "snp")
  )
```

## Predictive ability for non-imputed data
Peak at predictive abilities based on non-imputed predictor information and rank
the results by their predictive ability.
```{r Non-Imputed-Pred-Ability, cache = TRUE}
loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  mutate(Predictor = as.factor(Predictor)) %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  arrange(desc(Pred_Ability)) %>%
  rename("Predictive Ability" = Pred_Ability) %>%
  knitr::kable()
```

## Predicted vs. residuals
```{r Complete-Data-Residuals, cache = TRUE, dependson = "Rename-Traits"}
full_pred_resid <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
       geom_point(size = 0.5) +
       geom_ref_line(h = 0) +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Predicted") +
       ylab("Residuals") +
       ggtitle(.$Predictor))
invisible(lapply(full_pred_resid$plot, FUN = print))
```

## Predicted vs. Observed
```{r Complete-Data-y-yHat, cache = TRUE, dependson = "Rename-Traits"}
full_y_yhat <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = y, y = yhat)) +
       geom_point(size = 0.5) +
       geom_abline(slope = 1, linetype = "dashed", color = "blue") +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Observed") +
       ylab("Predicted") +
       ggtitle(.$Predictor))
invisible(lapply(full_y_yhat$plot, FUN = print))
```

### 2016-09-19 reports/na_fraction_pheno_predability.Rmd
Move report on Pred_Ability with NAs in parents
Bundle reports in one gitbook
Do not display code chunks in report

--- 2016-09-13 analysis/pheno_predability.Rmd
@@
----
-title: "Evaluation of LOOCV predictive abilities"
-output: 
-  html_document:
-    toc: true
-    toc_float: true
-    smooth_scroll: true
-    toc_depth: 3
-    number_sections: true
-    theme: lumen
-    highlight: tango
-    fig_width: 8
-    fig_height: 5
----
-
-# Data preparation
-Load packages
+# Evaluation of LOOCV predictive abilities
 ```{r message = FALSE}
@@
                "viridis", "tidyr", "gridExtra", "modelr", "MASS")
+knit_hooks$set(inline = function(x) {
+  prettyNum(x, big.mark = ",", digits = 2)
+})
+knitr::opts_knit$set(root.dir = "~/gamazon/",
+                     inline = TRUE)
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
+                      warning = FALSE, echo = FALSE)
 ```
 
-Set knitr options
-```{r message = FALSE}
-knitr::opts_knit$set(root.dir = "~/gamazon/")
-```
-
-Load log files and data
-```{r Load_Data, cache = TRUE}
-run_log <- fread("./data/derived/pred_log.txt")
-
-# LOOCV 
-loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
-loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
-  readRDS(paste0("./data/derived/predictions/", loocv_nms[i], ".RDS"))
-})
-loocv <- rbindlist(loocv)
-```
-
-Change trait names and compute residuals.
-```{r Rename_Traits, cache = TRUE}
-loocv_clean <- loocv %>%
-  mutate(Trait = fct_recode(Trait, 
-                            "FAT" = "FETT",
-                            "PRO" = "RPR",
-                            "DMY" = "GTM",
-                            "DMC" = "GTS"),
-         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
-                                      "XZ")),
-
-         Dent_NA_Fraction = as.factor(as.character(Dent_NA_Fraction)),
-         Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction)),
-         Residuals = y - yhat
-  )
-```
-
-## Predictive ability for non-imputed data
-Peak at predictive abilities based on non-imputed predictor information and rank
-the results by their predictive ability.
-```{r Non_Imputed_Pred_Ability, cache = TRUE}
-loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  mutate(Predictor = as.factor(Predictor)) %>%
-  droplevels() %>%
-  group_by(Predictor, Trait) %>%
-  summarize(Pred_Ability = cor(y, yhat)) %>%
-  arrange(desc(Pred_Ability))
-```
-
-# Plots
-## Complete data (no ssBLUP)
-### Predicted vs. residuals
-```{r Complete_Data_Residuals, cache = TRUE}
-full_pred_resid <- loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  group_by(Predictor) %>%
-  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
-       geom_point(size = 0.5) +
-       geom_ref_line(h = 0) +
-       facet_wrap(~ Trait, nrow = 2, scales = "free") +
-       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
-       xlab("Predicted") +
-       ylab("Residuals") +
-       ggtitle(.$Predictor))
-invisible(lapply(full_pred_resid$plot, FUN = print))
-```
-
-### Predicted vs. Observed
-```{r Complete_Data_y_yHat, cache = TRUE}
-full_y_yhat <- loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  group_by(Predictor) %>%
-  do(plot = ggplot(., aes(x = y, y = yhat)) +
-       geom_point(size = 0.5) +
-       geom_abline(slope = 1, linetype = "dashed", color = "blue") +
-       facet_wrap(~ Trait, nrow = 2, scales = "free") +
-       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
-       xlab("Predicted") +
-       ylab("Residuals") +
-       ggtitle(.$Predictor))
-invisible(lapply(full_y_yhat$plot, FUN = print))
-```
-
-## Imputed data
 ### Predicted vs. residuals
@@
 in Dent (rows) and Flint (columns) parents.
-```{r Imputed_Residual, cache = TRUE}
+```{r Imputed-Residual, cache = TRUE, dependson = "Rename-Traits"}
 resid_plots <- loocv_clean %>%
-  filter(Predictor == "mrna") %>%
+  filter(Predictor == "mRNA") %>%
   group_by(Trait) %>%
@@
 in the previous plot.
-```{r Imputed_y_yHat, cache = TRUE}
+```{r Imputed-y-yHat, cache = TRUE, dependson = "Rename-Traits"}
 y_yhat_plots <- loocv_clean %>%
-  filter(Predictor == "mrna") %>%
+  filter(Predictor == "mRNA") %>%
   group_by(Trait) %>%
@@
 ### mrna
-```{r DMY, cache = TRUE}
+```{r DMY, cache = TRUE, dependson = "Rename-Traits"}
 dmy <- loocv_clean %>%
@@
          Flint_NA_Fraction = as.numeric(as.character(Flint_NA_Fraction))) %>%
-  filter(Trait == "DMY", Predictor == "mrna") %>%
+  filter(Trait == "DMY", Predictor == "mRNA") %>%
   droplevels() %>%
@@
 ### Average r as a function of global missingness and missingness in Flint
-```{r DMY_Global_NA_Flint_NA, cache = TRUE}
+```{r DMY_Global_NA_Flint_NA, cache = TRUE, dependson = "DMY"}
 dmy %>%
@@
 ### Average r as a function of missingness in Dent and missingness in Flint
-```{r DMY_Dent_NA_Flint_NA, cache = TRUE}
+```{r DMY_Dent_NA_Flint_NA, cache = TRUE, dependson = "DMY"}
 dmy %>%
@@
 ### Flint for all traits
-```{r All_Traits_Dent_NA_Flint_NA, cache = TRUE} 
+```{r All_Traits_Dent_NA_Flint_NA, cache = TRUE, dependson = "Rename-Trait"} 
 loocv_clean %>%

### 2016-09-20 reports/agronomic_data.Rmd
Update references in report

@@
 pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "gridExtra", "modelr", "caret",
-               "knitr", "tibble")
+               "viridis", "tidyr", "knitr", "tibble")
 
@@
                      inline = TRUE)
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                       warning = FALSE, echo = FALSE)

### 2016-09-20 reports/conclusions.Rmd
Update references in report

# Conclusions and Outlook
We have seen that

1.    for the 'reduced' set of hybrids, predictive abilities for the traits 
      'DMY', 'PRO' and 'FAT' were higher when predicted via mRNA compared to SNP
      data (Table \@ref(tab:Non-Imputed-Pred-Ability)).
2.    for the 'full' set of hybrids, predictive abilities based on single-step
      prediction with transcriptomic data still outperformed SNP data for the
      traits 'DMY' and 'PRO' (Table \@ref(tab:Imputed-Pred-Ability)).
2.    an exploration of the effect of different levels of missingness &mdash; 
      with respect to transcriptomic BLUEs &mdash; in the parents is 
      computationally prohibitive 
      (Chapter \@ref(prediction-with-imputed-mrnas)).

Currently, our primary goal is to demonstrate the utility of imputing missing 
data in a quantitative predictor (exemplified by transcriptomic data) and 
using the resulting feature matrix in single-step hybrid prediction.

### 2016-09-20 reports/data_overview.Rmd
Update references in report

@@
 pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "knitr", "tibble", "VennDiagram", "grid",
-               "gridExtra", "purrr", "htmlTable")
+               "viridis", "tidyr", "knitr", "tibble", "purrr")
 
@@
                      inline = TRUE)
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                       warning = FALSE, echo = FALSE)
@@
 
-```{r Load_Agro_Data, message = FALSE, echo = FALSE}
+```{r Load-Agro-Data, message = FALSE, cache = TRUE}
 pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
@@
   filter(G %in% full_hybrid) %>%
-  select(G, Observed, Trait) %>%
+  dplyr::select(G, Observed, Trait) %>%
   filter(Trait != "ADL") %>%
@@
   bind_rows() %>%
-  select(-agro) %>%
+  dplyr::select(-agro) %>%
   mutate(Type = names(genos)) %>%
-  select(Type, snp, mrna) %>%
+  dplyr::select(Type, snp, mrna) %>%
   rename("Genomic Data" = snp,

### 2016-09-20 reports/index.Rmd
Update references in report

--- 2016-09-14 reports/index.Rmd
@@
 output:
-  bookdown::gitbook: default
+  bookdown::gitbook:
 documentclass: book
-bibliography: ["library.bib", "cheuer.bib"]
+bibliography: ["cheuer.bib"]
 biblio-style: apalike
@@
 # Preface {-}
-This book covers an extension of the methodology of single-step prediction 
+This report covers an extension of the methodology of single-step prediction 
 &mdash; which was developed for the imputation of SNP data from pedigree 
@@
 genomic information.
+
+Here, we are working with a "full" data set &mdash; referring to the maximum 
+number of hybrids for which agronomic and genomic, but not necessarily 
+transcriptomic data are available &mdash; and a "reduced" data set with less 
+hybrids but all of their parental inbred lines having complete genomic and 
+transcriptomic records.
+
+All our predictive abilities were evaluated using a leave-one-out cross 
+validation (LOOCV) routine and results are displayed for T0-hybrids only.
+Note, that T0-hybrids are defined as hybrids having zero parents that have been
+previously tested in other hybrid combinations.
 
@@
 ```
+
+## Objectives
+1.    Assess whether transcriptomic data with missing records for some genotypes         can be imputed successfully based on genomic information to improve 
+      prediction of unseen hybrids.
+2.    Explore the utility of various machine learning algorithms for the 
+      prediction of unseen hybrids.^[This is motivated by the fact that the 
+      success of hybrid prediction is a strongly convex function of predictive 
+      ability.]
+
+## Overview
+Chapter \@ref(introduction) gives a summary of the relevant literature.
+
+Chapter \@ref(data-overview) provides a brief summary of the available data 
+without delving into their properties.
+
+Chapter \@ref(agronomic-data) gives an overview of the distribution of the 
+agronomic traits as well as correlations among them, separately for the full
+set of hybrids as well as the reduced set of hybrids.
+
+Chapter \@ref(prediction-for-reduced-data) deals with the results of hybrid 
+prediction for the reduced set of data.
+
+Chapter \@ref(prediction-with-imputed-mrnas) examines the influence of certain levels of missingness 
+in the transcriptomic data.
+
+Chapter \@ref(conclusions-and-outlook) provides a summary of the analyses and 
+provides an outlook for future analyses.

### 2016-09-20 reports/loocv_reduced.Rmd
Update references in report

@@
 pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "gridExtra", "modelr", "MASS")
+               "viridis", "tidyr", "modelr")
 
@@
                      verbose = FALSE)
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                       warning = FALSE, echo = FALSE)
@@
 ```{r Non-Imputed-Pred-Ability, cache = TRUE}
+loocv_tbl_cap <- paste("Predictive abilities for the 'reduced' set of hybrids",
+                       "using two different predictors and seven agronomic",
+                       "traits. Results are ranked by predictive ability.")
 loocv_clean %>%
@@
   rename("Predictive Ability" = Pred_Ability) %>%
-  knitr::kable()
+  knitr::kable(caption = loocv_tbl_cap, booktabs = TRUE)
 ```

### 2016-09-20 reports/single_step_pred.Rmd
Update references in report

---
title: "Single-Step Hybrid Prediction Report"
author: "Matthias Westhues, Claas Heuer"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    smooth_scroll: true
    toc_depth: 3
    number_sections: true
    theme: lumen
    highlight: tango
    fig_width: 8
    fig_height: 5
    fig_caption: true
bibliography: cheuer.bib
---

# Preface <a name="Preface"></a>
This report covers an extension of the methodology of single-step prediction 
&mdash; which was developed for the imputation of SNP data from pedigree 
information &mdash; to the imputation of quantitative predictor variables using
genomic information.

Here, we are working with a "full" data set &mdash; referring to the maximum 
number of hybrids for which agronomic and genomic, but not necessarily 
transcriptomic data are available &mdash; and a "reduced" data set with less 
hybrids but all of their parental inbred lines having complete genomic and 
transcriptomic records.

All our predictive abilities were evaluated using a leave-one-out cross 
validation (LOOCV) routine and results are displayed for T0-hybrids only.
Note, that T0-hybrids are defined as hybrids having zero parents that have been
previously tested in other hybrid combinations.

The companion R-package for this analysis is [sspredr](https://github.com/mwesthues/sspredr).
The development version can be installed from Github:
```{r Install_sspred, eval = FALSE}
devtools::install_github("mwesthues/sspredr")
```

## Objectives
1.    Assess whether transcriptomic data with missing records for some genotypes         can be imputed successfully based on genomic information to improve 
      prediction of unseen hybrids.
2.    Explore the utility of various machine learning algorithms for the 
      prediction of unseen hybrids.^[This is motivated by the fact that the 
      success of hybrid prediction is a strongly convex function of predictive 
      ability.]

## Overview
[Chapter 2](#Introduction) gives a summary of the relevant literature.

[Chapter 3](#Data-Overview) provides a brief summary of the available data 
without delving into their properties.

[Chapter 4](#Agronomic-Data) gives an overview of the distribution of the 
agronomic traits as well as correlations among them, separately for the full
set of hybrids as well as the reduced set of hybrids.

[Chapter 5](#Prediction-for-reduced-data) deals with the results of hybrid 
prediction for the reduced set of data.

[Chapter 6](#Pred-NA) examines the influence of certain levels of missingness 
in the transcriptomic data.

[Chapter 7](#Full_Prediction) reveals results for predictions using all hybrids.

[Chapter 8](#Conclusions) provides a summary of the analyses and provides an
outlook for future analyses.

# Introduction <a name="Introduction"></a>
The integration of different sources of information for the prediction of 
breeding or production values is one of the major challenges in current genomic 
prediction research.
In the early years of genomic prediction, a 2-step procedure was the most 
intuitive choice for establishing a new selection scheme.
The first consists of conventionally estimating breeding values based on 
pedigree relationships and therefore the phenotypes of relatives.
Individuals with accurate breeding values out of the first step (i.e. high 
number of offspring), enter the second step in which SNP genotypes are regressed 
against the conventional breeding values.
Those markers can then be used to predict genomic breeding values for solely 
genotyped and potentially very young individuals.
Considering the two sources of information &mdash; pedigree and high density SNP 
genotypes &mdash; the process can be described as follows:

### First Step
Denote $\mathbf{y}$ as the vector of measured phenotypes for a normally 
distributed trait for all individuals available in the population.
The relationship &mdash; and therefore the strucure of the additive genetic 
covariance &mdash; across individuals is based on the numerator relationship 
matrix $\mathbf{A}$.
In order to obtain breeding values for any individual in the population, we 
consider the following model:

$$
\begin{equation}
	\mathbf{y} = \mathbf{Xb} + \mathbf{Za} + \mathbf{e},
	\label{eq:animal_model}
\end{equation}
$$

where $\mathbf{X}$ is a design matrix for fixed effects, $\mathbf{b}$ is a 
vector of fixed effects, $\mathbf{Z}$ is the design matrix for the random animal 
effects, which associate observations in $\mathbf{y}$ with elements in 
$\mathbf{A}$, $\mathbf{a}$ is the vector of random animal effects (=breeding 
values) and $\mathbf{e}$ is a vector of iid residual errors with expectation 
$0$ and variance $\sigma^2_e$.
The distribution of the random vector $\mathbf{u}$ is 
$MVN(0, \mathbf{A}\sigma^2_a)$.

### Second Step
In the second step, a subset of the individuals for which conventional breeding 
values have been estimated, is selected and enters the reference population for 
training a genomic prediction model.
That selection requires that the individual is genotyped and has 
$\text{PEV}(u_{\text{candidate}}) \leq \text{threshold}$.
As the breeding values from step one enter the genomic prediction model as 
phenotypes, one has to be careful about the residual error distribution of
such a model.
For instance, if the range of the prediction error variances is very broad in 
the reference population, the residual errors will not be identically 
distributed anymore, e.g. they should be weighted differently or estimated 
individually.
One popular approach to account for that problem is *Deregressing* the breeding 
values and weighting the residuals according to the prediction error variances 
[@Garrick2009].

$$
\begin{equation}
	\label{eq:gblup}
	\mathbf{\hat{a}} = \mathbf{1'b} + \mathbf{Mu} + \mathbf{e},
\end{equation}
$$

where $\mathbf{M}$ is the matrix of marker covariates and $\mathbf{u}$ is a 
vector of marker effects with iid normal distribution.

Genomic breeding values for any animal in the population can then be obtained 
by:

$$
\begin{equation}
	\label{eq:gebv}
	\mathbf{\hat{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
\end{equation}
$$

where $\mathbf{p}$ is a row-vector of allele frequencies for all the markers in 
$\mathbf{M}$ in the base population.
The prediction error variance of the vector $\mathbf{\hat{g}}$ is usually 
globally obtained using cross validation.
Alternatively they could be obtained individually by the posterior distribution 
of $(\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}}$ when a Bayesian model is used.

## Single Step Transciptomic prediction of Breeding Values
The breeding values in a mRNA prediction model are in general given by:
$$
\begin{equation}
	\label{eq:mrnaebv}
	\mathbf{\hat{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}
$$

where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model 
equaivalent to the GBLUP model but using mRNA covariates instead of 
marker genotypes.

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in @fernando_class_2014 and impute missing
mRNA data.
The difference here is that &mdash; instead of predicting SNP covariates using 
pedigree relationship &mdash; we impute MRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data are available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
$$
\begin{align}
	\label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}
$$

The distribution of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$.

Using those results from @legarra_relationship_2009 we can model the breeding 
values of the individuals for wich no mRNA information is available the 
expectation of the breeding values in group $2$.
The structure of the residual imputation/prediction error is known and can 
therefore be modelled.
The missing mRNA covariates can be predicted using the expectation of a 
multivariate normal random vector given correlated observations, which related 
to the Best Linear Predictor.

# Data Overview <a name="Data-Overview"></a>
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "purrr", "modelr",
               "htmlTable")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE, message = FALSE)
```

```{r Load-Agro-Data, message = FALSE, cache = TRUE}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
genos <- readRDS("./data/processed/common_genotypes.RDS")

full_hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% full_hybrid) %>%
  dplyr::select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
traits <- all_pheno %>%
  .$Trait %>%
  levels()

# Hybrids of parents with records for both endophenotypes
red_dent <- Reduce(intersect, genos[["Dent"]][c("snp", "mrna")])
red_flint <- Reduce(intersect, genos[["Flint"]][c("snp", "mrna")])
red_hybrid <- genos$Hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% red_dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% red_flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid
full_hybrid <- genos$Hybrid
genos$Hybrid <- NULL
genos$Hybrid$snp <- full_hybrid
genos$Hybrid$mrna <- red_hybrid
```

```{r Feature-Number, echo = FALSE}
snp <- readRDS("./data/processed/snp_mat.RDS")
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
```

Agronomic data for `r length(traits)` traits (including dry-matter yield (DMY))
and `r length(full_hybrid)` hybrids are available 
([Table 1](#Geno-Number-Table)).
After applying a series of quality checks, `r ncol(snp)` SNPs and 
best linear unbiased estimates of `r ncol(mrna)` mRNAs were available for 
further analyses.
Both, genomic and transcriptomic records, were collected for inbred parent lines
of `r length(red_hybrid)` hybrids.

<a name="Geno-Number-Table"></a>
```{r Geno-Number, echo = FALSE}
geno_tab_cap <- paste("**Table 1:** Number of genotypes for Dent and Flint", 
                      "parent lines and their hybrid progeny for which genomic",
                      "and/or transcriptomic data are available.",
                      "All genotypes for which transcriptomic data are", 
                      "available, also have genomic data.",
                      "Agronomic data are available for all 1,521 hybrid",
                      "genotypes.")
# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
# parental Dent and Flint lines, respectively.
genos %>% 
  at_depth(.depth = 2, .f = length) %>%
  bind_rows() %>%
  dplyr::select(-agro) %>%
  mutate(Type = names(genos)) %>%
  dplyr::select(Type, snp, mrna) %>%
  rename("Genomic Data" = snp,
         "Transcriptomic Data" = mrna) %>%
  htmlTable(caption = geno_tab_cap)
```

# Agronomic Data <a name="Agronomic-Data"></a>
```{r echo = FALSE}
# Load data
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")

# Prepare agronomic data.
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))

# Select hybrids for which both parents have genomic as well as transcriptomic
# records.
dent <- genos$Dent$mrna
flint <- genos$Flint$mrna
hybrid <- genos$Hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid

# All hybrids
all_pheno <- all_pheno %>%
  mutate(Reduced_Data = "no")

# Only hybrids for which parents have both, genomic and transcriptomic data
reduced_pheno <- all_pheno %>%
  mutate(Reduced_Data = ifelse(G %in% hybrid, yes = "yes", no = "no")) %>%
  filter(Reduced_Data == "yes")
```

## Skewness
A general rule of thumb to consider is that skewed data &mdash; whose ratio of 
the highest value to the lowest value is greater than 20 &mdash; have 
significant skewness.
The formula for the skewness statistic is
$$
\text{skewness}  = \frac{\sum (x_i - \bar{x})^{3}}{(n - 1)v^{3/2}},
$$
where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ being the predictor
variable, $n$ denoting the number of values and $\bar{x}$ denoting the sample 
mean of the predictor.
```{r Agro-Skewness, echo = FALSE}
skew_tab_cap <- paste("**Table 2:** Skewness of seven agronomic traits.")
all_skewness <- all_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
reduced_skewness <- reduced_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
full_join(x = all_skewness, y = reduced_skewness,
          by = "Trait",
          suffix = c(" (Full)", " (Reduced)")) %>%
  htmlTable(caption = skew_tab_cap)
```

## Distribution
```{r Agro-Distribution, message = FALSE, echo = FALSE, fig.cap = "**Figure 1:** Histogram of the distribution of seven agronomic traits. Black bins show the distribution of the reduced set of hybrids whereas light grey bins show the distribtution of the full set of hybrids."}
ggplot(all_pheno, aes(x = Observed)) +
  geom_histogram(fill = "grey", alpha = 0.5) +
  geom_histogram(data = reduced_pheno, color = "black") +
  facet_wrap(~ Trait, scales = "free") +
  theme_bw()
```

## Correlation
<a name="Correlation"></a>
```{r Agro-Correlation, echo = FALSE, fig.cap = "**Figure 2:** Pairwise correlations among seven agronomic traits for the full set of hybrids and the reduced set of hybrids, respectively."}
par(mar = c(0, 1, 0, 0),
    mfrow = c(1, 2))
all_pheno %>%
  select(-Reduced_Data) %>%
  spread(key = Trait, value = Observed) %>%
  as.data.frame() %>%
  remove_rownames() %>%
  column_to_rownames(var = "G") %>%
  cor() %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     cl.pos = "b",
                     tl.pos = "d",
                     tl.srt = 60,
                     order = "hclust",
                     addCoef.col = "black",
                     title = "Full",
                     mar = c(0, 1, 1, 0))
reduced_pheno %>%
  filter(Reduced_Data == "yes") %>%
  select(-Reduced_Data) %>%
  spread(key = Trait, value = Observed) %>%
  as.data.frame() %>%
  remove_rownames() %>%
  column_to_rownames(var = "G") %>%
  cor() %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     cl.pos = "b",
                     tl.pos = "d",
                     tl.srt = 60,
                     order = "hclust",
                     addCoef.col = "black",
                     title = "Reduced",
                     mar = c(0, 1, 1, 0))
```

# Predictions for Reduced Data <a name="Prediction-for-reduced-data"></a>
```{r Load-LOOCV-Data, cache = TRUE}
# Load log files and data
run_log <- fread("./data/derived/na_fraction_pred_log.txt")

# LOOCV 
loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
  readRDS(paste0("./data/derived/predictions/", loocv_nms[i], ".RDS"))
})
loocv <- rbindlist(loocv)
```

```{r Rename-Traits, cache = TRUE}
loocv_clean <- loocv %>%
  mutate(Trait = fct_recode(Trait, 
                            "FAT" = "FETT",
                            "PRO" = "RPR",
                            "DMY" = "GTM",
                            "DMC" = "GTS",
                            "SUG" = "XZ"),
         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
                                      "SUG")),
         Dent_NA_Fraction = as.factor(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction)),
         Residuals = y - yhat,
         Predictor = fct_recode(Predictor,
                                "mRNA" = "mrna",
                                "SNP" = "snp")
  )
```

## Predictive ability for non-imputed data
<a name="Pred-Table-Non-Imputed"></a>
```{r Non-Imputed-Pred-Ability, cache = TRUE}
loocv_tbl_cap <- paste("**Table 3:** Predictive abilities for the 'reduced' set",
                       "of hybrids using two different predictors and seven",
                       "agronomic traits.",
                       "Results are ranked by predictive ability.")
loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  mutate(Predictor = as.factor(Predictor)) %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
  arrange(desc(Pred_Ability)) %>%
  rename("Predictive Ability" = Pred_Ability) %>%
  htmlTable(caption = loocv_tbl_cap)
```

<a name="Pred-vs-Residual"></a>
```{r Complete-Data-Residuals, cache = TRUE, dependson = "Rename-Traits", fig.cap = "**Figure 3:** Predicted values (*x axis*) vs residuals (*y axis*) based on results from LOOCV with the 'reduced' set of hybrids."}
full_pred_resid <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
       geom_point(size = 0.5) +
       geom_ref_line(h = 0) +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Predicted") +
       ylab("Residuals") +
       ggtitle(.$Predictor))
invisible(lapply(full_pred_resid$plot, FUN = print))
```

<a name="Pred-vs-Observed"></a>
```{r Complete-Data-y-yHat, cache = TRUE, dependson = "Rename-Traits", fig.cap = "**Figure 4:** Observed (*x axis*) vs predicted (*y axis*) values based on results from LOOCV with the 'reduced' set of hybrids."}
full_y_yhat <- loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  group_by(Predictor) %>%
  do(plot = ggplot(., aes(x = y, y = yhat)) +
       geom_point(size = 0.5) +
       geom_abline(slope = 1, linetype = "dashed", color = "blue") +
       facet_wrap(~ Trait, nrow = 2, scales = "free") +
       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
       xlab("Observed") +
       ylab("Predicted") +
       ggtitle(.$Predictor))
invisible(lapply(full_y_yhat$plot, FUN = print))
```

# Predictions with Missingness in Inbred Parent Lines <a name="Pred-NA"></a> 
```{r DMY, cache = TRUE, dependson = "Rename-Traits"}
dmy <- loocv_clean %>%
  mutate(Dent_NA_Fraction = as.numeric(as.character(Dent_NA_Fraction)),
         Flint_NA_Fraction = as.numeric(as.character(Flint_NA_Fraction))) %>%
  filter(Trait == "DMY", Predictor == "mRNA") %>%
  droplevels() %>%
  unite(NA_Fraction, Dent_NA_Fraction, Flint_NA_Fraction, sep = "_") %>%
  group_by(Predictor, Trait, NA_Fraction) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  ungroup() %>%
  arrange(desc(Pred_Ability)) %>%
  separate(NA_Fraction,
           into = c("Dent_NA_Fraction", "Flint_NA_Fraction"), 
           sep = "_") %>%
  mutate(Dent_NA_Fraction = as.numeric(Dent_NA_Fraction),
         Flint_NA_Fraction = as.numeric(Flint_NA_Fraction)) %>%
  rowwise() %>%
  mutate(Avg_NA = mean(c(Dent_NA_Fraction, Flint_NA_Fraction)))
```

## Some mRNA-BLUEs set as missing
This first section deals with an initial attempt to show how well single-step 
hybrid prediction works for different levels of missingness in the mRNA data.
Hereto, we focussed on the "reduced" set of `r length(red_hybrid)` hybrids.
Prior to LOOCV, a fraction of existing transcriptomic BLUEs was set to `NA`
and then imputed using the `sspredr::impute_eta()` function, which is based on
equation 20 in @fernando_class_2014.
Unfortunately, the predictive abilities conditioned on different fractions of 
missingness in Dent and Flint material, are varying to a considerable degree
([Figure 5](#DMY-NA-Prediction), [Figure 6](#All-Traits-NA-Prediction)).
Our impression is that &mdash; in order to arrive at estimates unbiased through
missingness in certain parent lines &mdash; we would have to repeat every
LOOCV-run with a different parent set to `NA` (with respect to mRNA-BLUEs).
However, this is computationally prohibitive.

<a name="DMY-NA-Prediction"></a>
```{r DMY-Dent-NA-Flint-NA, message = FALSE, cache = TRUE, dependson = "DMY", fig.cap = "**Figure 5:** Predictive ability as a function of missingness in Dent and missingness in Flint for dry matter yield (DMY). The blue straight line denotes a 'robust linear regression' through all LOOCV-based predictive abilities. The 95% confidence interval around this regression line is indicated as a gray band."}
library("MASS")
dmy %>%
  mutate(Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction))) %>%
  mutate(Flint_NA_Fraction = fct_relevel(Flint_NA_Fraction, 
                                         "0", "0.1", "0.2", "0.3", "0.4",
                                         "0.5", "0.6")) %>%
  arrange(desc(Dent_NA_Fraction)) %>%
  ggplot(aes(x = Dent_NA_Fraction, y = Pred_Ability, 
             color = Flint_NA_Fraction)) +
  geom_point(size = 2) +
  geom_line() +
  geom_smooth(method = "rlm", aes(group = 1)) +
  theme_light() +
  ylab("r")
detach(package:MASS)
```

<a name="All-Traits-NA-Prediction"></a>
```{r All-Traits-Dent-NA-Flint-NA, cache = TRUE, dependson = "Rename-Trait", fig.cap = "**Figure 6:** Predictive ability as a function of missingness in Dent and missingness in Flint for all traits."} 
loocv_clean %>%
  filter(Dent_NA_Fraction != 0, Flint_NA_Fraction != 0) %>%
  unite(NA_Fraction, Dent_NA_Fraction, Flint_NA_Fraction, sep = "_") %>%
  group_by(Trait, NA_Fraction) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  ungroup() %>%
  separate(NA_Fraction, into = c("Dent_NA_Fraction",
                                 "Flint_NA_Fraction"), sep = "_") %>%
  mutate(Dent_NA_Fraction = as.numeric(Dent_NA_Fraction)) %>%
  ggplot(aes(x = Dent_NA_Fraction, y = Pred_Ability, 
             color = Flint_NA_Fraction)) +
  geom_line() +
  facet_wrap(~ Trait, scales = "free_x") +
  theme_light() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "top") +
  ylab("r")
```

# Single-Step Prediction for the 'full' Set <a name="Full_Prediction"></a>
<a name="Full-Prediction-Table"></a>
```{r Imputed-Pred-Ability, cache = TRUE}
run_log <- fread("./data/processed/prediction_log/log_list.txt")

# LOOCV 
full_loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
full_loocv <- lapply(seq_along(full_loocv_nms), FUN = function(i) {
  readRDS(paste0("./data/processed/predictions/", full_loocv_nms[i], ".RDS"))
})
full_loocv <- rbindlist(full_loocv)

full_loocv_clean <- full_loocv %>%
  mutate(Trait = as.factor(Trait),
         Trait = fct_recode(Trait, 
                            "FAT" = "FETT",
                            "PRO" = "RPR",
                            "DMY" = "GTM",
                            "DMC" = "GTS",
                            "SUG" = "XZ"),
         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
                                      "SUG")),
         Residuals = y - yhat,
         Predictor = as.factor(as.character(Imputation)),
         Predictor = fct_recode(Predictor,
                                "mRNA" = "TRUE",
                                "SNP" = "FALSE")
  )
full_loocv_tbl_cap <- paste(
  "**Table 4:** Predictive abilities for the 'full' set of hybrids using two",
  "different predictors and seven agronomic traits.", 
  "Results are ranked by predictive ability.")
full_loocv_clean %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
  arrange(desc(Pred_Ability)) %>%
  rename("Predictive Ability" = Pred_Ability) %>%
  htmlTable(caption = full_loocv_tbl_cap)
```

# Conclusions and Outlook <a name="Conclusions"></a>
We have seen that

1.    for the 'reduced' set of hybrids, predictive abilities for the traits 
      'DMY', 'PRO' and 'FAT' were higher when predicted via mRNA compared to SNP
      data ([Table 3](#Pred-Table-Non-Imputed)).
2.    for the 'full' set of hybrids, predictive abilities based on single-step
      prediction with transcriptomic data still outperformed SNP data for the
      traits 'DMY' and 'PRO' ([Table 4](#Full-Prediction-Table)).
2.    an exploration of the effect of different levels of missingness &mdash; 
      with respect to transcriptomic BLUEs &mdash; in the parents is 
      computationally prohibitive ([Chapter 6](#Pred-NA)).

Currently, our primary goal is to demonstrate the utility of imputing missing 
data in a quantitative predictor (exemplified by transcriptomic data) and 
using the resulting feature matrix in single-step hybrid prediction.

# References

### 2016-09-20 reports/fernando2014.R

@@ -1,111 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load("clusterGeneration", "ggplot2", "MASS")
-
-# Number of observations
-n_obs <- 50
-# Number of observations without genotypes.
-n_1 <- seq(from = 1, to = n_obs / 2, by = 1)
-# Number of observations with genotypes.
-n_2 <- seq(from = max(n_1) + 1, to = max(n_obs), by = 1)
-
-# Generate a positive definite numerator relationship matrix.
-A <- genPositiveDefMat(dim = n_obs, covMethod = "eigen",
-                       rangeVar = c(0, 2))$Sigma
-# Generate a positive definite genomic relationship matrix.
-G <- genPositiveDefMat(dim = n_obs, covMethod = "eigen", 
-                       rangeVar = c(0, 2))$Sigma
-
-# Block matrices
-## Numerator relationship matrix
-A11 <- A[n_1, n_1]
-A12 <- A[n_1, n_2]
-A21 <- A[n_2, n_1]
-A22 <- A[n_2, n_2]
-## Genomic relationship matrix
-G11 <- G[n_1, n_1]
-G12 <- G[n_1, n_2]
-G21 <- G[n_2, n_1]
-G22 <- G[n_2, n_2]
-
-
-# Legarra et al. (2009), Equation (4) 
-H11 <- A11 + A12 %*% solve(A22) %*% (G22 - A22) %*% solve(A22) %*% A21
-H12 <- A12 %*% solve(A22) %*% solve(G22)
-H21 <- G22 %*% solve(A22) %*% A21
-H22 <- G22
-H <- cbind(rbind(H11, H21), rbind(H12, H22))
-Hinv <- solve(H)
-
-
-# Legarra et al. (2009), Equation (6) 
-Diag <- diag(1, nrow = n_obs / 2, ncol = n_obs / 2)
-Zero <- matrix(0, nrow = n_obs / 2, ncol = n_obs / 2)
-block1 <- cbind(rbind(A12 %*% solve(A22),
-                      Zero),
-                rbind(Zero,
-                      Diag))
-block2 <- rbind(Diag, Diag)
-block3 <- G22 - A22
-block4 <- cbind(Diag, Diag)
-block5 <- cbind(rbind(solve(A22) %*% A21,
-                      Zero),
-                rbind(Zero, Diag))
[…]

### 2016-09-20 reports/intro.Rmd

@@
----
-output: html_document
----
 # Introduction
@@
 genotyped and potentially very young individuals.
-Considering the two sources of information, pedigree and high density SNP 
-genotypes, the process can be described as follows:
+Considering the two sources of information &mdash; pedigree and high density SNP 
+genotypes &mdash; the process can be described as follows:
 
@@
 distributed trait for all individuals available in the population.
-The relationship, and therefor the strucure of the additive genetic covariance, 
-across individuals is based on the numerator relationship matrix $\mathbf{A}$.
+The relationship &mdash; and therefore the strucure of the additive genetic 
+covariance &mdash; across individuals is based on the numerator relationship 
+matrix $\mathbf{A}$.
 In order to obtain breeding values for any individual in the population, we 
@@
 
+$$
 \begin{equation}
@@
 \end{equation}
+$$
 
-where $\mathbf{X}$ is a design matrix for fixed effects, $\mathbf{b}$ a vector 
-of fixed effects, $\mathbf{Z}$ is the design matrix for the random animal 
-effects, which related observations in $\mathbf{y}$ to elements in $\mathbf{A}$.
-$\mathbf{a}$ is the vector of random animal effects (=breeding values) and
-$\mathbf{e}$ is a vector of iid residual errors wich expectation $0$ and 
-variance $\sigma^2_e$.
+where $\mathbf{X}$ is a design matrix for fixed effects, $\mathbf{b}$ is a 
+vector of fixed effects, $\mathbf{Z}$ is the design matrix for the random animal 
+effects, which associate observations in $\mathbf{y}$ with elements in 
+$\mathbf{A}$, $\mathbf{a}$ is the vector of random animal effects (=breeding 
+values) and $\mathbf{e}$ is a vector of iid residual errors with expectation 
+$0$ and variance $\sigma^2_e$.
 The distribution of the random vector $\mathbf{u}$ is 
@@
 training a genomic prediction model.
-That selection includes as necessary condition, that the individual in genotyped 
-and has $\text{PEV}(u_{\text{candidate}}) \leq \text{threshold}$.
+That selection requires that the individual is genotyped and has 
+$\text{PEV}(u_{\text{candidate}}) \leq \text{threshold}$.
 As the breeding values from step one enter the genomic prediction model as 
-phenotypes, one has to be careful about distribution of the residual error of 
+phenotypes, one has to be careful about the residual error distribution of
 such a model.
-For instance, if the range of the prediction error variances is very broud in 
-the reference population, the residual errors will not be identically 
-distributed anymore, e.g. they should be weighted differently or estimated 
-individually.
-One popular approach to account for that problem is *Deregressing* the breeding 
-values and weighting the residuals according to the prediction error variances 
-[@Garrick2009].
-As the breeding values from step one enter the genomic prediction model as 
-phenotypes, one has to be careful about distribution of the residual error of 
-such a model.
-For instance, if the range of the prediction error variances is very broud in 
+For instance, if the range of the prediction error variances is very broad in 
 the reference population, the residual errors will not be identically 
@@
 \end{equation}
+$$
 
@@
 The breeding values in a mRNA prediction model are in general given by:
+$$
 \begin{equation}
@@
 \end{equation}
+$$
 
 where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model 
-equaivalent to equation \@ref(eq:gblup) but using MRNA covariates instead of 
+equaivalent to Eq. \@ref(eq:gblup) but using mRNA covariates instead of 
 marker genotypes.
@@
 transcriptomic information but all have SNP genotypes.
-We can take a similar route as in [@fernando_class_2014] and impute missing
+We can take a similar route as in @fernando_class_2014 and impute missing
 mRNA data.
-The difference here is, that instead of predicting SNP covariates using 
-pedigree relationship, we impute MRNA covariates using SNP or genomic 
+The difference here is that &mdash; instead of predicting SNP covariates using 
+pedigree relationship &mdash; we impute MRNA covariates using SNP or genomic 
 relationships.
 Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
-data is available.
+data are available.
 Individuals with subscript $2$ have both, SNP and mRNA information.
@@
 The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
-expectation given $\mathbf{g}_2$ and a residual:
+expectation given $\mathbf{g}_2$ and a residual error term:
+$$
 \begin{align}
@@
 \end{align}
+$$
 
@@
 
-Using those results from [@legarra_relationship_2009] we can model the breeding 
-values of the individuals for wich no MRNA information is available as expected 
-values from the breeding values in group $2$.
+Using those results from @legarra_relationship_2009 we can model the breeding 
+values of the individuals for wich no mRNA information is available the 
+expectation of the breeding values in group $2$.
 The structure of the residual imputation/prediction error is known and can 
 therefore be modelled.
-The missing MRNA covariates can be predicted using the expectation of a 
+The missing mRNA covariates can be predicted using the expectation of a 
 multivariate normal random vector given correlated observations, which related 
-to the Best Linear Predictor:
+to the Best Linear Predictor.

### 2016-09-20 reports/na_fraction_pheno_predability.Rmd

@@
-# Evaluation of LOOCV predictive abilities
+# Prediction with Imputed mRNAs
 ```{r message = FALSE}
@@
 pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "gridExtra", "modelr", "MASS")
+               "viridis", "tidyr")
 knit_hooks$set(inline = function(x) {
@@
 
-### Predicted vs. residuals
-Generate residual plots for each trait and each combination of missing values 
-in Dent (rows) and Flint (columns) parents.
-```{r Imputed-Residual, cache = TRUE, dependson = "Rename-Traits"}
-resid_plots <- loocv_clean %>%
-  filter(Predictor == "mRNA") %>%
-  group_by(Trait) %>%
-  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
-            geom_point(size = 0.5) +
-            geom_ref_line(h = 0) +
-            facet_grid(Dent_NA_Fraction ~ Flint_NA_Fraction) +
-            theme_light() +
-            theme(axis.text.x = element_text(angle = 60, 
-                                             hjust = 1, vjust = 1)) +
-            xlab("Predicted") +
-            ylab("Residuals") +
-            ggtitle(.$Trait))
-invisible(lapply(resid_plots$plot, print))
-```
-
-### Predicted vs. observed
-Likewise, compare observed with predicted values for combinations that were used
-in the previous plot.
-```{r Imputed-y-yHat, cache = TRUE, dependson = "Rename-Traits"}
-y_yhat_plots <- loocv_clean %>%
-  filter(Predictor == "mRNA") %>%
-  group_by(Trait) %>%
-  do(plot = ggplot(., aes(x = y, y = yhat)) +
-            geom_point(size = 0.5) +
-            geom_abline(slope = 1, linetype = "dashed", color = "blue") +
-            facet_grid(Dent_NA_Fraction ~ Flint_NA_Fraction) +
-            theme(axis.text.x = element_text(angle = 60,
-                                             hjust = 1, vjust = 1)) +
-            xlab("Observed") +
-            ylab("Predicted") +
-            ggtitle(.$Trait))
-invisible(lapply(y_yhat_plots$plot, print))
-```
-
-## LOOCV predictive ability (DMY)
-### mrna
 ```{r DMY, cache = TRUE, dependson = "Rename-Traits"}
@@
 
-### Average r as a function of global missingness and missingness in Flint
-```{r DMY_Global_NA_Flint_NA, cache = TRUE, dependson = "DMY"}
-dmy %>%
-  mutate(Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction))) %>%
-  mutate(Flint_NA_Fraction = fct_relevel(Flint_NA_Fraction, 
-                                         "0", "0.1", "0.2", "0.3", "0.4",
-                                         "0.5", "0.6")) %>%
-  arrange(desc(Avg_NA)) %>%
-  ggplot(aes(x = Avg_NA, y = Pred_Ability, color = Flint_NA_Fraction)) +
-  geom_point(size = 2) +
-  geom_line() +
-  geom_smooth(method = "rlm", aes(group = 1)) +
-  theme_light() +
-  ylab("r")
-```
+## Some mRNA-BLUEs set as missing
+This first section deals with an initial attempt to show how well single-step 
+hybrid prediction works for different levels of missingness in the mRNA data.
+Hereto, we focussed on the "reduced" set of `r length(red_hybrid)` hybrids.
+Prior to LOOCV, a fraction of existing transcriptomic BLUEs was set to `NA`
+and then imputed using the `sspredr::impute_eta()` function, which is based on
+equation 20 in @fernando_class_2014.
+Unfortunately, the predictive abilities conditioned on different fractions of 
+missingness in Dent and Flint material, are varying to a considerable degree
+(Fig. \@ref(fig:DMY-Dent-NA-Flint-NA), 
+Fig. \@ref(fig:All-Traits-Dent-NA-Flint-NA)).
+Our impression is that &mdash; in order to arrive at estimates unbiased through
+missingness in certain parent lines &mdash; we would have to repeat every
+LOOCV-run with a different parent set to `NA` (with respect to mRNA-BLUEs).
+However, this is computationally prohibitive.
 
-### Average r as a function of missingness in Dent and missingness in Flint
-```{r DMY_Dent_NA_Flint_NA, cache = TRUE, dependson = "DMY"}
+```{r DMY-Dent-NA-Flint-NA, cache = TRUE, dependson = "DMY", fig.cap = "Predictive ability as a function of missingness in Dent and missingness in Flint for dry matter yield (DMY).
+The blue straight line denotes a 'robust linear regression' through all LOOCV-based predictive abilities.
+The 95% confidence interval around this regression line is indicated as a gray band."}
+library("MASS")
 dmy %>%
@@
   ylab("r")
+detach(package:MASS)
 ```
 
-### Predictive ability as a function of missingness in Dent and missingness in 
-### Flint for all traits
-```{r All_Traits_Dent_NA_Flint_NA, cache = TRUE, dependson = "Rename-Trait"} 
+```{r All-Traits-Dent-NA-Flint-NA, cache = TRUE, dependson = "Rename-Trait", fig.cap = "Predictive ability as a function of missingness in Dent and missingness in Flint for all traits."} 
 loocv_clean %>%
@@
 ```
+
+## single-step Prediction for the 'full' Set
+```{r Imputed-Pred-Ability, cache = TRUE}
+run_log <- fread("./data/processed/prediction_log/log_list.txt")
+
+# LOOCV 
+full_loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
+full_loocv <- lapply(seq_along(full_loocv_nms), FUN = function(i) {
+  readRDS(paste0("./data/processed/predictions/", full_loocv_nms[i], ".RDS"))
+})
+full_loocv <- rbindlist(full_loocv)
+
+full_loocv_clean <- full_loocv %>%
+  mutate(Trait = as.factor(Trait),
+         Trait = fct_recode(Trait, 
+                            "FAT" = "FETT",
+                            "PRO" = "RPR",
+                            "DMY" = "GTM",
+                            "DMC" = "GTS",
+                            "SUG" = "XZ"),
+         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
+                                      "SUG")),
+         Residuals = y - yhat,
+         Predictor = as.factor(as.character(Imputation)),
+         Predictor = fct_recode(Predictor,
+                                "mRNA" = "TRUE",
+                                "SNP" = "FALSE")
+  )
+full_loocv_tbl_cap <- paste("Predictive abilities for the 'full' set of hybrids",
+                            "using two different predictors and seven agronomic",
+                            "traits. Results are ranked by predictive ability.")
+full_loocv_clean %>%
+  droplevels() %>%
+  group_by(Predictor, Trait) %>%
+  summarize(Pred_Ability = cor(y, yhat)) %>%
+  arrange(desc(Pred_Ability)) %>%
+  rename("Predictive Ability" = Pred_Ability) %>%
+  knitr::kable(caption = full_loocv_tbl_cap, booktabs = TRUE)
+```

### 2016-09-21 reports/appendix.Rmd
Init GitBook

```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("knitr")
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

# Additional Information
## Gamazon et al. (Nature Genetics, 2015)
## Rationale
*    Proposition: directly test the molecular mechanisms through which genomic 
     information affect phenotypic traits.
*    Gene regulation information is missing for some genotypes whereas genomic 
     information are complete.
     The latter was used to to impute missing gene expression information.
*    A major advantage of this approach is that less (imputed) features can be 
     used as surrogates for genomic information in phenotype prediction, thereby
     alleviating the multiple testing problem.
*    Estimate genetically regulated gene expression (GReX), which constitute 
     multiple-SNP prediction of expression levels.
*    Use GReX to impute missing transcriptomic data.

### Shortcomings
"We are aware of the attenuation bias that arises because of the error in the 
estimation of GRex.
This is a subject to be investigated in the future, but the bias does not 
invalidate our analysis as we only use the estimate of GReX as a discovery 
tool."

## Method
The additive model of gene expression is
$$
T = \sum_{k=1}^{K} w_{k} X_{k} + \epsilon,
$$
where $T$ denotes the vector of gene expression values, $k$ denotes the SNP
marker, $X_{k}$ is the number of reference alleles for $k$ and $w_{k}$ is the
coefficient of the model to be fitted.
$\mathbf{w}$ was estimated using LASSO and elastic net.

Missing gene expression values (GReX) are then imputed by multiplying the matrix 
of reference alleles for SNPs with the coefficients $\mathbf{w}$ from the 
training set.
Finally, GReX and the trait of interested are correlated.

## Garrick, Taylor and Fernando (2009)
1.    We do not have true genetic values at our disposal.
2.    The approach of @Gamazon2015 does not account for the non-diagonal 
      variance-covariance structure of the residuals.

Issues with BLUP:
1.    Superior genotypes tend to be underevaluated (i.e. have negative 
      prediction errors) whereas inferior animals tend to be overevaluated.
2.    Observations are shrunken towards the mean while the extent of shrinkage
      depends upon the amount of information, i.e. the reliability.
3.    Training genotypes, which were selected according to their estimated 
      breeding values, comprise individuals with varying reliability.

### 2016-09-21 reports/index.Rmd
Init GitBook

@@
 documentclass: book
-bibliography: ["cheuer.bib"]
+bibliography: ["library.bib", "cheuer.bib"]
 biblio-style: apalike

### 2016-09-21 reports/intro.Rmd
Init GitBook

@@
 [@Garrick2009].
-
-$$
-\begin{equation}
-	\label{eq:gblup}
-	\mathbf{\hat{a}} = \mathbf{1'b} + \mathbf{Mu} + \mathbf{e},
-\end{equation}
-$$
-
-where $\mathbf{M}$ is the matrix of marker covariates and $\mathbf{u}$ is a 
-vector of marker effects with iid normal distribution.
-
 Genomic breeding values for any animal in the population can then be obtained 
@@
 where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model 
-equaivalent to Eq. \@ref(eq:gblup) but using mRNA covariates instead of 
+equaivalent to the GBLUP model but using mRNA covariates instead of 
 marker genotypes.
@@
 data are available.
-Individuals with subscript $2$ have both, SNP and mRNA information.
-The covariates in $\mathbf{W}$ are centered.
-The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
-expectation given $\mathbf{g}_2$ and a residual error term:
-$$
-\begin{align}
-	\label{eq:mrna1}
-	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
-	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
-	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
-\end{align}
-$$
-
-The distribution of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$.
-
-Using those results from @legarra_relationship_2009 we can model the breeding 
-values of the individuals for wich no mRNA information is available the 
-expectation of the breeding values in group $2$.
-The structure of the residual imputation/prediction error is known and can 
-therefore be modelled.
-The missing mRNA covariates can be predicted using the expectation of a 
-multivariate normal random vector given correlated observations, which related 
-to the Best Linear Predictor.

### 2016-09-21 reports/agronomic_data.Rmd
Adjustments to kable() arguments

@@
 ## Skewness
-A general **rule of thumb** to consider is that skewed data whose ratio of the 
+A general rule of thumb to consider is that skewed data whose ratio of the 
 highest value to the lowest value is greater than 20 have significant skewness.
@@
 ```{r Agro-Skewness, echo = FALSE}
+skew_tab_cap <- paste(
+  "Skewness of seven agronomic traits for the full set of hybrids and the",
+  "reduced set of hybrids, respectively."
+)
 all_skewness <- all_pheno %>%
@@
           suffix = c(" (Full)", " (Reduced)")) %>%
-  kable(padding = 0)
+  kable(caption = skew_tab_cap, booktable = TRUE)
 ```

### 2016-09-21 reports/index.Rmd
Adjustments to kable() arguments

--- 2016-09-20 reports/index.Rmd
@@
 output:
-  bookdown::gitbook:
+  bookdown::html_document2:
 documentclass: book
-bibliography: ["cheuer.bib"]
+bibliography: ["library.bib", "cheuer.bib"]
 biblio-style: apalike

### 2016-09-21 reports/na_fraction_pheno_predability.Rmd
Adjustments to kable() arguments

@@
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
-                      warning = FALSE, echo = FALSE)
+                      warning = FALSE, echo = FALSE, message = FALSE)
 ```

### 2016-09-21 reports/single_step_pred.Rmd
Adjustments to kable() arguments

@@
 data are available.
-Individuals with subscript $2$ have both, SNP and mRNA information.
-The covariates in $\mathbf{W}$ are centered.
-The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
-expectation given $\mathbf{g}_2$ and a residual error term:
-$$
-\begin{align}
-	\label{eq:mrna1}
-	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
-	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
-	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
-\end{align}
-$$
-
-The distribution of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$.
-
-Using those results from @legarra_relationship_2009 we can model the breeding 
-values of the individuals for wich no mRNA information is available the 
-expectation of the breeding values in group $2$.
-The structure of the residual imputation/prediction error is known and can 
-therefore be modelled.
-The missing mRNA covariates can be predicted using the expectation of a 
-multivariate normal random vector given correlated observations, which related 
-to the Best Linear Predictor.
 

### 2016-09-21 reports/agronomic_data.Rmd
ss-pred report, single doc with toc

@@
 $$
-where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ is the predictor
-variable, $n$ is the number of values and $\bar{x}$ is the sample mean of the 
-predictor.
+where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ denoting the 
+predictor variable, $n$ denoting the number of values and $\bar{x}$ denoting the 
+sample mean of the predictor.
 ```{r Agro-Skewness, echo = FALSE}
@@
           suffix = c(" (Full)", " (Reduced)")) %>%
-  kable(caption = skew_tab_cap, booktable = TRUE)
+  kable(caption = skew_tab_cap, booktabs = TRUE)
 ```
@@
 ```{r Agro-Distribution, message = FALSE, echo = FALSE, fig.cap = "Histogram of the distribution of seven agronomic traits.
-Black bins show the distribution of the reduced set of hybrids whereas light grey bins show the distribtution of the full set of hybrids."}
+Black bins show the distribution of the reduced set of hybrids whereas light grey bins show the distribution of the full set of hybrids."}
 ggplot(all_pheno, aes(x = Observed)) +

### 2016-09-21 reports/index.Rmd
ss-pred report, single doc with toc

--- 2016-09-20 reports/index.Rmd
@@
 output:
-  bookdown::gitbook:
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
+    theme: readable
+    highlight: tango
 documentclass: book
-bibliography: ["cheuer.bib"]
+bibliography: ["library.bib", "cheuer.bib"]
 biblio-style: apalike
@@
 ---
+
+<script type="text/x-mathjax-config">
+MathJax.Hub.Config({
+  TeX: { equationNumbers: { autoNumber: "AMS" } }
+});
+</script>
 

### 2016-09-21 reports/intro.Rmd
ss-pred report, single doc with toc

@@
+---
+output: pdf_document
+---
 # Introduction
@@
 $0$ and variance $\sigma^2_e$.
-The distribution of the random vector $\mathbf{u}$ is 
+The distribution of the random vector $\mathbf{a}$ is 
 $MVN(0, \mathbf{A}\sigma^2_a)$.
@@
 That selection requires that the individual is genotyped and has 
-$\text{PEV}(u_{\text{candidate}}) \leq \text{threshold}$.
+$\text{PEV}(a_{\text{candidate}}) \leq \text{threshold}$.
 As the breeding values from step one enter the genomic prediction model as 
@@
 	\label{eq:gebv}
-	\mathbf{\hat{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
+	\mathbf{\tilde{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
 \end{equation}
@@
 
-where $\mathbf{p}$ is a row-vector of allele frequencies for all the markers in 
-$\mathbf{M}$ in the base population.
-The prediction error variance of the vector $\mathbf{\hat{g}}$ is usually 
+where $\mathbf{p}$ is a row-vector of allele frequencies in the base population,
+$\mathbf{M}$ is a matrix of marker covariates and $\mathbf{u}$ is a vector of
+marker effects.
+The prediction error variance of the vector $\mathbf{\tilde{g}}$ is usually 
 globally obtained using cross validation. 
@@
 	\label{eq:mrnaebv}
-	\mathbf{\hat{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
+	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
 \end{equation}
@@
 where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model 
-equaivalent to the GBLUP model but using mRNA covariates instead of 
+equaivalent to the GBLUP model but using mRNA covariates ($\mathbf{W}$) instead of 
 marker genotypes.
@@
 The difference here is that &mdash; instead of predicting SNP covariates using 
-pedigree relationship &mdash; we impute MRNA covariates using SNP or genomic 
+pedigree relationship &mdash; we impute mRNA covariates using SNP or genomic 
 relationships.
@@
 data are available.
+Individuals with subscript $2$ have both, SNP and mRNA information.
+The covariates in $\mathbf{W}$ are centered.
+The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
+expectation given $\mathbf{g}_2$ and a residual error term:
+$$
+\begin{align}
+	\label{eq:mrna1}
+	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
+	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
+	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
+\end{align}
+$$
+
+The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
+[@legarra_relationship_2009].
+
+The structure of the residual imputation/prediction error is known and can 
+therefore be modelled.
+The missing mRNA covariates can be predicted using the expectation of a 
+multivariate normal random vector given correlated observations, which are 
+related to the Best Linear Predictor.
+
+### Extending the single step model
+
+To this point, we have only taken the general idea behind single step genomic prediction and applied
+it to the situation in which the breeding values are defined as the sum of mRNA effects plus an *imputation error*
+term from using genomic relationships to impute missing mRNA data.
+It might also be interesting to include individuals which have neither mRNA nor SNP data available,
+but phenotypes.
+The general mRNA model for that situation could be written as:
+
+$$
+\begin{equation}
+\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Ped} \boldsymbol{\epsilon}_{Ped} 
++ \mathbf{U}_{SNP} \boldsymbol{\epsilon}_{SNP}+ \mathbf{e},
+\end{equation}
+$$
+
+with
+
+$$
+\begin{equation}
+\mathbf{X} = 
+\begin{bmatrix}
+  X_0 \\
+  X_1 \\
+  X_2 
+ \end{bmatrix},
+ \mathbf{W} = 
+\begin{bmatrix}
+  Z_0\hat{\mathbf{W}_0} \\
+  Z_1\hat{\mathbf{W}_1} \\
+  \mathbf{W}_2 
+ \end{bmatrix},
+ \mathbf{U}_{Ped} = 
+\begin{bmatrix}
+  Z_0 \\
+  0 \\
+  0 
+ \end{bmatrix},
+ \mathbf{U}_{SNP} = 
+\begin{bmatrix}
+  0 \\
+  Z_1 \\
+  0 
+ \end{bmatrix},
+ \end{equation}
+$$
+
+The subscript 0 denotes animals with Pedigree information and neither
+SNP nor mRNA data, 1 individuals with SNP but no mRNA data and 2 individuals with mRNA data.
+
+Following the notation in @fernando_class_2014 the phenotypes can be untangled like this:
+
+$$
+\begin{align}
+\begin{bmatrix}
+  y_0 \\
+  y_1 \\
+  y_2 
+ \end{bmatrix}
+& =
+ \begin{bmatrix}
+  X_0 \\
+  X_1 \\
+  X_2 
+ \end{bmatrix}
+ \boldsymbol{\beta} + 
+ \begin{bmatrix}
+  Z_0 & 0 & 0 \\
+  0 & Z_1 & 0 \\
+  0 & 0 & Z_2 
+ \end{bmatrix}
+\begin{bmatrix}
+  g_0 \\
+  g_1 \\
+  g_2 
+ \end{bmatrix}
+  + \mathbf{e} \\
+    & = 
+ \begin{bmatrix}
+  X_0 \\
+  X_1 \\
+  X_2 
+ \end{bmatrix}
+ \boldsymbol{\beta} + 
+ \begin{bmatrix}
+  Z_0 & 0 & 0 \\
+  0 & Z_1 & 0 \\
+  0 & 0 & Z_2 
+ \end{bmatrix}
+\begin{bmatrix}
+  \mathbf{A}_{02}\mathbf{A}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
+  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP}  \\
+  \mathbf{W}_2\boldsymbol{\alpha} \\
+ \end{bmatrix}
+  + \mathbf{e} \\
+    & = 
+ \begin{bmatrix}
+  X_0 \\
+  X_1 \\
+  X_2 
+ \end{bmatrix}
+ \boldsymbol{\beta} + 
+ \begin{bmatrix}
+  Z_0 & 0 & 0 \\
+  0 & Z_1 & 0 \\
+  0 & 0 & Z_2 
+ \end{bmatrix}
+\begin{bmatrix}
+  \hat{\mathbf{W}_0}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
+  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP} \\
+  \mathbf{W}_2\boldsymbol{\alpha} \\
+ \end{bmatrix}
+  + \mathbf{e} \\
+\end{align}
+$$
+
+The breeding values from that model are:
+
+$$
+\begin{equation}
+\tilde{\mathbf{g}} = 
+ \begin{bmatrix}
+  \hat{\mathbf{W}_0} \\
+  \hat{\mathbf{W}_1} \\
+  \mathbf{W}_2 
+ \end{bmatrix}
+ \hat{\boldsymbol{\alpha}}
+ + 
+ \begin{bmatrix}
+  \mathbf{Z}_0 \\
+  0 \\
+  0
+ \end{bmatrix}
+ \hat{\boldsymbol{\epsilon}}_{Ped}
+ + 
+ \begin{bmatrix}
+  0 \\
+  \mathbf{Z}_1 \\
+  0
+ \end{bmatrix}
+ \hat{\boldsymbol{\epsilon}}_{SNP} 
+ \end{equation}
+ $$
+
+When dealing with hybrid performance as phenotypes, we are interested in estimating
+general combining abilitiy (GCA) for every line of a specific subpopulation.
+In our case, the subpopulations comprise *dent* and *flint*.
+Hence, our final model looks like this:
+
+$$
+\begin{align}
+\mathbf{y} &= \mathbf{Xb} \\ 
+  &+ \mathbf{Z}_{dent}\mathbf{W}_{dent} \boldsymbol{\alpha}_{dent} + \mathbf{Z}_{dent}\mathbf{U}_{dent_{Ped}} \boldsymbol{\epsilon}_{dent_{Ped}}
++ \mathbf{Z}_{dent}\mathbf{U}_{dent_{SNP}} \boldsymbol{\epsilon}_{dent_{SNP}} \\
+  &+ \mathbf{Z}_{flint}\mathbf{W}_{flint} \boldsymbol{\alpha}_{flint} + \mathbf{Z}_{flint}\mathbf{U}_{flint_{Ped}} \boldsymbol{\epsilon}_{flint_{Ped}}
++ \mathbf{Z}_{flint}\mathbf{U}_{flint_{SNP}} \boldsymbol{\epsilon}_{flint_{SNP}} \\
+  &+ \mathbf{e},
+\end{align}
+$$
+
+where $\mathbf{Z}_{dent}$ and $\mathbf{Z}_{flint}$ are design matrices that map hybrids to their respective
+parents in the dent and flint lines.
+The obtained GCA effects for the inbred lines represent half their breeding values.
+The predicted hybrid performance is:
+
+$$
+\begin{equation}
+\tilde{\mathbf{h}} = \mathbf{Z}_{dent}\tilde{\mathbf{g}}_{dent} + \mathbf{Z}_{flint}\tilde{\mathbf{g}}_{flint} 
+\end{equation}
+$$

### 2016-09-21 reports/loocv_reduced.Rmd
ss-pred report, single doc with toc

@@
   group_by(Predictor, Trait) %>%
-  summarize(Pred_Ability = cor(y, yhat)) %>%
+  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
   arrange(desc(Pred_Ability)) %>%

### 2016-09-23 analysis/impute_single_mrna.Rmd
Init mRNA inspection

---
title: "Genotype-Wise mRNA-Imputation"
author: Matthias Westhues
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
---

Keep only genotypes for which both, transcriptomic and genomic data are 
available.
```{r}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/", inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, message = FALSE)
```

```{r mRNA-Data}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
dent <- as.character(genos$Dent$mrna)
flint <- as.character(genos$Flint$mrna)

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  select(-G)
```

Run a principal component analysis (PCA) on the "raw" mRNA-BLUEs.
```{r Raw-PCA}
mrna_df %>%
  split(.$Group) %>%
  map(~select(-Group)) %>%
  str()
  map(prcomp)
  by_slice(.f = prcomp)
  map(~ prcomp, data = .)
pca_obj <- prcomp(mrna)
perc_var <- pca_obj$sd ^ 2 / sum(pca_obj$sd ^ 2) * 100
perc_var[1:3]
```

`r sum(perc_var[seq_len(3)])`% of the variance in the "raw" mRNA-BLUEs is 
explained by the first three principal components.

## Data pre-processing
Pre-process the data in the following order:

1.    Near-zero-variance filtering.
2.    Box-Cox tranformation
3.    Centering
4.    Scaling
```{r Transform-Data}
trans <- preProcess(mrna, 
                    method = c("BoxCox", "center", "scale", "nzv"))
# Apply the transformation
mrna_trans <- predict(trans, mrna)
#correlations <- cor(mrna_trans)
#high_corr <- findCorrelation(correlations, cutoff = 0.9)
#mrna_trans <- mrna_trans[, -high_corr]

# PCA
trans_pca_obj <- prcomp(mrna_trans)
trans_perc_var <- trans_pca_obj$sd ^ 2 / sum(trans_pca_obj$sd ^ 2) * 100
trans_perc_var[1:3]
```

`r sum(trans_perc_var[seq_len(3)])`% of the variance in the transformed 
mRNA-BLUEs is explained by the first three principal components compared to
`r sum(perc_var[seq_len(3)])`% in the "raw" data.

## Tidy-up data
```{r Tidy-Data}
mrna_lst <- list(Raw = mrna, Trans = mrna_trans)
mrna_lst[] <- lapply(mrna_lst, FUN = as.data.frame)
tidy_mrna <- mrna_lst %>%
  bind_rows(.id = "Type") %>%
  as_tibble() %>%
  gather(key = "mRNA", value = "Expression", -Type) %>%
  mutate(Type = as.factor(Type))
```

Compute the skewness and the Shapiro-Wilk test of normality for each mRNA
```{r Check-Normality}
mrna_skew <- tidy_mrna %>%
  group_by(Type, mRNA) %>%
  summarize(Skewness = e1071::skewness(Expression),
            Shapiro = stats::shapiro.test(Expression)$statistic)
```

Plot the distribution of skewness values across all mRNAs
```{r}
mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  geom_density(alpha = 0.2) +
  theme(legend.position = "top")

mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  xlim(c(-2, 2)) +
  geom_density(alpha = 0.2) +
  theme(legend.position = "top")
```

Find the least skewed/non-normally distributed mRNAs.
```{r Association-Shapiro-Wilks-Skewness}
mrna_skew %>%
  mutate(Abs_Skewness = abs(Skewness)) %>%
  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
  xlim(c(0.8, 1)) +
  ylim(c(0, 2)) +
  geom_point() +
  facet_grid(. ~ Type) +
  xlab("Shapiro-Wilks Test Statistic") +
  ylab("Absolute Skewness")
```

## Function for the imputation of a single mRNA.
```{r Impute-mRNA-Function}
## Function for the imputation of a single mRNA
impute_one_mrna <- function(x, y) {
  geno <- rownames(x)
  # Names of genotypes for which transcriptomic records exist
  nm2 <- rownames(y)
  # Names of genotypes for which transcriptomic records are missing.
  nm1 <- setdiff(rownames(x), nm2)
  # Hybrid parents not in y.
  geno1 <- geno[geno %in% nm1]
  # Hybrid parents in y.
  geno2 <- geno[geno %in% nm2]
  x <- x[match(unique(geno), rownames(x)), ]
  # Design matrix mapping GCA effects from genotypes with transcriptomic
  # records to y2.
  Z2 <- Matrix::sparse.model.matrix(~-1 + factor(geno2),
                                    drop.unused.levels = FALSE)
  colnames(Z2) <- gsub("factor\\(geno2\\)", replacement = "", x = colnames(Z2))
  Z2 <- Z2[, match(nm2, colnames(Z2))]
  rownames(Z2) <- geno2
  # Hybrid parents not in y.
  geno1 <- geno[geno %in% nm1]
  # Hybrid parents in y.
  geno2 <- geno[geno %in% nm2]
  x <- x[match(unique(geno), rownames(x)), ]
  y <- y[nm2, ]
  M2 <- y[, matrixStats::colVars(y) != 0]
  x <- x[, matrixStats::colVars(x) != 0]
  A <- build_kernel(M = x, lambda = 0.01, algorithm = "RadenII")
  A11 <- A[nm1, nm1, drop = FALSE]
  A12 <- A[nm1, nm2, drop = FALSE]
  A21 <- A[nm2, nm1, drop = FALSE]
  A22 <- A[nm2, nm2, drop = FALSE]
  Ainv <- solve(A)
  dimnames(Ainv) <- dimnames(A)
  A_up11 <- Ainv[nm1, nm1]
  A_up12 <- Ainv[nm1, nm2]
  # Eq.21
  M1 <- A12 %*% solve(A22) %*% M2
  J2 <- matrix(-1, nrow = ncol(A12), ncol = 1)
  # Eq.22
  J1 <- A12 %*% solve(A22) %*% J2
  # Eq.10
  epsilon <- t(chol(solve(Ainv[nm1, nm1])))
  # Eq.20
  W2 <- Z2 %*% M2
  w1 <- A12 %*% solve(A22) %*% Z2 %*% M2
  as.numeric(w1)
}
```

Plot the distribution of correlations between vectors of original and imputed
mRNAs.
```{r Correlation-Imputed-vs-NonImputed}
#imp_mrna <- vapply(seq_len(nrow(y)), FUN = function(i) {
#  z <- y[-i, ]
#  imp_mrna <- impute_one_mrna(x, z)
#  cor(y[i, ], imp_mrna)
#}, FUN.VALUE = numeric(1))
#hist(imp_mrna)
```

### 2016-09-26 reports/impute_single_mrna.Rmd
Init mRNA report

--- 2016-09-23 analysis/impute_single_mrna.Rmd
@@
----
-title: "Genotype-Wise mRNA-Imputation"
-author: Matthias Westhues
-date: "`r Sys.Date()`"
-output: 
-  html_document:
-    theme: united
-    highlight: tango
-    toc: true
-    toc_float: true
----
-
-Keep only genotypes for which both, transcriptomic and genomic data are 
-available.
-```{r}
-pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
-knitr::knit_hooks$set(inline = function(x) {
-  prettyNum(x, big.mark = ",", digits = 2)
-})
-knitr::opts_knit$set(root.dir = "~/gamazon/", inline = TRUE)
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
-                      warning = FALSE, message = FALSE)
-```
-
-```{r mRNA-Data}
-# Common genotypes
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-dent <- as.character(genos$Dent$mrna)
-flint <- as.character(genos$Flint$mrna)
-
-# mRNA
-mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
-mrna <- t(mrna)
-mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
-mrna_df <- mrna %>%
-  as_data_frame() %>%
-  mutate(G = rownames(mrna)) %>%
-  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
-         Group = as.factor(Group)) %>%
-  select(-G)
-```
-
-Run a principal component analysis (PCA) on the "raw" mRNA-BLUEs.
-```{r Raw-PCA}
-mrna_df %>%
-  split(.$Group) %>%
-  map(~select(-Group)) %>%
-  str()
-  map(prcomp)
-  by_slice(.f = prcomp)
-  map(~ prcomp, data = .)
-pca_obj <- prcomp(mrna)
-perc_var <- pca_obj$sd ^ 2 / sum(pca_obj$sd ^ 2) * 100
-perc_var[1:3]
-```
-
-`r sum(perc_var[seq_len(3)])`% of the variance in the "raw" mRNA-BLUEs is 
-explained by the first three principal components.
-
-## Data pre-processing
-Pre-process the data in the following order:
-
-1.    Near-zero-variance filtering.
-2.    Box-Cox tranformation
-3.    Centering
-4.    Scaling
-```{r Transform-Data}
-trans <- preProcess(mrna, 
-                    method = c("BoxCox", "center", "scale", "nzv"))
-# Apply the transformation
-mrna_trans <- predict(trans, mrna)
-#correlations <- cor(mrna_trans)
-#high_corr <- findCorrelation(correlations, cutoff = 0.9)
-#mrna_trans <- mrna_trans[, -high_corr]
-
-# PCA
-trans_pca_obj <- prcomp(mrna_trans)
-trans_perc_var <- trans_pca_obj$sd ^ 2 / sum(trans_pca_obj$sd ^ 2) * 100
-trans_perc_var[1:3]
-```
-
-`r sum(trans_perc_var[seq_len(3)])`% of the variance in the transformed 
-mRNA-BLUEs is explained by the first three principal components compared to
-`r sum(perc_var[seq_len(3)])`% in the "raw" data.
-
-## Tidy-up data
-```{r Tidy-Data}
-mrna_lst <- list(Raw = mrna, Trans = mrna_trans)
-mrna_lst[] <- lapply(mrna_lst, FUN = as.data.frame)
-tidy_mrna <- mrna_lst %>%
-  bind_rows(.id = "Type") %>%
-  as_tibble() %>%
-  gather(key = "mRNA", value = "Expression", -Type) %>%
-  mutate(Type = as.factor(Type))
-```
-
-Compute the skewness and the Shapiro-Wilk test of normality for each mRNA
-```{r Check-Normality}
-mrna_skew <- tidy_mrna %>%
-  group_by(Type, mRNA) %>%
-  summarize(Skewness = e1071::skewness(Expression),
-            Shapiro = stats::shapiro.test(Expression)$statistic)
-```
-
-Plot the distribution of skewness values across all mRNAs
-```{r}
-mrna_skew %>%
-  ggplot(aes(x = Skewness, fill = Type)) +
-  geom_density(alpha = 0.2) +
-  theme(legend.position = "top")
-
-mrna_skew %>%
-  ggplot(aes(x = Skewness, fill = Type)) +
-  xlim(c(-2, 2)) +
-  geom_density(alpha = 0.2) +
-  theme(legend.position = "top")
-```
-
-Find the least skewed/non-normally distributed mRNAs.
-```{r Association-Shapiro-Wilks-Skewness}
-mrna_skew %>%
-  mutate(Abs_Skewness = abs(Skewness)) %>%
-  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
-  xlim(c(0.8, 1)) +
-  ylim(c(0, 2)) +
-  geom_point() +
-  facet_grid(. ~ Type) +
-  xlab("Shapiro-Wilks Test Statistic") +
-  ylab("Absolute Skewness")
-```
+# Genotype-Wise mRNA-Imputation
 

### 2016-09-26 reports/transcriptomic_data.Rmd
Init mRNA report

# Transcriptomic Data
Keep only genotypes for which both, transcriptomic and genomic data are 
available.
```{r echo = FALSE}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/", inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, message = FALSE)
```

```{r mRNA-Data, cache = TRUE}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
dent <- as.character(genos$Dent$mrna)
flint <- as.character(genos$Flint$mrna)

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  select(-G)
```

Separately for each heterotic group, run a principal component analysis (PCA) on 
the "raw" mRNA-BLUEs and return the importance of the first three PCs.
```{r Raw-PCA, cache = TRUE}
rm_grp <- function(x) {
  x$Group <- NULL
  x
}
pca_var_imp <- function(x) {
  var_imp <- x$sd ^ 2 / sum(x$sd ^ 2) * 100
  sum(var_imp[1:3])
}
# Variable importance in PCA (first three PCs)
raw_varimp <- mrna_df %>%
  split(.$Group) %>%
  map(rm_grp) %>%
  map(prcomp) %>%
  map(pca_var_imp) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type = "Raw")
```

## Data pre-processing
Pre-process the data in the following order:

1.    Near-zero-variance filtering.
2.    Box-Cox tranformation
3.    Centering
4.    Scaling
```{r Transform-Data, cache = TRUE}
mrna_lst <- mrna_df %>% 
  split(.$Group) %>%
  map(rm_grp) %>%
  map(~as.matrix(.))

# Data transformation
trans_lst <- mrna_lst %>%
  map(preProcess,
      method = c("BoxCox", "center", "scale", "nzv")) %>%
  map2(.y = mrna_lst, .f = predict)

# Variable importance in PCA (first three PCs)
trans_varimp <- trans_lst %>%
  map(prcomp) %>%
  map(pca_var_imp) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type = "Trans")

# Compare variable importances for raw and transformed mRNA data.
rbind(raw_varimp, trans_varimp) %>%
  knitr::kable(caption = "Importance of the first three PCs (in %)",
               digits = 2)
```

## Tidy-up data
```{r Tidy-Data, cache = TRUE}
mrna_lst <- list(Raw = mrna_lst, Trans = trans_lst)
tidy_mrna <- mrna_lst %>% 
  at_depth(.depth = 2, .f = ~as.data.frame(.)) %>%
  at_depth(.depth = 1, .f = ~bind_rows(., .id = "Group")) %>%
  bind_rows(.id = "Type") %>%
  as_tibble() %>%
  gather(key = "mRNA", value = "Expression", -Group, -Type) %>%
  mutate(Type = as.factor(Type),
         Group = as.factor(Group))
```

## Skewness
Compute the skewness and the Shapiro-Wilk test of normality for each mRNA
```{r Check-Normality, cache = TRUE}
mrna_skew <- tidy_mrna %>%
  group_by(Group, Type, mRNA) %>%
  summarize(Skewness = e1071::skewness(Expression),
            Shapiro = stats::shapiro.test(Expression)$statistic)
```

Plot the distribution of skewness values across all mRNAs
```{r, mRNA-Skewness, cache = TRUE}
mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top")

mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  xlim(c(-2, 2)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top")
```

Find the least skewed/non-normally distributed mRNAs.
```{r Association-Shapiro-Wilks-Skewness, cache = TRUE}
mrna_skew %>%
  mutate(Abs_Skewness = abs(Skewness)) %>%
  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
  xlim(c(0.8, 1)) +
  ylim(c(0, 2)) +
  geom_point() +
  facet_grid(Group ~ Type) +
  xlab("Shapiro-Wilks Test Statistic") +
  ylab("Absolute Skewness")
```

### 2016-09-26 reports/impute_single_mrna.Rmd
Report on cor(orig_mrna,imp_mrna)

@@
 # Genotype-Wise mRNA-Imputation
+```{r echo = FALSE}
+pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
+knitr::knit_hooks$set(inline = function(x) {
+  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
+})
+knitr::opts_knit$set(root.dir = "~/gamazon/", inline = TRUE)
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
+                      warning = FALSE, message = FALSE)
+```
+
+```{r Predictor-Data, cache = TRUE}
+# Common genotypes
+genos <- readRDS("./data/processed/common_genotypes.RDS")
+dent <- as.character(genos$Dent$mrna)
+flint <- as.character(genos$Flint$mrna)
+
+# Remove the 'Group' variable after splitting Dent and Flint data.
+rm_grp <- function(x) {
+  x$Group <- NULL
+  x
+}
+
+# mRNA
+mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
+mrna <- t(mrna)
+mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
+mrna_lst <- mrna %>%
+  as_data_frame() %>%
+  mutate(G = rownames(mrna)) %>%
+  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group)) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "G") %>%
+  split(.$Group) %>%
+  map(rm_grp) %>%
+  map(~as.matrix(.))
+
+# SNPs
+snp <- readRDS("./data/processed/snp_mat.RDS")
+snp_lst <- snp %>%
+  as_data_frame() %>%
+  mutate(G = rownames(snp)) %>%
+  filter(G %in% c(dent, flint)) %>%
+  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group)) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "G") %>%
+  split(.$Group) %>%
+  map(rm_grp) %>%
+  map(~as.matrix(.))
+```
 
@@
 ## Function for the imputation of a single mRNA
-impute_one_mrna <- function(x, y) {
+impute_mrna_genotype <- function(x, y) {
   geno <- rownames(x)
@@
 
-Plot the distribution of correlations between vectors of original and imputed
-mRNAs.
-```{r Correlation-Imputed-vs-NonImputed}
-#imp_mrna <- vapply(seq_len(nrow(y)), FUN = function(i) {
-#  z <- y[-i, ]
-#  imp_mrna <- impute_one_mrna(x, z)
-#  cor(y[i, ], imp_mrna)
-#}, FUN.VALUE = numeric(1))
-#hist(imp_mrna)
+## Correlations between vectors of original and imputed mRNAs.
+```{r Compute-Correlation-Imputed-vs-NonImputed, cache = TRUE}
+# Wrapper function for the imputation of single genotypes.
+imp_loop <- function(x, y) {
+  vapply(seq_len(nrow(y)), FUN = function(i) {
+    z <- y[-i, ]
+    imp_mrna <- impute_mrna_genotype(x = x, y = z)
+    cor(y[i, ], imp_mrna)
+  }, FUN.VALUE = numeric(1))
+}
+# Separately for each heterotic group, compute the correlation between each 
+# original mRNA-genotype and the corresponding imputed mRNA-genotype.
+imp_correlations <- map2(.x = snp_lst, .y = mrna_lst, .f = imp_loop)
 ```
+
+```{r Plot-Correlation-Imputed-vs-NonImputed, cache = TRUE, fig.cap = "Distribution of Pearson correlation coefficients between original and imputed mRNA genotypes for two heterotic groups."}
+# Plot the distribution of the resulting Pearson correlation coefficients.
+imp_correlations %>%
+  stack() %>%
+  rename(Value = values,
+         Group = ind) %>%
+  ggplot(aes(x = Value)) +
+  geom_density() +
+  geom_rug() +
+  facet_grid(. ~ Group) +
+  ylab("Density")
+```

### 2016-09-27 reports/index.Rmd
Init mRNA report
Implement efficient LOOCV

@@
 
-Chapter \@ref(prediction-for-reduced-data) deals with the results of hybrid 
-prediction for the reduced set of data.
+Chapter \@ref(transcriptomic-data) explores the distribution and correlations
+of/among transcriptomic predictors for "raw" and transformed mRNA BLUEs.
+
+Chapter \@ref(genotype-wise-mrna-imputation) examines the correlation between
+the original and imputed transcriptomic BLUEs.
+
+Chapter \@ref(loocv-prediction-without-prediction-error-estimates) deals with 
+the results of hybrid prediction for the reduced set of data and a comparison
+of predictive abilities with the full data set.
+
+Chapter \@ref(efficient-loocv-with-prediction-errors) is a first implementation
+of the approach suggested by Cheng, Garrick and Fernando (2016, unpublished)
+using a data set from the `BGLR` package.
 

### 2016-09-27 reports/loocv_example.Rmd
Implement efficient LOOCV

# Efficient LOOCV With Prediction Errors

The following code snippet is taken from the [BGLR Vignette](https://github.com/gdlc/BGLR-R/blob/master/inst/md/hyperparameters.md).
```{r Fit-BGLR-Model, cache = TRUE}
pacman::p_load("BGLR")
data("mice")
X <- scale(mice.X[, 1:2000])
h2 <- 0.5
QTL <- seq(from = 50, to = ncol(X), length = 20)
nQTL <- length(QTL)
n <- nrow(X)
b <- rep(1, nQTL) * sqrt(h2 / nQTL)
signal <- X[, QTL] %*% b
error <- rnorm(n, sd = sqrt(1 - h2))
y <- signal + error

## Using a flat prior
fm1 <- BGLR(y = y, ETA = list(list(X = X, model = "BayesB", counts0 = 1,
                                   counts1 = 1)), 
            nIter = 5000, burnIn = 500, verbose = FALSE)
```

All formulae are from Cheng, Garrick and Fernando (2016, unpublished):
"Efficient Strategies for Leave-one-out Cross Validation for Genomic Best Linear
Unbiased Prediction".

$$
\mathbf{X}^{*} = 
  \begin{bmatrix}
    \mathbf{1} & \mathbf{X}
  \end{bmatrix}
$$
```{r X-prime}
X_prime <- cbind(matrix(1, nrow = nrow(X), ncol = 1),
                 X)
```

$$
\hat{\beta}^{*} =
  \begin{bmatrix}
    \hat{\mu} \\
    \hat{\beta}
  \end{bmatrix}
$$
```{r}
b_hat <- fm1$ETA[[1]]$b
mu_hat <- fm1$mu
b_hat_prime <- matrix(c(mu_hat, b_hat), ncol = 1)
```

$\lambda = \frac{\sigma_{e}^{2}}{\sigma_{\beta}^{2}}$
```{r}
lambda <- c(fm1$varE / fm1$ETA[[1]]$varB, 1)
```

Diagonal matrix $\mathbf{D}_{[(p + 1) \times (p + 1)]}$.
```{r}
Diag <- cbind(diag(nrow = ncol(X), ncol = ncol(X)),
              matrix(1, nrow = ncol(X), ncol = 1))
Diag <- rbind(Diag,
              matrix(1, nrow = 1, ncol = ncol(Diag)))
```

$\mathbf{H} = \mathbf{X}^{*}(\mathbf{X}^{*'}\mathbf{X}^{*} +
\mathbf{D}\lambda)^{-1}\mathbf{X}^{*'}$
```{r H, cache = TRUE}
H <- X_prime %*% solve(t(X_prime) %*% X_prime + Diag * lambda) %*% t(X_prime)
```

$\hat{e_{j}} = \frac{y_{j} - 
\mathbf{x^{*'}}_{j}\hat{\beta}^{*}}{1-\mathbf{H_{jj}}}$
```{r LOOCV-Residuals}
e_hat <- vapply(seq_len(nrow(X_prime)), FUN = function(j) {
  t_x_prime <- X_prime[j, , drop = FALSE]
  numerator <- as.numeric(y[j] - t_x_prime %*% b_hat_prime)
  denominator <- 1 - H[j, j]
  numerator / denominator
}, FUN.VALUE = numeric(1))
```

$\text{PRESS} = \sum_{i=1}^{n} \hat{e}_{j}^{2}$
```{r LOOCV-Output-Summary}
press <- sum(e_hat ^ 2)
rmse <- sqrt(press / length(e_hat))
y_hat <- y - e_hat
# Leave-one-out prediction accuracy
r <- cor(y, y_hat)
```

The LOOCV-RMSE is `r round(rmse, digits = 2)` and the prediction accuracy is 
`r round(r, digits = 2)`.

### 2016-09-27 reports/loocv_reduced.Rmd
Implement efficient LOOCV

--- 2016-09-19 reports/loocv_reduced.Rmd
@@
-# Prediction For Reduced Data
+# LOOCV-Prediction Without Prediction Error Estimates
 ```{r message = FALSE, echo = FALSE}
@@
 pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "gridExtra", "modelr", "MASS")
+               "viridis", "tidyr", "modelr")
 
-knit_hooks$set(inline = function(x) {
+knitr::knit_hooks$set(inline = function(x) {
   prettyNum(x, big.mark = ",", digits = 2)
@@
 
-## Predictive ability for non-imputed data
-Peak at predictive abilities based on non-imputed predictor information and rank
-the results by their predictive ability.
-```{r Non-Imputed-Pred-Ability, cache = TRUE}
-loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  mutate(Predictor = as.factor(Predictor)) %>%
-  droplevels() %>%
-  group_by(Predictor, Trait) %>%
-  summarize(Pred_Ability = cor(y, yhat)) %>%
-  arrange(desc(Pred_Ability)) %>%
-  rename("Predictive Ability" = Pred_Ability) %>%
-  knitr::kable()
-```
-
-## Predicted vs. residuals
+## Reduced Data
+### Predicted vs. residuals
 ```{r Complete-Data-Residuals, cache = TRUE, dependson = "Rename-Traits"}
@@
 
-## Predicted vs. Observed
+### Predicted vs. Observed
 ```{r Complete-Data-y-yHat, cache = TRUE, dependson = "Rename-Traits"}
@@
 ```
+
+## Predictive Ability for 'Full' vs 'Reduced' Data
+```{r Non-Imputed-Pred-Ability, cache = TRUE}
+loocv_tbl_cap <- paste("Predictive abilities for the 'reduced' set of hybrids",
+                       "using two different predictors and seven agronomic",
+                       "traits. Results are ranked by predictive ability.")
+pred_reduced <- loocv_clean %>%
+  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
+  mutate(Predictor = as.factor(Predictor)) %>%
+  droplevels() %>%
+  group_by(Predictor, Trait) %>%
+  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
+  spread(key = "Predictor", value = "Pred_Ability") %>%
+  rename(Reduced_mRNA = mRNA,
+         Reduced_SNP = SNP)
+```
+
+```{r Imputed-Pred-Ability, cache = TRUE}
+run_log <- fread("./data/processed/prediction_log/log_list.txt")
+
+# LOOCV 
+full_loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
+full_loocv <- lapply(seq_along(full_loocv_nms), FUN = function(i) {
+  readRDS(paste0("./data/processed/predictions/", full_loocv_nms[i], ".RDS"))
+})
+full_loocv <- rbindlist(full_loocv)
+
+full_loocv_clean <- full_loocv %>%
+  mutate(Trait = as.factor(Trait),
+         Trait = fct_recode(Trait, 
+                            "FAT" = "FETT",
+                            "PRO" = "RPR",
+                            "DMY" = "GTM",
+                            "DMC" = "GTS",
+                            "SUG" = "XZ"),
+         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
+                                      "SUG")),
+         Residuals = y - yhat,
+         Predictor = as.factor(as.character(Imputation)),
+         Predictor = fct_recode(Predictor,
+                                "mRNA" = "TRUE",
+                                "SNP" = "FALSE")
+  )
+full_loocv_tbl_cap <- paste(
+  "Predictive abilities for the 'full' and the 'reduced' set of hybrids using",
+  "two different predictors and seven agronomic traits."
+  )
+full_loocv_clean %>%
+  droplevels() %>%
+  group_by(Predictor, Trait) %>%
+  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
+  spread(key = "Predictor", value = "Pred_Ability") %>%
+  rename(Full_mRNA = mRNA,
+         Full_SNP = SNP) %>%
+  full_join(y = pred_reduced, by = "Trait") %>%
+  knitr::kable(caption = full_loocv_tbl_cap, booktabs = TRUE)
+```

### 2016-09-27 reports/na_fraction_pheno_predability.Rmd
Implement efficient LOOCV

--- 2016-09-20 reports/na_fraction_pheno_predability.Rmd
@@
                "viridis", "tidyr")
-knit_hooks$set(inline = function(x) {
+knitr::knit_hooks$set(inline = function(x) {
   prettyNum(x, big.mark = ",", digits = 2)
@@
 ```
-
-## single-step Prediction for the 'full' Set
-```{r Imputed-Pred-Ability, cache = TRUE}
-run_log <- fread("./data/processed/prediction_log/log_list.txt")
-
-# LOOCV 
-full_loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
-full_loocv <- lapply(seq_along(full_loocv_nms), FUN = function(i) {
-  readRDS(paste0("./data/processed/predictions/", full_loocv_nms[i], ".RDS"))
-})
-full_loocv <- rbindlist(full_loocv)
-
-full_loocv_clean <- full_loocv %>%
-  mutate(Trait = as.factor(Trait),
-         Trait = fct_recode(Trait, 
-                            "FAT" = "FETT",
-                            "PRO" = "RPR",
-                            "DMY" = "GTM",
-                            "DMC" = "GTS",
-                            "SUG" = "XZ"),
-         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
-                                      "SUG")),
-         Residuals = y - yhat,
-         Predictor = as.factor(as.character(Imputation)),
-         Predictor = fct_recode(Predictor,
-                                "mRNA" = "TRUE",
-                                "SNP" = "FALSE")
-  )
-full_loocv_tbl_cap <- paste("Predictive abilities for the 'full' set of hybrids",
-                            "using two different predictors and seven agronomic",
-                            "traits. Results are ranked by predictive ability.")
-full_loocv_clean %>%
-  droplevels() %>%
-  group_by(Predictor, Trait) %>%
-  summarize(Pred_Ability = cor(y, yhat)) %>%
-  arrange(desc(Pred_Ability)) %>%
-  rename("Predictive Ability" = Pred_Ability) %>%
-  knitr::kable(caption = full_loocv_tbl_cap, booktabs = TRUE)
-```

### 2016-09-27 reports/transcriptomic_data.Rmd
Implement efficient LOOCV

@@
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
-                      warning = FALSE, message = FALSE)
+                      warning = FALSE, message = FALSE, echo = TRUE)
 ```

### 2016-09-27 analysis/impute_single_mrna.R

@@ -1,18 +1,12 @@
 # Genotype-Wise mRNA-Imputation
-```{r echo = FALSE}
-pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis",
-               "forcats", "knitr", "rmarkdown", "purrr", "rprojroot",
-               "methods", "parallel")
-knitr::knit_hooks$set(inline = function(x) {
-  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
-})
-knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
-                      warning = FALSE, message = FALSE, echo = TRUE)
-```
+pacman::p_load("sspredr", "tidyverse", "forcats", "purrr", "methods", 
+               "parallel")
+if (isTRUE(interactive())) {
+  use_cores <- 3L
+} else {
+  use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
+}  
 
-
-```{r Predictor-Data, cache = TRUE}
 # Common genotypes
 genos <- readRDS("./data/processed/common_genotypes.RDS")
 dent <- as.character(genos$Dent$mrna)
@@ -52,13 +46,9 @@ snp_lst <- snp %>%
   split(.$Group) %>%
   map(rm_grp) %>%
   map(~as.matrix(.))
-```
-
 
 
 ## Function for the imputation of a single mRNA.
-```{r Impute-mRNA-Function}
-## Function for the imputation of a single mRNA
 impute_mrna_genotype <- function(x, y) {
   geno <- rownames(x)
   # Names of genotypes for which transcriptomic records exist
@@ -123,46 +113,21 @@ shuffled_imputation <- function(x, y) {
     rownames_to_column(var = "Number_NA_Geno") %>%
     rename(Value = value)
 }
-```
 
 
 ## Correlations between vectors of original and imputed mRNAs.
 ### General idea:
-*    Impute transcriptomic data for different numbers of genotypes (from 1 to
-     number of genotypes with transcriptomic data).
-*    For each number of genotypes without transcriptomic data, randomly sample 
-     1000 sets of genotypes, which will be imputed.
-     This will allow us to aggregate and summarize the correlation between 
-     observed and imputed mRNAs with high confidence.
-
-
-The computations were run using multiple cores with 
[…]

### 2016-09-27 moab_scripts/impute_single_mrna.sh

@@ -1,0 +1,52 @@
+#!/bin/sh 
+########## Begin MOAB/Slurm header ##########
+#
+# Give job a reasonable name
+#MOAB -N impute_single_mrna_geno_report
+#
+# Request number of nodes and CPU cores per node for job
+#MOAB -l nodes=1:ppn=16
+#
+# Request memory per process
+#MOAB -l pmem=2000mb
+#
+# Estimated wallclock time for job
+#MOAB -l walltime=00:00:40:00
+#
+# Job submission directory
+#MOAB -d /pfs/work2/workspace/scratch/ho_westhues-gamazon-0/gamazon
+#
+# Standard output naming
+#MOAB -o $(JOBNAME)_$(JOBID)
+#
+# Queue: Use a single node 
+#MOAB -q singlenode
+#
+# Send email when job begins (b), aborts (a) and ends (e)
+#MOAB -m a
+#
+# Write standard output (o) and errors (e) to the same file
+#MOAB -j oe
+#
+##### **********************************************************************
+########### End MOAB header ##########
+
+echo "Working Directory:                    $PWD"
+echo "Running on host                       $HOSTNAME"
+echo "Job id:                               $MOAB_JOBID"
+echo "Job name:                             $MOAB_JOBNAME"
+echo "Number of nodes allocated to job:     $MOAB_NODECOUNT"
+echo "Number of cores allocated to job:     $MOAB_PROCCOUNT"
+
+# Setup R Environment
+module load math/R/3.2.1
+
+# Set-up program
+startprog="Rscript --no-save --no-restore --slave\
+           ./analysis/impute_single_mrna.R"
+
+# Start program
+echo $startprog
+echo $(date)
+exec $startprog
+echo $(date)

### 2016-10-13 reports/loocv_example.Rmd
Add Claas' fast-loocv code

@@
 `r round(r, digits = 2)`.
+
+# Claas Heuer, October 2016
+#
+# Fast implementation of leave-one-out cross-validation
+# without reestimating variance components in a ridge
+# regression framework.
+# Based on Rohan's paper
+
+library(pacman)
+p_load(rrBLUP, Matrix)
+
+LOOCVFast <- function(y, M) {
+
+	if(!is.vector(y) | !is.numeric(y)) stop("y has to be a numeric vector")
+	if(anyNA(y)) stop("no NAs allowed in y")
+	if(nrow(M) != length(y)) stop("dimensions for y and M dont match")
+	if(anyNA(M)) stop("no NAs allowed in M")
+
+	# check dimensions for most efficient scheme
+	if(nrow(M) < ncol(M)) {
+
+	G <- tcrossprod(M)
+	L <- t(chol(G))
+
+	} else {
+
+		L <- M
+
+	}
+
+	mod <- mixed.solve(y = y, Z = L)
+
+	bs <- c(mod$beta, mod$u)
+
+	vA <- mod$Vu
+	vE <- mod$Ve
+	lambda <- vE / vA
+	D <- c(0, rep(lambda, ncol(L)))
+
+	X <- cbind(1, L)
+	Htemp <- crossprod(X)
+	diag(Htemp) <- diag(Htemp) + D
+	H <- X %*% solve(Htemp) %*% t(X)
+
+	E <- (y - X %*% bs) / (1 - diag(H))
+	yhat <- y - as.numeric(E)
+
+	return(list(y = y, yhat = yhat, e = as.numeric(E), cor = cor(y, yhat)))
+
+}

### 2016-10-13 analysis/fast_loocv.R

@@ -1,0 +1,78 @@
+# Efficient LOOCV With Prediction Errors
+pacman::p_load("BGLR")
+data("mice")
+X <- scale(mice.X)
+h2 <- 0.5
+QTL <- seq(from = 50, to = ncol(X), length = 20)
+nQTL <- length(QTL)
+n <- nrow(X)
+b <- rep(1, nQTL) * sqrt(h2 / nQTL)
+signal <- X[, QTL] %*% b
+error <- rnorm(n, sd = sqrt(1 - h2))
+y <- signal + error
+
+## Using a flat prior
+fm1 <- BGLR(y = y, ETA = list(list(X = t(chol(tcrossprod(X))),
+                                   model = "BRR")), 
+            nIter = 10000, burnIn = 2000, verbose = TRUE)
+X_prime <- cbind(matrix(1, nrow = nrow(X), ncol = 1),
+                 X)
+b_hat <- fm1$ETA[[1]]$b
+mu_hat <- fm1$mu
+b_hat_prime <- matrix(c(mu_hat, b_hat), ncol = 1)
+lambda <- c(fm1$varE / fm1$ETA[[1]]$varB, 1)
+Diag <- cbind(diag(nrow = ncol(X), ncol = ncol(X)),
+              matrix(1, nrow = ncol(X), ncol = 1))
+Diag <- rbind(Diag,
+              matrix(1, nrow = 1, ncol = ncol(Diag)))
+H <- X_prime %*% solve(t(X_prime) %*% X_prime + Diag * lambda) %*% t(X_prime)
+e_hat <- vapply(seq_len(nrow(X_prime)), FUN = function(j) {
+t_x_prime <- X_prime[j, , drop = FALSE]
+numerator <- as.numeric(y[j] - t_x_prime %*% b_hat_prime)
+denominator <- 1 - H[j, j]
+numerator / denominator
+}, FUN.VALUE = numeric(1))
+press <- sum(e_hat ^ 2)
+rmse <- sqrt(press / length(e_hat))
+y_hat <- y - e_hat
+# Leave-one-out prediction accuracy
+r <- cor(y, y_hat)
+
+
+
+
+# Claas Heuer, October 2016
+#
+# Fast implementation of leave-one-out cross-validation
+# without reestimating variance components in a ridge
+# regression framework.
+# Based on Rohan's paper
+library(pacman)
+p_load(rrBLUP, Matrix)
[…]

### 2016-10-13 reports/loocv_issues.Rmd
Add report on my issues with fernando loocv

---
title: "Fernando LOOCV"
author: "Matthias Westhues"
date: "`r Sys.Date()`"
output: pdf_document
---

Meine Annahme ist, dass die effiziente Implementierung von LOOCV (Cheng et al.,
2016) den gesamten Stichprobenumfang (mit Ausnahme der einen Hybride im test 
set) als training set nutzt.
Allerdings wollen wir LOOCV ja nur auf T0-Hybride anwenden.
In diesem Fall kann nicht jede Hybride, die zuvor Teil des training sets fuer 
die Vorhersage einer anderen Hybride war, vom training in das test set 
ueberfuehrt werden.
Ich habe das nachfolgend mal anhand eines kleinen Beispiels illustriert:

|:---:|:----:|
|ID   |Hybride|
|1    |Aa     |
|2    |Ab     |
|3    |Ac     |
|4    |Ba     |
|5    |Bb     |
|6    |Bc     |
|7    |Ca     |
|8    |Cb     |
|9    |Cc     |

: Hybride und ihre korrespondierenden IDs.

|:--------:|:------------:|
| Test Set | Training Set |
| 1        | 5, 6, 8, 9   |        
| 2        | 4, 6, 8, 9   |        
| 3        | 4, 5, 7, 8   |        
| 4        | 2, 3, 8, 9   |        
| 5        | 1, 3, 7, 9   |        
| 6        | 1, 2, 7, 8   |        
| 7        | 2, 3, 5, 6   |        
| 8        | 1, 3, 4, 6   |        
| 9        | 1, 2, 4, 5   |

: Hybride im test set und korrespondierende T0-Hybride im training set. Die Hybride sind mit ihren IDs aus der vorherigen Tabelle gekennzeichnet.

Hybride \#1 kann mit den Hybriden 5, 6, 8 und 9 vorhergesagt werden, aber die
Hybride \#5 kann nicht mit den Hybriden 6 und 8 vorhergesagt werden, Hybride 
\#6 kann nicht mit den Hybriden 5, 6 und 9 vorhergesagt werden, Hybride \#8 
kann nicht mit den Hybriden 5 und 9 vorhergesagt werden, und Hybride \#9 kann 
nicht mit den Hybriden 6 und 8 vorhergesagt werden.

### 2016-10-13 analysis/common_genotypes.R
Add pedigree data to common geno filtering

@@ -2,86 +2,80 @@
 # ensure that for the hybrid progeny of these parent lines, agronomic data are
 # available.
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("data.table")
+pacman::p_load("data.table", "tidyverse")
 
 smp_conv <- fread("./data/input/tGTP.txt", sep = "\t")
 id_gtp <- smp_conv[, .(id_GTP, pool_2), 
                    ][pool_2 %in% c("D", "F"), , ]
 id_gtp[, id_GTP := as.character(id_GTP), ]
 
-#--- Genomic data
-snp <- fread("./data/processed/uhoh_smp_id_P235-update.txt")
-snp_dent <- id_gtp[match(snp$id.GTP, id_gtp$id_GTP), , 
-                   ][pool_2 == "D", id_GTP, ]
-snp_flint <- id_gtp[match(snp$id.GTP, id_gtp$id_GTP), , 
-                    ][pool_2 == "F", id_GTP, ]
-snp_geno <- c(snp_dent, snp_flint)
+#--- Predictor data
+# Genomic data
+snp <- fread("./data/raw/uhoh_smp_id_P235-update.txt")
+snp_geno <- id_gtp %>%
+  filter(id_GTP %in% snp$id.GTP) %>%
+  mutate(Data_Type = "snp")
 
+# Pedigree data
+ped <- readRDS("./data/processed/ped-datafull-GTP.RDS") 
+ped_geno <- id_gtp %>%
+  filter(id_GTP %in% rownames(ped)) %>%
+  mutate(Data_Type = "ped")
+
+
+# Transcriptomic data
+mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
+mrna_geno <- id_gtp %>%
+  filter(id_GTP %in% colnames(mrna)) %>%
+  mutate(Data_Type = "mrna")
+
+# Concatenate predictor data 
+pred_geno <- rbindlist(list(snp_geno, ped_geno, mrna_geno))
+pred_geno <- pred_geno %>%
+  rename(G = id_GTP,
+         Pool = pool_2) %>%
+  mutate(Pool = ifelse(Pool == "D", yes = "Dent", no = "Flint")) %>%
+  as_tibble()
 
 #--- Agronomic traits
 traits <- c("GTM", "GTS", "ADF", "FETT", "RPR", "STA", "XZ")
-Pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
[…]

### 2016-10-21 reports/agronomic_data.Rmd
remove code boxes from report

@@
 pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "knitr", "tibble")
+               "viridis", "tidyr", "knitr", "tibble", "purrr")
 
@@
 })
-knitr::opts_knit$set(root.dir = "~/gamazon/",
-                     inline = TRUE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
@@
 genos <- readRDS("./data/processed/common_genotypes.RDS")
-hybrid <- genos$Hybrid
+hybrid <- genos %>%
+  filter(Data_Type == "agro", Pool == "Hybrid") %>%
+  .$G
 all_pheno <- pheno %>%
@@
 # records.
-dent <- genos$Dent$mrna
-flint <- genos$Flint$mrna
-hybrid <- genos$Hybrid %>%
+mrna_inbreds <- genos %>%
+  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
+  split(.$Pool) %>%
+  map("G")
+hybrid <- hybrid %>%
   as_data_frame() %>%
   separate(value, into = c("DF", "Dent", "Flint")) %>%
-  mutate(Avail_Dent = ifelse(Dent %in% dent, yes = 1, no = 0),
-         Avail_Flint = ifelse(Flint %in% flint, yes = 1, no = 0),
+  mutate(Avail_Dent = ifelse(Dent %in% mrna_inbreds$Dent, yes = 1, no = 0),
+         Avail_Flint = ifelse(Flint %in% mrna_inbreds$Flint, yes = 1, no = 0),
          Tested_Parents = Avail_Dent + Avail_Flint) %>%

### 2016-10-21 reports/data_overview.Rmd
remove code boxes from report

@@
 })
-knitr::opts_knit$set(root.dir = "~/gamazon/",
-                     inline = TRUE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
@@
 
-full_hybrid <- genos$Hybrid
+full_hybrid <- genos %>%
+  filter(Pool == "Hybrid") %>%
+  .$G
 all_pheno <- pheno %>%
@@
 # Hybrids of parents with records for both endophenotypes
-red_dent <- Reduce(intersect, genos[["Dent"]][c("snp", "mrna")])
-red_flint <- Reduce(intersect, genos[["Flint"]][c("snp", "mrna")])
-red_hybrid <- genos$Hybrid %>%
+red_inbreds <- genos %>% 
+  filter(Pool != "Hybrid", Data_Type %in% c("snp", "mrna")) %>%
+  split(.$Pool) %>%
+  map(~split(., .$Data_Type)) %>%
+  at_depth(.depth = 2, "G") %>%
+  map(function(x) intersect(x$mrna, x$snp))
+
+red_hybrid <- full_hybrid %>%
   as_data_frame() %>%
   separate(value, into = c("DF", "Dent", "Flint")) %>%
-  mutate(Avail_Dent = ifelse(Dent %in% red_dent, yes = 1, no = 0),
-         Avail_Flint = ifelse(Flint %in% red_flint, yes = 1, no = 0),
+  mutate(Avail_Dent = ifelse(Dent %in% red_inbreds$Dent, yes = 1, no = 0),
+         Avail_Flint = ifelse(Flint %in% red_inbreds$Flint, yes = 1, no = 0),
          Tested_Parents = Avail_Dent + Avail_Flint) %>%
@@
   .$Hybrid
-full_hybrid <- genos$Hybrid
-genos$Hybrid <- NULL
-genos$Hybrid$snp <- full_hybrid
-genos$Hybrid$mrna <- red_hybrid
+genos <- genos %>%
+  rbind(tibble(G = red_hybrid,
+               Pool = "Hybrid",
+               Data_Type = "mrna")
+        )
 ```
@@
 analyses.
-Both, genomic and transcriptomic records, were collected for inbred parent lines
-of `r length(red_hybrid)` hybrids.
+Genomic, pedigree and transcriptomic records, were collected for inbred parent 
+lines of `r length(red_hybrid)` hybrids.
 
@@
 genos %>% 
-  at_depth(.depth = 2, .f = length) %>%
-  bind_rows() %>%
-  dplyr::select(-agro) %>%
-  mutate(Type = names(genos)) %>%
-  dplyr::select(Type, snp, mrna) %>%
-  rename("Genomic Data" = snp,
-         "Transcriptomic Data" = mrna) %>%
+  mutate(Data_Type = fct_recode(Data_Type,
+                                Agronomic = "agro",
+                                Transcriptomic = "mrna",
+                                Genomic = "snp",
+                                Pedigree = "ped")) %>%
+  group_by(Pool, Data_Type) %>%
+  count() %>%
+  rename(`Data Type` = Data_Type) %>%
   knitr::kable(caption = geno_tab_cap, booktabs = TRUE)

### 2016-10-21 reports/impute_single_mrna.Rmd
remove code boxes from report

--- 2016-09-26 reports/impute_single_mrna.Rmd
@@
 # Genotype-Wise mRNA-Imputation
+```{r echo = FALSE}
+pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis",
+               "forcats", "knitr", "rmarkdown", "purrr", "rprojroot",
+               "methods", "parallel")
+knitr::knit_hooks$set(inline = function(x) {
+  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
+})
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
+knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
+                      warning = FALSE, message = FALSE, echo = FALSE)
+```
 
-## Function for the imputation of a single mRNA.
+```{r Predictor-Data, cache = TRUE}
+# Common genotypes
+genos <- readRDS("./data/processed/common_genotypes.RDS")
+mrna_inbreds <- genos %>%
+  filter(Data_Type == "mrna") %>%
+  split(.$Pool) %>%
+  map("G")
+
+# Remove the 'Group' variable after splitting Dent and Flint data.
+rm_grp <- function(x) {
+  x$Group <- NULL
+  x
+}
+
+# mRNA
+mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
+mrna <- t(mrna)
+mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
+mrna_lst <- mrna %>%
+  as_data_frame() %>%
+  mutate(G = rownames(mrna)) %>%
+  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group)) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "G") %>%
+  split(.$Group) %>%
+  map(rm_grp) %>%
+  map(~as.matrix(.))
+
+# SNPs
+snp <- readRDS("./data/processed/snp_mat.RDS")
+snp_lst <- snp %>%
+  as_data_frame() %>%
+  mutate(G = rownames(snp)) %>%
+  filter(G %in% (mrna_inbreds %>% flatten_chr())) %>%
+  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group)) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "G") %>%
+  split(.$Group) %>%
+  map(rm_grp) %>%
+  map(~as.matrix(.))
+
+# Pedigree
+ped <- readRDS("./data/processed/ped-datafull-GTP.RDS")
+ped_lst <- ped %>%
+  mutate(G = rownames(ped)) %>%
+  filter(G %in% (mrna_inbreds %>% flatten_chr())) %>%
+  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group)) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "G") %>%
+  split(.$Group) %>%
+  map(rm_grp) %>%
+  map(as.matrix) %>%
+  map(function(x) x[, match(rownames(x), colnames(x))]) %>%
+  map(function(x) x * 2) %>%
+  map(function(x) {
+    x[is.na(x)] <- 0
+    x
+  })
+```
+
 ```{r Impute-mRNA-Function}
 ## Function for the imputation of a single mRNA
-impute_one_mrna <- function(x, y) {
+impute_mrna_genotype <- function(x, y) {
   geno <- rownames(x)
@@
   w1 <- A12 %*% solve(A22) %*% Z2 %*% M2
-  as.numeric(w1)
+  as.matrix(w1)
+}
+
+# Shuffled version of the imputation function above.
+shuffled_imputation <- function(x, y) {
+  cor_lst <- lapply(seq(from = 1, to = nrow(y) - 2), FUN = function(i) {
+    idx <- sample(nrow(y), size = i, replace = FALSE)
+    z <- y[-idx, ]
+    imp_mrna <- impute_mrna_genotype(x = x, y = z)
+    orig_mrna <- y[match(rownames(imp_mrna), rownames(y)), , drop = FALSE]
+    stopifnot(identical(rownames(imp_mrna), rownames(orig_mrna)))
+    cor(c(imp_mrna), c(orig_mrna))
+  })
+  cor_lst %>%
+    flatten_dbl() %>%
+    tbl_df() %>%
+    rownames_to_column(var = "Number_NA_Geno") %>%
+    rename(Value = value)
 }
@@
 
-Plot the distribution of correlations between vectors of original and imputed
-mRNAs.
-```{r Correlation-Imputed-vs-NonImputed}
-#imp_mrna <- vapply(seq_len(nrow(y)), FUN = function(i) {
-#  z <- y[-i, ]
-#  imp_mrna <- impute_one_mrna(x, z)
-#  cor(y[i, ], imp_mrna)
-#}, FUN.VALUE = numeric(1))
-#hist(imp_mrna)
+## Correlations between vectors of original and imputed mRNAs.
+### General idea:
+*    Impute transcriptomic data for different numbers of genotypes (from 1 to
+     number of genotypes with transcriptomic data).
+*    For each number of genotypes without transcriptomic data, randomly sample 
+     1000 sets of genotypes, which will be imputed.
+     This will allow us to aggregate and summarize the correlation between 
+     observed and imputed mRNAs with high confidence.
+
+The computations were run using multiple cores with 
+[this script](../analysis/impute_single_mrna.R).
+```{r Shuffle-Genotypes-Correlate-Imputation, eval = FALSE}
+repl <- 1000
+donors <- c("snp_lst", "ped_lst")
+imp_lst <- lapply(donors, FUN = function(type) {
+  par_lst <- vector(mode = "list", length = repl)
+  names(par_lst) <- paste0("Replication_", seq_len(repl))
+  par_lst[] <- mclapply(par_lst, FUN = function(iter, ...) {
+    try(map2(.x = get(type), .y = mrna_lst, .f = shuffled_imputation))
+  }, mc.cores = use_cores)
+  par_lst
+})
+names(imp_lst) <- donors
 ```
+
+```{r Plot-Correlation-Imputed-vs-NonImputed, cache = TRUE, fig.width = 12, fig.height = 12, fig.cap = "Pearson correlation coefficients between original and imputed mRNA genotypes for two heterotic groups and two sources of information (i.e.
+genomic and pedigree)."}
+par_lst <- readRDS("./data/derived/impute_single_mrna.RDS")
+par_lst %>%
+  map(~keep(., is_list)) %>%
+  at_depth(.depth = 2, ~bind_rows(., .id = "Group")) %>%
+  map(~bind_rows(., .id = "Replication")) %>%
+  bind_rows(.id = "Data_Type") %>%
+  mutate(Data_Type = fct_recode(Data_Type,
+                                Genomic = "snp_lst",
+                                Pedigree = "ped_lst")) %>%
+  mutate(Number_NA_Geno = as.factor(Number_NA_Geno)) %>%
+  mutate(Number_NA_Geno = fct_inorder(Number_NA_Geno)) %>%
+  ggplot(aes(x = Number_NA_Geno, y = Value)) +
+  geom_boxplot() +
+  facet_wrap(Data_Type ~ Group) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
+  xlab("Number of Genotypes Without mRNA-Data") +
+  ylab("Correlation Between Original and Imputed mRNA BLUEs")
+```

### 2016-10-21 reports/intro.Rmd
remove code boxes from report

@@
----
-output: pdf_document
----
 # Introduction

### 2016-10-21 reports/loocv_issues.Rmd
remove code boxes from report

@@
 
-Meine Annahme ist, dass die effiziente Implementierung von LOOCV (Cheng et al.,
-2016) den gesamten Stichprobenumfang (mit Ausnahme der einen Hybride im test 
-set) als training set nutzt.
-Allerdings wollen wir LOOCV ja nur auf T0-Hybride anwenden.
-In diesem Fall kann nicht jede Hybride, die zuvor Teil des training sets fuer 
-die Vorhersage einer anderen Hybride war, vom training in das test set 
-ueberfuehrt werden.
-Ich habe das nachfolgend mal anhand eines kleinen Beispiels illustriert:
+My assumption is that the efficient implementation of LOOCV proposed by Cheng
+et al. (2016) uses the entire sample space except for a single hybrid that 
+is used as a test set.
+However, our goal is to apply LOOCV only to the T0 hybrids.
+In this case, not very hybrid, that was previously part of the training set to
+predict another hybrid, can be moved to the test set (see the following 
+example).
 
 |:---:|:----:|
-|ID   |Hybride|
+|ID   |Hybrid|
 |1    |Aa    |
@@
 
-: Hybride und ihre korrespondierenden IDs.
+: Hybrids and their corresponding IDs.
 
@@
 
-: Hybride im test set und korrespondierende T0-Hybride im training set. Die Hybride sind mit ihren IDs aus der vorherigen Tabelle gekennzeichnet.
+: Hybrids in the test set and their corresponding T0-Hybrids in the training 
+set. The hybrids are denoted by their IDs from the previous table.
 
-Hybride \#1 kann mit den Hybriden 5, 6, 8 und 9 vorhergesagt werden, aber die
-Hybride \#5 kann nicht mit den Hybriden 6 und 8 vorhergesagt werden, Hybride 
-\#6 kann nicht mit den Hybriden 5, 6 und 9 vorhergesagt werden, Hybride \#8 
-kann nicht mit den Hybriden 5 und 9 vorhergesagt werden, und Hybride \#9 kann 
-nicht mit den Hybriden 6 und 8 vorhergesagt werden.
+Hybrid \#1 can be predicted with hybrids 5, 6, 8 and 9, however, hybrid \#5 
+cannot be predicted with hybrids 6 and 8, hybrid \#6 cannot be predicted with 
+hybrids 5, 6 and 9, hybrid \#8 cannot be predicted with hybrids 5 and 9, and 
+hybrid \#9 cannot be predicted with hybrids 6 und 8.
+
+## Proposed Solution

### 2016-10-21 reports/loocv_reduced.Rmd
remove code boxes from report

@@
 })
-knitr::opts_knit$set(root.dir = "~/gamazon/",
-                     inline = TRUE,
-                     verbose = FALSE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
@@
          Predictor = fct_recode(Predictor,
-                                "mRNA" = "mrna",
-                                "SNP" = "snp")
+                                "Transcriptomic" = "mrna",
+                                "Genomic" = "snp")
   )
@@
   spread(key = "Predictor", value = "Pred_Ability") %>%
-  rename(Reduced_mRNA = mRNA,
-         Reduced_SNP = SNP)
+  rename(Reduced_Transcriptomic = Transcriptomic,
+         Reduced_Genomic = Genomic)
 ```
@@
          Predictor = fct_recode(Predictor,
-                                "mRNA" = "TRUE",
-                                "SNP" = "FALSE")
+                                "Transcriptomic" = "TRUE",
+                                "Genomic" = "FALSE")
   )
@@
   spread(key = "Predictor", value = "Pred_Ability") %>%
-  rename(Full_mRNA = mRNA,
-         Full_SNP = SNP) %>%
+  rename(Full_Transcriptomic = Transcriptomic,
+         Full_Genomic = Genomic) %>%
   full_join(y = pred_reduced, by = "Trait") %>%

### 2016-10-21 reports/na_fraction_pheno_predability.Rmd
remove code boxes from report

@@
          Flint_NA_Fraction = as.numeric(as.character(Flint_NA_Fraction))) %>%
-  filter(Trait == "DMY", Predictor == "mRNA") %>%
+  filter(Trait == "DMY", Predictor == "Transcriptomic") %>%
   droplevels() %>%

### 2016-10-21 reports/transcriptomic_data.Rmd
remove code boxes from report

--- 2016-09-26 reports/transcriptomic_data.Rmd
@@
 })
-knitr::opts_knit$set(root.dir = "~/gamazon/", inline = TRUE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
-                      warning = FALSE, message = FALSE)
+                      warning = FALSE, message = FALSE, echo = FALSE)
 ```
@@
 genos <- readRDS("./data/processed/common_genotypes.RDS")
-dent <- as.character(genos$Dent$mrna)
-flint <- as.character(genos$Flint$mrna)
+mrna_inbreds <- genos %>%
+  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
+  split(.$Pool) %>%
+  map("G")
 
@@
   mutate(G = rownames(mrna)) %>%
-  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
+  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
          Group = as.factor(Group)) %>%
@@
 
-## Tidy-up data
 ```{r Tidy-Data, cache = TRUE}
@@
 ## Skewness
-Compute the skewness and the Shapiro-Wilk test of normality for each mRNA
 ```{r Check-Normality, cache = TRUE}
@@
 
-Plot the distribution of skewness values across all mRNAs
-```{r, mRNA-Skewness, cache = TRUE}
-mrna_skew %>%
-  ggplot(aes(x = Skewness, fill = Type)) +
-  geom_density(alpha = 0.2) +
-  facet_grid(. ~ Group) +
-  theme(legend.position = "top")
-
+```{r, mRNA-Skewness, cache = TRUE, fig.cap = "Distribution of skewness values across all mRNAs, separately for transformed and raw BLUEs."}
 mrna_skew %>%
@@
   facet_grid(. ~ Group) +
-  theme(legend.position = "top")
+  theme(legend.position = "top") +
+  ylab("Density")
 ```
 
-Find the least skewed/non-normally distributed mRNAs.
-```{r Association-Shapiro-Wilks-Skewness, cache = TRUE}
+```{r Association-Shapiro-Wilks-Skewness, cache = TRUE, fig.cap = "Shapiro-Wilks Test Statistic vs absolute skewness for transcriptomic BLUEs depending on whether the BLUEs were transformed or not."}
 mrna_skew %>%

### 2016-10-21 analysis/common_genotypes.R

@@ -16,11 +16,23 @@ snp_geno <- id_gtp %>%
   filter(id_GTP %in% snp$id.GTP) %>%
   mutate(Data_Type = "snp")
 
-# Pedigree data
+# There is only one genotype for which we have pedigree records that does not
+# also have genomic records. Hence, there is almost no difference between 
+# the extent of both datasets and we will reduce the pedigree data set to its
+# intersection with the genomic data set to alleviate further coding.
 ped <- readRDS("./data/processed/ped-datafull-GTP.RDS") 
 ped_geno <- id_gtp %>%
   filter(id_GTP %in% rownames(ped)) %>%
   mutate(Data_Type = "ped")
+split_pred <- snp_geno %>%
+  rbind(ped_geno) %>%
+  split(.$Data_Type) 
+common_snp_ped <- split_pred %>% map("id_GTP") %>%
+  reduce(intersect)
+snp_ped <- split_pred %>%
+  map(function(x) x[x$id_GTP %in% common_snp_ped, ]) %>%
+  bind_rows() %>%
+  as_tibble()
 
 
 # Transcriptomic data
@@ -30,7 +42,7 @@ mrna_geno <- id_gtp %>%
   mutate(Data_Type = "mrna")
 
 # Concatenate predictor data 
-pred_geno <- rbindlist(list(snp_geno, ped_geno, mrna_geno))
+pred_geno <- rbindlist(list(snp_ped, mrna_geno))
 pred_geno <- pred_geno %>%
   rename(G = id_GTP,
          Pool = pool_2) %>%
@@ -43,7 +55,10 @@ pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
 # Keep only factorial hybrids and unique records for each trait.
 agro_hybrid <- pheno %>%
   as_tibble() %>%
-  filter(Trait %in% traits, check == 0, dent.GTP != 0, flint.GTP != 0) %>%
+  filter(Trait %in% traits, 
+         check == 0,
+         dent.GTP %in% (pred_geno %>% filter(Pool == "Dent") %>% .$G),
+         flint.GTP %in% (pred_geno %>% filter(Pool == "Flint") %>% .$G)) %>%
   select(G, Trait) %>%
   split(.$Trait) %>%
   map("G") %>%

### 2016-10-21 analysis/fast_loocv.R

@@ -1,5 +1,5 @@
 # Efficient LOOCV With Prediction Errors
-pacman::p_load("BGLR")
+pacman::p_load("BGLR", "tidyverse")
 data("mice")
 X <- scale(mice.X)
 h2 <- 0.5
@@ -11,10 +11,20 @@ signal <- X[, QTL] %*% b
 error <- rnorm(n, sd = sqrt(1 - h2))
 y <- signal + error
 
+G <- X %>%
+  list() %>%
+  map(~tcrossprod(scale(.)) / ncol(.)) %>%
+  map(chol) %>%
+  map(t) %>%
+  .[[1]]
+
 ## Using a flat prior
-fm1 <- BGLR(y = y, ETA = list(list(X = t(chol(tcrossprod(X))),
-                                   model = "BRR")), 
-            nIter = 10000, burnIn = 2000, verbose = TRUE)
+system.time(fm1 <- BGLR(y = y, 
+                        ETA = list(list(X = G, model = "BRR")), 
+                        nIter = 30000, 
+                        burnIn = 15000,
+                        verbose = FALSE,
+                        saveAt = "./tmp/"))
 X_prime <- cbind(matrix(1, nrow = nrow(X), ncol = 1),
                  X)
 b_hat <- fm1$ETA[[1]]$b

### 2016-10-21 analysis/impute_single_mrna.R

@@ -1,16 +1,16 @@
 # Genotype-Wise mRNA-Imputation
 pacman::p_load("sspredr", "tidyverse", "forcats", "purrr", "methods", 
                "parallel")
-if (isTRUE(interactive())) {
-  use_cores <- 3L
-} else {
-  use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
-}  
+use_cores <- 3L
 
 # Common genotypes
 genos <- readRDS("./data/processed/common_genotypes.RDS")
-dent <- as.character(genos$Dent$mrna)
-flint <- as.character(genos$Flint$mrna)
+dent <- genos %>% 
+  filter(Pool == "Dent", Data_Type == "mrna") %>%
+  .$G
+flint <- genos %>%
+  filter(Pool == "Flint", Data_Type == "mrna") %>%
+  .$G
 
 # Remove the 'Group' variable after splitting Dent and Flint data.
 rm_grp <- function(x) {
@@ -46,6 +46,25 @@ snp_lst <- snp %>%
   split(.$Group) %>%
   map(rm_grp) %>%
   map(~as.matrix(.))
+
+# Pedigree
+ped <- readRDS("./data/processed/ped-datafull-GTP.RDS")
+ped_lst <- ped %>%
+  mutate(G = rownames(ped)) %>%
+  filter(G %in% c(dent, flint)) %>%
+  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group)) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "G") %>%
+  split(.$Group) %>%
+  map(rm_grp) %>%
+  map(as.matrix) %>%
+  map(function(x) x[, match(rownames(x), colnames(x))]) %>%
+  map(function(x) x * 2) %>%
+  map(function(x) {
+    x[is.na(x)] <- 0
+    x
+  })
 
 
 ## Function for the imputation of a single mRNA.
@@ -124,10 +143,17 @@ shuffled_imputation <- function(x, y) {
 #  This will allow us to aggregate and summarize the correlation between 
 #  observed and imputed mRNAs with high confidence.
 repl <- 1000
-par_lst <- vector(mode = "list", length = repl)
-names(par_lst) <- paste0("Replication_", seq_len(repl))
-par_lst[] <- mclapply(par_lst, FUN = function(iter, ...) {
[…]

### 2016-10-21 analysis/spaced_snps.R
select fixed number of equally spaced snps

@@ -1,0 +1,106 @@
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load("tidyverse", "data.table", "rvest", "magrittr")
+
+## -- DATA SELECTION -----------------------------------------------------
+genos <- readRDS("./data/processed/common_genotypes.RDS")
+
+# snp
+snp <- readRDS("./data/processed/snp_mat.RDS")
+snp_nms <- genos %>%
+  filter(Data_Type == "snp") %>% 
+  .$G
+snp <- snp %>%
+  .[match(snp_nms, rownames(.)), ]
+
+# Load data frame with meta information on marker names and their positions in 
+# the genome.
+snp_map <- fread("./data/raw/marker_map.txt", sep = "\t")
+snp_map <- snp_map[complete.cases(snp_map), ]
+snp_map <- snp_map[snp_map$chromosome != 0, ]
+map_marker_names <- as.character(snp_map[["markername"]])
+map_marker_names <- gsub(pattern = "[.]", replacement = "_",
+                         x = map_marker_names)
+map_marker_names <- gsub(pattern = "[-]", replacement = "_", 
+                         x = map_marker_names)
+snp_map[, markername := map_marker_names]
+
+# Get the length of each of the ten maize chromosomes. We need this information
+# so that we can align the chromosomes sequentially and treat them like an 
+# entire maize genome with a global length. Thereby, we can select equally 
+# spaced SNPs with respect to the genome instead of individual chromosomes.
+ensembl <- read_html("http://ensembl.gramene.org/Zea_mays/Info/Annotation/")
+chr_lengths <- ensembl %>%
+  html_nodes(., "#chromosome_table td:nth-child(2)") %>%
+  html_text(.) %>%
+  .[seq_len(10)] %>%
+  c(0, .) %>%
+  .[seq_len(10)] %>%
+  as.list() %>%
+  map(as.data.frame) %>%
+  bind_rows(.id = "chromosome") %>%
+  rename(Chr_Length = `.x[[i]]`) %>%
+  mutate(Chr_Length = gsub(",", replacement = "", x = Chr_Length),
+         Chr_Length = as.numeric(Chr_Length),
+         Chr_Length = cumsum(Chr_Length),
+         chromosome = as.integer(chromosome))
+
+# Create global SNP positions
+global_map <- snp_map %>%
+  left_join(y = chr_lengths, by = "chromosome") %>%
+  as_tibble() %>%
[…]

### 2016-10-21 analysis/predictor_subsetting.R
subset predictors to compare predabilities

@@ -1,0 +1,110 @@
+# Goal: Generate nine subsets of predictors for later comparisons between their
+# predictive abilities:
+# a. 100% SNP
+# b. 100% pedigree
+# c. 42% mRNA
+# d. 42% pedigree
+# e. 42% SNP
+# f. 100% pedigree & 77% SNP
+# g. 100% pedigree & 42% mRNA
+# h. 100% SNP & 42% mRNA
+# i. 100% pedigree & 77% SNP & 42% mRNA
+#
+# Make comparisons between the following scenarios:
+# a, b, f, g, h, i
+# c, d, e
+#
+# Throughout this script, digits behind the variable name of each predictor
+# (ped: pedigree, snp: genomic, mrna: transcriptomic) indicate the size of 
+# the subset relative to the full set of genotypes (n = 245). For example,
+# we have transcriptomic data for 103 genotypes, which are roughly 42% of 
+# 245 genotypes, hence the name 'mrna42' and so forth.
+
+pacman::p_load("tidyverse", "data.table")
+devtools::install_github("mwesthues/sspredr")
+
+
+
+# -- Genotype information.
+# Information on the set of genotypes.
+genos <- readRDS("./data/processed/common_genotypes.RDS")
+genos %>%
+  filter(Pool %in% c("Dent", "Flint")) %>%
+  group_by(Pool, Data_Type) %>%
+  count()
+
+# Get the names of the Dent Hybrid-parents.
+geno <- genos %>%
+  filter(Pool == "Hybrid") %>%
+  .$G %>%
+  map(~stringr::str_split(., pattern = "_")) %>%
+  at_depth(.depth = 2, .f = 2) %>%
+  flatten() %>%
+  as_vector()
+
+# -- Load predictor data.
+# pedigree
+ped100 <- readRDS("./data/processed/ped-datafull-GTP.RDS")
+ped_nms <- genos %>%
+  filter(Data_Type == "ped") %>%
+  .$G
+ped100 <- as.matrix(ped100) * 2
+ped100 <- ped100 %>%
+  .[rownames(.) %in% ped_nms, colnames(.) %in% ped_nms]
+ped100[is.na(ped100)] <- 0
+
+# snp
+snp100 <- readRDS("./data/processed/snp_mat.RDS")
+snp_nms <- genos %>%
+  filter(Data_Type == "snp") %>%
+  .$G
[…]

### 2016-10-21 analysis/ldocv.R
init potential ldocv implementation, Gianola and Schoen 2016

@@ -1,0 +1,193 @@
+# Leave-d-out cross-validation
+# Gianola et al. (2016), doi:10.1534/g3.116.033381
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load("tidyverse", "data.table", "sspredr", "BGLR", "Matrix")
+
+
+## -- DATA SELECTION -----------------------------------------------------
+genos <- readRDS("./data/processed/common_genotypes.RDS")
+# hybrids
+hybrid <- genos %>%
+  filter(Pool == "Hybrid") %>%
+  .$G
+
+# snp
+snp <- readRDS("./data/processed/snp_mat.RDS")
+snp_nms <- genos %>%
+  filter(Data_Type == "snp") %>% 
+  .$G
+snp <- snp %>%
+  .[match(snp_nms, rownames(.)), ]
+
+
+## -- PREDICTOR AND AGRONOMIC DATA PREPARATION ----------------------------
+# Agronomic data
+pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
+pheno <- pheno %>%
+  as_data_frame() %>%
+  filter(G %in% hybrid) %>%
+  dplyr::select(G, EST, Trait) %>%
+  spread(key = Trait, value = EST) %>%
+  dplyr::select(-ADL) %>%
+  as.data.frame %>%
+  remove_rownames() %>%
+  column_to_rownames(var = "G") %>%
+  as.matrix
+
+
+# Extract the names of the dent and flint parents, respectively.
+dent_flint_lst <- pheno %>%
+  rownames() %>%
+  as.list() %>%
+  map(~strsplit(., split = "_")) %>%
+  at_depth(.depth = 2, .f = function(x) x[c(2, 3)]) %>%
+  flatten() %>%
+  transpose()
+names(dent_flint_lst) <- c("Dent", "Flint")
+
+
+
+
+# Function for the setup of the feature matrix. 
+build_ZL <- function(M, Z, use_kernel = TRUE, bglr_model = "BRR") {
+  # Input tests
+  if (class(M) != "matrix") stop("'M' is not a matrix")
+  stopifnot(bglr_model %in% c("FIXED", "BRR", "BayesA", "BayesB", "BayesC",
+                              "RKHS"))
+  comgeno <- unique(rownames(Z))
+  M <- M[rownames(M) %in% comgeno, ]
+  # Sort
[…]

### 2016-10-21 analysis/predictor_subsetting.R
remove superfluous predictor subset

@@ -102,6 +102,7 @@ snp42 <- snp100 %>%
 # Store the new objects
 storage_dir <- "./data/derived/predictor_subsets/"
 ls() %>%
+  .[. != "snp35"] %>%
   .[stringr::str_detect(., pattern = "[:digit:]")] %>%
   as.list() %>%
   map_if(.p = ~get(.) %>% is.matrix,

### 2016-10-21 analysis/prediction.R
remove caret prediction option

@@ -12,11 +12,7 @@
 if (!require("pacman")) install.packages("pacman")
 if (!require("devtools")) install.packages("devtools")
 devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("BGLR", "caret", "corrplot", "Cubist", "data.table", "doMC", 
-               "dplyr", "e1071", "earth", "elasticnet", "gbm", "kernlab", 
-               "MASS", "Matrix", "matrixStats", "methods", "nnet", "parallel", 
-               "party", "partykit", "pls", "randomForest", "rpart", "RWeka", 
-               "testthat", "tibble", "tidyr")
+pacman::p_load("BGLR","data.table", "parallel", "tidyverse")
 pacman::p_load_gh("mwesthues/sspredr")
 
 
@@ -25,9 +21,6 @@ pacman::p_load_gh("mwesthues/sspredr")
 if (isTRUE(interactive())) {
   # Number of cores
   Sys.setenv("MOAB_PROCCOUNT" = "4")
-  # If TRUE, all agronomic data will be used, otherwise only agronomic data of
-  # hybrids whose parents have both, mrna and snp data, will be used.
-  Sys.setenv("ALL_PHENO" = "FALSE")
   Sys.setenv("TRAIT" = "GTM")
   # Number of iterations in BGLR()
   Sys.setenv("ITER" = "500")
@@ -37,8 +30,13 @@ if (isTRUE(interactive())) {
   Sys.setenv("VCOV" = "RadenII")
   Sys.setenv("PI" = "0.5")
   Sys.setenv("PRIOR_PI_COUNT" = "10")
-  # 'mrna' or 'snp' 
-  Sys.setenv("PREDICTOR" = "mrna")
+  # Main predictor. If 'Pred2' and 'Pred3' are empty, no imputation will take
+  # place.
+  Sys.setenv("PRED1" = "ped100")
+  # If 'Pred3' is empty, 'Pred2' will be imputed via information from 'Pred1'.
+  Sys.setenv("PRED2" = "snp77")
+  # If not empty, this predictor will be imputed.
+  Sys.setenv("PRED3" = "mrna42")
 }
 
 if (!interactive()) {
@@ -46,14 +44,15 @@ if (!interactive()) {
 } else job_id <- "interactive_00"
 
 use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
[…]

### 2016-11-22 reports/loocv_reduced.Rmd
Transfer prediction plots to report section

--- 2016-09-21 reports/loocv_reduced.Rmd
@@
-# Prediction For Reduced Data
+# LOOCV-Prediction
 ```{r message = FALSE, echo = FALSE}
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "modelr")
+pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
+               "stringr", "forcats")
 
-knit_hooks$set(inline = function(x) {
+knitr::knit_hooks$set(inline = function(x) {
   prettyNum(x, big.mark = ",", digits = 2)
 })
-knitr::opts_knit$set(root.dir = "~/gamazon/",
-                     inline = TRUE,
-                     verbose = FALSE)
+knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
 knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
@@
 ```{r Load-LOOCV-Data, cache = TRUE}
-# Load log files and data
-run_log <- fread("./data/derived/na_fraction_pred_log.txt")
+log_file <- fread("./data/derived/pred_log.txt")
 
-# LOOCV 
-loocv_nms <- run_log[CV == "LOOCV", Job_ID, ]
-loocv <- lapply(seq_along(loocv_nms), FUN = function(i) {
-  readRDS(paste0("./data/derived/predictions/", loocv_nms[i], ".RDS"))
-})
-loocv <- rbindlist(loocv)
+pred_df <- log_file %>%
+  .$Job_ID %>%
+  as.list() %>%
+  map(~readRDS(paste0("./data/derived/predictions/", ., ".RDS"))) %>%
+  bind_rows(.id = "id") %>%
+  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait, Iter),
+            by = "Job_ID") %>%
+  filter(Iter == 30000) %>%
+  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
+  mutate(Trait = fct_recode(Trait,
+    "DMY" = "GTM",
+    "DMC" = "GTS",
+    "FAT" = "FETT",
+    "PRO" = "RPR",
+    "SUG" = "XZ"
+  )) %>%
+  mutate(Trait = fct_relevel(Trait,
+                             "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"),
+         Predictor = as.factor(Predictor)) %>%
+  select(-Job_ID, -Iter)
+
+# Update the variable 'Trait' for subsequent joins with 'pred_df'.
+log_file <- log_file %>%
+  filter(Iter == 30000) %>%
+  mutate(Trait = fct_recode(Trait,
+    "DMY" = "GTM",
+    "DMC" = "GTS",
+    "FAT" = "FETT",
+    "PRO" = "RPR",
+    "SUG" = "XZ"
+  )) %>%
+  mutate(Trait = fct_relevel(Trait,
+                             "DMY", "DMC", "ADF", "FAT", "PRO",
+                             "STA", "SUG")) %>%
+  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
 ```
 
-```{r Rename-Traits, cache = TRUE}
-loocv_clean <- loocv %>%
-  mutate(Trait = fct_recode(Trait, 
-                            "FAT" = "FETT",
-                            "PRO" = "RPR",
-                            "DMY" = "GTM",
-                            "DMC" = "GTS",
-                            "SUG" = "XZ"),
-         Trait = fct_relevel(Trait, c("DMY", "DMC", "ADF", "FAT", "PRO", "STA",
-                                      "SUG")),
-         Dent_NA_Fraction = as.factor(as.character(Dent_NA_Fraction)),
-         Flint_NA_Fraction = as.factor(as.character(Flint_NA_Fraction)),
-         Residuals = y - yhat,
-         Predictor = fct_recode(Predictor,
-                                "mRNA" = "mrna",
-                                "SNP" = "snp")
-  )
+Compute the average standard deviation for each combination of predictors and 
+traits -> http://stats.stackexchange.com/a/26647
+```{r Coefficient-of-Variation, cache = TRUE}
+coefficient_of_variation <- function(x, y) {
+  sqrt(mean(x)) / mean(y)
+}
+cv_df <- pred_df %>%
+  split(list(.$Trait, .$Predictor)) %>%
+  map(~coefficient_of_variation(x = .$var_yhat, y = .$yhat)) %>%
+  stack() %>%
+  separate(col = ind, into = c("Trait", "Predictor"), sep = "[.]") %>%
+  rename(CV = values) %>%
+  tbl_dt() %>%
+  mutate(Predictor = as.factor(Predictor),
+         Trait = as.factor(Trait))
 ```
 
-## Predictive ability for non-imputed data
-Peak at predictive abilities based on non-imputed predictor information and rank
-the results by their predictive ability.
-```{r Non-Imputed-Pred-Ability, cache = TRUE}
-loocv_tbl_cap <- paste("Predictive abilities for the 'reduced' set of hybrids",
-                       "using two different predictors and seven agronomic",
-                       "traits. Results are ranked by predictive ability.")
-loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  mutate(Predictor = as.factor(Predictor)) %>%
-  droplevels() %>%
+Combine predictive abilities and coefficients of variation for all predictor
+and trait combinations for plotting.
+```{r Organize-Data, cache = TRUE}
+overview_df <- pred_df %>%
   group_by(Predictor, Trait) %>%
-  summarize(Pred_Ability = round(cor(y, yhat), digits = 2)) %>%
-  arrange(desc(Pred_Ability)) %>%
-  rename("Predictive Ability" = Pred_Ability) %>%
-  knitr::kable(caption = loocv_tbl_cap, booktabs = TRUE)
+  summarize(r = cor(y, yhat)) %>%
+  ungroup() %>%
+  as.data.frame() %>%
+  left_join(y = cv_df, by = c("Predictor", "Trait")) %>%
+  as_tibble() %>%
+  mutate(Predictor = as.factor(Predictor),
+         Predictor = forcats::fct_reorder(Predictor, CV),
+         Trait = fct_relevel(Trait,
+                             "DMY", "DMC", "ADF", "FAT", "PRO",
+                             "STA", "SUG"),
+         Group = ifelse(Predictor %in% c("snp42-none-none",
+                                         "ped42-none-none",
+                                         "mrna42-none-none"),
+                        yes = "Reduced", no = "Full")
+  ) %>%
+  mutate(Group = as.factor(Group))
+
+# Specify a unified predictor order.
+pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
+                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
+                "ped100-mrna42-none", "snp100-mrna42-none",
+                "ped100-snp77-mrna42")
 ```
 
-## Predicted vs. residuals
-```{r Complete-Data-Residuals, cache = TRUE, dependson = "Rename-Traits"}
-full_pred_resid <- loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  group_by(Predictor) %>%
-  do(plot = ggplot(., aes(x = yhat, y = Residuals)) +
-       geom_point(size = 0.5) +
-       geom_ref_line(h = 0) +
-       facet_wrap(~ Trait, nrow = 2, scales = "free") +
-       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
-       xlab("Predicted") +
-       ylab("Residuals") +
-       ggtitle(.$Predictor))
-invisible(lapply(full_pred_resid$plot, FUN = print))
+Distribution of predicted and observed values.
+Bear in mind that the distribution of the observed values will be different
+for different predictors because some predictors "xx42" and "xx77", where
+'xx' denotes a predictor, contain less hybrids than there are available 
+(i.e. "xx100").
+```{r Distribution-y-yhat, cache = TRUE}
+legend_labs <- expression(y, hat(y))
+distr_plot <- pred_df %>%
+  filter(Predictor != "snp77-none-none") %>%
+  gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
+  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
+  ggplot(aes(x = `Phenotypic Value`, color = Value)) +
+  geom_freqpoly() +
+  scale_color_manual(values = viridis(length(legend_labs), option = "C"),
+                     labels = legend_labs) +
+  facet_grid(Predictor ~ Trait, scales = "free") +
+  theme(legend.position = "top",
+        strip.text.y = element_text(size = 5)) +
+  ylab("Frequency")
 ```
 
-## Predicted vs. Observed
-```{r Complete-Data-y-yHat, cache = TRUE, dependson = "Rename-Traits"}
-full_y_yhat <- loocv_clean %>%
-  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
-  group_by(Predictor) %>%
-  do(plot = ggplot(., aes(x = y, y = yhat)) +
-       geom_point(size = 0.5) +
-       geom_abline(slope = 1, linetype = "dashed", color = "blue") +
-       facet_wrap(~ Trait, nrow = 2, scales = "free") +
-       theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
-       xlab("Observed") +
-       ylab("Predicted") +
-       ggtitle(.$Predictor))
-invisible(lapply(full_y_yhat$plot, FUN = print))
+Compare the coefficients of variation for all predictor-trait combinations.
+```{r Coefficient-of-Variation-Plot, cache = TRUE}
+cv_plot <- overview_df %>%
+  filter(Predictor != "snp77-none-none") %>%
+  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
+  ggplot(aes(x = CV, y = Predictor)) +
+  geom_segment(aes(yend = Predictor), xend = 0, color = "gray50") +
+  geom_point(aes(color = Predictor)) +
+  scale_color_viridis(discrete = TRUE) +
+  facet_wrap(~ Trait, scales = "free_x") +
+  theme_bw() +
+  theme(panel.grid.major.y = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        legend.position = c(0.7, 0.13)
+  ) +
+  guides(color = guide_legend(ncol = 2,
+                              reverse = FALSE)) +
+  xlab("Coefficient of Variation") +
+  ylab("Predictor Set") +
+  xlim(0, NA)
 ```
+
+Plot the predictive abilities for all predictor-trait combinations 
+separately for the reduced and the full data set, respectively.
+```{r Predictive-Ability-Plots, cache = TRUE}
+pred_plot <- overview_df %>%
+  as_tibble() %>%
+  filter(Predictor != "snp77-none-none") %>%
+  mutate(Predictor = fct_relevel(Predictor, pred_order),
+         Group = fct_relevel(Group, "Reduced", "Full")
+  ) %>%
+  ggplot(aes(x = Predictor, y = r)) +
+  geom_bar(stat = "identity", aes(fill = Predictor), color = "black") +
+  geom_text(aes(label = round(r, digits = 3)), vjust = -0.2) +
+  scale_fill_viridis(discrete = TRUE, option = "inferno") +
+  facet_grid(Trait ~ Group, space = "free", scales = "free_x") +
+  theme_bw() +
+  theme(
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.x = element_blank(),
+    legend.position = "bottom",
+    legend.key = element_rect(color = "white")
+  ) +
+  guides(fill = guide_legend(nrow = 3,
+                             byrow = TRUE,
+                             keywidth = 0.25,
+                             keyheight = 0.25,
+                             default.unit = "inch",
+                             title.position = "top")
+  ) +
+  ylim(c(0, 1))
+```

### 2016-11-22 analysis/plot_predictions.R

@@ -1,148 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
-               "stringr", "corrplot", "forcats")
-
-log_file <- fread("./data/derived/pred_log.txt")
-
-pred_df <- log_file %>%
-  .$Job_ID %>%
-  as.list() %>%
-  map(~readRDS(paste0("./data/derived/predictions/", ., ".RDS"))) %>%
-  bind_rows(.id = "id") %>%
-  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait),
-            by = "Job_ID") %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
-  mutate(Trait = fct_recode(Trait,
-    "DMY" = "GTM",
-    "DMC" = "GTS",
-    "FAT" = "FETT",
-    "PRO" = "RPR",
-    "SUG" = "XZ"
-  )) %>%
-  mutate(Trait = fct_relevel(Trait,
-                             "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"),
-         Predictor = as.factor(Predictor)) %>%
-  select(-Job_ID)
-
-
-# Update the variable 'Trait' for subsequent joins with 'pred_df'.
-log_file <- log_file %>%
-  mutate(Trait = fct_recode(Trait,
-    "DMY" = "GTM",
-    "DMC" = "GTS",
-    "FAT" = "FETT",
-    "PRO" = "RPR",
-    "SUG" = "XZ"
-  )) %>%
-  mutate(Trait = fct_relevel(Trait,
-                             "DMY", "DMC", "ADF", "FAT", "PRO",
-                             "STA", "SUG")) %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
-
-
-# Compute the average standard deviation for each combination of predictors and 
-# traits -> http://stats.stackexchange.com/a/26647
-coefficient_of_variation <- function(x, y) {
-  sqrt(mean(x)) / mean(y)
-}
-cv_df <- pred_df %>%
-  split(list(.$Trait, .$Predictor)) %>%
-  map(~coefficient_of_variation(x = .$var_yhat, y = .$yhat)) %>%
-  stack() %>%
[…]

### 2016-11-22 tmp/dummy.R

@@ -1,239 +1,0 @@
-# Single-step prediction of hybrid performance based on Fernando et al.: A
-# class of Bayesian methods to combine large numbers of genotyped and
-# non-genotyped animals for whole-genome analyses. Genetics Selection Evolution
-# 2014 46:50
-#
-# The approach will be adjusted in a sense that we do have complete genomic
-# instead of complete pedigree data (i.e. A -> G) and that we have incomplete
-# transcriptomic instead of incomplete genomic data (i.e. G -> T).
-#
-#
-#
-if (!require("pacman")) install.packages("pacman")
-if (!require("devtools")) install.packages("devtools")
-devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("data.table", "dplyr", "tidyr", "methods", "BGLR", "Matrix",
-               "parallel", "matrixStats", "dtplyr", "tibble", "testthat")
-pacman::p_load_gh("mwesthues/sspredr")
-
-
-## --------------------------------------------------------------------------
-## COLLECT AND CHECK ARGUMENTS 
-if (isTRUE(interactive())) {
-  # Number of cores
-  Sys.setenv("MOAB_PROCCOUNT" = "4")
-  Sys.setenv("TRAIT" = "")
-  # Number of iterations in BGLR()
-  Sys.setenv("ITER" = "10000")
-  # Prediction model in BGLR()
-  Sys.setenv("MODEL" = "BRR")
-  # Algorithm to generate variance-covariance matrices.
-  Sys.setenv("VCOV" = "RadenII")
-  Sys.setenv("PI" = "0.5")
-  Sys.setenv("PRIOR_PI_COUNT" = "10")
-  # 'mrna' or 'snp' 
-  Sys.setenv("PREDICTOR" = "mrna")
-  # Which fraction of Dent genotypes shall be set to 'NA' for cross-validation?
-  Sys.setenv("DENT_NA_FRACTION" = "0.05")
-  # Which fraction of Flint genotypes shall be set to 'NA' for cross-validation?
-  Sys.setenv("FLINT_NA_FRACTION" = "0.05")
-}
-
-if (!interactive()) {
-  job_id <- Sys.getenv("MOAB_JOBID")
-} else job_id <- "interactive_00"
-
-use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
-init_traits <- as.character(Sys.getenv("TRAIT"))
[…]

### 2016-11-22 reports/agronomic_data.Rmd
Add legends to prediction legends

@@
 # Agronomic Data
-```{r echo = FALSE, message = FALSE}
+```{r setup-agro, echo = FALSE, message = FALSE}
 if (!require("pacman")) install.packages("pacman")

### 2016-11-22 reports/data_overview.Rmd
Add legends to prediction legends

@@
 # Data Overview
-```{r message = FALSE, echo = FALSE}
+```{r setup-overview, message = FALSE, echo = FALSE}
 if (!require("pacman")) install.packages("pacman")

### 2016-11-22 reports/impute_single_mrna.Rmd
Add legends to prediction legends

@@
 # Genotype-Wise mRNA-Imputation
-```{r echo = FALSE}
+```{r setup-impute, echo = FALSE}
 pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis",
@@
 knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
+knitr::opts_chunk$set(fig.width = 14, fig.height = 10, 
                       warning = FALSE, message = FALSE, echo = FALSE)
@@
 ```{r Plot-Correlation-Imputed-vs-NonImputed, cache = TRUE, fig.width = 12, fig.height = 12, fig.cap = "Pearson correlation coefficients between original and imputed mRNA genotypes for two heterotic groups and two sources of information (i.e.
-genomic and pedigree)."}
+genomic and pedigree), depending on the number of genotypes without mRNA records.
+Genotypes without mRNA records were declared as such at random."}
 par_lst <- readRDS("./data/derived/impute_single_mrna.RDS")
@@
   geom_boxplot() +
-  facet_wrap(Data_Type ~ Group) +
+  facet_wrap(Data_Type ~ Group, scales = "free") +
+  theme_bw() +
   theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +

### 2016-11-22 reports/index.Rmd
remove code boxes from report
Add legends to prediction legends

--- 2016-09-21 reports/index.Rmd
@@
 
-Chapter \@ref(prediction-for-reduced-data) deals with the results of hybrid 
-prediction for the reduced set of data.
+Chapter \@ref(transcriptomic-data) explores the distribution and correlations
+of/among transcriptomic predictors for "raw" and transformed mRNA BLUEs.
 
-Chapter \@ref(prediction-with-imputed-mrnas) examines the influence of certain levels of missingness 
-in the transcriptomic data.
+Chapter \@ref(genotype-wise-mrna-imputation) examines the correlation between
+the original and imputed transcriptomic BLUEs.
+
+Chapter \@ref(loocv-prediction) deals with the results of hybrid prediction for 
+the reduced set of data and a comparison of predictive abilities with the full 
+data set.
 

### 2016-11-22 reports/loocv_reduced.Rmd
Add legends to prediction legends

@@
 # LOOCV-Prediction
-```{r message = FALSE, echo = FALSE}
+```{r setup-loocv, message = FALSE, echo = FALSE}
 if (!require("pacman")) install.packages("pacman")
@@
 (i.e. "xx100").
-```{r Distribution-y-yhat, cache = TRUE}
+```{r Distribution-y-yhat, cache = TRUE, fig.height = 12, fig.cap = "Frequency polygon of observed and predicted values for seven agronomic traits (columns) and nine sets of predictors (rows)."}
 legend_labs <- expression(y, hat(y))
-distr_plot <- pred_df %>%
+pred_df %>%
   filter(Predictor != "snp77-none-none") %>%
@@
   geom_freqpoly() +
-  scale_color_manual(values = viridis(length(legend_labs), option = "C"),
-                     labels = legend_labs) +
+  scale_color_manual(values = c("#1f78b4", "#e31a1c"), labels = legend_labs) +
   facet_grid(Predictor ~ Trait, scales = "free") +
+  theme_bw() +
   theme(legend.position = "top",
-        strip.text.y = element_text(size = 5)) +
+        strip.text.y = element_text(size = 7)) +
+  guides(color = guide_legend(override.aes = list(size = 2))) +
   ylab("Frequency")
@@
 Compare the coefficients of variation for all predictor-trait combinations.
-```{r Coefficient-of-Variation-Plot, cache = TRUE}
-cv_plot <- overview_df %>%
+```{r Coefficient-of-Variation-Plot, cache = TRUE, fig.cap = "Coefficient of variation for seven agronomic traits and nine sets of predictors."}
+overview_df %>%
   filter(Predictor != "snp77-none-none") %>%
@@
   geom_segment(aes(yend = Predictor), xend = 0, color = "gray50") +
-  geom_point(aes(color = Predictor)) +
-  scale_color_viridis(discrete = TRUE) +
+  geom_point(aes(color = Predictor), size = 3) +
+  scale_color_viridis(discrete = TRUE, option = "inferno") +
   facet_wrap(~ Trait, scales = "free_x") +
@@
 separately for the reduced and the full data set, respectively.
-```{r Predictive-Ability-Plots, cache = TRUE}
-pred_plot <- overview_df %>%
+```{r Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for seven agronomic traits (rows) and nine predictor sets (columns).
+The predictor sets are split into the categories 'Reduced' (i.e.
+685 hybrids) and 'Full' (i.e.
+1,521 hybrids)."}
+overview_df %>%
   as_tibble() %>%

### 2016-11-22 reports/na_fraction_pheno_predability.Rmd
Add legends to prediction legends

@@
 # Prediction with Imputed mRNAs
-```{r message = FALSE}
+```{r setup-fraction, message = FALSE}
 if (!require("pacman")) install.packages("pacman")

### 2016-11-22 reports/transcriptomic_data.Rmd
Add legends to prediction legends

@@
 available.
-```{r echo = FALSE}
+```{r setup-transcriptomic, echo = FALSE}
 pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")

### 2016-11-22 reports/agronomic_data.Rmd
Move predictor_subsetting.R content into reports file

@@
 ## Correlation
-```{r Agro-Correlation, echo = FALSE, fig.cap = "Pairwise correlations among seven agronomic traits for the full set of hybrids and the reduced set of hybrids, respectively."}
-par(mar = c(0, 1, 0, 0),
-    mfrow = c(1, 2))
-all_pheno %>%
-  select(-Reduced_Data) %>%
+```{r Agro-Correlation, echo = FALSE, fig.cap = "Pairwise correlations among seven agronomic traits for the full set of hybrids (upper diagonal) and the reduced set of hybrids (lower diagonal), respectively."}
+cor_coeffs <- all_pheno %>%
+  rbind(., reduced_pheno) %>%
   spread(key = Trait, value = Observed) %>%
-  as.data.frame() %>%
-  remove_rownames() %>%
-  column_to_rownames(var = "G") %>%
-  cor() %>%
+  split(.$Reduced_Data) %>%
+  map(~select(., -Reduced_Data)) %>%
+  map(as.data.frame) %>%
+  map(remove_rownames) %>%
+  map(column_to_rownames, var = "G") %>%
+  map(cor) %>%
+  map_at(.at = "no", .f = ~.[upper.tri(., diag = FALSE)]) %>%
+  map_at(.at = "yes", .f = ~.[lower.tri(., diag = FALSE)])
+traits <- levels(all_pheno$Trait)
+cor_mat <- diag(nrow = 7, ncol = 7)
+cor_mat[upper.tri(cor_mat, diag = FALSE)] <- cor_coeffs$no
+cor_mat[lower.tri(cor_mat, diag = FALSE)] <- cor_coeffs$yes
+dimnames(cor_mat) <- list(traits, traits)
+cor_mat %>%
   corrplot::corrplot(col = viridis(256),
@@
                      addCoef.col = "black",
-                     title = "Full",
-                     mar = c(0, 1, 1, 0))
-reduced_pheno %>%
-  filter(Reduced_Data == "yes") %>%
-  select(-Reduced_Data) %>%
-  spread(key = Trait, value = Observed) %>%
-  as.data.frame() %>%
-  remove_rownames() %>%
-  column_to_rownames(var = "G") %>%
-  cor() %>%
-  corrplot::corrplot(col = viridis(256),
-                     method = "color",
-                     diag = FALSE,
-                     cl.pos = "b",
-                     tl.pos = "d",
-                     tl.srt = 60,
-                     order = "hclust",
-                     addCoef.col = "black",
-                     title = "Reduced",
                      mar = c(0, 1, 1, 0))

### 2016-11-22 reports/loocv_reduced.Rmd
Move predictor_subsetting.R content into reports file

@@
 
-Compute the average standard deviation for each combination of predictors and 
-traits -> http://stats.stackexchange.com/a/26647
+## Distribution of predicted and observed values
+```{r Distribution-y-yhat, cache = TRUE, fig.height = 12, fig.cap = "Frequency polygon of observed and predicted values for seven agronomic traits (columns) and nine sets of predictors (rows)."}
+# Specify a unified predictor order.
+pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
+                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
+                "ped100-mrna42-none", "snp100-mrna42-none",
+                "ped100-snp77-mrna42")
+legend_labs <- expression(y, hat(y))
+pred_df %>%
+  filter(Predictor != "snp77-none-none") %>%
+  gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
+  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
+  ggplot(aes(x = `Phenotypic Value`, color = Value)) +
+  geom_freqpoly() +
+  scale_color_manual(values = c("#1f78b4", "#e31a1c"), labels = legend_labs) +
+  facet_grid(Predictor ~ Trait, scales = "free") +
+  theme_bw() +
+  theme(legend.position = "top",
+        strip.text.y = element_text(size = 7)) +
+  guides(color = guide_legend(override.aes = list(size = 2))) +
+  ylab("Frequency")
+```
+
+## Coefficient of variation for predicted values
+We are interested in the variation in the predicted values relative to the 
+arithmetic mean of the predicted values, separately for each agronomic trait and
+each set of predictors:
+
+$$
+\begin{equation}
+  \label{eq:CV}
+CV = \sqrt{\frac{\sum_{i = 1}^{N} \sigma^{2}_{\hat{y}(i)} / N}{\sum_{i=1}^{N}\hat{y}_{i} / N}},
+\end{equation}
+$$
+
+where $N$ denotes the number of hybrids.
+
 ```{r Coefficient-of-Variation, cache = TRUE}
@@
          Trait = as.factor(Trait))
-```
 
-Combine predictive abilities and coefficients of variation for all predictor
-and trait combinations for plotting.
-```{r Organize-Data, cache = TRUE}
 overview_df <- pred_df %>%
@@
   mutate(Group = as.factor(Group))
-
-# Specify a unified predictor order.
-pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
-                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
-                "ped100-mrna42-none", "snp100-mrna42-none",
-                "ped100-snp77-mrna42")
 ```
 
-Distribution of predicted and observed values.
-Bear in mind that the distribution of the observed values will be different
-for different predictors because some predictors "xx42" and "xx77", where
-'xx' denotes a predictor, contain less hybrids than there are available 
-(i.e. "xx100").
-```{r Distribution-y-yhat, cache = TRUE, fig.height = 12, fig.cap = "Frequency polygon of observed and predicted values for seven agronomic traits (columns) and nine sets of predictors (rows)."}
-legend_labs <- expression(y, hat(y))
-pred_df %>%
-  filter(Predictor != "snp77-none-none") %>%
-  gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
-  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
-  ggplot(aes(x = `Phenotypic Value`, color = Value)) +
-  geom_freqpoly() +
-  scale_color_manual(values = c("#1f78b4", "#e31a1c"), labels = legend_labs) +
-  facet_grid(Predictor ~ Trait, scales = "free") +
-  theme_bw() +
-  theme(legend.position = "top",
-        strip.text.y = element_text(size = 7)) +
-  guides(color = guide_legend(override.aes = list(size = 2))) +
-  ylab("Frequency")
-```
-
-Compare the coefficients of variation for all predictor-trait combinations.
+## Predictive Abilities
 ```{r Coefficient-of-Variation-Plot, cache = TRUE, fig.cap = "Coefficient of variation for seven agronomic traits and nine sets of predictors."}

### 2016-11-22 reports/predictor_subsetting.Rmd
Move predictor_subsetting.R content into reports file

# Predictor Subsetting
```{r setup-predictor-subsets, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
               "stringr", "forcats")

devtools::install_github("mwesthues/sspredr")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

Goal: Generate nine subsets of predictors for later comparisons between their
predictive abilities:

a.    100% SNP
b.    100% pedigree
c.    42% mRNA
d.    42% pedigree
e.    42% SNP
f.    100% pedigree & 77% SNP
g.    100% pedigree & 42% mRNA
h.    100% SNP & 42% mRNA
i.    100% pedigree & 77% SNP & 42% mRNA

Make comparisons between the following scenarios:

1.    a, b, f, g, h, i
2.    c, d, e

Throughout this manuscript, digits behind the variable name of each predictor
(ped: pedigree, snp: genomic, mrna: transcriptomic) indicate the size of 
the subset relative to the full set of genotypes (n = 245).
For example, we have transcriptomic data for 103 genotypes, which are roughly 
42% of 245 genotypes, hence the name 'mrna42' and so forth.
For the set "snp77", all genotypes &mdash; for which transcriptomic records are 
available &mdash; were used and the remaining genotypes were sampled at random
with equally large proportions being sampled from the Dent and the Flint 
heterotic group, respectively.

```{r Prepare-Predictor-Subsetting, cache = TRUE, results = "hide"}
# -- Genotype information.
# Information on the set of genotypes.
genos <- readRDS("./data/processed/common_genotypes.RDS")
genos %>%
  filter(Pool %in% c("Dent", "Flint")) %>%
  group_by(Pool, Data_Type) %>%
  count()

# Get the names of the Dent Hybrid-parents.
geno <- genos %>%
  filter(Pool == "Hybrid") %>%
  .$G %>%
  map(~stringr::str_split(., pattern = "_")) %>%
  at_depth(.depth = 2, .f = 2) %>%
  flatten() %>%
  as_vector()

# -- Load predictor data.
# pedigree
ped100 <- readRDS("./data/processed/ped-datafull-GTP.RDS")
ped_nms <- genos %>%
  filter(Data_Type == "ped") %>%
  .$G
ped100 <- as.matrix(ped100) * 2
ped100 <- ped100 %>%
  .[rownames(.) %in% ped_nms, colnames(.) %in% ped_nms]
ped100[is.na(ped100)] <- 0

# snp
snp100 <- readRDS("./data/processed/snp_mat.RDS")
snp_nms <- genos %>%
  filter(Data_Type == "snp") %>%
  .$G
snp100 <- snp100 %>%
  .[rownames(.) %in% snp_nms, ]

# mrna
mrna42 <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna_nms <- genos %>%
  filter(Data_Type == "mrna") %>%
  .$G
mrna42 <- mrna42 %>%
  as.matrix() %>%
  t() %>%
  .[rownames(.) %in% mrna_nms, ]
```

```{r Resample-Predictor-Data, cache = TRUE, results = "hide"}
## Resample predictor data 
# Remove some genomic records so that we can impute them via pedigree 
# information.
set.seed(314)
snp35 <- snp100 %>%
  as_tibble() %>%
  mutate(G = rownames(snp100)) %>%
  filter(!G %in% rownames(mrna42)) %>%
  left_join(genos %>% filter(Data_Type == "snp") %>% select(-Data_Type),
            by = "G") %>%
  group_by(Pool) %>%
  sample_frac(size = 0.6) %>%
  ungroup() %>%
  select(-Pool) %>%
  as.data.frame() %>%
  column_to_rownames(var = "G") %>%
  as.matrix()
snp77 <- snp100 %>%
  .[rownames(.) %in% rownames(mrna42), ] %>%
  rbind(snp35)
stopifnot(all(rownames(mrna42) %in% rownames(snp77)))

ped42 <- ped100 %>%
  .[match(rownames(mrna42), rownames(.)), 
    match(rownames(mrna42), colnames(.))]
snp42 <- snp100 %>%
  .[match(rownames(mrna42), rownames(.)), ]

# Store the new objects
storage_dir <- "./data/derived/predictor_subsets/"
ls() %>%
  .[. != "snp35"] %>%
  .[stringr::str_detect(., pattern = "[:digit:]")] %>%
  as.list() %>%
  map_if(.p = ~get(.) %>% is.matrix,
         .f = ~saveRDS(get(.), file = paste0(storage_dir, ., ".RDS"),
                       compress = TRUE))
```

### 2016-11-22 reports/loocv_reduced.Rmd
Remove conclusions from reports

@@
 
-## Predictive Abilities
 ```{r Coefficient-of-Variation-Plot, cache = TRUE, fig.cap = "Coefficient of variation for seven agronomic traits and nine sets of predictors."}
@@
 
-Plot the predictive abilities for all predictor-trait combinations 
-separately for the reduced and the full data set, respectively.
+## Predictive Abilities
 ```{r Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for seven agronomic traits (rows) and nine predictor sets (columns).

### 2016-12-05 reports/index.Rmd
Remove conclusions from reports
Investigate kernel coefficients for single step predictions

@@
 
+Chapter \@ref(predictor-subsetting) introduces the different predictor sets that
+were used for predictions.
+
+Chapter \@ref(kernel-inspection) gives the distribution of coefficients in the 
+different kernel matrices used in hybrid prediction.
+
 Chapter \@ref(loocv-prediction) deals with the results of hybrid prediction for 
@@
 data set.
-
-Chapter \@ref(conclusions-and-outlook) provides a summary of the analyses and 
-provides an outlook for future analyses.

### 2016-12-05 reports/single_step_kernels.Rmd
Investigate kernel coefficients for single step predictions

# Kernel Inspection
```{r setup-kernel-inspection, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
               "purrr", "testthat", "methods", "tibble", "ggplot2", "stringr",
               "stringi", "lubridate", "readr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

Goal: Inspection of input matrices for single-step hybrid prediction using 
BGLR.

```{r Prepare-Kernel-Inspection, cache = TRUE}
## -- DATA SELECTION -----------------------------------------------------
# Specify the predictor sets.
preds <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
           "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
           "ped100-mrna42-none", "snp100-mrna42-none", "ped100-snp77-mrna42")

# Load the predictor data.
pred_lst <- preds %>%
  as.list() %>%
  map(~str_split(string = ., pattern = "-")) %>%
  at_depth(.depth = 2, ~discard(., . == "none")) %>%
  at_depth(.depth = 2, as.list) %>%
  at_depth(.depth = 2, 
           .f = ~paste0("./data/derived/predictor_subsets/", ., ".RDS")) %>%
  at_depth(.depth = 3, .f = readRDS) %>%
  flatten()
names(pred_lst) <- preds

# For each predictor set and for each heterotic group, select the corresponding
# hybrids for which phenotpyic data are available.
# Then, store the names of all Dent and Flint hybrid parents in a list in order 
# to correctly augment the predictor matrices with their help.
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos %>%
  filter(Pool == "Hybrid") %>%
  mutate(G = stringr::str_replace(G, pattern = "DF_", replacement = "")) %>%
  separate(G, into = c("Dent", "Flint"), sep = "_")
geno_nms <- pred_lst %>%
  at_depth(.depth = 2, .f = rownames) %>%
  map(.f = ~reduce(., union))
hybrid_parent_nms <- lapply(seq_along(geno_nms), FUN = function(i) {
  ind_nms <- geno_nms[[i]] 
  hybrid %>%
    filter(Dent %in% ind_nms, Flint %in% ind_nms) %>%
    unite(col = Hybrid, Dent, Flint, sep = "_", remove = FALSE) %>%
    mutate(Hybrid = paste0("DF_", Hybrid)) %>%
    gather(key = "Group", value = "G", Dent, Flint) %>%
    split(.$Group) %>%
    map("G")
})
names(hybrid_parent_nms) <- names(pred_lst)
hybrid_parent_nms <- transpose(hybrid_parent_nms)

## -- PREPARE DATA FOR KERNEL BUILDING ------------------------------------
# Split the predictor matrices into Dent and Flint.
# Then, transform the data frames back to matrices in order to be able build 
# kernels from them.
merge_pred <- function(x, genos) {
  y <- genos %>%
    select(G, Pool) %>%
    distinct(G, .keep_all = TRUE)
  x %>%
    as.data.frame() %>%
    rownames_to_column(., var = "G") %>%
    as_tibble() %>%
    left_join(., y = y, by = "G") %>%
    split(., .$Pool)
}
pred_df_to_matrix <- function(x) {
  x %>%
    select(-Pool) %>%
    as.data.frame() %>%
    column_to_rownames(var = "G") %>%
    as.matrix()
}
feat_lst <- pred_lst %>%
  at_depth(.depth = 2,
           .f = merge_pred, genos = genos) %>%
  map(transpose) %>%
  at_depth(.depth = 3, .f = pred_df_to_matrix)
names(feat_lst) <- str_replace_all(preds, pattern = "-", replacement = "_")
feat_lst <- transpose(feat_lst)

# As a requirement of the kernel-generating functions, add the names of the 
# hybrid parents to each predictor-set/group combination.
cat_lst <- lapply(seq_along(feat_lst), FUN = function(i) {
  x_lvl1 <- feat_lst[[i]] 
  y_lvl1 <- hybrid_parent_nms[[i]]
  geno_lst <- lapply(seq_along(x_lvl1), FUN = function(j) {
    x_lvl2 <- x_lvl1[[j]]
    y_lvl2 <- y_lvl1[[j]]
    x_lvl2[["geno"]] <- y_lvl2
    x_lvl2
  })
  names(geno_lst) <- names(x_lvl1)
  geno_lst
})

## -- BUILD KERNELS --------------------------------------------------------
# Apply kernel building functions to different predictor sets.
# The choice of function depends on the number of predictors as well as on the
# type of predictor (are pedigree data involved or not?).
ped100_snp77_mrna42 <- cat_lst %>%
  transpose() %>%
  .[["ped100_snp77_mrna42"]] %>%
  map(~impute2(ped = .[[1]],
               snp = .[[2]],
               mrna = .[[3]],
               geno = .[["geno"]],
               as_kernel = TRUE,
               bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

snp100_mrna42_none <- cat_lst %>%
  transpose() %>%
  .[["snp100_mrna42_none"]] %>%
  map(~impute_eta(
     x = .[[1]],
     y = .[[2]],
     geno = .[["geno"]],
     as_kernel = TRUE,
     is_pedigree = FALSE,
     bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

ped100_x_none <- cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("ped100_mrna42_none", "ped100_snp77_none")] %>%
  at_depth(.depth = 2, .f = ~impute_eta(
    x = .[[1]],
    y = .[[2]],
    geno = .[["geno"]], 
    as_kernel = TRUE,
    is_pedigree = TRUE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 2) %>%
  at_depth(.depth = 2, .f = "X")

pedx_none_none <- cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("ped100_none_none", "ped42_none_none")] %>%
  at_depth(.depth = 2, .f = ~complete_eta(
    x = .[[1]],
    geno = .[["geno"]],
    as_kernel = FALSE,
    is_pedigree = TRUE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 1) %>%
  at_depth(.depth = 2, .f = "X")

single_pred <- cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("snp42_none_none", "mrna42_none_none", 
                    "snp100_none_none")] %>%
  at_depth(.depth = 2, .f = ~complete_eta(
    x = .[[1]],
    geno = .[["geno"]],
    as_kernel = TRUE,
    is_pedigree = FALSE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 1) %>%
  at_depth(.depth = 2, .f = "X")

# Combine the results in a single list to reduce code replications (the
# following steps are the same for all predictor sets).
W_lst <- c(list(ped100_snp77_mrna42 = ped100_snp77_mrna42),
           list(snp100_mrna42_none = snp100_mrna42_none),
           pedx_none_none,
           ped100_x_none,
           single_pred)
```

```{r Plot-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the kernel matrices used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

# For each predictor set and each heterotic group, plot the distribution of the
# coefficients from the kernels.
W_df %>%
  ggplot(aes(x = Values, fill = Group)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 80) +
  facet_wrap(~ Predictor, scales = "free") +
  theme_bw() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#669933", "#FFCC66")) +
  xlim(c(-1.0, 1.0))
```

```{r Extreme-Kernel-Coefficients, cache = TRUE}
extreme_cap <- paste("Most extreme values for each predictor set and each", 
                     "heterotic group.")
W_df %>%
  group_by(Predictor) %>%
  summarize(Min = round(min(Values), digits = 3),
            Max = round(max(Values), digits = 3)) %>%
  ungroup() %>%
  knitr::kable(caption = extreme_cap, booktabs = TRUE)
```

```{r Kernel-Closeup, cache = TRUE, fig.cap = "Distribution of kernel coefficients used in single step hybrid prediction for a scenario with complete genomic and incomplete transcriptomic information ('snp100-mrna42-none') as well as a scenario with complete pedigree, incomplete genomic and incomplete transcriptomic information ('ped100-snp77-mrna42'). The focus was placed on coefficients in the range from -0.1 to +0.1."}
W_df %>%
  filter(Predictor %in% c("snp100_mrna42_none", "ped100_snp77_mrna42")) %>%
  ggplot(aes(x = Values, fill = Group)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 80) +
  facet_wrap(~ Predictor, scales = "free") +
  theme_bw() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#669933", "#FFCC66")) +
  xlim(c(-0.1, 0.1))
```

### 2016-12-05 analysis/helper_prediction.R

@@ -1,226 +1,0 @@
-# Single-step prediction of hybrid performance based on Fernando et al.: A
-# class of Bayesian methods to combine large numbers of genotyped and
-# non-genotyped animals for whole-genome analyses. Genetics Selection Evolution
-# 2014 46:50
-#
-# The approach will be adjusted in a sense that we do have complete genomic
-# instead of complete pedigree data (i.e. A -> G) and that we have incomplete
-# transcriptomic instead of incomplete genomic data (i.e. G -> T).
-#
-#
-#
-if (!require("pacman")) install.packages("pacman")
-if (!require("devtools")) install.packages("devtools")
-devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
-               "purrr", "testthat", "methods", "tibble", "ggplot2", "stringr",
-               "stringi", "lubridate", "readr")
-pacman::p_load_gh("mwesthues/sspredr")
-
-
-## --------------------------------------------------------------------------
-## COLLECT AND CHECK ARGUMENTS 
-if (isTRUE(interactive())) {
-  # Number of cores
-  Sys.setenv("MOAB_PROCCOUNT" = "3")
-  Sys.setenv("TRAIT" = "GTM")
-  # Number of iterations in BGLR()
-  Sys.setenv("ITER" = "30000")
-  # Prediction model in BGLR()
-  Sys.setenv("MODEL" = "BRR")
-  # Algorithm to generate variance-covariance matrices.
-  Sys.setenv("VCOV" = "RadenII")
-  Sys.setenv("PI" = "0.5")
-  Sys.setenv("PRIOR_PI_COUNT" = "10")
-  # Main predictor. If 'Pred2' and 'Pred3' are empty, no imputation will take
-  # place.
-  Sys.setenv("PRED1" = "ped100")
-  # If 'Pred3' is empty, 'Pred2' will be imputed via information from 'Pred1'.
-  Sys.setenv("PRED2" = "snp77")
-  # If not empty, this predictor will be imputed.
-  Sys.setenv("PRED3" = "")
-  # Number of genotypes to predict (only for testing!)
-  Sys.setenv("RUNS" = "")
-}
-
-if (!interactive()) {
-  job_id <- Sys.getenv("MOAB_JOBID")
[…]

### 2016-12-05 new_factorization.R

@@ -18,7 +18,11 @@ pred_df %>%
   unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
   select(-Job_ID, -Iter) %>%
   group_by(Predictor, Trait) %>%
-  summarize(PredAbility = cor(y, yhat))
+  summarize(PredAbility = cor(y, yhat)) %>%
+  as.data.frame()
 
 
 
+readRDS("./data/derived/predictions/10094280.RDS") %>%
+  plot(.$y, .$yhat)
+  .[, .(cor(y, yhat)), ]

### 2016-12-06 reports/single_step_kernels.Rmd
Try predictions with scaled input matrices

@@
 })
+saveRDS(cat_lst,
+        file = "./data/derived/pred_sub_list.RDS")
 

### 2016-12-06 analysis/scaled_mrna_predictions.R

@@ -1,0 +1,196 @@
+if (!require("pacman")) install.packages("pacman")
+if (!require("devtools")) install.packages("devtools")
+devtools::install_github("mwesthues/sspredr", update = TRUE)
+pacman::p_load("BGLR", "data.table", "tidyverse", "dtplyr", "parallel", 
+               "caret", "e1071", "viridis")
+pacman::p_load_gh("mwesthues/sspredr")
+
+
+# Common genotypes
+genos <- readRDS("./data/processed/common_genotypes.RDS")
+mrna_inbreds <- genos %>%
+  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
+  split(.$Pool) %>%
+  map("G")
+
+# mRNA
+mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
+mrna <- t(mrna)
+mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
+mrna_df <- mrna %>%
+  as_data_frame() %>%
+  mutate(G = rownames(mrna)) %>%
+  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
+         Group = as.factor(Group))
+
+## Data pre-processing
+# Pre-process the data in the following order:
+# 1.    Near-zero-variance filtering.
+# 2.    Box-Cox tranformation
+# 3.    Centering
+# 4.    Scaling
+mrna_lst <- mrna_df %>% 
+  split(.$Group) %>%
+  map(as.data.frame) %>%
+  map(~column_to_rownames(., var = "G")) %>%
+  map(.f = function(x) {
+    x$Group <- NULL
+    x
+  }) %>%
+  map(~as.matrix(.))
+
+# Data transformation
+trans_lst <- mrna_lst %>%
+  map(preProcess,
+      method = c("BoxCox", "center", "scale", "nzv")) %>%
+  map2(.y = mrna_lst, .f = predict)
+
+
+pred_lst <- readRDS("./data/derived/pred_sub_list.RDS")
+snp77 <- pred_lst %>% 
+  transpose() %>%
+  .[names(.) == "ped100_snp77_none"] %>%
+  at_depth(.depth = 2, .f = ~.[c(2, 3)]) %>%
+  .[[1]]
+bare_snp77 <- snp77 %>%
+  map(.f = 1)
+snp77_cor <- bare_snp77 %>%
+  map(cor)
+snp77_zv <- snp77_cor %>%
[…]

### 2016-12-06 reports/single_step_kernels.Rmd
Working version of scaled prediction with impute2()

@@
 devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
-               "purrr", "testthat", "methods", "tibble", "ggplot2", "stringr",
-               "stringi", "lubridate", "readr")
+pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071")
 pacman::p_load_gh("mwesthues/sspredr")

### 2016-12-06 analysis/scaled_mrna_predictions.R

@@ -46,8 +46,8 @@ trans_lst <- mrna_lst %>%
   map2(.y = mrna_lst, .f = predict)
 
 
-pred_lst <- readRDS("./data/derived/pred_sub_list.RDS")
-snp77 <- pred_lst %>% 
+pred_sub_lst <- readRDS("./data/derived/pred_sub_list.RDS")
+snp77 <- pred_sub_lst %>% 
   transpose() %>%
   .[names(.) == "ped100_snp77_none"] %>%
   at_depth(.depth = 2, .f = ~.[c(2, 3)]) %>%
@@ -77,12 +77,11 @@ snp77_mrna42 <- lapply(seq_along(thinned_snp77), FUN = function(i) {
 })
 names(snp77_mrna42) <- c("Dent", "Flint")
 
-ped100 <- pred_lst %>% 
+ped100 <- pred_sub_lst %>% 
   transpose() %>%
   .[names(.) == "ped100_snp77_none"] %>%
   at_depth(.depth = 2, .f = 1) %>%
-  .[[1]] %>%
-  map(~.[, match(rownames(.), rownames(.))])
+  .[[1]]
 
 ped100_snp77_mrna42 <- snp77_mrna42
 ped100_snp77_mrna42[["Dent"]][["ped"]] <- ped100[[1]]
@@ -98,19 +97,19 @@ ped100_snp77_mrna42_eta <- ped100_snp77_mrna42 %>%
     )
   )
 
-#snp77_mrna42_eta %>%
-#  map(2) %>%
-#  map("X") %>%
-#  map(~.[match(colnames(.), rownames(.)), ]) %>%
-#  map(~.[lower.tri(., diag = FALSE)]) %>%
-#  stack() %>%
-#  rename(Values = values,
-#         Group = ind) %>%
-#  ggplot(aes(x = Values, fill = Group)) +
-#  geom_histogram(alpha = 0.5, position = "identity") +
-#  theme_bw() +
-#  theme(legend.position = "top") +
-#  scale_fill_manual(values = c("#669933", "#FFCC66"))
+ped100_snp77_mrna42_eta %>%
+  map(2) %>%
+  map("X") %>%
+  map(~.[match(colnames(.), rownames(.)), ]) %>%
+  map(~.[lower.tri(., diag = FALSE)]) %>%
+  stack() %>%
+  rename(Values = values,
+         Group = ind) %>%
+  ggplot(aes(x = Values, fill = Group)) +
[…]

### 2016-12-06 reports/single_step_kernels.Rmd
Plot coefficients of transformed kernels in report

@@
 ```
-
-```{r Extreme-Kernel-Coefficients, cache = TRUE}
-extreme_cap <- paste("Most extreme values for each predictor set and each", 
-                     "heterotic group.")
-W_df %>%
-  group_by(Predictor) %>%
-  summarize(Min = round(min(Values), digits = 3),
-            Max = round(max(Values), digits = 3)) %>%
-  ungroup() %>%
-  knitr::kable(caption = extreme_cap, booktabs = TRUE)
-```
-
-```{r Kernel-Closeup, cache = TRUE, fig.cap = "Distribution of kernel coefficients used in single step hybrid prediction for a scenario with complete genomic and incomplete transcriptomic information ('snp100-mrna42-none') as well as a scenario with complete pedigree, incomplete genomic and incomplete transcriptomic information ('ped100-snp77-mrna42').
-The focus was placed on coefficients in the range from -0.1 to +0.1."}
-W_df %>%
-  filter(Predictor %in% c("snp100_mrna42_none", "ped100_snp77_mrna42")) %>%
-  ggplot(aes(x = Values, fill = Group)) +
-  geom_histogram(alpha = 0.5, position = "identity", bins = 80) +
-  facet_wrap(~ Predictor, scales = "free") +
-  theme_bw() +
-  theme(legend.position = "top") +
-  scale_fill_manual(values = c("#669933", "#FFCC66")) +
-  xlim(c(-0.1, 0.1))
-```

### 2016-12-06 reports/transformed_kernels.Rmd
Plot coefficients of transformed kernels in report

# Kernel Inspection
```{r setup-kernel-transformation, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

Goal: Tranformation of predictor data for improved single-step prediction using 
BGLR.

```{r Raw-Kernel-Transformation, cache = TRUE}
genos <- readRDS("./data/processed/common_genotypes.RDS")
cat_lst <- readRDS("./data/derived/pred_sub_list.RDS")

mrna_inbreds <- genos %>%
  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group))

## Data pre-processing
# Pre-process the data in the following order:
# 1.    Box-Cox tranformation
# 2.    Centering
# 3.    Scaling
# 4.    Remove highly correlated features
mrna_lst <- mrna_df %>% 
  split(.$Group) %>%
  map(as.data.frame) %>%
  map(~column_to_rownames(., var = "G")) %>%
  map(.f = function(x) {
    x$Group <- NULL
    x
  }) %>%
  map(~as.matrix(.))

# Data transformation
trans_lst <- mrna_lst %>%
  map(preProcess,
      method = c("BoxCox", "center", "scale")) %>%
  map2(.y = mrna_lst, .f = predict)

mrna_cor_idx <- trans_lst %>%
  map(cor) %>%
  map(~findCorrelation(x = ., cutoff = 0.999))

mrna42 <- lapply(seq_along(trans_lst), FUN = function(i) {
  mat <- trans_lst[[i]]
  high_cor_idx <- mrna_cor_idx[[i]]
  if (isTRUE(length(high_cor_idx) != 0)) {
    mat <- mat[, -high_cor_idx]
  }
  mat
})
names(mrna42) <- names(mrna_lst)

# Remove perfect LD
snp42 <- cat_lst %>% 
  transpose() %>%
  .[names(.) == "snp42_none_none"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]

snp77 <- cat_lst %>% 
  transpose() %>%
  .[names(.) == "ped100_snp77_none"] %>%
  at_depth(.depth = 2, .f = 2) %>%
  .[[1]]

snp100 <- cat_lst %>%
  transpose() %>%
  .[names(.) == "snp100_none_none"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]

snp_lst <- list(snp100 = snp100, snp77 = snp77, snp42 = snp42)
snp_lst[] <- lapply(snp_lst , FUN = function(x) {
  thinned_snp <- x %>%
    map(unique, MARGIN = 2)
  names(thinned_snp) <- c("Dent", "Flint")
  thinned_snp
})
invisible(list2env(snp_lst, envir = globalenv()))

# Get pedigree data
ped100 <- cat_lst %>% 
  transpose() %>%
  .[names(.) == "ped100_snp77_none"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]
names(ped100) <- c("Dent", "Flint")

ped42 <- cat_lst %>% 
  transpose() %>%
  .[names(.) == "ped42_none_none"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]
names(ped42) <- c("Dent", "Flint")
```

```{r Kernel-assembly, cache = TRUE}
nms100 <- cat_lst %>%
  transpose() %>%
  .[names(.) == "ped100_none_none"] %>%
  at_depth(.depth = 2, "geno") %>%
  .[[1]]

nms42 <- cat_lst %>%
  transpose() %>%
  .[names(.) == "mrna42_none_none"] %>%
  at_depth(.depth = 2, "geno") %>%
  .[[1]]

# Specify the predictor sets.
pred_vec <- c("snp42-none-none", "mrna42-none-none", "snp100-none-none",
              "ped100-snp77-none", "ped100-mrna42-none", "snp100-mrna42-none",
              "ped100-snp77-mrna42")
pred_sets <- pred_vec %>%
  strsplit(., split = "[-]") %>%
  map(.f = function(x) {
    x[x != "none"] 
  })

trans_cat_lst <- lapply(pred_sets, FUN = function(x) {
  cmb_lst <- x %>% 
    map(~get(.))
  names(cmb_lst) <- x
  cmb_lst <- cmb_lst %>% 
    transpose()
  largest_pred_size <- x %>%
    str_extract(., pattern = "[0-9]+") %>%
    as.integer() %>% 
    max()

  if (isTRUE(largest_pred_size == 100)) {
    hyb_names_pred_size <- nms100
  } else {
    hyb_names_pred_size <- nms42
  }
  names(hyb_names_pred_size) <- c("Dent", "Flint")

  cmb_lst[] <- lapply(seq_along(cmb_lst), FUN = function(i) {
    nm <- names(cmb_lst)[i]
    dat <- cmb_lst[[nm]]
    dat$geno <- hyb_names_pred_size[[nm]]
    dat
  })
  cmb_lst
})
names(trans_cat_lst) <- pred_sets %>%
  map(~paste(., collapse = "_")) %>%
  flatten_chr()
trans_cat_lst <- trans_cat_lst %>%
  transpose()

## -- BUILD KERNELS --------------------------------------------------------
# Apply kernel building functions to different predictor sets.
# The choice of function depends on the number of predictors as well as on the
# type of predictor (are pedigree data involved or not?).
ped100_snp77_mrna42 <- trans_cat_lst %>%
  transpose() %>%
  .[["ped100_snp77_mrna42"]] %>%
  map(~impute2(ped = .[[1]],
               snp = .[[2]],
               mrna = .[[3]],
               geno = .[["geno"]],
               as_kernel = TRUE,
               bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

snp100_mrna42 <- trans_cat_lst %>%
  transpose() %>%
  .[["snp100_mrna42"]] %>%
  map(~impute_eta(
     x = .[[1]],
     y = .[[2]],
     geno = .[["geno"]],
     as_kernel = TRUE,
     is_pedigree = FALSE,
     bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

ped100_x <- trans_cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("ped100_mrna42", "ped100_snp77")] %>%
  at_depth(.depth = 2, .f = ~impute_eta(
    x = .[[1]],
    y = .[[2]],
    geno = .[["geno"]], 
    as_kernel = TRUE,
    is_pedigree = TRUE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 2) %>%
  at_depth(.depth = 2, .f = "X")

single_pred <- trans_cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("snp42", "mrna42", "snp100")] %>%
  at_depth(.depth = 2, .f = ~complete_eta(
    x = .[[1]],
    geno = .[["geno"]],
    as_kernel = TRUE,
    is_pedigree = FALSE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 1) %>%
  at_depth(.depth = 2, .f = "X")

# Combine the results in a single list to reduce code replications (the
# following steps are the same for all predictor sets).
W_lst <- c(list(ped100_snp77_mrna42 = ped100_snp77_mrna42),
           list(snp100_mrna42 = snp100_mrna42),
           ped100_x,
           single_pred)
```

```{r Plot-transformed-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the kernel matrices used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

# For each predictor set and each heterotic group, plot the distribution of the
# coefficients from the kernels.
W_df %>%
  ggplot(aes(x = Values, fill = Group)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 80) +
  facet_wrap(~ Predictor, scales = "free") +
  theme_bw() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#669933", "#FFCC66")) +
  xlim(c(-1.0, 1.0))
```

### 2016-12-06 analysis/scaled_mrna_predictions.R

@@ -25,10 +25,10 @@ mrna_df <- mrna %>%
 
 ## Data pre-processing
 # Pre-process the data in the following order:
-# 1.    Near-zero-variance filtering.
-# 2.    Box-Cox tranformation
-# 3.    Centering
-# 4.    Scaling
+# 1.    Box-Cox tranformation
+# 2.    Centering
+# 3.    Scaling
+# 4.    Remove highly correlated features
 mrna_lst <- mrna_df %>% 
   split(.$Group) %>%
   map(as.data.frame) %>%

### 2016-12-07 reports/single_step_kernels.Rmd
Match names of original to names of transformed kernels

@@
 devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071")
+pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
+               "stringr")
 pacman::p_load_gh("mwesthues/sspredr")
@@
 })
+names(cat_lst) <- names(feat_lst)
+
+# Remove empty names (i.e. 'none') from predictor set names.
+cat_lst <- cat_lst %>%
+  map(function(x) {
+    names(x) <- str_replace_all(names(x), pattern = "_none", replacement = "")
+    x
+  })
 saveRDS(cat_lst,
@@
 
-snp100_mrna42_none <- cat_lst %>%
+snp100_mrna42 <- cat_lst %>%
   transpose() %>%
-  .[["snp100_mrna42_none"]] %>%
+  .[["snp100_mrna42"]] %>%
   map(~impute_eta(
@@
 
-ped100_x_none <- cat_lst %>%
+ped100_x <- cat_lst %>%
   transpose() %>%
-  .[names(.) %in% c("ped100_mrna42_none", "ped100_snp77_none")] %>%
+  .[names(.) %in% c("ped100_mrna42", "ped100_snp77")] %>%
   at_depth(.depth = 2, .f = ~impute_eta(
@@
 
-pedx_none_none <- cat_lst %>%
+pedx <- cat_lst %>%
   transpose() %>%
-  .[names(.) %in% c("ped100_none_none", "ped42_none_none")] %>%
+  .[names(.) %in% c("ped100", "ped42")] %>%
   at_depth(.depth = 2, .f = ~complete_eta(
@@
   transpose() %>%
-  .[names(.) %in% c("snp42_none_none", "mrna42_none_none", 
-                    "snp100_none_none")] %>%
+  .[names(.) %in% c("snp42", "mrna42", "snp100")] %>%
   at_depth(.depth = 2, .f = ~complete_eta(
@@
 W_lst <- c(list(ped100_snp77_mrna42 = ped100_snp77_mrna42),
-           list(snp100_mrna42_none = snp100_mrna42_none),
-           pedx_none_none,
-           ped100_x_none,
+           list(snp100_mrna42 = snp100_mrna42),
+           pedx,
+           ped100_x,
            single_pred)

### 2016-12-07 reports/transformed_kernels.Rmd
Match names of original to names of transformed kernels

@@
   transpose() %>%
-  .[names(.) == "snp42_none_none"] %>%
+  .[names(.) == "snp42"] %>%
   at_depth(.depth = 2, .f = 1) %>%
@@
   transpose() %>%
-  .[names(.) == "ped100_snp77_none"] %>%
+  .[names(.) == "ped100_snp77"] %>%
   at_depth(.depth = 2, .f = 2) %>%
@@
   transpose() %>%
-  .[names(.) == "snp100_none_none"] %>%
+  .[names(.) == "snp100"] %>%
   at_depth(.depth = 2, .f = 1) %>%
@@
   transpose() %>%
-  .[names(.) == "ped100_snp77_none"] %>%
+  .[names(.) == "ped100_snp77"] %>%
   at_depth(.depth = 2, .f = 1) %>%
@@
   transpose() %>%
-  .[names(.) == "ped42_none_none"] %>%
+  .[names(.) == "ped42"] %>%
   at_depth(.depth = 2, .f = 1) %>%
@@
   transpose() %>%
-  .[names(.) == "ped100_none_none"] %>%
+  .[names(.) == "ped100"] %>%
   at_depth(.depth = 2, "geno") %>%
@@
   transpose() %>%
-  .[names(.) == "mrna42_none_none"] %>%
+  .[names(.) == "mrna42"] %>%
   at_depth(.depth = 2, "geno") %>%
@@
 # Specify the predictor sets.
-pred_vec <- c("snp42-none-none", "mrna42-none-none", "snp100-none-none",
-              "ped100-snp77-none", "ped100-mrna42-none", "snp100-mrna42-none",
-              "ped100-snp77-mrna42")
-pred_sets <- pred_vec %>%
-  strsplit(., split = "[-]") %>%
-  map(.f = function(x) {
-    x[x != "none"] 
-  })
+pred_sets <- c(
+  "snp42", "mrna42", "snp100", "ped100-snp77", "ped100-mrna42",
+  "snp100-mrna42", "ped100-snp77-mrna42"
+  ) %>%
+  strsplit(., split = "[-]")
 

### 2016-12-07 analysis/prediction.R

@@ -76,38 +76,31 @@ test_that("kernel method is defined", {
 })
 
 
-
-
 ## -- DATA SELECTION -----------------------------------------------------
 # Load the predictor data.
-pred_nms <- list(pred1 = pred1, pred2 = pred2, pred3 = pred3)
-pred_lst <- pred_nms %>%
-  keep(~nchar(.) != 0) %>%
-  map(~paste0("./data/derived/predictor_subsets/", ., ".RDS")) %>%
-  map(readRDS)
+pred_combi <- paste(c(pred1, pred2, pred3), collapse = "_")
+if (isTRUE(transformation)) {
+  pred_lst <- readRDS("./data/derived/transformed_pred_sub_list.RDS")
+} else {
+  pred_lst <- readRDS("./data/derived/pred_sub_list.RDS")
+}
 
-# Names of inbred lines for which at least one predictor type has records.
-pred_geno_nms <- pred_lst %>%
-  map(rownames) %>%
-  reduce(union)
+# Select the requested set of predictors.
+pre_eta <- pred_lst %>%
+  map(~keep(., names(.) == pred_combi))
 
 # Select hybrids for whose parent lines at least one predictor has records.
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-hybrid <- genos %>%
-  filter(Pool == "Hybrid") %>%
-  mutate(G = stringr::str_replace(G, pattern = "DF_", replacement = "")) %>%
+hybrid <- pre_eta %>%
+  at_depth(2, "geno") %>%
+  flatten() %>%
+  transpose() %>%
+  map(paste, collapse = "_") %>%
+  flatten_chr() %>%
+  as_data_frame() %>%
+  rename(G = value) %>%
   separate(G, into = c("Dent", "Flint"), sep = "_") %>%
-  filter(Dent %in% pred_geno_nms, Flint %in% pred_geno_nms) %>%
   unite(col = Hybrid, Dent, Flint, sep = "_", remove = FALSE) %>%
   mutate(Hybrid = paste0("DF_", Hybrid))
-
-# Store the names of all Dent and Flint hybrid parents in a list in order to
-# correctly augment the predictor matrices with their help.
-hybrid_parent_nms <- hybrid %>%
-  gather(key = "Group", value = "G", Dent, Flint) %>%
-  split(.$Group) %>%
-  map("G")
-
 
[…]

### 2016-12-07 analysis/prediction.R
Add new transformation-related parameters to prediction scripts

@@ -78,7 +78,10 @@ test_that("kernel method is defined", {
 
 ## -- DATA SELECTION -----------------------------------------------------
 # Load the predictor data.
-pred_combi <- paste(c(pred1, pred2, pred3), collapse = "_")
+pred_combi <- list(pred1, pred2, pred3) %>%
+  keep(nchar(.) != 0) %>%
+  flatten_chr() %>%
+  paste(., collapse = "_")
 if (isTRUE(transformation)) {
   pred_lst <- readRDS("./data/derived/transformed_pred_sub_list.RDS")
 } else {
@@ -196,11 +199,13 @@ eta <- lapply(eta, FUN = function(x) {
 # Determine how much time (hh:mm:ss format) has elapsed since script initiation.
 get_elapsed_time <- function(start_time, tz = "CEST") {
   start_time <- as.POSIXct(start_time)
-  dt <- difftime(Sys.time(), start_time, units = "secs")
-  # Since you only want the H:M:S, we can ignore the date...
-  # but you have to be careful about time-zone issues
-  format(.POSIXct(dt, tz = tz), "%H:%M:%S")
+  sec <- Sys.time() %>%
+    difftime(time1 = ., time2 = start_time, units = "secs") %>%
+    lubridate::seconds_to_period()
+  paste0(sprintf("%02d", c(day(sec), hour(sec), minute(sec), second(sec))),
+         collapse = ":")
 }
+
 
 # Define which runs shall be used, i.e., which genotypes shall be included as 
 # test sets.
@@ -217,8 +222,7 @@ param_df <- expand.grid(Trait = init_traits,
                         Iter = init_iter,
                         Run = run_length)
 param_df$Trait <- as.character(param_df$Trait)
-
-param_df <- param_df[sample(rownames(param_df), size = 150), ]
+param_df <- param_df[sample(rownames(param_df), size = 30), ]
 
 start_time <- Sys.time()
 # Keep track of how long a job is running.
@@ -241,13 +245,15 @@ yhat_lst <- mclapply(seq_len(nrow(param_df)), FUN = function(i) {
   cbind(pred, Iter = iter)
 }, mc.cores = use_cores)
 res <- rbindlist(yhat_lst)
-elapsed_time <- get_elapsed_time(start_time)
-pred_nms <- pred_nms %>% 
-  map_if(., .p = nchar(.) == 0, .f = function(x) {
-    x <- "none"
-    x
-  }) %>%
+
+# If not all three predictors were used, declare the missing ones as "none" and
[…]

### 2016-12-07 moab_scripts/prediction.sh

@@ -28,7 +28,7 @@
 # Write standard output (o) and errors (e) to the same file
 #MOAB -j oe
 #
-#MOAB -v TRAIT=GTM
+#MOAB -v TRAIT=
 #
 #MOAB -v ITER=30000
 #
@@ -39,6 +39,8 @@
 #MOAB -v PI=0.5
 #
 #MOAB -v PRIOR_PI_COUNT=10
+#
+#MOAB -v TRANSFORMATION=
 #
 #MOAB -v PRED1=ped100
 #
@@ -68,6 +70,7 @@ echo "Trait=${TRAIT}\
       VCOV=${VCOV}\
       Pi=${PI}\
       PriorPiCount=${PRIOR_PI_COUNT}\
+      Transformation=${TRANSFORMATION}\
       Pred1=${PRED1}\
       Pred2=${PRED2}\
       Pred3=${PRED3}\
@@ -77,7 +80,8 @@ echo "Trait=${TRAIT}\
 startprog="Rscript --no-save --no-restore --slave\
            ./analysis/prediction.R\
            ${TRAIT} ${ITER} ${MODEL} ${VCOV} ${PI}\
-           ${PRIOR_PI_COUNT} ${PRED1} ${PRED2} ${PRED3} ${RUNS}"
+           ${PRIOR_PI_COUNT} ${TRANSFORMATION} ${PRED1} ${PRED2} ${PRED3}\
+           ${RUNS}"
 
 # Start program
 echo $startprog

### 2016-12-07 reports/transformed_kernels.Rmd
Fix merge conflict

@@
   transpose()
+saveRDS(trans_cat_lst, 
+        file = "./data/derived/transformed_pred_sub_list.RDS")
 

### 2016-12-07 analysis/prediction.R

@@ -32,6 +32,8 @@ if (isTRUE(interactive())) {
   Sys.setenv("VCOV" = "RadenII")
   Sys.setenv("PI" = "0.5")
   Sys.setenv("PRIOR_PI_COUNT" = "10")
+  # Were the data transformed (TRUE) or not?
+  Sys.setenv("TRANSFORMATION" = "TRUE")
   # Main predictor. If 'Pred2' and 'Pred3' are empty, no imputation will take
   # place.
   Sys.setenv("PRED1" = "ped100")
@@ -54,6 +56,7 @@ hypred_model <- as.character(Sys.getenv("MODEL"))
 g_method <- as.character(Sys.getenv("VCOV"))
 Pi <- as.numeric(Sys.getenv("PI"))
 PriorPiCount <- as.integer(Sys.getenv("PRIOR_PI_COUNT"))
+transformation <- as.logical(Sys.getenv("TRANSFORMATION"))
 pred1 <- as.character(Sys.getenv("PRED1"))
 pred2 <- as.character(Sys.getenv("PRED2"))
 pred3 <- as.character(Sys.getenv("PRED3"))
@@ -222,7 +225,8 @@ param_df <- expand.grid(Trait = init_traits,
                         Iter = init_iter,
                         Run = run_length)
 param_df$Trait <- as.character(param_df$Trait)
-param_df <- param_df[sample(rownames(param_df), size = 30), ]
+
+
 
 start_time <- Sys.time()
 # Keep track of how long a job is running.

### 2016-12-07 analysis/prediction.R
Add session_info logging to prediction script

@@ -17,6 +17,9 @@ pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
                "stringi", "lubridate", "readr")
 pacman::p_load_gh("mwesthues/sspredr")
 
+# Load function to save session info on the software used for the analysis in
+# this script.
+source("./software/session_info.R")
 
 ## --------------------------------------------------------------------------
 ## COLLECT AND CHECK ARGUMENTS 
@@ -48,6 +51,10 @@ if (isTRUE(interactive())) {
 if (!interactive()) {
   job_id <- Sys.getenv("MOAB_JOBID")
 } else job_id <- "interactive_00"
+
+# Save session info
+write_session_info(directory = "./data/derived/session_info/",
+                   job_id = job_id)
 
 use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
 init_traits <- as.character(Sys.getenv("TRAIT"))

### 2016-12-07 software/session_info.R

@@ -1,0 +1,20 @@
+write_session_info <- function(directory, job_id) {
+  # Goal: Store session info for the software environment in which this
+  # function is being called.
+  require("devtools")
+  # Ensure that the path to the storage directory ends with a forward slash, 
+  # otherwise add one to the path.
+  path_length <- nchar(directory)
+  if (!isTRUE(substr(directory, start = path_length,
+                     stop = path_length) == "/")) {
+    directory <- c(directory, "/")
+  }
+  writeLines(c(paste("Session info", paste0(rep("-", 68), collapse = "")),
+               capture.output(session_info()[[1]]),
+               "\n",
+               paste("Packages", paste0(rep("-", 68), collapse = "")),
+               capture.output(session_info()[[2]])),
+             con = paste0(directory, job_id, ".txt"))
+}
+
+

### 2016-12-07 reports/intro.Rmd
Update equation numbering after updates to bookdown package

@@
 
-$$
 \begin{equation}
+	(\#eq:animal-model)
 	\mathbf{y} = \mathbf{Xb} + \mathbf{Za} + \mathbf{e},
-	\label{eq:animal_model}
 \end{equation}
-$$
 
@@
 
-$$
 \begin{equation}
-	\label{eq:gebv}
+  (\#eq:gebv)
 	\mathbf{\tilde{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
 \end{equation}
-$$
 
@@
 The breeding values in a mRNA prediction model are in general given by:
-$$
 \begin{equation}
-	\label{eq:mrnaebv}
+	(\#eq:mrnaebv)
 	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
 \end{equation}
-$$
 
@@
 expectation given $\mathbf{g}_2$ and a residual error term:
-$$
 \begin{align}
-	\label{eq:mrna1}
+	(\#eq:mrna1)
 	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
@@
 \end{align}
-$$
 
@@
 
-$$
 \begin{equation}
+  (\#eq:mrna-model)
 \mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Ped} \boldsymbol{\epsilon}_{Ped} 
@@
 \end{equation}
-$$
 
@@
 
-$$
 \begin{equation}
+ (\#eq:mrna-submatrices)
 \mathbf{X} = 
@@
 
-$$
 \begin{align}
+ (\#eq:entangled-augmented-mrna-model)
 \begin{bmatrix}
@@
 
-$$
 \begin{equation}
+ (\#eq:breeding-values)
 \tilde{\mathbf{g}} = 
@@
 \end{equation}
- $$
 
@@
 
-$$
 \begin{align}
+ (\#eq:final-model)
 \mathbf{y} &= \mathbf{Xb} \\ 
@@
 
-$$
 \begin{equation}
+ (\#eq:predicted-performance)
 \tilde{\mathbf{h}} = \mathbf{Z}_{dent}\tilde{\mathbf{g}}_{dent} + \mathbf{Z}_{flint}\tilde{\mathbf{g}}_{flint} 
 \end{equation}
-$$

### 2016-12-07 reports/single_step_kernels.Rmd
Update equation numbering after updates to bookdown package

@@
 BGLR.
+With the original configuration of the predictors, very low predictive abilities
+were obtained for the amalgamation of pedigree, genomic and transcriptomic 
+information, regardless of the phenotypic trait.
+One of the possible causes were inconsistencies in the functions `impute_eta()`
+and `impute2` from the `sspredr` package regarding the factorization of 
+matrices.
+After tackling those bugs, we wanted to revisit the distribution of the 
+coefficients from the matrix $\mathbf{W}$ in \@ref(eq:mrna-model).
 

### 2016-12-07 analysis/prediction.R
Disable transformation option for ped only pred"

@@ -92,6 +92,13 @@ pred_combi <- list(pred1, pred2, pred3) %>%
   keep(nchar(.) != 0) %>%
   flatten_chr() %>%
   paste(., collapse = "_")
+
+if (isTRUE(transformation)) {
+  if (nchar(pred2) == 0 && nchar(pred3) == 0 && grepl("ped", x = pred1)) {
+    stop("No transformation was applied to 'ped100'. Choose a different set!")
+  }
+}
+
 if (isTRUE(transformation)) {
   pred_lst <- readRDS("./data/derived/transformed_pred_sub_list.RDS")
 } else {

### 2016-12-07 reports/index.Rmd
Update equation numbering after updates to bookdown package
Describe kernel transformation in report

@@
 ---
-
-<script type="text/x-mathjax-config">
-MathJax.Hub.Config({
-  TeX: { equationNumbers: { autoNumber: "AMS" } }
-});
-</script>
 
@@
 
-Chapter \@ref(kernel-inspection) gives the distribution of coefficients in the 
+Chapter \@ref(original-kernels) gives the distribution of coefficients in the 
 different kernel matrices used in hybrid prediction.
+
+Chapter \@ref(transformed-kernels) gives the distribution of coefficients in 
+$\mathbf{W}$ after applying centering, scaling, transformations and feature 
+filtering on the original matrices.
 

### 2016-12-07 reports/single_step_kernels.Rmd
Describe kernel transformation in report

@@
-# Kernel Inspection
+# Original Kernels
 ```{r setup-kernel-inspection, message = FALSE, echo = FALSE}
@@
 
-Goal: Inspection of input matrices for single-step hybrid prediction using 
-BGLR.
 With the original configuration of the predictors, very low predictive abilities
@@
 
-```{r Plot-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the kernel matrices used in single step or regular prediction of hybrid performance."}
+```{r Plot-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the unmodified kernel matrices ($\\mathbf{W}$) used in single step or regular prediction of hybrid performance."}
 ## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------

### 2016-12-07 reports/transformed_kernels.Rmd
Describe kernel transformation in report

@@
-# Kernel Inspection
+# Transformed Kernels
 ```{r setup-kernel-transformation, message = FALSE, echo = FALSE}
@@
 
-Goal: Tranformation of predictor data for improved single-step prediction using 
-BGLR.
+After noticing extremely narrow distributions in the coefficients of 
+$\mathbf{W}$ (defined in Eq. \@ref(eq:mrna-model), Fig.
+\@ref(fig:Plot-kernel-coefficients)), we decided to explore the effect of several
+data pre-processing procedures on the distribution of the coefficients.
+Separately for each heterotic group, the following data pre-processing 
+procedures were applied to the transcriptomic data: i) Box-Cox transformation,
+ii) centering, iii) scaling, removal of features with a correlation 
+$\rho \geq 0.999$, using the function `preProcess` from the R package `caret` 
+[@Kuhn2013].
+Separately for each heterotic group and each subset of features (i.e. `snp100`, 
+`snp77` and `snp42`), only SNPs that were not in perfect linkage disequilibrium
+with another marker were kept. 
+From Fig. \@ref(fig:Plot-transformed-kernel-coefficients) we can glean that 
+data processing had a considerable, positive effect on the distribution of 
+coefficients in $\mathbf{W}$, which became wider in the case of predictor 
+combinations such as `ped100_snp77_mrna42` compared to unprocessed features 
+(Fig \@ref(fig:Plot-kernel-coefficients)).
 
@@
 
-```{r Plot-transformed-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the kernel matrices used in single step or regular prediction of hybrid performance."}
+```{r Plot-transformed-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the transformed kernel matrices $\\mathbf{W}$ used in single step or regular prediction of hybrid performance."}
 ## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------

### 2016-12-09 reports/transformed_kernel_predictions.Rmd
Init plots for transformed kernel predictions

--- 2016-11-22 reports/loocv_reduced.Rmd
@@
 log_file <- fread("./data/derived/pred_log.txt")
+ped_log <- log_file %>%
+  filter(Pred1 %in% c("ped100", "ped77", "ped42"),
+         Pred2 == "none",
+         Pred3 == "none")
+log_file %>%
+  filter(Transformation == TRUE)
+trait_order <- levels(log_file$Trait)
+log_file <- log_file %>%
+  select(-Trait) %>%
+  distinct()
 
-pred_df <- log_file %>%
-  .$Job_ID %>%
-  as.list() %>%
-  map(~readRDS(paste0("./data/derived/predictions/", ., ".RDS"))) %>%
-  bind_rows(.id = "id") %>%
-  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait, Iter),
-            by = "Job_ID") %>%
-  filter(Iter == 30000) %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
-  mutate(Trait = fct_recode(Trait,
-    "DMY" = "GTM",
-    "DMC" = "GTS",
-    "FAT" = "FETT",
-    "PRO" = "RPR",
-    "SUG" = "XZ"
-  )) %>%
+pred_lst <- lapply(seq_len(nrow(log_file)), FUN = function(i) {
+  job_id <- log_file[i, Job_ID, ]
+  pred_obj <- readRDS(paste0("./data/derived/predictions/", job_id, ".RDS"))
+  geno_number <- pred_obj %>%
+    .$Geno %>%
+    unique() %>%
+    length()
+  geno_trait_labels <- rep(trait_order, times = geno_number)
+  pred_obj$Trait <- geno_trait_labels
+  pred_obj$Predictor <- log_file[i, Predictor, ]
+  pred_obj 
+})
+
+pred_df <- pred_lst %>%
+  bind_rows() %>% 
+  mutate(Trait = as.factor(Trait)) %>%
   mutate(Trait = fct_relevel(Trait,
                              "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"),
-         Predictor = as.factor(Predictor)) %>%
-  select(-Job_ID, -Iter)
-
-# Update the variable 'Trait' for subsequent joins with 'pred_df'.
-log_file <- log_file %>%
-  filter(Iter == 30000) %>%
-  mutate(Trait = fct_recode(Trait,
-    "DMY" = "GTM",
-    "DMC" = "GTS",
-    "FAT" = "FETT",
-    "PRO" = "RPR",
-    "SUG" = "XZ"
-  )) %>%
-  mutate(Trait = fct_relevel(Trait,
-                             "DMY", "DMC", "ADF", "FAT", "PRO",
-                             "STA", "SUG")) %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
+         Predictor = as.factor(Predictor))
 ```
@@
 # Specify a unified predictor order.
-pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
-                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
+pred_order <- c("snp42-none-none", "mrna42-none-none",
+                "snp100-none-none", "ped100-snp77-none",
                 "ped100-mrna42-none", "snp100-mrna42-none",
@@
   gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
-  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
+  mutate(Predictor = fct_relevel(Predictor, pred_order),
+         Value = as.factor(Value)) %>%
   ggplot(aes(x = `Phenotypic Value`, color = Value)) +
@@
          Group = ifelse(Predictor %in% c("snp42-none-none",
-                                         "ped42-none-none",
                                          "mrna42-none-none"),

### 2016-12-09 analysis/prediction.R

@@ -292,7 +292,7 @@ res <- res %>%
   as.data.table()
 
 log_file <- unique(res[, .(Job_ID, Pred1, Pred2, Pred3, Transformation, 
-                           Elapsed_Time, Trait, Iter, CV, Model, PI,
+                           Elapsed_Time, Iter, CV, Model, PI,
                            PriorPiCount, Start_Time, Runs, Cores), ])
 
 # Reduce the size of the prediction object to the minimum possible size.
@@ -302,7 +302,6 @@ res[, `:=` (Iter = NULL,
             Pred3 = NULL, 
             Transformation = NULL,
             CV = NULL,
-            Trait = NULL,
             Dent = NULL,
             Flint = NULL,
             Runs = NULL,

### 2016-12-09 reports/single_step_kernels.Rmd
Add nested pred-names to untransformed kernels

@@
   y_lvl1 <- hybrid_parent_nms[[i]]
+  lvl1_nms <- x_lvl1 %>%
+    names() %>%
+    str_replace_all(., pattern = "_none", replacement = "") %>%
+    strsplit(., split = "_")
   geno_lst <- lapply(seq_along(x_lvl1), FUN = function(j) {
@@
     y_lvl2 <- y_lvl1[[j]]
+    names(x_lvl2) <- lvl1_nms[[j]]
     x_lvl2[["geno"]] <- y_lvl2

### 2016-12-12 reports/index.Rmd
Update report

@@
 
-Chapter \@ref(loocv-prediction) deals with the results of hybrid prediction for 
+Chapter \@ref(loocv-transformed-kernel-prediction) deals with the results of hybrid prediction for 
 the reduced set of data and a comparison of predictive abilities with the full 

### 2016-12-12 reports/loocv_reduced.Rmd
Update report

@@
 ```{r Load-LOOCV-Data, cache = TRUE}
-log_file <- fread("./data/derived/pred_log.txt")
+log_file <- read_delim("./data/derived/pred_log.txt", delim = "\t") %>%
+  as.data.table() %>%
+  mutate(Job_ID = as.character(Job_ID))
 
 pred_df <- log_file %>%
+  filter(Transformation == FALSE) %>%
   .$Job_ID %>%
@@
   bind_rows(.id = "id") %>%
-  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait, Iter),
+  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Iter),
             by = "Job_ID") %>%
-  filter(Iter == 30000) %>%
   unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
@@
   select(-Job_ID, -Iter)
-
-# Update the variable 'Trait' for subsequent joins with 'pred_df'.
-log_file <- log_file %>%
-  filter(Iter == 30000) %>%
-  mutate(Trait = fct_recode(Trait,
-    "DMY" = "GTM",
-    "DMC" = "GTS",
-    "FAT" = "FETT",
-    "PRO" = "RPR",
-    "SUG" = "XZ"
-  )) %>%
-  mutate(Trait = fct_relevel(Trait,
-                             "DMY", "DMC", "ADF", "FAT", "PRO",
-                             "STA", "SUG")) %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
 ```

### 2016-12-12 reports/transformed_kernel_predictions.Rmd
Update report

--- 2016-11-22 reports/loocv_reduced.Rmd
@@
-# LOOCV-Prediction
-```{r setup-loocv, message = FALSE, echo = FALSE}
+# LOOCV-Transformed-Kernel-Prediction
+```{r setup-transformed-kernel-prediction, message = FALSE, echo = FALSE}
 if (!require("pacman")) install.packages("pacman")
@@
 log_file <- fread("./data/derived/pred_log.txt")
+ped_log <- log_file %>%
+  filter(Pred1 %in% c("ped100", "ped77", "ped42"),
+         Pred2 == "none",
+         Pred3 == "none")
+log_file <- log_file %>%
+  filter(Transformation == TRUE) %>%
+  rbind(., ped_log)
 
@@
   as.list() %>%
-  map(~readRDS(paste0("./data/derived/predictions/", ., ".RDS"))) %>%
-  bind_rows(.id = "id") %>%
-  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait, Iter),
+  map(~paste0("./data/derived/predictions/", ., ".RDS")) %>%
+  map(readRDS) %>%
+  bind_rows() %>%
+  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Iter),
             by = "Job_ID") %>%
-  filter(Iter == 30000) %>%
   unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%

### 2016-12-12 new_factorization.R

@@ -1,28 +1,0 @@
-pacman::p_load("data.table", "tidyverse", "dtplyr", "lubridate", "forcats")
-log_file <- fread("./data/derived/pred_log.txt")
-
-
-log_file <- log_file %>%
-  mutate(Date = as.POSIXct(Date, format = "%Y-%m-%d %H:%M:%S")) %>%
-  filter(Date >= as.POSIXct("2016-11-27"))
-
-pred_df <- log_file %>%
-  .$Job_ID %>%
-  as.list() %>%
-  map(~readRDS(paste0("./data/derived/predictions/", ., ".RDS"))) %>%
-  bind_rows(.id = "id")
-
-pred_df %>%
-  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Trait, Iter),
-            by = "Job_ID") %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
-  select(-Job_ID, -Iter) %>%
-  group_by(Predictor, Trait) %>%
-  summarize(PredAbility = cor(y, yhat)) %>%
-  as.data.frame()
-
-
-
-readRDS("./data/derived/predictions/10094280.RDS") %>%
-  plot(.$y, .$yhat)
-  .[, .(cor(y, yhat)), ]

### 2016-12-12 reports/transformed_kernels.Rmd
Init repeated prediction with different snp77

@@
 names(mrna42) <- names(mrna_lst)
+saveRDS(mrna42, 
+        file = "./data/derived/predictor_subsets/transformed-mrna42.RDS")
 
@@
 snp42 <- cat_lst %>% 
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "snp42"] %>%
@@
 snp77 <- cat_lst %>% 
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "ped100_snp77"] %>%
@@
 snp100 <- cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "snp100"] %>%
@@
 ped100 <- cat_lst %>% 
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "ped100_snp77"] %>%
@@
 ped42 <- cat_lst %>% 
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "ped42"] %>%
@@
 nms100 <- cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "ped100"] %>%
@@
 nms42 <- cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) == "mrna42"] %>%
@@
   cmb_lst <- cmb_lst %>% 
-    transpose()
+    purrr::transpose()
   largest_pred_size <- x %>%
@@
 trans_cat_lst <- trans_cat_lst %>%
-  transpose()
+  purrr::transpose()
 saveRDS(trans_cat_lst, 
@@
 ped100_snp77_mrna42 <- trans_cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[["ped100_snp77_mrna42"]] %>%
@@
 snp100_mrna42 <- trans_cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[["snp100_mrna42"]] %>%
@@
 ped100_x <- trans_cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) %in% c("ped100_mrna42", "ped100_snp77")] %>%
@@
 single_pred <- trans_cat_lst %>%
-  transpose() %>%
+  purrr::transpose() %>%
   .[names(.) %in% c("snp42", "mrna42", "snp100")] %>%

### 2016-12-12 analysis/repeated_snp77_prediction.R

@@ -1,0 +1,377 @@
+# Goal: Generate nine subsets of predictors for later comparisons between their
+# predictive abilities:
+
+if (!require("pacman")) install.packages("pacman")
+if (!require("devtools")) install.packages("devtools")
+pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
+               "purrr", "testthat", "methods", "tibble", "ggplot2", "stringr",
+               "stringi", "lubridate", "readr")
+devtools::install_github("mwesthues/sspredr")
+
+# Load function to save session info on the software used for the analysis in
+# this script.
+source("./software/session_info.R")
+
+
+
+## --------------------------------------------------------------------------
+## COLLECT AND CHECK ARGUMENTS 
+if (isTRUE(interactive())) {
+  # Number of cores
+  Sys.setenv("MOAB_PROCCOUNT" = "3")
+  Sys.setenv("TRAIT" = "")
+  # Number of iterations in BGLR()
+  Sys.setenv("ITER" = "30000")
+  # Main predictor. If 'Pred2' and 'Pred3' are empty, no imputation will take
+  # place.
+  Sys.setenv("PRED1" = "ped100")
+  # If 'Pred3' is empty, 'Pred2' will be imputed via information from 'Pred1'.
+  Sys.setenv("PRED2" = "snp77")
+  # If not empty, this predictor will be imputed.
+  Sys.setenv("PRED3" = "mrna42")
+  # Which sampling fold shall be used for snp77 (1 to 20)?
+  Sys.setenv("REPLICATION" = "1")
+  # Number of genotypes to predict (only for testing!)
+  Sys.setenv("RUNS" = "1")
+}
+
+if (!interactive()) {
+  job_id <- Sys.getenv("MOAB_JOBID")
+} else job_id <- "interactive_00"
+
+# Save session info
+write_session_info(directory = "./data/derived/session_info/",
+                   job_id = job_id)
+
+use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
+init_traits <- as.character(Sys.getenv("TRAIT"))
+init_iter <- as.integer(Sys.getenv("ITER"))
+pred1 <- as.character(Sys.getenv("PRED1"))
+pred2 <- as.character(Sys.getenv("PRED2"))
+pred3 <- as.character(Sys.getenv("PRED3"))
[…]

### 2016-12-12 moab_scripts/repeated_snp77_prediction.sh

@@ -1,0 +1,77 @@
+#!/bin/sh 
+########## Begin MOAB/Slurm header ##########
+#
+# Give job a reasonable name
+#MOAB -N ssBLUP
+#
+# Request number of nodes and CPU cores per node for job
+#MOAB -l nodes=1:ppn=16
+#
+# Request memory per process
+#MOAB -l pmem=3800mb
+#
+# Estimated wallclock time for job
+#MOAB -l walltime=00:22:00:00
+#
+# Job submission directory
+#MOAB -d /pfs/work2/workspace/scratch/ho_westhues-gamazon-0/gamazon
+#
+# Standard output naming
+#MOAB -o $(JOBNAME)_$(JOBID)
+#
+# Queue: Use a single node 
+#MOAB -q singlenode
+#
+# Send email when job begins (b), aborts (a) and ends (e)
+#MOAB -m a
+#
+# Write standard output (o) and errors (e) to the same file
+#MOAB -j oe
+#
+#MOAB -v TRAIT=
+#
+#MOAB -v ITER=30000
+#
+#MOAB -v PRED1=ped100
+#
+#MOAB -v PRED2=
+#
+#MOAB -v PRED3=
+#
+#MOAB -v REPLICATION=
+#
+#MOAB -v RUNS=
+#
+##### **********************************************************************
+########### End MOAB header ##########
+
+echo "Working Directory:                    $PWD"
+echo "Running on host                       $HOSTNAME"
+echo "Job id:                               $MOAB_JOBID"
+echo "Job name:                             $MOAB_JOBNAME"
+echo "Number of nodes allocated to job:     $MOAB_NODECOUNT"
+echo "Number of cores allocated to job:     $MOAB_PROCCOUNT"
+
+# Setup R Environment
+module load math/R/3.3.1
+
+# Echo input variables
+echo "Trait=${TRAIT}\
+      Iter=${ITER}\
+      Pred1=${PRED1}\
+      Pred2=${PRED2}\
+      Pred3=${PRED3}\
+      Transformation=${REPLICATION}\
+      Runs=${RUNS}"
+
+# Set-up program
+startprog="Rscript --no-save --no-restore --slave\
+           ./analysis/prediction.R\
+           ${TRAIT} ${ITER} ${REPLICATION} ${PRED1} ${PRED2} ${PRED3}\
+           ${RUNS}"
+
+# Start program
+echo $startprog
+echo $date
+exec $startprog
+echo $date

### 2017-01-09 reports/methods.Rmd
Use tmp path for temporary BGLR output

--- 2016-10-21 reports/data_overview.Rmd
@@
-# Data Overview
-```{r message = FALSE, echo = FALSE}
+# Methods
+## Genetic Material and Phenotyping
+```{r setup-overview, message = FALSE, echo = FALSE}
 if (!require("pacman")) install.packages("pacman")

### 2017-01-09 analysis/prediction.R

@@ -46,6 +46,8 @@ if (isTRUE(interactive())) {
   Sys.setenv("PRED3" = "mrna42")
   # Number of genotypes to predict (only for testing!)
   Sys.setenv("RUNS" = "")
+  # Output directory for temporary BGLR files
+  Sys.setenv("TMP" = "./tmp")
 }
 
 if (!interactive()) {
@@ -68,6 +70,7 @@ pred1 <- as.character(Sys.getenv("PRED1"))
 pred2 <- as.character(Sys.getenv("PRED2"))
 pred3 <- as.character(Sys.getenv("PRED3"))
 runs <- as.character(Sys.getenv("RUNS"))
+temp <- as.character(Sys.getenv("TMP"))
 
 
 # Input tests
@@ -259,7 +262,7 @@ yhat_lst <- mclapply(seq_len(nrow(param_df)), FUN = function(i) {
                     speed_tst = FALSE,
                     run = run,
                     verbose = FALSE,
-                    out_loc = "./tmp/")
+                    out_loc = temp)
   cbind(pred, Iter = iter)
 }, mc.cores = use_cores)
 res <- rbindlist(yhat_lst)

### 2017-01-09 analysis/repeated_snp77_prediction.R

@@ -34,6 +34,8 @@ if (isTRUE(interactive())) {
   Sys.setenv("REPLICATION" = "2")
   # Number of genotypes to predict (only for testing!)
   Sys.setenv("RUNS" = "1-3")
+  # Output directory for temporary BGLR files
+  Sys.setenv("TMP" = "./tmp")
 }
 
 if (!interactive()) {
@@ -52,6 +54,8 @@ pred2 <- as.character(Sys.getenv("PRED2"))
 pred3 <- as.character(Sys.getenv("PRED3"))
 runs <- as.character(Sys.getenv("RUNS"))
 replication <- as.integer(Sys.getenv("REPLICATION"))
+temp <- as.character(Sys.getenv("TMP"))
+
 
 # Input tests
 poss_traits <- c("GTM", "GTS", "FETT", "RPR", "STA", "XZ", "ADF")
@@ -277,7 +281,7 @@ yhat_lst <- mclapply(seq_len(nrow(param_df)), FUN = function(i) {
                     speed_tst = FALSE,
                     run = run,
                     verbose = FALSE,
-                    out_loc = "./tmp/")
+                    out_loc = temp)
   cbind(pred, Iter = iter)
 }, mc.cores = use_cores)
 res <- rbindlist(yhat_lst)

### 2017-01-09 moab_scripts/prediction.sh

@@ -14,7 +14,7 @@
 #MOAB -l walltime=00:22:00:00
 #
 # Job submission directory
-#MOAB -d /pfs/work2/workspace/scratch/ho_westhues-gamazon-0/gamazon
+#MOAB -d /pfs/work2/workspace/scratch/ho_westhues-single_step-0/gamazon
 #
 # Standard output naming
 #MOAB -o $(JOBNAME)_$(JOBID)

### 2017-01-09 moab_scripts/repeated_snp77_prediction.sh

@@ -14,7 +14,7 @@
 #MOAB -l walltime=00:22:00:00
 #
 # Job submission directory
-#MOAB -d /pfs/work2/workspace/scratch/ho_westhues-gamazon-0/gamazon
+#MOAB -d /pfs/work2/workspace/scratch/ho_westhues-single_step-0/gamazon
 #
 # Standard output naming
 #MOAB -o $(JOBNAME)_$(JOBID)
@@ -67,8 +67,8 @@ echo "Trait=${TRAIT}\
 # Set-up program
 startprog="Rscript --no-save --no-restore --slave\
            ./analysis/repeated_snp77_prediction.R\
-           ${TRAIT} ${ITER} ${REPLICATION} ${PRED1} ${PRED2} ${PRED3}\
-           ${RUNS}"
+           ${TRAIT} ${ITER} ${PRED1} ${PRED2} ${PRED3}\
+           ${REPLICATION} ${RUNS}"
 
 # Start program
 echo $startprog

### 2017-01-31 reports/transformed_kernel_predictions.Rmd
Plot prediction results based on snp77 resampling

@@
 ```{r Load-LOOCV-Data, cache = TRUE}
-log_file <- fread("./data/derived/pred_log.txt")
-ped_log <- log_file %>%
-  filter(Pred1 %in% c("ped100", "ped77", "ped42"),
-         Pred2 == "none",
-         Pred3 == "none")
-log_file <- log_file %>%
-  filter(Transformation == TRUE) %>%
-  rbind(., ped_log)
+# Specify which predictions are based on the updated single-step prediction
+# algorithms in 'sspredr'.
+commit <-  "mwesthues/sspredr@26447a1"
+session_files <- list.files("./data/derived/session_info/")
+session_files <- session_files[grep(session_files,
+                                    pattern = "interactive", invert = TRUE)]
+session_filter <- session_files %>%
+  map(~readLines(paste0("./data/derived/session_info/", .))) %>%
+  grepl(., pattern = commit)
+upd_algo_runs <- session_files[session_filter]
+upd_algo_runs <- gsub(upd_algo_runs, pattern = ".txt", replacement = "")
+rep_log_file <- fread("./data/derived/repeated_snp77_prediction_log.txt") %>%
+  filter(Cores == 16,
+         Job_ID %in% upd_algo_runs)
 
-pred_df <- log_file %>%
+# Load all prediction results based on the resampling of 'snp77'.
+rep_df <- rep_log_file %>%
   .$Job_ID %>%
@@
   bind_rows() %>%
-  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Iter),
+  left_join(y = rep_log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Iter,
+                                        Replication),
             by = "Job_ID") %>%
-  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
-  mutate(Trait = fct_recode(Trait,
+  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
+
+# Load all other predictions, which did not involve 'snp77'.
+reg_log <- fread("./data/derived/pred_log.txt") %>%
+  select(-Model, -PI, -PriorPiCount) %>%
+  filter(Cores == 16,
+        # Job_ID %in% upd_algo_runs,
+         Transformation == TRUE | Pred1 %in% c("ped42", "ped100"),
+         Pred1 != "snp77",
+         Pred2 != "snp77",
+         Pred3 != "snp77") %>% 
+  select(-Transformation)
+reg_df <- reg_log %>%
+  .$Job_ID %>%
+  as.list() %>%
+  map(~paste0("./data/derived/predictions/", ., ".RDS")) %>%
+  map(readRDS) %>%
+  bind_rows() %>%
+  left_join(y = reg_log %>% select(Job_ID, Pred1, Pred2, Pred3, Iter),
+            by = "Job_ID") %>%
+  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")
+
+# Combine all data for ease of handling.
+pred_df <- rbind(reg_df, rep_df %>% select(-Replication))
+pred_df <- pred_df %>%
+  mutate(Trait = as.factor(Trait),
+         Trait = fct_recode(Trait,
     "DMY" = "GTM",
@@
   select(-Job_ID, -Iter)
+```
+
+```{r Reduce-Hybrid-Data, dependson="Load-Agro-Data}
+pred_df <- pred_df %>%
+  filter(Geno %in% red_hybrid)
 ```
@@
 685 hybrids) and 'Full' (i.e.
-1,521 hybrids)."}
+1,521 hybrids).
+For both categories, predictive abilities are based solely on the set of 685 common hybrids."}
 overview_df %>%
@@
 ```
+
+```{r Resampled-Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for two scenarios and seven agronomic traits where the composition of the set of genomic predictors 'snp77' is based on sampling at random for 20 times.
+Predictive abilities are based solely on the set of 685 common hybrids."}
+# Plot the distribution of predictive abilities over 20 rounds of resampled
+# 'snp77'.
+rep_df %>%
+  mutate(Trait = as.factor(Trait),
+         Trait = fct_recode(Trait,
+    "DMY" = "GTM",
+    "DMC" = "GTS",
+    "FAT" = "FETT",
+    "PRO" = "RPR",
+    "SUG" = "XZ"
+  )) %>%
+  mutate(Trait = fct_relevel(
+    Trait, "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"
+    )
+  ) %>%
+  group_by(Trait, Replication, Predictor) %>%
+  summarize(r = cor(y, yhat)) %>%
+  ggplot(aes(x = Predictor, y = r, fill = Predictor)) +
+  geom_boxplot() +
+  facet_wrap(~Trait) +
+  theme_bw(base_size = 10) +
+  theme(axis.text.x = element_blank(),
+        legend.position = "top")
+```

### 2017-01-31 analysis/prediction.R

@@ -47,7 +47,7 @@ if (isTRUE(interactive())) {
   # Number of genotypes to predict (only for testing!)
   Sys.setenv("RUNS" = "1-3")
   # Output directory for temporary BGLR files
-  Sys.setenv("TMP" = "./tmp/")
+  Sys.setenv("TMP" = "./tmp")
 }
 
 if (!interactive()) {

### 2017-01-31 analysis/repeated_snp77_prediction.R

@@ -54,7 +54,7 @@ pred2 <- as.character(Sys.getenv("PRED2"))
 pred3 <- as.character(Sys.getenv("PRED3"))
 runs <- as.character(Sys.getenv("RUNS"))
 replication <- as.integer(Sys.getenv("REPLICATION"))
-temp <- as.character(Sys.getenv("TMP"))
+temp <- paste0(as.character(Sys.getenv("TMP")), "/")
 
 
 # Input tests

### 2017-01-31 reports/data_overview.Rmd
Save names of reduced set of 685 hybrids

@@
   .$Hybrid
+saveRDS(red_hybrid, "./data/processed/hybrids_685.RDS")
 genos <- genos %>%

### 2017-02-01 reports/transformed_kernel_predictions.Rmd
Save names of reduced set of 685 hybrids
Init interpretation of predictive abilities

@@
 
-```{r Reduce-Hybrid-Data, dependson="Load-Agro-Data}
+```{r Reduce-Hybrid-Data, dependson="Load-Agro-Data"}
+red_hybrid <- readRDS("./data/processed/hybrids_685.RDS")
 pred_df <- pred_df %>%
@@
 
+### Findings
+
+a)   Reduced
+
+   -   For dry matter yield, transcriptomic data clearly perform best.
+
+   -   Genomic do not clearly perform better than pedigree data except for ADF
+     and FAT.
+
+   -   Transcripts perform poorly for DMC and STA.
+
+b)   Full
+
+   -   SNP data perform better for all traits compared to pedigree information.
+
+   -   Predictive abilities for the combination of pedigree, SNP and 
+     transcriptomic data "ped100-snp77-mrna42" seem to be the result of the
+     predictive abilities of the individual predictors weighted by their share
+     of the genotypes. 
+     For example 
+     $$r(\text{ped100-snp77-mrna42}) = r(\text{ped100}) * (1 - 0.77) +
+     r(\text{snp77}) * (0.77 - 0.42) + r(\text{mrna42}) * 0.42$$
+
+   -   The predictive ability of only two predictors is typically higher than a
+     weighted average of the predictive abilities of the consituting
+     individual predictors.
+
+   -   Predictive abilities for FAT and the predictor combination
+     "ped100-snp77-mrna42" are difficult to explain.
+
 ```{r Resampled-Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for two scenarios and seven agronomic traits where the composition of the set of genomic predictors 'snp77' is based on sampling at random for 20 times.

### 2017-02-01 reports/transformed_kernel_predictions.Rmd
Expand predictive ability interpretation

@@
 ## Predictive Abilities
+For an introduction to the abbreviations used for the different predictors, and
+combinations thereof, please refer to chapter \@ref(predictor-subsetting).
+
 ```{r Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for seven agronomic traits (rows) and nine predictor sets (columns).
@@
 1,521 hybrids).
-For both categories, predictive abilities are based solely on the set of 685 common hybrids."}
+For both categories, predictive abilities are based solely on the set of 685 common hybrids.
+Predictions including 'snp77' are based on 20 runs where each run was assembled by selecting a set of genotypes at random.
+The resampling procedure took into account the difference in absolute size between sets of Dent and the Flint parent lines.
+In these instances, the arithmetic mean of predictive abilities across all 20 runs is displayed."}
 overview_df %>%
@@
 ### Findings
-
 a)   Reduced
@@
 
-   -   Genomic do not clearly perform better than pedigree data except for ADF
-     and FAT.
+   -   Genomic data do not clearly perform better than pedigree data except for 
+     ADF and FAT.
 
@@
    -   The predictive ability of only two predictors is typically higher than a
-     weighted average of the predictive abilities of the consituting
+     weighted average of the predictive abilities of the constituting
      individual predictors.
@@
 
+c)   General
+
+   -   If a reduced predictor performed better than another predictor in the 
+     reduced data set (e.g. 'mrna42' vs. 'snp42' for DMY), it will boost the 
+     performance of the other predictor (here 'snp100') in the full set when the 
+     two predictors are combined in a single-step approach (here 
+     'snp100-mrna-42-none').
+
+#### Predictive abilities for predictor sets with resampling.
 ```{r Resampled-Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for two scenarios and seven agronomic traits where the composition of the set of genomic predictors 'snp77' is based on sampling at random for 20 times.

### 2017-02-01 reports/outlook.Rmd
Add outlook on Bustos-Korts_2016

# Outlook
## Size of individual predictors
We have not yet investigated the influence of the size (in terms of the number
of parental genotypes) of an individual predictor on the predictive ability
achieved with a combination of multiple predictors in a single-step prediction
settting.
Consider the following predictors in single-step prediction, where nothing
changes but the coverage of genotypes through genomic data in increments of 10%:

1.   'ped100-snp90' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp90-mrna42'

2.   'ped100-snp80' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp80-mrna42'

3.   $\dots$        or  &nbsp;&nbsp;&nbsp;&nbsp; $\dots$

4.   'ped100-snp10' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp10-mrna42'

Using such an iterative approach, we might get closer to describing the
contribution of individual predictors to the predictive ability in a combined
single-step predictor.

One caveat here is that, so far, we have sampled 'snp77' at random (except for
the adjustment to sampling, in relative terms, equally large fractions from the
Dent and the Flint group, respectively).
If we run the scenario proposed above, we would have a high computational load.
What I am looking for is a way to sample 'snpXX' --- where XX reprents the
fraction of all possible genotypes covered --- such that it covers the genetic
target space (i.e. the whole set of 142 parental Dent and 103 parental Flint
lines) equally well.
@Bustos-Korts2016 have described a uniform sampling procedure that we might be
able to use for this purpose.
To evaluate the soundness of the method described in this publication, we could
compare the predictive ability obtained with 'snp77' from uniform sampling with
the mean (or median) of the predictive ability for 'snp77' from repeated random
sampling (20 replications, see Figure
\@ref(fig:Resampled-Predictive-Ability-Plots)).

## Algorithm validation
To validate the algorithms from @Fernando2014 and its extension in
\@ref(introduction), we could perhaps have a look at an independent data set,
such as the one described by @Guo2016, who used data from the [Yan
lab](http://www.maizego.org/Resources.html).

### 2017-02-01 reports/transformed_kernel_predictions.Rmd
Add outlook on Bustos-Korts_2016

@@
 
-   -   Genomic data do not clearly perform better than pedigree data except for 
-     ADF and FAT.
+   -   Genomic data perform decisively better than pedigree data for DMY, ADF 
+     and FAT.
 
-   -   Transcripts perform poorly for DMC and STA.
+   -   Relative to pedigree and genomic data, transcripts perform poorly for 
+     DMC and STA.
 
@@
    -   Predictive abilities for the combination of pedigree, SNP and 
-     transcriptomic data "ped100-snp77-mrna42" seem to be the result of the
+     transcriptomic data "ped100-snp77-mrna42" seem to be the sum of the
      predictive abilities of the individual predictors weighted by their share
@@
    -   Predictive abilities for FAT and the predictor combination
-     "ped100-snp77-mrna42" are difficult to explain.
+     "ped100-snp77-mrna42" are difficult to explain. **Suggestions???**
 

### 2017-03-13 reports/single-step-hypred-report.Rmd
Init report

---
title: Single-step Hybrid Predition
author: Matthias Westhues, Claas Heuer
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
documentclass: book
bibliography: ["library.bib", "cheuer.bib"]
---

# Preface {-}
This report covers an extension of the methodology of single-step prediction 
&mdash; which was developed for the imputation of SNP data from pedigree 
information &mdash; to the imputation of quantitative predictor variables using
genomic information.

Here, we are working with a "full" data set &mdash; referring to the maximum 
number of hybrids for which agronomic and genomic, but not necessarily 
transcriptomic data are available &mdash; and a "reduced" data set with less 
hybrids but all of their parental inbred lines having complete genomic and 
transcriptomic records.

All our predictive abilities were evaluated using a leave-one-out cross 
validation (LOOCV) routine and results are displayed for T0-hybrids only.
Note, that T0-hybrids are defined as hybrids having zero parents that have been
previously tested in other hybrid combinations.

The companion R-package for this analysis is [sspredr](https://github.com/mwesthues/sspredr).
The development version can be installed from Github:
```{r Install_sspred, eval = FALSE}
devtools::install_github("mwesthues/sspredr")
```

## Objectives
1.    Assess whether transcriptomic data with missing records for some genotypes         can be imputed successfully based on genomic information to improve 
      prediction of unseen hybrids.
2.    Explore the utility of various machine learning algorithms for the 
      prediction of unseen hybrids.^[This is motivated by the fact that the 
      success of hybrid prediction is a strongly convex function of predictive 
      ability.]

## Overview
Chapter \@ref(introduction) gives a summary of the relevant literature.

Chapter \@ref(data-overview) provides a brief summary of the available data 
without delving into their properties.

Chapter \@ref(agronomic-data) gives an overview of the distribution of the 
agronomic traits as well as correlations among them, separately for the full
set of hybrids as well as the reduced set of hybrids.

Chapter \@ref(transcriptomic-data) explores the distribution and correlations
of/among transcriptomic predictors for "raw" and transformed mRNA BLUEs.

Chapter \@ref(genotype-wise-mrna-imputation) examines the correlation between
the original and imputed transcriptomic BLUEs.

Chapter \@ref(predictor-subsetting) introduces the different predictor sets that
were used for predictions.

Chapter \@ref(original-kernels) gives the distribution of coefficients in the 
different kernel matrices used in hybrid prediction.

Chapter \@ref(transformed-kernels) gives the distribution of coefficients in 
$\mathbf{W}$ after applying centering, scaling, transformations and feature 
filtering on the original matrices.

Chapter \@ref(loocv-transformed-kernel-prediction) deals with the results of hybrid prediction for 
the reduced set of data and a comparison of predictive abilities with the full 
data set.

<!--chapter:end:index.Rmd-->

# Introduction
The integration of different sources of information for the prediction of 
breeding or production values is one of the major challenges in current genomic 
prediction research.
In the early years of genomic prediction, a 2-step procedure was the most 
intuitive choice for establishing a new selection scheme.
The first consists of conventionally estimating breeding values based on 
pedigree relationships and therefore the phenotypes of relatives.
Individuals with accurate breeding values out of the first step (i.e. high 
number of offspring), enter the second step in which SNP genotypes are regressed 
against the conventional breeding values.
Those markers can then be used to predict genomic breeding values for solely 
genotyped and potentially very young individuals.
Considering the two sources of information &mdash; pedigree and high density SNP 
genotypes &mdash; the process can be described as follows:

### First Step
Denote $\mathbf{y}$ as the vector of measured phenotypes for a normally 
distributed trait for all individuals available in the population.
The relationship &mdash; and therefore the strucure of the additive genetic 
covariance &mdash; across individuals is based on the numerator relationship 
matrix $\mathbf{A}$.
In order to obtain breeding values for any individual in the population, we 
consider the following model:

\begin{equation}
	(\#eq:animal-model)
	\mathbf{y} = \mathbf{Xb} + \mathbf{Za} + \mathbf{e},
\end{equation}

where $\mathbf{X}$ is a design matrix for fixed effects, $\mathbf{b}$ is a 
vector of fixed effects, $\mathbf{Z}$ is the design matrix for the random animal 
effects, which associate observations in $\mathbf{y}$ with elements in 
$\mathbf{A}$, $\mathbf{a}$ is the vector of random animal effects (=breeding 
values) and $\mathbf{e}$ is a vector of iid residual errors with expectation 
$0$ and variance $\sigma^2_e$.
The distribution of the random vector $\mathbf{a}$ is 
$MVN(0, \mathbf{A}\sigma^2_a)$.

### Second Step
In the second step, a subset of the individuals for which conventional breeding 
values have been estimated, is selected and enters the reference population for 
training a genomic prediction model.
That selection requires that the individual is genotyped and has 
$\text{PEV}(a_{\text{candidate}}) \leq \text{threshold}$.
As the breeding values from step one enter the genomic prediction model as 
phenotypes, one has to be careful about the residual error distribution of
such a model.
For instance, if the range of the prediction error variances is very broad in 
the reference population, the residual errors will not be identically 
distributed anymore, e.g. they should be weighted differently or estimated 
individually.
One popular approach to account for that problem is *Deregressing* the breeding 
values and weighting the residuals according to the prediction error variances 
[@Garrick2009].
Genomic breeding values for any animal in the population can then be obtained 
by:

\begin{equation}
  (\#eq:gebv)
	\mathbf{\tilde{g}} =  (\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}},
\end{equation}

where $\mathbf{p}$ is a row-vector of allele frequencies in the base population,
$\mathbf{M}$ is a matrix of marker covariates and $\mathbf{u}$ is a vector of
marker effects.
The prediction error variance of the vector $\mathbf{\tilde{g}}$ is usually 
globally obtained using cross validation. 
Alternatively they could be obtained individually by the posterior distribution 
of $(\mathbf{M} - 2\mathbf{p})\mathbf{\hat{u}}$ when a Bayesian model is used.

## Single Step Transciptomic prediction of Breeding Values
The breeding values in a mRNA prediction model are in general given by:
\begin{equation}
	(\#eq:mrnaebv)
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model 
equaivalent to the GBLUP model but using mRNA covariates ($\mathbf{W}$) instead of 
marker genotypes.

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in @fernando_class_2014 and impute missing
mRNA data.
The difference here is that &mdash; instead of predicting SNP covariates using 
pedigree relationship &mdash; we impute mRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data are available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align}
	(\#eq:mrna1)
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
[@legarra_relationship_2009].

The structure of the residual imputation/prediction error is known and can 
therefore be modelled.
The missing mRNA covariates can be predicted using the expectation of a 
multivariate normal random vector given correlated observations, which are 
related to the Best Linear Predictor.

### Extending the single step model

To this point, we have only taken the general idea behind single step genomic prediction and applied
it to the situation in which the breeding values are defined as the sum of mRNA effects plus an *imputation error*
term from using genomic relationships to impute missing mRNA data.
It might also be interesting to include individuals which have neither mRNA nor SNP data available,
but phenotypes.
The general mRNA model for that situation could be written as:

\begin{equation}
  (\#eq:mrna-model)
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Ped} \boldsymbol{\epsilon}_{Ped} 
+ \mathbf{U}_{SNP} \boldsymbol{\epsilon}_{SNP}+ \mathbf{e},
\end{equation}

with

\begin{equation}
 (\#eq:mrna-submatrices)
\mathbf{X} = 
\begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_0\hat{\mathbf{W}_0} \\
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}_{Ped} = 
\begin{bmatrix}
  Z_0 \\
  0 \\
  0 
 \end{bmatrix},
 \mathbf{U}_{SNP} = 
\begin{bmatrix}
  0 \\
  Z_1 \\
  0 
 \end{bmatrix},
\end{equation}

The subscript 0 denotes animals with Pedigree information and neither
SNP nor mRNA data, 1 individuals with SNP but no mRNA data and 2 individuals with mRNA data.

Following the notation in @fernando_class_2014 the phenotypes can be untangled like this:

\begin{align}
 (\#eq:entangled-augmented-mrna-model)
\begin{bmatrix}
  y_0 \\
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_0 \\
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{A}_{02}\mathbf{A}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_0 \\
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_0 & 0 & 0 \\
  0 & Z_1 & 0 \\
  0 & 0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_0}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Ped} \\
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{SNP} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

The breeding values from that model are:

\begin{equation}
 (\#eq:breeding-values)
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_0} \\
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_0 \\
  0 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{Ped}
 + 
 \begin{bmatrix}
  0 \\
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{SNP} 
\end{equation}

When dealing with hybrid performance as phenotypes, we are interested in estimating
general combining abilitiy (GCA) for every line of a specific subpopulation.
In our case, the subpopulations comprise *dent* and *flint*.
Hence, our final model looks like this:

\begin{align}
 (\#eq:final-model)
\mathbf{y} &= \mathbf{Xb} \\ 
  &+ \mathbf{Z}_{dent}\mathbf{W}_{dent} \boldsymbol{\alpha}_{dent} + \mathbf{Z}_{dent}\mathbf{U}_{dent_{Ped}} \boldsymbol{\epsilon}_{dent_{Ped}}
+ \mathbf{Z}_{dent}\mathbf{U}_{dent_{SNP}} \boldsymbol{\epsilon}_{dent_{SNP}} \\
  &+ \mathbf{Z}_{flint}\mathbf{W}_{flint} \boldsymbol{\alpha}_{flint} + \mathbf{Z}_{flint}\mathbf{U}_{flint_{Ped}} \boldsymbol{\epsilon}_{flint_{Ped}}
+ \mathbf{Z}_{flint}\mathbf{U}_{flint_{SNP}} \boldsymbol{\epsilon}_{flint_{SNP}} \\
  &+ \mathbf{e},
\end{align}

where $\mathbf{Z}_{dent}$ and $\mathbf{Z}_{flint}$ are design matrices that map hybrids to their respective
parents in the dent and flint lines.
The obtained GCA effects for the inbred lines represent half their breeding values.
The predicted hybrid performance is:

\begin{equation}
 (\#eq:predicted-performance)
\tilde{\mathbf{h}} = \mathbf{Z}_{dent}\tilde{\mathbf{g}}_{dent} + \mathbf{Z}_{flint}\tilde{\mathbf{g}}_{flint} 
\end{equation}

<!--chapter:end:intro.Rmd-->

# Data Overview
```{r setup-overview, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "purrr")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load-Agro-Data, message = FALSE, cache = TRUE}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
genos <- readRDS("./data/processed/common_genotypes.RDS")

full_hybrid <- genos %>%
  filter(Pool == "Hybrid") %>%
  .$G
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% full_hybrid) %>%
  dplyr::select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
traits <- all_pheno %>%
  .$Trait %>%
  levels()

# Hybrids of parents with records for both endophenotypes
red_inbreds <- genos %>% 
  filter(Pool != "Hybrid", Data_Type %in% c("snp", "mrna")) %>%
  split(.$Pool) %>%
  map(~split(., .$Data_Type)) %>%
  at_depth(.depth = 2, "G") %>%
  map(function(x) intersect(x$mrna, x$snp))

red_hybrid <- full_hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% red_inbreds$Dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% red_inbreds$Flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid
saveRDS(red_hybrid, "./data/processed/hybrids_685.RDS")
genos <- genos %>%
  rbind(tibble(G = red_hybrid,
               Pool = "Hybrid",
               Data_Type = "mrna")
        )
```

```{r Feature-Number, echo = FALSE}
snp <- readRDS("./data/processed/snp_mat.RDS")
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
```

Agronomic data for `r length(traits)` traits (including dry-matter yield (DMY))
and `r length(full_hybrid)` hybrids are available 
(Table \@ref(tab:Geno-Number)).
After applying a series of quality checks, `r ncol(snp)` SNPs and 
best linear unbiased estimates of `r ncol(mrna)` were available for further 
analyses.
Genomic, pedigree and transcriptomic records, were collected for inbred parent 
lines of `r length(red_hybrid)` hybrids.

```{r Geno-Number, echo = FALSE}
geno_tab_cap <- paste("Number of genotypes for Dent and Flint parent lines and",
                      "their hybrid progeny for which genomic and/or",
                      "transcriptomic data are available.",
                      "All genotypes for which transcriptomic data are", 
                      "available, also have genomic data.",
                      "Agronomic data are available for all 1,521 hybrid",
                      "genotypes.")
# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
# parental Dent and Flint lines, respectively.
genos %>% 
  mutate(Data_Type = fct_recode(Data_Type,
                                Agronomic = "agro",
                                Transcriptomic = "mrna",
                                Genomic = "snp",
                                Pedigree = "ped")) %>%
  group_by(Pool, Data_Type) %>%
  count() %>%
  rename(`Data Type` = Data_Type) %>%
  knitr::kable(caption = geno_tab_cap, booktabs = TRUE)
```

<!--chapter:end:data_overview.Rmd-->

# Agronomic Data
```{r setup-agro, echo = FALSE, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "purrr")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, echo = FALSE)
```

```{r echo = FALSE}
# Load data
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")

# Prepare agronomic data.
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos %>%
  filter(Data_Type == "agro", Pool == "Hybrid") %>%
  .$G
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))

# Select hybrids for which both parents have genomic as well as transcriptomic
# records.
mrna_inbreds <- genos %>%
  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")
hybrid <- hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% mrna_inbreds$Dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% mrna_inbreds$Flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid

# All hybrids
all_pheno <- all_pheno %>%
  mutate(Reduced_Data = "no")

# Only hybrids for which parents have both, genomic and transcriptomic data
reduced_pheno <- all_pheno %>%
  mutate(Reduced_Data = ifelse(G %in% hybrid, yes = "yes", no = "no")) %>%
  filter(Reduced_Data == "yes")
```

## Skewness
A general rule of thumb to consider is that skewed data whose ratio of the 
highest value to the lowest value is greater than 20 have significant skewness.
The formula for the skewness statistic is
$$
\text{skewness}  = \frac{\sum (x_i - \bar{x})^{3}}{(n - 1)v^{3/2}},
$$
where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ denoting the 
predictor variable, $n$ denoting the number of values and $\bar{x}$ denoting the 
sample mean of the predictor.
```{r Agro-Skewness, echo = FALSE}
skew_tab_cap <- paste(
  "Skewness of seven agronomic traits for the full set of hybrids and the",
  "reduced set of hybrids, respectively."
)
all_skewness <- all_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
reduced_skewness <- reduced_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
full_join(x = all_skewness, y = reduced_skewness,
          by = "Trait",
          suffix = c(" (Full)", " (Reduced)")) %>%
  kable(caption = skew_tab_cap, booktabs = TRUE)
```

## Distribution
```{r Agro-Distribution, message = FALSE, echo = FALSE, fig.cap = "Histogram of the distribution of seven agronomic traits. Black bins show the distribution of the reduced set of hybrids whereas light grey bins show the distribution of the full set of hybrids."}
ggplot(all_pheno, aes(x = Observed)) +
  geom_histogram(fill = "grey", alpha = 0.5) +
  geom_histogram(data = reduced_pheno, color = "black") +
  facet_wrap(~ Trait, scales = "free") +
  theme_bw()
```

## Correlation
```{r Agro-Correlation, echo = FALSE, fig.cap = "Pairwise correlations among seven agronomic traits for the full set of hybrids (upper diagonal) and the reduced set of hybrids (lower diagonal), respectively."}
cor_coeffs <- all_pheno %>%
  rbind(., reduced_pheno) %>%
  spread(key = Trait, value = Observed) %>%
  split(.$Reduced_Data) %>%
  map(~select(., -Reduced_Data)) %>%
  map(as.data.frame) %>%
  map(remove_rownames) %>%
  map(column_to_rownames, var = "G") %>%
  map(cor) %>%
  map_at(.at = "no", .f = ~.[upper.tri(., diag = FALSE)]) %>%
  map_at(.at = "yes", .f = ~.[lower.tri(., diag = FALSE)])
traits <- levels(all_pheno$Trait)
cor_mat <- diag(nrow = 7, ncol = 7)
cor_mat[upper.tri(cor_mat, diag = FALSE)] <- cor_coeffs$no
cor_mat[lower.tri(cor_mat, diag = FALSE)] <- cor_coeffs$yes
dimnames(cor_mat) <- list(traits, traits)
cor_mat %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     cl.pos = "b",
                     tl.pos = "d",
                     tl.srt = 60,
                     order = "hclust",
                     addCoef.col = "black",
                     mar = c(0, 1, 1, 0))
```

<!--chapter:end:agronomic_data.Rmd-->

# Transcriptomic Data
Keep only genotypes for which both, transcriptomic and genomic data are 
available.
```{r setup-transcriptomic, echo = FALSE}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, message = FALSE, echo = FALSE)
```

```{r mRNA-Data, cache = TRUE}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
mrna_inbreds <- genos %>%
  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  select(-G)
```

Separately for each heterotic group, run a principal component analysis (PCA) on 
the "raw" mRNA-BLUEs and return the importance of the first three PCs.
```{r Raw-PCA, cache = TRUE}
rm_grp <- function(x) {
  x$Group <- NULL
  x
}
pca_var_imp <- function(x) {
  var_imp <- x$sd ^ 2 / sum(x$sd ^ 2) * 100
  sum(var_imp[1:3])
}
# Variable importance in PCA (first three PCs)
raw_varimp <- mrna_df %>%
  split(.$Group) %>%
  map(rm_grp) %>%
  map(prcomp) %>%
  map(pca_var_imp) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type = "Raw")
```

## Data pre-processing
Pre-process the data in the following order:

1.    Near-zero-variance filtering.
2.    Box-Cox tranformation
3.    Centering
4.    Scaling
```{r Transform-Data, cache = TRUE}
mrna_lst <- mrna_df %>% 
  split(.$Group) %>%
  map(rm_grp) %>%
  map(~as.matrix(.))

# Data transformation
trans_lst <- mrna_lst %>%
  map(preProcess,
      method = c("BoxCox", "center", "scale", "nzv")) %>%
  map2(.y = mrna_lst, .f = predict)

# Variable importance in PCA (first three PCs)
trans_varimp <- trans_lst %>%
  map(prcomp) %>%
  map(pca_var_imp) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type = "Trans")

# Compare variable importances for raw and transformed mRNA data.
rbind(raw_varimp, trans_varimp) %>%
  knitr::kable(caption = "Importance of the first three PCs (in %)",
               digits = 2)
```

```{r Tidy-Data, cache = TRUE}
mrna_lst <- list(Raw = mrna_lst, Trans = trans_lst)
tidy_mrna <- mrna_lst %>% 
  at_depth(.depth = 2, .f = ~as.data.frame(.)) %>%
  at_depth(.depth = 1, .f = ~bind_rows(., .id = "Group")) %>%
  bind_rows(.id = "Type") %>%
  as_tibble() %>%
  gather(key = "mRNA", value = "Expression", -Group, -Type) %>%
  mutate(Type = as.factor(Type),
         Group = as.factor(Group))
```

## Skewness
```{r Check-Normality, cache = TRUE}
mrna_skew <- tidy_mrna %>%
  group_by(Group, Type, mRNA) %>%
  summarize(Skewness = e1071::skewness(Expression),
            Shapiro = stats::shapiro.test(Expression)$statistic)
```

```{r, mRNA-Skewness, cache = TRUE, fig.cap = "Distribution of skewness values across all mRNAs, separately for transformed and raw BLUEs."}
mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  xlim(c(-2, 2)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top") +
  ylab("Density")
```

```{r Association-Shapiro-Wilks-Skewness, cache = TRUE, fig.cap = "Shapiro-Wilks Test Statistic vs absolute skewness for transcriptomic BLUEs depending on whether the BLUEs were transformed or not."}
mrna_skew %>%
  mutate(Abs_Skewness = abs(Skewness)) %>%
  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
  xlim(c(0.8, 1)) +
  ylim(c(0, 2)) +
  geom_point() +
  facet_grid(Group ~ Type) +
  xlab("Shapiro-Wilks Test Statistic") +
  ylab("Absolute Skewness")
```

<!--chapter:end:transcriptomic_data.Rmd-->

# Genotype-Wise mRNA-Imputation
```{r setup-impute, echo = FALSE}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis",
               "forcats", "knitr", "rmarkdown", "purrr", "rprojroot",
               "methods", "parallel")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 14, fig.height = 10, 
                      warning = FALSE, message = FALSE, echo = FALSE)
```

```{r Predictor-Data, cache = TRUE}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
mrna_inbreds <- genos %>%
  filter(Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")

# Remove the 'Group' variable after splitting Dent and Flint data.
rm_grp <- function(x) {
  x$Group <- NULL
  x
}

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_lst <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  as.data.frame() %>%
  column_to_rownames(var = "G") %>%
  split(.$Group) %>%
  map(rm_grp) %>%
  map(~as.matrix(.))

# SNPs
snp <- readRDS("./data/processed/snp_mat.RDS")
snp_lst <- snp %>%
  as_data_frame() %>%
  mutate(G = rownames(snp)) %>%
  filter(G %in% (mrna_inbreds %>% flatten_chr())) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  as.data.frame() %>%
  column_to_rownames(var = "G") %>%
  split(.$Group) %>%
  map(rm_grp) %>%
  map(~as.matrix(.))

# Pedigree
ped <- readRDS("./data/processed/ped-datafull-GTP.RDS")
ped_lst <- ped %>%
  mutate(G = rownames(ped)) %>%
  filter(G %in% (mrna_inbreds %>% flatten_chr())) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  as.data.frame() %>%
  column_to_rownames(var = "G") %>%
  split(.$Group) %>%
  map(rm_grp) %>%
  map(as.matrix) %>%
  map(function(x) x[, match(rownames(x), colnames(x))]) %>%
  map(function(x) x * 2) %>%
  map(function(x) {
    x[is.na(x)] <- 0
    x
  })
```

```{r Impute-mRNA-Function}
## Function for the imputation of a single mRNA
impute_mrna_genotype <- function(x, y) {
  geno <- rownames(x)
  # Names of genotypes for which transcriptomic records exist
  nm2 <- rownames(y)
  # Names of genotypes for which transcriptomic records are missing.
  nm1 <- setdiff(rownames(x), nm2)
  # Hybrid parents not in y.
  geno1 <- geno[geno %in% nm1]
  # Hybrid parents in y.
  geno2 <- geno[geno %in% nm2]
  x <- x[match(unique(geno), rownames(x)), ]
  # Design matrix mapping GCA effects from genotypes with transcriptomic
  # records to y2.
  Z2 <- Matrix::sparse.model.matrix(~-1 + factor(geno2),
                                    drop.unused.levels = FALSE)
  colnames(Z2) <- gsub("factor\\(geno2\\)", replacement = "", x = colnames(Z2))
  Z2 <- Z2[, match(nm2, colnames(Z2))]
  rownames(Z2) <- geno2
  # Hybrid parents not in y.
  geno1 <- geno[geno %in% nm1]
  # Hybrid parents in y.
  geno2 <- geno[geno %in% nm2]
  x <- x[match(unique(geno), rownames(x)), ]
  y <- y[nm2, ]
  M2 <- y[, matrixStats::colVars(y) != 0]
  x <- x[, matrixStats::colVars(x) != 0]
  A <- build_kernel(M = x, lambda = 0.01, algorithm = "RadenII")
  A11 <- A[nm1, nm1, drop = FALSE]
  A12 <- A[nm1, nm2, drop = FALSE]
  A21 <- A[nm2, nm1, drop = FALSE]
  A22 <- A[nm2, nm2, drop = FALSE]
  Ainv <- solve(A)
  dimnames(Ainv) <- dimnames(A)
  A_up11 <- Ainv[nm1, nm1]
  A_up12 <- Ainv[nm1, nm2]
  # Eq.21
  M1 <- A12 %*% solve(A22) %*% M2
  J2 <- matrix(-1, nrow = ncol(A12), ncol = 1)
  # Eq.22
  J1 <- A12 %*% solve(A22) %*% J2
  # Eq.10
  epsilon <- t(chol(solve(Ainv[nm1, nm1])))
  # Eq.20
  W2 <- Z2 %*% M2
  w1 <- A12 %*% solve(A22) %*% Z2 %*% M2
  as.matrix(w1)
}

# Shuffled version of the imputation function above.
shuffled_imputation <- function(x, y) {
  cor_lst <- lapply(seq(from = 1, to = nrow(y) - 2), FUN = function(i) {
    idx <- sample(nrow(y), size = i, replace = FALSE)
    z <- y[-idx, ]
    imp_mrna <- impute_mrna_genotype(x = x, y = z)
    orig_mrna <- y[match(rownames(imp_mrna), rownames(y)), , drop = FALSE]
    stopifnot(identical(rownames(imp_mrna), rownames(orig_mrna)))
    cor(c(imp_mrna), c(orig_mrna))
  })
  cor_lst %>%
    flatten_dbl() %>%
    tbl_df() %>%
    rownames_to_column(var = "Number_NA_Geno") %>%
    rename(Value = value)
}
```

## Correlations between vectors of original and imputed mRNAs.
### General idea:
*    Impute transcriptomic data for different numbers of genotypes (from 1 to
     number of genotypes with transcriptomic data).
*    For each number of genotypes without transcriptomic data, randomly sample 
     1000 sets of genotypes, which will be imputed.
     This will allow us to aggregate and summarize the correlation between 
     observed and imputed mRNAs with high confidence.

The computations were run using multiple cores with 
[this script](../analysis/impute_single_mrna.R).
```{r Shuffle-Genotypes-Correlate-Imputation, eval = FALSE}
repl <- 1000
donors <- c("snp_lst", "ped_lst")
imp_lst <- lapply(donors, FUN = function(type) {
  par_lst <- vector(mode = "list", length = repl)
  names(par_lst) <- paste0("Replication_", seq_len(repl))
  par_lst[] <- mclapply(par_lst, FUN = function(iter, ...) {
    try(map2(.x = get(type), .y = mrna_lst, .f = shuffled_imputation))
  }, mc.cores = use_cores)
  par_lst
})
names(imp_lst) <- donors
```

```{r Plot-Correlation-Imputed-vs-NonImputed, cache = TRUE, fig.width = 12, fig.height = 12, fig.cap = "Pearson correlation coefficients between original and imputed mRNA genotypes for two heterotic groups and two sources of information (i.e. genomic and pedigree), depending on the number of genotypes without mRNA records. Genotypes without mRNA records were declared as such at random."}
par_lst <- readRDS("./data/derived/impute_single_mrna.RDS")
par_lst %>%
  map(~keep(., is_list)) %>%
  at_depth(.depth = 2, ~bind_rows(., .id = "Group")) %>%
  map(~bind_rows(., .id = "Replication")) %>%
  bind_rows(.id = "Data_Type") %>%
  mutate(Data_Type = fct_recode(Data_Type,
                                Genomic = "snp_lst",
                                Pedigree = "ped_lst")) %>%
  mutate(Number_NA_Geno = as.factor(Number_NA_Geno)) %>%
  mutate(Number_NA_Geno = fct_inorder(Number_NA_Geno)) %>%
  ggplot(aes(x = Number_NA_Geno, y = Value)) +
  geom_boxplot() +
  facet_wrap(Data_Type ~ Group, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1)) +
  xlab("Number of Genotypes Without mRNA-Data") +
  ylab("Correlation Between Original and Imputed mRNA BLUEs")
```

<!--chapter:end:impute_single_mrna.Rmd-->

# Predictor Subsetting
```{r setup-predictor-subsets, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
               "stringr", "forcats")

devtools::install_github("mwesthues/sspredr")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

Goal: Generate nine subsets of predictors for later comparisons between their
predictive abilities:

a.    100% SNP (`snp100`)
b.    100% pedigree (`ped100`)
c.    42% mRNA (`mrna42`)
d.    42% pedigree (`ped42`)
e.    42% SNP (`snp42`)
f.    100% pedigree & 77% SNP (`ped100_snp77`)
g.    100% pedigree & 42% mRNA (`ped100_mrna42`)
h.    100% SNP & 42% mRNA (`snp100_mrna42`)
i.    100% pedigree & 77% SNP & 42% mRNA (`ped100_snp77_mrna42`)

Make comparisons between the following scenarios:

1.    a, b, f, g, h, i
2.    c, d, e

Throughout this manuscript, digits behind the variable name of each predictor
(ped: pedigree, snp: genomic, mrna: transcriptomic) indicate the size of 
the subset relative to the full set of genotypes (n = 245).
For example, we have transcriptomic data for 103 genotypes, which are roughly 
42% of 245 genotypes, hence the name 'mrna42' and so forth.
For the set "snp77", all genotypes &mdash; for which transcriptomic records are 
available &mdash; were used and the remaining genotypes were sampled at random
with equally large proportions being sampled from the Dent and the Flint 
heterotic group, respectively.

```{r Prepare-Predictor-Subsetting, cache = TRUE, results = "hide"}
# -- Genotype information.
# Information on the set of genotypes.
genos <- readRDS("./data/processed/common_genotypes.RDS")
genos %>%
  filter(Pool %in% c("Dent", "Flint")) %>%
  group_by(Pool, Data_Type) %>%
  count()

# Get the names of the Dent Hybrid-parents.
geno <- genos %>%
  filter(Pool == "Hybrid") %>%
  .$G %>%
  map(~stringr::str_split(., pattern = "_")) %>%
  at_depth(.depth = 2, .f = 2) %>%
  flatten() %>%
  as_vector()

# -- Load predictor data.
# pedigree
ped100 <- readRDS("./data/processed/ped-datafull-GTP.RDS")
ped_nms <- genos %>%
  filter(Data_Type == "ped") %>%
  .$G
ped100 <- as.matrix(ped100) * 2
ped100 <- ped100 %>%
  .[rownames(.) %in% ped_nms, colnames(.) %in% ped_nms]
ped100[is.na(ped100)] <- 0

# snp
snp100 <- readRDS("./data/processed/snp_mat.RDS")
snp_nms <- genos %>%
  filter(Data_Type == "snp") %>%
  .$G
snp100 <- snp100 %>%
  .[rownames(.) %in% snp_nms, ]

# mrna
mrna42 <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna_nms <- genos %>%
  filter(Data_Type == "mrna") %>%
  .$G
mrna42 <- mrna42 %>%
  as.matrix() %>%
  t() %>%
  .[rownames(.) %in% mrna_nms, ]
```

```{r Resample-Predictor-Data, cache = TRUE, results = "hide"}
## Resample predictor data 
# Remove some genomic records so that we can impute them via pedigree 
# information.
set.seed(314)
snp35 <- snp100 %>%
  as_tibble() %>%
  mutate(G = rownames(snp100)) %>%
  filter(!G %in% rownames(mrna42)) %>%
  left_join(genos %>% filter(Data_Type == "snp") %>% select(-Data_Type),
            by = "G") %>%
  group_by(Pool) %>%
  sample_frac(size = 0.6) %>%
  ungroup() %>%
  select(-Pool) %>%
  as.data.frame() %>%
  column_to_rownames(var = "G") %>%
  as.matrix()
snp77 <- snp100 %>%
  .[rownames(.) %in% rownames(mrna42), ] %>%
  rbind(snp35)
stopifnot(all(rownames(mrna42) %in% rownames(snp77)))

ped42 <- ped100 %>%
  .[match(rownames(mrna42), rownames(.)), 
    match(rownames(mrna42), colnames(.))]
snp42 <- snp100 %>%
  .[match(rownames(mrna42), rownames(.)), ]

# Store the new objects
storage_dir <- "./data/derived/predictor_subsets/"
ls() %>%
  .[. != "snp35"] %>%
  .[stringr::str_detect(., pattern = "[:digit:]")] %>%
  as.list() %>%
  map_if(.p = ~get(.) %>% is.matrix,
         .f = ~saveRDS(get(.), file = paste0(storage_dir, ., ".RDS"),
                       compress = TRUE))
```

<!--chapter:end:predictor_subsetting.Rmd-->

# Original Kernels
```{r setup-kernel-inspection, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

With the original configuration of the predictors, very low predictive abilities
were obtained for the amalgamation of pedigree, genomic and transcriptomic 
information, regardless of the phenotypic trait.
One of the possible causes were inconsistencies in the functions `impute_eta()`
and `impute2` from the `sspredr` package regarding the factorization of 
matrices.
After tackling those bugs, we wanted to revisit the distribution of the 
coefficients from the matrix $\mathbf{W}$ in \@ref(eq:mrna-model).

```{r Prepare-Kernel-Inspection, cache = TRUE}
## -- DATA SELECTION -----------------------------------------------------
# Specify the predictor sets.
preds <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
           "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
           "ped100-mrna42-none", "snp100-mrna42-none", "ped100-snp77-mrna42")

# Load the predictor data.
pred_lst <- preds %>%
  as.list() %>%
  map(~str_split(string = ., pattern = "-")) %>%
  at_depth(.depth = 2, ~discard(., . == "none")) %>%
  at_depth(.depth = 2, as.list) %>%
  at_depth(.depth = 2, 
           .f = ~paste0("./data/derived/predictor_subsets/", ., ".RDS")) %>%
  at_depth(.depth = 3, .f = readRDS) %>%
  flatten()
names(pred_lst) <- preds

# For each predictor set and for each heterotic group, select the corresponding
# hybrids for which phenotpyic data are available.
# Then, store the names of all Dent and Flint hybrid parents in a list in order 
# to correctly augment the predictor matrices with their help.
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos %>%
  filter(Pool == "Hybrid") %>%
  mutate(G = stringr::str_replace(G, pattern = "DF_", replacement = "")) %>%
  separate(G, into = c("Dent", "Flint"), sep = "_")
geno_nms <- pred_lst %>%
  at_depth(.depth = 2, .f = rownames) %>%
  map(.f = ~reduce(., union))
hybrid_parent_nms <- lapply(seq_along(geno_nms), FUN = function(i) {
  ind_nms <- geno_nms[[i]] 
  hybrid %>%
    filter(Dent %in% ind_nms, Flint %in% ind_nms) %>%
    unite(col = Hybrid, Dent, Flint, sep = "_", remove = FALSE) %>%
    mutate(Hybrid = paste0("DF_", Hybrid)) %>%
    gather(key = "Group", value = "G", Dent, Flint) %>%
    split(.$Group) %>%
    map("G")
})
names(hybrid_parent_nms) <- names(pred_lst)
hybrid_parent_nms <- transpose(hybrid_parent_nms)

## -- PREPARE DATA FOR KERNEL BUILDING ------------------------------------
# Split the predictor matrices into Dent and Flint.
# Then, transform the data frames back to matrices in order to be able build 
# kernels from them.
merge_pred <- function(x, genos) {
  y <- genos %>%
    select(G, Pool) %>%
    distinct(G, .keep_all = TRUE)
  x %>%
    as.data.frame() %>%
    rownames_to_column(., var = "G") %>%
    as_tibble() %>%
    left_join(., y = y, by = "G") %>%
    split(., .$Pool)
}
pred_df_to_matrix <- function(x) {
  x %>%
    select(-Pool) %>%
    as.data.frame() %>%
    column_to_rownames(var = "G") %>%
    as.matrix()
}
feat_lst <- pred_lst %>%
  at_depth(.depth = 2,
           .f = merge_pred, genos = genos) %>%
  map(transpose) %>%
  at_depth(.depth = 3, .f = pred_df_to_matrix)
names(feat_lst) <- str_replace_all(preds, pattern = "-", replacement = "_")
feat_lst <- transpose(feat_lst)

# As a requirement of the kernel-generating functions, add the names of the 
# hybrid parents to each predictor-set/group combination.
cat_lst <- lapply(seq_along(feat_lst), FUN = function(i) {
  x_lvl1 <- feat_lst[[i]] 
  y_lvl1 <- hybrid_parent_nms[[i]]
  lvl1_nms <- x_lvl1 %>%
    names() %>%
    str_replace_all(., pattern = "_none", replacement = "") %>%
    strsplit(., split = "_")
  geno_lst <- lapply(seq_along(x_lvl1), FUN = function(j) {
    x_lvl2 <- x_lvl1[[j]]
    y_lvl2 <- y_lvl1[[j]]
    names(x_lvl2) <- lvl1_nms[[j]]
    x_lvl2[["geno"]] <- y_lvl2
    x_lvl2
  })
  names(geno_lst) <- names(x_lvl1)
  geno_lst
})
names(cat_lst) <- names(feat_lst)

# Remove empty names (i.e. 'none') from predictor set names.
cat_lst <- cat_lst %>%
  map(function(x) {
    names(x) <- str_replace_all(names(x), pattern = "_none", replacement = "")
    x
  })
saveRDS(cat_lst,
        file = "./data/derived/pred_sub_list.RDS")

## -- BUILD KERNELS --------------------------------------------------------
# Apply kernel building functions to different predictor sets.
# The choice of function depends on the number of predictors as well as on the
# type of predictor (are pedigree data involved or not?).
ped100_snp77_mrna42 <- cat_lst %>%
  transpose() %>%
  .[["ped100_snp77_mrna42"]] %>%
  map(~impute2(ped = .[[1]],
               snp = .[[2]],
               mrna = .[[3]],
               geno = .[["geno"]],
               as_kernel = TRUE,
               bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

snp100_mrna42 <- cat_lst %>%
  transpose() %>%
  .[["snp100_mrna42"]] %>%
  map(~impute_eta(
     x = .[[1]],
     y = .[[2]],
     geno = .[["geno"]],
     as_kernel = TRUE,
     is_pedigree = FALSE,
     bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

ped100_x <- cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("ped100_mrna42", "ped100_snp77")] %>%
  at_depth(.depth = 2, .f = ~impute_eta(
    x = .[[1]],
    y = .[[2]],
    geno = .[["geno"]], 
    as_kernel = TRUE,
    is_pedigree = TRUE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 2) %>%
  at_depth(.depth = 2, .f = "X")

pedx <- cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("ped100", "ped42")] %>%
  at_depth(.depth = 2, .f = ~complete_eta(
    x = .[[1]],
    geno = .[["geno"]],
    as_kernel = FALSE,
    is_pedigree = TRUE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 1) %>%
  at_depth(.depth = 2, .f = "X")

single_pred <- cat_lst %>%
  transpose() %>%
  .[names(.) %in% c("snp42", "mrna42", "snp100")] %>%
  at_depth(.depth = 2, .f = ~complete_eta(
    x = .[[1]],
    geno = .[["geno"]],
    as_kernel = TRUE,
    is_pedigree = FALSE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 1) %>%
  at_depth(.depth = 2, .f = "X")

# Combine the results in a single list to reduce code replications (the
# following steps are the same for all predictor sets).
W_lst <- c(list(ped100_snp77_mrna42 = ped100_snp77_mrna42),
           list(snp100_mrna42 = snp100_mrna42),
           pedx,
           ped100_x,
           single_pred)
```

```{r Plot-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the unmodified kernel matrices ($\\mathbf{W}$) used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

# For each predictor set and each heterotic group, plot the distribution of the
# coefficients from the kernels.
W_df %>%
  ggplot(aes(x = Values, fill = Group)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 80) +
  facet_wrap(~ Predictor, scales = "free") +
  theme_bw() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#669933", "#FFCC66")) +
  xlim(c(-1.0, 1.0))
```

<!--chapter:end:single_step_kernels.Rmd-->

# Transformed Kernels
```{r setup-kernel-transformation, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("mwesthues/sspredr", update = TRUE)
pacman::p_load("BGLR","data.table", "tidyverse", "dtplyr", "caret", "e1071",
               "stringr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

After noticing extremely narrow distributions in the coefficients of 
$\mathbf{W}$ (defined in Eq. \@ref(eq:mrna-model), Fig.
\@ref(fig:Plot-kernel-coefficients)), we decided to explore the effect of several
data pre-processing procedures on the distribution of the coefficients.
Separately for each heterotic group, the following data pre-processing 
procedures were applied to the transcriptomic data: i) Box-Cox transformation,
ii) centering, iii) scaling, removal of features with a correlation 
$\rho \geq 0.999$, using the function `preProcess` from the R package `caret` 
[@Kuhn2013].
Separately for each heterotic group and each subset of features (i.e. `snp100`, 
`snp77` and `snp42`), only SNPs that were not in perfect linkage disequilibrium
with another marker were kept. 
From Fig. \@ref(fig:Plot-transformed-kernel-coefficients) we can glean that 
data processing had a considerable, positive effect on the distribution of 
coefficients in $\mathbf{W}$, which became wider in the case of predictor 
combinations such as `ped100_snp77_mrna42` compared to unprocessed features 
(Fig \@ref(fig:Plot-kernel-coefficients)).

```{r Raw-Kernel-Transformation, cache = TRUE}
genos <- readRDS("./data/processed/common_genotypes.RDS")
cat_lst <- readRDS("./data/derived/pred_sub_list.RDS")

mrna_inbreds <- genos %>%
  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group))

## Data pre-processing
# Pre-process the data in the following order:
# 1.    Box-Cox tranformation
# 2.    Centering
# 3.    Scaling
# 4.    Remove highly correlated features
mrna_lst <- mrna_df %>% 
  split(.$Group) %>%
  map(as.data.frame) %>%
  map(~column_to_rownames(., var = "G")) %>%
  map(.f = function(x) {
    x$Group <- NULL
    x
  }) %>%
  map(~as.matrix(.))

# Data transformation
trans_lst <- mrna_lst %>%
  map(preProcess,
      method = c("BoxCox", "center", "scale")) %>%
  map2(.y = mrna_lst, .f = predict)

mrna_cor_idx <- trans_lst %>%
  map(cor) %>%
  map(~findCorrelation(x = ., cutoff = 0.999))

mrna42 <- lapply(seq_along(trans_lst), FUN = function(i) {
  mat <- trans_lst[[i]]
  high_cor_idx <- mrna_cor_idx[[i]]
  if (isTRUE(length(high_cor_idx) != 0)) {
    mat <- mat[, -high_cor_idx]
  }
  mat
})
names(mrna42) <- names(mrna_lst)
saveRDS(mrna42, 
        file = "./data/derived/predictor_subsets/transformed-mrna42.RDS")

# Remove perfect LD
snp42 <- cat_lst %>% 
  purrr::transpose() %>%
  .[names(.) == "snp42"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]

snp77 <- cat_lst %>% 
  purrr::transpose() %>%
  .[names(.) == "ped100_snp77"] %>%
  at_depth(.depth = 2, .f = 2) %>%
  .[[1]]

snp100 <- cat_lst %>%
  purrr::transpose() %>%
  .[names(.) == "snp100"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]

snp_lst <- list(snp100 = snp100, snp77 = snp77, snp42 = snp42)
snp_lst[] <- lapply(snp_lst , FUN = function(x) {
  thinned_snp <- x %>%
    map(unique, MARGIN = 2)
  names(thinned_snp) <- c("Dent", "Flint")
  thinned_snp
})
invisible(list2env(snp_lst, envir = globalenv()))

# Get pedigree data
ped100 <- cat_lst %>% 
  purrr::transpose() %>%
  .[names(.) == "ped100_snp77"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]
names(ped100) <- c("Dent", "Flint")

ped42 <- cat_lst %>% 
  purrr::transpose() %>%
  .[names(.) == "ped42"] %>%
  at_depth(.depth = 2, .f = 1) %>%
  .[[1]]
names(ped42) <- c("Dent", "Flint")
```

```{r Kernel-assembly, cache = TRUE}
nms100 <- cat_lst %>%
  purrr::transpose() %>%
  .[names(.) == "ped100"] %>%
  at_depth(.depth = 2, "geno") %>%
  .[[1]]

nms42 <- cat_lst %>%
  purrr::transpose() %>%
  .[names(.) == "mrna42"] %>%
  at_depth(.depth = 2, "geno") %>%
  .[[1]]

# Specify the predictor sets.
pred_sets <- c(
  "snp42", "mrna42", "snp100", "ped100-snp77", "ped100-mrna42",
  "snp100-mrna42", "ped100-snp77-mrna42"
  ) %>%
  strsplit(., split = "[-]")

trans_cat_lst <- lapply(pred_sets, FUN = function(x) {
  cmb_lst <- x %>% 
    map(~get(.))
  names(cmb_lst) <- x
  cmb_lst <- cmb_lst %>% 
    purrr::transpose()
  largest_pred_size <- x %>%
    str_extract(., pattern = "[0-9]+") %>%
    as.integer() %>% 
    max()

  if (isTRUE(largest_pred_size == 100)) {
    hyb_names_pred_size <- nms100
  } else {
    hyb_names_pred_size <- nms42
  }
  names(hyb_names_pred_size) <- c("Dent", "Flint")

  cmb_lst[] <- lapply(seq_along(cmb_lst), FUN = function(i) {
    nm <- names(cmb_lst)[i]
    dat <- cmb_lst[[nm]]
    dat$geno <- hyb_names_pred_size[[nm]]
    dat
  })
  cmb_lst
})
names(trans_cat_lst) <- pred_sets %>%
  map(~paste(., collapse = "_")) %>%
  flatten_chr()
trans_cat_lst <- trans_cat_lst %>%
  purrr::transpose()
saveRDS(trans_cat_lst, 
        file = "./data/derived/transformed_pred_sub_list.RDS")

## -- BUILD KERNELS --------------------------------------------------------
# Apply kernel building functions to different predictor sets.
# The choice of function depends on the number of predictors as well as on the
# type of predictor (are pedigree data involved or not?).
ped100_snp77_mrna42 <- trans_cat_lst %>%
  purrr::transpose() %>%
  .[["ped100_snp77_mrna42"]] %>%
  map(~impute2(ped = .[[1]],
               snp = .[[2]],
               mrna = .[[3]],
               geno = .[["geno"]],
               as_kernel = TRUE,
               bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

snp100_mrna42 <- trans_cat_lst %>%
  purrr::transpose() %>%
  .[["snp100_mrna42"]] %>%
  map(~impute_eta(
     x = .[[1]],
     y = .[[2]],
     geno = .[["geno"]],
     as_kernel = TRUE,
     is_pedigree = FALSE,
     bglr_model = "BRR"
    )
  ) %>%
  map(2) %>%
  map("X")

ped100_x <- trans_cat_lst %>%
  purrr::transpose() %>%
  .[names(.) %in% c("ped100_mrna42", "ped100_snp77")] %>%
  at_depth(.depth = 2, .f = ~impute_eta(
    x = .[[1]],
    y = .[[2]],
    geno = .[["geno"]], 
    as_kernel = TRUE,
    is_pedigree = TRUE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 2) %>%
  at_depth(.depth = 2, .f = "X")

single_pred <- trans_cat_lst %>%
  purrr::transpose() %>%
  .[names(.) %in% c("snp42", "mrna42", "snp100")] %>%
  at_depth(.depth = 2, .f = ~complete_eta(
    x = .[[1]],
    geno = .[["geno"]],
    as_kernel = TRUE,
    is_pedigree = FALSE,
    bglr_model = "BRR"
    )
  ) %>%
  at_depth(.depth = 2, .f = 1) %>%
  at_depth(.depth = 2, .f = "X")

# Combine the results in a single list to reduce code replications (the
# following steps are the same for all predictor sets).
W_lst <- c(list(ped100_snp77_mrna42 = ped100_snp77_mrna42),
           list(snp100_mrna42 = snp100_mrna42),
           ped100_x,
           single_pred)
```

```{r Plot-transformed-kernel-coefficients, cache = TRUE, fig.cap = "Distribution of coefficients that are elements of the transformed kernel matrices $\\mathbf{W}$ used in single step or regular prediction of hybrid performance."}
## -- PLOT KERNEL COEFFICIENTS ---------------------------------------------
# Extract the coefficients from the lower triangle of the kernels for
# subsequent plotting of their distributions.
W_df <- W_lst %>%
  at_depth(.depth = 2, .f = ~.[match(colnames(.), rownames(.)), ]) %>%
  at_depth(.depth = 2, .f = ~.[lower.tri(., diag = FALSE)]) %>%
  map(function(x) {
    names(x) <- c("Dent", "Flint")
    x 
  }) %>%
  map(.f = stack) %>%
  bind_rows(.id = "id") %>%
  rename(Predictor = id, 
         Values = values,
         Group = ind) %>%
  as_tibble()

# For each predictor set and each heterotic group, plot the distribution of the
# coefficients from the kernels.
W_df %>%
  ggplot(aes(x = Values, fill = Group)) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 80) +
  facet_wrap(~ Predictor, scales = "free") +
  theme_bw() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#669933", "#FFCC66")) +
  xlim(c(-1.0, 1.0))
```

<!--chapter:end:transformed_kernels.Rmd-->

# LOOCV-Transformed-Kernel-Prediction
```{r setup-transformed-kernel-prediction, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
               "stringr", "forcats")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load-LOOCV-Data, cache = TRUE}
# Specify which predictions are based on the updated single-step prediction
# algorithms in 'sspredr'.
commit <-  "mwesthues/sspredr@26447a1"
session_files <- list.files("./data/derived/session_info/")
session_files <- session_files[grep(session_files,
                                    pattern = "interactive", invert = TRUE)]
session_filter <- session_files %>%
  map(~readLines(paste0("./data/derived/session_info/", .))) %>%
  grepl(., pattern = commit)
upd_algo_runs <- session_files[session_filter]
upd_algo_runs <- gsub(upd_algo_runs, pattern = ".txt", replacement = "")
rep_log_file <- fread("./data/derived/repeated_snp77_prediction_log.txt") %>%
  filter(Cores == 16,
         Job_ID %in% upd_algo_runs)

# Load all prediction results based on the resampling of 'snp77'.
rep_df <- rep_log_file %>%
  .$Job_ID %>%
  as.list() %>%
  map(~paste0("./data/derived/predictions/", ., ".RDS")) %>%
  map(readRDS) %>%
  bind_rows() %>%
  left_join(y = rep_log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Iter,
                                        Replication),
            by = "Job_ID") %>%
  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")

# Load all other predictions, which did not involve 'snp77'.
reg_log <- fread("./data/derived/pred_log.txt") %>%
  select(-Model, -PI, -PriorPiCount) %>%
  filter(Cores == 16,
        # Job_ID %in% upd_algo_runs,
         Transformation == TRUE | Pred1 %in% c("ped42", "ped100"),
         Pred1 != "snp77",
         Pred2 != "snp77",
         Pred3 != "snp77") %>% 
  select(-Transformation)
reg_df <- reg_log %>%
  .$Job_ID %>%
  as.list() %>%
  map(~paste0("./data/derived/predictions/", ., ".RDS")) %>%
  map(readRDS) %>%
  bind_rows() %>%
  left_join(y = reg_log %>% select(Job_ID, Pred1, Pred2, Pred3, Iter),
            by = "Job_ID") %>%
  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-")

# Combine all data for ease of handling.
pred_df <- rbind(reg_df, rep_df %>% select(-Replication))
pred_df <- pred_df %>%
  mutate(Trait = as.factor(Trait),
         Trait = fct_recode(Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  mutate(Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"),
         Predictor = as.factor(Predictor)) %>%
  select(-Job_ID, -Iter)
```

```{r Reduce-Hybrid-Data, dependson="Load-Agro-Data"}
red_hybrid <- readRDS("./data/processed/hybrids_685.RDS")
pred_df <- pred_df %>%
  filter(Geno %in% red_hybrid)
```

## Distribution of predicted and observed values
```{r Distribution-y-yhat, cache = TRUE, fig.height = 12, fig.cap = "Frequency polygon of observed and predicted values for seven agronomic traits (columns) and nine sets of predictors (rows)."}
# Specify a unified predictor order.
pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
                "ped100-mrna42-none", "snp100-mrna42-none",
                "ped100-snp77-mrna42")
legend_labs <- expression(y, hat(y))
pred_df %>%
  filter(Predictor != "snp77-none-none") %>%
  gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order),
         Value = as.factor(Value)) %>%
  ggplot(aes(x = `Phenotypic Value`, color = Value)) +
  geom_freqpoly() +
  scale_color_manual(values = c("#1f78b4", "#e31a1c"), labels = legend_labs) +
  facet_grid(Predictor ~ Trait, scales = "free") +
  theme_bw() +
  theme(legend.position = "top",
        strip.text.y = element_text(size = 7)) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  ylab("Frequency")
```

## Coefficient of variation for predicted values
We are interested in the variation in the predicted values relative to the 
arithmetic mean of the predicted values, separately for each agronomic trait and
each set of predictors:

$$
\begin{equation}
  \label{eq:CV}
CV = \sqrt{\frac{\sum_{i = 1}^{N} \sigma^{2}_{\hat{y}(i)} / N}{\sum_{i=1}^{N}\hat{y}_{i} / N}},
\end{equation}
$$

where $N$ denotes the number of hybrids.

```{r Coefficient-of-Variation, cache = TRUE}
coefficient_of_variation <- function(x, y) {
  sqrt(mean(x)) / mean(y)
}
cv_df <- pred_df %>%
  split(list(.$Trait, .$Predictor)) %>%
  map(~coefficient_of_variation(x = .$var_yhat, y = .$yhat)) %>%
  stack() %>%
  separate(col = ind, into = c("Trait", "Predictor"), sep = "[.]") %>%
  rename(CV = values) %>%
  tbl_dt() %>%
  mutate(Predictor = as.factor(Predictor),
         Trait = as.factor(Trait))

overview_df <- pred_df %>%
  group_by(Predictor, Trait) %>%
  summarize(r = cor(y, yhat)) %>%
  ungroup() %>%
  as.data.frame() %>%
  left_join(y = cv_df, by = c("Predictor", "Trait")) %>%
  as_tibble() %>%
  mutate(Predictor = as.factor(Predictor),
         Predictor = forcats::fct_reorder(Predictor, CV),
         Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO",
                             "STA", "SUG"),
         Group = ifelse(Predictor %in% c("snp42-none-none",
                                         "ped42-none-none",
                                         "mrna42-none-none"),
                        yes = "Reduced", no = "Full")
  ) %>%
  mutate(Group = as.factor(Group))
```

```{r Coefficient-of-Variation-Plot, cache = TRUE, fig.cap = "Coefficient of variation for seven agronomic traits and nine sets of predictors."}
overview_df %>%
  filter(Predictor != "snp77-none-none") %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
  ggplot(aes(x = CV, y = Predictor)) +
  geom_segment(aes(yend = Predictor), xend = 0, color = "gray50") +
  geom_point(aes(color = Predictor), size = 3) +
  scale_color_viridis(discrete = TRUE, option = "inferno") +
  facet_wrap(~ Trait, scales = "free_x") +
  theme_bw() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(0.7, 0.13)
  ) +
  guides(color = guide_legend(ncol = 2,
                              reverse = FALSE)) +
  xlab("Coefficient of Variation") +
  ylab("Predictor Set") +
  xlim(0, NA)
```

## Predictive Abilities
For an introduction to the abbreviations used for the different predictors, and
combinations thereof, please refer to chapter \@ref(predictor-subsetting).

```{r Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for seven agronomic traits (rows) and nine predictor sets (columns). The predictor sets are split into the categories 'Reduced' (i.e. 685 hybrids) and 'Full' (i.e. 1,521 hybrids). For both categories, predictive abilities are based solely on the set of 685 common hybrids. Predictions including 'snp77' are based on 20 runs where each run was assembled by selecting a set of genotypes at random. The resampling procedure took into account the difference in absolute size between sets of Dent and the Flint parent lines. In these instances, the arithmetic mean of predictive abilities across all 20 runs is displayed."}
overview_df %>%
  as_tibble() %>%
  filter(Predictor != "snp77-none-none") %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order),
         Group = fct_relevel(Group, "Reduced", "Full")
  ) %>%
  ggplot(aes(x = Predictor, y = r)) +
  geom_bar(stat = "identity", aes(fill = Predictor), color = "black") +
  geom_text(aes(label = round(r, digits = 3)), vjust = -0.2) +
  scale_fill_viridis(discrete = TRUE, option = "inferno") +
  facet_grid(Trait ~ Group, space = "free", scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "bottom",
    legend.key = element_rect(color = "white")
  ) +
  guides(fill = guide_legend(nrow = 3,
                             byrow = TRUE,
                             keywidth = 0.25,
                             keyheight = 0.25,
                             default.unit = "inch",
                             title.position = "top")
  ) +
  ylim(c(0, 1))
```

### Findings
a)   Reduced

   -   For dry matter yield, transcriptomic data clearly perform best.

   -   Genomic data perform decisively better than pedigree data for DMY, ADF 
     and FAT.

   -   Relative to pedigree and genomic data, transcripts perform poorly for 
     DMC and STA.

b)   Full

   -   SNP data perform better for all traits compared to pedigree information.

   -   Predictive abilities for the combination of pedigree, SNP and 
     transcriptomic data "ped100-snp77-mrna42" seem to be the sum of the
     predictive abilities of the individual predictors weighted by their share
     of the genotypes. 
     For example 
     $$r(\text{ped100-snp77-mrna42}) = r(\text{ped100}) * (1 - 0.77) +
     r(\text{snp77}) * (0.77 - 0.42) + r(\text{mrna42}) * 0.42$$

   -   The predictive ability of only two predictors is typically higher than a
     weighted average of the predictive abilities of the constituting
     individual predictors.

   -   Predictive abilities for FAT and the predictor combination
     "ped100-snp77-mrna42" are difficult to explain. **Suggestions???**

c)   General

   -   If a reduced predictor performed better than another predictor in the 
     reduced data set (e.g. 'mrna42' vs. 'snp42' for DMY), it will boost the 
     performance of the other predictor (here 'snp100') in the full set when the 
     two predictors are combined in a single-step approach (here 
     'snp100-mrna-42-none').

#### Predictive abilities for predictor sets with resampling.
```{r Resampled-Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for two scenarios and seven agronomic traits where the composition of the set of genomic predictors 'snp77' is based on sampling at random for 20 times. Predictive abilities are based solely on the set of 685 common hybrids."}
# Plot the distribution of predictive abilities over 20 rounds of resampled
# 'snp77'.
rep_df %>%
  mutate(Trait = as.factor(Trait),
         Trait = fct_recode(Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  mutate(Trait = fct_relevel(
    Trait, "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"
    )
  ) %>%
  group_by(Trait, Replication, Predictor) %>%
  summarize(r = cor(y, yhat)) %>%
  ggplot(aes(x = Predictor, y = r, fill = Predictor)) +
  geom_boxplot() +
  facet_wrap(~Trait) +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_blank(),
        legend.position = "top")
```

<!--chapter:end:transformed_kernel_predictions.Rmd-->

# Outlook
## Size of individual predictors
We have not yet investigated the influence of the size (in terms of the number
of parental genotypes) of an individual predictor on the predictive ability
achieved with a combination of multiple predictors in a single-step prediction
settting.
Consider the following predictors in single-step prediction, where nothing
changes but the coverage of genotypes through genomic data in increments of 10%:

1.   'ped100-snp90' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp90-mrna42'

2.   'ped100-snp80' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp80-mrna42'

3.   $\dots$        or  &nbsp;&nbsp;&nbsp;&nbsp; $\dots$

4.   'ped100-snp10' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp10-mrna42'

Using such an iterative approach, we might get closer to describing the
contribution of individual predictors to the predictive ability in a combined
single-step predictor.

One caveat here is that, so far, we have sampled 'snp77' at random (except for
the adjustment to sampling, in relative terms, equally large fractions from the
Dent and the Flint group, respectively).
If we run the scenario proposed above, we would have a high computational load.
What I am looking for is a way to sample 'snpXX' --- where XX reprents the
fraction of all possible genotypes covered --- such that it covers the genetic
target space (i.e. the whole set of 142 parental Dent and 103 parental Flint
lines) equally well.
@Bustos-Korts2016 have described a uniform sampling procedure that we might be
able to use for this purpose.
To evaluate the soundness of the method described in this publication, we could
compare the predictive ability obtained with 'snp77' from uniform sampling with
the mean (or median) of the predictive ability for 'snp77' from repeated random
sampling (20 replications, see Figure
\@ref(fig:Resampled-Predictive-Ability-Plots)).

## Algorithm validation
To validate the algorithms from @Fernando2014 and its extension in
\@ref(introduction), we could perhaps have a look at an independent data set,
such as the one described by @Guo2016, who used data from the [Yan
lab](http://www.maizego.org/Resources.html).

<!--chapter:end:outlook.Rmd-->

# References

<!--chapter:end:references.Rmd-->

### 2017-03-13 analysis/maizego_corehunter.R

@@ -4,7 +4,8 @@ if (!require("pacman")) install.packages("pacman")
 if (!require("devtools")) install.packages("devtools")
 devtools::install_github("mwesthues/sspredr", update = TRUE)
 options(java.parameters = "-Xmx25G")
-pacman::p_load("tidyverse", "data.table", "dtplyr", "corehunter")
+pacman::p_load("tidyverse", "data.table", "dtplyr", "corehunter", "LEA", 
+               "ggthemes")
 pacman::p_load_gh("mwesthues/sspredr")
 
 # For the second scenario, keep only names of genotypes, which are covered by 
@@ -122,7 +123,7 @@ pc_df <- pc_mat %>%
 pc_df %>%
   ggplot(aes(x = PC_1, y = PC_2, color = Core_Group, shape = Core_Group)) +
   geom_point() +
-  scale_color_tableau() +
+  ggthemes::scale_color_tableau() +
   facet_wrap(~Fraction) +
   theme_bw()
 

### 2017-03-14 analysis/maizego_corehunter.R
Init ADMIXTURE analysis of TST lines

@@ -160,3 +160,9 @@ all_geno_pc %>%
   scale_color_tableau() +
   theme_bw() +
   theme(legend.position = "top")
+
+
+
+# STRUCTURE ANALYSIS ------------------------------------------------------
+
+

### 2017-03-14 analysis/maizego_snp_analyses.R

@@ -1,10 +1,11 @@
 # Data and packages -------------------------------------------------------
 if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "LEA", "stringr", "viridis")
+
+
 snp <- readRDS("./data/processed/maizego/imputed_snp_mat.RDS")
-unique_genotypes <- readRDS(
-  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
-  ) %>%
+unique_genotypes <- "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS" %>%
+  readRDS() %>%
   keep(names(.) == "snp") %>%
   flatten_chr()
 
@@ -12,8 +13,7 @@ unique_genotypes <- readRDS(
 # PCA ---------------------------------------------------------------------
 write.lfmm(snp, "./data/derived/maizego/snp.lfmm")
 
-# Determine the structure of the data using genotypic and gene expression data,
-# respectively.
+# Determine the structure of the data using genotypic data.
 snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                    scale = TRUE)
 # Perform Tracy-Widom tests on all eigenvalues to determine the optimal number
@@ -24,7 +24,7 @@ tw_K <- snp_tw %>%
   flatten_dbl() %>%
   keep(~ .x <= 0.0001) %>%
   length()
-tw_K <- min(c(tw_K, 10))
+tw_K <- min(c(tw_K, 5))
 
 snp_tw %>%
   .["percentage"] %>%
@@ -38,7 +38,7 @@ lfmm2geno("./data/derived/maizego/snp.lfmm",
 project <- snmf(input.file = "./data/derived/maizego/snp.geno",
                 K = seq_len(tw_K),
                 project = "new",
-                repetitions = 10,
+                repetitions = 25,
                 CPU = 3,
                 entropy = TRUE)
 
@@ -75,11 +75,47 @@ pc_with_assignment %>%
   geom_point() 
 
 
-# Evaluate the population structure when considering the first K principal 
-# components.
-K <- 10
-ce <- cross.entropy(project, K = K)
-best <- which.min(ce)
-barplot(t(Q(project, K = K, run = best)), col = viridis(n = K))
 
+# ADMIXTURE ---------------------------------------------------------------
+K_lst <- seq(from = 2, to = tw_K, by = 1) %>%
+  as.list()
[…]

### 2017-03-14 reports/index.Rmd
Remove chapter overview from report index

--- 2016-12-05 reports/index.Rmd
@@
 title: Single-step Hybrid Predition
-author: Matthias Westhues, Claas Heuer
+author:
+  - name: Matthias Westhues
+    email: matthias.westhues@uni-hohenheim.de
+    affiliation: University of Hohenheim
+  - name: Claas Heuer
+    email: cheuer@tierzucht.uni-kiel.de
+    affiliation: University of Kiel
+  - name: Georg Thaller
+    email: thaller@tierzucht.uni-kiel.de
+    affiliation: University of Kiel
+  - name: Rohan Fernando
+    email: rohan@iastate.edu
+    affiliation: Iowa State University
+  - name: Albrecht E. Melchinger
+    email: melchinger@uni-hohenheim.de
+    affiliation: University of Hohenheim
+    footnote: Corresponding Author
+address:
+  - code: University of Hohenheim
+    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
+  - code: University of Kiel
+    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
+  - code: Iowa State University
+    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
 date: "`r Sys.Date()`"
@@
 output:
-  bookdown::html_document2:
-    toc: true
-    toc_float: true
-    theme: readable
-    highlight: tango
-documentclass: book
+  bookdown::pdf_book:
+    base_format: rticles::elsevier_article
 bibliography: ["library.bib", "cheuer.bib"]
-biblio-style: apalike
-link-citations: yes
+abstract: |
+  This is the abstract.
+
+  It consists of two paragraphs.
 ---
-
-<script type="text/x-mathjax-config">
-MathJax.Hub.Config({
-  TeX: { equationNumbers: { autoNumber: "AMS" } }
-});
-</script>
-
-# Preface {-}
-This report covers an extension of the methodology of single-step prediction 
-&mdash; which was developed for the imputation of SNP data from pedigree 
-information &mdash; to the imputation of quantitative predictor variables using
-genomic information.
-
-Here, we are working with a "full" data set &mdash; referring to the maximum 
-number of hybrids for which agronomic and genomic, but not necessarily 
-transcriptomic data are available &mdash; and a "reduced" data set with less 
-hybrids but all of their parental inbred lines having complete genomic and 
-transcriptomic records.
-
-All our predictive abilities were evaluated using a leave-one-out cross 
-validation (LOOCV) routine and results are displayed for T0-hybrids only.
-Note, that T0-hybrids are defined as hybrids having zero parents that have been
-previously tested in other hybrid combinations.
-
-The companion R-package for this analysis is [sspredr](https://github.com/mwesthues/sspredr).
-The development version can be installed from Github:
-```{r Install_sspred, eval = FALSE}
-devtools::install_github("mwesthues/sspredr")
-```
-
-## Objectives
-1.    Assess whether transcriptomic data with missing records for some genotypes         can be imputed successfully based on genomic information to improve 
-      prediction of unseen hybrids.
-2.    Explore the utility of various machine learning algorithms for the 
-      prediction of unseen hybrids.^[This is motivated by the fact that the 
-      success of hybrid prediction is a strongly convex function of predictive 
-      ability.]
-
-## Overview
-Chapter \@ref(introduction) gives a summary of the relevant literature.
-
-Chapter \@ref(data-overview) provides a brief summary of the available data 
-without delving into their properties.
-
-Chapter \@ref(agronomic-data) gives an overview of the distribution of the 
-agronomic traits as well as correlations among them, separately for the full
-set of hybrids as well as the reduced set of hybrids.
-
-Chapter \@ref(transcriptomic-data) explores the distribution and correlations
-of/among transcriptomic predictors for "raw" and transformed mRNA BLUEs.
-
-Chapter \@ref(genotype-wise-mrna-imputation) examines the correlation between
-the original and imputed transcriptomic BLUEs.
-
-Chapter \@ref(predictor-subsetting) introduces the different predictor sets that
-were used for predictions.
-
-Chapter \@ref(kernel-inspection) gives the distribution of coefficients in the 
-different kernel matrices used in hybrid prediction.
-
-Chapter \@ref(loocv-prediction) deals with the results of hybrid prediction for 
-the reduced set of data and a comparison of predictive abilities with the full 
-data set.

### 2017-03-14 reports/methods_data.R
Bundle all info on data for methods description

@@ -1,0 +1,130 @@
+# Goal: Gather all data on the number of genotypes and features that were used
+# throughout this study in a comprehensive manner so that it will we easy to 
+# write and update the manuscript.
+
+
+
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load("tidyverse", "dtplyr", "data.table")
+
+
+
+# UHOH --------------------------------------------------------------------
+# *** ./analysis/common_genotypes.R ***
+# Number of Dent and Flint parental inbred lines as well as hybrids per 
+# predictor.
+common_genotypes <- "./data/processed/common_genotypes.RDS" %>%
+  readRDS()
+common_genotypes %>%
+  group_by(Pool, Data_Type) %>%
+  count()
+
+# *** ./analysis/uhoh_data_preparation.R ***
+# Number of transcripts."
+"./data/derived/uhoh/mrna.RDS" %>%
+  readRDS() %>%
+  ncol()
+
+# Number of pedigree records.
+"./data/derived/uhoh/pedigree_matrix.RDS" %>%
+  readRDS() %>%
+  ncol()
+
+# SNP data
+# *** ./analysis/snp_preparation.R ***
+## Number of SNP marker before applying quality checks.
+"./data/raw/marker_sample_matrix_FWD_P235-update.txt" %>%
+  data.table::fread() %>%
+  nrow()
+
+# *** ./analysis/uhoh_data_preparation.R ***
+## MAF: 5%, Call rate: 95%, Heterozygosity rate: 5%
+uhoh_imp_snp <- "./data/derived/uhoh/snp_matrix.RDS" %>%
+  readRDS()
+
+# Number of SNPs for all genotypes.
+uhoh_imp_snp %>%
+  ncol()
+
+# *** ./analysis/prediction.R ***
+# 1) Number of SNPs for parental inbred lines of all 1,521 hybrids.
+# 2) Number of SNPs for parental inbred lines covered by mRNA data.
+common_genotypes %>%
+  filter(Data_Type %in% c("snp", "mrna")) %>%
+  split(.$Pool, .$Data_Type) %>%
+  map(., ~split(., .$Data_Type)) %>%
+  at_depth(.depth = 2, "G") %>%
+  at_depth(.depth = 2, function(x) {
+    uhoh_imp_snp[rownames(uhoh_imp_snp) %in% x, ]
+  }) %>%
[…]

### 2017-03-14 reports/methods.Rmd
Remove analyses from methods

--- 2016-10-21 reports/data_overview.Rmd
@@
-# Data Overview
-```{r message = FALSE, echo = FALSE}
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
-               "viridis", "tidyr", "knitr", "tibble", "purrr")
+# Materials and Methods
 
-knit_hooks$set(inline = function(x) {
-  prettyNum(x, big.mark = ",", digits = 2)
-})
-knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
-knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
-                      warning = FALSE, echo = FALSE)
-```
+## Data
+<!--
+./reports/methods_data.R
+-->
 
-```{r Load-Agro-Data, message = FALSE, cache = TRUE}
-pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
-genos <- readRDS("./data/processed/common_genotypes.RDS")
+### Agronomic dat
+The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
+mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
+Evaluations of six agronomically important traits were recorded on all hybrids
+in field experiments at three or more agro-ecologically diverse locations 
+across Germany.
 
-full_hybrid <- genos %>%
-  filter(Pool == "Hybrid") %>%
-  .$G
-all_pheno <- pheno %>%
-  rename(Observed = EST) %>%
-  as_data_frame() %>%
-  filter(G %in% full_hybrid) %>%
-  dplyr::select(G, Observed, Trait) %>%
-  filter(Trait != "ADL") %>%
-  mutate(Trait = fct_recode(Trait, 
-    "FAT" = "FETT",
-    "DMY" = "GTM",
-    "DMC" = "GTS",
-    "PRO" = "RPR",
-    "SUG" = "XZ"),
-         Trait = fct_relevel(Trait, 
-    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
-    ))
-traits <- all_pheno %>%
-  .$Trait %>%
-  levels()
+The maize inbred line data set originally comprised 513 maize lines representing 
+the global maize diversity and was reduced to the set of tropical and
+subtropical lines ($n = 211$) given that it is the largest of the four
+pre-classified subgroups [@Yang2014].
+All inbred lines were evaluated in five Chinese environments ranging from
+$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
+[@Yang20114].
+Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
+we will evaluate six in our prediction models.
 
-# Hybrids of parents with records for both endophenotypes
-red_inbreds <- genos %>% 
-  filter(Pool != "Hybrid", Data_Type %in% c("snp", "mrna")) %>%
-  split(.$Pool) %>%
-  map(~split(., .$Data_Type)) %>%
-  at_depth(.depth = 2, "G") %>%
-  map(function(x) intersect(x$mrna, x$snp))
+#### Predictor data
+All parental inbred lines of the maize hybrids were genotyped using the Illumina 
+SNP BeadChip MaizeSNP50 [@Ganal2011].
 
-red_hybrid <- full_hybrid %>%
-  as_data_frame() %>%
-  separate(value, into = c("DF", "Dent", "Flint")) %>%
-  mutate(Avail_Dent = ifelse(Dent %in% red_inbreds$Dent, yes = 1, no = 0),
-         Avail_Flint = ifelse(Flint %in% red_inbreds$Flint, yes = 1, no = 0),
-         Tested_Parents = Avail_Dent + Avail_Flint) %>%
-  filter(Tested_Parents == 2) %>%
-  unite(Hybrid, DF, Dent, Flint) %>%
-  .$Hybrid
-genos <- genos %>%
-  rbind(tibble(G = red_hybrid,
-               Pool = "Hybrid",
-               Data_Type = "mrna")
-        )
-```
+MaizeSNP50 BeadChip-data were available for the entire set of lines from the
+diversity panel.
+Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
+days after sowing.
+By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
+overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
+for these lines [@Fu2013,Li2013].
+For the remaining 145 maize lines without RNA-seq data, high density markers
+were inferred via projection of IBD regions onto the BeadChip, using this core 
+set of markers [@Yang2014].
+Gene expression data were normalized using a normal quantile transformation to
+satisfy modelling assumptions [@Fu2013].
 
-```{r Feature-Number, echo = FALSE}
-snp <- readRDS("./data/processed/snp_mat.RDS")
-mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
-mrna <- t(mrna)
-mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
-```
+After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
+($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
+using the Beagle software [@Browning2009].
+The final number of polymorphic marker loci was 7,013 for parental Dent and
+6,212 for parental Flint lines, respectively.
 
-Agronomic data for `r length(traits)` traits (including dry-matter yield (DMY))
-and `r length(full_hybrid)` hybrids are available 
-(Table \@ref(tab:Geno-Number)).
-After applying a series of quality checks, `r ncol(snp)` SNPs and 
-best linear unbiased estimates of `r ncol(mrna)` were available for further 
-analyses.
-Genomic, pedigree and transcriptomic records, were collected for inbred parent 
-lines of `r length(red_hybrid)` hybrids.
+### Pedigree Data (Hybrid)
 
-```{r Geno-Number, echo = FALSE}
-geno_tab_cap <- paste("Number of genotypes for Dent and Flint parent lines and",
-                      "their hybrid progeny for which genomic and/or",
-                      "transcriptomic data are available.",
-                      "All genotypes for which transcriptomic data are", 
-                      "available, also have genomic data.",
-                      "Agronomic data are available for all 1,521 hybrid",
-                      "genotypes.")
-# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
-# parental Dent and Flint lines, respectively.
-genos %>% 
-  mutate(Data_Type = fct_recode(Data_Type,
-                                Agronomic = "agro",
-                                Transcriptomic = "mrna",
-                                Genomic = "snp",
-                                Pedigree = "ped")) %>%
-  group_by(Pool, Data_Type) %>%
-  count() %>%
-  rename(`Data Type` = Data_Type) %>%
-  knitr::kable(caption = geno_tab_cap, booktabs = TRUE)
-```
+## Structure
+
+### PCA
+
+### Population Structure Analysis
+
+## Core Sampling
+
+## Prediction
+
+### Kernels
+
+### Imputation
+
+### Models
+
+### Leave-one-out cross-validation (LOOCV)

### 2017-03-14 reports/methods.Rmd
Describe UHOH predictor data

@@
 
-### Agronomic dat
+### Hybrid data
 The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
 mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
-Evaluations of six agronomically important traits were recorded on all hybrids
-in field experiments at three or more agro-ecologically diverse locations 
-across Germany.
+Best linear unbiased estimates (BLUEs) were computed across all factorials and 
+three or more agro-ecologically diverse locations across Germany for six 
+agronomically important traits.
+All 245 parent lines have pedigree records back to the generation of their 
+grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
+SNP BeadChip MaizeSNP50 [@Ganal2011].
+After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
+($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
+using the Beagle software [@Browning2009].
+The final number of polymorphic marker loci was 7,013 for parental Dent and
+6,212 for parental Flint lines, respectively.
+Gene expression data for seedlings seven days after sowing were obtained for a 
+subset of 60 parental Dent and 43 parental Flint lines through two-color
+hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
+After applying established normalization procedures [@Smyth2003,Ritchie2007], 
+gene expression BLUEs were computed for 1,323 transcripts in reference to 
+established protocols [@Symth2003,Ritchie2007,Frisch2010] using the R-package 
+*limma* [@Ritchie2015].
 
+#### Diversity panel
 The maize inbred line data set originally comprised 513 maize lines representing 
@@
 we will evaluate six in our prediction models.
-
-#### Predictor data
-All parental inbred lines of the maize hybrids were genotyped using the Illumina 
-SNP BeadChip MaizeSNP50 [@Ganal2011].
-
 MaizeSNP50 BeadChip-data were available for the entire set of lines from the
@@
 set of markers [@Yang2014].
+The same SNP quality checks as for the parental inbred lines of the hybrids
+were applied to the 211 inbred lines from the tropical/subtropical subset,
+yielding 37,760 SNPs.
 Gene expression data were normalized using a normal quantile transformation to
 satisfy modelling assumptions [@Fu2013].
-
-After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
-($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
-using the Beagle software [@Browning2009].
-The final number of polymorphic marker loci was 7,013 for parental Dent and
-6,212 for parental Flint lines, respectively.
+A subset of 28,850 annotated genes was kept for further analyses [@Li2013].
 

### 2017-03-14 reports/methods_data.R

@@ -107,7 +107,7 @@ maizego_imp_snp %>%
 # *** ./analysis/prediction.R ***
 # 1) Number of SNPs for inbred lines covered by mRNA data.
 # 2) Number of SNPs for all inbred lines.
-common_genotypes %>%
+maizego_genotypes %>%
   discard(names(.) == "pheno") %>%
   map(function(x) {
     maizego_imp_snp[rownames(maizego_imp_snp) %in% x, ]

### 2017-03-14 reports/methods.Rmd
Methods: Core sampling, PCA, Admixture

@@
 satisfy modelling assumptions [@Fu2013].
-A subset of 28,850 annotated genes was kept for further analyses [@Li2013].
+A subset of 28,850 annotated genes was available for 149 out of the 211
+genotypes and was kept for further analyses [@Li2013].
 
-### Pedigree Data (Hybrid)
-
-## Structure
-
-### PCA
-
-### Population Structure Analysis
+## Population structure
+Principal component analyses (PCA) were run on scaled and centered predictor 
+matrices using the *pca()* function from the R-package *LEA* [@Frichot2015].
+Estimates of ancestry coefficients were estimated based on parse nonnegative
+matrix factorization algorithms [@Frichot2014] using the *snmf()* function 
+in *LEA*.
+Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
+repetitions for each value of $K$. 
+For each $K$, only the repetition with the lowest cross-entropy was kept for 
+further analyses.
 
 ## Core Sampling
+In order to evaluate the influence of the genetic constitution of the set of
+genotypes that has only data on one out of two predictors, we generated core 
+samples using the R-package *corehunter* [@Thachuk2009].
+These core samples were based on the 149 maize inbred lines from the diversity
+panel which had gene expression as well as genomic information.
+The size of the core set was varied in increments of ten percentage points and 
+ranged from 10% to 90% of all 149 inbred lines.
+For assembling the various core sets, we first computed the Modifed Rogers (MR)
+distance between each pair of genotypes and then aimed at maximizing the
+average distance between each member of the selected core.
+MR was computed using the *rogers.dist()*-function from the R-pacakge *poppr*
+[@Kamvar2015].
 

### 2017-03-14 reports/methods.Rmd
Methods: Kernels and LOOCV

@@
 ### Kernels
+BLUP models are computationally attractive when the number of features $p$
+exceeds the number of genotypes $n$ and are equivalent to a selection index
+when fixed effect have been accomodated by the dependent variable
+(@Mrode2014, pp. 34, 311, 312).
+Depending on the data set, up to three predictors were available for agronomic
+trait predictions, namely pedigree data (P), genomic data (G) and
+transcriptompic data (T).
+The corresponding feature matrix of the $l$-th group of inbred lines and the 
+$o$-th predictor ($\mathbf{W_{lo}}$) has dimensions $n \times p$ where $n$
+pertains to the number of gneotypes in the $l$-th group and $p$ pertains to 
+the number of features.
+For the diversity panel maize lines, $n$ corresponds to the number of genotypes
+whereas, in the case of the hybrid data, $n$ corresponds to the number of
+parent lines in the corresponding heterotic group.
+In generating kernels for each predictor, all features in $\mathbf{W_{lo}}$ were
+centered and standardized to unit variance.
+Then, each kernel can be defined as
+
+$$
+\begin{equation} \label{eq:GenomicRelationship}
+  \mathbf{G}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
+\end{equation}
+$$
+where $W$ denotes the number of features [@VanRaden2008].
+In the case of pedigree data (P), coancestry coefficients were used directly
+for $\mathbf{G}_{lo}$.
+
+### Models
 
@@
 
-### Models
+### Leave-one-out cross-validation
+In the case of the diversity panel maize inbred lines, leave-one-out cross
+validation (LOOCV) was done by simply using a single genotype as a hold-out
+sample, which will subsequently be predicted, while using all other inbred 
+lines for model training.
+This process is repeated until all inbred lines have been used once for testing
+and $n - 1$ times for model training.
 
-### Leave-one-out cross-validation (LOOCV)
+For the hybrid data, LOOCV was carried out as follows:
+Let $D$ and $F$ denote the set of parental inbred lines from the Dent and the 
+Flint group, respectively.
+Further, let $H \cap [D \times F]$ denote the set of hybrids from crosses 
+between $D$ and $F$.
+The training set ($H_{TRN}$) for the hybrid to be predicted ($H_{ij}$) --- 
+where $i \in D_{TRN}$ and $j \in F_{TRN}$ --- was assembled as 
+$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$, where 
+$D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.

### 2017-03-14 reports/index.Rmd
Methods: Init (single step) prediction models

@@
 ---
-title: Single-step Hybrid Predition
+title: Omics-based Single-step Trait Predition
 author:
@@
     base_format: rticles::elsevier_article
-bibliography: ["library.bib", "cheuer.bib"]
+bibliography: ["~/Dropbox/Paper/library.bib"]
 abstract: |

### 2017-03-14 reports/intro.Rmd
Methods: Init (single step) prediction models

@@
 Individuals with accurate breeding values out of the first step (i.e. high 
-number of offspring), enter the second step in which SNP genotypes are regressed 
-against the conventional breeding values.
+number of offspring), enter the second step in which conventional breeding 
+values are regressed onto SNP genotypes.
 Those markers can then be used to predict genomic breeding values for solely 
@@
 transcriptomic information but all have SNP genotypes.
-We can take a similar route as in @fernando_class_2014 and impute missing
+We can take a similar route as in @Fernando2014 and impute missing
 mRNA data.
@@
 The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
-[@legarra_relationship_2009].
+[@Legarra2009].
 
@@
 
-Following the notation in @fernando_class_2014 the phenotypes can be untangled like this:
+Following the notation in @Fernando2014 the phenotypes can be untangled like this:
 

### 2017-03-14 reports/methods.Rmd
Methods: Init (single step) prediction models

@@
 SNP BeadChip MaizeSNP50 [@Ganal2011].
-After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
-($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
+After filtering for minor allele frequency ($\geq 5$%), heterozygosity rate
+($\geq 5$%) and call frequency ($\geq 95$%), missing genotypes were imputed 
 using the Beagle software [@Browning2009].
@@
 hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
-After applying established normalization procedures [@Smyth2003,Ritchie2007], 
+After applying established normalization procedures [@Smyth2003;@Ritchie2007], 
 gene expression BLUEs were computed for 1,323 transcripts in reference to 
-established protocols [@Symth2003,Ritchie2007,Frisch2010] using the R-package 
-*limma* [@Ritchie2015].
+established protocols [@Smyth2003;@Ritchie2007;@Frisch2010] using the R-package 
+*limma* [@Ritchie2015a].
 
@@
 $18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
-[@Yang20114].
+[@Yang2014].
 Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
@@
 overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
-for these lines [@Fu2013,Li2013].
+for these lines [@Fu2013;@Li2013].
 For the remaining 145 maize lines without RNA-seq data, high density markers
@@
 $o$-th predictor ($\mathbf{W_{lo}}$) has dimensions $n \times p$ where $n$
-pertains to the number of gneotypes in the $l$-th group and $p$ pertains to 
+pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
 the number of features.
@@
 
-$$
 \begin{equation} \label{eq:GenomicRelationship}
@@
 \end{equation}
-$$
+
 where $W$ denotes the number of features [@VanRaden2008].
@@
 ### Models
+The universal model for breeding values was:
 
-### Imputation
+\begin{equation} \label{eq:GBLUPModel}
+  \mathbf{y} = \mu + 
+  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
+  \mathbf{\epsilon},
+\end{equation}
 
-### Leave-one-out cross-validation
-In the case of the diversity panel maize inbred lines, leave-one-out cross
-validation (LOOCV) was done by simply using a single genotype as a hold-out
-sample, which will subsequently be predicted, while using all other inbred 
-lines for model training.
+where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
+respectively, $\mu$ is the fixed model intercept, $\mathbf{Z}_{lo}$ is the 
+corresponding design matrix associating the random genotype effects (*i.e.*, 
+breeding values) of the lines in the $l$-th inbred line group and the $o$-th 
+predictor type with $\mathbf{y}$.
+Note that, in the case of hybrid data, $\mathbf{Z}_{lo}$ associates the random
+general combining ability effects in the parental inbred lines (either Dent or
+Flint, as expressed by $l$) with the vector of observed hybrid performance
+$\mathbf{y}$.
+The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
+to $\mathbf{G}_{lo} \sigma^{2}_{{u}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
+for the residual error.
+All prediction models were implemented using the R-package *BGLR* [@Perez2014].
+
+### Single Step Prediction
+
+The breeding values in a prediction model are in general given by:
+\begin{equation} \label{eq:mrnaebv}
+	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
+\end{equation}
+
+where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model,
+which is equivalent to the BLUP model.
+
+Consider now the situation in which one predictor is complete in the sense that
+it contains information on the whole population but a second predictor is
+incomplete and covers only a subset of the population.
+We can take a similar route as in @Fernando2014 and impute covariates of the 
+incomplete predictor by using covariates of the complete predictor.
+Let the subscript $1$ denote individuals covered only by the complete
+predictor whereas individuals covered by both, the complete and the
+incomplete predictor, are indicated by the subscript $2$.
+Furhter, the covariates in $\mathbf{W}$ are centered.
+The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
+expectation given $\mathbf{g}_2$ and a residual error term:
+\begin{align} \label{eq:mrna1}
+	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
+	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
+	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
+\end{align}
+
+The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
+[@Legarra2009].
+
+The structure of the residual imputation/prediction error is known and can 
+therefore be modelled.
+The covariates not covered by the incomplete predictor can be predicted using 
+the expectation of a multivariate normal random vector given correlated 
+observations, which are related to the Best Linear Predictor.
+
+The general model for the single step procedure can then be written as
+
+\begin{equation}
+  (\#eq:single-step-model)
+\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Full} \boldsymbol{\epsilon}_{Full} + \mathbf{e},
+\end{equation}
+
+with
+
+\begin{equation}
+ (\#eq:single-step-submatrices)
+\mathbf{X} = 
+\begin{bmatrix}
+  X_1 \\
+  X_2 
+ \end{bmatrix},
+ \mathbf{W} = 
+\begin{bmatrix}
+  Z_1\hat{\mathbf{W}_1} \\
+  \mathbf{W}_2 
+ \end{bmatrix},
+ \mathbf{U}_{Full} = 
+\begin{bmatrix}
+  Z_1 \\
+  0 
+ \end{bmatrix}
+\end{equation}
+
+Following the notation in @Fernando2014 the phenotypes can be untangled like this:
+
+\begin{align}
+ (\#eq:entangled-augmented-single-step-model)
+\begin{bmatrix}
+  y_1 \\
+  y_2 
+ \end{bmatrix}
+& =
+ \begin{bmatrix}
+  X_1 \\
+  X_2 
+ \end{bmatrix}
+ \boldsymbol{\beta} + 
+ \begin{bmatrix}
+  Z_1 & 0 \\
+  0 & Z_2 
+ \end{bmatrix}
+\begin{bmatrix}
+  g_1 \\
+  g_2 
+ \end{bmatrix}
+  + \mathbf{e} \\
+    & = 
+ \begin{bmatrix}
+  X_1 \\
+  X_2 
+ \end{bmatrix}
+ \boldsymbol{\beta} + 
+ \begin{bmatrix}
+  Z_1 & 0 \\
+  0 & Z_2 
+ \end{bmatrix}
+\begin{bmatrix}
+  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full}  \\
+  \mathbf{W}_2\boldsymbol{\alpha} \\
+ \end{bmatrix}
+  + \mathbf{e} \\
+    & = 
+ \begin{bmatrix}
+  X_1 \\
+  X_2 
+ \end{bmatrix}
+ \boldsymbol{\beta} + 
+ \begin{bmatrix}
+  Z_1 & 0 \\
+  0 & Z_2 
+ \end{bmatrix}
+\begin{bmatrix}
+  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full} \\
+  \mathbf{W}_2\boldsymbol{\alpha} \\
+ \end{bmatrix}
+  + \mathbf{e} \\
+\end{align}
+
+The breeding values from that model are:
+
+\begin{equation}
+ (\#eq:breeding-values)
+\tilde{\mathbf{g}} = 
+ \begin{bmatrix}
+  \hat{\mathbf{W}_1} \\
+  \mathbf{W}_2 
+ \end{bmatrix}
+ \hat{\boldsymbol{\alpha}}
+ + 
+ \begin{bmatrix}
+  \mathbf{Z}_1 \\
+  0
+ \end{bmatrix}
+ \hat{\boldsymbol{\epsilon}}_{Full} 
+\end{equation}
+
+Our final model looks like this:
+
+\begin{align}
+ (\#eq:final-model)
+\mathbf{y} &= \mathbf{Xb} +
+\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
+\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l{Full}} \boldsymbol{\epsilon}_{l{Full}} +
+\mathbf{e}.
+\end{align}
+
+For the hybrid data, the obtained GCA effects for the inbred lines represent 
+half their breeding values.
+The predicted agronomic performance is:
+
+\begin{equation}
+ (\#eq:predicted-performance)
+\tilde{\mathbf{h}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
+\end{equation}
+
+### Model validation
+For the validation of our predictions, we employed leave-one-out
+cross-validation (LOOCV) routines.
+In the case of the diversity panel maize inbred lines, LOOCV was done by simply 
+using a single genotype as a hold-out sample, which will subsequently be 
+predicted, while using all other inbred lines for model training.
 This process is repeated until all inbred lines have been used once for testing
@@
 where $i \in D_{TRN}$ and $j \in F_{TRN}$ --- was assembled as 
-$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$, where 
+$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$ with
 $D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.

### 2017-03-15 reports/index.Rmd
Fix typos in Methods

@@
 abstract: |
-  This is the abstract.
-
-  It consists of two paragraphs.
+  $\dots$
 ---

### 2017-03-15 reports/methods.Rmd
Fix typos in Methods

@@
 grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
-SNP BeadChip MaizeSNP50 [@Ganal2011].
+BeadChip MaizeSNP50 [@Ganal2011].
 After filtering for minor allele frequency ($\geq 5$%), heterozygosity rate
-($\geq 5$%) and call frequency ($\geq 95$%), missing genotypes were imputed 
+($\leq 5$%) and call frequency ($\geq 95$%), missing genotypes were imputed 
 using the Beagle software [@Browning2009].
@@
 
-#### Diversity panel
+### Diversity panel
 The maize inbred line data set originally comprised 513 maize lines representing 
@@
 All inbred lines were evaluated in five Chinese environments ranging from
-$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
-[@Yang2014].
+$18$ to $30^{\circ}$N and from $102$ to $110^{\circ}$E [@Yang2014].
 Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
-we will evaluate six in our prediction models.
+six were used for this study.
 MaizeSNP50 BeadChip-data were available for the entire set of lines from the
@@
 matrices using the *pca()* function from the R-package *LEA* [@Frichot2015].
-Estimates of ancestry coefficients were estimated based on parse nonnegative
+Estimates of ancestry coefficients were estimated based on sparse nonnegative
 matrix factorization algorithms [@Frichot2014] using the *snmf()* function 
@@
 For assembling the various core sets, we first computed the Modifed Rogers (MR)
-distance between each pair of genotypes and then aimed at maximizing the
-average distance between each member of the selected core.
-MR was computed using the *rogers.dist()*-function from the R-pacakge *poppr*
+distance between each pair of genotypes and then selected a predefined subset
+of genotypes that display maximum average distance between each other.
+MR was computed using the *rogers.dist()* function from the R-pacakge *poppr*
 [@Kamvar2015].
@@
 exceeds the number of genotypes $n$ and are equivalent to a selection index
-when fixed effect have been accomodated by the dependent variable
-(@Mrode2014, pp. 34, 311, 312).
+when fixed effects have been accomodated by the dependent variable [@Mrode2014].
 Depending on the data set, up to three predictors were available for agronomic
@@
 The corresponding feature matrix of the $l$-th group of inbred lines and the 
-$o$-th predictor ($\mathbf{W_{lo}}$) has dimensions $n \times p$ where $n$
+$o$-th predictor ($\mathbf{W}_{lo}$) has dimensions $n \times p$ where $n$
 pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
@@
 parent lines in the corresponding heterotic group.
-In generating kernels for each predictor, all features in $\mathbf{W_{lo}}$ were
+In generating kernels for each predictor, all features in $\mathbf{W}_{lo}$ were
 centered and standardized to unit variance.
@@
 
-\begin{equation} \label{eq:GenomicRelationship}
-  \mathbf{G}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
+\begin{equation} (\#eq:GenomicRelationship)
+  \mathbf{K}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
 \end{equation}
 
-where $W$ denotes the number of features [@VanRaden2008].
+where $W_{lo}$ denotes the number of features [@VanRaden2008].
 In the case of pedigree data (P), coancestry coefficients were used directly
-for $\mathbf{G}_{lo}$.
+for $\mathbf{K}_{lo}$.
 
@@
 
-\begin{equation} \label{eq:GBLUPModel}
+\begin{equation} (\#eq:KBLUPModel)
   \mathbf{y} = \mu + 
@@
 The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
-to $\mathbf{G}_{lo} \sigma^{2}_{{u}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
+to $\mathbf{K}_{lo} \sigma^{2}_{{g}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
 for the residual error.
@@
 The breeding values in a prediction model are in general given by:
-\begin{equation} \label{eq:mrnaebv}
+\begin{equation} (\#eq:mrnaebv)
 	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
@@
 Consider now the situation in which one predictor is complete in the sense that
-it contains information on the whole population but a second predictor is
+it contains information on the whole population whereas a second predictor is
 incomplete and covers only a subset of the population.
@@
 incomplete predictor, are indicated by the subscript $2$.
-Furhter, the covariates in $\mathbf{W}$ are centered.
+Further, the covariates in $\mathbf{W}$ are centered.
 The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
 expectation given $\mathbf{g}_2$ and a residual error term:
-\begin{align} \label{eq:mrna1}
+\begin{align} (\#eq:mrna1)
 	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
-	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
+	&= \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
 	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
@@
 
-The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
+The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{K}_{11} - \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{K}_{21} = (\mathbf{K}^{11})^{-1}$
 [@Legarra2009].
@@
   (\#eq:single-step-model)
-\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Full} \boldsymbol{\epsilon}_{Full} + \mathbf{e},
+\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}^{Complete} \boldsymbol{\epsilon}^{Complete} + \mathbf{e},
 \end{equation}
@@
  \end{bmatrix},
- \mathbf{U}_{Full} = 
+ \mathbf{U}^{Complete} = 
 \begin{bmatrix}
@@
 \begin{bmatrix}
-  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full}  \\
+  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete}  \\
   \mathbf{W}_2\boldsymbol{\alpha} \\
@@
 \begin{bmatrix}
-  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full} \\
+  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete} \\
   \mathbf{W}_2\boldsymbol{\alpha} \\
@@
  \end{bmatrix}
- \hat{\boldsymbol{\epsilon}}_{Full} 
+ \hat{\boldsymbol{\epsilon}}^{Complete} 
 \end{equation}
@@
 \sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
-\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l{Full}} \boldsymbol{\epsilon}_{l{Full}} +
+\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l}^{Complete} \boldsymbol{\epsilon}_{l}^{Complete} +
 \mathbf{e}.
@@
 cross-validation (LOOCV) routines.
-In the case of the diversity panel maize inbred lines, LOOCV was done by simply 
+In the case of the diversity panel maize inbred lines, LOOCV was performed by 
 using a single genotype as a hold-out sample, which will subsequently be 

### 2017-03-15 reports/index.Rmd
Methods: Describe CV and predictive ability computation

@@
 ---
-title: Omics-based Single-step Trait Predition
+title: Omics-based Single Step Trait Prediction
 author:

### 2017-03-15 reports/methods.Rmd
Methods: Describe CV and predictive ability computation

@@
  (\#eq:predicted-performance)
-\tilde{\mathbf{h}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
+\hat{\mathbf{y}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
 \end{equation}
 
-### Model validation
+### Predictive ability and model validation
 For the validation of our predictions, we employed leave-one-out
@@
 $D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.
+
+We judged the performance of each model by looking at its predictive ability,
+which is calculated as $\rho(\mathbf{y}, \mathbf{\hat{y}})$, where 
+$\mathbf{\hat{y}}$ is the vector of predicted values from each LOOCV run.
+Our confidence in the performance of each model was evalutaed by computing the
+coefficient of variation for the predicted values:
+
+\begin{equation}
+  (\#eq:CV)
+CV = \sqrt{\frac{\sum_{i = 1}^{n} \sigma^{2}_{\hat{y}(i)} / n}{\sum_{i=1}^{n}\hat{y}_{i} / n}},
+\end{equation}
+
+where $n$ denotes the number of genotypes (*i.e.*, inbred lines or hybrids).

### 2017-03-15 analysis/ensure_snp_quality.R
Delete superfluous files from old analyses

@@ -1,207 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-if (!require("devtools")) install.packages("devtools")
-devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("tidyverse", "data.table", "dtplyr", "viridis", "stringr",
-               "corehunter")
-pacman::p_load_gh("mwesthues/sspredr")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("LEA")
-library("LEA")
-
-
-# Get labels for the different subpopulations of the genotypes.
-pop_struc <- fread("./data/input/maizego/genotype_annotation.txt")
-
-# Keep only genotypes from the tropical/subtropical (TST) subset to avoid 
-# hassles with extreme population structure.
-tst_genotypes <- pop_struc %>%
-  filter(Subpopulations == "TST") %>%
-  select(Lines) %>%
-  flatten_chr()
-
-# Load the phenotypic BLUPs.
-pheno_path <- "./data/input/maizego/blup_traits_final.csv"
-pheno <- pheno_path %>%
-  read_csv() %>%
-  rename(Genotype = `<Trait>`) %>%
-  gather(key = Trait, value = Value, -Genotype) %>%
-  filter(Genotype %in% tst_genotypes)
-
-
-# Load the gene expression data.
-mrna_path <-  "./data/input/maizego/Expression_kernel_finalNormalized_28850.txt"
-mrna <- mrna_path %>%
-  fread(header = TRUE) %>%
-  gather(key = Genotype, value = Expression, -Gene_ID) %>%
-  as.data.table() %>%
-  filter(Genotype %in% tst_genotypes)
-
-
-
-# Explore the kinship matrix.
-snp_path <- "./data/input/maizego/513allchr-imputation.txt"
-snp <- snp_path %>%
-  fread(header = TRUE) %>%
-  as.data.table()
-
-
-# Collect meta information on the genomic data such as the marker name, the 
-# chromosome and the position of the marker on the chromosome.
-snp_meta_vars <- c("rs#", "alleles", "chrom", "pos", "strand", "assembly#",
-                   "center", "protLSID", "assayLSID", "paneLSID", "QCcode")
-snp_selection_vars <- c(tst_genotypes, snp_meta_vars)
-snp_meta_info <- snp %>%
-  select(one_of(snp_meta_vars))
-
-# Keep 
-snp_mat <- snp %>%
-  select(-one_of(snp_meta_vars)) %>%
[…]

### 2017-03-15 analysis/impute_single_mrna.R

@@ -1,159 +1,0 @@
-# Genotype-Wise mRNA-Imputation
-pacman::p_load("sspredr", "tidyverse", "forcats", "purrr", "methods", 
-               "parallel")
-use_cores <- 3L
-
-# Common genotypes
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-dent <- genos %>% 
-  filter(Pool == "Dent", Data_Type == "mrna") %>%
-  .$G
-flint <- genos %>%
-  filter(Pool == "Flint", Data_Type == "mrna") %>%
-  .$G
-
-# Remove the 'Group' variable after splitting Dent and Flint data.
-rm_grp <- function(x) {
-  x$Group <- NULL
-  x
-}
-
-# mRNA
-mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
-mrna <- t(mrna)
-mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
-mrna_lst <- mrna %>%
-  as_data_frame() %>%
-  mutate(G = rownames(mrna)) %>%
-  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
-         Group = as.factor(Group)) %>%
-  as.data.frame() %>%
-  column_to_rownames(var = "G") %>%
-  split(.$Group) %>%
-  map(rm_grp) %>%
-  map(~as.matrix(.))
-
-# SNPs
-snp <- readRDS("./data/processed/snp_mat.RDS")
-snp_lst <- snp %>%
-  as_data_frame() %>%
-  mutate(G = rownames(snp)) %>%
-  filter(G %in% c(dent, flint)) %>%
-  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
-         Group = as.factor(Group)) %>%
-  as.data.frame() %>%
-  column_to_rownames(var = "G") %>%
-  split(.$Group) %>%
-  map(rm_grp) %>%
-  map(~as.matrix(.))
-
-# Pedigree
-ped <- readRDS("./data/processed/ped-datafull-GTP.RDS")
-ped_lst <- ped %>%
-  mutate(G = rownames(ped)) %>%
-  filter(G %in% c(dent, flint)) %>%
-  mutate(Group = ifelse(G %in% dent, yes = "Dent", no = "Flint"),
-         Group = as.factor(Group)) %>%
-  as.data.frame() %>%
-  column_to_rownames(var = "G") %>%
[…]

### 2017-03-15 analysis/predictions_server_time.R

@@ -1,98 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load("data.table", "tidyverse", "dtplyr")
-log_file <- fread("./data/derived/pred_log.txt")
- 
-convert_runs <- function(x) {
-  x %>%
-  str_split(., pattern = "-") %>%
-  as_vector() %>%
-  as.integer() %>%
-  (function(y) seq(from = y[1], to = y[2])) %>%
-  length()
-}           
-
-compute_runtime <- function(cu, ca, ru, ra, hu, tol = 1) {
-  ### Input:
-  # cu: number of cores used for development run
-  # ca: number of cores used for production run
-  # ru: number of development runs
-  # ra: number of production runs
-  # hu: number of hours needed in development
-  # tol: multiplier for time tolerance (1 = no tolerance,
-  #                                     1.2 i.e. 20% tolerance)
-  ### Output:
-  # number of hours needed in production in DD:HH:MM:SS format
-  ha <- tol * (cu * ra * hu) / (ru * ca)
-  td <- seconds_to_period(ceiling(ha))
-  sprintf('%02d:%02d:%02d:%02d', day(td), td@hour, minute(td), second(td))
-}
-
-run_df <- log_file %>%
-  filter(grepl("2016-11-09|2016-11-08", Date),
-         Job_ID != "interactive_00",
-         Cores < 16,
-         Iter == 30000,
-         Job_ID != 10076266,
-         Job_ID != 10075993) %>%
-  mutate(Elapsed = seconds(hms(Elapsed_Time)),
-         Elapsed = str_replace(Elapsed, pattern = "S", replacement = ""),
-         Elapsed = as.numeric(Elapsed)) %>%
-  rowwise() %>%
-  mutate(Runs = convert_runs(Runs),
-         Needed = compute_runtime(cu = Cores,
-                                  ca = 16,
-                                  ru = Runs,
-                                  ra = 1521,
-                                  hu = Elapsed,
-                                  tol = 2))
-
-setup_moab_cmd <- function(pred1, pred2, pred3, trait, iter, model, varcov,
-                           pi_param, prior_pi_cnt, walltime, pmem, nodes, 
-                           cores) {
[…]

### 2017-03-15 analysis/predictor_subsetting.R

@@ -1,111 +1,0 @@
-# Goal: Generate nine subsets of predictors for later comparisons between their
-# predictive abilities:
-# a. 100% SNP
-# b. 100% pedigree
-# c. 42% mRNA
-# d. 42% pedigree
-# e. 42% SNP
-# f. 100% pedigree & 77% SNP
-# g. 100% pedigree & 42% mRNA
-# h. 100% SNP & 42% mRNA
-# i. 100% pedigree & 77% SNP & 42% mRNA
-#
-# Make comparisons between the following scenarios:
-# a, b, f, g, h, i
-# c, d, e
-#
-# Throughout this script, digits behind the variable name of each predictor
-# (ped: pedigree, snp: genomic, mrna: transcriptomic) indicate the size of 
-# the subset relative to the full set of genotypes (n = 245). For example,
-# we have transcriptomic data for 103 genotypes, which are roughly 42% of 
-# 245 genotypes, hence the name 'mrna42' and so forth.
-
-pacman::p_load("tidyverse", "data.table")
-devtools::install_github("mwesthues/sspredr")
-
-
-
-# -- Genotype information.
-# Information on the set of genotypes.
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-genos %>%
-  filter(Pool %in% c("Dent", "Flint")) %>%
-  group_by(Pool, Data_Type) %>%
-  count()
-
-# Get the names of the Dent Hybrid-parents.
-geno <- genos %>%
-  filter(Pool == "Hybrid") %>%
-  .$G %>%
-  map(~stringr::str_split(., pattern = "_")) %>%
-  at_depth(.depth = 2, .f = 2) %>%
-  flatten() %>%
-  as_vector()
-
-# -- Load predictor data.
-# pedigree
-ped100 <- readRDS("./data/processed/ped-datafull-GTP.RDS")
-ped_nms <- genos %>%
-  filter(Data_Type == "ped") %>%
-  .$G
-ped100 <- as.matrix(ped100) * 2
-ped100 <- ped100 %>%
-  .[rownames(.) %in% ped_nms, colnames(.) %in% ped_nms]
-ped100[is.na(ped100)] <- 0
-
-# snp
-snp100 <- readRDS("./data/processed/snp_mat.RDS")
-snp_nms <- genos %>%
-  filter(Data_Type == "snp") %>%
-  .$G
[…]

### 2017-03-15 analysis/repeated_snp77_prediction.R

@@ -1,357 +1,0 @@
-# Goal: Generate nine subsets of predictors for later comparisons between their
-# predictive abilities:
-
-if (!require("pacman")) install.packages("pacman")
-if (!require("devtools")) install.packages("devtools")
-pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
-               "purrr", "testthat", "methods", "tibble", "ggplot2", "stringr",
-               "stringi", "lubridate", "readr")
-devtools::install_github("mwesthues/sspredr")
-pacman::p_load_gh("mwesthues/sspredr")
-
-# Load function to save session info on the software used for the analysis in
-# this script.
-source("./software/session_info.R")
-
-
-
-## --------------------------------------------------------------------------
-## COLLECT AND CHECK ARGUMENTS 
-if (isTRUE(interactive())) {
-  # Number of cores
-  Sys.setenv("MOAB_PROCCOUNT" = "3")
-  Sys.setenv("TRAIT" = "")
-  # Number of iterations in BGLR()
-  Sys.setenv("ITER" = "30000")
-  # Main predictor. If 'Pred2' and 'Pred3' are empty, no imputation will take
-  # place.
-  Sys.setenv("PRED1" = "ped100")
-  # If 'Pred3' is empty, 'Pred2' will be imputed via information from 'Pred1'.
-  Sys.setenv("PRED2" = "snp77")
-  # If not empty, this predictor will be imputed.
-  Sys.setenv("PRED3" = "mrna42")
-  # Which sampling fold shall be used for snp77 (1 to 20)?
-  Sys.setenv("REPLICATION" = "2")
-  # Number of genotypes to predict (only for testing!)
-  Sys.setenv("RUNS" = "1-3")
-  # Output directory for temporary BGLR files
-  Sys.setenv("TMP" = "./tmp")
-}
-
-if (!interactive()) {
-  job_id <- Sys.getenv("MOAB_JOBID")
-} else job_id <- "interactive_00"
-
-# Save session info
-write_session_info(directory = "./data/derived/session_info/",
-                   job_id = job_id)
-
-use_cores <- as.integer(Sys.getenv("MOAB_PROCCOUNT"))
-init_traits <- as.character(Sys.getenv("TRAIT"))
-init_iter <- as.integer(Sys.getenv("ITER"))
-pred1 <- as.character(Sys.getenv("PRED1"))
[…]

### 2017-03-15 analysis/repeated_snp77_subsetting.R

@@ -1,94 +1,0 @@
-# Goal: Generate 20 subsets of genomic data to average out the influence of 
-# sampling on all prediction scenarios including 'snp77'. 
-
-if (!require("pacman")) install.packages("pacman")
-if (!require("devtools")) install.packages("devtools")
-pacman::p_load("BGLR","data.table", "parallel", "magrittr", "dplyr", "tidyr",
-               "purrr", "testthat", "methods", "tibble", "ggplot2", "stringr",
-               "stringi", "lubridate", "readr")
-devtools::install_github("mwesthues/sspredr")
-pacman::p_load_gh("mwesthues/sspredr")
-
-
-
-# -- GENOTYPE INFORMATION -----------------------------------------------
-# Information on the set of genotypes.
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-
-# -- LOAD PREDICTOR DATA ------------------------------------------------
-# pedigree
-cat_lst <- readRDS("./data/derived/pred_sub_list.RDS")
-hybrids <- cat_lst %>%
-  purrr::transpose() %>%
-  .[names(.) == "ped100_snp77"] %>%
-  at_depth(.depth = 2, .f = "geno") %>%
-  .[[1]]
-
-ped100 <- cat_lst %>% 
-  purrr::transpose() %>%
-  .[names(.) == "ped100_snp77"] %>%
-  at_depth(.depth = 2, .f = 1) %>%
-  .[[1]]
-names(ped100) <- c("Dent", "Flint")
-
-# snp
-snp100 <- readRDS("./data/processed/snp_mat.RDS")
-snp_nms <- genos %>%
-  filter(Data_Type == "snp") %>%
-  .$G
-snp100 <- snp100 %>%
-  .[rownames(.) %in% snp_nms, ]
-
-# mrna
-mrna42 <- readRDS("./data/derived/predictor_subsets/transformed-mrna42.RDS")
-
-
-
-# -- RESAMPLE PREDICTOR DATA -------------------------------------------- 
-# Remove some genomic records so that we can impute them via pedigree
-# information.
-mrna_genos <- genos %>%
-  filter(Data_Type == "mrna") %>%
-  split(.$Pool) %>%
-  map("G")
-
-set.seed(314)
-genos %>%
-  filter(Pool %in% c("Dent", "Flint"),
[…]

### 2017-03-15 analysis/scaled_mrna_predictions.R

@@ -1,213 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-if (!require("devtools")) install.packages("devtools")
-devtools::install_github("mwesthues/sspredr", update = TRUE)
-pacman::p_load("BGLR", "data.table", "tidyverse", "dtplyr", "parallel", 
-               "caret", "e1071", "viridis")
-pacman::p_load_gh("mwesthues/sspredr")
-
-
-# Common genotypes
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-mrna_inbreds <- genos %>%
-  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
-  split(.$Pool) %>%
-  map("G")
-
-# mRNA
-mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
-mrna <- t(mrna)
-mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
-mrna_df <- mrna %>%
-  as_data_frame() %>%
-  mutate(G = rownames(mrna)) %>%
-  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
-         Group = as.factor(Group))
-
-## Data pre-processing
-# Pre-process the data in the following order:
-# 1.    Box-Cox tranformation
-# 2.    Centering
-# 3.    Scaling
-# 4.    Remove highly correlated features
-mrna_lst <- mrna_df %>% 
-  split(.$Group) %>%
-  map(as.data.frame) %>%
-  map(~column_to_rownames(., var = "G")) %>%
-  map(.f = function(x) {
-    x$Group <- NULL
-    x
-  }) %>%
-  map(~as.matrix(.))
-
-# Data transformation
-trans_lst <- mrna_lst %>%
-  map(preProcess,
-      method = c("BoxCox", "center", "scale")) %>%
-  map2(.y = mrna_lst, .f = predict)
-
-mrna_cor_idx <- trans_lst %>%
-  map(cor) %>%
-  map(~findCorrelation(x = ., cutoff = 0.999))
-
-trans_lst <- lapply(seq_along(trans_lst), FUN = function(i) {
-  mat <- trans_lst[[i]]
-  high_cor_idx <- mrna_cor_idx[[i]]
-  if (isTRUE(length(high_cor_idx) != 0)) {
-    mat <- mat[, -high_cor_idx]
-  }
-  mat
-})
-
-
[…]

### 2017-03-15 analysis/spaced_snps.R

@@ -1,106 +1,0 @@
-if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "data.table", "rvest", "magrittr")
-
-## -- DATA SELECTION -----------------------------------------------------
-genos <- readRDS("./data/processed/common_genotypes.RDS")
-
-# snp
-snp <- readRDS("./data/processed/snp_mat.RDS")
-snp_nms <- genos %>%
-  filter(Data_Type == "snp") %>% 
-  .$G
-snp <- snp %>%
-  .[match(snp_nms, rownames(.)), ]
-
-# Load data frame with meta information on marker names and their positions in 
-# the genome.
-snp_map <- fread("./data/raw/marker_map.txt", sep = "\t")
-snp_map <- snp_map[complete.cases(snp_map), ]
-snp_map <- snp_map[snp_map$chromosome != 0, ]
-map_marker_names <- as.character(snp_map[["markername"]])
-map_marker_names <- gsub(pattern = "[.]", replacement = "_",
-                         x = map_marker_names)
-map_marker_names <- gsub(pattern = "[-]", replacement = "_", 
-                         x = map_marker_names)
-snp_map[, markername := map_marker_names]
-
-# Get the length of each of the ten maize chromosomes. We need this information
-# so that we can align the chromosomes sequentially and treat them like an 
-# entire maize genome with a global length. Thereby, we can select equally 
-# spaced SNPs with respect to the genome instead of individual chromosomes.
-ensembl <- read_html("http://ensembl.gramene.org/Zea_mays/Info/Annotation/")
-chr_lengths <- ensembl %>%
-  html_nodes(., "#chromosome_table td:nth-child(2)") %>%
-  html_text(.) %>%
-  .[seq_len(10)] %>%
-  c(0, .) %>%
-  .[seq_len(10)] %>%
-  as.list() %>%
-  map(as.data.frame) %>%
-  bind_rows(.id = "chromosome") %>%
-  rename(Chr_Length = `.x[[i]]`) %>%
-  mutate(Chr_Length = gsub(",", replacement = "", x = Chr_Length),
-         Chr_Length = as.numeric(Chr_Length),
-         Chr_Length = cumsum(Chr_Length),
-         chromosome = as.integer(chromosome))
-
-# Create global SNP positions
-global_map <- snp_map %>%
-  left_join(y = chr_lengths, by = "chromosome") %>%
-  as_tibble() %>%
[…]

### 2017-03-15 paper/genetics.tex
Init LaTeX paper

\documentclass[12pt,titlepage]{article}

%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
\usepackage{graphicx}
\usepackage{endfloat}
\usepackage{amsfonts}
\usepackage{mychicago}
\usepackage{subfigure}
\usepackage{genetics_manu_style}
\usepackage{authblk} % improved citations
\usepackage{setspace} %double spacing
\usepackage{lineno} % line numbers
\usepackage{booktabs} % book-quality tables
\usepackage{multicol} % list corresponding authors in two separate columns

%% Probably unnecessary packages.
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

\usepackage{xr}
\externaldocument{supp_info}

%% THE MANUSCRIPT TITLE
\title{Omics-based Hybrid Prediction in Maize}

% Allow for multiple authors to share the same institution.
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\author{
  Matthias Westhues\thanks{equally contributing}\thanks{Institute of Plant Breeding, Seed Science and Population Genetics, University of Hohenheim, D-70599 Stuttgart, Germany},
  Tobias A. Schrag\samethanks[1]\samethanks[2],
  Claas Heuer\thanks{Institute of Animal Breeding and Husbandry, Christian-Albrechts-University Kiel, D-24098 Kiel, Germany},
  Georg Thaller\samethanks[3],
  H. Friedrich Utz\samethanks[2],
  Wolfgang Schipprack\samethanks[2],
  Alexander Thiemann\thanks{Biocenter Klein Flottbeck, Developmental Biology and Biotechnology, University of Hamburg, D-22609 Hamburg, Germany},
  Felix Seifert\samethanks,
  Anita Ehret\samethanks[3],
  Armin Schlereth\thanks{Max-Planck Institute of Molecular Plant Physiology,
  D-14476 Potsdam, Germany},
  Mark Stitt\samethanks,
  Zoran Nikoloski\samethanks,
  Lothar Willmitzer\samethanks,
  Chris C. Sch{\"o}n\thanks{Plant Breeding, Technische Universit{\"a}t
  M{\"u}nchen, D-85354 Freising, Germany},
  Stefan Scholten\samethanks[4],
  Albrecht E. Melchinger\samethanks[2]
}

\renewcommand{\CorrespondingAddressA}{%
    University of Hohenheim \\
    Institute of Plant Breeding, Seed Science and Population Genetics \\ 
    Fruwirthstr. 21 \\ 
    Stuttgart 70599, GERMANY \\
    Tel.: (+49) 0711 459-22334 \\
    Fax: (+49) 0711 459-22343 \\ 
    \texttt{melchinger@uni-hohenheim.de} \vfill}

\renewcommand{\CorrespondingAddressB}{%
    University of Hamburg \\
    Biocenter Klein Flottbeck, Developmental Biology and Biotechnology \\ 
    Ohnhorststr. 18 \\ 
    Hamburg 22609, GERMANY \\ 
    Tel.: (+49) 40 42816-286 \\
    Fax: (+49) 40 42816-252 \\ 
    \texttt{stefan.scholten@uni-hamburg.de} \vfill}

\renewcommand{\RunningHead}{Omics-based Hybrid Prediction}
\renewcommand{\CorrespondingAuthorA}{Prof. Dr. A. E. Melchinger}
\renewcommand{\CorrespondingAuthorB}{PD Dr. Stefan Scholten}
\renewcommand{\KeyWords}{omics, hybrid breeding, prediction of complex traits, maize, endophenotype}

%% SOME COMMANDS FOR THE CONTENT OF THIS FILE.  NOT NECESSARY FOR
%% GENETIC_MANU_STYLE
\newcommand{\bp}{\mathbf{p}}
\newcommand{\LLL}{\mathcal{L}}

%% BEGIN DOC
\begin{document}

\maketitle
\doublespacing
\linenumbers

\begin{abstract}
Accurate prediction of traits with complex genetic architecture is crucial for 
selecting superior candidates in animal and plant breeding and for guiding 
decisions in personalized medicine.
Whole-genome prediction (WGP) has revolutionized these areas but has inherent 
limitations in incorporating intricate epistatic interactions.
Downstream 'omics' data are expected to integrate interactions within and 
between preceding biological strata and provide the opportunity to improve trait 
prediction.
Yet, predicting traits from parents to progeny has not been addressed by a 
combination of 'omics' data. 
Here, we evaluate several 'omics' predictors --- genomic, transcriptomic and 
metabolic data --- measured on parent lines at early developmental stages, and 
demonstrate that the integration of transcriptomic with genomic data leads to 
higher success rates in the correct prediction of untested hybrid progeny in 
maize.
Despite the high predictive ability of genomic data, transcriptomic data alone 
outperformed them and other predictors for the most complex heterotic trait, 
dry matter yield.
An eQTL analysis revealed that transcriptomic data integrate genomic information
from both, adjacent and distant sites relative to the expressed genes.
Together, these findings suggest that downstream predictors capture
physiological epistasis that is transmitted from parents to their hybrid
offspring.
We conclude that the use of downstream 'omics' data in prediction can exploit 
important information beyond structural genomics for leveraging the efficiency 
of hybrid breeding.
\end{abstract}

% Please add here a significance statement to explain the relevance of your work
\section{Article Summary}
The efficiency of modern hybrid breeding programs could be greatly improved
by pre-selecting the very best out of millions of theoretically possible
candidates using molecular 'omics'-features of their parents.
% Data for the following statement from 
% '/reduced_pred/100_Combined_Analyses/integrate.R'
Here, we show that transcriptomic data alone increased the predictive ability
for yield by 14.7\% compared to the gold standard, genomic data.
This improvement translates into a substantial increase in the probability of
success for predicting the best hybrids.
A combination of transcriptomic with genomic data yielded optimal and stable
predictive abilities across a wide range of traits.

% Introduction
\section{Introduction} 
Hybrid breeding, which entails crossing of lines from two genetically distant
germplasm collections --- called heterotic groups \cite{Melchinger1998} --- 
has emerged as a prime strategy to meet demands for a sustainable
intensification of agricultural production \cite{Duvick2005}.
However, unlocking the full potential of hybrid breeding requires accurate
prediction methods to efficiently identify the superior candidates out of the
millions of possible hybrids that could potentially be produced in each
cycle of an ordinary-sized breeding program.
With the advent of the doubled haploid (DH) technology \cite{Wedzony2009}
this prediction problem has become even more challenging because, based on 
breeder's experience, the vast majority ($\approx 90$\%) of competing lines in 
each heterotic group are "new" lines without any phenotypic records on hybrid
progeny from previous breeding cycles.
Consequently, among all hybrid combinations possible between lines from two
heterotic groups, about 81\% are T0 hybrids, 18\% are T1 hybrids and 1\% are T2
hybrids having zero, one or two parents, respectively, that have been previously
tested in other hybrid combinations.
Preselection of a few hundred of the most favorable hybrids with high success
rate could significantly reduce the labor-intensive and time-consuming
field-testing \cite{Kadam2016,Xu2016}.
This could greatly impact the efficiency of hybrid breeding and boost the
annual selection gain \cite{Longin2015}.

Whereas yield and other heterotic traits of hybrids are generally poorly
predicted by the performance of their parent lines \cite{Melchinger1998}, WGP
has emerged as a major tool for tackling this challenge
\cite{Massman2013,Technow2014}.
Nevertheless, there is evidence that, even with complete sequence information,
genomic prediction may not capture complex interactions between biological 
strata or downstream regulation that act through the entire cascade from 
genotype to phenotype \cite{Dalchau2011,Zhu2012,Rudd2015,Ritchie2015b}.
Most studies have evaluated predictive ability by looking at only one kind of 
endophenotype (intermediaries between genotype and phenotype 
\cite{Gottesman2003,Mackay2009}) such as the transcriptome 
\cite{Swanson-Wagner2006,Frisch2010,Xu2016} or the metabolome 
\cite{Meyer2007,Xu2016,Dan2016}.
The integration of different endophenotypic and genomic data is expected to
reflect more closely the phenotypic variability across individuals 
\cite{Mackay2009,Patti2012,Civelek2014}. 
Two recent studies that integrated multiple biological strata in predicting 
breast cancer risk \cite{Vazquez2016} and performance of maize inbred lines
\cite{Guo2016} demonstrated the benefit of this strategy.
However, unlike forecasting clinical or agronomic traits from endophenotypes of 
the same individual, hybrid breeding requires the prediction of the genotypic 
values (GV) of hybrid progeny based on parental information.
To achieve this objective, we used the BLUP approach --- originally developed in
animal breeding \cite{Henderson1984} --- for the more complex setting of hybrids 
between parents from two heterotic groups 
\cite{Bernardo1994,Bernardo1995,Bernardo1996,Massman2013}.
Here, we measured endophenotypes of parent lines to forecast the GV of
T0, T1 and T2 hybrid progeny by using prediction equations trained with "omics" 
information from related parent lines and phenotypic information on their
hybrid offspring.

\section{Materials and Methods}
\subsection{Genetic material and phenotyping}
The entire genetic material consisted of a set of 1,536 hybrids, denoted as 
$H_{Tot}$, produced in 16 factorial mating designs between 142 Dent and 103 
Flint lines from the maize breeding program at the University of Hohenheim, on 
which agronomic data for silage maize production, as well as pedigree and 
genomic data were available.
A subset of this material, albeit from different trials, has been used for
genomic prediction of traits related to grain maize production
\cite{Technow2014}.
For hybrid prediction, we used a core set $H \subset H_{\text{Tot}}$ of 617 
hybrids, produced in six factorials with hybrid sets $H_{\text{FAC}^{(i)}} (i =
1, 2, \dots, 6; H = \bigcup_{i=1}^{6} H_{\text{FAC}^{(i)}})$ from crosses 
between 57 Dent and 41 Flint inbred lines, denoted as $D = \{1, 2, \dots, 57 \}$ 
and $F = \{1, 2, \dots, 41 \}$ (File S1).
All hybrids were evaluated in field experiments at three or more 
agro-ecologically diverse locations across Germany.
In the trials of each factorial, which included at least five common check 
genotypes, the entries were randomized in $\alpha$ lattice designs and planted 
in 2-row plots.
Dry matter yield (DMY, $t/ha$) and dry matter content (DMC, \%) of whole-plant
aboveground biomass were determined by established procedures 
\cite{Riedelsheimer2012b}. 
For quality traits, contents of fiber (ADF, \%), fat (FAT, \textperthousand), 
protein (PRO, \textperthousand), starch (STA, \%), and sugars (SUG,
\textperthousand) in dry matter were measured in the harvested plant material 
using calibrated near-infrared spectroscopy (NIRS; \citeNP{Grieder2011}, 
File S1).

\subsection{Pedigree-based relationship coefficients}
Pedigree records were gathered for all parent lines at least back to the 
generation of their grandparents.
Coancestry coefficients were calculated using SAS (version 9.4, SAS Institute) 
for all possible pairs of lines in each heterotic group according to established 
rules \cite{Falconer1996} under the following assumptions \cite{Cox1986}:
(i) all lines in a pedigree are genetically homogeneous and homozygous, (ii) 
pairs of genotypes with no known common parentage are unrelated, and (iii) a 
line derived from a cross or backcross obtained a proportional fraction of the 
genome from each parent, as expected under Mendelian inheritance in the absence 
of selection.

\subsection{Genotyping}
Genotyping of all inbred lines was performed with the Illumina SNP chip
MaizeSNP50 \cite{Ganal2011}.
After performing a commonly used quality check \cite{Technow2014} and 
imputation of missing data \cite{Browning2009}, a total of 21,565 polymorphic 
SNPs was available and used for all further analyses.

\subsection{Metabolite profiling}
Seedlings of all parental inbred lines were grown under controlled conditions 
inside climate chambers to quantify their root metabolite profiles as detailed
by \citeN{DeAbreueLima2017}.
The experiment was laid out as a randomized incomplete block design with
replicated germination boxes.
For leaf metabolic profiles (known metabolites and unannotated chromatographic
peaks), a field experiment was carried out in an $\alpha$ lattice design with 
two replications at one location in southern Germany in the spring of 2012.
Excision of leaves at the third leaf stage was performed according to an 
established protocol \cite{Riedelsheimer2012b} and finalized within 45 minutes 
for the entire experiment. 
For both profiling procedures, all material was transferred directly into 
containers with dry ice and then into liquid nitrogen to quench metabolic 
activity.

\subsection{Transcriptome profiling}
For transcriptome profiling, five seeds per parent line were taken from the 
same seed lot as used for metabolite profiling and laid out inside a climate 
chamber in a randomized complete block design with five replications.
Seedlings were sampled seven days after sowing, snap-frozen in liquid nitrogen,
and stored at -80$^{\circ}$C until use.
Prior to mRNA extraction, all replicates of a genotype were pooled and
homogenized.
A custom 2K-microarray (GPL22267) was assembled from a subset of the 47K maize 
oligonucleotide array (GPL6438).
Two-color hybridizations were carried out separately for each of the six 
factorials using interwoven loop designs \cite{Kerr2001}.
The average number of shared genotypes between factorials was 4.5 and ranged 
from 2 to 10.

\subsection{Statistical analysis of agronomic traits}
Agronomic data were analyzed in two stages, following \citeN{Technow2014}. 
In the first stage, and separately for each environment, adjusted-entry means 
were computed for every hybrid using REML-based linear mixed-model analyses.
In the second stage, best linear unbiased estimates (BLUEs) were computed for 
all hybrids in $H_{Tot}$.
The BLUEs of hybrids in the core set $H$ served as response variables in our 
hybrid prediction models and cross-validation routines. 
For all predictions we used computationally efficient best linear unbiased 
predictor (BLUP) models, which have the same properties as those of a selection 
index because we previously accounted for fixed effects 
(\citeNP{Mrode2014}, pp. 34, 311, 312).
For general and specific combining abilities (GCA and SCA) of parent lines,
we used ASReml \cite{asreml} to compute best linear unbiased predictors (BLUPs), 
variance components ($\sigma^2_{\text{GCA}^{D}}$, $\sigma^2_{\text{GCA}^{F}}$, 
$\sigma^2_{\text{SCA}}$) and entry-mean heritabilities ($H^2$) of all hybrids
in $H_{Tot}$, treating all effects in the model as random.
The covariance matrices of the GCA and SCA effects were defined by multiplying
the variance components with their respective genomic relationship matrices 
(File S1).

\subsection{Statistical analysis of endophenotypes}
Raw data were normalized using established procedures for metabolites
\cite{VandenBerg2006} and transcripts \cite{Smyth2003,Ritchie2007}.
From these data, we obtained BLUEs for metabolite levels and transcript 
abundance of each line using REML-based mixed-model analyses.
The statistical models for the analysis of metabolite profiles accounted for 
various experimental effects as detailed by \citeN{DeAbreueLima2017}.
After applying quality checks and computing BLUEs, 92 leaf metabolic analytes 
and 283 root metabolic analytes remained for further analyses.
BLUEs for transcriptomic data were computed using the R-package \emph{limma} 
\cite{Ritchie2015a} in reference to established protocols 
\cite{Smyth2003,Ritchie2007,Frisch2010}.
The design matrix for the linear model was based on the dye-labeling of a 
reference genotype and included a fixed effect for each factorial design 
in order to account for the use of biological replicates of some genotypes in 
different arrays. 
All gene expression values were subsequently computed, based on the log-ratio 
relative to this common genotype \cite{Smyth2004}.
In total, 1,323 gene expression profiles were available.
Repeatabilities ($w^2$) were estimated for each endophenotype at the inbred line 
level using the same models as for the computation of BLUEs, but treating the 
genotype effect as random.
This analysis was performed jointly for the Dent and Flint lines allowing for
different means and heterogeneous genotypic variances of the heterotic groups,
but assuming a common error variance.
Variance components were estimated by Gibbs sampling using the \emph{R} package
\emph{MCMCglmm} \cite{Hadfield2010}.

\subsection{Prediction models and model evaluation}
Predictions of hybrid performance were compared on the basis of the core set
of hybrids $H$ and the corresponding sets of parent lines $D$ and $F$ on which
data for all five predictors (P, pedigree; G, genomic; T, transcriptomic; L,
leaf metabolic; R, root metabolic data) were available with the exception of
missing data in a few lines for R.
The matrices $\mathbf{W}_{D}$ and $\mathbf{W}_{F}$ are matrices of standardized 
feature measurements for the various predictors (G, T, L, R).
The matrix $\mathbf{W}$ has dimension 'number of parent lines in the 
corresponding heterotic group' times 'number of features'.
The columns in $\mathbf{W}_{D}$ and $\mathbf{W}_{F}$ are centered and
standardized to unit variance, respectively.

The kernels pertaining to each predictor and lines from each heterotic group
--- corresponding to genomic relationship matrices in the case of SNPs --- can 
then be defined as
\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{G}_{D} = \frac{1}{W} \mathbf{W}_{D} \mathbf{W}_{D}^{\top}, \
  \mathbf{G}_{F} = \frac{1}{W} \mathbf{W}_{F} \mathbf{W}_{F}^{\top},
\end{equation}

where $W$ denotes the number of features \cite{VanRaden2008}.
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{G}_{D}$ and $\mathbf{G}_{F}$, respectively.

The universal model for GCA and SCA effects was:

\begin{equation} \label{eq:HypredDominance}
  \mathbf{y} = \mu + 
  \sum_{p=1}^{P} \mathbf{Z}_{D} \mathbf{g}_{Dp} +
  \sum_{p=1}^{P} \mathbf{Z}_{F} \mathbf{g}_{Fp} +
  \sum_{p=1}^{P} \mathbf{Z}_{S} \mathbf{s}_{p} +
  \mathbf{\epsilon},
\end{equation}

where $\mathbf{y}$ is the vector of observed hybrid performance (BLUEs), $\mu$ 
is the fixed model intercept, $\mathbf{Z}_{D}$ is the
corresponding design matrix associating the random GCA effects of the lines in
$D$ ($\mathbf{g}_{Dp}$) with $\mathbf{y}$, $\mathbf{Z}_{F}$ is the 
corresponding design matrix associating the random GCA effects of the lines in 
$F$ ($\mathbf{g}_{Fp}$) with $\mathbf{y}$ and $\mathbf{Z}_{S}$ is a design 
matrix associating the SCA effects ($\mathbf{s}_{p}$), pertaining to hybrid 
combinations for the $p$-th predictor data type with the corresponding hybrid 
measurements in $\mathbf{y}$.
The random effects ($\mathbf{g}_{Dp}$ and $\mathbf{g}_{Fp}$) have expectation 
zero and variance equal to $\mathbf{G}_{Dp} \sigma^{2}_{\text{GCA}^{Dp}}$ and 
$\mathbf{G}_{Fp} \sigma^{2}_{\text{GCA}^{Fp}}$for the GCA effects of the Dent
and Flint lines, respectively,
$\mathbf{S}_{p} \sigma^{2}_{\text{SCA}^{p}}$ for the SCA effects and
$\mathbf{I} \sigma^2_{\epsilon}$ for the residual error.
For each pair of hybrids $i \times k$ and $j \times l$, the corresponding
elements in $\mathbf{S}_{p}$ were obtained as the product of the respective
elements $f_{ij}$ in $\mathbf{G}_{Dp}$ and $f_{kl}$ in $\mathbf{G}_{Fp}$,
respectively \cite{Schnell1965,Bernardo1996,Massman2013} (File S1).
In the absence of epistasis, this model is equivalent to a feature model
accounting for dually defined additive effects in each heterotic group and
dominance effects between them.
Extensions of the single-predictor models were made by adding GCA and SCA
effects for any additional predictor assuming stochastic independence of
effects.
In order to obtain unbiased estimates of the predictive ability and to compare
different models and predictor combinations, following \citeN{Technow2014}, we 
devised a cross-validation (CV) scheme, stratified by the parent lines and using 
1,000 runs (CV1000, File S1). All prediction models were implemented using the 
\emph{R} package \emph{BGLR} \cite{Perez2014}.

\subsection{Comparison of predictive abilities}
Predictive abilities were obtained by calculating Pearson correlations between
predicted ($\hat{y}$) and observed phenotypes ($y$), separately for three
test set partitions (T0, T1 and T2 hybrids).
For each CV run, the training and validation sets were stored to ensure the
validity of comparisons between any predictor and combinations thereof.
For any two predictors, say $A$ and $B$, we have the corresponding vectors 
with predictive abilities $p_{A}$ and $p_{B}$ of length 'number of cross
validation runs'.
For comparing predictive abilities obtained for different predictors and models,
we first ensured that the difference between any two scenarios (\textit{i.e.}, 
$p_{A} - p_{B}$) is normally distributed \cite{Dietterich1998}.
Then, we applied multiple contrast tests on the mean of the distribution of 
predictive abilities, obtained from the 1,000 CV runs, using the \emph{R} 
package \emph{multcomp} \cite{Hothorn2008} and Tukey contrasts.
This procedure is similar to the pairwise \textit{t}-test \cite{Guo2016} but 
accounts for multiple comparisons.

\subsection{Evaluation of a pre-selection bias in transcriptomic data}
A custom 2K-microarray (GPL22267) was assembled from a subset of the 47K maize 
oligonucleotide array (GPL6438), based on association of genes with hybrid
performance or mid-parent heterosis for grain yield and grain dry matter
content of maize. 
These two traits were evaluated in separate grain-yield trials with hybrids from 
factorial $H_{\text{FAC}^{(1)}}$ (\citeNP{Frisch2010,Thiemann2010}, File S1). 
To ensure that no pre-selection bias was introduced in hybrid prediction using 
these transcriptomic data, we compared predictive abilities for $H$ with 
predictive abilities when excluding $H_{\text{FAC}^{(1)}}$.

\subsection{Association mapping}
For each of the seven agronomic traits, we performed a genome-wide association
study (GWAS) with GCA effects of all 142 Dent and 103 Flint parent lines as
response variables using the EMMAX-method \cite{Kang2010} as implemented in 
\emph{cpgen} \cite{Heuer2015}.
To avoid using the marker data twice, GCA effects were calculated using only
pedigree information.
Furthermore, an eQTL analysis was carried out to examine statistically 
significant associations between genomic and transcriptomic data for the parent
lines ($D$ and $F$) of the core set $H$ plus five additional lines.
This was accomplished in the same way as in the GWAS for agronomic traits, but
here the BLUPs of the transcriptomic data of each mRNA were used as the 
response variables.
Associations in each GWAS were declared statistically significant at 
$\alpha = 0.05$ after Bonferroni correction.

\subsection{Probability of success}
Calculations of the conditional probability ($P[r, \beta]$) of a hybrid, 
selected at random from the upper $\beta$ percent fraction of the distribution 
of predicted values having a phenotypic value contained in the upper $\beta$ 
percent of the distribution of observed phenotypic values, assumed a bivariate
normal distribution

\begin{align}
  \begin{bmatrix}
    \hat{y} \\
    y
   \end{bmatrix}
  & \sim
  N \Bigg(
  \begin{bmatrix}
    0 \\
    0
  \end{bmatrix},
  \begin{bmatrix}
    1 & r \\
    r & 1
  \end{bmatrix}
  \Bigg).
\end{align}

and were performed within the \emph{R} statistical environment to solve the 
required integrals \cite{Robson1967}.

\subsection{Principal component analysis}
For principal component analyses of predictor data, all variables were scaled 
and centered.
Clusters represent two component mixtures of bivariate t-distributions, which
were estimated using Maximum Likelihood.
Ellipses were drawn based on the 0.95 quantiles of the respective bivariate
t-distributions.
Unless stated otherwise, all statistical analyses were carried out inside the 
\emph{R} environment for statistical computing \cite{Team2016}.

\subsection{Data availability}
Adjusted means of hybrid performances are provided in File S2, genomic marker
data are provided in File S3, genomic marker positions are provided in File
S4, transcriptomic BLUEs are provided in File S5, leaf metabolic BLUEs are
provided in File S6 and root metabolic BLUEs are provided in File S7.
Genotypes can be assigned to either the Dent or the Flint heterotic group using
a classification table provided in File S8.
Code used to analyze the data is available upon request.

\section*{Results}
\Genetics2level{Agronomic data}
Mean values of the 1,536 hybrids for the seven evaluated agronomic traits, with 
relevance to animal feed and biogas production, were of the same magnitude as 
reported by \citeN{Riedelsheimer2012b} and \citeN{Grieder2011}.
For all traits, $\sigma^{2}_{\text{GCA}^{D}}$ and
$\sigma^{2}_{\text{GCA}^{F}}$, describing the main effects of the parents from
each heterotic group, together explained more than 93\% of the genotypic
variance among hybrids (Table \ref{table:VarianceComponents}).
Heritabilities were moderate to high for all agronomic traits, indicating a 
high precision of field experiments and data collection 
(Table \ref{table:VarianceComponents}).

% -- TABLE 1
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[tbhp]
\centering
\caption{
  Summary of agronomic traits.
  Traits are characterized by overall mean ($\mu$), variance components of GCA
  effects for Dent ($\sigma^2_{\text{GCA}^D}$) and Flint lines
  ($\sigma^{2}_{\text{GCA}^{F}}$) and SCA effects ($\sigma^{2}_{SCA}$)
  (followed by s.e.m.) as well as entry mean heritabilities ($H^2$).
}
\label{table:VarianceComponents}
\begin{tabular}{@{}llllll@{}}
\toprule
Trait                   & $\mu$ & $\sigma^{2}_{\text{GCA}^{D}}$ & $\sigma^{2}_{\text{GCA}^{F}}$ & $\sigma^{2}_{\text{SCA}}$ & $H^{2}$         \\ \midrule
  DMY (t/ha)                    & 19.00 & $1.51 \pm 0.22$               & $1.00 \pm 0.18$               & $0.17 \pm 0.03$           & $0.82$ \\
  DMC (\%)                    & 34.13 & $4.17 \pm 0.59$               & $5.03 \pm 0.79$               & $0.49 \pm 0.07$           & $0.91$ \\
  ADF (\%)                    & 20.93 & $0.27 \pm 0.06$               & $0.40 \pm 0.09$               & $0.02 \pm 0.02$           & $0.43$ \\
  FAT (\textperthousand)                    & 30.02 & $1.01 \pm 0.19$               & $2.10 \pm 0.37$               & $0.17 \pm 0.05$           & $0.73$ \\
  PRO (\textperthousand)                    & 69.65 & $3.11 \pm 0.53$               & $2.77 \pm 0.51$               & $0.29 \pm 0.10$           & $0.70$ \\
  STA (\%)                    & 35.56 & $2.95 \pm 0.51$               & $3.33 \pm 0.63$               & $0.24 \pm 0.10$           & $0.69$ \\
  SUG (\textperthousand)                    & 38.12 & $31.33 \pm 5.15$              & $31.90 \pm 5.73$              & $4.12 \pm 1.02$           & $0.77$ \\ \bottomrule
\end{tabular}
\end{table}

\Genetics2level{Predictor data}
% Repeatabilities
Repeatabilities ($w^{2}$) for endophenotypes varied considerably in both groups
of parents (Fig. S\ref{fig:Repeatability}\textit{A}) with average values
ranging from 0.31 to 0.41, except for transcriptomic data in Flint material 
where the average repeatability was only 0.18.
Nevertheless, in the latter case, 291 out of 1,323 transcripts still
exceeded a threshold of $w^{2} = 0.4$.

Dent and Flint lines were clearly separated in principal component analyses of 
genomic and transcriptomic data (Fig. \ref{fig:PredictorProperties}\textit{A}).
However, they overlapped for leaf metabolic and, to an even greater extent, for
root metabolic data.
Off-diagonal elements of the kernels $\mathbf{G}_{D}$ and $\mathbf{G}_{F}$, 
respectively, showed moderate correlations between genomic and transcriptomic 
data ($\rho_{D} \approx 0.56, \rho_{F} \approx 0.44$, 
Fig. S\ref{fig:Rel_Assoc}). 
Correlations between the off-diagonal elements of the $\mathbf{G}$-matrices were 
highest for the comparison between genomic and pedigree data 
($\rho_{D} \approx 0.72$, $\rho_{F} \approx 0.63$).
Intriguingly, the associations between the $\mathbf{G}$-matrices for the root
and leaf metabolic data were very low ($\rho_{D} \approx 0.12$,
$\rho_{F} \approx 0.06$).

% Figure 1
\begin{figure}[H]
\centering
\includegraphics[width=0.90\linewidth]{./figures/predictor_properties_vertical.pdf}
\caption{
  Properties of predictor data for Dent (red) and Flint (teal)
  parent lines.
  (\textbf{A}) Principal component analysis (PCA) of both
  groups for genomic (G), transcriptomic (T), leaf metabolic (L) and root
  metabolic (R) data. The variance explained by PC 1 (x axis) and PC 2 (y
  axis) are shown in the caption of each facet.
  (\textbf{B}) Linkage disequilibrium decay as a function of the distance
  between two loci using 40 bins of 0.125 Mbp width, each.
  The median $r^2$ is depicted as a horizontal bar whereas the mean $r^2$ is 
  depicted as a white diamond.
  }
\label{fig:PredictorProperties}
\end{figure}

We observed high median pairwise linkage disequilibrium (LD) between
SNP markers ($r^{2} \approx 0.39$ in Dent and $r^{2} \approx 0.37$ in Flint 
material) at a distance of $\Delta \text{Mbp} \leq 0.125$ 
(Fig. \ref{fig:PredictorProperties}\textit{B}). 
After an initial drop in $r^{2}$ for $\Delta > 0.125$, substantial long-range
LD remained.
Large differences in allele frequencies in the two heterotic groups were 
present for 57\% of SNPs (Fig. \ref{fig:Circos}A,B) --- particularly in the 
telomeric regions of the genome.
An eQTL analysis performed with the parent lines suggests that transcript 
abundance integrates variegated genetic information given the fact that i) on 
the same chromosome, significant associations not only occurred between 
adjacent but also between distant pairs of expressed genes and SNPs and ii) 50\% 
of the significant associations ($\alpha = 0.05$, Bonferroni-corrected) occurred
between expressed genes and SNPs on different chromosomes
(Fig. \ref{fig:Circos}).

% Figure 2
\begin{figure}[H]
\centering
\includegraphics[width=0.95\linewidth]{./figures/86_13_eQTL_Circos.pdf}
\caption{
  Distribution and relationship of genomic and transcriptomic
  data for the ten maize chromosomes.
  Centromeres for each chromosome are depicted as vertical red lines.
  (\textbf{A,C}) Density of SNPs and mRNAs, respectively, across the ten maize
  chromosomes.
  (\textbf{B,D}) Statistical significance ($-\log_{10} \text{p-value}$) for
  differences in SNP allele frequencies and expression levels of mRNAs,
  respectively, between Dent and Flint lines.
  (\textbf{Center}) Links between any statistically significant
  ($\alpha = 0.05$ after Bonferroni correction) association between SNPs and
  mRNAs.
  Associations are displayed as links for SNPs on chromosome 5, for which the
  distribution of associations is representative for the entire genome, using 
  red color for Dent parent lines and teal color for Flint parent lines.
}
\label{fig:Circos}
\end{figure}

\Genetics2level{Predictive abilities}
Assuming a polygenic architecture for all traits, as suggested by results from
a GWAS (Fig. S\ref{fig:GCA_GWAS_Manhattan}), we chose the best linear unbiased
predictor (BLUP) method as a baseline for prediction of T0, T1 and T2 hybrids.
Given that we corrected for fixed effects in advance, this method is equivalent
to a selection index.
A cross-validation scheme with 1,000 runs (CV1000), stratified by the parent 
lines, was devised (File S1, Fig. S\ref{fig:CV_Overview}).
Our main emphasis was on predicting T0 hybrids given the fact that they 
constitute the majority of possible hybrids in practical breeding programs 
\cite{Kadam2016}.

For predictive abilities ($r$) of T0 hybrids, transcriptomic data alone were the
best predictor ($\alpha = 0.01$) for the most complex and highly heterotic
trait, DMY, as well as for PRO (Fig. \ref{fig:PredAbility}\textit{A}).
Intriguingly, owing to the marked non-linear relationship between $r$ and
the probability of successfully selecting the best hybrid candidates
$P[r, \beta]$ (Fig. \ref{fig:PredAbility}\textit{B}), the 14.7\% higher $r$ for
DMY, when using transcriptomic over genomic data, yields an increase in
$P[r, \beta]$ of 85\% at a selection intensity $\beta = 0.01\%$.
This selection intensity corresponds to picking the top 100 out of $10^{6}$ 
predicted hybrids for production and intensive testing in field trials.

% Figure 3
\begin{figure*}[tbph]
\centering
\includegraphics[width=17.8cm,height=8.89cm]{./figures/t0_pred_and_success.pdf}
  \caption{
  Predictive abilities (r) from BLUP models using a CV scheme with 
  sampling of $|H_{TRN}| = 200$ hybrids, $|D_{TRN}|=40$ Dent and $|F_{TRN}|=33$ 
  Flint parent lines for various predictors and combinations thereof 
  (P, pedigree; G, genome; T, transcriptome; L, leaf metabolome; R, root 
  metabolome).
  (\textbf{A}) Comparison of $r$ values from 1,000 CV runs for T0 hybrids and
  seven agronomic traits.
  Differences among the means of predictive abilities were tested by Tukey's HSD 
  and the same letter was used for non-significant ($\alpha = 0.01$) 
  differences.
  (\textbf{B}) Success rate of selecting superior hybrids ($P[r,\beta]$).
  $P[r, \beta]$ is a function of the predictive ability $r=r(y,\hat{y})$ and
  refers to the conditional probability of a hybrid, selected at random from the 
  upper $\beta \text{\%}$ fraction of the distribution of predicted values 
  ($\hat{y}$), having a phenotypic value contained in the upper $\beta \text{\%}$ 
  of the distribution of phenotypic values $y$.
  Observed predictive abilities ($r$) for T0 hybrids and the trait DMY are 
  displayed as vertical, colored lines for three predictors.
  }
\label{fig:PredAbility}
\end{figure*}

Compared to other individual predictors, predictive abilities obtained with
genomic data alone were significantly ($\alpha = 0.01$) higher for FAT and SUG.
Root metabolites displayed moderate to high predictive abilities for DMY and
FAT, but did not perform well otherwise.
Leaf metabolites performed relatively poorly for all traits.
Regardless of the trait, combinations of genomic and transcriptomic information 
displayed robust and consistently high predictive abilities.
Except for PRO, incorporating additional endophenotypes as predictors into our 
models did not yield statistically significant ($\alpha = 0.01$) improvements 
but remained at the same level compared to combining genomic and transcriptomic 
data.
Incorporating SCA effects into our models did not further improve predictive
abilties ($\alpha = 0.05$, Fig. S\ref{fig:GCA_vs_SCA_PredAbility}).
Results for the combination of other predictors with metabolic data are not
presented because no improvement of predictive abilities over the combination 
of genomic data with transcriptomic data and pedigree data could be achieved.
Finally, we assessed the influence of the number of SNPs and mRNAs on predictive 
abilities. 
For genomic data, a subset of 5,000 SNPs already yielded the same predictive 
ability as when using the entire available set and for transcriptomic data, the
predictive ability improved only marginally with subsets larger than 50\% of 
the available transcripts (Fig. S\ref{fig:Feature_Filter}).

\section*{Discussion}
\Genetics2level{A paradigm shift in hybrid breeding}
Hybrid breeding programs are generally based on genetically divergent heterotic 
groups \cite{Bernardo2010}.
Their use enables a better exploitation of heterosis when conducting crosses 
between them \cite{Melchinger1998} and is expected to reduce the ratio of 
specific to general combining ability variance 
($\sigma^{2}_{\text{SCA}} : \sigma^{2}_{\text{GCA}}$) in the crosses, thereby 
allowing for the selection of hybrids largely on the basis of GCA of their 
parent lines \cite{Reif2007}.
However, obtaining accurate estimates of GCA requires the evaluation of 
new lines in combinations with testers from the opposite heterotic group in
multi-environment field trials.
The promise of hybrid prediction is to accelerate breeding programs by skipping 
a large share of these tests in favor of selecting the most promising hybrids
before they are even produced \cite{Technow2014}.
This approach involves the prediction of an impressive number of putative hybrid 
candidates ($n^{2}$) using predictor data collected on only 2$n$ parent lines.
Crucial for hybrid prediction are predictors, which not only reflect the 
relationship between parental inbred lines but also the interaction of the two 
parental genomes in their hybrid progeny.

\subsection{Heterotic groups}
Because of genetic drift and selection for hybrid performance, allele
frequencies are expected to diverge in the two heterotic groups, thereby
enlarging their genetic distance \cite{Falconer1996,Reif2007,Lariepe2017}.
Consistent with this hypothesis and two pilot studies with U.S. maize lines
\cite{Gerke2015,Hall2016}, Dent and Flint lines in our study were clearly 
separated in principal component analyses of genomic and transcriptomic data.
With large differences in allele frequencies $p^{I}$ and $p^{II}$ in the two
heterotic groups, as observed for 57\% of SNPs, dominance variance
$\sigma_{D}^{2}$ becomes very small because it is a function of the product
$p^{I} ( 1 - p^{I} ) p^{II} ( 1 - p^{II} )$ (\citeNP{Stuber1966},
File S1).
$\sigma_{D}^{2}$ is the main component contributing to the variance of 
the specific combining ability  effects ($\sigma_{\text{SCA}}^{2}$), describing 
interactions among the parental genomes in hybrid combinations.
It was therefore not surprising that the variances of the general combining
ability (GCA) effects ($\sigma_{\text{GCA}^{D}}^{2}$ and 
$\sigma_{\text{GCA}^{F}}^{2}$), describing the main effects of the parents from 
each heterotic group, together explained more than 93\% of the genotypic 
variance among hybrids for agronomic traits, which is consistent with an earlier
study on silage maize of the Dent $\times$ Flint heterotic pattern
\cite{Argillier2000}.
The importance of GCA in our material was further corroborated by statistically 
non-significant ($\alpha = 0.05$) differences in predictive abilities between 
models using only GCA effects and those that additionally incorporated SCA 
effects (Fig. S\ref{fig:GCA_vs_SCA_PredAbility}).
Nevertheless, in crops such as wheat, with no clearly defined heterotic groups
\cite{Zhao2015a} and greater importance of SCA, inclusion of SCA effects in
the model should improve predictive abilities.

\subsection{Properties of well-established predictors}
While pedigree data reflect the expected relationship between individuals, they 
do not necessarily depict their realized relationship.
Genomic data and downstream endophenotypes offer to improve upon this
pedigree-based approximation by more closely mirroring the transmission of 
genes between individuals and their interactions.
Genomic data have the advantage of reliably capturing Mendelian sampling, 
thereby improving pedigree-based prediction for many traits.
However, genomic data alone may not be the final answer for the prediction of
complex traits for two major reasons:
First, the number of samples in most studies is considerably smaller than the
number of genetic markers or even nucleotides of a genome.
This implies that just modeling additive effects already necessitates effect 
shrinkage.
More importantly, however, interactions between loci throughout the genome 
can be frequent \cite{Brem2005,Brown2014}, but attempts to incorporate this 
epistasis for the prediction of heterotic traits using genomic data have been 
disappointing when the prediction and the training set did not share the same
or closely related parents \cite{Jiang2015}.
This was true even when using recently developed, efficient models 
\cite{Jarquin2014,Martini2016} and suggests that genomic data capture 
only statistical epistasis, referring to genetic variation at the population 
level \cite{Sackton2016}, which is generally of negligible magnitude 
\cite{Hill2008,Mackay2014,Guo2016,Vazquez2016}.

\Genetics2level{Complementation of predictors}
\subsection{Flow of biological information}
It is well-known that genetic effects on the phenotype are mediated through
multiple layers of endophenotypes \cite{Civelek2014,Ritchie2015b} with 
information mainly flowing from the genome toward the phenotype via the 
transcriptome, the proteome and the metabolome with metabolite fluxes 
ultimately governing energy production and growth \cite{Cornish-Bowden2001}.
For most traits in our material, metabolite- and pedigree-based predictive
abilities were lower than those obtained with either transcriptomic or genomic
information.
However, consistently high predictive abilities across multiple traits could be
realized when combining multiple predictors, as has been reported previously in 
humans \cite{Vazquez2016} and maize inbred lines \cite{Guo2016}.
This suggests complementary properties of the different predictors resulting in 
better proxies for the complex interplay in gene networks than solely genomic 
information.
Such an advantage is particularly important for hybrid prediction when parents 
of prediction set hybrids are not closely related to parents of training set
hybrids (File S1) as was shown by the relative excellence of transcriptomic 
data and the use of multiple predictors for the prediction of traits in T0 
hybrids compared to T1 and T2 hybrids.

\subsection{Tapping new sources of information}
Whereas pedigree and genomic information are static, subsequent endophenotypes 
are characterized by pervasive interactions among and between each other 
\cite{Dalchau2011,Zhu2012} and are, to varying degrees, influenced by biotic 
\cite{Rudd2015,Tzin2015} and abiotic perturbations \cite{Caldana2011,Witt2012}.
So while endophenotypes do not exclusively report on physiological epistasis but 
also on non-heritable effects, they do seem to capture important information not
represented by the genome given their intermediate position in the
genotype-phenotype cascade.
We get support for this hypothesis from (i) merely low to moderate correlations 
between off-diagonal elements of the kernels of different predictors in our 
study, (ii) mounting evidence for further improvements of predictive abilities 
when complementing genomic prediction with other endophenotypes despite 
sufficient marker densities (Fig. S\ref{fig:Feature_Filter}, \citeNP{Guo2016})
and (iii) the integration of SNP information from close and distant eQTL in the 
transcripts analyzed in our study.
However, we concede that the number of parental genotypes in our mRNA assays
was too small to warrant a reasonable statistical power for detecting epistasis
in the expression of transcripts.
While two other studies also attempted to model interactions between different
predictors, we refrained from this approach given that their reported predictive
abilities based on interactions were not different from those in additive
models despite using much larger sample sizes \cite{Vazquez2016,Guo2016}.

\Genetics2level{Transcriptomic data}
\subsection{Utility of transcriptomic data for trait predictions}
Of particular note was the excellent performance of transcriptomic data in 
predicting dry matter yield and protein.
Evidence that parental gene expression patterns might be predictive of hybrid 
performance is given by (i) prevailing additive expression patterns in maize 
hybrids \cite{Springer2007,Stupar2008}, (ii) a positive correlation of the 
proportion of additive gene expression with the yield of hybrids \cite{Guo2006}, 
and (iii) co-localization of additively expressed genes with heterotic QTL 
\cite{Thiemann2014}.
According to metabolic flux theory, gene expression in hybrids at the 
mid-parent level can generate hybrid vigor by counterbalancing opposing 
detrimental expression levels in their parent lines on a genome-wide scale 
\cite{Kacser1981,Springer2007a}.
The same concept is expected to apply to other quantitative endophenotypes 
\cite{Lisec2011}.

\subsection{Pre-selection bias}
As pointed out earlier, our transcripts were pre-selected based on associations
with grain dry matter yield and grain dry matter content in hybrids, using a 
subset of the data included in our study ($H_{\text{FAC}^{(1)}}$).
Two findings indicate that no pre-selection was introduced:
(i) Relative differences in predictive abilities between transcriptomic and
pedigree or genomic data did not change when excluding genotypes from 
$H_{\text{FAC}^{(1)}}$ from the data (Fig. S\ref{fig:NoExp1}) and 
(ii) transcriptomic data performed rather poorly in predicting dry matter 
content although this trait was also among the criteria for the pre-selection 
procedure.
Finally, an independent study using RNA-Seq data for the prediction of traits
in maize inbred lines also reported exceptionally good performance of 
transcriptomic data in the prediction of multiple yield-related traits 
\cite{Guo2016}.

\Genetics2level{Relative excellence of predictors for different traits}
\subsection{Tissue and sampling time}
Despite the great prospects of using endophenotypes for trait predictions, some
aspects require careful consideration when using this approach.
A particular challenge in endophenotype-based prediction efforts is the selection
of a suitable tissue and sampling time.
Tissue-related effects regarding gene expression were found in studies on humans 
\cite{Yang2015,Mele2015,Searle2016} and \textit{A. thaliana} \cite{Schmid2005}
and in maize hybrids with respect to metabolome composition and metabolite 
abundance \cite{Witt2012}.
Moreover, the age of an organism can selectively influence the expression of 
genes as observed in studies on humans \cite{Mele2015,Yang2015} and 
\textit{C. elegans} \cite{Vinuela2010,Francesconi2014}.
The low correlations between the off-diagonal elements of the kernels calculated 
from root and leaf metabolites might therefore be a reflection of highly dynamic 
processes differing between tissues and during different developmental stages.
Whereas root metabolic data and transcriptomic data were obtained from
seedlings germinated in standard controlled conditions, leaf metabolic data were
derived from field-grown plants at a much later developmental stage, thereby
increasing the possibility of environmentally-induced modifications.
One might hypothesize, that the choice of sampling time and tissue could
influence the chances of successful trait prediction if such age- or 
tissue-dependent transcripts and metabolites are associated with a phenotypic 
or clinical trait.
Generally, sampling from the seedling would be highly desirable in breeding when 
trying to scale the use of endophenotypic data to even larger numbers of inbred 
lines and assay genotypes independent of the season.

\subsection{Feature selection}
Another explanation for trait-dependent excellence of any predictor might lie
in the sampling of features.
In this study, only a small subset of metabolites was sampled and even very 
recent technologies \cite{Xu2016,Dan2016} capture only a fraction of the 
estimated set of metabolites \cite{Fernie2007}.
Moreover, the smaller differences in metabolite levels between both heterotic 
groups (Fig. \ref{fig:PredictorProperties}) were most likely not conducive to
capturing basic components underlying complex heterotic traits.
It is also possible that transcriptomic data are associated with more
biological processes than metabolite data and better capture the genetic effects
relevant for the prediction of T0 hybrids.

\subsection{Prospects for metabolites}
Previously observed moderate metabolite-based predictive abilities for T1
hybrids \cite{Riedelsheimer2012b} were confirmed in our study
(Fig. S\ref{fig:PredAbilityT0-T2}), but for the majority of traits, root 
metabolites reached only medium and leaf metabolites even lower predictive 
abilities when predicting T0 hybrids.
Despite the aforementioned shortcomings of metabolites, they have shown to be 
intriguing predictors due to their phsyiological proximity to the phenotype,
which provides information that is impossible to infer from DNA or proteins
\cite{Fernie2012}, as well as encouraging results from other studies
\cite{Guo2016,Dan2016}.
A recently introduced technology, allowing for live-measurements of small 
molecules in the blood of living and awake animals \cite{Arroyo-Curras2017}, 
might overcome the problem of poorly time-resolved snap-shots of some 
metabolites with extremely fast turnover rates \cite{Arrivault2009} if modified
to properly work in plants.

\section*{Conclusions}
The use of whole-genome information has considerably advanced trait prediction
over traditional pedigree-based BLUP by incorporating previously unobservable 
Mendelian sampling.
Combining variegated sources of information promises to capture complex
interactions between genes and endophenotypes, leading to stable predictions
across traits.
Especially if an extremely small fraction of the candidates is selected from
the millions of possible new hybrids from each breeding cycle, the success of
forecasts is a strongly convex function of predictive ability
(Fig. \ref{fig:PredAbility}\textit{B}).
Therefore, considering endophenotypes could have a substantial effect on the 
success and economics of hybrid breeding.
Given the anticipated technological improvements in RNA-Seq and metabolite 
profiling, as well as the forthcoming adoption of the DH-technology for many
crops \cite{Kelliher2017}, a paradigm shift from exclusively genomic prediction 
models to more inclusive approaches seems imminent.

\section{Acknowledgments} 
We thank the staff of the Agricultural Experimental Research station,
University of Hohenheim, for excellent technical assistance in
conducting the field experiments. We are indebted to the group of R.
Fries from Technische Universit{\"a}t M{\"u}nchen for the SNP genotyping of the
parent inbred lines, to X. Mi for his assistance in preparing auxiliary
figures based on the Mathematica software, to C. Zenke for advice on the
computation of transcriptomic BLUEs and to P. Schopp for advice on
prediction models. The authors acknowledge support by the state of
Baden-W{\"u}rttemberg through bwHPC. This project was funded by the German
Federal Ministry of Education and Research (BMBF) within the projects
OPTIMAL (FKZ: 0315958B,0315958F), SYNBREED (FKZ: 0315528D) and by the
German Research Foundation (DFG, Grants No. ME 2260/5-1 and SCHO 764/6-1).
Financial support for M.W. was provided by the Fiat Panis foundation, Ulm,
Germany.

\nolinenumbers
% Bibliography
\bibliography{library}
\bibliographystyle{mychicago}
\end{document}

### 2017-03-15 paper/manuscript.tex
Init LaTeX paper

\documentclass[12pt,titlepage]{article}

%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
\usepackage{graphicx}
\usepackage{endfloat}
\usepackage{amsfonts}
\usepackage{mychicago}
\usepackage{subfigure}
\usepackage{genetics_manu_style}
\usepackage{authblk} % improved citations
\usepackage{setspace} %double spacing
\usepackage{lineno} % line numbers
\usepackage{booktabs} % book-quality tables
\usepackage{multicol} % list corresponding authors in two separate columns

%% Probably unnecessary packages.
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

%% THE MANUSCRIPT TITLE
\title{Omics-based Single Step Trait Prediction}

% Allow for multiple authors to share the same institution.
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\author{
  Matthias Westhues\thanks{Institute of Plant Breeding, Seed Science and Population Genetics, University of Hohenheim, 70599 Stuttgart, Germany},
  Claas Heuer\thanks{Institute of Animal Breeding and Husbandry, Christian-Albrechts-University Kiel, 24098 Kiel, Germany},
  Georg Thaller\samethanks[2],
  Rohan Fernando\thanks{Department of Animal Science, Iowa State University, 50011 Ames, Iowa, U.S.A.},
  Albrecht E. Melchinger\samethanks[1]
}

\renewcommand{\CorrespondingAddress}{%
    University of Hohenheim \\
    Institute of Plant Breeding, Seed Science and Population Genetics \\ 
    Fruwirthstr. 21 \\ 
    Stuttgart 70599, GERMANY \\
    Tel.: (+49) 0711 459-22334 \\
    Fax: (+49) 0711 459-22343 \\ 
    \texttt{melchinger@uni-hohenheim.de} \vfill}

\renewcommand{\RunningHead}{Single Step Prediction}
\renewcommand{\CorrespondingAuthor}{Prof. Dr. A. E. Melchinger}
\renewcommand{\KeyWords}{single step, omics, prediction, maize}

%% SOME COMMANDS FOR THE CONTENT OF THIS FILE.  NOT NECESSARY FOR
%% GENETIC_MANU_STYLE
\newcommand{\bp}{\mathbf{p}}
\newcommand{\LLL}{\mathcal{L}}

%% BEGIN DOC
\begin{document}

\maketitle
\doublespacing
\linenumbers

\begin{abstract}

  $\dots$

\end{abstract}

% Please add here a significance statement to explain the relevance of your work
\section{Article Summary}
$\dots$

% Introduction
\section{Introduction}

$\dots$

Our objectives were threefold: (1) Examine whether the single step prediction
framework can be transfered to hybrids, (2) explore the utility of single step 
predition with a quantitative predictor and (3) evaluate the impact of the
particular genetic space covered by the incomplete predictor on predictive
ability.

\section{Materials and Methods}
\subsection{Maize hybrids}
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines \cite{Westhues2017}. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper \cite{Westhues2017} and were genotyped using the Illumina 
BeadChip MaizeSNP50 \cite{Ganal2011}.
After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed 
using the Beagle software \cite{Browning2009}.
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) \cite{Westhues2017}.
After applying established normalization procedures 
\cite{Smyth2003,Ritchie2007}, gene expression BLUEs were computed for 1,323 
transcripts in reference to established protocols 
\cite{Smyth2003,Ritchie2007,Frisch2010} using the R-package \textit{limma}
\cite{Ritchie2015a}.

\subsection{Maize diversity panel}
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups \cite{Yang2014}.
All inbred lines were evaluated in five Chinese environments ranging from
$18$ to $30^{\circ}$N and from $102$ to $110^{\circ}$E \cite{Yang2014}.
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
six were used for this study.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines \cite{Fu2013,Li2013}.
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers \cite{Yang2014}.
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions \cite{Fu2013}.
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes and was kept for further analyses \cite{Li2013}.

\subsection{Population structure}
Principal component analyses (PCA) were run on scaled and centered predictor 
matrices using the \emph{pca} function from the R-package \textit{LEA} 
\cite{Frichot2015}.
Estimates of ancestry coefficients were estimated based on sparse nonnegative
matrix factorization algorithms \cite{Frichot2014} using the \emph{snmf} 
function in \textit{LEA}.
Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
repetitions for each value of $K$. 
For each $K$, only the repetition with the lowest cross-entropy was kept for 
further analyses.

\subsection{Core sampling}
In order to evaluate the influence of the genetic constitution of the set of
genotypes that has only data on one out of two predictors, we generated core 
samples using the R-package \textit{corehunter} \cite{Thachuk2009}.
These core samples were based on the 149 maize inbred lines from the diversity
panel which had gene expression as well as genomic information.
The size of the core set was varied in increments of ten percentage points and 
ranged from 10\% to 90\% of all 149 inbred lines.
For assembling the various core sets, we first computed the Modifed Rogers (MR)
distance between each pair of genotypes and then selected a predefined subset
of genotypes that display maximum average distance between each other.
MR was computed using the \emph{rogers.dist} function from the R-pacakge
\textit{poppr} \cite{Kamvar2015}.

\subsection{Prediction}
\subsubsection{Kernels}
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effects have been accomodated by the dependent variable 
\cite{Mrode2014}.
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the $l$-th group of inbred lines and the 
$o$-th predictor ($\mathbf{W}_{lo}$) has dimensions $n \times p$ where $n$
pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
the number of features.
For the diversity panel maize lines, $n$ corresponds to the number of genotypes
whereas, in the case of the hybrid data, $n$ corresponds to the number of
parent lines in the corresponding heterotic group.
In generating kernels for each predictor, all features in $\mathbf{W}_{lo}$ were
centered and standardized to unit variance.
Then, each kernel can be defined as

\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{K}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
\end{equation}

where $W_{lo}$ denotes the number of features \cite{VanRaden2008}.
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{K}_{lo}$.

\subsubsection{Universal breeding value model}
The universal model for breeding values was:

\begin{equation} \label{eq:KBLUPModel}
  \mathbf{y} = \mu + 
  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
  \mathbf{\epsilon},
\end{equation}

where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
respectively, $\mu$ is the fixed model intercept, $\mathbf{Z}_{lo}$ is the 
corresponding design matrix associating the random genotype effects 
(\textit{i.e.}, breeding values) of the lines in the $l$-th inbred line group 
and the $o$-th predictor type with $\mathbf{y}$.
Note that, in the case of hybrid data, $\mathbf{Z}_{lo}$ associates the random
general combining ability effects in the parental inbred lines (either Dent or
Flint, as expressed by $l$) with the vector of observed hybrid performance
$\mathbf{y}$.
The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
to $\mathbf{K}_{lo} \sigma^{2}_{{g}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
for the residual error.
All prediction models were implemented using the R-package \textit{BGLR}
\cite{Perez2014}.

\subsubsection{Single step prediction}
The breeding values in a prediction model are in general given by:
\begin{equation} \label{eq:mrnaebv}
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model,
which is equivalent to the BLUP model.

Consider now the situation in which one predictor is complete in the sense that
it contains information on the whole population whereas a second predictor is
incomplete and covers only a subset of the population.
We can take a similar route as in \citeNP{Fernando2014} and impute covariates 
of the incomplete predictor by using covariates of the complete predictor.
Let the subscript $1$ denote individuals covered only by the complete
predictor whereas individuals covered by both, the complete and the
incomplete predictor, are indicated by the subscript $2$.
Further, the covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align} \label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{K}_{11} - \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{K}_{21} = (\mathbf{K}^{11})^{-1}$
\cite{Legarra2009}.

The structure of the residual imputation/prediction error is known and can 
therefore be modelled.
The covariates not covered by the incomplete predictor can be predicted using 
the expectation of a multivariate normal random vector given correlated 
observations, which are related to the Best Linear Predictor.

The general model for the single step procedure can then be written as

\begin{equation} \label{eq:single-step-model}
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}^{Complete} \boldsymbol{\epsilon}^{Complete} + \mathbf{e},
\end{equation}

with

\begin{equation} \label{eq:single-step-submatrices}
\mathbf{X} = 
\begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}^{Complete} = 
\begin{bmatrix}
  Z_1 \\
  0 
 \end{bmatrix}
\end{equation}

Following the notation in \citeNP{Fernando2014} the phenotypes can be untangled 
like this:

\begin{align} \label{eq:entangled-augmented-single-step-model}
\begin{bmatrix}
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

The breeding values from that model are:

\begin{equation} \label{eq:breeding-values}
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}^{Complete} 
\end{equation}

Our final model looks like this:

\begin{align} \label{eq:final-model}
\mathbf{y} &= \mathbf{Xb} +
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l}^{Complete} \boldsymbol{\epsilon}_{l}^{Complete} +
\mathbf{e}.
\end{align}

For the hybrid data, the obtained GCA effects for the inbred lines represent 
half their breeding values.
The predicted agronomic performance is:

\begin{equation} \label{eq:predicted-performance}
\hat{\mathbf{y}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
\end{equation}

\subsubsection{Predictive ability and model validation}
For the validation of our predictions, we employed leave-one-out
cross-validation (LOOCV) routines.
In the case of the diversity panel maize inbred lines, LOOCV was performed by 
using a single genotype as a hold-out sample, which will subsequently be 
predicted, while using all other inbred lines for model training.
This process is repeated until all inbred lines have been used once for testing
and $n - 1$ times for model training.

For the hybrid data, LOOCV was carried out as follows:
Let $D$ and $F$ denote the set of parental inbred lines from the Dent and the 
Flint group, respectively.
Further, let $H \cap [D \times F]$ denote the set of hybrids from crosses 
between $D$ and $F$.
The training set ($H_{TRN}$) for the hybrid to be predicted ($H_{ij}$) --- 
where $i \in D_{TRN}$ and $j \in F_{TRN}$ --- was assembled as 
$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$ with
$D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.

We judged the performance of each model by looking at its predictive ability,
which is calculated as $\rho(\mathbf{y}, \mathbf{\hat{y}})$, where 
$\mathbf{\hat{y}}$ is the vector of predicted values from each LOOCV run.
Our confidence in the performance of each model was evalutaed by computing the
coefficient of variation for the predicted values:

\begin{equation} \label{eq:CV}
CV = \sqrt{\frac{\sum_{i = 1}^{n} \sigma^{2}_{\hat{y}(i)} / n}{\sum_{i=1}^{n}\hat{y}_{i} / n}},
\end{equation}

where $n$ denotes the number of genotypes (*i.e.*, inbred lines or hybrids).

\section*{Results}

\begin{figure}[H]
\centering
\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
\caption{
Influence of core set composition on predictive abilities.
Predictive abilities are based on single step predition with genomic data as the 
complete predictor and transcriptomic data as the incomplete predictor.
Results are shown for six agronomic traits and nine core sets of the maize 
diversity panel.
}
\label{fig:CoreSetInfluence}
\end{figure}

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Wed Mar 15 10:35:35 2017
% Script: ./analysis/maizego_plot_predictive_abilities.R
\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for the set of 211 tropical/subtropical lines from the maize diversity panel.} 
\begin{tabular}{rllllll}
  \toprule
 & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: no}}\\
snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: yes}}\\
mrna-none & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
  snp-none1 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Wed Mar 15 11:11:37 2017
% Script: ./analysis/uhoh_plot_predictive_abilities.R
\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids} 
\begin{tabular}{rllllll}
  \toprule
 & DMY & DMC & FAT & PRO & STA & SUG \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
ped-none & 0.54 (0.0493) & 0.48 (0.0521) & 0.21 (0.0365) & 0.6 (0.0187) & 0.35 (0.0427) & 0.42 (0.117) \\ 
  snp-none & 0.74 (0.0517) & 0.63 (0.0521) & 0.43 (0.0356) & 0.65 (0.0223) & 0.46 (0.0458) & 0.47 (0.1261) \\ 
  ped-snp & 0.71 (0.054) & 0.57 (0.0545) & 0.35 (0.0375) & 0.62 (0.0235) & 0.39 (0.0478) & 0.42 (0.1298) \\ 
  ped-mrna & 0.73 (0.056) & 0.53 (0.0594) & 0.35 (0.0398) & 0.61 (0.0254) & 0.31 (0.0504) & 0.37 (0.1396) \\ 
  snp-mrna & 0.75 (0.0533) & 0.61 (0.0553) & 0.43 (0.0374) & 0.65 (0.0239) & 0.44 (0.048) & 0.44 (0.1342) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
mrna-none & 0.75 (0.0552) & 0.34 (0.0619) & 0.4 (0.0412) & 0.47 (0.0293) & 0.16 (0.0496) & 0.27 (0.1447) \\ 
  ped-none1 & 0.46 (0.0469) & 0.53 (0.05) & 0.23 (0.0368) & 0.46 (0.0234) & 0.38 (0.0391) & 0.34 (0.1316) \\ 
  snp-none1 & 0.67 (0.0534) & 0.47 (0.0552) & 0.43 (0.0384) & 0.48 (0.027) & 0.36 (0.0457) & 0.35 (0.132) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Wed Mar 15 11:36:08 2017
% Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel.} 
\begin{tabular}{rllllll}
  \toprule
 & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
mrna-none\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
  snp-mrna\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
  snp-mrna\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
  snp-mrna\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
  snp-mrna\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
  snp-mrna\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
  snp-mrna\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
  snp-mrna\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
  snp-none\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

\Genetics2level{Some results section}

$\dots$

\Genetics2level{Another results section}

$\dots$

\section*{Discussion}
\Genetics2level{Some discussion section}

$\dots$

\subsection{More detailed discussion}

$\dots$

\Genetics2level{Another discussion section}

$\dots$

\section*{Conclusions}

$\dots$

\section{Acknowledgments} 
We thank the staff of the Agricultural Experimental Research station,
University of Hohenheim, for excellent technical assistance in
conducting the field experiments. We are indebted to the group of R.
Fries from Technische Universit{\"a}t M{\"u}nchen for the SNP genotyping of the
parent inbred lines, to X. Mi for his assistance in preparing auxiliary
figures based on the Mathematica software, to C. Zenke for advice on the
computation of transcriptomic BLUEs and to P. Schopp for advice on
prediction models. The authors acknowledge support by the state of
Baden-W{\"u}rttemberg through bwHPC. This project was funded by the German
Federal Ministry of Education and Research (BMBF) within the projects
OPTIMAL (FKZ: 0315958B,0315958F), SYNBREED (FKZ: 0315528D) and by the
German Research Foundation (DFG, Grants No. ME 2260/5-1 and SCHO 764/6-1).
Financial support for M.W. was provided by the Fiat Panis foundation, Ulm,
Germany.

\nolinenumbers
% Bibliography
\bibliography{library}
\bibliographystyle{mychicago}
\end{document}

### 2017-03-15 reports/Paper.Rmd
Init LaTeX paper

--- 2017-03-15 reports/methods.Rmd
@@
+---
+title: Omics-based Single Step Trait Prediction
+author:
+  - name: Matthias Westhues
+    email: matthias.westhues@uni-hohenheim.de
+    affiliation: University of Hohenheim
+  - name: Claas Heuer
+    email: cheuer@tierzucht.uni-kiel.de
+    affiliation: University of Kiel
+  - name: Georg Thaller
+    email: thaller@tierzucht.uni-kiel.de
+    affiliation: University of Kiel
+  - name: Rohan Fernando
+    email: rohan@iastate.edu
+    affiliation: Iowa State University
+  - name: Albrecht E. Melchinger
+    email: melchinger@uni-hohenheim.de
+    affiliation: University of Hohenheim
+    footnote: Corresponding Author
+address:
+  - code: University of Hohenheim
+    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
+  - code: University of Kiel
+    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
+  - code: Iowa State University
+    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
+date: "`r Sys.Date()`"
+knit: "bookdown::render_book"
+output:
+  bookdown::pdf_book:
+    base_format: rticles::elsevier_article
+bibliography: ["~/Dropbox/Paper/library.bib"]
+abstract: |
+  $\dots$
+---
+
+<!--chapter:end:index.Rmd-->
+
+# Introduction
+$\dots$
+
+Our objectives were threefold: (1) Examine whether the single step prediction
+framework can be transfered to hybrids, (2) explore the utility of single step 
+predition with a quantitative predictor and (3) evaluate the impact of the
+particular genetic space covered by the incomplete predictor on predictive
+ability.
+
+<!--chapter:end:introduction.Rmd-->
+
 # Materials and Methods
@@
 where $n$ denotes the number of genotypes (*i.e.*, inbred lines or hybrids).
+
+<!--chapter:end:methods.Rmd-->
+
+# Results
+
+![Bullshit]("./tabs_figs/core_fraction_predictive_ability_trend.pdf")
+
+<!--
+\begin{figure}[H]
+\centering
+\includegraphics{./tabs_figs/core_fraction_predictive_ability_trend.pdf}
+\caption{
+Influence of core set composition on predictive abilities.
+Predictive abilities are based on single step predition with genomic data as the 
+complete predictor and transcriptomic data as the incomplete predictor.
+Results are shown for six agronomic traits and nine core sets of the maize 
+diversity panel.
+}
+\label{fig:CoreSetInfluence}
+\end{figure}
+-->
+
+<!--
+% latex table generated in R 3.3.3 by xtable 1.8-2 package
+% Wed Mar 15 10:35:35 2017
+# Script: ./analysis/maizego_plot_predictive_abilities.R
+-->
+\begin{table}[ht]
+\centering
+\caption{Predictive abilities and corresponding coefficients of variation for the set of 211 tropical/subtropical lines from the maize diversity panel.} 
+\begin{tabular}{rllllll}
+  \toprule
+ & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
+  \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: no}}\\
+snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
+  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
+   \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: yes}}\\
+mrna-none & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
+  snp-none1 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
+   \bottomrule
+\multicolumn{6}{l}{}\\
+\end{tabular}
+\end{table}
+
+<!--
+% latex table generated in R 3.3.3 by xtable 1.8-2 package
+% Wed Mar 15 11:11:37 2017
+# Script: ./analysis/uhoh_plot_predictive_abilities.R
+-->
+\begin{table}[ht]
+\centering
+\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids} 
+\begin{tabular}{rllllll}
+  \toprule
+ & DMY & DMC & FAT & PRO & STA & SUG \\ 
+  \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
+ped-none & 0.54 (0.0493) & 0.48 (0.0521) & 0.21 (0.0365) & 0.6 (0.0187) & 0.35 (0.0427) & 0.42 (0.117) \\ 
+  snp-none & 0.74 (0.0517) & 0.63 (0.0521) & 0.43 (0.0356) & 0.65 (0.0223) & 0.46 (0.0458) & 0.47 (0.1261) \\ 
+  ped-snp & 0.71 (0.054) & 0.57 (0.0545) & 0.35 (0.0375) & 0.62 (0.0235) & 0.39 (0.0478) & 0.42 (0.1298) \\ 
+  ped-mrna & 0.73 (0.056) & 0.53 (0.0594) & 0.35 (0.0398) & 0.61 (0.0254) & 0.31 (0.0504) & 0.37 (0.1396) \\ 
+  snp-mrna & 0.75 (0.0533) & 0.61 (0.0553) & 0.43 (0.0374) & 0.65 (0.0239) & 0.44 (0.048) & 0.44 (0.1342) \\ 
+   \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
+mrna-none & 0.75 (0.0552) & 0.34 (0.0619) & 0.4 (0.0412) & 0.47 (0.0293) & 0.16 (0.0496) & 0.27 (0.1447) \\ 
+  ped-none1 & 0.46 (0.0469) & 0.53 (0.05) & 0.23 (0.0368) & 0.46 (0.0234) & 0.38 (0.0391) & 0.34 (0.1316) \\ 
+  snp-none1 & 0.67 (0.0534) & 0.47 (0.0552) & 0.43 (0.0384) & 0.48 (0.027) & 0.36 (0.0457) & 0.35 (0.132) \\ 
+   \bottomrule
+\multicolumn{6}{l}{}\\
+\end{tabular}
+\end{table}
+
+<!--
+% latex table generated in R 3.3.3 by xtable 1.8-2 package
+% Wed Mar 15 11:36:08 2017
+# Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
+-->
+\begin{table}[ht]
+\centering
+\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel.} 
+\begin{tabular}{rllllll}
+  \toprule
+ & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
+  \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
+snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
+  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
+   \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
+mrna-none\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
+  snp-mrna\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
+  snp-mrna\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
+  snp-mrna\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
+  snp-mrna\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
+  snp-mrna\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
+  snp-mrna\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
+  snp-mrna\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
+  snp-none\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
+   \bottomrule
+\multicolumn{6}{l}{}\\
+\end{tabular}
+\end{table}
+
+<!--chapter:end:results.Rmd-->
+
+# Discussion
+
+<!--chapter:end:discussion.Rmd-->
+
+# References
+
+<!--chapter:end:references.Rmd-->

### 2017-03-15 reports/Paper.tex
Init LaTeX paper

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
%\makeatletter
%\def\ps@pprintTitle{%
% \let\@oddhead\@empty
% \let\@evenhead\@empty
% \def\@oddfoot{\it \hfill\today}%
% \let\@evenfoot\@oddfoot}
%\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Omics-based Single Step Trait Prediction},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)
% Pandoc header

\usepackage[nomarkers]{endfloat}

\begin{document}
\begin{frontmatter}

  \title{Omics-based Single Step Trait Prediction}
    \author[University of Hohenheim]{Matthias Westhues}
   \ead{matthias.westhues@uni-hohenheim.de}

    \author[University of Kiel]{Claas Heuer}
   \ead{cheuer@tierzucht.uni-kiel.de}

    \author[University of Kiel]{Georg Thaller}
   \ead{thaller@tierzucht.uni-kiel.de}

    \author[Iowa State University]{Rohan Fernando}
   \ead{rohan@iastate.edu}

    \author[University of Hohenheim]{Albrecht E. Melchinger\corref{c1}}
   \ead{melchinger@uni-hohenheim.de} 
   \cortext[c1]{Corresponding Author}
      \address[University of Hohenheim]{Institute of Plant Breeding, Seed Science and Population Genetics,
Fruwirthstr. 21, 70599 Stuttgart, Germany}
    \address[University of Kiel]{Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6,
24118 Kiel, Germany}
    \address[Iowa State University]{Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.}

  \begin{abstract}
  \(\dots\)
  \end{abstract}

 \end{frontmatter}

\section{Introduction}\label{introduction}

\(\dots\)

Our objectives were threefold: (1) Examine whether the single step
prediction framework can be transfered to hybrids, (2) explore the
utility of single step predition with a quantitative predictor and (3)
evaluate the impact of the particular genetic space covered by the
incomplete predictor on predictive ability.

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Data}\label{data}

\subsubsection{Hybrid data}\label{hybrid-data}

The maize hybrid data set comprises 1,521 hybrids produced in 16
factorial mating designs between 142 Dent and 103 Flint parent lines
(Westhues et al. 2017). Best linear unbiased estimates (BLUEs) were
computed across all factorials and three or more agro-ecologically
diverse locations across Germany for six agronomically important traits.
All 245 parent lines have pedigree records back to the generation of
their grandparents or deeper (Westhues et al. 2017) and were genotyped
using the Illumina BeadChip MaizeSNP50 (Ganal et al. 2011). After
filtering for minor allele frequency (\(\geq 5\)\%), heterozygosity rate
(\(\leq 5\)\%) and call frequency (\(\geq 95\)\%), missing genotypes
were imputed using the Beagle software (B. L. Browning and Browning
2009). The final number of polymorphic marker loci was 7,013 for
parental Dent and 6,212 for parental Flint lines, respectively. Gene
expression data for seedlings seven days after sowing were obtained for
a subset of 60 parental Dent and 43 parental Flint lines through
two-color hybridizations using a custom 2K microarray (GPL22267)
(Westhues et al. 2017). After applying established normalization
procedures (G. K. Smyth and Speed 2003; Matthew E Ritchie et al. 2007),
gene expression BLUEs were computed for 1,323 transcripts in reference
to established protocols (G. K. Smyth and Speed 2003; Matthew E Ritchie
et al. 2007; Frisch et al. 2010) using the R-package \emph{limma} (M. E.
Ritchie et al. 2015).

\subsubsection{Diversity panel}\label{diversity-panel}

The maize inbred line data set originally comprised 513 maize lines
representing the global maize diversity and was reduced to the set of
tropical and subtropical lines (\(n = 211\)) given that it is the
largest of the four pre-classified subgroups (N. Yang et al. 2014). All
inbred lines were evaluated in five Chinese environments ranging from
\(18\) to \(30^{\circ}\)N and from \(102\) to \(110^{\circ}\)E (N. Yang
et al. 2014). Best linear unbiased predictors (BLUPs) were calculated
for 17 traits of which six were used for this study. MaizeSNP50
BeadChip-data were available for the entire set of lines from the
diversity panel. Additionally, RNA sequencing was performed for 368
lines on young seedlings 15 days after sowing. By exploting identify by
descent (IBD) between 49,728 SNPs on the BeadChip overlapping with the
RNA-seq data, 556,809 high quality SNPs could be inferred for these
lines (Fu et al. 2013; H. Li et al. 2013). For the remaining 145 maize
lines without RNA-seq data, high density markers were inferred via
projection of IBD regions onto the BeadChip, using this core set of
markers (N. Yang et al. 2014). The same SNP quality checks as for the
parental inbred lines of the hybrids were applied to the 211 inbred
lines from the tropical/subtropical subset, yielding 37,760 SNPs. Gene
expression data were normalized using a normal quantile transformation
to satisfy modelling assumptions (Fu et al. 2013). A subset of 28,850
annotated genes was available for 149 out of the 211 genotypes and was
kept for further analyses (H. Li et al. 2013).

\subsection{Population structure}\label{population-structure}

Principal component analyses (PCA) were run on scaled and centered
predictor matrices using the \emph{pca()} function from the R-package
\emph{LEA} (Frichot and François 2015). Estimates of ancestry
coefficients were estimated based on sparse nonnegative matrix
factorization algorithms (Frichot et al. 2014) using the \emph{snmf()}
function in \emph{LEA}. Algorithms were run for \(K=2, \dots 5\)
putative ancestral gene pools with 25 repetitions for each value of
\(K\). For each \(K\), only the repetition with the lowest cross-entropy
was kept for further analyses.

\subsection{Core Sampling}\label{core-sampling}

In order to evaluate the influence of the genetic constitution of the
set of genotypes that has only data on one out of two predictors, we
generated core samples using the R-package \emph{corehunter} (Thachuk et
al. 2009). These core samples were based on the 149 maize inbred lines
from the diversity panel which had gene expression as well as genomic
information. The size of the core set was varied in increments of ten
percentage points and ranged from 10\% to 90\% of all 149 inbred lines.
For assembling the various core sets, we first computed the Modifed
Rogers (MR) distance between each pair of genotypes and then selected a
predefined subset of genotypes that display maximum average distance
between each other. MR was computed using the \emph{rogers.dist()}
function from the R-pacakge \emph{poppr} (Kamvar, Brooks, and Grünwald
2015).

\subsection{Prediction}\label{prediction}

\subsubsection{Kernels}\label{kernels}

BLUP models are computationally attractive when the number of features
\(p\) exceeds the number of genotypes \(n\) and are equivalent to a
selection index when fixed effects have been accomodated by the
dependent variable (Mrode 2014). Depending on the data set, up to three
predictors were available for agronomic trait predictions, namely
pedigree data (P), genomic data (G) and transcriptompic data (T). The
corresponding feature matrix of the \(l\)-th group of inbred lines and
the \(o\)-th predictor (\(\mathbf{W}_{lo}\)) has dimensions
\(n \times p\) where \(n\) pertains to the number of genotypes in the
\(l\)-th group and \(p\) pertains to the number of features. For the
diversity panel maize lines, \(n\) corresponds to the number of
genotypes whereas, in the case of the hybrid data, \(n\) corresponds to
the number of parent lines in the corresponding heterotic group. In
generating kernels for each predictor, all features in
\(\mathbf{W}_{lo}\) were centered and standardized to unit variance.
Then, each kernel can be defined as

\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{K}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
\end{equation}

where \(W_{lo}\) denotes the number of features (VanRaden 2008). In the
case of pedigree data (P), coancestry coefficients were used directly
for \(\mathbf{K}_{lo}\).

\subsubsection{Models}\label{models}

The universal model for breeding values was:

\begin{equation} \label{eq:KBLUPModel}
  \mathbf{y} = \mu + 
  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
  \mathbf{\epsilon},
\end{equation}

where \(\mathbf{y}\) is the vector of observed inbred line or hybrid
performance, respectively, \(\mu\) is the fixed model intercept,
\(\mathbf{Z}_{lo}\) is the corresponding design matrix associating the
random genotype effects (\emph{i.e.}, breeding values) of the lines in
the \(l\)-th inbred line group and the \(o\)-th predictor type with
\(\mathbf{y}\). Note that, in the case of hybrid data,
\(\mathbf{Z}_{lo}\) associates the random general combining ability
effects in the parental inbred lines (either Dent or Flint, as expressed
by \(l\)) with the vector of observed hybrid performance \(\mathbf{y}\).
The random effects (\(\mathbf{g}_{lo}\)) have expectation zero and
variance equal to \(\mathbf{K}_{lo} \sigma^{2}_{{g}^{lo}}\) and
\(\mathbf{I} \sigma^2_{\epsilon}\) for the residual error. All
prediction models were implemented using the R-package \emph{BGLR}
(Pérez and de Los Campos 2014).

\subsubsection{Single Step Prediction}\label{single-step-prediction}

The breeding values in a prediction model are in general given by:

\begin{equation} \label{eq:mrnaebv}
    \mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

where \(\boldsymbol{\hat{\alpha}}\) is the solution to a ridge
regression model, which is equivalent to the BLUP model.

Consider now the situation in which one predictor is complete in the
sense that it contains information on the whole population whereas a
second predictor is incomplete and covers only a subset of the
population. We can take a similar route as in Fernando, Dekkers, and
Garrick (2014) and impute covariates of the incomplete predictor by
using covariates of the complete predictor. Let the subscript \(1\)
denote individuals covered only by the complete predictor whereas
individuals covered by both, the complete and the incomplete predictor,
are indicated by the subscript \(2\). Further, the covariates in
\(\mathbf{W}\) are centered. The vector \(\mathbf{g}_1\) can be written
as the sum of the conditional expectation given \(\mathbf{g}_2\) and a
residual error term:

\begin{align} \label{eq:mrna1}
    \mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
    &= \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
    &= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of \(\boldsymbol{\epsilon}\) is
\(\mathbf{K}_{11} - \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{K}_{21} = (\mathbf{K}^{11})^{-1}\)
(Legarra, Aguilar, and Misztal 2009).

The structure of the residual imputation/prediction error is known and
can therefore be modelled. The covariates not covered by the incomplete
predictor can be predicted using the expectation of a multivariate
normal random vector given correlated observations, which are related to
the Best Linear Predictor.

The general model for the single step procedure can then be written as

\begin{equation}
  \label{eq:single-step-model}
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}^{Complete} \boldsymbol{\epsilon}^{Complete} + \mathbf{e},
\end{equation}

with

\begin{equation}
 \label{eq:single-step-submatrices}
\mathbf{X} = 
\begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}^{Complete} = 
\begin{bmatrix}
  Z_1 \\
  0 
 \end{bmatrix}
\end{equation}

Following the notation in Fernando, Dekkers, and Garrick (2014) the
phenotypes can be untangled like this:

\begin{align}
 \label{eq:entangled-augmented-single-step-model}
\begin{bmatrix}
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

The breeding values from that model are:

\begin{equation}
 \label{eq:breeding-values}
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}^{Complete} 
\end{equation}

Our final model looks like this:

\begin{align}
 \label{eq:final-model}
\mathbf{y} &= \mathbf{Xb} +
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l}^{Complete} \boldsymbol{\epsilon}_{l}^{Complete} +
\mathbf{e}.
\end{align}

For the hybrid data, the obtained GCA effects for the inbred lines
represent half their breeding values. The predicted agronomic
performance is:

\begin{equation}
 \label{eq:predicted-performance}
\hat{\mathbf{y}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
\end{equation}

\subsubsection{Predictive ability and model
validation}\label{predictive-ability-and-model-validation}

For the validation of our predictions, we employed leave-one-out
cross-validation (LOOCV) routines. In the case of the diversity panel
maize inbred lines, LOOCV was performed by using a single genotype as a
hold-out sample, which will subsequently be predicted, while using all
other inbred lines for model training. This process is repeated until
all inbred lines have been used once for testing and \(n - 1\) times for
model training.

For the hybrid data, LOOCV was carried out as follows: Let \(D\) and
\(F\) denote the set of parental inbred lines from the Dent and the
Flint group, respectively. Further, let \(H \cap [D \times F]\) denote
the set of hybrids from crosses between \(D\) and \(F\). The training
set (\(H_{TRN}\)) for the hybrid to be predicted (\(H_{ij}\)) --- where
\(i \in D_{TRN}\) and \(j \in F_{TRN}\) --- was assembled as
\(H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]\) with
\(D_{TRN}^{C} = D \setminus D_{TRN}\) and
\(F_{TRN}^{C} = F \setminus F_{TRN}\).

We judged the performance of each model by looking at its predictive
ability, which is calculated as \(\rho(\mathbf{y}, \mathbf{\hat{y}})\),
where \(\mathbf{\hat{y}}\) is the vector of predicted values from each
LOOCV run. Our confidence in the performance of each model was evalutaed
by computing the coefficient of variation for the predicted values:

\begin{equation}
  \label{eq:CV}
CV = \sqrt{\frac{\sum_{i = 1}^{n} \sigma^{2}_{\hat{y}(i)} / n}{\sum_{i=1}^{n}\hat{y}_{i} / n}},
\end{equation}

where \(n\) denotes the number of genotypes (\emph{i.e.}, inbred lines
or hybrids).

\section{Results}\label{results}

\begin{figure}[htbp]
\centering
\includegraphics{"./tabs_figs/core_fraction_predictive_ability_trend.pdf"}
\caption{Bullshit}
\end{figure}

\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for the set of 211 tropical/subtropical lines from the maize diversity panel.} 
\begin{tabular}{rllllll}
  \toprule
 & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: no}}\\
snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: yes}}\\
mrna-none & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
  snp-none1 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids} 
\begin{tabular}{rllllll}
  \toprule
 & DMY & DMC & FAT & PRO & STA & SUG \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
ped-none & 0.54 (0.0493) & 0.48 (0.0521) & 0.21 (0.0365) & 0.6 (0.0187) & 0.35 (0.0427) & 0.42 (0.117) \\ 
  snp-none & 0.74 (0.0517) & 0.63 (0.0521) & 0.43 (0.0356) & 0.65 (0.0223) & 0.46 (0.0458) & 0.47 (0.1261) \\ 
  ped-snp & 0.71 (0.054) & 0.57 (0.0545) & 0.35 (0.0375) & 0.62 (0.0235) & 0.39 (0.0478) & 0.42 (0.1298) \\ 
  ped-mrna & 0.73 (0.056) & 0.53 (0.0594) & 0.35 (0.0398) & 0.61 (0.0254) & 0.31 (0.0504) & 0.37 (0.1396) \\ 
  snp-mrna & 0.75 (0.0533) & 0.61 (0.0553) & 0.43 (0.0374) & 0.65 (0.0239) & 0.44 (0.048) & 0.44 (0.1342) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
mrna-none & 0.75 (0.0552) & 0.34 (0.0619) & 0.4 (0.0412) & 0.47 (0.0293) & 0.16 (0.0496) & 0.27 (0.1447) \\ 
  ped-none1 & 0.46 (0.0469) & 0.53 (0.05) & 0.23 (0.0368) & 0.46 (0.0234) & 0.38 (0.0391) & 0.34 (0.1316) \\ 
  snp-none1 & 0.67 (0.0534) & 0.47 (0.0552) & 0.43 (0.0384) & 0.48 (0.027) & 0.36 (0.0457) & 0.35 (0.132) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel.} 
\begin{tabular}{rllllll}
  \toprule
 & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
mrna-none\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
  snp-mrna\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
  snp-mrna\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
  snp-mrna\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
  snp-mrna\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
  snp-mrna\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
  snp-mrna\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
  snp-mrna\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
  snp-none\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

\section{Discussion}\label{discussion}

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Browning2009}{}
Browning, Brian L, and Sharon R Browning. 2009. ``A unified approach to
genotype imputation and haplotype-phase inference for large data sets of
trios and unrelated individuals.'' \emph{Am. J. Hum. Genet.} 84 (2). The
American Society of Human Genetics: 210--23.
doi:\href{https://doi.org/10.1016/j.ajhg.2009.01.005}{10.1016/j.ajhg.2009.01.005}.

\hypertarget{ref-Fernando2014}{}
Fernando, Rohan L, Jack Cm Dekkers, and Dorian J Garrick. 2014. ``A
class of Bayesian methods to combine large numbers of genotyped and
non-genotyped animals for whole-genome analyses.'' \emph{Genetics,
Selection, Evolution : GSE} 46 (1): 50.
doi:\href{https://doi.org/10.1186/1297-9686-46-50}{10.1186/1297-9686-46-50}.

\hypertarget{ref-Frichot2015}{}
Frichot, Eric, and Olivier François. 2015. ``LEA: An R package for
landscape and ecological association studies.'' \emph{Methods in Ecology
and Evolution} 6 (8): 925--29.
doi:\href{https://doi.org/10.1111/2041-210X.12382}{10.1111/2041-210X.12382}.

\hypertarget{ref-Frichot2014}{}
Frichot, Eric, François Mathieu, Théo Trouillon, Guillaume Bouchard, and
Olivier François. 2014. ``Fast and efficient estimation of individual
ancestry coefficients.'' \emph{Genetics} 196 (4): 973--83.
doi:\href{https://doi.org/10.1534/genetics.113.160572}{10.1534/genetics.113.160572}.

\hypertarget{ref-Frisch2010}{}
Frisch, Matthias, Alexander Thiemann, Junjie Fu, Tobias a. Schrag,
Stefan Scholten, and Albrecht E. Melchinger. 2010. ``Transcriptome-based
distance measures for grouping of germplasm and prediction of hybrid
performance in maize.'' \emph{Theor. Appl. Genet.} 120 (2): 441--50.
doi:\href{https://doi.org/10.1007/s00122-009-1204-1}{10.1007/s00122-009-1204-1}.

\hypertarget{ref-Fu2013}{}
Fu, Junjie, Yanbing Cheng, Jingjing Linghu, Xiaohong Yang, Lin Kang,
Zuxin Zhang, Jie Zhang, et al. 2013. ``RNA sequencing reveals the
complex regulatory network in the maize kernel.'' \emph{Nat. Comm.} 4.
Nature Publishing Group: 2832.
doi:\href{https://doi.org/10.1038/ncomms3832}{10.1038/ncomms3832}.

\hypertarget{ref-Ganal2011}{}
Ganal, Martin W, Gregor Durstewitz, Andreas Polley, Aurélie Bérard,
Edward S Buckler, Alain Charcosset, Joseph D Clarke, et al. 2011. ``A
large maize (Zea mays L.) SNP genotyping array: development and
germplasm genotyping, and genetic mapping to compare with the B73
reference genome.'' \emph{PloS ONE} 6 (12): e28334.
doi:\href{https://doi.org/10.1371/journal.pone.0028334}{10.1371/journal.pone.0028334}.

\hypertarget{ref-Kamvar2015}{}
Kamvar, Zhian N., Jonah C. Brooks, and Niklaus J. Grünwald. 2015.
``Novel R tools for analysis of genome-wide population genetic data with
emphasis on clonality.'' \emph{Frontiers in Genetics} 6: 1--10.
doi:\href{https://doi.org/10.3389/fgene.2015.00208}{10.3389/fgene.2015.00208}.

\hypertarget{ref-Legarra2009}{}
Legarra, a, I Aguilar, and I Misztal. 2009. ``A relationship matrix
including full pedigree and genomic information.'' \emph{Journal of
Dairy Science} 92 (9). Elsevier: 4656--63.
doi:\href{https://doi.org/10.3168/jds.2009-2061}{10.3168/jds.2009-2061}.

\hypertarget{ref-Li2013}{}
Li, Hui, Zhiyu Peng, Xiaohong Yang, Weidong Wang, Junjie Fu, Jianhua
Wang, Yingjia Han, et al. 2013. ``Genome-wide association study dissects
the genetic architecture of oil biosynthesis in maize kernels.''
\emph{Nat. Genet.} 45 (1). Nature Publishing Group: 43--50.
doi:\href{https://doi.org/10.1038/ng.2484}{10.1038/ng.2484}.

\hypertarget{ref-Mrode2014}{}
Mrode, Raphael A. 2014. \emph{Linear Models for the Prediction of Animal
Breeding Values}. 3rd ed. Oxfordshire: CABI.
doi:\href{https://doi.org/10.1017/CBO9781107415324.004}{10.1017/CBO9781107415324.004}.

\hypertarget{ref-Perez2014}{}
Pérez, Paulino, and Gustavo de Los Campos. 2014. ``Genome-wide
regression \& prediction with the BGLR statistical package.''
\emph{Genetics} 198 (October): 483--95.
doi:\href{https://doi.org/10.1534/genetics.114.164442}{10.1534/genetics.114.164442}.

\hypertarget{ref-Ritchie2015a}{}
Ritchie, M. E., B. Phipson, D. Wu, Y. Hu, C. W. Law, W. Shi, and G. K.
Smyth. 2015. ``Limma powers differential expression analyses for
RNA-sequencing and microarray studies.'' \emph{Nucleic Acids Res.} 43
(7): e47.
doi:\href{https://doi.org/10.1093/nar/gkv007}{10.1093/nar/gkv007}.

\hypertarget{ref-Ritchie2007}{}
Ritchie, Matthew E, Jeremy Silver, Alicia Oshlack, Melissa Holmes,
Dileepa Diyagama, Andrew Holloway, and Gordon K Smyth. 2007. ``A
comparison of background correction methods for two-colour
microarrays.'' \emph{Bioinformatics} 23 (20): 2700--2707.
doi:\href{https://doi.org/10.1093/bioinformatics/btm412}{10.1093/bioinformatics/btm412}.

\hypertarget{ref-Smyth2003}{}
Smyth, Gordon K, and Terry Speed. 2003. ``Normalization of cDNA
microarray data.'' \emph{Methods} 31 (4): 265--73.
doi:\href{https://doi.org/10.1016/S1046-2023(03)00155-5}{10.1016/S1046-2023(03)00155-5}.

\hypertarget{ref-Thachuk2009}{}
Thachuk, Chris, José Crossa, Jorge Franco, Susanne Dreisigacker, Marilyn
Warburton, and Guy F Davenport. 2009. ``Core Hunter: an algorithm for
sampling genetic resources based on multiple genetic measures.''
\emph{BMC Bioinformatics} 10 (243): 1--13.
doi:\href{https://doi.org/10.1186/1471-2105-10-243}{10.1186/1471-2105-10-243}.

\hypertarget{ref-VanRaden2008}{}
VanRaden, P M. 2008. ``Efficient methods to compute genomic
predictions.'' \emph{J. Dairy Sci.} 91 (11). Elsevier: 4414--23.
doi:\href{https://doi.org/10.3168/jds.2007-0980}{10.3168/jds.2007-0980}.

\hypertarget{ref-Westhues2017}{}
Westhues, Matthias, Tobias A Schrag, Claas Heuer, Georg Thaller, H.F.
Utz, Wolfgang Schipprack, Alexander Thiemann, et al. 2017. ``Omics-based
Hybrid Prediction in Maize.''

\hypertarget{ref-Yang2014}{}
Yang, Ning, Yanli Lu, Xiaohong Yang, Juan Huang, Yang Zhou, Farhan Ali,
Weiwei Wen, Jie Liu, Jiansheng Li, and Jianbing Yan. 2014. ``Genome Wide
Association Studies Using a New Nonparametric Model Reveal the Genetic
Architecture of 17 Agronomic Traits in an Enlarged Maize Association
Panel.'' \emph{PLoS Genet.} 10 (9).
doi:\href{https://doi.org/10.1371/journal.pgen.1004573}{10.1371/journal.pgen.1004573}.

\end{document}

### 2017-03-15 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -1,7 +1,8 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
-               "stringr", "forcats")
+pacman::p_load("tidyverse", "data.table", "dtplyr", "xtable", "stringr",
+               "forcats", "ggthemes", "viridis")
+source("./software/prediction_helper_functions.R")
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
 prediction_path <- "./data/derived/predictions/"
@@ -31,39 +32,73 @@ pred_df <- filtered_log_file %>%
   droplevels() %>%
   select(-Job_ID)
   
-pred_df %>%
-  group_by(Predictor, Core_Fraction) %>%
-  count()
   
 
-## Predictive ability for seven agronomic traits (rows) and nine predictor sets 
-# (columns). The predictor sets are split into the categories 'Reduced' 
-# (i.e. 685 hybrids) and 'Full' (i.e. 1,521 hybrids)."
+pred_lst <- pred_df %>%
+  mutate(
+    Reduced = if_else(Core_Fraction == "Full", true = FALSE, false = TRUE)
+  ) %>%
+  group_by(Trait, Predictor, Core_Fraction, Reduced) %>%
+  summarize(
+    r = cor(y, yhat),
+    CV = coefficient_of_variation(var_yhat, yhat)
+  ) %>%
+  ungroup() %>%
+  mutate(
+    r = round(r, digits = 2),
+    CV = round(CV, digits = 4),
+    Value = paste0(
+      r, " (", CV, ")"
+    )
+  ) %>%
+  select(-r, -CV) %>%
+  spread(key = Trait, value = Value) %>%
+  split(.$Reduced) %>%
+  map(., ~select(., -Reduced)) %>%
+  map_at("TRUE", .f = ~mutate(
+    ., Predictor = paste0(Predictor, "_", Core_Fraction)
+  )) %>%
+  map(., ~select(., -Core_Fraction)) %>%
+  map(., ~column_to_rownames(., var = "Predictor"))
+
+# Prepare the split data for being displayed inside a LaTeX table.
+names(pred_lst)
+coverage_subheading <- "Only incomplete predictor:"
+attr(pred_lst, "subheadings") <- paste(coverage_subheading, names(pred_lst))
+core_caption <- paste(
+  "Predictive abilities and corresponding coefficients of variation for",
[…]

### 2017-03-15 analysis/maizego_plot_predictive_abilities.R

@@ -1,12 +1,19 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
-               "stringr", "forcats")
+pacman::p_load("tidyverse", "data.table", "dtplyr", "stringr", "forcats",
+               "xtable")
+source("./software/prediction_helper_functions.R")
+
+# xtable options
+options(xtable.floating = FALSE)
+options(xtable.timestamp = "")
+options(width = 60)
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
 prediction_path <- "./data/derived/predictions/"
 log_file <- fread(log_path)
 
+# Specify which prediction results shall be loaded.
 filtered_log_file <- log_file %>%
   filter(Runs == "",
          Data_Type == "Inbred") %>%
@@ -19,6 +26,8 @@ filtered_log_file <- log_file %>%
          Reduced = if_else(Pred1 == "mrna", true = "yes", false = Reduced)
   )
 
+# Read the results from the prediction models based on the log files and 
+# concatenate them in a common data frame.
 pred_df <- filtered_log_file %>%
   select(Job_ID) %>%
   flatten_chr() %>%
@@ -31,72 +40,50 @@ pred_df <- filtered_log_file %>%
   unite(col = Predictor, Pred1, Pred2, sep = "-") %>%
   droplevels() %>%
   select(-Job_ID)
-  
-  
-
-## Distribution of predicted and observed values
-legend_labs <- expression(y, hat(y))
-g1 <- pred_df %>%
-  unite(col = "Unique_Predictor", Reduced, Predictor, sep = "_") %>%
-  ggplot(aes(x = y, y = yhat)) +
-  geom_point() +
-  facet_wrap(Unique_Predictor ~ Trait, scales = "free") +
-  theme_pander() 
-
-## Predictive ability for seven agronomic traits (rows) and nine predictor sets 
-# (columns). The predictor sets are split into the categories 'Reduced' 
-# (i.e. 685 hybrids) and 'Full' (i.e. 1,521 hybrids)."
-g2 <- pred_df %>%
-  group_by(Trait, Predictor, Reduced) %>%
-  summarize(r = cor(y, yhat)) %>%
-  ggplot(aes(x = Predictor, y = r)) +
[…]

### 2017-03-15 analysis/uhoh_plot_predictive_abilities.R

@@ -1,7 +1,8 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
-               "stringr", "forcats")
+pacman::p_load("tidyverse", "data.table", "dtplyr", "xtable", "stringr",
+               "forcats")
+source("./software/prediction_helper_functions.R")
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
 prediction_path <- "./data/derived/predictions/"
@@ -41,80 +42,47 @@ raw_pred_df <- log_file %>%
   select(-Job_ID)
   
   
-pred_df <- raw_pred_df %>%
+pred_lst <- raw_pred_df %>%
   mutate(
     Reduced = ifelse(Predictor == "mrna-none", yes = TRUE, no = FALSE),
     Reduced = ifelse(Core_Fraction == "1.0", yes = TRUE, no = Reduced) 
   ) %>%
-  select(-Core_Fraction)
-
-
-
-## Distribution of predicted and observed values
-legend_labs <- expression(y, hat(y))
-g1 <- pred_df %>%
-  ggplot(aes(x = y, y = yhat)) +
-  geom_point() +
-  facet_wrap(Predictor ~ Trait, scales = "free") +
-  theme_pander() 
-
-## Predictive ability for seven agronomic traits (rows) and nine predictor sets 
-# (columns). The predictor sets are split into the categories 'Reduced' 
-# (i.e. 685 hybrids) and 'Full' (i.e. 1,521 hybrids)."
-g2 <- pred_df %>%
+  select(-Core_Fraction) %>%
   group_by(Trait, Predictor, Reduced) %>%
-  summarize(r = cor(y, yhat)) %>%
-  mutate(Predictor = fct_relevel(Predictor, pred_order)
+  summarize(
+    r = cor(y, yhat),
+    CV = coefficient_of_variation(var_yhat, yhat)
   ) %>%
-  ggplot(aes(x = Predictor, y = r)) +
-  geom_bar(stat = "identity", aes(fill = Predictor), color = "black") +
-  geom_text(aes(label = round(r, digits = 3)), vjust = -0.2) +
-  scale_fill_viridis(discrete = TRUE, option = "inferno") +
-  facet_grid(Trait ~ Reduced, space = "free", scales = "free_x") +
-  theme_bw() +
-  theme(
[…]

### 2017-03-15 analysis/uhoh_snp_analyses.R

@@ -1,0 +1,121 @@
+# Data and packages -------------------------------------------------------
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load("tidyverse", "LEA", "stringr", "viridis")
+
+
+snp <- readRDS("./data/processed/maizego/imputed_snp_mat.RDS")
+unique_genotypes <- "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS" %>%
+  readRDS() %>%
+  keep(names(.) == "snp") %>%
+  flatten_chr()
+
+
+# PCA ---------------------------------------------------------------------
+write.lfmm(snp, "./data/derived/maizego/snp.lfmm")
+
+# Determine the structure of the data using genotypic data.
+snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
+                   scale = TRUE)
+# Perform Tracy-Widom tests on all eigenvalues to determine the optimal number
+# of principal components.
+snp_tw <- tracy.widom(snp_pc)
+tw_K <- snp_tw %>%
+  .["pvalues"] %>%
+  flatten_dbl() %>%
+  keep(~ .x <= 0.0001) %>%
+  length()
+tw_K <- min(c(tw_K, 5))
+
+snp_tw %>%
+  .["percentage"] %>%
+  flatten_dbl() %>%
+  plot()
+
+# Qualitatively assign the genotypes to the previously determined 
+# subpopulations.
+lfmm2geno("./data/derived/maizego/snp.lfmm", 
+          output.file = "./data/derived/maizego/snp.geno")
+project <- snmf(input.file = "./data/derived/maizego/snp.geno",
+                K = seq_len(tw_K),
+                project = "new",
+                repetitions = 25,
+                CPU = 3,
+                entropy = TRUE)
+
+# Based on the results of a PCA, decide whether it is even necessary to treat
+# subpopulations as such or whether all inbred lines can be treated as a single
+# group.
+project <- load.snmfProject("./data/derived/maizego/snp.snmfProject")
+
+
+# Get labels for the different subpopulations of the genotypes.
+pop_struc <- fread("./data/input/maizego/genotype_annotation.txt")
+
+pc_mat <- snp_pc$projections
+rownames(pc_mat) <- rownames(snp)
+pc_with_assignment <- pc_mat %>%
[…]

### 2017-03-15 software/prediction_helper_functions.R

@@ -41,3 +41,10 @@ create_named_list <- function(...) {
     if (any(nonames <- nm == "")) nm[nonames] <- snm[nonames]
     setNames(L, nm)
 }
+
+
+
+# Function for the computation of the coefficient of variation for each scenario.
+coefficient_of_variation <- function(x, y) {
+  sqrt(mean(x)) / mean(y)
+}

### 2017-03-15 analysis/maizego_corehunter.R
Move Maizego SNP vs. mRNA PCA to general SNP analysis

@@ -115,54 +115,11 @@ pc_df <- pc_mat %>%
   rownames_to_column(var = "G") %>%
   gather(key = PC, value = Score, -G) %>%
   as_data_frame() %>%
-  filter(PC %in% paste0("PC_", seq_len(3))) %>%
+  filter(PC %in% paste0("PC_", seq_len(2))) %>%
   left_join(y = group_membership, by = "G") %>%
-  spread(key = PC, value = Score)
-
-
-pc_df %>%
-  ggplot(aes(x = PC_1, y = PC_2, color = Core_Group, shape = Core_Group)) +
-  geom_point() +
-  ggthemes::scale_color_tableau() +
-  facet_wrap(~Fraction) +
-  theme_bw()
-
-
-
-# SNP VS MRNA PCA ---------------------------------------------------------
-# Explore the kinship matrix.
-all_geno_snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
-  readRDS() %>%
-  sspredr::ensure_snp_quality(
-    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
-    any_missing = FALSE, remove_duplicated = TRUE
-  )
-
-write.lfmm(all_geno_snp, "./data/derived/maizego/all_tst_snp.lfmm")
-all_geno_pc <- LEA::pca("./data/derived/maizego/all_tst_snp.lfmm", scale = TRUE)
-
-all_geno_pc <- all_geno_pc$projections
-rownames(all_geno_pc) <- rownames(all_geno_snp)
-colnames(all_geno_pc) <- paste0("PC_", seq_len(ncol(all_geno_pc)))
-
-all_geno_pc %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "G") %>%
-  gather(key = PC, value = Score, -G) %>%
-  as_data_frame() %>%
-  filter(PC %in% paste0("PC_", seq_len(3))) %>%
-  mutate(Group = if_else(
-    G %in% common_genotypes, true = "mRNA", false = "All")
+  mutate(
+    PC = gsub("PC_", replacement = "PC", x = PC)
   ) %>%
-  spread(key = PC, value = Score) %>%
-  ggplot(aes(x = PC_1, y = PC_2, color = Group, shape = Group)) +
-  geom_point() +
-  scale_color_tableau() +
-  theme_bw() +
-  theme(legend.position = "top")
-
-
-
[…]

### 2017-03-15 analysis/maizego_snp_analyses.R

@@ -1,6 +1,7 @@
 # Data and packages -------------------------------------------------------
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "LEA", "stringr", "viridis")
+pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "cowplot", 
+               "ggthemes")
 
 
 snp <- readRDS("./data/processed/maizego/imputed_snp_mat.RDS")
@@ -48,31 +49,52 @@ project <- snmf(input.file = "./data/derived/maizego/snp.geno",
 project <- load.snmfProject("./data/derived/maizego/snp.snmfProject")
 
 
-# Get labels for the different subpopulations of the genotypes.
-pop_struc <- fread("./data/input/maizego/genotype_annotation.txt")
+# SNP VS MRNA PCA ---------------------------------------------------------
+# For the second scenario, keep only names of genotypes, which are covered by 
+# all data types (i.e. phenotypic, genotypic and transcriptomic).
+common_genotypes <- readRDS(
+  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
+)
+common_genotypes <- common_genotypes %>%
+  reduce(intersect)
 
-pc_mat <- snp_pc$projections
-rownames(pc_mat) <- rownames(snp)
-pc_with_assignment <- pc_mat %>%
+# Explore the kinship matrix.
+all_geno_snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
+  readRDS() %>%
+  sspredr::ensure_snp_quality(
+    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
+    any_missing = FALSE, remove_duplicated = TRUE
+  )
+
+write.lfmm(all_geno_snp, "./data/derived/maizego/all_tst_snp.lfmm")
+all_geno_pc <- LEA::pca("./data/derived/maizego/all_tst_snp.lfmm", scale = TRUE)
+
+all_geno_pc <- all_geno_pc$projections
+rownames(all_geno_pc) <- rownames(all_geno_snp)
+colnames(all_geno_pc) <- paste0("PC_", seq_len(ncol(all_geno_pc)))
+
+g1 <- all_geno_pc %>%
   as.data.frame() %>%
-  rownames_to_column(var = "Lines") %>%
[…]

### 2017-03-15 analysis/maizego_plot_predictive_abilities.R
Abbreviate predictor names in tables

@@ -1,12 +1,10 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "data.table", "dtplyr", "stringr", "forcats",
-               "xtable")
+               "xtable", "gsubfn")
 source("./software/prediction_helper_functions.R")
 
 # xtable options
-options(xtable.floating = FALSE)
-options(xtable.timestamp = "")
 options(width = 60)
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
@@ -22,8 +20,8 @@ filtered_log_file <- log_file %>%
   ) %>%
   mutate(Core_Fraction = as.numeric(Core_Fraction)) %>%
   filter(Core_Fraction >= 1) %>%
-  mutate(Reduced = if_else(Core_Fraction == 1, true = "yes", false = "no"),
-         Reduced = if_else(Pred1 == "mrna", true = "yes", false = Reduced)
+  mutate(Reduced = if_else(Core_Fraction == 1, true = TRUE, false = FALSE),
+         Reduced = if_else(Pred1 == "mrna", true = TRUE, false = Reduced)
   )
 
 # Read the results from the prediction models based on the log files and 
@@ -40,6 +38,20 @@ pred_df <- filtered_log_file %>%
   unite(col = Predictor, Pred1, Pred2, sep = "-") %>%
   droplevels() %>%
   select(-Job_ID)
+
+# Replace long predictor names by short, one-character versions.
+pred_df <- pred_df %>%
+  mutate(
+    Predictor = gsubfn(
+      pattern = "\\S+", 
+      replacement = list(
+        "snp-none" = "G",
+        "mrna-none" = "T",
+        "snp-mrna" = "GT"
+      ),
+      x = Predictor
+    )
+  )
 
 
 # For each combination of trait, genotype coverage and predictor, compute the 
@@ -58,7 +70,7 @@ pred_lst <- pred_df %>%
   ungroup() %>%
   mutate(
     r = round(r, digits = 2),
-    CV = round(CV, digits = 4),
+    CV = round(CV, digits = 2),
     Value = paste0(
       r, " (", CV, ")"
   )) %>%

### 2017-03-15 analysis/uhoh_plot_predictive_abilities.R

@@ -1,16 +1,21 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "data.table", "dtplyr", "xtable", "stringr",
-               "forcats")
+               "forcats", "gsubfn")
+# xtable options
+options(width = 60)
+
 source("./software/prediction_helper_functions.R")
+
+
+
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
 prediction_path <- "./data/derived/predictions/"
 log_file <- fread(log_path)
 
 # Specify a unified predictor order.
-pred_order <- c("mrna-none", "ped-none", "snp-none",
-                "ped-snp", "ped-mrna", "snp-mrna")
+pred_order <- c("T", "P", "G", "PG", "PT", "GT")
 
 # Specify a unique trait order.
 trait_order <- c("DMY", "DMC", "FAT", "PRO", "STA", "SUG")
@@ -27,7 +32,6 @@ raw_pred_df <- log_file %>%
     ) %>%
   unite(col = Predictor, Pred1, Pred2, sep = "-") %>%
   filter(Trait != "ADL") %>%
-  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
   droplevels() %>%
   mutate(Trait = fct_recode(Trait,
     "DMY" = "GTM",
@@ -36,11 +40,29 @@ raw_pred_df <- log_file %>%
     "PRO" = "RPR",
     "SUG" = "XZ"
   )) %>%
-  mutate(Trait = fct_relevel(Trait, trait_order),
-         Predictor = as.factor(Predictor)
+  mutate(
+    Trait = fct_relevel(Trait, trait_order)
   ) %>%
   select(-Job_ID)
-  
+
+# Replace long predictor names by short, one-character versions.
+raw_pred_df <- raw_pred_df %>%
+  mutate(
+    Predictor = gsubfn(
+      pattern = "\\S+", 
+      replacement = list(
+        "mrna-none" = "T",
+        "ped-none" = "P",
+        "snp-none" = "G",
+        "ped-snp" = "PG",
+        "ped-mrna" = "PT",
+        "snp-mrna" = "GT"
+      ),
+      x = Predictor
+    ),
+    Predictor = fct_relevel(Predictor, pred_order)
+  )
+ 
   
 pred_lst <- raw_pred_df %>%
   mutate(

### 2017-03-15 paper/manuscript.tex
Update tables and acknowledgments

@@
 \section{Materials and Methods}
-\subsection{Maize hybrids}
+\Genetics2level{Maize hybrids}
 The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
@@
 
-\subsection{Maize diversity panel}
+\Genetics2level{Maize diversity panel}
 The maize inbred line data set originally comprised 513 maize lines representing 
@@
 
-\subsection{Population structure}
+\Genetics2level{Population structure}
 Principal component analyses (PCA) were run on scaled and centered predictor 
@@
 
-\subsection{Core sampling}
+\Genetics2level{Core sampling}
 In order to evaluate the influence of the genetic constitution of the set of
@@
 
-\subsection{Prediction}
-\subsubsection{Kernels}
+\Genetics2level{Prediction}
+\subsection{Kernels}
 BLUP models are computationally attractive when the number of features $p$
@@
 
-\subsubsection{Universal breeding value model}
+\subsection{Universal breeding value model}
 The universal model for breeding values was:
@@
 
-\subsubsection{Single step prediction}
+\subsection{Single step prediction}
 The breeding values in a prediction model are in general given by:
@@
 
-\subsubsection{Predictive ability and model validation}
+\subsection{Predictive ability and model validation}
 For the validation of our predictions, we employed leave-one-out
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/uhoh_dent_flint_pca.pdf}
+\caption{
+  Principal component (PC) analysis of the two heterotic groups of the maize 
+  hybrids.
+  All inbred lines, which are covered by genomic and transcriptomic features
+  are depicted as orange triangles whereas genotypes for which only genomic
+  information are available are depicted as blue dots.
+}
+\label{fig:UHOH-PCA}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/maizego_admixture_pca.pdf}
+\caption{
+  (\textbf{A}) Ancestry coefficients for 211 tropical/subtropical inbred lines
+  from the maize diversity panel considering $K=2 \dots 5$ ancestral
+  populations.
+  (\textbf{B}) Principal component analysis for nine different core sets of the
+  set of tropical/subtropical maize inbred lines.
+  For each size of the core set, its members are plotted as orange dots whereas 
+  all genotypes that are not elements of the cores sets (\textit{i.e.},
+  'Complement') are depicted as blue dots.
+}
+\label{fig:PopStructure}
+\end{figure}
+
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 10:35:35 2017
+% Wed Mar 15 17:33:34 2017
 % Script: ./analysis/maizego_plot_predictive_abilities.R
@@
   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: no}}\\
-snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
-  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
+G & 0.43 (0.07) & 0.37 (0.17) & 0.39 (0.05) & 0.34 (0.05) & 0.45 (0.08) & 0.53 (0.03) \\ 
+  GT & 0.44 (0.07) & 0.33 (0.17) & 0.4 (0.05) & 0.34 (0.04) & 0.4 (0.07) & 0.53 (0.03) \\ 
    \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: yes}}\\
-mrna-none & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
-  snp-none1 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
+G & 0.36 (0.08) & 0.21 (0.17) & 0.3 (0.05) & 0.34 (0.05) & 0.41 (0.08) & 0.45 (0.03) \\ 
+  T & 0.4 (0.07) & 0.15 (0.16) & 0.37 (0.05) & 0.36 (0.04) & 0.32 (0.07) & 0.49 (0.03) \\ 
    \bottomrule
@@
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 11:11:37 2017
+% Wed Mar 15 17:31:28 2017
 % Script: ./analysis/uhoh_plot_predictive_abilities.R
@@
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-ped-none & 0.54 (0.0493) & 0.48 (0.0521) & 0.21 (0.0365) & 0.6 (0.0187) & 0.35 (0.0427) & 0.42 (0.117) \\ 
-  snp-none & 0.74 (0.0517) & 0.63 (0.0521) & 0.43 (0.0356) & 0.65 (0.0223) & 0.46 (0.0458) & 0.47 (0.1261) \\ 
-  ped-snp & 0.71 (0.054) & 0.57 (0.0545) & 0.35 (0.0375) & 0.62 (0.0235) & 0.39 (0.0478) & 0.42 (0.1298) \\ 
-  ped-mrna & 0.73 (0.056) & 0.53 (0.0594) & 0.35 (0.0398) & 0.61 (0.0254) & 0.31 (0.0504) & 0.37 (0.1396) \\ 
-  snp-mrna & 0.75 (0.0533) & 0.61 (0.0553) & 0.43 (0.0374) & 0.65 (0.0239) & 0.44 (0.048) & 0.44 (0.1342) \\ 
+T & 0.75 (0.0552) & 0.34 (0.0619) & 0.4 (0.0412) & 0.47 (0.0293) & 0.16 (0.0496) & 0.27 (0.1447) \\ 
+  P & 0.54 (0.0493) & 0.48 (0.0521) & 0.21 (0.0365) & 0.6 (0.0187) & 0.35 (0.0427) & 0.42 (0.117) \\ 
+  G & 0.74 (0.0517) & 0.63 (0.0521) & 0.43 (0.0356) & 0.65 (0.0223) & 0.46 (0.0458) & 0.47 (0.1261) \\ 
+  PG & 0.71 (0.054) & 0.57 (0.0545) & 0.35 (0.0375) & 0.62 (0.0235) & 0.39 (0.0478) & 0.42 (0.1298) \\ 
+  PT & 0.73 (0.056) & 0.53 (0.0594) & 0.35 (0.0398) & 0.61 (0.0254) & 0.31 (0.0504) & 0.37 (0.1396) \\ 
+  GT & 0.75 (0.0533) & 0.61 (0.0553) & 0.43 (0.0374) & 0.65 (0.0239) & 0.44 (0.048) & 0.44 (0.1342) \\ 
    \midrule
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-mrna-none & 0.75 (0.0552) & 0.34 (0.0619) & 0.4 (0.0412) & 0.47 (0.0293) & 0.16 (0.0496) & 0.27 (0.1447) \\ 
-  ped-none1 & 0.46 (0.0469) & 0.53 (0.05) & 0.23 (0.0368) & 0.46 (0.0234) & 0.38 (0.0391) & 0.34 (0.1316) \\ 
-  snp-none1 & 0.67 (0.0534) & 0.47 (0.0552) & 0.43 (0.0384) & 0.48 (0.027) & 0.36 (0.0457) & 0.35 (0.132) \\ 
+P & 0.46 (0.0469) & 0.53 (0.05) & 0.23 (0.0368) & 0.46 (0.0234) & 0.38 (0.0391) & 0.34 (0.1316) \\ 
+  G & 0.67 (0.0534) & 0.47 (0.0552) & 0.43 (0.0384) & 0.48 (0.027) & 0.36 (0.0457) & 0.35 (0.132) \\ 
    \bottomrule
@@
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 11:36:08 2017
+% Wed Mar 15 17:36:40 2017
 % Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
@@
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-snp-mrna & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
-  snp-none & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
+G & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
+  GT & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
    \midrule
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-mrna-none\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
-  snp-mrna\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
-  snp-mrna\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
-  snp-mrna\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
-  snp-mrna\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
-  snp-mrna\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
-  snp-mrna\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
-  snp-mrna\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
-  snp-none\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
+G\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
+  GT\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
+  GT\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
+  GT\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
+  GT\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
+  GT\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
+  GT\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
+  GT\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
+  T\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
    \bottomrule
@@
 \section{Acknowledgments} 
-We thank the staff of the Agricultural Experimental Research station,
-University of Hohenheim, for excellent technical assistance in
-conducting the field experiments. We are indebted to the group of R.
-Fries from Technische Universit{\"a}t M{\"u}nchen for the SNP genotyping of the
-parent inbred lines, to X. Mi for his assistance in preparing auxiliary
-figures based on the Mathematica software, to C. Zenke for advice on the
-computation of transcriptomic BLUEs and to P. Schopp for advice on
-prediction models. The authors acknowledge support by the state of
-Baden-W{\"u}rttemberg through bwHPC. This project was funded by the German
-Federal Ministry of Education and Research (BMBF) within the projects
-OPTIMAL (FKZ: 0315958B,0315958F), SYNBREED (FKZ: 0315528D) and by the
-German Research Foundation (DFG, Grants No. ME 2260/5-1 and SCHO 764/6-1).
+This project was funded by the German Federal Ministry of Education and 
+Research (BMBF) within the projects OPTIMAL (FKZ: 0315958B,0315958F),
+SYNBREED (FKZ: 0315528D) and by the German Research Foundation 
+(DFG, Grants No. ME 2260/5-1 and SCHO 764/6-1).
+The authors acknowledge support by the state of Baden-W{\"u}rttemberg through 
+bwHPC.
 Financial support for M.W. was provided by the Fiat Panis foundation, Ulm,

### 2017-03-16 paper/manuscript.tex
Cosmetic changes to tables and figures

@@
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 17:33:34 2017
+% Wed Mar 15 17:31:28 2017
 % Script: ./analysis/maizego_plot_predictive_abilities.R
-\begin{table}[ht]
-\centering
-\caption{Predictive abilities and corresponding coefficients of variation for the set of 211 tropical/subtropical lines from the maize diversity panel.} 
-\begin{tabular}{rllllll}
-  \toprule
- & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
-  \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-G & 0.43 (0.07) & 0.37 (0.17) & 0.39 (0.05) & 0.34 (0.05) & 0.45 (0.08) & 0.53 (0.03) \\ 
-  GT & 0.44 (0.07) & 0.33 (0.17) & 0.4 (0.05) & 0.34 (0.04) & 0.4 (0.07) & 0.53 (0.03) \\ 
-   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-G & 0.36 (0.08) & 0.21 (0.17) & 0.3 (0.05) & 0.34 (0.05) & 0.41 (0.08) & 0.45 (0.03) \\ 
-  T & 0.4 (0.07) & 0.15 (0.16) & 0.37 (0.05) & 0.36 (0.04) & 0.32 (0.07) & 0.49 (0.03) \\ 
-   \bottomrule
-\multicolumn{6}{l}{}\\
-\end{tabular}
-\end{table}
-
-% latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 17:31:28 2017
-% Script: ./analysis/uhoh_plot_predictive_abilities.R
 \begin{table}[ht]
@@
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 17:36:40 2017
+% Thu Mar 16 08:16:30 2017
+% Script: ./analysis/uhoh_plot_predictive_abilities.R
+\begin{table}[ht]
+\centering
+\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids.
+As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.}
+\begin{tabular}{rllllll}
+  \toprule
+ & DMY & DMC & FAT & PRO & STA & SUG \\ 
+  \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
+P & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
+  G & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
+  PG & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
+  PT & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
+  GT & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
+   \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
+T & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
+  P & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
+  G & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
+   \bottomrule
+\multicolumn{6}{l}{}\\
+\end{tabular}
+\end{table}
+
+% latex table generated in R 3.3.3 by xtable 1.8-2 package
+% Thu Mar 16 08:03:02 2017
 % Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
@@
 \centering
-\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel.} 
+\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel.
+As predictors, genomic (G) data, transcriptomic (T) data and their combination (GT) were used.
+The decimal after the predictor name indicates the size of the genetic material that was covered by inbred lines in the core set.}
 \begin{tabular}{rllllll}
@@
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-G & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
-  GT & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
+G\_Full & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
+  GT\_Full & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
    \midrule
@@
 G\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
+  T\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
+  GT\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
+  GT\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
+  GT\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
+  GT\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
+  GT\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
+  GT\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
   GT\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
-  GT\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
-  GT\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
-  GT\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
-  GT\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
-  GT\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
-  GT\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
-  T\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
    \bottomrule

### 2017-03-16 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -1,7 +1,7 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "data.table", "dtplyr", "xtable", "stringr",
-               "forcats", "ggthemes", "viridis")
+               "forcats", "ggthemes", "viridis", "gsubfn")
 source("./software/prediction_helper_functions.R")
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
@@ -19,7 +19,7 @@ filtered_log_file <- log_file %>%
     (Pred1 == "mrna" & Pred2 == "none"), true = "1.0", false = Core_Fraction
   ))
 
-pred_df <- filtered_log_file %>%
+raw_pred_df <- filtered_log_file %>%
   select(Job_ID) %>%
   flatten_chr() %>%
   map(~readRDS(paste0(prediction_path, ., ".RDS"))) %>%
@@ -32,9 +32,20 @@ pred_df <- filtered_log_file %>%
   droplevels() %>%
   select(-Job_ID)
   
+abbrev_pred_df <- raw_pred_df %>%
+  mutate(
+    Predictor = gsubfn(
+      pattern = "\\S+",
+      replacement = list(
+        "snp-none" = "G",
+        "mrna-none" = "T",
+        "snp-mrna" = "GT"
+      ),
+      x = Predictor
+    )
+  )
   
-
-pred_lst <- pred_df %>%
+pred_df <- abbrev_pred_df %>%
   mutate(
     Reduced = if_else(Core_Fraction == "Full", true = FALSE, false = TRUE)
   ) %>%
@@ -53,11 +64,14 @@ pred_lst <- pred_df %>%
   ) %>%
   select(-r, -CV) %>%
   spread(key = Trait, value = Value) %>%
+  arrange(Reduced, desc(Core_Fraction), Predictor) %>%
+  mutate(
+    Predictor = paste0(Predictor, "_", Core_Fraction)
+  )
+  
+pred_lst <- pred_df %>%
   split(.$Reduced) %>%
   map(., ~select(., -Reduced)) %>%
-  map_at("TRUE", .f = ~mutate(
-    ., Predictor = paste0(Predictor, "_", Core_Fraction)
-  )) %>%
   map(., ~select(., -Core_Fraction)) %>%
   map(., ~column_to_rownames(., var = "Predictor"))
 
[…]

### 2017-03-16 analysis/maizego_plot_predictive_abilities.R

@@ -86,7 +86,9 @@ coverage_subheading <- "Only incomplete predictor:"
 attr(pred_lst, "subheadings") <- paste(coverage_subheading, names(pred_lst))
 inbred_caption <- paste(
   "Predictive abilities and corresponding coefficients of variation for the",
-  "set of 211 tropical/subtropical lines from the maize diversity panel."
+  "set of 211 tropical/subtropical lines from the maize diversity panel.",
+  "As predictors, genomic data (G), transcriptomic data (T) and their",
+  "combination (GT) were used."
 )
 xtable_lst <- xtableList(pred_lst,
                          caption = inbred_caption)

### 2017-03-16 analysis/maizego_snp_analyses.R

@@ -153,6 +153,7 @@ g3 <- pc_df %>%
   rename(`Core Group` = Core_Group) %>%
   ggplot(aes(x = PC1, y = PC2, color = `Core Group`, shape = `Core Group`)) +
   geom_point(size = 0.5) +
+  guides(color = guide_legend(override.aes = list(size = 3))) +
   ggthemes::scale_color_tableau() +
   facet_wrap(~Fraction, nrow = 3, ncol = 3) +
   theme_bw(base_size = 10) +

### 2017-03-16 analysis/uhoh_pca.R

@@ -52,6 +52,7 @@ quality_snps %>%
 # Determine the structure of the data using genotypic data.
 pca_lst <- lfmm_path %>%
   list.files(pattern = "lfmm") %>%
+  discard(. == "snp_Hybrid.lfmm") %>%
   map(., ~paste0(lfmm_path, .)) %>%
   map(., ~LEA::pca(input.file = ., scale = TRUE)) %>%
   map(function(x) x$projections) %>%
@@ -94,6 +95,7 @@ g1 <- pca_df %>%
   spread(key = PC, value = Score) %>%
   ggplot(aes(x = PC1, y = PC2, color = Reduced, shape = Reduced)) +
   geom_point() +
+  guides(color = guide_legend(override.aes = list(size = 3))) +
   scale_color_tableau() +
   facet_grid(. ~ Group) +
   theme_bw(base_size = 10) +

### 2017-03-16 analysis/uhoh_plot_predictive_abilities.R

@@ -63,10 +63,10 @@ raw_pred_df <- raw_pred_df %>%
     Predictor = fct_relevel(Predictor, pred_order)
   )
  
-  
+ 
 pred_lst <- raw_pred_df %>%
   mutate(
-    Reduced = ifelse(Predictor == "mrna-none", yes = TRUE, no = FALSE),
+    Reduced = ifelse(Predictor == "T", yes = TRUE, no = FALSE),
     Reduced = ifelse(Core_Fraction == "1.0", yes = TRUE, no = Reduced) 
   ) %>%
   select(-Core_Fraction) %>%
@@ -78,7 +78,7 @@ pred_lst <- raw_pred_df %>%
   ungroup() %>%
   mutate(
     r = round(r, digits = 2),
-    CV = round(CV, digits = 4),
+    CV = round(CV, digits = 3),
     Value = paste0(
       r, " (", CV, ")"
     )
@@ -95,7 +95,8 @@ coverage_subheading <- "Only incomplete predictor:"
 attr(pred_lst, "subheadings") <- paste(coverage_subheading, names(pred_lst))
 hybrid_caption <- paste(
   "Predictive abilities and corresponding coefficients of variation for the",
-  "set of maize hybrids"
+  "set of maize hybrids. As predictors, pedigree (P), genomic (G),",
+  "transcriptomic (T) data and combinations thereof we used."
 )
 xtable_lst <- xtableList(pred_lst,
                          caption = hybrid_caption)
@@ -108,3 +109,4 @@ print.xtableList(xtable_lst,
                  booktabs = TRUE,
                  sanitize.subheadings.function = bold,
                  caption.placement = "top")
+

### 2017-03-23 analysis/maizego_core_fraction_plot_predictive_ability.R
Reorder core set predability line plot

@@ -54,7 +54,9 @@ pred_df <- abbrev_pred_df %>%
     r = cor(y, yhat),
     CV = coefficient_of_variation(var_yhat, yhat)
   ) %>%
-  ungroup() %>%
+  ungroup()
+
+xtable_pred_df <- pred_df %>%
   mutate(
     r = round(r, digits = 2),
     CV = round(CV, digits = 4),
@@ -69,7 +71,7 @@ pred_df <- abbrev_pred_df %>%
     Predictor = paste0(Predictor, "_", Core_Fraction)
   )
   
-pred_lst <- pred_df %>%
+pred_lst <- xtable_pred_df %>%
   split(.$Reduced) %>%
   map(., ~select(., -Reduced)) %>%
   map(., ~select(., -Core_Fraction)) %>%
@@ -103,14 +105,17 @@ print.xtableList(xtable_lst,
 
 # PLOT THE TREND IN R OVER CORE SETS --------------------------------------
 g1 <- pred_df %>%
-  filter(!Core_Fraction %in% c("Full", "1.0")) %>%
+  filter(Reduced == TRUE,
+         Predictor == "GT") %>%
   group_by(Trait, Core_Fraction) %>%
-  summarize(r = cor(y, yhat)) %>%
   rename(`Core Fraction` = Core_Fraction) %>%
-  ggplot(aes(x = `Core Fraction`, y = r, color = Trait, group = Trait)) +
+  ggplot(aes(x = `Core Fraction`, y = r, group = Trait, 
+             color = forcats::fct_reorder2(Trait, `Core Fraction`, r))) +
   geom_line() +
+  labs(color = "Trait") +
   scale_color_viridis(discrete = TRUE) +
-  theme_bw(base_size = 10)
+  theme_bw(base_size = 10) +
+  guides(color = guide_legend(override.aes = list(size = 3)))
 plot_name <- "./paper/tables_figures/core_fraction_predictive_ability_trend.pdf"
 ggsave(plot = g1, 
        filename = plot_name,

### 2017-03-23 analysis/maizego_snp_analyses.R
Restrict admixture to 149 core genotypes

@@ -3,12 +3,21 @@ if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "cowplot", 
                "ggthemes")
 
+# For the second scenario, keep only names of genotypes, which are covered by 
+# all data types (i.e. phenotypic, genotypic and transcriptomic).
+common_genotypes <- readRDS(
+  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
+)
+common_genotypes <- common_genotypes %>%
+  reduce(intersect)
 
-snp <- readRDS("./data/processed/maizego/imputed_snp_mat.RDS")
-unique_genotypes <- "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS" %>%
+snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
   readRDS() %>%
-  keep(names(.) == "snp") %>%
-  flatten_chr()
+  .[rownames(.) %in% common_genotypes, ] %>%
+  sspredr::ensure_snp_quality(
+    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05, 
+    any_missing = FALSE, remove_duplicated = TRUE
+  )
 
 
 # PCA ---------------------------------------------------------------------
@@ -36,12 +45,14 @@ snp_tw %>%
 # subpopulations.
 lfmm2geno("./data/derived/maizego/snp.lfmm", 
           output.file = "./data/derived/maizego/snp.geno")
-project <- snmf(input.file = "./data/derived/maizego/snp.geno",
-                K = seq_len(tw_K),
-                project = "new",
-                repetitions = 25,
-                CPU = 3,
-                entropy = TRUE)
+# This code snippet is commented out because it takes a long time to run and 
+# can easily and inadvertantly be overridden.
+#project <- snmf(input.file = "./data/derived/maizego/snp.geno",
+#                K = seq_len(tw_K),
+#                project = "new",
+#                repetitions = 25,
+#                CPU = 3,
+#                entropy = TRUE)
 
 # Based on the results of a PCA, decide whether it is even necessary to treat
 # subpopulations as such or whether all inbred lines can be treated as a single
[…]

### 2017-03-23 software/prediction_helper_functions.R
Move prediction timer function to helper script

@@ -1,3 +1,15 @@
+# Determine how much time (hh:mm:ss format) has elapsed since script initiation.
+get_elapsed_time <- function(start_time, tz = "CEST") {
+  start_time <- as.POSIXct(start_time)
+  sec <- Sys.time() %>%
+    difftime(time1 = ., time2 = start_time, units = "secs") %>%
+    lubridate::seconds_to_period()
+  paste0(sprintf("%02d", c(day(sec), hour(sec), minute(sec), second(sec))),
+         collapse = ":")
+}
+
+
+
 # In the case of hybrid data, we still need to split all predictor matrices
 # into Flint and Dent components first.
 # In the case of pedigree data, we need to split the data by genotypes in the

### 2017-03-23 analysis/prediction.R
Add admixture-based core sets to predictions

@@ -32,10 +32,10 @@ if (isTRUE(interactive())) {
   Sys.setenv("MOAB_PROCCOUNT" = "3")
   # Are you using inbred ("Inbred") data from the Yan-lab or hybrid ("Hybrid")
   # data from the UHOH-group?
-  Sys.setenv("DATA_TYPE" = "Hybrid")
+  Sys.setenv("DATA_TYPE" = "Inbred")
   # Which agronomic trait do you want to evaluate? Leave blank if you want to
   # analyze all traits.
-  Sys.setenv("TRAIT" = "GTM")
+  Sys.setenv("TRAIT" = "100grainweight")
   # Number of iterations in BGLR()
   Sys.setenv("ITER" = "3000") 
   # Prediction model in BGLR()
@@ -46,11 +46,11 @@ if (isTRUE(interactive())) {
   Sys.setenv("PRIOR_PI_COUNT" = "10")
   # Main predictor. If 'Pred2' and 'Pred3' are empty, no imputation will take
   # place.
-  Sys.setenv("PRED1" = "ped")
+  Sys.setenv("PRED1" = "snp")
   # If 'Pred3' is empty, 'Pred2' will be imputed via information from 'Pred1'.
-  Sys.setenv("PRED2" = "")
+  Sys.setenv("PRED2" = "mrna")
   # Fraction of genotypes to be included in the core set.
-  Sys.setenv("CORE_FRACTION" = "1.0")
+  Sys.setenv("CORE_SET" = "A1")
   # Number of genotypes to predict (only for testing!)
   Sys.setenv("RUNS" = "1-3")
   # Output directory for temporary BGLR files
@@ -75,7 +75,7 @@ Pi <- as.numeric(Sys.getenv("PI"))
 PriorPiCount <- as.integer(Sys.getenv("PRIOR_PI_COUNT"))
 pred1 <- as.character(Sys.getenv("PRED1"))
 pred2 <- as.character(Sys.getenv("PRED2"))
-core_fraction <- as.character(Sys.getenv("CORE_FRACTION"))
+core_set <- as.character(Sys.getenv("CORE_SET"))
 runs <- as.character(Sys.getenv("RUNS"))
 temp <- paste0(as.character(Sys.getenv("TMP")), "/")
 
@@ -94,20 +94,12 @@ pred_sets <- pred_combi %>%
 
 
 ## -- INPUT CHECKS AND UPDATES ------------------------------------------
[…]

### 2017-03-23 moab_scripts/prediction.sh

@@ -50,7 +50,7 @@
 #
 # Specify the core set set size, as a fraction of genotypes covered by mrna
 # data
-#MOAB -v CORE_FRACTION=
+#MOAB -v CORE_SET=
 #
 #MOAB -v RUNS=
 #
@@ -77,14 +77,14 @@ echo "Data_Type=${DATA_TYPE}\
       PriorPiCount=${PRIOR_PI_COUNT}\
       Pred1=${PRED1}\
       Pred2=${PRED2}\
-      Core_Fraction=${CORE_FRACTION}\
+      Core_Set=${CORE_SET}\
       Runs=${RUNS}"
 
 # Set-up program
 startprog="Rscript --no-save --no-restore --slave\
            ./analysis/prediction.R\
            ${DATA_TYPE} ${TRAIT} ${ITER} ${MODEL} ${VCOV} ${PI}\
-           ${PRIOR_PI_COUNT} ${PRED1} ${PRED2} ${CORE_FRACTION}\
+           ${PRIOR_PI_COUNT} ${PRED1} ${PRED2} ${CORE_SET}\
            ${RUNS}"
 
 # Start program

### 2017-03-23 paper/manuscript.tex
Assemble admixture-based core sets

@@
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 15 17:31:28 2017
-% Script: ./analysis/maizego_plot_predictive_abilities.R
-\begin{table}[ht]
-\centering
-\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids} 
-\begin{tabular}{rllllll}
-  \toprule
- & DMY & DMC & FAT & PRO & STA & SUG \\ 
-  \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-T & 0.75 (0.0552) & 0.34 (0.0619) & 0.4 (0.0412) & 0.47 (0.0293) & 0.16 (0.0496) & 0.27 (0.1447) \\ 
-  P & 0.54 (0.0493) & 0.48 (0.0521) & 0.21 (0.0365) & 0.6 (0.0187) & 0.35 (0.0427) & 0.42 (0.117) \\ 
-  G & 0.74 (0.0517) & 0.63 (0.0521) & 0.43 (0.0356) & 0.65 (0.0223) & 0.46 (0.0458) & 0.47 (0.1261) \\ 
-  PG & 0.71 (0.054) & 0.57 (0.0545) & 0.35 (0.0375) & 0.62 (0.0235) & 0.39 (0.0478) & 0.42 (0.1298) \\ 
-  PT & 0.73 (0.056) & 0.53 (0.0594) & 0.35 (0.0398) & 0.61 (0.0254) & 0.31 (0.0504) & 0.37 (0.1396) \\ 
-  GT & 0.75 (0.0533) & 0.61 (0.0553) & 0.43 (0.0374) & 0.65 (0.0239) & 0.44 (0.048) & 0.44 (0.1342) \\ 
-   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-P & 0.46 (0.0469) & 0.53 (0.05) & 0.23 (0.0368) & 0.46 (0.0234) & 0.38 (0.0391) & 0.34 (0.1316) \\ 
-  G & 0.67 (0.0534) & 0.47 (0.0552) & 0.43 (0.0384) & 0.48 (0.027) & 0.36 (0.0457) & 0.35 (0.132) \\ 
-   \bottomrule
-\multicolumn{6}{l}{}\\
-\end{tabular}
-\end{table}
-
-% latex table generated in R 3.3.3 by xtable 1.8-2 package
 % Thu Mar 16 08:16:30 2017
@@
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Thu Mar 16 08:03:02 2017
+% Wed Mar 22 08:40:11 2017
 % Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
@@
   GT\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
+  GT\_0.2 & 0.35 (0.072) & 0.23 (0.1645) & 0.26 (0.0519) & 0.36 (0.0474) & 0.38 (0.0722) & 0.44 (0.0273) \\ 
+  GT\_0.1 & 0.29 (0.0762) & 0.21 (0.1659) & 0.23 (0.0522) & 0.32 (0.0471) & 0.4 (0.0743) & 0.39 (0.0302) \\ 
    \bottomrule

### 2017-03-23 analysis/maizego_manual_corehunter.R

@@ -1,0 +1,62 @@
+# Goal: Add three new core sets (based on a previous admixture analysis) to the
+# existing list of core samples.
+# our predictions in increments of ten percentage points.
+if (!require("pacman")) install.packages("pacman")
+if (!require("devtools")) install.packages("devtools")
+devtools::install_github("mwesthues/sspredr", update = TRUE)
+options(java.parameters = "-Xmx25G")
+pacman::p_load("tidyverse", "data.table", "dtplyr", "corehunter", "LEA", 
+               "ggthemes")
+pacman::p_load_gh("mwesthues/sspredr")
+
+# This data frame contains the ancestry coefficients of the genotypes that are
+# covered by both, the complete (SNP) and the incomplete (mRNA) predictors, 
+# based on an admixture analysis assuming three ancestral populations.
+ancestry_df <- "./data/derived/maizego/ancestry_k3.RDS" %>%
+  readRDS()
+
+# Genotypes with a share >= 50% of the second ancestral population (A2) comprise
+# a far larger set (n = 109) than genotypes with either a share >= 50% of the
+# first or the third ancestral population (n = 17, n = 19).
+# To reduce the bias of the sample size on predictions using these three 
+# ancestral populations, we will build a core set from the A2-genotypes of 
+# similar size as the two other sets (n = 19).
+a2 <- ancestry_df %>%
+  filter(Ancestor == "A2") %>% 
+  select(G) %>% 
+  flatten_chr()
+
+a2_core <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
+  readRDS() %>%
+  .[rownames(.) %in% a2, ] %>%
+  sspredr::ensure_snp_quality(
+    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
+    any_missing = FALSE, remove_duplicated = TRUE
+  ) %>%
+  genotypes(format = "biparental") %>%
+  sampleCore(
+    ., size = 19, obj = objective(type = "EN", measure = "MR", weight = 1)
+  )
+
+# Genotypes belonging largely to the first or third ancestral population will
+# not be modified and simply be assigned to the list with the names of core set
+# genotypes.
+a1_a3 <- ancestry_df %>% 
+  split(.$Ancestor) %>% 
+  map("G") %>% 
[…]

### 2017-03-24 paper/manuscript.tex
Update plot legends for manuscript

@@
 \section*{Results}
-
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
-\caption{
-Influence of core set composition on predictive abilities.
-Predictive abilities are based on single step predition with genomic data as the 
-complete predictor and transcriptomic data as the incomplete predictor.
-Results are shown for six agronomic traits and nine core sets of the maize 
-diversity panel.
-}
-\label{fig:CoreSetInfluence}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/uhoh_dent_flint_pca.pdf}
-\caption{
-  Principal component (PC) analysis of the two heterotic groups of the maize 
-  hybrids.
-  All inbred lines, which are covered by genomic and transcriptomic features
-  are depicted as orange triangles whereas genotypes for which only genomic
-  information are available are depicted as blue dots.
-}
-\label{fig:UHOH-PCA}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/maizego_admixture_pca.pdf}
-\caption{
-  (\textbf{A}) Ancestry coefficients for 211 tropical/subtropical inbred lines
-  from the maize diversity panel considering $K=2 \dots 5$ ancestral
-  populations.
-  (\textbf{B}) Principal component analysis for nine different core sets of the
-  set of tropical/subtropical maize inbred lines.
-  For each size of the core set, its members are plotted as orange dots whereas 
-  all genotypes that are not elements of the cores sets (\textit{i.e.},
-  'Complement') are depicted as blue dots.
-}
-\label{fig:PopStructure}
-\end{figure}
-
-% latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Thu Mar 16 08:16:30 2017
-% Script: ./analysis/uhoh_plot_predictive_abilities.R
-\begin{table}[ht]
-\centering
-\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids.
-As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.}
-\begin{tabular}{rllllll}
-  \toprule
- & DMY & DMC & FAT & PRO & STA & SUG \\ 
-  \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-P & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
-  G & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
-  PG & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
-  PT & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
-  GT & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
-   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-T & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
-  P & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
-  G & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
-   \bottomrule
-\multicolumn{6}{l}{}\\
-\end{tabular}
-\end{table}
+\Genetics2level{Inbred lines}
 
@@
 \begin{table}[ht]
+\label{tbl:InbredTable}
 \centering
@@
 
-\Genetics2level{Some results section}
+\Genetics2level{Hybrids}
 
-$\dots$
+% latex table generated in R 3.3.3 by xtable 1.8-2 package
+% Thu Mar 16 08:16:30 2017
+% Script: ./analysis/uhoh_plot_predictive_abilities.R
+\begin{table}[ht]
+\label{tbl:HybridTable}
+\centering
+\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids.
+As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.}
+\begin{tabular}{rllllll}
+  \toprule
+ & DMY & DMC & FAT & PRO & STA & SUG \\ 
+  \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
+P & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
+  G & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
+  PG & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
+  PT & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
+  GT & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
+   \midrule
+\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
+T & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
+  P & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
+  G & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
+   \bottomrule
+\multicolumn{6}{l}{}\\
+\end{tabular}
+\end{table}
 
-\Genetics2level{Another results section}
+\Genetics2level{Coverage of the genetic space}
 
-$\dots$
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
+\caption{
+Influence of core set composition on predictive abilities.
+Predictive abilities are based on single step predition with genomic data as the 
+complete predictor and transcriptomic data as the incomplete predictor.
+Results are shown for six agronomic traits and nine core sets of the maize 
+diversity panel.
+}
+\label{fig:CoreSetInfluence}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/maizego_core_sampling_pca.pdf}
+\caption{
+  Coverage of the genetic space by nine different core sets. 
+  Genotypes are distributed according to their scores based on the first two
+  principal components (PC).
+  The total genetic space was spanned by 149 genotypes from the maize diversity 
+  panel for which both, genomic and transcriptomic data, were available.
+  Genotypes that were members of the core set (orange triangles) were declared 
+  as being covered by both predictors whereas members of the complement (blue,
+  filled circles) were declared as only being covered by genomic information.
+  Assignments of genotypes to the core set was done by selecting a subset of
+  genotypes that maximized the average genetic distance in the core set based 
+  on the Modified Rogers distance between all genotypes.
+  Core set sizes ranged from 10\% to 90\% of the 149 available genotypes and
+  are given as decimals above each facet of the plot.
+}
+\label{fig:CoreSetPCA}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/maizego_admixture_k3_pca.pdf}
+\caption{
+  Population structure of 149 maize inbred lines from the maize diversity panel
+  covered by genomic and transcriptomic data.
+  (\textbf{A}) Ancestry coefficients from an admixture analysis when 
+  considering $K=3$ ancestral populations.
+  (\textbf{B}) Principal component analysis with genotypes colored according to 
+  the most likely ancestral population (\textit{i.e.}, share $\geq 0.5$) as 
+  indicated by the admixture analysis.
+  Genotypes for which none of the three ancestry coefficients reached a value
+  of 0.5 were classified as "mixed".
+}
+\label{fig:PopStructure}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/uhoh_dent_flint_pca.pdf}
+\caption{
+  Principal component (PC) analysis of the two heterotic groups of the maize 
+  hybrids.
+  All inbred lines, which are covered by genomic and transcriptomic features
+  are depicted as orange triangles whereas genotypes for which only genomic
+  information are available are depicted as blue dots.
+}
+\label{fig:UHOH-PCA}
+\end{figure}
 

### 2017-03-24 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -56,20 +56,29 @@ pred_df <- abbrev_pred_df %>%
   ) %>%
   ungroup()
 
+# Custom order for predictors.
+pred_order <- c(
+  "G_Full", "GT_Full", "G_1.0", "T_1.0", 
+  paste0("GT_", seq(from = 0.1, to = 0.9, by = 0.1)),
+  paste0("GT_A", seq_len(3))
+)
+
 xtable_pred_df <- pred_df %>%
   mutate(
     r = round(r, digits = 2),
-    CV = round(CV, digits = 4),
+    CV = round(CV, digits = 3),
     Value = paste0(
       r, " (", CV, ")"
     )
   ) %>%
   select(-r, -CV) %>%
   spread(key = Trait, value = Value) %>%
-  arrange(Reduced, desc(Core_Fraction), Predictor) %>%
   mutate(
     Predictor = paste0(Predictor, "_", Core_Fraction)
   )
+
+xtable_pred_df %>%
+  mutate(Predictor = factor(Predictor, levels = pred_order))
   
 pred_lst <- xtable_pred_df %>%
   split(.$Reduced) %>%

### 2017-03-24 analysis/prediction.R

@@ -50,7 +50,7 @@ if (isTRUE(interactive())) {
   # If 'Pred3' is empty, 'Pred2' will be imputed via information from 'Pred1'.
   Sys.setenv("PRED2" = "mrna")
   # Fraction of genotypes to be included in the core set.
-  Sys.setenv("CORE_SET" = "A1")
+  Sys.setenv("CORE_SET" = "A3")
   # Number of genotypes to predict (only for testing!)
   Sys.setenv("RUNS" = "1-3")
   # Output directory for temporary BGLR files

### 2017-03-24 paper/notes.Rmd
Init notes for paper

# Aguilar 2009 JDairySci
-   "A typical evaluation erquires 1) traditional evaluation with an animal model,
    2) extraction of pseudo-observations such as deregressed evaluations or
    daughter deviations (DD), 3) estimation of genomic effects for genotypes
    animals usually using simple sire models, and possibly 4) combining the genomic
    index with traditional parent averages (PA) and EBV (Hayes 2009, VanRaden
    2009b)."

-   "Genomic evaluations are more accurate than PA and approach the accuracy of
  evaluations for progeny-tested bulls, but they also seem inflated (VanRaden
  2009a). Although their inflation is lower than that of current PA, the
  potentially great utilization of top genomically evaluated young sires
  increases the improtance of high accuracy and minimum bias. Inflation of
  genetic evalutions by genomic information causes top young bulls to have an
  unfair advantage over older progeny-tested bulls. Some of the problems with
  genomic evaluations may be caused by incorrect parameters and strong
  assumptions used in multiple-step procedures. However, effects of those
  parameters and assumptions are extremely difficult to verify, particularly in
  the presence of selection."

-   "A more serious problem is when pseudo-observations are poorly defined or
  of poor quality (e.g. for animals with small progeny numbers), which is often
  the case for monogastric species and for beef cattle."

-   "The objective of this study was to use a single-step procedure for genomic
    evaluation in a national evaluation setting and compare its performance to
    a multi-step procedure."

# Ashraf 2016 CropSci
-   "The objective of this study was to evaluate single-step GP in a wheat
  program. We compared the performance of pedigree-based, marker-based and
  single-step models (ABLUP, GBLUP and HBLUP, respectively). Data consisted of
  1,176 genotypes (via genotyping-by-sequencing) and 11,131 nongentyped wheat
  lines."

-   In a dairy cattle study, Su et al. (2012) reported that this single-step
  approach gave a more accurate GEBV than a multistep procedure."

-   "The main aim of this study was to investigate the impact on prediction
  accuracy when some wheat lines are not genotyped. In additiona, we compared
  the performance of different prediction models and determined the impact of
  different weights for the pedigree relationship on prediction accuracy."

-   "The single-step method (Christensen et al. 2012, Legarra et al. 2009)
  includes information from genotyped and nongentoyped individuals by blending
  the genomic relationship matrix $\mathbf{G}$ and the pedigree-based matrix
  $\mathbf{A}$. The HBLUP method uses the same linear model as the GBLUP
  methods with the exception that $u \texttilde N(0,
  \mathbf{H}\sigma^{2}_{g}$), where $\mathbf{H}$ is a combined relationship
  matrix (Christensen et al. 2012, Legarra et al. 2009)."

-   "We used $w = 0.05$ to place the majority of the weight on the genomic
  relationship matrix as this matrix should be able to track Mendelian
  segregation and we therefore assumed that it would be a more accurate
  estimate of the true relationship. A similar $w$ was also tested in
  Christensen et al. (2012)."

-   Optimized the weight of genomic and pedigree information in blending the
  two matrices by fitting them simultaneous as random effects in a mixed model.

-   "The polygenic effect weight, $w$, optimized in the HBLUP-weight model,
  outperformed the fixed weight used in HBLUP for all traits.

-   "First, HBLUP accuracy was greater than GBLUP accuracy in all caes. Similar
  results were reported by Gao et al. (2012) in Nordic Holstein population. The
  likely reason for this superiority is that HBLUP was able to use much more
  phenotypic information than GBLUP. From Table 1, we can see than on average
  over all traits, there were close to 10,000 observations on nongenotyped
  lines. These observations were not available to GBLUP, for which there were
  generally < 7000 observations across six environments."

-   "Optimzing the weight $w$ for combining the pedigree and the genomic
  relationship matrices in scenario 2 (termed HBLUP-weight here) gave higher
  accuracies than fixed weight HBLUP in all traits. The optimal weighting
  factors differed between the traits."

-   "One interpretation of the different weights between traits is that this
  parameter is affected by the fraction of the variance caused by low minor
  allele frequency (MAF) loci whose effects are poorly captured by markers. For
  traits with a high fraction of the variance generated by such low MAF causal
  loci, the weight would be high on the pedigree component."

# Christensen and Lund 2010, GSE
-   "The extension of the genomic relationship matrix to non-gneotyped animals
  together with the addition of the polygenic effect provide a natural one-step
  procedure to blend the information from relatives and the genomic information
  into a combined genomically enhanced breeding value (GEBV)."

# Christensen et al. 2012, Animal
-   "The initial methods for incorporating genomic information into genetic
  evaluation were multi-step. These methods consist of (1) constructing a
  response variable for the genotyped animals that summarises the phenotypic
  information, (2) a genomic prediction method that associates the responsive
  variable to the marker information and (3) blending genomic predictions with
  parent average estimated breeding values (EBV)."

-   "Single-step methods provide predictions for all animals in the pedigree."

-   "Furthermore, the single-step methods include information about Mendelian
  inheritance through markers, thereby providing a more accurate modelling, and
  we therefore conjecture that for non-gneotyped animals single-step methods
  also produce more accurate predictions than the pedigree-based method."

-   "The single-step method in theory requires that the genomic relationship
  matrix is constructed using the allele frequencies in the founder population
  of the pedigree, but these are not feasible to compute with sufficient
  accuracy in practice since founder animals are ussually not genotyped. Two
  obvious alternatives are to use either observed allele frequencies or assume
  that allele frequencies are equal to 0.5 instead. Forni et al. (2011) and
  Chen et al. (2011) concluded that using observed allele frequencies resulted
  in the most accurate predictions."

-   "The specific aim of this study on DG and FCR in Danish Duroc pigs were to
  demonstrate that (1) single-step methods are at least as accurate as GBLUP
  for prediction of genotyped animals, and that both outperform the
  pedigree-based method in this respect, (2) single-step methods provide more
  accurate predictions than the pedigree-based method for non-genotyped
  animals, (3) adjusted single-step method produces more accureate predictions
  and behave better than the original single-step method, and 94) bivariate
  models improve accuracy of predictions for FCR, no matter which of the
  methods are used."

-   "Here we combined these two ideas and adjusted $\mathbf{G}$ to
  $\mathbf{G}_{a} = \beta \mathbf{G} + \alpha$, where $\beta$ and $\alpha$
  solved the system of equations $\text{Avg}(diag(\mathbf{G}))\beta + \alpha =
  \text{Avg}(diag(\mathbf{A}_{11}))$."

-   "The comparison of the different methods was therefore based on these
  validation correlations, and differences were assessed using the
  Hotelling-Williams *t*-test at confidence level 5%."

-   Contrary to what the authors claim, the differences in $r$ between regular
  single step prediction and single step predictions with adjustments to the
  genomic relationship matrix are minor (third digit after the comma, see Table
  2).

-   "The results show that the three methods with marker information (two
  single-step methods and GBLUP) produce more accurate predictions of genotyped
  animals than the pedigree-based method. In addition, single-step methods
  produce more accurate predictions for non-genotyped animals."

-   "In this paper, differences in validation correlation between methods are
  assessed with the Hotelling-Williams *t*-test. Most previous paper on genomic
  evaluation have not made any assessment of statistical significance of
  differences in accuracy between methods. The likely reason for this is that
  confidence intervals for accuracies of different methods are often highly
  overlapping, which falsely gives the impression of statistical
  non-significance. The Hotelling-Williams *t*-test takes into account the
  strong correlation of predictions from different methods, and is therefore a
  much more powerful test. The test is known in psychology (Revelle, 2010), but
  does not seem to be much known in other scientific fields."

# Christensen 2012, GSE
-   "In this paper, I propose a new method to address this issue by 1)
  adjusting the pedigree-based relationship matrix to be compatible with the
  marker-based relationship matrix instead of the reverse and 2) extending the
  single-step genetic evaluation using a joint likelihood of observed
  phenotypes and observed marker genotypes."

-   "A single-step method extends the usual pedigree-based method by replacing
  the additive relationship matrix constructed from pedigree by an additive
  relatoinship matrix that combines the marker-based relationship matrix for
  genotyped animals with the pedigree-based relationship matrix."

-   "An issue with a single-step method is compatability between the
  marker-based relationship matrix and the pedigree-based relationship matrix.
  To handle this problem, it is necessary to determine which allele frequencies
  should be used in the marker-based relationship matrix and to adjust this
  matrix to the pedigree-based relationship matrix. In theory, one should use
  the allele frequencies in the founder population of the pedigree (base
  animals) for the marker-based and pedigree-based relationships to be
  compatible, but these allele frequencies are rarely availalbe in practice
  since base animals are not genotyped."

-   "This results in a single-step method in which the marker-based
  relationship matrix is constructed assuming all allele frequencies are equal
  to 0.5 and the pedigree-based relationship matrix is constructed using the
  unusual assumption that animals in the base population are related and inbred
  with a relationship coefficient $\gamma$ and an inbreeding coefficient
  $\gamma / 2$."

# Christensen et al. 2014, GSE
-   "Production systems based on crossbreeding are predominant in pig and
  chicken breeding and take advantage of the increased performance of crossbred
  animals compared to purebred animals."

-   "Genomic selection of purebreds for crossbred performance was proposed by
  Ibanez-Escriche et al. (2009) that used phenotypes on crossbreds only, and a
  genomic model with a breed of origin specific allele substitution effects."

-   "A single-step method for genomic evaluation of both purebred and crossbred
  performances has not been developed, yet."

-   "A partial relationship matrix is a relationship matrix that describes the
  relationships only according to genetic origin."

# Fernando, Dekkers and Garrick 2014, GSE
-   "Phenotypic values or deregressed estimated breeding values (EBV) on these
  genotyped individuals are used to estimate the effects of the SNPs using
  Bayesian multiple-regression models in which the marker effects are treated
  as random. We refer to such models as marker effect models (MEM)."

-   "Fernando [7] showed that the MEM and BVM provide equivalent predictions of
  BV and that predictions could be more easily obtained from the MEM than from
  BVM because at the time of their study the number of markers was much smaller
  than the number of genotyped animals."

-   "Thus, their method is called single-step BLUP (SS-BLUP), and is expected
  to yield unbiased predictions under multivariate normality, even in
  populations that are undergoing selection and non-random mating. This is
  improtant, because genotypes are usually collected only on superior animals
  and this can lead to biased evaluations. A properly calculated BLUP
  evaluation has been shown to be free of this selection bias [15-21]."

-   "Conceptually, the remaining relationship coefficients constructed from
  pedigree are modified to provide consistency with $\mathbf{G}$ [11]."

MANY MORE NOTES

# Fernando et al. (2016), GSE
-   "The objective of this paper is to present computational strategies for
  whole-genome analyses based on the SS-MEM that avoid storing large, dense
  blocks of the MME that involve imputed genotypes. First, we will show this
  for the MME given in [10]. Second, we will present what we refer to as a
  hybrid type model (HM) that uses a MEM for the animals with marker genotypes
  and a BVM for animals without genotypes."

-   "$\mathbf{M}_{g}$ is the matrix of centered marker covariates for animals
  with genotypes, $\mathbf{M}_{n} =
  \mathbf{A}_{ng}\mathbf{A}_{gg}^{-1}\mathbf{M}_{g}$, is the matrix of imputed
  marker covariates for animals with missing genotypes, $\mathbf{\alpha}$ is
  the vector of random marker effects, $\mathbf{\epsilon}$ is the vector of
  imputation residuals with null means and covariance matrix proportional to
  the inverse of $\mathbf{A}^{nn}$, the sub-matrix corresponding to animals
  with missing genotypes in the inverse of the matrix $\mathbf{A}$ of
  pedigree-based additive relationships, and $\mathbf{e}$ is a vector of
  residuals. The matrix of imputed genotypes can be more efficiently computed
  by solving the sparse system of equations [10]:
  $\mathbf{A}^{nn}\mathbf{M}_{n} = - \mathbf{A}^{ng}\mathbf{M}_{g}$."

-   "The large dense matrix $\mathbf{M}_{n}$ appears in the MEM given by Model
  (1). This is avoided here by using a BVM for animals with missing genotypes
  rather than expressing their breeding values as the sum of the effects of
  their imputed marker gneotypes plus their separate imputation residuals. The
  advantages of the MEM such as allowing for alternative priors for marker
  effects are retained by still fitting a MEM but only for animals with
  genotypes."

-   Correlations between estimates of breeding values from the hybrid model and
  the regular model were 1.0 for non-genotyped animals and over 0.99 for
  genotyped animals.

### 2017-03-24 paper/manuscript.tex
Init results on inbred lines

@@
 \Genetics2level{Inbred lines}
+At the level of the 149 genotypes for which both, genomic (G) and 
+transcriptomic (T) information, were available, the use of T yielded higher
+predictive abilities than the use of G for four out of six traits
+\ref{table:InbredTable}.
+In terms of predictive ability, the absolute advantage of T over G was as large 
+as 0.04 as observed for the traits '100grainweight', 'Eardiameter' and 
+'Silkingtime'.
+Except for 'Kernelwidth', the predictive ability for predictions based on
+G and the full set of inbred lines ($n = 211$) was always higher than for the 
+reduced set of inbred lines ($n = 149$).
+The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
+prediction of all 211 lines performed at least as well as G alone when T was 
+superior over G for the reduced set of 149 lines.
+In the case of '100grainweight' and 'Eardiameter', the use of GT yielded a 
+slight improvement in predictive ability over G.
+Where T was a worse predictor for the reduced set of 149 lines than G,
+predictive abilities of GT for the full set of 211 lines were also worse than 
+those based on G alone.
 
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Wed Mar 22 08:40:11 2017
+% Fri Mar 24 11:34:47 2017
 % Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
 \begin{table}[ht]
-\label{tbl:InbredTable}
+\label{table:InbredTable}
 \centering
@@
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-G\_Full & 0.43 (0.0733) & 0.37 (0.1734) & 0.39 (0.0535) & 0.34 (0.0455) & 0.45 (0.0755) & 0.53 (0.0297) \\ 
-  GT\_Full & 0.44 (0.0748) & 0.33 (0.1727) & 0.4 (0.0543) & 0.34 (0.0425) & 0.4 (0.0707) & 0.53 (0.0285) \\ 
+G Full & 0.43 (0.073) & 0.37 (0.173) & 0.39 (0.053) & 0.34 (0.045) & 0.45 (0.076) & 0.53 (0.03) \\ 
+  GT Full & 0.44 (0.075) & 0.33 (0.173) & 0.4 (0.054) & 0.34 (0.043) & 0.4 (0.071) & 0.53 (0.028) \\ 
    \midrule
 \multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-G\_1.0 & 0.36 (0.0773) & 0.21 (0.1678) & 0.3 (0.0527) & 0.34 (0.0472) & 0.41 (0.078) & 0.45 (0.0302) \\ 
-  T\_1.0 & 0.4 (0.0747) & 0.15 (0.1568) & 0.37 (0.0509) & 0.36 (0.0423) & 0.32 (0.0692) & 0.49 (0.0279) \\ 
-  GT\_0.9 & 0.39 (0.0792) & 0.21 (0.1692) & 0.38 (0.0561) & 0.41 (0.0459) & 0.29 (0.075) & 0.49 (0.0311) \\ 
-  GT\_0.8 & 0.4 (0.0774) & 0.21 (0.1581) & 0.36 (0.0515) & 0.41 (0.0458) & 0.29 (0.0737) & 0.49 (0.0288) \\ 
-  GT\_0.7 & 0.37 (0.0734) & 0.1 (0.1555) & 0.31 (0.0508) & 0.37 (0.0422) & 0.33 (0.0755) & 0.52 (0.028) \\ 
-  GT\_0.6 & 0.4 (0.0739) & 0.13 (0.1563) & 0.27 (0.0497) & 0.37 (0.0417) & 0.42 (0.075) & 0.55 (0.0272) \\ 
-  GT\_0.5 & 0.41 (0.0671) & 0.16 (0.1549) & 0.31 (0.0486) & 0.4 (0.0414) & 0.44 (0.0748) & 0.49 (0.0278) \\ 
-  GT\_0.4 & 0.39 (0.0697) & 0.2 (0.1558) & 0.33 (0.0486) & 0.35 (0.0428) & 0.38 (0.0727) & 0.52 (0.028) \\ 
-  GT\_0.3 & 0.36 (0.0711) & 0.2 (0.1566) & 0.28 (0.0503) & 0.37 (0.044) & 0.42 (0.0736) & 0.5 (0.028) \\ 
-  GT\_0.2 & 0.35 (0.072) & 0.23 (0.1645) & 0.26 (0.0519) & 0.36 (0.0474) & 0.38 (0.0722) & 0.44 (0.0273) \\ 
-  GT\_0.1 & 0.29 (0.0762) & 0.21 (0.1659) & 0.23 (0.0522) & 0.32 (0.0471) & 0.4 (0.0743) & 0.39 (0.0302) \\ 
+G 1.0 & 0.36 (0.077) & 0.21 (0.168) & 0.3 (0.053) & 0.34 (0.047) & 0.41 (0.078) & 0.45 (0.03) \\ 
+  T 1.0 & 0.4 (0.075) & 0.15 (0.157) & 0.37 (0.051) & 0.36 (0.042) & 0.32 (0.069) & 0.49 (0.028) \\ 
+  GT 0.1 & 0.29 (0.076) & 0.21 (0.166) & 0.23 (0.052) & 0.32 (0.047) & 0.4 (0.074) & 0.39 (0.03) \\ 
+  GT 0.2 & 0.35 (0.072) & 0.23 (0.164) & 0.26 (0.052) & 0.36 (0.047) & 0.38 (0.072) & 0.44 (0.027) \\ 
+  GT 0.3 & 0.36 (0.071) & 0.2 (0.157) & 0.28 (0.05) & 0.37 (0.044) & 0.42 (0.074) & 0.5 (0.028) \\ 
+  GT 0.4 & 0.39 (0.07) & 0.2 (0.156) & 0.33 (0.049) & 0.35 (0.043) & 0.38 (0.073) & 0.52 (0.028) \\ 
+  GT 0.5 & 0.41 (0.067) & 0.16 (0.155) & 0.31 (0.049) & 0.4 (0.041) & 0.44 (0.075) & 0.49 (0.028) \\ 
+  GT 0.6 & 0.4 (0.074) & 0.13 (0.156) & 0.27 (0.05) & 0.37 (0.042) & 0.42 (0.075) & 0.55 (0.027) \\ 
+  GT 0.7 & 0.37 (0.073) & 0.1 (0.156) & 0.31 (0.051) & 0.37 (0.042) & 0.33 (0.075) & 0.52 (0.028) \\ 
+  GT 0.8 & 0.4 (0.077) & 0.21 (0.158) & 0.36 (0.051) & 0.41 (0.046) & 0.29 (0.074) & 0.49 (0.029) \\ 
+  GT 0.9 & 0.39 (0.079) & 0.21 (0.169) & 0.38 (0.056) & 0.41 (0.046) & 0.29 (0.075) & 0.49 (0.031) \\ 
+  GT A1 & 0.32 (0.079) & 0.03 (0.155) & 0.28 (0.053) & 0.32 (0.045) & 0.31 (0.073) & 0.46 (0.03) \\ 
+  GT A2 & 0.4 (0.081) & 0.18 (0.164) & 0.32 (0.055) & 0.33 (0.046) & 0.36 (0.073) & 0.47 (0.03) \\ 
+  GT A3 & 0.3 (0.071) & 0.1 (0.164) & 0.34 (0.054) & 0.41 (0.041) & 0.33 (0.072) & 0.41 (0.029) \\ 
    \bottomrule
@@
 \begin{table}[ht]
-\label{tbl:HybridTable}
+\label{table:HybridTable}
 \centering

### 2017-03-24 paper/manuscript.tex
Init hybrid results in manuscript

@@
 \Genetics2level{Hybrids}
+The maize hybrid data set contains information on six agronomic traits and
+three predictors, namely pedigree data (P), genomic information (G), 
+transcriptomic information (T), and combinations thereof.
+Pedigree records and genomic information were available for the 245 parent 
+lines of all 1,521 hybrids and we denote these data as the 'Full' set whereas
+transcriptomic data were available for 103 parent lines of 685 hybrids, which
+we denote as 'Core'.
+For dry matter yield (DMY), T was the best predictor in the core set of 685 
+hybrids with a predictive ability that was 0.08 points higher than for genomic 
+data, as the second best predictor for this trait.
+In the core set, P had the highest predictive ability for starch ('STA').
+For the other four traits, G had the best predictive abilities.
+In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
+markedly for all traits with the exception of 'FAT'.
+The combination of genomic and transriptomic information (GT) in a single-step
+model yielded a small improvement of G alone for the trait 'DMY'.
+All traits, where T was not superior to G in the core set, did not benefit from 
+the single-step approach when compared to genomic information alone.
+Predictive abilities resulting from combinations of pedigree with genomic (PG)
+information were better than P alone for all traits and PT improved upon P in
+the core set for four traits.
 
 % latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Thu Mar 16 08:16:30 2017
+% Fri Mar 24 12:24:30 2017
 % Script: ./analysis/uhoh_plot_predictive_abilities.R
@@
 As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.}
-\begin{tabular}{rllllll}
+\begin{tabular}{lllllll}
   \toprule
- & DMY & DMC & FAT & PRO & STA & SUG \\ 
+Predictor & DMY & DMC & FAT & PRO & STA & SUG \\ 
   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-P & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
-  G & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
-  PG & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
-  PT & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
-  GT & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
-   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-T & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
-  P & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
-  G & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
+P Full & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
+  G Full & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
+  PG Full & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
+  PT Full & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
+  GT Full & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
+  P Core & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
+  G Core & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
+  T Core & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
    \bottomrule
-\multicolumn{6}{l}{}\\
 \end{tabular}

### 2017-03-24 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -75,10 +75,13 @@ xtable_pred_df <- pred_df %>%
   spread(key = Trait, value = Value) %>%
   mutate(
     Predictor = paste0(Predictor, "_", Core_Fraction)
-  )
-
-xtable_pred_df %>%
-  mutate(Predictor = factor(Predictor, levels = pred_order))
+  ) %>% 
+  mutate(
+    Pred_Order = match(Predictor, pred_order)
+  ) %>% 
+  arrange(Pred_Order) %>% 
+  mutate(Predictor = gsub("_", replacement = " ", x = Predictor)) %>% 
+  select(-Pred_Order)
   
 pred_lst <- xtable_pred_df %>%
   split(.$Reduced) %>%

### 2017-03-24 analysis/uhoh_plot_predictive_abilities.R

@@ -64,18 +64,25 @@ raw_pred_df <- raw_pred_df %>%
   )
  
  
-pred_lst <- raw_pred_df %>%
+pred_df <- raw_pred_df %>%
   mutate(
-    Reduced = ifelse(Predictor == "T", yes = TRUE, no = FALSE),
-    Reduced = ifelse(Core_Fraction == "1.0", yes = TRUE, no = Reduced) 
+    Group = ifelse(Predictor == "T", yes = "Core", no = "Full"),
+    Group = ifelse(Core_Fraction == "1.0", yes = "Core", no = Group) 
   ) %>%
   select(-Core_Fraction) %>%
-  group_by(Trait, Predictor, Reduced) %>%
+  group_by(Trait, Predictor, Group) %>%
   summarize(
     r = cor(y, yhat),
     CV = coefficient_of_variation(var_yhat, yhat)
   ) %>%
-  ungroup() %>%
+  ungroup()
+
+ext_pred_order <- c(
+  "P_Full", "G_Full", "PG_Full", "PT_Full", "GT_Full",
+  "P_Core", "G_Core", "T_Core"
+)
+
+xtable_pred_df <- pred_df %>%
   mutate(
     r = round(r, digits = 2),
     CV = round(CV, digits = 3),
@@ -85,28 +92,34 @@ pred_lst <- raw_pred_df %>%
   ) %>%
   select(-r, -CV) %>%
   spread(key = Trait, value = Value) %>%
-  split(.$Reduced) %>%
-  map(., ~select(., -Reduced)) %>%
-  map(., ~column_to_rownames(., var = "Predictor"))
+  mutate(
+    Predictor = paste0(Predictor, "_", Group),
+    Pred_Order = match(Predictor, ext_pred_order)
+  ) %>% 
+  arrange(Pred_Order) %>% 
+  mutate(Predictor = gsub("_", replacement = " ", x = Predictor)) %>% 
+  select(-Group, -Pred_Order)
+ 
 
-# Prepare the split data for being displayed inside a LaTeX table.
-names(pred_lst)
-coverage_subheading <- "Only incomplete predictor:"
-attr(pred_lst, "subheadings") <- paste(coverage_subheading, names(pred_lst))
 hybrid_caption <- paste(
   "Predictive abilities and corresponding coefficients of variation for the",
   "set of maize hybrids. As predictors, pedigree (P), genomic (G),",
   "transcriptomic (T) data and combinations thereof we used."
 )
-xtable_lst <- xtableList(pred_lst,
[…]

### 2017-03-24 reports/methods_data.R

@@ -21,9 +21,28 @@ common_genotypes %>%
 
 # *** ./analysis/uhoh_data_preparation.R ***
 # Number of transcripts."
-"./data/derived/uhoh/mrna.RDS" %>%
-  readRDS() %>%
+uhoh_mrna <- "./data/derived/uhoh/mrna.RDS" %>%
+  readRDS() 
+
+uhoh_mrna %>%
   ncol()
+
+# Names of parental inbred lines in the UHOH maize hybrid data set that are 
+# covered by transcriptomic data.
+uhoh_mrna_nms <- rownames(uhoh_mrna)
+
+# Number of hybrids whose parents are covered by transcriptomic data.
+"./data/derived/uhoh/agro_tibble.RDS" %>%
+  readRDS() %>%
+  split(.$Trait) %>% 
+  map(., ~select(., Genotype)) %>% 
+  map(flatten_chr) %>% 
+  reduce(intersect) %>% 
+  as_data_frame() %>% 
+  separate(value, into = c("DF", "Dent", "Flint")) %>% 
+  filter(Dent %in% uhoh_mrna_nms,
+         Flint %in% uhoh_mrna_nms) %>% 
+  dim()
 
 # Number of pedigree records.
 "./data/derived/uhoh/pedigree_matrix.RDS" %>%

### 2017-03-24 paper/manuscript.tex
Init genetic space coverage in manuscript

@@
 of genotypes that display maximum average distance between each other.
-MR was computed using the \emph{rogers.dist} function from the R-pacakge
-\textit{poppr} \cite{Kamvar2015}.
 
@@
 predictive abilities than the use of G for four out of six traits
-\ref{table:InbredTable}.
+(Table \ref{table:InbredTable}).
 In terms of predictive ability, the absolute advantage of T over G was as large 
@@
 \begin{table}[ht]
-\label{table:InbredTable}
 \centering
@@
 The decimal after the predictor name indicates the size of the genetic material that was covered by inbred lines in the core set.}
+\label{table:InbredTable}
 \begin{tabular}{rllllll}
@@
 hybrids with a predictive ability that was 0.08 points higher than for genomic 
-data, as the second best predictor for this trait.
+data, as the second best predictor for this trait 
+(Table \ref{table:HybridTable}).
 In the core set, P had the highest predictive ability for starch ('STA').
@@
 \begin{table}[ht]
-\label{table:HybridTable}
 \centering
@@
 As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.}
+\label{table:HybridTable}
 \begin{tabular}{lllllll}
@@
 \Genetics2level{Coverage of the genetic space}
-
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
-\caption{
-Influence of core set composition on predictive abilities.
-Predictive abilities are based on single step predition with genomic data as the 
-complete predictor and transcriptomic data as the incomplete predictor.
-Results are shown for six agronomic traits and nine core sets of the maize 
-diversity panel.
-}
-\label{fig:CoreSetInfluence}
-\end{figure}
+Single-step approaches are based on the use of one complete predictor that
+covers all genotypes and one incomplete predictor that covers only a subset of
+all available genotypes.
+Here, we explore the impact of which data are covered by the incomplete 
+predictor on the predictive abilities obtained via single-step approaches by
+focusing exclusively on the 149 inbred lines from the maize diversity panel
+that are covered by both, genomic and transcriptomic information.
+First, we assmbled nine different core sets, which we defined as subsets
+of genotypes that maximize the average genetic distance between genotypes of
+the full set.
+The nine core sets differ in that each covers a different number of genotypes,
+raning from only 10\% to to 90\% of all genotypes covered by the incomplete
+predictor (Fig. \ref{fig:CoreSetPCA}).
 
@@
 
+Predictive abilities obtained with the different core sets varied widely within
+traits (Fig. \ref{fig:CoreSetInfluence}, Table \ref{table:InbredTable}).
+For the traits 'cobweight' and 'Plantheight', which were better predicted via G
+than via T in the reduced set of lines, predictive abilities were higher when
+only a small number of genotypes was included in the core set
+For most other traits, an initial increase in predictive abilities when moving
+from a core set size of 0.1 to 0.5/0.6 was observed, followed by a drop around
+0.6/0.7 and a subsequent hike up to a core set size of 0.9.
+
+\begin{figure}[H]
+\centering
+\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
+\caption{
+Influence of core set composition on predictive abilities.
+Predictive abilities are based on single step predition with genomic data as the 
+complete predictor and transcriptomic data as the incomplete predictor.
+Results are shown for six agronomic traits and nine core sets of the maize 
+diversity panel.
+}
+\label{fig:CoreSetInfluence}
+\end{figure}
+
+As an alternative to exploring the impact of genotypes sampled based on the
+maximization of genetic distance within the core set, we wanted to explore a
+more extreme scenario where we select core sets based on population structure.
+Hereto, we ran an admixture analysis on the same set of 149 inbred lines and
+display the estimated ancestry coefficients for the genotypes assuming $K=3$
+ancestral populations (Fig. \ref{fig:PopStructure}).
+
 \begin{figure}[H]
@@
 \end{figure}
+
+Of the 149 genotypes, 17 belonged to more than 50\% to the first ancestral
+population (A1), 109 belonged to more than 50\% to the second ancestral
+population (A2) and 19 belonged to more than 50\% to the third ancestral
+population while four genotypes could not be unequivocally assigned to either
+ancestral population (mixed).
+To adjust the size of the set of genotypes largely belonging to A2 to the other
+two sets, we selected 19 genotypes from A2 using previously described core
+sampling.
+For single-step prediction, each of the three genotype sets were then declared
+as having genomic and transcriptomic information whereas their complements were
+covered by transcriptomic data only.
+Genotypes from A1 and A3, which were at the far ends in the PCA plot (Fig.
+\ref{fig:PopStructure}B), were estimated with less precision than genotypes
+from the central cluster of genotypes belonging largely to the A2 population
+(Fig. \ref{fig:PopStructure}A, Table \ref{table:InbredTable}).
+
+A principal component analysis, carried out separately on the two parental
+heterotic groups of the maize hybrids, showed that the entire genetic space in
+each group was well covered by genotypes that had both, transcriptomic and
+genomic data (Fig. \ref{fig:UHOH-PCA}).
 

### 2017-03-24 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -131,7 +131,7 @@ g1 <- pred_df %>%
 plot_name <- "./paper/tables_figures/core_fraction_predictive_ability_trend.pdf"
 ggsave(plot = g1, 
        filename = plot_name,
-       width = 4,
+       width = 7,
        height = 4,
        units = "in")
   

### 2017-03-27 paper/manuscript.tex
Init manuscript intro

@@
 \section{Introduction} 
+% Necessity for predictions.
+Genomic prediction, pioneered by \citeNP{Meuwissen2001}, has revolutionized
+multiple areas of genetic analyses such as animal and plant breeding
+\cite{DeLosCampos2013} as well as disease diagnostics in humans 
+\cite{DeLosCampos2010,Vazquez2016}.
+The primary application of genomic prediction in breeding programs is for the 
+accelerated selection of promising candidates without phenotypes.
 
-$\dots$
+% -   Describe the two step procedure and its shortcomings.
+% Incomplete coverage of individuals with genomic data.
+In animal breeding, only a small fraction of all animals that are recorded in
+the pedigree also have genotypes.
+To utilize all available phenotpyic information for model training methods were 
+needed that would allow for a combined use of genomic and pedigree information 
+\cite{Hayes2009a,VanRaden2009}.
 
+% Two-step prediction.
+In the early years of genomic prediction, a two-step procedure was the most
+intuitive choice for establishing a new selection scheme.
+The first step constists of conventionally estimating breeding values based on
+pedigree relationships and therefore the phenotypes of relatives.
+Individuals with accurate breeding values out of the first step (\textit{i.e.}
+high number of offspring), enter the second step in which conventional breeding
+values are regressed onto SNP genotypes. 
+Those markers can then be used to predict genomic breeding values for solely
+genotyped and potentially very young individuals.
+
+% Shortcomings of the two-step procedure.
+\begin{itemize}
+  \item Bias introduced by genomic information giving an advantage to young,
+    genotyped individuals over old animals with only pedigree records
+    \cite{Aguilar2009}.
+  \item Pedigree depth varies among individuals; some animals have more
+    ancestors and therefore more precise estimates \cite{Aguilar2009}.
+\end{itemize}
+
+% Problems with SS-BLUP
+\begin{itemize}
+  \item Weight $w$ for $\mathbf{A}$ and $\mathbf{G}$.
+  \item The optimal weight differs between traits \cite{Ashraf2016}.
+  \item compatibility between $\mathbf{A}$ and $\mathbf{G}$
+    \cite{Christensen2012}
+\end{itemize}
+
+% Workings and advantages of the single step procedure.
+Independently from each other, \citeNP{Legarra2009} and
+\citeNP{Christensen2010} developed a BLUP method that blends the numerator
+relationship matrix $\mathbf{A}$ with the genomic relationship matrix
+$\mathbf{G}$ in a mutual matrix $\mathbf{H}$.
+
+% Physiological epistasis (added benefits of downstream 'omics' predictors)
+Compared to pedigree-based prediction, genomic prediction captures Mendelian
+sampling, which is the difference between the expected and the realized
+relationship between individuals.
+Despite its great success and benefits, genomic prediction does not seem to
+capture all physiological epistasis \cite{Sackton2016,Reif2015}, which rather 
+manifests itself in the form of main and interaction effects of and between 
+downstream 'omics' strata, such as the transcriptome, the proteome and the 
+metabolome.
+This has lead to the exploration of
+
+Besides genomic data, other predictors have proven their value in the
+prediction of complex traits in plant populations \cite{Guo2016,Westhues2017}
+and in humans \cite{Vazquez2016}.
+
+% Special situation in hybrid breeding.
+Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
+maize breeding --- is a particularly challenging application field for
+prediction tools.
+Here, $2n$ parent individuals from two genetically distinct heterotic groups are
+crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
+require intensive field testing.
+In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
+technology \cite{Wedzony2009} now allows for an annual production of 1,000 
+parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
+progeny.
+
+% -   Mention scarcity of data on single step prediction in plant populations.
+Single step prediction has been a hot topic in animal breeding for a few years,
+but so far, to the best of our knowledge, has only been considered in plants
+only in one study on wheat lines from CIMMYT \cite{Ashraf2016}. 
+So far, its applications in hybrid populations are limited to pig breeding
+\cite{Xiang2015,Xiang2016,Tusell2016}.
+% Are there applications in chicken breeding? (Hints in Christensen et al.
+% 2014, GSE).
 Our objectives were threefold: (1) Examine whether the single step prediction
@@
 ability.
+
+% -   How extensive are pedigree records in commercial breeding companies?
 
@@
 \section*{Discussion}
-\Genetics2level{Some discussion section}
+\Genetics2level{Advantages of marker effect models}
+% Describe the many hacks that have been developed to tackle problems with BLUP
+% models.
 
@@
 
-\subsection{More detailed discussion}
+\Genetics2level{Superiority of single step models}
 
-$\dots$
+\Genetics2level{Coverage of the genetic space}
+Animal populations are so large that it does not truly matter, which animals are 
+being genotyped, but in plant breeding the situation is different and we would 
+be interested in determining which individuals should be complemented with 
+information on another predictor.
 
-\Genetics2level{Another discussion section}
+Prediction of phenotypic performance is particularly crucial in hybrid breeding
+programs because the number of potential selection candidates is $n^{2}$ where
+$n$ corresponds to the number of parent lines.
 
-$\dots$
+Given recent results on the utility of downstream 'omics' predictors in a study
+on hybrid prediction in a European hybrid maize panel \cite{Westhues2017}, our 
+objectives were i) to explore the benefits of a single step procedure for the 
+same material and ii) to examine the influence on the composition of the set of
+genotypes that is covered by more than one predictor on the success of single
+step prediction.
 

### 2017-03-27 paper/notes.Rmd
Check whether repo transfer to Github worked
Init manuscript intro

@@
 -   Optimized the weight of genomic and pedigree information in blending the
-  two matrices by fitting them simultaneous as random effects in a mixed model.
+  two matrices by fitting them simultaneously as random effects in a mixed 
+  model.
 
@@
   the additive relationship matrix constructed from pedigree by an additive
-  relatoinship matrix that combines the marker-based relationship matrix for
+  relationship matrix that combines the marker-based relationship matrix for
   genotyped animals with the pedigree-based relationship matrix."
@@
   populations that are undergoing selection and non-random mating. This is
-  improtant, because genotypes are usually collected only on superior animals
+  important, because genotypes are usually collected only on superior animals
   and this can lead to biased evaluations. A properly calculated BLUP

### 2017-03-27 paper/manuscript.tex
First, complete draft of intro

--- 2017-03-24 paper/manuscript.tex
@@
 \section{Introduction} 
+% Rationale for predictions.
+Genomic prediction, pioneered by \citeNP{Meuwissen2001}, has revolutionized
+animal and plant breeding by accelerating the selection of promising candidates 
+without phenotypes \cite{DeLosCampos2013,Garcia-Ruiz2016} and recently emerged
+as an important tool in human diagnostics \cite{DeLosCampos2010,Vazquez2016}.
+Its use is largely motivated by the high costs and oftentimes long duration of 
+phenotyping, which can be reduced when covering all genotypes with a predictor 
+while limiting phenotyping to a few genotypes with high precision.
+The performance of all other genotypes is then forecast using model parameters
+obtained from thorough (repeated) $k$-fold cross-validation routines applied to 
+the subset of individuals with full information.
+In animal breeding, genomic selection has increased the annual selection gain 
+by $\approx$ 50-100\% for yield traits and by 300-400\% for traits with low 
+heritability \cite{Garcia-Ruiz2016}.
+In plant breeding, the relative excellence of a breeding strategy largely 
+depends on the prediction accuracy, which may vary widely among traits and 
+crops \cite{Longin2015}.
 
-$\dots$
+% Search for better predictors
+Compared to pedigree information, which represent the expected relationship
+among individuals, genomic information capture Mendelian sampling and thereby 
+provide an improved proxy of the realized relationship among individuals.
+Nevertheless, genomic data do not exhaustively capture physiological epistasis 
+\cite{Jiang2015,Guo2016,Vazquez2016}, describing the interactions 
+within and between different biological strata \cite{Sackton2016}.
+Such interactions were found to be pervasive throughout the genomes of yeast 
+\cite{Brem2005} and humans \cite{Brown2014} and have motivated studies on the 
+utility of downstream 'omics' predictors for integrating such interactions.
+So far, encouraging results have been found for humans \cite{Vazquez2016}, maize 
+inbred lines \cite{Guo2016} and maize hybrids \cite{Westhues2017}.
 
+% Incomplete predictors.
+Breeding programs --- particularly in animal breeding --- are characterized by
+the availability of one predictor type, which is available for all recorded
+genotypes while other predictors are incomplete in that they cover only a 
+subset of all genotypes.
+Usually, the complete predictor is pedigree information while genomic data are
+the most prevalent incomplete predictor.
+To utilize all available phenotypic information for model training, methods were 
+needed that would allow for a combined use of genomic and pedigree information 
+\cite{Hayes2009a,VanRaden2009}.
+
+% Two-step prediction.
+In the early years of genomic prediction, a two-step procedure was the most
+intuitive choice for establishing a new selection scheme.
+The first step constists of conventionally estimating breeding values based on
+pedigree relationships and the phenotypes of relatives.
+Individuals with accurate breeding values out of the first step (\textit{i.e.}
+high number of offspring), enter the second step in which conventional breeding
+values are regressed onto SNP genotypes.
+The resulting model coefficients can then be used to predict genomic breeding 
+values for solely genotyped and potentially very young individuals.
+However, as the breeding values from step one enter the genomic prediction
+model as phenotypes, one has to be careful about the residual error
+distribution of such a model.
+For instance, if the range of the prediction errors will not be identically
+distributed anymore, \textit{i.e.} they should be weighted differently or
+estimated individually \cite{Aguilar2010}.
+One popular approach to account for this problem is deregressing the breeding
+values and weighting the residuals according to the prediction error variances
+\cite{Garrick2009}.
+
+% Workings and advantages of the single step procedure.
+Independently from each other, \citeNP{Legarra2009} and
+\citeNP{Christensen2010} developed a single step BLUP method that blends the 
+numerator relationship matrix $\mathbf{A}$ with the genomic relationship matrix
+$\mathbf{G}$ in a mutual matrix $\mathbf{H}$ to use all predictor information
+simultaneously.
+Issues with this H-BLUP approach were the compatability of $\mathbf{A}$ with
+$\mathbf{G}$ \cite{Christensen2012} and their weighting, which is not
+trait-independent \cite{Ashraf2016}.
+\citeNP{Fernando2014} derived the equivalent single step marker-effect-model to 
+the single step breeding value model of \citeNP{Legarra2009} and 
+\citeNP{Christensen2010}, which has the advantage of not requiring weights for
+$\mathbf{A}$ and $\mathbf{G}$ as well as the ability to model the structure of 
+the imputation error.
+
+% -   Mention scarcity of data on single step prediction in plant populations.
+Single step prediction has already become a routine procedure in animal
+breeding but, to the best of our knowledge, has only been considered in plants
+in a study on wheat lines \cite{Ashraf2016}. 
 Our objectives were threefold: (1) Examine whether the single step prediction
-framework can be transfered to hybrids, (2) explore the utility of single step 
-predition with a quantitative predictor and (3) evaluate the impact of the
-particular genetic space covered by the incomplete predictor on predictive
-ability.
+framework can be transfered to plant hybrids, (2) explore the utility of single 
+step predition with a quantitative predictor and (3) evaluate the impact of the
+genetic space covered by the incomplete predictor on predictive ability.
 
@@
 
-\subsection{More detailed discussion}
+\Genetics2level{Superiority of single step models}
 
-$\dots$
+\Genetics2level{Coverage of the genetic space}
+Animal populations are so large that it does not truly matter, which animals are 
+being genotyped, but in plant breeding the situation is different and we would 
+be interested in determining which individuals should be complemented with 
+information on another predictor.
 
-\Genetics2level{Another discussion section}
+Prediction of phenotypic performance is particularly crucial in hybrid breeding
+programs because the number of potential selection candidates is $n^{2}$ where
+$n$ corresponds to the number of parent lines.
 
-$\dots$
+Given recent results on the utility of downstream 'omics' predictors in a study
+on hybrid prediction in a European hybrid maize panel \cite{Westhues2017}, our 
+objectives were i) to explore the benefits of a single step procedure for the 
+same material and ii) to examine the influence on the composition of the set of
+genotypes that is covered by more than one predictor on the success of single
+step prediction.
+
+% Special situation in hybrid breeding.
+Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
+maize breeding --- is a particularly challenging application field for
+prediction tools.
+Here, $2n$ parent individuals from two genetically distinct heterotic groups are
+crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
+require intensive field testing.
+In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
+technology \cite{Wedzony2009} now allows for an annual production of 1,000 
+parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
+progeny.
+
+So far, its applications in hybrid populations are limited to pig breeding
+\cite{Xiang2015,Xiang2016,Tusell2016}.
+% Are there applications in chicken breeding? (Hints in Christensen et al.
+% 2014, GSE).
 

### 2017-03-28 paper/manuscript.tex
Add minor points to discussion

@@
 \section*{Discussion}
+
 \Genetics2level{Advantages of marker effect models}
@@
 \Genetics2level{Superiority of single step models}
+\subsection{Inbred lines}
+Predictive abilities for the maize diversity panel are not in complete
+agreement with those reported by \citeNP{Guo2016}.
+Reasons for this may be threefold: 
+1) In our study, we reduced the dataset to the 211 tropical/subtropical lines 
+and excluded all other genotypes whereas \citeNP{Guo2016} used genotypes from 
+four subpopulations and only a subset of 157 genotypes from the set of 
+tropical/subtropical lines while correcting for population structure by
+including a fixed effect in the model for predicting agronomic performance.
+2) We applied quality checks for the predictor data after generating the subset
+of 211 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
+genotypes.
+3) While \citeNP{Guo2016} used five-fold cross-validation with 500 repetitions,
+we employed a leave-one-out cross-validation scheme (LOOCV).
 
-\Genetics2level{Coverage of the genetic space}
-Animal populations are so large that it does not truly matter, which animals are 
-being genotyped, but in plant breeding the situation is different and we would 
-be interested in determining which individuals should be complemented with 
-information on another predictor.
-
-Prediction of phenotypic performance is particularly crucial in hybrid breeding
-programs because the number of potential selection candidates is $n^{2}$ where
-$n$ corresponds to the number of parent lines.
-
-Given recent results on the utility of downstream 'omics' predictors in a study
-on hybrid prediction in a European hybrid maize panel \cite{Westhues2017}, our 
-objectives were i) to explore the benefits of a single step procedure for the 
-same material and ii) to examine the influence on the composition of the set of
-genotypes that is covered by more than one predictor on the success of single
-step prediction.
-
-% Special situation in hybrid breeding.
+\subsection{Maize hybrids}
 Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
@@
 progeny.
-
 So far, its applications in hybrid populations are limited to pig breeding
@@
 % 2014, GSE).
+
+\begin{itemize}
+  \item Generally, it is important that, on the same set of genotypes, the 
+    incomplete predictor yields larger predictive abilities than the complete 
+    predictor.
+  \item In hybrid breeding, the probability of successfully selecting the best
+    observed genotypes based on the best predicted candidates is a strongly 
+    convex function of the predictive ability.
+    Hence, even the minor gains when using transcriptomic and genomic data in a
+    single step might justify additional investments in RNA-seq at least of a
+    subste of genotypes that represents the genetic space of the breeding 
+    material well.
+  \item For the hybrid data, the predictive abilities of genomic information
+    for the whole set of available parent lines rose considerably for several
+    agronomic traits.
+    Hence, the inclusion of trancsriptomic data could not further improve upon
+    this result.
+\end{itemize}
+
+\Genetics2level{Coverage of the genetic space}
+Animal populations are so large that it does not truly matter, which animals are 
+being genotyped, but in plant breeding the situation is different and we would 
+be interested in determining which individuals should be complemented with 
+information on another predictor.
+If one can cover the genetic space sufficently with the incomplete predictor,
 

### 2017-03-28 paper/manuscript.tex
Substitute tables with figures, use r from imputed genos only

@@
 predictive abilities than the use of G for four out of six traits
-(Table \ref{table:InbredTable}).
+(Table \ref{table:InbredResults}).
 In terms of predictive ability, the absolute advantage of T over G was as large 
@@
 
-% latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Fri Mar 24 11:34:47 2017
 % Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
-\begin{table}[ht]
+\begin{figure}[H]
   \centering
-\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel.
-As predictors, genomic (G) data, transcriptomic (T) data and their combination (GT) were used.
-The decimal after the predictor name indicates the size of the genetic material that was covered by inbred lines in the core set.}
-\label{table:InbredTable}
-\begin{tabular}{rllllll}
-  \toprule
- & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
-  \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
-G Full & 0.43 (0.073) & 0.37 (0.173) & 0.39 (0.053) & 0.34 (0.045) & 0.45 (0.076) & 0.53 (0.03) \\ 
-  GT Full & 0.44 (0.075) & 0.33 (0.173) & 0.4 (0.054) & 0.34 (0.043) & 0.4 (0.071) & 0.53 (0.028) \\ 
-   \midrule
-\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
-G 1.0 & 0.36 (0.077) & 0.21 (0.168) & 0.3 (0.053) & 0.34 (0.047) & 0.41 (0.078) & 0.45 (0.03) \\ 
-  T 1.0 & 0.4 (0.075) & 0.15 (0.157) & 0.37 (0.051) & 0.36 (0.042) & 0.32 (0.069) & 0.49 (0.028) \\ 
-  GT 0.1 & 0.29 (0.076) & 0.21 (0.166) & 0.23 (0.052) & 0.32 (0.047) & 0.4 (0.074) & 0.39 (0.03) \\ 
-  GT 0.2 & 0.35 (0.072) & 0.23 (0.164) & 0.26 (0.052) & 0.36 (0.047) & 0.38 (0.072) & 0.44 (0.027) \\ 
-  GT 0.3 & 0.36 (0.071) & 0.2 (0.157) & 0.28 (0.05) & 0.37 (0.044) & 0.42 (0.074) & 0.5 (0.028) \\ 
-  GT 0.4 & 0.39 (0.07) & 0.2 (0.156) & 0.33 (0.049) & 0.35 (0.043) & 0.38 (0.073) & 0.52 (0.028) \\ 
-  GT 0.5 & 0.41 (0.067) & 0.16 (0.155) & 0.31 (0.049) & 0.4 (0.041) & 0.44 (0.075) & 0.49 (0.028) \\ 
-  GT 0.6 & 0.4 (0.074) & 0.13 (0.156) & 0.27 (0.05) & 0.37 (0.042) & 0.42 (0.075) & 0.55 (0.027) \\ 
-  GT 0.7 & 0.37 (0.073) & 0.1 (0.156) & 0.31 (0.051) & 0.37 (0.042) & 0.33 (0.075) & 0.52 (0.028) \\ 
-  GT 0.8 & 0.4 (0.077) & 0.21 (0.158) & 0.36 (0.051) & 0.41 (0.046) & 0.29 (0.074) & 0.49 (0.029) \\ 
-  GT 0.9 & 0.39 (0.079) & 0.21 (0.169) & 0.38 (0.056) & 0.41 (0.046) & 0.29 (0.075) & 0.49 (0.031) \\ 
-  GT A1 & 0.32 (0.079) & 0.03 (0.155) & 0.28 (0.053) & 0.32 (0.045) & 0.31 (0.073) & 0.46 (0.03) \\ 
-  GT A2 & 0.4 (0.081) & 0.18 (0.164) & 0.32 (0.055) & 0.33 (0.046) & 0.36 (0.073) & 0.47 (0.03) \\ 
-  GT A3 & 0.3 (0.071) & 0.1 (0.164) & 0.34 (0.054) & 0.41 (0.041) & 0.33 (0.072) & 0.41 (0.029) \\ 
-   \bottomrule
-\multicolumn{6}{l}{}\\
-\end{tabular}
-\end{table}
+  \includegraphics{./tables_figures/inbred_line_combi_plot.pdf}
+  \caption{
+  Predictive abilities and corresponding coefficients of variation for 
+  different core sets from the tropical/subtropical material of the maize 
+  diversity panel.
+  As predictors, genomic (G) data, transcriptomic (T) data and their 
+  combination (GT) were used.
+  The decimal after the predictor name indicates the size of the genetic 
+  material that was covered by inbred lines in the core set.
+  Influence of core set composition on predictive abilities.
+  Predictive abilities are based on single step predition with genomic data as the 
+  complete predictor and transcriptomic data as the incomplete predictor.
+  Results are shown for six agronomic traits and nine core sets of the maize 
+  diversity panel.
+  } 
+  \label{fig:InbredResults}
+\end{figure}
 
@@
 data, as the second best predictor for this trait 
-(Table \ref{table:HybridTable}).
+(Table \ref{fig:HybridResults}).
 In the core set, P had the highest predictive ability for starch ('STA').
@@
 % Script: ./analysis/uhoh_plot_predictive_abilities.R
-\begin{table}[ht]
+\begin{figure}[H]
 \centering
-\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids.
-As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.}
-\label{table:HybridTable}
-\begin{tabular}{lllllll}
-  \toprule
-Predictor & DMY & DMC & FAT & PRO & STA & SUG \\ 
-  \midrule
-P Full & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
-  G Full & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
-  PG Full & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
-  PT Full & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
-  GT Full & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
-  P Core & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
-  G Core & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
-  T Core & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
-   \bottomrule
-\end{tabular}
-\end{table}
+  \includegraphics{./tables_figures/hybrid_predictive_ability.pdf}
+  \caption{
+    Predictive abilities and corresponding coefficients of variation for the 
+    set of maize hybrids.
+    As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
+    combinations thereof we used.
+  } 
+  \label{fig:HybridResults}
+\end{figure}
 
@@
 Predictive abilities obtained with the different core sets varied widely within
-traits (Fig. \ref{fig:CoreSetInfluence}, Table \ref{table:InbredTable}).
+traits (Fig. \ref{fig:InbredResults}C).
 For the traits 'cobweight' and 'Plantheight', which were better predicted via G
@@
 0.6/0.7 and a subsequent hike up to a core set size of 0.9.
-
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
-\caption{
-Influence of core set composition on predictive abilities.
-Predictive abilities are based on single step predition with genomic data as the 
-complete predictor and transcriptomic data as the incomplete predictor.
-Results are shown for six agronomic traits and nine core sets of the maize 
-diversity panel.
-}
-\label{fig:CoreSetInfluence}
-\end{figure}
 
@@
 from the central cluster of genotypes belonging largely to the A2 population
-(Fig. \ref{fig:PopStructure}A, Table \ref{table:InbredTable}).
+(Fig. \ref{fig:PopStructure}A, Table \ref{fig:InbredTable}).
 

### 2017-03-28 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -1,7 +1,7 @@
 # LOOCV-Prediction
 if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "data.table", "dtplyr", "xtable", "stringr",
-               "forcats", "ggthemes", "viridis", "gsubfn")
+               "forcats", "ggthemes", "viridis", "gsubfn", "cowplot")
 source("./software/prediction_helper_functions.R")
 
 log_path <- "./data/derived/uhoh_maizego_prediction_log.txt"
@@ -32,7 +32,7 @@ raw_pred_df <- filtered_log_file %>%
   droplevels() %>%
   select(-Job_ID)
   
-abbrev_pred_df <- raw_pred_df %>%
+pred_df <- raw_pred_df %>%
   mutate(
     Predictor = gsubfn(
       pattern = "\\S+",
@@ -42,144 +42,123 @@ abbrev_pred_df <- raw_pred_df %>%
         "snp-mrna" = "GT"
       ),
       x = Predictor
+    ),
+    Trait = gsubfn(
+      pattern = "\\S+",
+      replacement = list(
+        "Plantheight" = "PH",
+        "Eardiameter" = "ED",
+        "100grainweight" = "GW",
+        "cobweight" = "CW",
+        "Kernelwidth" = "KW",
+        "Silkingtime" = "DS"
+      ),
+      x = Trait
     )
   )
 
-
-abbrev_pred_df %>% 
-  filter(Core_Fraction %in% seq(from = 0.1, to = 0.9, by = 0.1)) %>% 
+# Get the names of genotypes that were imputed based on core genotypes.
+imputed_genotypes <- pred_df %>% 
+  filter(Core_Fraction %in% c("Full", "1.0")) %>% 
   split(.$Core_Fraction) %>% 
   map(., ~select(., Geno)) %>% 
-  map(., ~unique(.)) %>% 
-  str()
-
- # Determine which genotypes were imputed and which weren't.
-core_genotypes <- abbrev_pred_df %>% 
-  filter(Core_Fraction == "1.0") %>% 
-  select(Geno) %>% 
-  unique() %>% 
-  flatten_chr()
-
-core_lst <- "./data/derived/maizego/augmented_core_list.RDS" %>% 
-  readRDS()
-
-imputed_genotypes <- core_lst %>% 
-  set_names(gsub("selected_fraction_", replacement = "", x = names(.))) %>% 
-  keep(names(.) %in% seq(from = 0.1, to = 0.9, by = 0.1)) %>% 
[…]

### 2017-03-28 analysis/maizego_snp_analyses.R

@@ -213,7 +213,7 @@ g4 <- structure_df %>%
   ggplot(aes(x = G, y = Value, fill = Component)) +
   geom_bar(stat = "identity", width = 1) +
   facet_wrap(~K, ncol = 1) +
-  scale_fill_viridis(discrete = TRUE) +
+  scale_fill_manual(values = c("#258039", "#F5BE41", "#31A9B8")) +
   ggthemes::theme_pander(base_size = 10) +
   theme(
     axis.text.x = element_blank(),
@@ -233,7 +233,7 @@ g5 <- pc_df %>%
   )) %>% 
   ggplot(aes(x = PC1, y = PC2, color = Ancestor, shape = Ancestor)) +
   geom_point(size = 2) +
-  scale_color_manual(values = c(viridis(n = 3), "red")) +
+  scale_color_manual(values = c("#258039", "#F5BE41", "#31A9B8", "red")) +
   scale_shape_manual(values = c(15, 17, 19, 3)) +
   guides(color = guide_legend(override.aes = list(size = 3))) +
   ggthemes::theme_pander(base_size = 10)

### 2017-03-28 analysis/uhoh_plot_predictive_abilities.R

@@ -97,7 +97,11 @@ plot_data <- pred_df %>%
   set_names(c("Core", "Imputed")) %>% 
   bind_rows(.id = "Group") %>% 
   unite(col = Exact_Predictor, Predictor, Group, sep = "_", remove = FALSE) %>% 
-  mutate(Pred_Order = match(Exact_Predictor, ext_pred_order)) %>% 
+  mutate(
+    Pred_Order = match(Exact_Predictor, ext_pred_order),
+    Predictor = factor(Predictor, levels = c(
+      "P", "G", "T", "PG", "PT", "GT"
+    ))) %>% 
   arrange(Pred_Order)
 
 # Define the top and bottom of the errorbars
@@ -105,16 +109,24 @@ limits <- aes(ymax = r + CV, ymin = r - CV)
 # Because the bars and errorbars have different widths
 # we need to specify how wide the objects we are dodging are
 dodge <- position_dodge(width = 0.9)
+# Predefine the color scheme.
+mycols <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6)
+names(mycols) <- c("P", "G", "T", "PG", "PT", "GT")
 
-plot_data %>% 
+
+g1 <- plot_data %>% 
   ggplot(aes(x = Trait, y = r, fill = Predictor)) +
-  geom_bar(stat = "identity", position = dodge) +
+  geom_bar(stat = "identity", position = dodge, color = "black") +
   geom_errorbar(limits, position = dodge, width = 0.25) +
   facet_grid(. ~ Group) +
-  scale_fill_viridis(discrete = TRUE) +
+  scale_fill_manual(values = mycols) +
   theme_base()
   
-
+ggsave(plot = g1, 
+       filename = "./paper/tables_figures/hybrid_predictive_ability.pdf",
+       width = 7,
+       height = 4,
+       units = "in")
 
 
 

### 2017-03-28 analysis/maizego_core_fraction_plot_predictive_ability.R
Use predability from all genotypes, not just imputed ones

@@ -58,13 +58,13 @@ pred_df <- raw_pred_df %>%
   )
 
 # Get the names of genotypes that were imputed based on core genotypes.
-imputed_genotypes <- pred_df %>% 
-  filter(Core_Fraction %in% c("Full", "1.0")) %>% 
-  split(.$Core_Fraction) %>% 
-  map(., ~select(., Geno)) %>% 
-  map(flatten_chr) %>% 
-  .[c("Full", "1.0")] %>% 
-  reduce(setdiff)
+#imputed_genotypes <- pred_df %>% 
+#  filter(Core_Fraction %in% c("Full", "1.0")) %>% 
+#  split(.$Core_Fraction) %>% 
+#  map(., ~select(., Geno)) %>% 
+#  map(flatten_chr) %>% 
+#  .[c("Full", "1.0")] %>% 
+#  reduce(setdiff)
  
 
 
@@ -78,8 +78,8 @@ dodge <- position_dodge(width = 0.9)
 plot_a_data <- pred_df %>% 
   filter(Core_Fraction %in% c("Full", "1.0")) %>% 
   split(.$Core_Fraction) %>% 
-  set_names(c("Core", "Imputed")) %>% 
-  map_at("Imputed", .f = ~filter(., Geno %in% imputed_genotypes)) %>% 
+  set_names(c("Core", "Full")) %>% 
+#  map_at("Imputed", .f = ~filter(., Geno %in% imputed_genotypes)) %>% 
   bind_rows(.id = "Group") %>% 
   group_by(Trait, Predictor, Group) %>% 
   summarize(
@@ -143,7 +143,6 @@ plot_c <- plot_c_data %>%
   ggplot(aes(x = `Core Fraction`, y = r, color = Trait, group = Trait)) +
   geom_line(stat = "identity", position = dodge_c) +
   geom_errorbar(limits, position = dodge_c, width = 0.25) +
-  #facet_wrap(~ Trait) +
   scale_color_tableau() +
   theme_pander(base_size = 10) +
   theme(legend.position = "top") +

### 2017-03-28 analysis/maizego_snp_analyses.R

@@ -91,7 +91,7 @@ g1 <- all_geno_pc %>%
   ggplot(aes(x = PC1, y = PC2, color = Group, shape = Group)) +
   geom_point() +
   scale_color_tableau() +
-  theme_bw(base_size = 10) +
+  theme_pander(base_size = 10) +
   theme(legend.position = "top")
 ggsave(plot = g1,
        filename = "./paper/tables_figures/maizego_pca.pdf",
@@ -142,9 +142,9 @@ g2 <- structure_df %>%
   geom_bar(stat = "identity", width = 1) +
   facet_wrap(~K, ncol = 1, strip.position = "right") +
   scale_fill_viridis(discrete = TRUE) +
-  ggthemes::theme_pander(base_size = 10) +
+  theme_pander(base_size = 10) +
   theme(axis.text.x = element_blank(),
-        #legend.position = "top") +
+        axis.ticks = element_blank(),
         legend.position = "none") +
   xlab("Genotype") +
   ylab("Ancestry Coefficient")
@@ -159,9 +159,9 @@ g3 <- pc_df %>%
   ggplot(aes(x = PC1, y = PC2, color = `Core Group`, shape = `Core Group`)) +
   geom_point(size = 1) +
   guides(color = guide_legend(override.aes = list(size = 3))) +
-  ggthemes::scale_color_tableau() +
+  scale_color_tableau() +
   facet_wrap(~Fraction, nrow = 3, ncol = 3) +
-  theme_bw(base_size = 10) +
+  theme_pander(base_size = 10) +
   theme(
     legend.position = "top",
     strip.background = element_blank(),
@@ -214,9 +214,10 @@ g4 <- structure_df %>%
   geom_bar(stat = "identity", width = 1) +
   facet_wrap(~K, ncol = 1) +
   scale_fill_manual(values = c("#258039", "#F5BE41", "#31A9B8")) +
-  ggthemes::theme_pander(base_size = 10) +
+  theme_pander(base_size = 10) +
   theme(
     axis.text.x = element_blank(),
+    axis.ticks = element_blank(),
     legend.position = "none",
     strip.background = element_blank(),
     strip.text.x = element_blank()
@@ -236,7 +237,7 @@ g5 <- pc_df %>%
   scale_color_manual(values = c("#258039", "#F5BE41", "#31A9B8", "red")) +
   scale_shape_manual(values = c(15, 17, 19, 3)) +
[…]

### 2017-03-28 analysis/uhoh_plot_predictive_abilities.R

@@ -86,15 +86,15 @@ compute_r_and_cv <- function(x) {
 }
 
 ext_pred_order <- c(
-  "P_Imputed", "G_Imputed", "PG_Imputed", "PT_Imputed", "GT_Imputed",
+  "P_Full", "G_Full", "PG_Full", "PT_Full", "GT_Full",
   "P_Core", "G_Core", "T_Core"
 )
 
 plot_data <- pred_df %>% 
   split(.$Group) %>% 
-  map_at("Full", .f = ~filter(., Geno %in% imputed_genotypes)) %>% 
+  #map_at("Full", .f = ~filter(., Geno %in% imputed_genotypes)) %>% 
   map(compute_r_and_cv) %>% 
-  set_names(c("Core", "Imputed")) %>% 
+  set_names(c("Core", "Full")) %>% 
   bind_rows(.id = "Group") %>% 
   unite(col = Exact_Predictor, Predictor, Group, sep = "_", remove = FALSE) %>% 
   mutate(
@@ -120,7 +120,7 @@ g1 <- plot_data %>%
   geom_errorbar(limits, position = dodge, width = 0.25) +
   facet_grid(. ~ Group) +
   scale_fill_manual(values = mycols) +
-  theme_base()
+  theme_pander(base_size = 10)
   
 ggsave(plot = g1, 
        filename = "./paper/tables_figures/hybrid_predictive_ability.pdf",

### 2017-03-29 paper/manuscript.tex
Expand discussion, add results on manual core sampling

@@
 \usepackage{multicol} % list corresponding authors in two separate columns
+\usepackage{hyperref} % web links
 
@@
 phenotyping, which can be reduced when covering all genotypes with a predictor 
-while limiting phenotyping to a few genotypes with high precision.
+while limiting phenotyping to a few genotypes with high precision
+\cite{Kadam2016}.
 The performance of all other genotypes is then forecast using model parameters
 obtained from thorough (repeated) $k$-fold cross-validation routines applied to 
-the subset of individuals with full information.
+the subset of individuals with full information \cite{Technow2014}.
 In animal breeding, genomic selection has increased the annual selection gain 
@@
 
-% Search for better predictors
-Compared to pedigree information, which represent the expected relationship
-among individuals, genomic information capture Mendelian sampling and thereby 
-provide an improved proxy of the realized relationship among individuals.
-Nevertheless, genomic data do not exhaustively capture physiological epistasis 
-\cite{Jiang2015,Guo2016,Vazquez2016}, describing the interactions 
-within and between different biological strata \cite{Sackton2016}.
-Such interactions were found to be pervasive throughout the genomes of yeast 
-\cite{Brem2005} and humans \cite{Brown2014} and have motivated studies on the 
-utility of downstream 'omics' predictors for integrating such interactions.
-So far, encouraging results have been found for humans \cite{Vazquez2016}, maize 
-inbred lines \cite{Guo2016} and maize hybrids \cite{Westhues2017}.
-
 % Incomplete predictors.
@@
 genotypes while other predictors are incomplete in that they cover only a 
-subset of all genotypes.
+subset of all genotypes \cite{Fragomeni2015}.
 Usually, the complete predictor is pedigree information while genomic data are
@@
 In the early years of genomic prediction, a two-step procedure was the most
-intuitive choice for establishing a new selection scheme.
-The first step constists of conventionally estimating breeding values based on
+intuitive choice for establishing a new selection scheme \cite{VanRaden2009}.
+The first step consists of conventionally estimating breeding values based on
 pedigree relationships and the phenotypes of relatives.
@@
 % -   Mention scarcity of data on single step prediction in plant populations.
-Single step prediction has already become a routine procedure in animal
-breeding but, to the best of our knowledge, has only been considered in plants
-in a study on wheat lines \cite{Ashraf2016}. 
+In animal breeding, single step prediction has already become a routine 
+procedure \cite{Legarra2014} but, to the best of our knowledge, has only been 
+considered in plants in a study on wheat lines \cite{Ashraf2016}. 
 Our objectives were threefold: (1) Examine whether the single step prediction
@@
 Gene expression data for seedlings seven days after sowing were obtained for a 
-subset of 60 parental Dent and 43 parental Flint lines through two-color
-hybridizations using a custom 2K microarray (GPL22267) \cite{Westhues2017}.
+subset of 60 parental Dent and 43 parental Flint lines, that are parents of 685
+hybrid progeny, through two-color hybridizations using a custom 2K microarray 
+(GPL22267) \cite{Westhues2017}.
 After applying established normalization procedures 
@@
 genotypes that has only data on one out of two predictors, we generated core
-samples using the R-package \textit{corehunter} \cite{Thachuk2009}.
-These core samples were based on the 149 maize inbred lines from the diversity
+samples based on two different approaches.
+All core sets were assembled from the 149 maize inbred lines from the diversity
 panel which had gene expression as well as genomic information.
+We defined core samples as a subset of genotypes that is covered by two
+predictors whereas all other genotypes (\textit{i.e.}, the complement) are only
+covered by a single predictor.
+The first application involved assembling core sets of varying sizes by 
+maximizing the average genetic distance among core set members using the 
+Modified Rogers (MR) distance as the criterion \cite{Thachuk2009}.
 The size of the core set was varied in increments of ten percentage points and 
 ranged from 10\% to 90\% of all 149 inbred lines.
-For assembling the various core sets, we first computed the Modifed Rogers (MR)
-distance between each pair of genotypes and then selected a predefined subset
-of genotypes that display maximum average distance between each other.
+For the second application, we first ran an admixture analysis in combination
+with a PCA and kept $K=3$ as the number of putative ancestral population, which 
+seemed to describe the population structure best.
+Genotypes were then assigned to one of the three ancestral populations based on
+whether any of the three relationship coefficients (A1-A3) was equal to or
+greater than 0.5.
+For example, a genotype with an admixture coefficient for A2 $\geq 0.5$ was
+assigned to the second ancestral population.
+Thereby, we obtained $n_{A1} = 17$, $n_{A2} = 109$, $n_{A3} = 19$ and four 
+genotypes that could not be unambiguously assigned to any putative ancestral 
+population based on this criterion.
+To ensure commensurability of the three subpopulations within the set of 149
+genotypes, the subset of 109 genotypes (\textit{i.e.}, members of A2) was
+reduced to $n_{A2} = 19$ genotypes, again using the method of maximizing the
+average genetic distance among selected genotypes described above.
 
@@
 
-\subsection{Universal breeding value model}
-The universal model for breeding values was:
+\subsection{Breeding value model}
+The general model for breeding values was:
 
@@
 corresponding design matrix associating the random genotype effects
-(\textit{i.e.}, breeding values) of the lines in the $l$-th inbred line group 
-and the $o$-th predictor type with $\mathbf{y}$.
-Note that, in the case of hybrid data, $\mathbf{Z}_{lo}$ associates the random
+$\mathbf{g}_{lo}$ (\textit{i.e.}, breeding values) of the lines in the $l$-th 
+inbred line group and the $o$-th predictor type with $\mathbf{y}$.
+Note that, in the case of hybrid data, $\mathbf{g}_{lo}$ pertains to the
 general combining ability effects in the parental inbred lines (either Dent or
-Flint, as expressed by $l$) with the vector of observed hybrid performance
-$\mathbf{y}$.
+Flint, as expressed by $l$) which are associated with the vector of
+observed hybrid performance $\mathbf{y}$ through $\mathbf{Z}_{lo}$.
 The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
@@
 for the residual error.
-All prediction models were implemented using the R-package \textit{BGLR}
-\cite{Perez2014}.
+All prediction models were implemented using the R-packages \textit{BGLR}
+\cite{Perez2014} and \textit{sspredr}
+\url{https://github.com/mwesthues/sspredr}.
 
@@
 \begin{equation} \label{eq:mrnaebv}
-	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
+	\mathbf{\hat{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
 \end{equation}
@@
 \begin{equation} \label{eq:single-step-model}
-\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}^{Complete} \boldsymbol{\epsilon}^{Complete} + \mathbf{e},
+\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U} \boldsymbol{\epsilon} + \mathbf{e},
 \end{equation}
@@
  \end{bmatrix},
- \mathbf{U}^{Complete} = 
+ \mathbf{U} = 
 \begin{bmatrix}
@@
 \begin{bmatrix}
-  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete}  \\
+  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}  \\
   \mathbf{W}_2\boldsymbol{\alpha} \\
@@
 \begin{bmatrix}
-  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}^{Complete} \\
+  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon} \\
   \mathbf{W}_2\boldsymbol{\alpha} \\
@@
 
-The breeding values from that model are:
+Our final model looks like this:
+
+\begin{align} \label{eq:final-model}
+\mathbf{y} &= \mathbf{Xb} +
+\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
+\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l} \boldsymbol{\epsilon}_{l} +
+\mathbf{e}.
+\end{align}
+
+Breeding values can be estimated as:
 
 \begin{equation} \label{eq:breeding-values}
-\tilde{\mathbf{g}} = 
+\hat{\mathbf{g}} = 
  \begin{bmatrix}
@@
  \end{bmatrix}
- \hat{\boldsymbol{\epsilon}}^{Complete} 
+ \hat{\boldsymbol{\epsilon}} 
 \end{equation}
-
-Our final model looks like this:
-
-\begin{align} \label{eq:final-model}
-\mathbf{y} &= \mathbf{Xb} +
-\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
-\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l}^{Complete} \boldsymbol{\epsilon}_{l}^{Complete} +
-\mathbf{e}.
-\end{align}
 
@@
 \begin{equation} \label{eq:predicted-performance}
-\hat{\mathbf{y}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
+\hat{\mathbf{y}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\hat{\mathbf{g}}_{l} 
 \end{equation}
@@
 
-where $n$ denotes the number of genotypes (*i.e.*, inbred lines or hybrids).
+where $n$ denotes the number of genotypes (\textit{i.e.}, inbred lines or 
+hybrids).
 
@@
 predictive abilities than the use of G for four out of six traits
-(Table \ref{table:InbredResults}).
+(Fig. \ref{fig:InbredResults}A).
 In terms of predictive ability, the absolute advantage of T over G was as large 
-as 0.04 as observed for the traits '100grainweight', 'Eardiameter' and 
-'Silkingtime'.
-Except for 'Kernelwidth', the predictive ability for predictions based on
-G and the full set of inbred lines ($n = 211$) was always higher than for the 
-reduced set of inbred lines ($n = 149$).
+as 0.04 as observed for the traits DS, ED, GW and KW.
+Except for KW, the predictive ability for predictions based on G were always
+higher for the full set of inbred lines ($n = 211$) than for the reduced set of 
+inbred lines ($n = 149$).
 The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
@@
 superior over G for the reduced set of 149 lines.
-In the case of '100grainweight' and 'Eardiameter', the use of GT yielded a 
-slight improvement in predictive ability over G.
+In the case of GW and ED, the use of GT yielded a slight improvement in predictive 
+ability over G.
 Where T was a worse predictor for the reduced set of 149 lines than G,
@@
   \caption{
-  Predictive abilities and corresponding coefficients of variation for 
-  different core sets from the tropical/subtropical material of the maize 
-  diversity panel.
+  Predictive abilities and average coefficient of variation for
+  tropical/subtropical inbred lines from the maize diversity panel and six 
+  agronomic traits.
   As predictors, genomic (G) data, transcriptomic (T) data and their 
   combination (GT) were used.
-  The decimal after the predictor name indicates the size of the genetic 
-  material that was covered by inbred lines in the core set.
-  Influence of core set composition on predictive abilities.
-  Predictive abilities are based on single step predition with genomic data as the 
-  complete predictor and transcriptomic data as the incomplete predictor.
-  Results are shown for six agronomic traits and nine core sets of the maize 
-  diversity panel.
+  (\textbf{A}) The 'Core' panel includes a subset of 149 lines covered by
+  genomic and transcriptomic information whereas the 'Full' panel comprises all
+  211 genotypes. In the 'Full' set of genotypes, transcriptomic records were
+  imputed for 62 lines.
+  (\textbf{B}) Influence of structural missingness in the coverage of genotypes 
+  with transcriptomic information on single step-based predictive abilities 
+  using a 'Core' set of 149 inbred lines.
+  Transcriptomic records were deleted for 17 to 19 genotypes that were
+  primarily associated with a particular subgroup in the material, as defined
+  by an admixture analysis and a PCA (Fig. \ref{fig:PopStructure}).
+  (\textbf{C}) Influence of core set sampling on predictive abilities using
+  nine different sub set sizes where 'Core Fraction' indicates the fraction
+  of 149 inbred lines that is covered by both, genomic and transcriptomic
+  data.
   } 
@@
 data, as the second best predictor for this trait 
-(Table \ref{fig:HybridResults}).
-In the core set, P had the highest predictive ability for starch ('STA').
-For the other four traits, G had the best predictive abilities.
+(Fig. \ref{fig:HybridResults}).
+In the core set, P had the highest predictive ability for starch ('STA') and 
+dry matter content ('DMC').
+For the other three traits, inside the core set G was the superior predictor.
 In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
@@
 
-% latex table generated in R 3.3.3 by xtable 1.8-2 package
-% Fri Mar 24 12:24:30 2017
 % Script: ./analysis/uhoh_plot_predictive_abilities.R
@@
   \caption{
-    Predictive abilities and corresponding coefficients of variation for the 
-    set of maize hybrids.
+    Predictive abilities and average coefficients of variation for the set of 
+    maize hybrids and six agronomic traits.
     As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
     combinations thereof we used.
+    The 'Full' set of genotypes comprises 1,521 hybrid progeny based on 245
+    parental inbred lines, whereas the 'Core' set comprises 685 hybrid progeny
+    based on 103 parental inbred lines.
   } 
@@
 The nine core sets differ in that each covers a different number of genotypes,
-raning from only 10\% to to 90\% of all genotypes covered by the incomplete
+ranging from only 10\% to to 90\% of all genotypes covered by the incomplete
 predictor (Fig. \ref{fig:CoreSetPCA}).
@@
 traits (Fig. \ref{fig:InbredResults}C).
-For the traits 'cobweight' and 'Plantheight', which were better predicted via G
-than via T in the reduced set of lines, predictive abilities were higher when
-only a small number of genotypes was included in the core set
+For the traits CW and PH, which were better predicted via G than via T in the 
+reduced set of lines, predictive abilities were higher when only a small number 
+of genotypes was included in the core set.
+The latter corresponds to a situation where a large number of genotypes is
+being imputed and the majority of the information is derived from genomic
+information. 
 For most other traits, an initial increase in predictive abilities when moving
@@
 0.6/0.7 and a subsequent hike up to a core set size of 0.9.
+A principal component analysis depicts the coverage of the genetic space by
+genotypes (Fig. \ref{fig:PopStructure}).
+With a core set size of 50-60\% of the 149 genotypes, the congruence between
+core set and complement was high whereas genotypes in the south-west direction
+of the PCA plot were overrepresented compared to genotypes in the east 
+direction at a core set size of 0.7.
 
@@
 from the central cluster of genotypes belonging largely to the A2 population
-(Fig. \ref{fig:PopStructure}A, Table \ref{fig:InbredTable}).
+(Fig. \ref{fig:PopStructure}A, Fig. \ref{fig:InbredResults}B).
 
@@
 \section*{Discussion}
-
-\Genetics2level{Advantages of marker effect models}
-% Describe the many hacks that have been developed to tackle problems with BLUP
-% models.
-
-$\dots$
+\Genetics2level{Addition of new predictors}
+Pedigree records are the most ubiquitous predictor in animal and plant breeding
+programs, because they are easy and inexpensive to collect.
+Compared to pedigree information, which represent the expected relationship
+among individuals, genomic information capture Mendelian sampling and thereby 
+provide an improved proxy of the realized relationship among individuals.
+Nevertheless, genomic data do not exhaustively capture physiological epistasis 
+\cite{Jiang2015,Guo2016,Vazquez2016}, describing the interactions 
+within and between different biological strata \cite{Sackton2016}.
+Such interactions were found to be pervasive throughout the genomes of yeast 
+\cite{Brem2005} and humans \cite{Brown2014} and have motivated studies on the 
+utility of downstream 'omics' predictors for integrating such interactions.
+Recently, encouraging results have been found for humans \cite{Vazquez2016}, 
+maize inbred lines \cite{Guo2016} and maize hybrids \cite{Westhues2017}.
+While single step prediction has become an established method when complete
+pedigree but only incomplete genomic records are available, its application has
+not yet been considered for the imputation of a complete predictor based on 
+incomplete downstream 'omics' predictors such as transcriptomic, proteomic or 
+metabolomic data.
+Here, we considered single step approaches for a maize inbred line diversity
+panel as well as for a collection of maize hybrids using pedigree, genomic and
+transcriptomic data as predictors.
 
@@
 we employed a leave-one-out cross-validation scheme (LOOCV).
+When using a single step approach based on complete genomic and incomplete
+transcriptomic information, gains in predictive ability for the inbred lines
+were at best just slightly higher than when using complete genomic information
+alone.
+One of the reasons might be that the subset of genotypes that was covered by
+transcriptomic data was not a good representation of the full genetic space of
+all available genotypes.
+A second possibility is that the number of additional phenotypes in the full
+set compared to the core set ($\Delta(n) = 62$) was not large enough to provide
+considerably more useful information.
+The latter was the case for \citeNP{Ashraf2016}, who imputed genotypes for about 
+10,000 wheat lines using pedigree information and observed greater accuracy of
+the single step method over genomic BLUP for all four evaluated traits.
 
@@
 % 2014, GSE).
-
-\begin{itemize}
-  \item Generally, it is important that, on the same set of genotypes, the 
-    incomplete predictor yields larger predictive abilities than the complete 
-    predictor.
-  \item In hybrid breeding, the probability of successfully selecting the best
+For the hybrid data, the predictive abilities of genomic information
+for the whole set of available parent lines rose considerably for several
+agronomic traits.
+Hence, the inclusion of trancsriptomic data could not further improve upon
+this result.
+Predictive abilities obtained when using pedigree data could be 
+improved considerably when, depending on the agronomic trait, combined with 
+either transcriptomic or genomic information.
+Evaluations of core sets for the parental inbred lines of the maize hybrids
+used throughout this study were outside the scope of this paper, but
+results on the inbred lines have shown that homogeneous coverage of the
+genetic space by the incomplete predictor can considerably leverage the
+predictive ability in single step prediction compared to using only the
+complete predictor.
+In hybrid breeding, the probability of successfully selecting the best
 observed genotypes based on the best predicted candidates is a strongly 
@@
 single step might justify additional investments in RNA-seq at least of a
-    subste of genotypes that represents the genetic space of the breeding 
+subset of genotypes that represents the genetic space of the breeding 
 material well.
-  \item For the hybrid data, the predictive abilities of genomic information
-    for the whole set of available parent lines rose considerably for several
-    agronomic traits.
-    Hence, the inclusion of trancsriptomic data could not further improve upon
-    this result.
-\end{itemize}
 
@@
 
-\section*{Conclusions}
+When using core sets of varying sizes, which were established with the
+constraint of maximizing the average genetic distance among their members, we
+noticed strong fluctuations in predictive abilities when moving from small core
+sets to large core sets.
+For traits that were better predicted by transcriptomic than by genomic
+information in the core set, predictive abilities typically peaked around
+core set size of 0.5/0.6, then dropped sharply at 0.7 and started to increase
+again with a size of 0.9.
+A principal component analysis showed that, at a core set size of 0.7, 
+genotypes assigned to the A3 ancestral population (Fig. \ref{fig:PopStructure})
+were largely elements of the core set whereas genotypes with a large share at 
+the A1 ancestral population were predominantly excluded from the core set.
+At a core set size of 90\% of the original size, the majority of the A1
+genotypes were finally covered by the core set as well.
+This result suggests a negative bias in the imputation of genotypes from the A1
+population and could explain the drop in predictive ability at a core set size
+of 0.7, which is in agreement with the reduced predictive abilities observed
+when stripping A1 genotypes from their transcriptomic information (Fig.
+\ref{fig:InbredResults}B).
+
+\Genetics2level{Advantages of marker effect models}
+% Describe the many hacks that have been developed to tackle problems with BLUP
+% models.
 
 $\dots$
+
+\section*{Conclusions}
+We have successfully applied single step prediction to a hybrid plant data set 
+while imputing with a quantitative downstream 'omics' predictor.
+By declaring different subsets of individuals as covered by one or two 
+predictors, respectively, we could show the importance of homogeneous coverage
+of the genetic space by genotypes covered by both predictors on predictive
+ability.
+Extensions to more than two predictors were outside the scope of this study
+but, with mounting interest in systems genetics, should be considered in the
+future.
 

### 2017-03-29 analysis/uhoh_pca.R

@@ -6,7 +6,7 @@
 
 # Data and packages -------------------------------------------------------
 if (!require("pacman")) install.packages("pacman")
-pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "devtools")
+pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "devtools", "ggthemes")
 devtools::install_github("mwesthues/sspredr")
 pacman::p_load_gh("mwesthues/sspredr")
 
@@ -77,29 +77,34 @@ pca_df <- pca_lst %>%
   map(., ~rownames_to_column(., var = "G")) %>%
   map(., ~gather(., key = PC, value = Score, -G)) %>%
   set_names(names(quality_snps)) %>%
-  bind_rows(.id = "Group") %>%
+  bind_rows(.id = "Heterotic_Group") %>%
   as_data_frame() %>%
   mutate(PC = str_replace_all(PC, pattern = "V", replacement = "PC")) %>%
   filter(PC %in% paste0("PC", seq_len(2))) %>%
   mutate(
-    Reduced = if_else(
+    Group = if_else(
       G %in% predictor_coverage$mrna,
-      true = TRUE,
-      false = FALSE
+      true = "Core",
+      false = "Full"
     )
-  )
+  ) %>% 
+  rename(`Heterotic Group` = Heterotic_Group)
 
 
 # PCA plot
 g1 <- pca_df %>%
   spread(key = PC, value = Score) %>%
-  ggplot(aes(x = PC1, y = PC2, color = Reduced, shape = Reduced)) +
+  ggplot(aes(x = PC1, y = PC2, color = Group, shape = Group)) +
   geom_point() +
   guides(color = guide_legend(override.aes = list(size = 3))) +
   scale_color_tableau() +
-  facet_grid(. ~ Group) +
-  theme_bw(base_size = 10) +
-  theme(legend.position = "top")
+  facet_grid(. ~ `Heterotic Group`) +
+  theme_pander(base_size = 10) +
+  theme(
+    legend.position = "top",
+    panel.border = element_blank(),
+    axis.line = element_line()
+    )
 ggsave(plot = g1, 
        filename = "./paper/tables_figures/uhoh_dent_flint_pca.pdf",
        height = 4, 

### 2017-03-30 paper/manuscript.tex
Discuss SS-BLUP with SSBR, Add trait acronym table

@@
 The performance of all other genotypes is then forecast using model parameters
-obtained from thorough (repeated) $k$-fold cross-validation routines applied to 
-the subset of individuals with full information \cite{Technow2014}.
+obtained from thorough resampling routines applied to the subset of individuals 
+with full information \cite{Hastie2009}.
 In animal breeding, genomic selection has increased the annual selection gain 
@@
 $\mathbf{G}$ \cite{Christensen2012} and their weighting, which is not
-trait-independent \cite{Ashraf2016}.
+trait-independent \cite{Vitezica2011,Ashraf2016}.
 \citeNP{Fernando2014} derived the equivalent single step marker-effect-model to 
@@
 three or more agro-ecologically diverse locations across Germany for six 
-agronomically important traits.
+agronomically important traits (Table \ref{table:TraitKey}).
 All 245 parent lines have pedigree records back to the generation of their 
@@
 Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
-six were used for this study.
+six were used for this study (Table \ref{table:TraitKey}).
 MaizeSNP50 BeadChip-data were available for the entire set of lines from the
@@
 genotypes and was kept for further analyses \cite{Li2013}.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\centering
+\caption{Acronyms for the trait names}
+\label{table:TraitKey}
+\begin{tabular}{@{}lll@{}}
+\toprule
+Trait              & Acronym & Data Set  \\ \midrule
+Cob weight         & CW      & Diversity \\
+Days to silking    & DS      & Diversity \\
+Ear diameter       & ED      & Diversity \\
+100-grain weight   & GW      & Diversity \\
+Kernel width       & KW      & Diversity \\
+Plant height       & PH      & Diversity \\
+Dry matter yield   & DMY     & Hybrid    \\
+Dry matter content & DMC     & Hybrid    \\
+Fat                & FAT     & Hybrid    \\
+Protein            & PRO     & Hybrid    \\
+Starch             & STA     & Hybrid    \\
+Sugar              & SUG     & Hybrid    \\ \bottomrule
+\end{tabular}
+\end{table}
 
@@
 The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
-prediction of all 211 lines performed at least as well as G alone when T was 
+prediction of all 211 lines, performed at least as well as G alone when T was 
 superior over G for the reduced set of 149 lines.
@@
 we denote as 'Core'.
-For dry matter yield (DMY), T was the best predictor in the core set of 685 
-hybrids with a predictive ability that was 0.08 points higher than for genomic 
-data, as the second best predictor for this trait 
-(Fig. \ref{fig:HybridResults}).
-In the core set, P had the highest predictive ability for starch ('STA') and 
-dry matter content ('DMC').
-For the other three traits, inside the core set G was the superior predictor.
+For DMY, T was the best predictor in the core set of 685 hybrids with a 
+predictive ability that was 0.08 points higher than for genomic data, as the 
+second best predictor for this trait (Fig. \ref{fig:HybridResults}).
+In the core set, P had the highest predictive ability for STA and DMC.
+For the other three traits, G was the superior predictor inside the core set.
 In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
-markedly for all traits with the exception of 'FAT'.
+markedly for all traits with the exception of FAT.
 The combination of genomic and transriptomic information (GT) in a single-step
-model yielded a small improvement of G alone for the trait 'DMY'.
-All traits, where T was not superior to G in the core set, did not benefit from 
-the single-step approach when compared to genomic information alone.
+model yielded a small improvement over G alone for the traits DMY and PRO.
+Except for PRO, compared to using only G, trait prediction did not benefit from 
+the single step approach when T was not superior to G in the core set.
 Predictive abilities resulting from combinations of pedigree with genomic (PG)
@@
 that are covered by both, genomic and transcriptomic information.
-First, we assmbled nine different core sets, which we defined as subsets
+First, we assembled nine different core sets, which we defined as subsets
 of genotypes that maximize the average genetic distance between genotypes of
@@
 The nine core sets differ in that each covers a different number of genotypes,
-ranging from only 10\% to to 90\% of all genotypes covered by the incomplete
+ranging from only 10\% to 90\% of all genotypes covered by the incomplete
 predictor (Fig. \ref{fig:CoreSetPCA}).
@@
 Predictive abilities obtained with the different core sets varied widely within
-traits (Fig. \ref{fig:InbredResults}C).
+traits (Fig. \ref{fig:InbredResults}B).
 For the traits CW and PH, which were better predicted via G than via T in the 
@@
 of genotypes was included in the core set.
-The latter corresponds to a situation where a large number of genotypes is
-being imputed and the majority of the information is derived from genomic
-information. 
+This corresponds to a situation where a large number of genotypes is being 
+imputed and the majority of the information is derived from genomic information. 
 For most other traits, an initial increase in predictive abilities when moving
@@
 more extreme scenario where we select core sets based on population structure.
-Hereto, we ran an admixture analysis on the same set of 149 inbred lines and
-display the estimated ancestry coefficients for the genotypes assuming $K=3$
-ancestral populations (Fig. \ref{fig:PopStructure}).
+Hereto, we ran an admixture analysis on the same set of 149 inbred lines 
+assuming $K=3$ ancestral populations (Fig. \ref{fig:PopStructure}).
 
@@
 ancestral population (mixed).
-To adjust the size of the set of genotypes largely belonging to A2 to the other
-two sets, we selected 19 genotypes from A2 using previously described core
-sampling.
+To reconcile the large size of the A2 set compared to A1 and A3, we selected 19 
+genotypes from A2 using previously described core sampling.
 For single-step prediction, each of the three genotype sets were then declared
@@
 from the central cluster of genotypes belonging largely to the A2 population
-(Fig. \ref{fig:PopStructure}A, Fig. \ref{fig:InbredResults}B).
+(Fig. \ref{fig:PopStructure}A, Fig. \ref{fig:InbredResults}C).
 
-A principal component analysis, carried out separately on the two parental
-heterotic groups of the maize hybrids, showed that the entire genetic space in
-each group was well covered by genotypes that had both, transcriptomic and
-genomic data (Fig. \ref{fig:UHOH-PCA}).
+Regarding the hybrid material, a PCA carried out separately for the two groups
+of parental inbred lines showed that the entire genetic space in each group 
+was covered well by genotypes that had both, transcriptomic and genomic data 
+(Fig. \ref{fig:UHOH-PCA}).
 
@@
 \section*{Discussion}
+\Genetics2level{Advantages of marker effect models}
+Single step approaches in hybrid populations have so far been limited to pig
+breeding \cite{Xiang2015,Xiang2016,Tusell2016}.
+All of these studies used breeding value-based single step BLUP methods
+(SS-BLUP) conceived by \citeNP{Legarra2009} and \citeNP{Christensen2010} with
+extensions to hybrid populations \cite{Christensen2014,Christensen2015}.
+An attactive property of SS-BLUP models is their computational advantage over 
+marker-effect single step Bayesian regression models (SSBR) when the number of 
+genotypes is smaller than the number of features of the incomplete predictor.
+Nevertheless, SS-BLUP suffers from a variety of issues, which we will present
+briefly:
+First, the combination of the numerator with the genomic relationship matrix
+requires commensurability of the two matrices, which necessitates some form of
+scaling and weighting of $\mathbf{A}$ and $\mathbf{G}$
+\cite{Christensen2012,Christensen2012a}.
+Second, the fact that mostly recent individuals have genotypes whereas the 
+majority of old individuals has only pedigree records introduces a bias that 
+has to be accounted for \cite{Vitezica2011,Legarra2015,Garcia-Baccino2017}.
+Third, the addition of new genotypes requires updates of $\mathbf{G}$ that
+demand additional measures such as computing the inverse of $\mathbf{G}$ via
+recursion
+\cite{Misztal2014,Misztal2016,Misztal2016a,Fragomeni2015,Masuda2016,Pocrnic2016}.
+Advantages of the SSBR algorithm, conceived by \citeNP{Fernando2014}, include
+that i) it does not require commensurability of $\mathbf{A}$ and $\mathbf{G}$, 
+ii) it models the bias incurred by the incomplete predictor explicitly as 
+$\mathbf{\epsilon}$ and iii) its speed depends largely on the number of 
+features.
+To our best knowledge, this study presents the first implementation of the SSBR 
+algorithm in a hybrid data set.
+
 \Genetics2level{Addition of new predictors}
@@
 progeny.
-So far, its applications in hybrid populations are limited to pig breeding
-\cite{Xiang2015,Xiang2016,Tusell2016}.
 % Are there applications in chicken breeding? (Hints in Christensen et al.
@@
 Animal populations are so large that it does not truly matter, which animals are 
-being genotyped, but in plant breeding the situation is different and we would 
-be interested in determining which individuals should be complemented with 
-information on another predictor.
-If one can cover the genetic space sufficently with the incomplete predictor,
-
+being genotyped once more than 10,000 animals have pedigree records 
+\cite{Fragomeni2015,Lourenco2015,Masuda2016}, but in plant breeding programs,
+pedigrees are typically considerably smaller.
+Single step prediction offers the promise of leveraging the predictive ability
+for all genotypes by borrowing the superior performance of an auxiliary
+predictor on only a subset of genotypes, thereby reducing predictor costs.
+Hence, we were interested in determining which individuals should be 
+complemented with information on another predictor.
 When using core sets of varying sizes, which were established with the
@@
 \ref{fig:InbredResults}B).
-
-\Genetics2level{Advantages of marker effect models}
-% Describe the many hacks that have been developed to tackle problems with BLUP
-% models.
-
-$\dots$
 

### 2017-03-30 paper/manuscript.tex
Rearrange inbred combi plot and PCA plot

@@
 was covered well by genotypes that had both, transcriptomic and genomic data 
-(Fig. \ref{fig:UHOH-PCA}).
+(Fig. \ref{fig:PCA}).
 
@@
 \centering
-\includegraphics{./tables_figures/uhoh_dent_flint_pca.pdf}
+\includegraphics{./tables_figures/diversity_dent_flint_pca.pdf}
 \caption{
-  Principal component (PC) analysis of the two heterotic groups of the maize 
-  hybrids.
+  Principal component (PC) analysis i) of the 211 tropical/subtropical lines 
+  from the maize diversity panel, ii) of the 142 Dent parent lines of the 1,521
+  maize hybrids and iii) of the 103 Flint parent lines of the maize hybrids.
   All inbred lines, which are covered by genomic and transcriptomic features
@@
 }
-\label{fig:UHOH-PCA}
+\label{fig:PCA}
 \end{figure}

### 2017-03-30 analysis/maizego_core_fraction_plot_predictive_ability.R

@@ -106,26 +106,6 @@ plot_a <- plot_a_data %>%
   
 
 plot_b_data <- pred_df %>% 
-  filter(Core_Fraction %in% paste0("A", seq_len(3))) %>% 
-  group_by(Trait, Core_Fraction, Predictor) %>% 
-  summarize(
-    r = cor(y, yhat),
-    CV = coefficient_of_variation(var_yhat, yhat)
-  ) %>%
-  ungroup() %>% 
-  rename(`Imputed Ancestral Population` = Core_Fraction)
-
-plot_b <- plot_b_data %>% 
-  ggplot(aes(x = Trait, y = r, fill = `Imputed Ancestral Population`)) +
-  geom_bar(stat = "identity", position = dodge) +
-  geom_errorbar(limits, position = dodge, width = 0.25) +
-  scale_fill_manual(values = c("#258039", "#F5BE41", "#31A9B8")) +
-  theme_pander(base_size = 10) +
-  theme(legend.position = "top")
-plot_b
-
-
-plot_c_data <- pred_df %>% 
   filter(Core_Fraction %in% seq(from = 0.1, to = 0.9, by = 0.1)) %>% 
   group_by(Trait, Core_Fraction, Predictor) %>% 
   summarize(
@@ -135,18 +115,35 @@ plot_c_data <- pred_df %>%
   ungroup() %>% 
   rename(`Core Fraction` = Core_Fraction)
   
-
 # Because the bars and errorbars have different widths
 # we need to specify how wide the objects we are dodging are
-dodge_c <- position_dodge(width = 0.3)
-plot_c <- plot_c_data %>% 
+dodge_b <- position_dodge(width = 0.3)
+plot_b <- plot_b_data %>% 
   ggplot(aes(x = `Core Fraction`, y = r, color = Trait, group = Trait)) +
-  geom_line(stat = "identity", position = dodge_c) +
-  geom_errorbar(limits, position = dodge_c, width = 0.25) +
+  geom_line(stat = "identity", position = dodge_b) +
+  geom_errorbar(limits, position = dodge_b, width = 0.25) +
   scale_color_tableau() +
   theme_pander(base_size = 10) +
   theme(legend.position = "top") +
   guides(color = guide_legend(override.aes = list(size = 3)))
+
+plot_c_data <- pred_df %>% 
+  filter(Core_Fraction %in% paste0("A", seq_len(3))) %>% 
[…]

### 2017-03-30 analysis/maizego_snp_analyses.R

@@ -76,29 +76,8 @@ all_geno_pc <- LEA::pca("./data/derived/maizego/all_tst_snp.lfmm", scale = TRUE)
 all_geno_pc <- all_geno_pc$projections
 rownames(all_geno_pc) <- rownames(all_geno_snp)
 colnames(all_geno_pc) <- paste0("PC_", seq_len(ncol(all_geno_pc)))
-
-g1 <- all_geno_pc %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "G") %>%
-  gather(key = PC, value = Score, -G) %>%
-  as_data_frame() %>%
-  filter(PC %in% paste0("PC_", seq_len(3))) %>%
-  mutate(Group = if_else(
-    G %in% common_genotypes, true = "mRNA", false = "All"),
-    PC = gsub("PC_", replacement = "PC", x = PC)
-  ) %>%
-  spread(key = PC, value = Score) %>%
-  ggplot(aes(x = PC1, y = PC2, color = Group, shape = Group)) +
-  geom_point() +
-  scale_color_tableau() +
-  theme_pander(base_size = 10) +
-  theme(legend.position = "top")
-ggsave(plot = g1,
-       filename = "./paper/tables_figures/maizego_pca.pdf",
-       width = 6,
-       height = 4, 
-       units = "in")
-
+all_geno_pc %>% 
+  saveRDS(., file = "./data/derived/maizego/maizego_pca.RDS")
 
 
 
@@ -137,7 +116,7 @@ structure_df <- K_lst %>%
 
 # For each genotype and for each run with K ancestral populations, plot the 
 # admixture coefficients by genotype.
-g2 <- structure_df %>%
+g1 <- structure_df %>%
   ggplot(aes(x = G, y = Value, fill = Component)) +
   geom_bar(stat = "identity", width = 1) +
   facet_wrap(~K, ncol = 1, strip.position = "right") +
@@ -153,7 +132,7 @@ g2 <- structure_df %>%
 pc_df <- "./data/derived/maizego/core_sampling_pca.RDS" %>%
   readRDS()
 
-g3 <- pc_df %>%
+g2 <- pc_df %>%
   filter(Fraction != "1.0") %>%
   rename(`Core Group` = Core_Group) %>%
   ggplot(aes(x = PC1, y = PC2, color = `Core Group`, shape = `Core Group`)) +
[…]

### 2017-03-30 analysis/pca.R

@@ -72,41 +72,77 @@ predictor_coverage <- common_genotypes %>%
 
 # Convert the PCA list for Dent and Flint material to a data frame and add 
 # information on predictor coverage of the genotypes.
-pca_df <- pca_lst %>%
+hybrids <- pca_lst %>%
   map(as.data.frame) %>%
   map(., ~rownames_to_column(., var = "G")) %>%
   map(., ~gather(., key = PC, value = Score, -G)) %>%
   set_names(names(quality_snps)) %>%
-  bind_rows(.id = "Heterotic_Group") %>%
+  bind_rows(.id = "Group") %>%
   as_data_frame() %>%
   mutate(PC = str_replace_all(PC, pattern = "V", replacement = "PC")) %>%
   filter(PC %in% paste0("PC", seq_len(2))) %>%
   mutate(
-    Group = if_else(
+    Cluster = if_else(
       G %in% predictor_coverage$mrna,
       true = "Core",
       false = "Full"
     )
-  ) %>% 
-  rename(`Heterotic Group` = Heterotic_Group)
+  )
 
 
 # PCA plot
 g1 <- pca_df %>%
   spread(key = PC, value = Score) %>%
-  ggplot(aes(x = PC1, y = PC2, color = Group, shape = Group)) +
+  ggplot(aes(x = PC1, y = PC2, color = Cluster, shape = Cluster)) +
   geom_point() +
   guides(color = guide_legend(override.aes = list(size = 3))) +
   scale_color_tableau() +
-  facet_grid(. ~ `Heterotic Group`) +
+  facet_grid(. ~ Group) +
   theme_pander(base_size = 10) +
   theme(
     legend.position = "top",
     panel.border = element_blank(),
     axis.line = element_line()
     )
+
+
+# DIVERSITY INBRED LINE CORE SET ------------------------------------------
+all_geno_pc <- "./data/derived/maizego/maizego_pca.RDS" %>% 
+  readRDS()
+common_inbreds <- "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS" %>% 
+  readRDS() %>% 
+  reduce(intersect)
+inbreds <- all_geno_pc %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "G") %>%
+  gather(key = PC, value = Score, -G) %>%
+  as_data_frame() %>%
+  filter(PC %in% paste0("PC_", seq_len(2))) %>%
+  mutate(Cluster = if_else(
[…]

### 2017-03-30 paper/manuscript.tex
Reorder paragraphs in discussion

@@
 To our best knowledge, this study presents the first implementation of the SSBR 
-algorithm in a hybrid data set.
+algorithm on a hybrid data set.
 
@@
 \subsection{Inbred lines}
-Predictive abilities for the maize diversity panel are not in complete
-agreement with those reported by \citeNP{Guo2016}.
-Reasons for this may be threefold: 
-1) In our study, we reduced the dataset to the 211 tropical/subtropical lines 
-and excluded all other genotypes whereas \citeNP{Guo2016} used genotypes from 
-four subpopulations and only a subset of 157 genotypes from the set of 
-tropical/subtropical lines while correcting for population structure by
-including a fixed effect in the model for predicting agronomic performance.
-2) We applied quality checks for the predictor data after generating the subset
-of 211 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
-genotypes.
-3) While \citeNP{Guo2016} used five-fold cross-validation with 500 repetitions,
-we employed a leave-one-out cross-validation scheme (LOOCV).
 When using a single step approach based on complete genomic and incomplete
@@
 all available genotypes.
-A second possibility is that the number of additional phenotypes in the full
+Another possibility is that the number of additional phenotypes in the full
 set compared to the core set ($\Delta(n) = 62$) was not large enough to provide
@@
 the single step method over genomic BLUP for all four evaluated traits.
+Predictive abilities reported for the maize diversity panel in this study were
+slightly different from those reported by \citeNP{Guo2016}, for three
+possible reasons:
+1) We reduced the dataset to the 211 tropical/subtropical lines and excluded 
+all other genotypes from our analyses whereas \citeNP{Guo2016} used genotypes 
+from four different subpopulations, modeled as a fixed effect, and used only a 
+subset of 157 genotypes from the set of tropical/subtropical lines.
+2) We applied quality checks for the predictor data after generating the subset
+of 211 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
+genotypes.
+3) While \citeNP{Guo2016} used five-fold cross-validation with 500 repetitions,
+we employed a leave-one-out cross-validation scheme (LOOCV).
+Notwithstanding, the relative differences between predictive abilities for G 
+and T, respectively, were similar in both studies.
 
 \subsection{Maize hybrids}
-Hybrid breeding --- with noteworthy commercial applications in pig, chicken and 
-maize breeding --- is a particularly challenging application field for
-prediction tools.
+Hybrid breeding --- with noteworthy commercial applications in pig
+\cite{Xiang2016,Tusell2016} and maize breeding \cite{TheRoyalSociety2009} --- 
+is a particularly challenging application field for prediction tools 
+\cite{Kadam2016}.
 Here, $2n$ parent individuals from two genetically distinct heterotic groups are
@@
 progeny.
-% Are there applications in chicken breeding? (Hints in Christensen et al.
-% 2014, GSE).
+\citeNP{Westhues2017} have shown that, in hybrid breeding, the probability 
+of successfully selecting the best observed genotypes based on the best 
+predicted candidates is a strongly convex function of the predictive ability.
+Thus, even the minor gains when using transcriptomic and genomic data in a
+single step might justify additional investments in RNA-seq at least of a
+subset of genotypes that represents the genetic space of the breeding 
+material well.
 For the hybrid data, the predictive abilities of genomic information
@@
 complete predictor.
-In hybrid breeding, the probability of successfully selecting the best
-observed genotypes based on the best predicted candidates is a strongly 
-convex function of the predictive ability.
-Hence, even the minor gains when using transcriptomic and genomic data in a
-single step might justify additional investments in RNA-seq at least of a
-subset of genotypes that represents the genetic space of the breeding 
-material well.
 
 \Genetics2level{Coverage of the genetic space}
-Animal populations are so large that it does not truly matter, which animals are 
-being genotyped once more than 10,000 animals have pedigree records 
+Animal populations are so large that it the assembly of a core set of
+individuals with records on multiple predictors can effectively be done at 
+random once more than 10,000 animals have pedigree records 
 \cite{Fragomeni2015,Lourenco2015,Masuda2016}, but in plant breeding programs,

### 2017-03-30 paper/manuscript.tex
Fix typos in manuscript

@@
 either transcriptomic or genomic information.
-Evaluations of core sets for the parental inbred lines of the maize hybrids
-used throughout this study were outside the scope of this paper, but
-results on the inbred lines have shown that homogeneous coverage of the
-genetic space by the incomplete predictor can considerably leverage the
-predictive ability in single step prediction compared to using only the
-complete predictor.
+Evaluations of core sets for the parental inbred lines of the maize hybrids were 
+outside the scope of this study, but results on the inbred lines have shown that 
+homogeneous coverage of the genetic space by the incomplete predictor can 
+considerably leverage the predictive ability in single step prediction compared 
+to using only the complete predictor.
 
 \Genetics2level{Coverage of the genetic space}
-Animal populations are so large that it the assembly of a core set of
-individuals with records on multiple predictors can effectively be done at 
-random once more than 10,000 animals have pedigree records 
-\cite{Fragomeni2015,Lourenco2015,Masuda2016}, but in plant breeding programs,
-pedigrees are typically considerably smaller.
+Whereas animal breeding populations are oftentimes so large that the assembly 
+of a core set of individuals with records on multiple predictors can effectively 
+be done at random once more than 10,000 animals have pedigree records 
+\cite{Fragomeni2015,Lourenco2015,Masuda2016}, pedigrees in plant breeding 
+programs are typically considerably smaller.
 Single step prediction offers the promise of leveraging the predictive ability

### 2017-05-10 reports/xgboost.Rmd
Check out XGBoosting

---
title: "Hybrid Prediction with Extreme Gradient Boosting"
author: Matthias Westhues
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

```{r setup}
if (!require("pacman")) install.packages("pacman")
# Load 'xgboost' before 'dplyr' because the former relies on 'plyr' functions,
# which cause conflicts with 'dplyr' functions when not not loaded first.
pacman::p_load("xgboost")
pacman::p_load_gh("tidyverse/dplyr")
pacman::p_load(
  "purrr", "readr", "stringr", "ggplot2", "tibble", "forcats", "tidyr", 
  "assertr", "knitr", "rprojroot", "caret", "DMwR", "propagate"
)
knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load_Data}
# Response variable.
pheno <- "./data/derived/maizego/tst_pheno_tibble.RDS" %>% 
  readRDS() %>% 
  filter(Trait == "Silkingtime")

# Predictor variables
mrna <- "./data/derived/maizego/mrna.RDS" %>% 
  readRDS()

# Extract genotypes for which predictor data (i.e. mRNA information) are
# available.
genotypes <- mrna %>% 
  rownames()
```

```{r Data_Inspection}
transform_boxcox <- function(x) {
  x %>% 
    BoxCoxTrans() %>% 
    predict(., x)
}

pheno %>% 
  rename(
    y = Value
  ) %>% 
  mutate(
    box_cox_y = transform_boxcox(y)
  ) %>% 
  gather(key = Preprocessing, value = Value, -Genotype, -Trait) %>% 
  ggplot(aes(x = Value)) +
  geom_histogram(bins = 80) +
  facet_wrap(~ Preprocessing, scales = "free")
```

In the case of the trait "days to silking" pre-processing of the response 
variable seems to be pointless so we will just stick with the original values, 
which are approximately normally distributed.

```{r}
# mRNA correlations
mrna_corr <- mrna %>% 
  bigcor(fun = "cor", verbose = FALSE) %>% 
  .[1:nrow(.), 1:ncol(.)]
```

```{r Prediction}
train_control <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 5
)

xgb_grid <- expand.grid(
  nrounds = 1e3,
  eta = 0.01,
  max_depth = c(2L, 10L),
  gamma = c(0, 10),
  colsample_bytree = c(0.1, 0.4),
  min_child_weight = c(1L, 10L),
  subsample = c(0.5, 1)
)

# Extract the agronomic BLUEs, ordered according to the mRNA matrix..
response <- pheno %>% 
  dplyr::slice(., match(Genotype, genotypes)) %>% 
  select(Value) %>% 
  flatten_dbl()

set.seed(340934)
xgb_res <- train(
  x = mrna,
  y = response,
  method = "xgbTree",
  tuneGrid = xgb_grid,
  trControl = train_control,
  trace = FALSE
)

yhat <- predict(xgb_res, newdata = mrna)
plot(response, yhat)

set.seed(340934)
train_idx <- seq_len(80)
xgb_fit <- train(
  x = mrna[train_idx, ],
  y = response[train_idx],
  method = "xgbTree",
  tuneGrid = data.frame(
    nrounds = 1000,
    max_depth = 10,
    eta = 0.01,
    gamma = 10,
    colsample_bytree = 0.1,
    min_child_weight = 1,
    subsample = 1   
  ),
  trControl = train_control,
  verbose = FALSE
)

cor(response[-train_idx], predict(xgb_fit, newdata = mrna[-train_idx, ]))
```

### 2017-05-10 paper/manuscript.tex
Add AEM remarks up to single step methods

@@
 Genomic prediction, pioneered by \citeNP{Meuwissen2001}, has revolutionized
-animal and plant breeding by accelerating the selection of promising candidates 
+animal and plant breeding by a more efficient selection of promising candidates 
 without phenotypes \cite{DeLosCampos2013,Garcia-Ruiz2016} and recently emerged
@@
 heritability \cite{Garcia-Ruiz2016}.
-In plant breeding, the relative excellence of a breeding strategy largely 
-depends on the prediction accuracy, which may vary widely among traits and 
-crops \cite{Longin2015}.
 
@@
 Breeding programs --- particularly in animal breeding --- are characterized by
-the availability of one predictor type, which is available for all recorded
+the presence of one predictor type, which is available for all recorded
 genotypes while other predictors are incomplete in that they cover only a 
@@
 To utilize all available phenotypic information for model training, methods were 
-needed that would allow for a combined use of genomic and pedigree information 
+developed that allowed for a combined use of genomic and pedigree information 
 \cite{Hayes2009a,VanRaden2009}.
@@
 % Two-step prediction.
-In the early years of genomic prediction, a two-step procedure was the most
-intuitive choice for establishing a new selection scheme \cite{VanRaden2009}.
+Initially, a two-step procedure was the most intuitive choice for accomplishing
+this goal \cite{VanRaden2009}.
 The first step consists of conventionally estimating breeding values based on
@@
 values are regressed onto SNP genotypes.
-The resulting model coefficients can then be used to predict genomic breeding 
+The estimated model coefficients can then be used to predict genomic breeding 
 values for solely genotyped and potentially very young individuals.
@@
 distribution of such a model.
-For instance, if the range of the prediction errors will not be identically
-distributed anymore, \textit{i.e.} they should be weighted differently or
-estimated individually \cite{Aguilar2010}.
+For instance, the range of the prediction errors will no longer be identically
+distributed, \textit{i.e.}, they should be weighted differently or estimated 
+individually \cite{Aguilar2010}.
 One popular approach to account for this problem is deregressing the breeding
@@
 the single step breeding value model of \citeNP{Legarra2009} and 
-\citeNP{Christensen2010}, which has the advantage of not requiring weights for
-$\mathbf{A}$ and $\mathbf{G}$ as well as the ability to model the structure of 
-the imputation error.
+\citeNP{Christensen2010}, which does not require weights for $\mathbf{A}$ and 
+$\mathbf{G}$.
+An additional advantage of this alternative formulation is the ability to 
+implicitly model the structure of the imputation error.
 
 % -   Mention scarcity of data on single step prediction in plant populations.
-In animal breeding, single step prediction has already become a routine 
-procedure \cite{Legarra2014} but, to the best of our knowledge, has only been 
-considered in plants in a study on wheat lines \cite{Ashraf2016}. 
+In animal breeding, single step prediction, particularly H-BLUP, has already 
+become a routine procedure \cite{Legarra2014} but, to the best of our knowledge, 
+has only been considered in plants in a study on wheat \cite{Ashraf2016} and
+white spruce lines, respectively \cite{Ratcliffe2017}. 
 Our objectives were threefold: (1) Examine whether the single step prediction
@@
 \Genetics2level{Maize hybrids}
-The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
-mating designs between 142 Dent and 103 Flint parent lines \cite{Westhues2017}. 
+The maize hybrid data set, subsequently denoted as \textit{Exp1}, comprised 
+1,521 hybrids produced in 16 factorial mating designs between 142 Dent and 103 
+Flint parent lines \cite{Westhues2017}. 
 Best linear unbiased estimates (BLUEs) were computed across all factorials and 
 three or more agro-ecologically diverse locations across Germany for six 
-agronomically important traits (Table \ref{table:TraitKey}).
+agronomically important traits by accounting for year, location, field 
+replication, block and genotype effects as well as their interactions
+(Table~\ref{table:TraitKey}).
 All 245 parent lines have pedigree records back to the generation of their 
@@
 using the Beagle software \cite{Browning2009}.
-The final number of polymorphic marker loci was 7,013 for parental Dent and
-6,212 for parental Flint lines, respectively.
+The final number of polymorphic marker loci was 7,013 for the Dent and 6,212 for 
+the Flint lines, respectively.
 Gene expression data for seedlings seven days after sowing were obtained for a 
 subset of 60 parental Dent and 43 parental Flint lines, that are parents of 685
-hybrid progeny, through two-color hybridizations using a custom 2K microarray 
-(GPL22267) \cite{Westhues2017}.
-After applying established normalization procedures 
-\cite{Smyth2003,Ritchie2007}, gene expression BLUEs were computed for 1,323 
-transcripts in reference to established protocols 
-\cite{Smyth2003,Ritchie2007,Frisch2010} using the R-package \textit{limma}
-\cite{Ritchie2015a}.
+hybrid progeny.
+These data were obtained through two-color hybridizations using a custom 2K 
+microarray (GPL22267) \cite{Westhues2017}.
+After applying established normalization procedures \cite{Smyth2003,Ritchie2007},
+gene expression BLUEs were computed for 1,323 transcripts while adjusting for 
+putative effects arising from the presence of multiple factorials (\textit{cf.}
+\citeNP{Westhues2017}) using the R-package \textit{limma} \cite{Ritchie2015a}.
+
+This dataset was included to address objectives 1 and 2.
 
 \Genetics2level{Maize diversity panel}
-The maize inbred line data set originally comprised 513 maize lines representing 
-the global maize diversity and was reduced to the set of tropical and
-subtropical lines ($n = 211$) given that it is the largest of the four
-pre-classified subgroups \cite{Yang2014}.
+The maize inbred line data set, subsequently referred to as \textit{Exp2}, 
+originally comprised 513 maize lines representing the global maize diversity 
+and was reduced to the set of tropical and subtropical lines ($n = 211$) given 
+that it is the largest of the four pre-classified subgroups \cite{Yang2014}.
+We reduced the original dataset to this subset to avoid inflated predictive 
+abilities due to large differences in the average performance of the different
+subpopulations \cite{Windhausen2012}.
 All inbred lines were evaluated in five Chinese environments ranging from
@@
 Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
-six were used for this study (Table \ref{table:TraitKey}).
+six are presented in this study (Table~\ref{table:TraitKey}).
+The selection of these traits was done so that four traits were better 
+predicted by transcriptomic data and two further traits were better predicted 
+by genomic data.
 MaizeSNP50 BeadChip-data were available for the entire set of lines from the
@@
 genotypes and was kept for further analyses \cite{Li2013}.
+
+This dataset was included to address objectives 2 and 3.
 
@@
 \Genetics2level{Population structure}
-Principal component analyses (PCA) were run on scaled and centered predictor 
-matrices using the \emph{pca} function from the R-package \textit{LEA} 
-\cite{Frichot2015}.
-Estimates of ancestry coefficients were estimated based on sparse nonnegative
-matrix factorization algorithms \cite{Frichot2014} using the \emph{snmf} 
-function in \textit{LEA}.
+Separately for each data set, principal component analyses (PCA) were run on 
+scaled and centered predictor matrices using the \emph{pca} function from the 
+R-package \textit{LEA} \cite{Frichot2015}.
+Least-squares estimates of ancestry proportions were estimated based on sparse 
+nonnegative matrix factorization algorithms \cite{Frichot2014} using the 
+\emph{snmf} function in \textit{LEA}.
 Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
@@
 The corresponding feature matrix of the $l$-th group of inbred lines and the 
-$o$-th predictor ($\mathbf{W}_{lo}$) has dimensions $n \times p$ where $n$
-pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
-the number of features.
-For the diversity panel maize lines, $n$ corresponds to the number of genotypes
-whereas, in the case of the hybrid data, $n$ corresponds to the number of
-parent lines in the corresponding heterotic group.
+$o$-th predictor ($\mathbf{W}_{lo}$) has dimensions $n_{l} \times p_{o}$ where 
+$n_{l}$ pertains to the number of genotypes in the $l$-th group and $p_{o}$
+pertains to the number of features.
+Note that $l$ is only defined for \textit{Exp1} where a distinction between 
+the two heterotic groups is necessary whereas $l$ can be ignored for all 
+models pertaining to the inbred lines of the maize diversity panel in 
+\textit{Exp2}.
+To avoid mentioning special cases, we will use the general notation with 
+the index $l$ throughout this section, though.
+For the diversity panel maize lines, $n_{l}$ corresponds to the number of 
+genotypes whereas, in the case of the hybrid data, $n_{l}$ corresponds to the 
+number of parent lines in the $l$-th heterotic group.
 In generating kernels for each predictor, all features in $\mathbf{W}_{lo}$ were
@@
 where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
-respectively, $\mu$ is the fixed model intercept, $\mathbf{Z}_{lo}$ is the 
-corresponding design matrix associating the random genotype effects
-$\mathbf{g}_{lo}$ (\textit{i.e.}, breeding values) of the lines in the $l$-th 
-inbred line group and the $o$-th predictor type with $\mathbf{y}$.
-Note that, in the case of hybrid data, $\mathbf{g}_{lo}$ pertains to the
-general combining ability effects in the parental inbred lines (either Dent or
-Flint, as expressed by $l$) which are associated with the vector of
-observed hybrid performance $\mathbf{y}$ through $\mathbf{Z}_{lo}$.
+respectively, $\mu$ is the fixed model intercept, and $\epsilon$ is the residual
+error.
+For the hybrid data from \textit{Exp1}, we fit separate additive effects for 
+each of the $L$ heterotic groups and $\mathbf{g}_{lo}$ now pertains to the 
+general combining ability effects of the parental inbred lines (either Dent or 
+Flint, as expressed by $l$).
+For the maize diversity panel inbred lines from \textit{Exp2}, we fit random 
+genotype effects $\mathbf{g}_{lo}$ (\textit{i.e.}, breeding values) of the 
+lines for the $o$-th predictor.
+For both experiments, genetic effects $\mathbf_{g}_{lo}$ are associated with 
+the vector of phenotypic performance $\mathbf{y}$ through a corresponding 
+design matrix $\mathbf{Z}_{lo}$.
 The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 

### 2017-05-12 paper/manuscript.tex
Update methods

@@
 transcriptompic data (T).
-The corresponding feature matrix of the $l$-th group of inbred lines and the 
-$o$-th predictor ($\mathbf{W}_{lo}$) has dimensions $n_{l} \times p_{o}$ where 
-$n_{l}$ pertains to the number of genotypes in the $l$-th group and $p_{o}$
-pertains to the number of features.
+The corresponding feature matrix of the $l$-th group of inbred lines 
+($\mathbf{W}_{l}$) has dimensions $n_{l} \times p$ where $n_{l}$ pertains to 
+the number of genotypes in the $l$-th group and $p$ pertains to the number of 
+features.
 Note that $l$ is only defined for \textit{Exp1} where a distinction between 
@@
 number of parent lines in the $l$-th heterotic group.
-In generating kernels for each predictor, all features in $\mathbf{W}_{lo}$ were
+In generating kernels for each predictor, all features in $\mathbf{W}_{l}$ were
 centered and standardized to unit variance.
@@
 \begin{equation} \label{eq:GenomicRelationship}
-  \mathbf{K}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
+  \mathbf{K}_{l} = \frac{1}{W_{l}} \mathbf{W}_{l} \mathbf{W}_{l}^{\top},
 \end{equation}
 
-where $W_{lo}$ denotes the number of features \cite{VanRaden2008}.
+where $W_{l}$ denotes the number of features \cite{VanRaden2008}.
 In the case of pedigree data (P), coancestry coefficients were used directly
-for $\mathbf{K}_{lo}$.
+for $\mathbf{K}_{l}$.
 
@@
   \mathbf{y} = \mu + 
-  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
-  \mathbf{\epsilon},
+  \sum_{l=1}^{L} \mathbf{Z}_{l} \mathbf{g}_{l} + \mathbf{e},
 \end{equation}
@@
 where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
-respectively, $\mu$ is the fixed model intercept, and $\epsilon$ is the residual
-error.
+respectively, $\mu$ is the fixed model intercept, and $\mathbf{e}$ is the 
+residual error.
 For the hybrid data from \textit{Exp1}, we fit separate additive effects for 
-each of the $L$ heterotic groups and $\mathbf{g}_{lo}$ now pertains to the 
+each of the $L$ heterotic groups and $\mathbf{g}_{l}$ now pertains to the 
 general combining ability effects of the parental inbred lines (either Dent or 
@@
 For the maize diversity panel inbred lines from \textit{Exp2}, we fit random 
-genotype effects $\mathbf{g}_{lo}$ (\textit{i.e.}, breeding values) of the 
+genotype effects $\mathbf{g}_{l}$ (\textit{i.e.}, breeding values) of the 
 lines for the $o$-th predictor.
-For both experiments, genetic effects $\mathbf_{g}_{lo}$ are associated with 
+For both experiments, genetic effects $\mathbf{g}_{l}$ are associated with 
 the vector of phenotypic performance $\mathbf{y}$ through a corresponding 
-design matrix $\mathbf{Z}_{lo}$.
-The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
-to $\mathbf{K}_{lo} \sigma^{2}_{{g}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
+design matrix $\mathbf{Z}_{l}$.
+The random effects ($\mathbf{g}_{l}$) have expectation zero and variance equal 
+to $\mathbf{K}_{l} \sigma^{2}_{{g}^{l}}$ and $\mathbf{I} \sigma^2_{e}$ 
 for the residual error.
+Here, we did not consider the inclusion of dominance effects for the hybrid
+material after a previous investigation into the predictive ability of various
+"omics" predictors in the same material concluded no considerably gains 
+\cite{Westhues2017}.
 All prediction models were implemented using the R-packages \textit{BGLR}
@@
 \subsection{Single step prediction}
-The breeding values in a prediction model are in general given by:
+For simplicity, let us assume throughout this section that we have $L=1$.
+If the breeding values are fully captured by the features in $\mathbf{W}$,
+Eq.~\ref{eq:KBLUPModel} can then be expressed as
+
+\begin{align} \label{eq:KBLUPModelSimplified}
+	\mathbf{y} &= \mu + \mathbf{Z} \mathbf{g} + \boldsymbol{e} \\
+	&= \mu + \mathbf{Z} \mathbf{W} \boldsymbol{\alpha} + \boldsymbol{e}
+\end{align}
+\cite{Fernando2014} since the ridge regression model is equivalent to the BLUP
+model \cite{Ruppert2003}.
+
+Hence, the breeding values in a prediction model are given by:
 \begin{equation} \label{eq:mrnaebv}
@@
 
-where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model,
-which is equivalent to the BLUP model.
+where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model.
 
@@
 Further, the covariates in $\mathbf{W}$ are centered.
+In this case we need to partition the vector of breeding values $\mathbf{g}$
+into a component $\mathbf{g}_{1}$ (\textit{i.e.,} breeding values of animals 
+covered only by the complete predictor) and a component 
+$\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} breeding 
+values of animals covered by both predictors).
 The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
@@
 	&= \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
-	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
+	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
 \end{align}
+
+where $\mathbf{K}_{ij}$ are partitions of $\mathbf{K}$ that correspond to 
+$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively \cite{Fernando2014}.
 
@@
   0 
- \end{bmatrix}
+ \end{bmatrix},
 \end{equation}
+
+where $\hat{\mathbf{W}}_{1} = \mathbf{K}_{12} \mathbf{K}_{22}^{-1} \mathbf{W}_{2}$.
 

### 2017-06-15 reports/select_subpopulation.Rmd
Prepare random subsampling for prediction

---
title: Selection of a subpopulation with little structure
author: Matthias Westhues
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    fig_caption: yes
bibliography: ["~/Dropbox/Paper/library.bib"]
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
# Data and packages -------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "cowplot", 
               "ggthemes", "pophelper", "corehunter", "forcats")
#devtools::install_github("mwesthues/sspredr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 6, warning = FALSE, echo = FALSE, 
  message = FALSE
  )
```

# Rationale {-}
Based on our discussion with G. Thaller, R.L. Fernando and C. Heuer, we 
concluded that the pre-selection of a subset of individuals might be necessary
to mitigate the impact of population structure on predictive ability.

```{r pca-structure, cache=TRUE, results='hide'}
# For the second scenario, keep only names of genotypes, which are covered by 
# all data types (i.e. phenotypic, genotypic and transcriptomic).
common_genotypes <- readRDS(
  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
)
common_genotypes <- common_genotypes %>%
  reduce(intersect)

snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
  readRDS() %>%
  #.[rownames(.) %in% common_genotypes, ] %>%
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05, 
    any_missing = FALSE, remove_duplicated = TRUE
  )

# PCA ---------------------------------------------------------------------
write.lfmm(snp, "./data/derived/maizego/snp.lfmm")

# Determine the structure of the data using genotypic data.
snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                   scale = TRUE)

pc_mat <- snp_pc$projections
rownames(pc_mat) <- rownames(snp)
colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))

pc_df <- pc_mat %>%
  as.data.frame() %>%
  rownames_to_column(var = "G") %>%
  gather(key = PC, value = Score, -G) %>%
  as_data_frame() %>%
  filter(PC %in% paste0("PC_", seq_len(2))) %>%
  mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  spread(key = PC, value = Score)

# PCA ---------------------------------------------------------------------
genos <- read_tsv(
  "./data/processed/maizego/popstruc.names",
  col_names = FALSE
  ) %>% 
  select(2) %>% 
  flatten_chr()

# Select K=4
# For each individual, return the cluster with the largest coefficient.
# Assign a color to each cluster.
# Plot the PCA results with colors corresponding to the cluster with the largest
# coefficient for that individual.

get_max_cluster <- function(x) {
  x %>%   
    mutate(G = genos) %>% 
    gather(key = "Cluster", value = "AncCoef", -G) %>% 
    group_by(G) %>% 
    summarize(main_cluster = which.max(AncCoef)) %>% 
    mutate(main_cluster = main_cluster %>% as.character() %>% as.factor()) %>% 
    ungroup()
}

# STRUCTURE FILES ---------------------------------------------------------
slist <- list.files(
  path = "./data/processed/maizego",
  pattern = "popstruc[0-9]",
  full.names = TRUE
  ) %>% 
  readQ(
    files = .,
    indlabfromfile = TRUE,
    filetype = "structure"
  )

cluster_df <- slist %>% 
  map(get_max_cluster) %>% 
  map(.f = ~right_join(., y = pc_df, by = "G")) %>% 
  bind_rows(.id = "K")

change_to_k <- function(x) {
  x %>% 
    gsub(x = ., replacement = "", pattern = "_f") %>% 
    gsub(x = ., replacement = "", pattern = "popstruc")
}
```

# Analysis {-}
First, I ran STRUCTURE [@Pritchard2000] to estimate the ancestry coefficients
for each of the `r nrow(snp)` genotypes from the subset 'tropical/subtropical' 
of the [maize diversity panel](http://www.maizego.org/Resources.html).
In total, I have examined 8 putative ancestral populations (K=2 to K=9) and, 
based on combinations of bar plots and PCA plots 
(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of 
the data for each K, I concluded that K=4 represents the population structure 
best (Fig. \@ref(fig:pca-structure-plot)).

```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel. ($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations. ($\\textbf{B}$) Principal Component Analysis. Individual points are colored according to the cluster with which they have the highest overlap."}
k <- 4
g1 <- slist %>% 
  set_names(., change_to_k(names(.))) %>% 
  map(.f = ~mutate(., G = genos)) %>% 
  keep(names(.) == k) %>% 
  map(.f = ~mutate(., Central_Cluster = Cluster1 + Cluster4)) %>% 
  map(.f = ~gather(
    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
    ) %>% 
  bind_rows(.id = "K") %>% 
  mutate(G = as.factor(G)) %>% 
  ggplot(aes(
    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
    ) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  xlab("Genotype") +
  ylab("Ancestry Coefficient")

g2 <- cluster_df %>%
  mutate(K = change_to_k(K)) %>% 
  filter(K == k) %>% 
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Cluster"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )

plot_grid(
  g1, g2, labels = c("A", "B"), nrow = 2
)
```

The set of selected individuals (Fig. \@ref(fig:pca-plot)) roughly corresponds
to clusters 1 and 4 in Fig. \@ref(fig:pca-structure-plot) and excludes 
individuals in the tails of the PCA plot.

```{r pca-plot, fig.cap="Selecting individuals based on PCA"}
selected_df <- slist %>% 
  set_names(., change_to_k(names(.))) %>% 
  map(.f = ~mutate(., G = genos)) %>% 
  keep(names(.) == k) %>% 
  .[[1]] %>% 
  left_join(y = pc_df, by = "G") %>% 
  mutate(main_cluster = case_when(
    PC1 >= -30 & PC1 <= 10 & PC2 >= -25 & PC2 <= 40 ~ "yes",
    PC1 != 0 ~ "no"
    ))

selected_df %>%
  mutate(main_cluster = as.factor(main_cluster))  %>% 
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Selected"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )
```

```{r sub-pca, cache=TRUE, results='hide', fig.cap="PCA with all 211 inbred lines from the tropical/subtropical subset of the maize diversity panel. Individuals with seemingly negligible structure are being selected (blue dots)."}
selected_geno <- selected_df %>% 
  filter(main_cluster == "yes") %>% 
  select(G) %>% 
  flatten_chr()

sel_snps <- snp %>% 
  .[rownames(snp) %in% selected_geno, ] %>% 
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05, 
    any_missing = FALSE, remove_duplicated = TRUE
  )
write.lfmm(sel_snps, "./data/derived/maizego/sel_snps.lfmm")
# Determine the structure of the data using genotypic data.
sel_snp_pc <- LEA::pca(
  input.file = "./data/derived/maizego/sel_snps.lfmm",
  scale = TRUE
  )

sel_pc_mat <- sel_snp_pc$projections
rownames(sel_pc_mat) <- rownames(sel_snps)
colnames(sel_pc_mat) <- paste0("PC_", seq_len(ncol(sel_pc_mat)))
```

```{r sub-pca-plot, fig.cap="PCA of the selcted inviduals from Fig. 2."}
sel_pc_df <- sel_pc_mat %>%
  as.data.frame() %>%
  rownames_to_column(var = "G") %>%
  gather(key = PC, value = Score, -G) %>%
  as_data_frame() %>%
  filter(PC %in% paste0("PC_", seq_len(2))) %>%
  mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  spread(key = PC, value = Score)

sel_pc_df %>% 
  ggplot(aes(PC1, PC2)) +
  geom_point() +
  theme_pander(base_size = 15)
```

Finally, I ran a PCA on the subset of `r nrow(sel_snps)` selected individuals 
to check whether there is still structure in the population 
(Fig. \@ref(fig:sub-pca-plot)).
I'm not pleased with this result as this plot still suggests structure in the
subpopulation.
We might argue, as G. Thaller suggested, that a breeding population will always 
be structured to an extent, however, I doubt that a reviewer would be convinced
by our rationale of subsetting a population when the subset itself still 
displays this structure.

## Possible solution {-}
Should we omit Fig. \@ref(fig:sub-pca-plot) and present only 
Fig. \@ref(fig:pca-plot)?

# References

### 2017-06-15 analysis/maizego_snp_analyses.R

@@ -1,7 +1,9 @@
 # Data and packages -------------------------------------------------------
 if (!require("pacman")) install.packages("pacman")
 pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "cowplot", 
-               "ggthemes")
+               "ggthemes", "pophelper", "corehunter", "forcats")
+#devtools::install_github("mwesthues/sspredr")
+pacman::p_load_gh("mwesthues/sspredr")
 
 # For the second scenario, keep only names of genotypes, which are covered by 
 # all data types (i.e. phenotypic, genotypic and transcriptomic).
@@ -13,7 +15,7 @@ common_genotypes <- common_genotypes %>%
 
 snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
   readRDS() %>%
-  .[rownames(.) %in% common_genotypes, ] %>%
+  #.[rownames(.) %in% common_genotypes, ] %>%
   sspredr::ensure_snp_quality(
     ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05, 
     any_missing = FALSE, remove_duplicated = TRUE
@@ -26,120 +28,114 @@ write.lfmm(snp, "./data/derived/maizego/snp.lfmm")
 # Determine the structure of the data using genotypic data.
 snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                    scale = TRUE)
-# Perform Tracy-Widom tests on all eigenvalues to determine the optimal number
-# of principal components.
-snp_tw <- tracy.widom(snp_pc)
-tw_K <- snp_tw %>%
-  .["pvalues"] %>%
-  flatten_dbl() %>%
-  keep(~ .x <= 0.0001) %>%
-  length()
-tw_K <- min(c(tw_K, 5))
 
-snp_tw %>%
-  .["percentage"] %>%
-  flatten_dbl() %>%
-  plot()
-
-# Qualitatively assign the genotypes to the previously determined 
-# subpopulations.
-lfmm2geno("./data/derived/maizego/snp.lfmm", 
-          output.file = "./data/derived/maizego/snp.geno")
-# This code snippet is commented out because it takes a long time to run and 
-# can easily and inadvertantly be overridden.
[…]

### 2017-06-15 analysis/prepare_subsamples.R

@@ -1,0 +1,92 @@
+# Goal: Sample core sets of genotypes for the Yan-lab data and scenario two of
+# our predictions in increments of ten percentage points.
+if (!require("pacman")) install.packages("pacman")
+if (!require("devtools")) install.packages("devtools")
+#devtools::install_github("mwesthues/sspredr", update = FALSE)
+pacman::p_load("tidyverse", "data.table", "dtplyr")
+pacman::p_load_gh("mwesthues/sspredr")
+
+# For the second scenario, keep only names of genotypes, which are covered by 
+# all data types (i.e. phenotypic, genotypic and transcriptomic).
+common_genotypes <- readRDS(
+  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
+)
+common_genotypes <- common_genotypes %>%
+  reduce(intersect)
+
+## --- Scenario 2 SNP preparation.
+# Explore the kinship matrix.
+genos <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
+  readRDS() %>%
+  rownames()
+
+runs <- 10
+fractions = seq(from = 0.1, to = 1, by = 0.1)
+
+
+
+#### Functions
+compute_totals_from_fractions <- function(nm, frac) {
+  # nm: vector of names
+  # frac: fractions
+  totals <- nm %>%
+    length() %>%
+    map(function(i) i * frac) %>%
+    flatten_dbl() %>%
+    ceiling()
+}
+
+
+sample_fraction <- function(nm, frac) {
+  # nm: names of genotypes
+  # frac: vector of genotypes fractions
+  totals <- compute_totals_from_fractions(
+    nm = common_genotypes,
+    frac = frac
+  )
+  totals %>%
+    map(.f = ~sample(
+      x = nm,
+      size = .,
+      replace = FALSE
+    )) %>%
+  set_names(., nm = as.character(frac)) %>%
+  stack()
+}
+
+
+sample_loo <- function(geno, frac, iter) {
+  train_geno <- map(geno, .f = ~ setdiff(geno, .))
+  names(train_geno) <- geno
+  train_lst <- rerun(.n = iter, {
+    sub_train_lst <- map(train_geno, .f = ~sample_fraction(., frac = frac))
+    sub_train_df <- bind_rows(sub_train_lst, .id = "TST_Geno")
[…]

### 2017-06-15 analysis/sample_and_predict.R

@@ -194,6 +194,7 @@ pred_lst <- pred_lst[match(names(pred_lst), pred_sets)]
 
 ## -- CORE SET SUBSAMPLING OF SNPS -----------------------------------------
 if (isTRUE(nchar(core_set) != 0) && data_type == "Inbred") {
+  #### BUG !!!!
   core_lst <- readRDS("./data/derived/maizego/augmented_core_list.RDS")
   names(core_lst) <- gsub(names(core_lst), pattern = "selected_fraction_",
                           replacement = "")

### 2017-09-01 reports/resampling_report.Rmd
init resampling pred report

---
title: Selection of a subpopulation with little structure
author: Matthias Westhues
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: readable
    fig_caption: yes
highlight: tango
documentclass: book
bibliography: ["~/Dropbox/Paper/library.bib"]
biblio-style: apalike
link-citations: yes
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
# Data and packages -------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  "tidyverse",
  "viridis",
  "ggthemes",
  "forcats",
  "LEA",
  "pophelper",
  "cowplot",
  "dtplyr",
  "data.table",
  "gsubfn"
  )
#devtools::install_github("mwesthues/sspredr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 6, warning = FALSE, echo = FALSE,
  message = FALSE
  )
```

# Overview
The purpose of this project is to explore the utility of single-step prediction
of phenotype performance in a set of maize inbred lines and maize hybrids,
respectively.
The methods used herein are based on the seminal papers by @Legarra2009 and
@Christensen2010 with extensions introduced by @Fernando2014.

# Data Preparation
We analyze hybrid maize data from the public breeding program of the University
of Hohenheim as well as [data on inbred lines](http://www.maizego.org/Resources.html) from the lab of doctor Jianbing Yan.
The data on inbred lines were downloaded on 2017-01-31 from the [Baidu Cloud](https://pan.baidu.com/s/1eQH3hfW#list/path=%2F)
set up by the Yan lab and are stored under `data/input/maizego`.
The Hohenheim data were generated as part of the publication "Omics-based Hybrid
Prediction in Maize" [@Westhues2017].

## Agronomic Data
For the maize inbred lines we focused on the following traits that were
previously described in [Guo et al. (2016)](https://link.springer.com/article/10.1007/s00122-016-2780-5):

*   `Cob weight` (CW)
*   `Days to silking` (DS)
*   `Ear diameter` (ED)
*   `100-grain weight` (GW)
*   `Kernel width` (KW)
*   `Plant height` (PH)

and explored the explored the distribution (including skewness) of each trait
and the phenotypic correlation between traits.

In the case of maize hybrids we analyzed the traits

*   `Dry matter yield` (DMY)
*   `Dry matter content` (DMC)
*   `Fat` (FAT)
*   `Protein` (PRO)
*   `Starch` (STA)
*   `Sugar` (SUG)

## Genomic Data
The preparation of genomic marker data for hybrids and inbred lines,
respectively, involved the following five steps (for hybrid data these were
applied separately to each of the two heterotic groups):

1.   Remove loci with a callfrequency $\geq 0.95$.

2.   Remove loci with heterozygosity $\geq 0.05$.

3.   Remove loci with minor allele frequency $\geq 0.05$.

4.   Impute remaining missing values using BEAGLE using 25 iterations and 20
     samples per iterations.

5.   Remove copies (i.e. loci in perfect LD with a previous one) of marker
     loci.

## Transcriptomic Data
For both material types transcriptomic data were already sufficiently
pre-processed.

## Population Structure
Principal component analyses were run for the [hybrid](analysis/pca.R) and
[inbred](analysis/maizego_snp_analyses.R) material.

The separation between the Dent and the Flint heterotic group for the hybrid
material [looked as expected](analysis/pca.R) but for the inbred lines we
found unexpected clustering of the lines, as determined via a [STRUCTURE](analysis/maizego_structure.R),
analysis and a [PCA](analysis/maizego_snp_anlyses.R), albeit it with very low
fractions of variance explained by the first five principal components.
See \@ref(population-structure) for a more detailed report.
Inbred lines that belonged to the clusters `1` and `4`, which were rather
homogeneous, were assigned to scenario `A`.
We defined a second scenario `B` which simply comprised all available inbred
lines and ran it to have predictive abilities for all genotypes in case
population structure was no issue for this material.

# Predictions
## Preparation
### Previous work for this manuscript
In the first draft we wanted to evaluate the influence of the genetic
constitutation of the set of genotypes that has only data on one out of two
predictors by generating core samples  (https://github.com/mwesthues/single_step/commit/8f8b39198be7d5091954f84ec0c5834afd1c3cfa).
We defined core samples as a subset of genotypes that is covered by two
predictors whereas all other genotypes (*i.e.* the complement) are only covered
by a single predictor.
The first application involved assembling core sets of varying sizes by
maximizing the average genetic distance among core set members using the
Modifed Rogers (MR) distance as the criterion ([Thachuk et al. (2009)](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-10-243)).
The size of the core set was varied in increments of ten percentage points and
ranged from 10% to 90% of all inbred lines.
The issue with this approch is that we could not preclude effects of population
structure in our material on predictive abilities.
Therefore, we decided to replace the core sampling procedure by a nested random
subsampling procedure.

### Nested subsampling scheme
To ensure that a potential bias due to population structure would be
represented by an increased standard error of the predictive abilities we
conceived a nested resampling scheme.
A first level of sampling was applied to each combination of `Material` (*i.e.*
"Hybrid" or "Inbred"), `Scenario` ("A" and "B" for inbred lines; "None" for
hybrids), `Extent` ("Core" if every predictor was available for all genotypes
and "Full for the entire set of genotypes") and `Core_Fraction`, representing
the share of core-set genotypes for which data on all predictors were assumed
to be available.
The notion of `Core_Fraction` may sound contradictory to the concept of "Core"
and "Full" sets of genotypes but note that for the variable `Core_Fraction` we
artificially removed information on a predictor for some genotypes to implore
the influence of genetic space-coverage through incomplete predictors on
predictive ability.
This random subsampling procedure was repeated 20 times and stored as
`Rnd_Level2`.

For all core-set inbred lines we applied a second, nested randomization scheme
where `n * frac` genotypes were declared to have both data on the complete and
the incomplete predictor whereas `n - n * frac` genotypes had only information
on the complete predictor.
Here `frac` denotes the fraction of all genotypes having information on both
predictors.
This second randomization was nested within each randomization stored in
`Rnd_Level2` and applied 20 times per randomization in `Rnd_Level2`.
Therefore, we ended up with 20 sets of predictions for each combination of
`"Material" * "Extent" * "Scenario"` when `frac = 1` and `20 ** 2 = 400`
sets of predictions for each combination of `"Material" * "Extent" *
"Scenario" * "Core_Fraction"` when `frac != 1`.

### Prediction template
In order to use the resources on the server as efficiently as possible,
similar combinations of the various parameters considered (`Trait`,
`Predictor`, `Material`, `Extent`, `Scenario`, `Rnd_Level1`, `Rnd_Level2`,
`Core_Fraction`) were clustered together (coded as `Interval`) in a
prediction template.

After running some speed tests for some of these combinations a
script for the automatic generation of `msub` commands was built.

## Execution
All predictions were run on the [bwunicluster](https://www.bwhpc-c5.de/wiki/index.php/Category:BwUniCluster)
by entering the msub commands into the terminal on the cluster.
For information on how to use the `bwunicluster`, in case you have access to
it, can be found [here](https://mwesthues.github.io/bwunicluster.html).

After running the prediction, the results from all clusters were concatenated.

## Bootstrap
Separately for each combination of `Material`, `Extent`, `Scenario`, `Trait`,
`Core_Fraction`, `Predictor`, `Rnd_Level1` and `Rnd_Level2`, a
bootstrap of 10,000 runs was applied to a data frame comprising observed and
predicted phenotypic values.
This yielded 10,000 bootstrap predictive abilities for each of the
aforementioned combinations together with a corresponding standard error.
The final predictive ability was calculated by taking the arithmetic mean
of bootstrapped predictive abilities over `Rnd_Level1` and `Rnd_Level2`.
The corresponding standard error was calculated across the average bootstrapped
standard error across `Rnd_Level1` and `Rnd_Level`.

## Visualization
```{r prediction-data, cache=FALSE, results='hide'}
boot_df <- "./data/processed/predictions/bootstrapped_predictions.RDS" %>%
  readRDS() %>%
  dplyr::group_by(Combi, Trait, Core_Fraction, Predictor) %>%
  dplyr::summarize(
    avg_r = mean(r),
    se = mean(std.error)
  ) %>%
  dplyr::ungroup()

## -- HYBRIDS -----------------------------------------------------------------
# Specify a unified predictor order.
pred_order <- c("P", "G", "T", "PG", "PT", "GT")

# Specify a unique trait order.
trait_order <- c("DMY", "DMC", "ADL", "FAT", "PRO", "STA", "SUG")

hyb_df <- boot_df %>%
  dplyr::filter(Combi %in% c("CHN", "FHN")) %>%
  dplyr::mutate(Trait = forcats::fct_recode(
    Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  dplyr::mutate(Combi = forcats::fct_recode(
    Combi,
    "Core" = "CHN",
    "Full" = "FHN"
  )) %>%
  dplyr::mutate(Trait = forcats::fct_relevel(Trait, trait_order)) %>%
  dplyr::mutate(
    Predictor = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "mrna" = "T",
        "ped" = "P",
        "snp" = "G",
        "ped_snp" = "PG",
        "ped_mrna" = "PT",
        "snp_mrna" = "GT"
      ),
      x = Predictor
    ),
    Predictor = forcats::fct_relevel(Predictor, pred_order)
  )

# Because the bars and errorbars have different widths
# we need to specify how wide the objects we are dodging are
dodge <- position_dodge(width = 0.9)

# Predefine the color scheme.
mycols <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6)
names(mycols) <- c("P", "G", "T", "PG", "PT", "GT")
limits <- aes(
  ymax = `Predictive Ability` + se,
  ymin = `Predictive Ability` - se
  )

## -- INBREDS CORE_FRACTION == 1 ----------------------------------------------
inbred_df <- boot_df %>%
  dplyr::filter(!Combi %in% c("CHN", "FHN")) %>%
  dplyr::mutate(
    Predictor = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "snp" = "G",
        "mrna" = "T",
        "snp_mrna" = "GT"
      ),
      x = Predictor
    ),
    Trait = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "Plantheight" = "PH",
        "Eardiameter" = "ED",
        "100grainweight" = "GW",
        "cobweight" = "CW",
        "Kernelwidth" = "KW",
        "Silkingtime" = "DS"
      ),
      x = Trait
    )
  ) %>%
  dplyr::mutate(Combi = forcats::fct_recode(
    Combi,
    "Core" = "CHN",
    "Full" = "FHN"
  ))

# Predefine the color scheme.
# Use also the levels of predictors in the hybrid data set to ensure that,
# regardless of the material, the same color is picked for the same predictor.
a_colors <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6) %>%
  set_names(c("P", "G", "T", "PG", "PT", "GT")) %>%
  .[names(.) %in% c("G", "T", "GT")]
```

```{r hybrid-plot, cache=FALSE, fig.cap="Hybrid Plot"}
hyb_df %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
  geom_bar(stat = "identity", position = dodge, color = "black") +
  geom_errorbar(limits, position = dodge, width = 0.25) +
  facet_grid(. ~ Combi) +
  scale_fill_manual(values = mycols) +
  ggthemes::theme_pander(base_size = 10)
```

```{r inbred-plot, cache=FALSE, fig.cap="Inbred Plot"}
inbred_df %>%
  dplyr::filter(Core_Fraction == "1") %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
  geom_bar(stat = "identity", position = dodge) +
  geom_errorbar(limits, position = dodge, width = 0.25) +
  facet_grid(Combi ~ .) +
  scale_fill_manual(values = a_colors) +
  theme_pander(base_size = 10)
```

```{r core-inbred-plot, cache=FALSE, fig.cap="Core Plot"}
inbred_df %>%
  dplyr::filter(Predictor == "GT", Combi %in% c("CIA", "CIB", "CIC")) %>%
  dplyr::mutate_at(vars(Predictor), funs(as.factor)) %>%
  dplyr::rename(`Core Fraction` = "Core_Fraction") %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(
    x = `Core Fraction`,
    y = `Predictive Ability`,
    color = Trait,
    group = Trait
    )
  ) +
  geom_line(stat = "identity", position = position_dodge(width = 0.3)) +
  facet_grid(~ Combi) +
  ggthemes::scale_color_tableau() +
  ggthemes::theme_pander(base_size = 10) +
  theme(legend.position = "right") +
  guides(color = guide_legend(override.aes = list(size = 3)))
```

# Appendix {-}
## Population Structure {-}
Based on our last discussion we concluded that the pre-selection of a subset
of individuals might be necessary to mitigate the impact of population structure
on predictive ability.

```{r pca-structure, cache=TRUE, results='hide'}
# For the second scenario, keep only names of genotypes, which are covered by
# all data types (i.e. phenotypic, genotypic and transcriptomic).
common_genotypes <- readRDS(
  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
)
common_genotypes <- common_genotypes %>%
  purrr::reduce(intersect)

snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
  readRDS() %>%
  #.[rownames(.) %in% common_genotypes, ] %>%
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
    any_missing = FALSE, remove_duplicated = TRUE
  )

# PCA ---------------------------------------------------------------------
write.lfmm(snp, "./data/derived/maizego/snp.lfmm")

# Determine the structure of the data using genotypic data.
snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                   scale = TRUE)

pc_mat <- snp_pc$projections
rownames(pc_mat) <- rownames(snp)
colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))

pc_df <- pc_mat %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "G") %>%
  tidyr::gather(key = PC, value = Score, -G) %>%
  tibble::as_data_frame() %>%
  dplyr::filter(PC %in% paste0("PC_", seq_len(2))) %>%
  dplyr::mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  tidyr::spread(key = PC, value = Score)

# PCA ---------------------------------------------------------------------
genos <- readr::read_tsv(
  "./data/processed/maizego/popstruc.names",
  col_names = FALSE
  ) %>%
  dplyr::select(2) %>%
  purrr::flatten_chr()

# Select K=4
# For each individual, return the cluster with the largest coefficient.
# Assign a color to each cluster.
# Plot the PCA results with colors corresponding to the cluster with the largest
# coefficient for that individual.

get_max_cluster <- function(x) {
  x %>%
    dplyr::mutate(G = genos) %>%
    tidyr::gather(key = "Cluster", value = "AncCoef", -G) %>%
    dplyr::group_by(G) %>%
    dplyr::summarize(main_cluster = which.max(AncCoef)) %>%
    dplyr::mutate(main_cluster = main_cluster %>%
      as.character() %>%
      as.factor()
    ) %>%
    dplyr::ungroup()
}

# STRUCTURE FILES ---------------------------------------------------------
slist <- list.files(
  path = "./data/processed/maizego",
  pattern = "popstruc[0-9]",
  full.names = TRUE
  ) %>%
  pophelper::readQ(
    files = .,
    indlabfromfile = TRUE,
    filetype = "structure"
  )

cluster_df <- slist %>%
  purrr::map(get_max_cluster) %>%
  purrr::map(.f = ~dplyr::right_join(., y = pc_df, by = "G")) %>%
  dplyr::bind_rows(.id = "K")

change_to_k <- function(x) {
  x %>%
    gsub(x = ., replacement = "", pattern = "_f") %>%
    gsub(x = ., replacement = "", pattern = "popstruc")
}
```

First, I ran STRUCTURE [@Pritchard2000] to estimate the ancestry coefficients
for each of the `r nrow(snp)` genotypes from the subset 'tropical/subtropical'
of the [maize diversity panel](http://www.maizego.org/Resources.html).
In total, I have examined 8 putative ancestral populations (K=2 to K=9) and,
based on combinations of bar plots and PCA plots
(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of
the data for each K, I concluded that K=4 represents the population structure
best (Fig. \@ref(fig:pca-structure-plot)).

```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel. ($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations. ($\\textbf{B}$) Principal Component Analysis. Individual points are colored according to the cluster with which they have the highest overlap."}
k <- 4
g1 <- slist %>%
  purrr::set_names(., change_to_k(names(.))) %>%
  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
  purrr::keep(names(.) == k) %>%
  purrr::map(.f = ~dplyr::mutate(., Central_Cluster = Cluster1 + Cluster4)) %>%
  purrr::map(.f = ~tidyr::gather(
    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
    ) %>%
  dplyr::bind_rows(.id = "K") %>%
  dplyr::mutate(G = as.factor(G)) %>%
  ggplot(aes(
    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
    ) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  xlab("Genotype") +
  ylab("Ancestry Coefficient")

g2 <- cluster_df %>%
  dplyr::mutate(K = change_to_k(K)) %>%
  dplyr::filter(K == k) %>%
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Cluster"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )

cowplot::plot_grid(
  g1, g2, labels = c("A", "B"), nrow = 2
)
```

The set of selected individuals (Fig. \@ref(fig:pca-plot)) roughly corresponds
to clusters 1 and 4 in Fig. \@ref(fig:pca-structure-plot) and excludes
individuals in the tails of the PCA plot.

```{r pca-plot, fig.cap="Selecting individuals based on PCA"}
selected_df <- slist %>%
  purrr::set_names(., change_to_k(names(.))) %>%
  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
  purrr::keep(names(.) == k) %>%
  .[[1]] %>%
  dplyr::left_join(y = pc_df, by = "G") %>%
  dplyr::mutate(main_cluster = dplyr::case_when(
    PC1 >= -30 & PC1 <= 10 & PC2 >= -25 & PC2 <= 40 ~ "yes",
    PC1 != 0 ~ "no"
    ))

selected_df %>%
  dplyr::mutate(main_cluster = as.factor(main_cluster))  %>%
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Selected"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )
```

# References

### 2017-09-01 reports/resampling_report.Rmd
legend for sampling scheme fig

@@
 ### Nested subsampling scheme
+```{r sampling-prediction-scheme, fig.show='hold', fig.cap="Prediction sampling scheme.
+($\\textbf{a}$) Level 1 sampling scheme.
+During the first round of sampling, 70% of the genotypes were drawn at random (black symbols) and assigned to the training set (TRN) whereas phenotypes for the other 30% of genotypes (yellow symbols) were set to missing for this particular random sample.
+The random samples drawn at this level were subsequently denoted as s~1~ to s~20~.
+($\\textbf{b}$) Level 2 sampling scheme.
+For maize inbred lines from the diversity panel, some genotypes had both genotypic and transcriptomic predictor data (black symbols).
+For each level of s from the first round of random sampling, a second round of random sub-sampling was applied where x% of the genotypes from the training set of s were declared to have both, genotypic and transcriptomic data (black symbols) whereas 1-x% of the genotypes were declared to have only genotypic information (gray symbols).
+This nested subsampling process was repeated 20 times and for nine different values of $x = (0.1, 0.2, 0.3, \\dots, 0.9)$.
+($\\textbf{c}$) Leave-one-out cross-validation (LOOCV) scheme for maize diversity panel inbred lines.
+For each of the subsets from level 1 (and level 2) LOOCV was applied.
+Thereto, $n$ TST were generated (yellow symbols) where all genotypes eligible for TRN membership based on $\\textbf{a}$ were assigned to TRN (black symbols).
+($\\textbf{d}$) LOOCV scheme for maize hybrids.
+This scheme is based solely on level 1 randomization in $\\textbf{a}$ and, for each level of s, places only T0 hybrids, which do not share a common parent with a TST hybrid, inside the TRN."}
+knitr::include_graphics("compound_plot.svg", auto_pdf = FALSE)
+```
+
 To ensure that a potential bias due to population structure would be

### 2017-09-01 reports/resampling_report.Rmd
send resampling pred report

@@
 ---
-title: Selection of a subpopulation with little structure
+title: Single-Step Prediction with Resampling
 author: Matthias Westhues
@@
   bookdown::html_document2:
+    self_contained: true
     toc: true
@@
 
-# Overview
-The purpose of this project is to explore the utility of single-step prediction
-of phenotype performance in a set of maize inbred lines and maize hybrids,
-respectively.
-The methods used herein are based on the seminal papers by @Legarra2009 and
-@Christensen2010 with extensions introduced by @Fernando2014.
-
 # Data Preparation
-We analyze hybrid maize data from the public breeding program of the University
+We analyzed hybrid maize data from the public breeding program of the University
 of Hohenheim as well as [data on inbred lines](http://www.maizego.org/Resources.html) from the lab of doctor Jianbing Yan.
 The data on inbred lines were downloaded on 2017-01-31 from the [Baidu Cloud](https://pan.baidu.com/s/1eQH3hfW#list/path=%2F)
-set up by the Yan lab and are stored under `data/input/maizego`.
+set up by the Yan lab.
 The Hohenheim data were generated as part of the publication "Omics-based Hybrid
@@
 For the maize inbred lines we focused on the following traits that were
-previously described in [Guo et al. (2016)](https://link.springer.com/article/10.1007/s00122-016-2780-5):
+previously described in @Guo2016:
 
@@
 
-and explored the explored the distribution (including skewness) of each trait
-and the phenotypic correlation between traits.
-
 In the case of maize hybrids we analyzed the traits
@@
 *   `Dry matter content` (DMC)
+*   `Lignin` (ADL)
 *   `Fat` (FAT)
@@
 
-1.   Remove loci with a callfrequency $\geq 0.95$.
+1.   Remove loci with a callfrequency $<0.95$.
 
@@
 
-3.   Remove loci with minor allele frequency $\geq 0.05$.
+3.   Remove loci with minor allele frequency $<0.05$.
 
@@
 ## Population Structure
-Principal component analyses were run for the [hybrid](analysis/pca.R) and
-[inbred](analysis/maizego_snp_analyses.R) material.
-
+Principal component analyses were run for the hybrid and inbred material.
 The separation between the Dent and the Flint heterotic group for the hybrid
-material [looked as expected](analysis/pca.R) but for the inbred lines we
-found unexpected clustering of the lines, as determined via a [STRUCTURE](analysis/maizego_structure.R),
-analysis and a [PCA](analysis/maizego_snp_anlyses.R), albeit it with very low
-fractions of variance explained by the first five principal components.
-See \@ref(population-structure) for a more detailed report.
+material looked as expected but for the inbred lines we found unexpected
+clustering of the lines, as determined via a STRUCTURE analysis [@Pritchard2000]
+and a PCA, albeit it with very low fractions of variance explained by the first
+five principal components.
+In total, I have examined 8 putative ancestral populations (K=2 to K=9) and,
+based on combinations of bar plots and PCA plots
+(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of
+the data for each K, I chose K=4 to represent the population structure
+(Fig. \@ref(fig:pca-structure-plot)).
+Based on our last discussion we concluded that the pre-selection of a subset
+of individuals might be necessary to mitigate the impact of population structure
+on predictive ability.
+The set of selected individuals entails individuals assigned to clusters 1 and 4
+in Fig. \@ref(fig:pca-structure-plot) and excludes individuals in the tails of
+the PCA plot.
 Inbred lines that belonged to the clusters `1` and `4`, which were rather
@@
 population structure was no issue for this material.
+Finally, we defined scenario `C` as a subset of scenario `B`, comprising the
+same set of genotypes as `B` but including a fixed effect for the cluster to
+which each genotype's ancestry was primarily assigned.
+
+```{r pca-structure, cache=TRUE, results='hide'}
+# For the second scenario, keep only names of genotypes, which are covered by
+# all data types (i.e. phenotypic, genotypic and transcriptomic).
+common_genotypes <- readRDS(
+  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
+)
+common_genotypes <- common_genotypes %>%
+  purrr::reduce(intersect)
+
+snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
+  readRDS() %>%
+  #.[rownames(.) %in% common_genotypes, ] %>%
+  sspredr::ensure_snp_quality(
+    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
+    any_missing = FALSE, remove_duplicated = TRUE
+  )
+
+# PCA ---------------------------------------------------------------------
+write.lfmm(snp, "./data/derived/maizego/snp.lfmm")
+
+# Determine the structure of the data using genotypic data.
+snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
+                   scale = TRUE)
+
+pc_mat <- snp_pc$projections
+rownames(pc_mat) <- rownames(snp)
+colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))
+
+pc_df <- pc_mat %>%
+  as.data.frame() %>%
+  tibble::rownames_to_column(var = "G") %>%
+  tidyr::gather(key = PC, value = Score, -G) %>%
+  tibble::as_data_frame() %>%
+  dplyr::filter(PC %in% paste0("PC_", seq_len(2))) %>%
+  dplyr::mutate(
+    PC = gsub("PC_", replacement = "PC", x = PC)
+  ) %>%
+  tidyr::spread(key = PC, value = Score)
+
+# PCA ---------------------------------------------------------------------
+genos <- readr::read_tsv(
+  "./data/processed/maizego/popstruc.names",
+  col_names = FALSE
+  ) %>%
+  dplyr::select(2) %>%
+  purrr::flatten_chr()
+
+# Select K=4
+# For each individual, return the cluster with the largest coefficient.
+# Assign a color to each cluster.
+# Plot the PCA results with colors corresponding to the cluster with the largest
+# coefficient for that individual.
+
+get_max_cluster <- function(x) {
+  x %>%
+    dplyr::mutate(G = genos) %>%
+    tidyr::gather(key = "Cluster", value = "AncCoef", -G) %>%
+    dplyr::group_by(G) %>%
+    dplyr::summarize(main_cluster = which.max(AncCoef)) %>%
+    dplyr::mutate(main_cluster = main_cluster %>%
+      as.character() %>%
+      as.factor()
+    ) %>%
+    dplyr::ungroup()
+}
+
+# STRUCTURE FILES ---------------------------------------------------------
+slist <- list.files(
+  path = "./data/processed/maizego",
+  pattern = "popstruc[0-9]",
+  full.names = TRUE
+  ) %>%
+  pophelper::readQ(
+    files = .,
+    indlabfromfile = TRUE,
+    filetype = "structure"
+  )
+
+cluster_df <- slist %>%
+  purrr::map(get_max_cluster) %>%
+  purrr::map(.f = ~dplyr::right_join(., y = pc_df, by = "G")) %>%
+  dplyr::bind_rows(.id = "K")
+
+change_to_k <- function(x) {
+  x %>%
+    gsub(x = ., replacement = "", pattern = "_f") %>%
+    gsub(x = ., replacement = "", pattern = "popstruc")
+}
+```
+
+```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel.
+($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations.
+($\\textbf{B}$) Principal Component Analysis.
+Individual points are colored according to the cluster with which they have the highest overlap."}
+k <- 4
+g1 <- slist %>%
+  purrr::set_names(., change_to_k(names(.))) %>%
+  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
+  purrr::keep(names(.) == k) %>%
+  purrr::map(.f = ~dplyr::mutate(., Central_Cluster = Cluster1 + Cluster4)) %>%
+  purrr::map(.f = ~tidyr::gather(
+    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
+    ) %>%
+  dplyr::bind_rows(.id = "K") %>%
+  dplyr::mutate(G = as.factor(G)) %>%
+  ggplot(aes(
+    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
+    ) +
+  geom_bar(stat = "identity") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_pander(base_size = 15) +
+  theme(
+    axis.text.x = element_blank(),
+    axis.ticks = element_blank(),
+    legend.title = element_blank(),
+    legend.position = "top"
+  ) +
+  guides(fill = guide_legend(ncol = 2)) +
+  xlab("Genotype") +
+  ylab("Ancestry Coefficient")
+
+g2 <- cluster_df %>%
+  dplyr::mutate(K = change_to_k(K)) %>%
+  dplyr::filter(K == k) %>%
+  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
+  geom_point(size = 1) +
+  guides(color = guide_legend(
+    override.aes = list(size = 3),
+    title = "Cluster"
+    )) +
+  scale_color_brewer(palette = "Set1") +
+  theme_pander(base_size = 15) +
+  theme(
+    legend.position = "top",
+    strip.background = element_blank(),
+    strip.placement = "outside"
+  )
+
+cowplot::plot_grid(
+  g1, g2, labels = c("A", "B"), nrow = 2
+)
+```
 
@@
 In the first draft we wanted to evaluate the influence of the genetic
-constitutation of the set of genotypes that has only data on one out of two
-predictors by generating core samples  (https://github.com/mwesthues/single_step/commit/8f8b39198be7d5091954f84ec0c5834afd1c3cfa).
+constitution of the set of genotypes that has only data on one out of two
+predictors by generating core samples.
 We defined core samples as a subset of genotypes that is covered by two
@@
 maximizing the average genetic distance among core set members using the
-Modifed Rogers (MR) distance as the criterion ([Thachuk et al. (2009)](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-10-243)).
+Modifed Rogers (MR) distance as the criterion [@Thachuk2009].
 The size of the core set was varied in increments of ten percentage points and
@@
 ### Nested subsampling scheme
+A first level of sampling was applied to each combination of `Material` (*i.e.*
+"Hybrid" or "Inbred"), `Scenario` ("A", "B" and "C" for inbred lines; "None" for
+hybrids), `Extent` ("Core" if every predictor was available for all genotypes
+and "Full for the entire set of genotypes") and `Core_Fraction`, representing
+the share of core-set genotypes for which data on all predictors were assumed
+to be available (Fig. \@ref(fig:sampling-prediction-scheme) **a**,
+Table \@ref(tab:genotype-count)).
+For the variable `Core_Fraction` we artificially removed information on a
+predictor for some genotypes to implore the influence of genetic space-coverage
+through incomplete predictors on predictive ability.
+This random subsampling procedure was repeated 20 times and stored as
+`Rnd_Level2`.
+
+```{r genotype-count, cache=TRUE}
+html_table_width <- function(kable_output, width) {
+  width_html <- paste0(paste0('<col width = "', width, '">'), collapse = "\n")
+  sub("<table>", paste0("<table>\n", width_html), kable_output)
+}
+
+"./data/processed/predictions/predictions.RDS" %>%
+  readRDS() %>%
+  dplyr::mutate(Combi = substr(Interval, start = 1, stop = 3)) %>%
+  dplyr::group_by(Combi) %>%
+  dplyr::distinct(TST_Geno) %>%
+  dplyr::count() %>%
+  dplyr::rename(`Extent-Material-Scenario` = "Combi") %>%
+  knitr::kable(
+    x = .,
+    booktabs = TRUE,
+    caption = "Number of genotypes (n) for each combination of extent (F = full, C = core), material (I = inbred, H = hybrid) and scenario."
+  ) %>%
+  html_table_width(c(200, 100))
+```
+
+For all core-set inbred lines we applied a second, nested randomization scheme
+where $n * x\%$ genotypes were declared to have both, data on the complete and
+the incomplete predictor, whereas $n - n * x\%$ genotypes had only information
+on the complete predictor (Fig. \@ref(fig:sampling-prediction-scheme) **b**).
+Here $x\%$ denotes the fraction of all genotypes having information on both
+predictors.
+This second randomization was nested within each randomization stored in
+`Rnd_Level2` and applied 20 times per randomization in `Rnd_Level2`.
+Therefore, we ended up with 20 sets of predictions for each combination of
+`"Material" * "Extent" * "Scenario"` when $x\% = 100$ and `20 ** 2 = 400`
+sets of predictions for each combination of `"Material" * "Extent" *
+"Scenario" * "Core_Fraction"` when $x\% \neq 100$.
+
 ```{r sampling-prediction-scheme, fig.show='hold', fig.cap="Prediction sampling scheme.
@@
 For maize inbred lines from the diversity panel, some genotypes had both genotypic and transcriptomic predictor data (black symbols).
-For each level of s from the first round of random sampling, a second round of random sub-sampling was applied where x% of the genotypes from the training set of s were declared to have both, genotypic and transcriptomic data (black symbols) whereas 1-x% of the genotypes were declared to have only genotypic information (gray symbols).
-This nested subsampling process was repeated 20 times and for nine different values of $x = (0.1, 0.2, 0.3, \\dots, 0.9)$.
+For each level of s from the first round of random sampling, a second round of random sub-sampling was applied where x% of the genotypes from the training set of s were declared to have both, genotypic and transcriptomic data (black symbols) whereas 100-x% of the genotypes were declared to have only genotypic information (gray symbols).
+This nested subsampling process was repeated 20 times and for nine different values of $x = (10\\%, 20\\%, 30\\%, \\dots, 90\\%)$.
 ($\\textbf{c}$) Leave-one-out cross-validation (LOOCV) scheme for maize diversity panel inbred lines.
@@
 ```
-
-To ensure that a potential bias due to population structure would be
-represented by an increased standard error of the predictive abilities we
-conceived a nested resampling scheme.
-A first level of sampling was applied to each combination of `Material` (*i.e.*
-"Hybrid" or "Inbred"), `Scenario` ("A" and "B" for inbred lines; "None" for
-hybrids), `Extent` ("Core" if every predictor was available for all genotypes
-and "Full for the entire set of genotypes") and `Core_Fraction`, representing
-the share of core-set genotypes for which data on all predictors were assumed
-to be available.
-The notion of `Core_Fraction` may sound contradictory to the concept of "Core"
-and "Full" sets of genotypes but note that for the variable `Core_Fraction` we
-artificially removed information on a predictor for some genotypes to implore
-the influence of genetic space-coverage through incomplete predictors on
-predictive ability.
-This random subsampling procedure was repeated 20 times and stored as
-`Rnd_Level2`.
-
-For all core-set inbred lines we applied a second, nested randomization scheme
-where `n * frac` genotypes were declared to have both data on the complete and
-the incomplete predictor whereas `n - n * frac` genotypes had only information
-on the complete predictor.
-Here `frac` denotes the fraction of all genotypes having information on both
-predictors.
-This second randomization was nested within each randomization stored in
-`Rnd_Level2` and applied 20 times per randomization in `Rnd_Level2`.
-Therefore, we ended up with 20 sets of predictions for each combination of
-`"Material" * "Extent" * "Scenario"` when `frac = 1` and `20 ** 2 = 400`
-sets of predictions for each combination of `"Material" * "Extent" *
-"Scenario" * "Core_Fraction"` when `frac != 1`.
-
-### Prediction template
-In order to use the resources on the server as efficiently as possible,
-similar combinations of the various parameters considered (`Trait`,
-`Predictor`, `Material`, `Extent`, `Scenario`, `Rnd_Level1`, `Rnd_Level2`,
-`Core_Fraction`) were clustered together (coded as `Interval`) in a
-prediction template.
-
-After running some speed tests for some of these combinations a
-script for the automatic generation of `msub` commands was built.
-
-## Execution
-All predictions were run on the [bwunicluster](https://www.bwhpc-c5.de/wiki/index.php/Category:BwUniCluster)
-by entering the msub commands into the terminal on the cluster.
-For information on how to use the `bwunicluster`, in case you have access to
-it, can be found [here](https://mwesthues.github.io/bwunicluster.html).
-
-After running the prediction, the results from all clusters were concatenated.
 
@@
 
-```{r hybrid-plot, cache=FALSE, fig.cap="Hybrid Plot"}
+```{r hybrid-plot, cache=FALSE, fig.cap="Predictive abilities for seven agronomic traits in maize hybrids based on three different predictors and combinations thereof.
+The bootstrapped standard error is displayed through error bars.
+The predictors are pedigree (P), genomic data (G) and transcriptomic data (T).
+Two letters indicate single step prediction where the first letter denotes the complete predictor whereas the second letter denotes the incomplete predictor.
+The 'Core' set of 685 hybrids entails all hybrids with records for all predictors.
+The 'Full' set comprises 1521 hybrids with data for P and G and the subset of 685 hybrids that additionally has information on T."}
 hyb_df %>%
@@
 
-```{r inbred-plot, cache=FALSE, fig.cap="Inbred Plot"}
+```{r inbred-plot, cache=FALSE, fig.cap="Predictive abilities for six agronomic traits in maize inbred lines based on two different predictors and their combination.
+The bootstrapped standard error is displayed through error bars.
+The predictors are genomic data (G) and transcriptomic data (T).
+Two letters indicate single step prediction where the first letter denotes the complete predictor and the second letter denotes the incomplete predictor.
+Different combinations of extent (C = 'Core' and F = 'Full') with prediction scenario (A, B, or C) are plotted separately along the y-axis where the fist letter indicates the extent of the material and the third letter denotes the scenario.
+The 'Core' set of 149 inbred lines entails all genotypes with records for both predictors.
+The 'Full' set comprises 211 genotypes with complete data for G and the subset of 149 inbred lines that additionally has information on T.
+Note that scenario A only entails the 110 inbred lines that belong to clusters 1 and 4 in the PCA (Fig.
+3.5)."}
 inbred_df %>%
@@
 
-```{r core-inbred-plot, cache=FALSE, fig.cap="Core Plot"}
+```{r core-inbred-plot, cache=FALSE, fig.cap="Influence of the size of the subset of genotypes with both, genomic and transcriptomic data on predictive abilities.
+Results are shown for six agronomic traits in the core set of 149 maize inbred lines ('CIB' and 'CIC') and the core set of 110 genotypes ('CIA'), respectively.
+The fraction of genotypes with information on both predictors is denoted as 'Core Fraction' and displayed on the x-axis."}
 inbred_df %>%
@@
 
-# Appendix {-}
-## Population Structure {-}
-Based on our last discussion we concluded that the pre-selection of a subset
-of individuals might be necessary to mitigate the impact of population structure
-on predictive ability.
-
-```{r pca-structure, cache=TRUE, results='hide'}
-# For the second scenario, keep only names of genotypes, which are covered by
-# all data types (i.e. phenotypic, genotypic and transcriptomic).
-common_genotypes <- readRDS(
-  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
-)
-common_genotypes <- common_genotypes %>%
-  purrr::reduce(intersect)
-
-snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
-  readRDS() %>%
-  #.[rownames(.) %in% common_genotypes, ] %>%
-  sspredr::ensure_snp_quality(
-    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
-    any_missing = FALSE, remove_duplicated = TRUE
-  )
-
-# PCA ---------------------------------------------------------------------
-write.lfmm(snp, "./data/derived/maizego/snp.lfmm")
-
-# Determine the structure of the data using genotypic data.
-snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
-                   scale = TRUE)
-
-pc_mat <- snp_pc$projections
-rownames(pc_mat) <- rownames(snp)
-colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))
-
-pc_df <- pc_mat %>%
-  as.data.frame() %>%
-  tibble::rownames_to_column(var = "G") %>%
-  tidyr::gather(key = PC, value = Score, -G) %>%
-  tibble::as_data_frame() %>%
-  dplyr::filter(PC %in% paste0("PC_", seq_len(2))) %>%
-  dplyr::mutate(
-    PC = gsub("PC_", replacement = "PC", x = PC)
-  ) %>%
-  tidyr::spread(key = PC, value = Score)
-
-# PCA ---------------------------------------------------------------------
-genos <- readr::read_tsv(
-  "./data/processed/maizego/popstruc.names",
-  col_names = FALSE
-  ) %>%
-  dplyr::select(2) %>%
-  purrr::flatten_chr()
-
-# Select K=4
-# For each individual, return the cluster with the largest coefficient.
-# Assign a color to each cluster.
-# Plot the PCA results with colors corresponding to the cluster with the largest
-# coefficient for that individual.
-
-get_max_cluster <- function(x) {
-  x %>%
-    dplyr::mutate(G = genos) %>%
-    tidyr::gather(key = "Cluster", value = "AncCoef", -G) %>%
-    dplyr::group_by(G) %>%
-    dplyr::summarize(main_cluster = which.max(AncCoef)) %>%
-    dplyr::mutate(main_cluster = main_cluster %>%
-      as.character() %>%
-      as.factor()
-    ) %>%
-    dplyr::ungroup()
-}
-
-# STRUCTURE FILES ---------------------------------------------------------
-slist <- list.files(
-  path = "./data/processed/maizego",
-  pattern = "popstruc[0-9]",
-  full.names = TRUE
-  ) %>%
-  pophelper::readQ(
-    files = .,
-    indlabfromfile = TRUE,
-    filetype = "structure"
-  )
-
-cluster_df <- slist %>%
-  purrr::map(get_max_cluster) %>%
-  purrr::map(.f = ~dplyr::right_join(., y = pc_df, by = "G")) %>%
-  dplyr::bind_rows(.id = "K")
-
-change_to_k <- function(x) {
-  x %>%
-    gsub(x = ., replacement = "", pattern = "_f") %>%
-    gsub(x = ., replacement = "", pattern = "popstruc")
-}
-```
-
-First, I ran STRUCTURE [@Pritchard2000] to estimate the ancestry coefficients
-for each of the `r nrow(snp)` genotypes from the subset 'tropical/subtropical'
-of the [maize diversity panel](http://www.maizego.org/Resources.html).
-In total, I have examined 8 putative ancestral populations (K=2 to K=9) and,
-based on combinations of bar plots and PCA plots
-(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of
-the data for each K, I concluded that K=4 represents the population structure
-best (Fig. \@ref(fig:pca-structure-plot)).
-
-```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel.
-($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations.
-($\\textbf{B}$) Principal Component Analysis.
-Individual points are colored according to the cluster with which they have the highest overlap."}
-k <- 4
-g1 <- slist %>%
-  purrr::set_names(., change_to_k(names(.))) %>%
-  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
-  purrr::keep(names(.) == k) %>%
-  purrr::map(.f = ~dplyr::mutate(., Central_Cluster = Cluster1 + Cluster4)) %>%
-  purrr::map(.f = ~tidyr::gather(
-    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
-    ) %>%
-  dplyr::bind_rows(.id = "K") %>%
-  dplyr::mutate(G = as.factor(G)) %>%
-  ggplot(aes(
-    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
-    ) +
-  geom_bar(stat = "identity") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_pander(base_size = 15) +
-  theme(
-    axis.text.x = element_blank(),
-    axis.ticks = element_blank(),
-    legend.title = element_blank(),
-    legend.position = "top"
-  ) +
-  guides(fill = guide_legend(ncol = 2)) +
-  xlab("Genotype") +
-  ylab("Ancestry Coefficient")
-
-g2 <- cluster_df %>%
-  dplyr::mutate(K = change_to_k(K)) %>%
-  dplyr::filter(K == k) %>%
-  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
-  geom_point(size = 1) +
-  guides(color = guide_legend(
-    override.aes = list(size = 3),
-    title = "Cluster"
-    )) +
-  scale_color_brewer(palette = "Set1") +
-  theme_pander(base_size = 15) +
-  theme(
-    legend.position = "top",
-    strip.background = element_blank(),
-    strip.placement = "outside"
-  )
-
-cowplot::plot_grid(
-  g1, g2, labels = c("A", "B"), nrow = 2
-)
-```
-
-The set of selected individuals (Fig. \@ref(fig:pca-plot)) roughly corresponds
-to clusters 1 and 4 in Fig. \@ref(fig:pca-structure-plot) and excludes
-individuals in the tails of the PCA plot.
-
-```{r pca-plot, fig.cap="Selecting individuals based on PCA"}
-selected_df <- slist %>%
-  purrr::set_names(., change_to_k(names(.))) %>%
-  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
-  purrr::keep(names(.) == k) %>%
-  .[[1]] %>%
-  dplyr::left_join(y = pc_df, by = "G") %>%
-  dplyr::mutate(main_cluster = dplyr::case_when(
-    PC1 >= -30 & PC1 <= 10 & PC2 >= -25 & PC2 <= 40 ~ "yes",
-    PC1 != 0 ~ "no"
-    ))
-
-selected_df %>%
-  dplyr::mutate(main_cluster = as.factor(main_cluster))  %>%
-  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
-  geom_point(size = 1) +
-  guides(color = guide_legend(
-    override.aes = list(size = 3),
-    title = "Selected"
-    )) +
-  scale_color_brewer(palette = "Set1") +
-  theme_pander(base_size = 15) +
-  theme(
-    legend.position = "top",
-    strip.background = element_blank(),
-    strip.placement = "outside"
-  )
-```
-
 # References

### 2017-09-06 paper/manuscript.tex
fix latex warnings in manuscript

@@
     Stuttgart 70599, GERMANY \\
-    Tel.: (+49) 0711 459-22334 \\
-    Fax: (+49) 0711 459-22343 \\ 
+    Tel.: (+49) 0711 459--22334 \\
+    Fax: (+49) 0711 459--22343 \\ 
     \texttt{melchinger@uni-hohenheim.de} \vfill}
@@
 animal and plant breeding by a more efficient selection of promising candidates 
-without phenotypes \cite{DeLosCampos2013,Garcia-Ruiz2016} and recently emerged
-as an important tool in human diagnostics \cite{DeLosCampos2010,Vazquez2016}.
+without phenotypes~\cite{DeLosCampos2013,Garcia-Ruiz2016} and recently emerged
+as an important tool in human diagnostics~\cite{DeLosCampos2010,Vazquez2016}.
 Its use is largely motivated by the high costs and oftentimes long duration of 
@@
 obtained from thorough resampling routines applied to the subset of individuals 
-with full information \cite{Hastie2009}.
+with full information~\cite{Hastie2009}.
 In animal breeding, genomic selection has increased the annual selection gain 
-by $\approx$ 50-100\% for yield traits and by 300-400\% for traits with low 
-heritability \cite{Garcia-Ruiz2016}.
+by $\approx$ 50--100\% for yield traits and by 300--400\% for traits with low 
+heritability~\cite{Garcia-Ruiz2016}.
 
@@
 genotypes while other predictors are incomplete in that they cover only a 
-subset of all genotypes \cite{Fragomeni2015}.
+subset of all genotypes~\cite{Fragomeni2015}.
 Usually, the complete predictor is pedigree information while genomic data are
@@
 Initially, a two-step procedure was the most intuitive choice for accomplishing
-this goal \cite{VanRaden2009}.
+this goal~\cite{VanRaden2009}.
 The first step consists of conventionally estimating breeding values based on
 pedigree relationships and the phenotypes of relatives.
-Individuals with accurate breeding values out of the first step (\textit{i.e.}
+Individuals with accurate breeding values out of the first step (that is a
 high number of offspring), enter the second step in which conventional breeding
@@
 For instance, the range of the prediction errors will no longer be identically
-distributed, \textit{i.e.}, they should be weighted differently or estimated 
-individually \cite{Aguilar2010}.
+distributed, that is they should be weighted differently or estimated 
+individually~\cite{Aguilar2010}.
 One popular approach to account for this problem is deregressing the breeding
@@
 % Workings and advantages of the single step procedure.
-Independently from each other, \citeNP{Legarra2009} and
+Independently from each other,~\citeNP{Legarra2009} and
 \citeNP{Christensen2010} developed a single step BLUP method that blends the 
@@
 Issues with this H-BLUP approach were the compatability of $\mathbf{A}$ with
-$\mathbf{G}$ \cite{Christensen2012} and their weighting, which is not
-trait-independent \cite{Vitezica2011,Ashraf2016}.
+$\mathbf{G}$~\cite{Christensen2012} and their weighting, which is not
+trait-independent~\cite{Vitezica2011,Ashraf2016}.
 \citeNP{Fernando2014} derived the equivalent single step marker-effect-model to 
@@
 \citeNP{Christensen2010}, which does not require weights for $\mathbf{A}$ and 
-$\mathbf{G}$.
-An additional advantage of this alternative formulation is the ability to 
-implicitly model the structure of the imputation error.
+$\mathbf{G}$ and allows implicit modeling of the imputation error's structure.
 
@@
 In animal breeding, single step prediction, particularly H-BLUP, has already 
-become a routine procedure \cite{Legarra2014} but, to the best of our knowledge, 
-has only been considered in plants in a study on wheat \cite{Ashraf2016} and
-white spruce lines, respectively \cite{Ratcliffe2017}. 
+become a routine procedure~\cite{Legarra2014} but, to the best of our knowledge, 
+has only been considered in plants in a study on wheat~\cite{Ashraf2016} and
+white spruce lines, respectively~\cite{Ratcliffe2017}.
 Our objectives were threefold: (1) Examine whether the single step prediction
@@
 step predition with a quantitative predictor and (3) evaluate the impact of the
-genetic space covered by the incomplete predictor on predictive ability.
+proporition of genotypes not covered by the incomplete predictor on predictive
+ability.
 
@@
 1,521 hybrids produced in 16 factorial mating designs between 142 Dent and 103 
-Flint parent lines \cite{Westhues2017}. 
+Flint parent lines~\cite{Westhues2017}. 
 Best linear unbiased estimates (BLUEs) were computed across all factorials and 
@@
 All 245 parent lines have pedigree records back to the generation of their 
-grandparents or deeper \cite{Westhues2017} and were genotyped using the Illumina 
-BeadChip MaizeSNP50 \cite{Ganal2011}.
+grandparents or deeper~\cite{Westhues2017} and were genotyped using the Illumina 
+BeadChip MaizeSNP50~\cite{Ganal2011}.
 After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
 ($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed 
-using the Beagle software \cite{Browning2009}.
+using the Beagle software~\cite{Browning2009}.
 The final number of polymorphic marker loci was 7,013 for the Dent and 6,212 for 
@@
 These data were obtained through two-color hybridizations using a custom 2K 
-microarray (GPL22267) \cite{Westhues2017}.
-After applying established normalization procedures \cite{Smyth2003,Ritchie2007},
+microarray (GPL22267)~\cite{Westhues2017}.
+After applying established normalization procedures~\cite{Smyth2003,Ritchie2007},
 gene expression BLUEs were computed for 1,323 transcripts while adjusting for 
 putative effects arising from the presence of multiple factorials (\textit{cf.}
-\citeNP{Westhues2017}) using the R-package \textit{limma} \cite{Ritchie2015a}.
+\citeNP{Westhues2017}) using the R-package \textit{limma}~\cite{Ritchie2015a}.
 
@@
 and was reduced to the set of tropical and subtropical lines ($n = 211$) given 
-that it is the largest of the four pre-classified subgroups \cite{Yang2014}.
+that it is the largest of the four pre-classified subgroups~\cite{Yang2014}.
 We reduced the original dataset to this subset to avoid inflated predictive 
 abilities due to large differences in the average performance of the different
-subpopulations \cite{Windhausen2012}.
+subpopulations~\cite{Windhausen2012}.
 All inbred lines were evaluated in five Chinese environments ranging from
-$18$ to $30^{\circ}$N and from $102$ to $110^{\circ}$E \cite{Yang2014}.
+$18$ to $30^{\circ}$N and from $102$ to $110^{\circ}$E~\cite{Yang2014}.
 Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
@@
 overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
-for these lines \cite{Fu2013,Li2013}.
+for these lines~\cite{Fu2013,Li2013}.
 For the remaining 145 maize lines without RNA-seq data, high density markers
 were inferred via projection of IBD regions onto the BeadChip, using this core 
-set of markers \cite{Yang2014}.
+set of markers~\cite{Yang2014}.
 The same SNP quality checks as for the parental inbred lines of the hybrids
@@
 Gene expression data were normalized using a normal quantile transformation to
-satisfy modelling assumptions \cite{Fu2013}.
+satisfy modelling assumptions~\cite{Fu2013}.
 A subset of 28,850 annotated genes was available for 149 out of the 211
-genotypes and was kept for further analyses \cite{Li2013}.
+genotypes and was kept for further analyses~\cite{Li2013}.
 
@@
 scaled and centered predictor matrices using the \emph{pca} function from the 
-R-package \textit{LEA} \cite{Frichot2015}.
+R-package \textit{LEA}~\cite{Frichot2015}.
 Least-squares estimates of ancestry proportions were estimated based on sparse 
-nonnegative matrix factorization algorithms \cite{Frichot2014} using the 
+nonnegative matrix factorization algorithms~\cite{Frichot2014} using the 
 \emph{snmf} function in \textit{LEA}.
@@
 maximizing the average genetic distance among core set members using the 
-Modified Rogers (MR) distance as the criterion \cite{Thachuk2009}.
+Modified Rogers (MR) distance as the criterion~\cite{Thachuk2009}.
 The size of the core set was varied in increments of ten percentage points and 
@@
 
-where $W_{l}$ denotes the number of features \cite{VanRaden2008}.
+where $W_{l}$ denotes the number of features~\cite{VanRaden2008}.
 In the case of pedigree data (P), coancestry coefficients were used directly
@@
 material after a previous investigation into the predictive ability of various
-"omics" predictors in the same material concluded no considerably gains 
+``omics'' predictors in the same material concluded no considerably gains 
 \cite{Westhues2017}.
@@
 \cite{Fernando2014} since the ridge regression model is equivalent to the BLUP
-model \cite{Ruppert2003}.
+model~\cite{Ruppert2003}.
 
@@
 where $\mathbf{K}_{ij}$ are partitions of $\mathbf{K}$ that correspond to 
-$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively \cite{Fernando2014}.
+$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014}.
 
@@
 predictive abilities than the use of G for four out of six traits
-(Fig. \ref{fig:InbredResults}A).
+(Fig.~\ref{fig:InbredResults}A).
 In terms of predictive ability, the absolute advantage of T over G was as large 
-as 0.04 as observed for the traits DS, ED, GW and KW.
+as 0.04 as observed for the traits DS, ED, GW and KW\@.
 Except for KW, the predictive ability for predictions based on G were always
@@
   combination (GT) were used.
-  (\textbf{A}) The 'Core' panel includes a subset of 149 lines covered by
-  genomic and transcriptomic information whereas the 'Full' panel comprises all
-  211 genotypes. In the 'Full' set of genotypes, transcriptomic records were
+  (\textbf{A}) The `Core' panel includes a subset of 149 lines covered by
+  genomic and transcriptomic information whereas the `Full' panel comprises all
+  211 genotypes. In the `Full' set of genotypes, transcriptomic records were
   imputed for 62 lines.
@@
   with transcriptomic information on single step-based predictive abilities 
-  using a 'Core' set of 149 inbred lines.
+  using a `Core' set of 149 inbred lines.
   Transcriptomic records were deleted for 17 to 19 genotypes that were
   primarily associated with a particular subgroup in the material, as defined
-  by an admixture analysis and a PCA (Fig. \ref{fig:PopStructure}).
+  by an admixture analysis and a PCA (Fig.~\ref{fig:PopStructure}).
   (\textbf{C}) Influence of core set sampling on predictive abilities using
-  nine different sub set sizes where 'Core Fraction' indicates the fraction
+  nine different sub set sizes where `Core Fraction' indicates the fraction
   of 149 inbred lines that is covered by both, genomic and transcriptomic
@@
 Pedigree records and genomic information were available for the 245 parent 
-lines of all 1,521 hybrids and we denote these data as the 'Full' set whereas
+lines of all 1,521 hybrids and we denote these data as the `Full' set whereas
 transcriptomic data were available for 103 parent lines of 685 hybrids, which
-we denote as 'Core'.
+we denote as `Core'.
 For DMY, T was the best predictor in the core set of 685 hybrids with a 
 predictive ability that was 0.08 points higher than for genomic data, as the 
-second best predictor for this trait (Fig. \ref{fig:HybridResults}).
-In the core set, P had the highest predictive ability for STA and DMC.
+second best predictor for this trait (Fig.~\ref{fig:HybridResults}).
+In the core set, P had the highest predictive ability for STA and DMC\@.
 For the other three traits, G was the superior predictor inside the core set.
 In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
-markedly for all traits with the exception of FAT.
+markedly for all traits with the exception of FAT\@.
 The combination of genomic and transriptomic information (GT) in a single-step
-model yielded a small improvement over G alone for the traits DMY and PRO.
+model yielded a small improvement over G alone for the traits DMY and PRO\@.
 Except for PRO, compared to using only G, trait prediction did not benefit from 
@@
     combinations thereof we used.
-    The 'Full' set of genotypes comprises 1,521 hybrid progeny based on 245
-    parental inbred lines, whereas the 'Core' set comprises 685 hybrid progeny
+    The `Full' set of genotypes comprises 1,521 hybrid progeny based on 245
+    parental inbred lines, whereas the `Core' set comprises 685 hybrid progeny
     based on 103 parental inbred lines.
@@
 ranging from only 10\% to 90\% of all genotypes covered by the incomplete
-predictor (Fig. \ref{fig:CoreSetPCA}).
+predictor (Fig.~\ref{fig:CoreSetPCA}).
 
@@
 Predictive abilities obtained with the different core sets varied widely within
-traits (Fig. \ref{fig:InbredResults}B).
+traits (Fig.~\ref{fig:InbredResults}B).
 For the traits CW and PH, which were better predicted via G than via T in the 
@@
 A principal component analysis depicts the coverage of the genetic space by
-genotypes (Fig. \ref{fig:PopStructure}).
-With a core set size of 50-60\% of the 149 genotypes, the congruence between
+genotypes (Fig.~\ref{fig:PopStructure}).
+With a core set size of 50--60\% of the 149 genotypes, the congruence between
 core set and complement was high whereas genotypes in the south-west direction
@@
 Hereto, we ran an admixture analysis on the same set of 149 inbred lines 
-assuming $K=3$ ancestral populations (Fig. \ref{fig:PopStructure}).
+assuming $K=3$ ancestral populations (Fig.~\ref{fig:PopStructure}).
 
@@
   Genotypes for which none of the three ancestry coefficients reached a value
-  of 0.5 were classified as "mixed".
+  of 0.5 were classified as ``mixed''.
 }
@@
 covered by transcriptomic data only.
-Genotypes from A1 and A3, which were at the far ends in the PCA plot (Fig.
-\ref{fig:PopStructure}B), were estimated with less precision than genotypes
+Genotypes from A1 and A3, which were at the far ends in the PCA plot 
+(Fig.~\ref{fig:PopStructure}B), were estimated with less precision than genotypes
 from the central cluster of genotypes belonging largely to the A2 population
-(Fig. \ref{fig:PopStructure}A, Fig. \ref{fig:InbredResults}C).
+(Fig.~\ref{fig:PopStructure}A, Fig.~\ref{fig:InbredResults}C).
 
@@
 was covered well by genotypes that had both, transcriptomic and genomic data 
-(Fig. \ref{fig:PCA}).
+(Fig.~\ref{fig:PCA}).
 
@@
 Single step approaches in hybrid populations have so far been limited to pig
-breeding \cite{Xiang2015,Xiang2016,Tusell2016}.
+breeding~\cite{Xiang2015,Xiang2016,Tusell2016}.
 All of these studies used breeding value-based single step BLUP methods
 (SS-BLUP) conceived by \citeNP{Legarra2009} and \citeNP{Christensen2010} with
-extensions to hybrid populations \cite{Christensen2014,Christensen2015}.
+extensions to hybrid populations~\cite{Christensen2014,Christensen2015}.
 An attactive property of SS-BLUP models is their computational advantage over 
@@
 majority of old individuals has only pedigree records introduces a bias that 
-has to be accounted for \cite{Vitezica2011,Legarra2015,Garcia-Baccino2017}.
+has to be accounted for~\cite{Vitezica2011,Legarra2015,Garcia-Baccino2017}.
 Third, the addition of new genotypes requires updates of $\mathbf{G}$ that
@@
 \cite{Jiang2015,Guo2016,Vazquez2016}, describing the interactions 
-within and between different biological strata \cite{Sackton2016}.
+within and between different biological strata~\cite{Sackton2016}.
 Such interactions were found to be pervasive throughout the genomes of yeast 
-\cite{Brem2005} and humans \cite{Brown2014} and have motivated studies on the 
-utility of downstream 'omics' predictors for integrating such interactions.
-Recently, encouraging results have been found for humans \cite{Vazquez2016}, 
-maize inbred lines \cite{Guo2016} and maize hybrids \cite{Westhues2017}.
+\cite{Brem2005} and humans~\cite{Brown2014} and have motivated studies on the 
+utility of downstream `omics' predictors for integrating such interactions.
+Recently, encouraging results have been found for humans~\cite{Vazquez2016}, 
+maize inbred lines~\cite{Guo2016} and maize hybrids~\cite{Westhues2017}.
 While single step prediction has become an established method when complete
@@
 not yet been considered for the imputation of a complete predictor based on 
-incomplete downstream 'omics' predictors such as transcriptomic, proteomic or 
+incomplete downstream `omics' predictors such as transcriptomic, proteomic or 
 metabolomic data.
@@
 Hybrid breeding --- with noteworthy commercial applications in pig
-\cite{Xiang2016,Tusell2016} and maize breeding \cite{TheRoyalSociety2009} --- 
+\cite{Xiang2016,Tusell2016} and maize breeding~\cite{TheRoyalSociety2009} --- 
 is a particularly challenging application field for prediction tools 
@@
 In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
-technology \cite{Wedzony2009} now allows for an annual production of 1,000 
+technology~\cite{Wedzony2009} now allows for an annual production of 1,000 
 parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
@@
 A principal component analysis showed that, at a core set size of 0.7, 
-genotypes assigned to the A3 ancestral population (Fig. \ref{fig:PopStructure})
+genotypes assigned to the A3 ancestral population (Fig.~\ref{fig:PopStructure})
 were largely elements of the core set whereas genotypes with a large share at 
@@
 of 0.7, which is in agreement with the reduced predictive abilities observed
-when stripping A1 genotypes from their transcriptomic information (Fig.
-\ref{fig:InbredResults}B).
+when stripping A1 genotypes from their transcriptomic information
+(Fig.~\ref{fig:InbredResults}B).
 
@@
 We have successfully applied single step prediction to a hybrid plant data set 
-while imputing with a quantitative downstream 'omics' predictor.
+while imputing with a quantitative downstream `omics' predictor.
 By declaring different subsets of individuals as covered by one or two 
@@
 The authors acknowledge support by the state of Baden-W{\"u}rttemberg through 
-bwHPC.
+bwHPC\@.
 Financial support for M.W. was provided by the Fiat Panis foundation, Ulm,

### 2017-09-06 paper/manuscript.tex
paper, describe new subsampling scheme

@@
 \Genetics2level{Maize hybrids}
-The maize hybrid data set, subsequently denoted as \textit{Exp1}, comprised 
-1,521 hybrids produced in 16 factorial mating designs between 142 Dent and 103 
-Flint parent lines~\cite{Westhues2017}. 
+To address the objectives 1 and 2, a maize hybrid data set, subsequently
+denoted as \textit{Exp1}, was used.
+The material comprised 1,521 hybrids produced in 16 factorial mating designs
+between 142 Dent and 103 Flint parent lines~\cite{Westhues2017}. 
 Best linear unbiased estimates (BLUEs) were computed across all factorials and 
@@
 After applying established normalization procedures~\cite{Smyth2003,Ritchie2007},
-gene expression BLUEs were computed for 1,323 transcripts while adjusting for 
-putative effects arising from the presence of multiple factorials (\textit{cf.}
-\citeNP{Westhues2017}) using the R-package \textit{limma}~\cite{Ritchie2015a}.
-
-This dataset was included to address objectives 1 and 2.
+BLUEs for gene expression were computed for 1,323 transcripts while adjusting
+for putative effects arising from the presence of multiple factorials
+(\textit{cf.} \citeNP{Westhues2017}) using the R-package
+\textit{limma}~\cite{Ritchie2015a}.
 
 \Genetics2level{Maize diversity panel}
-The maize inbred line data set, subsequently referred to as \textit{Exp2}, 
-originally comprised 513 maize lines representing the global maize diversity 
-and was reduced to the set of tropical and subtropical lines ($n = 211$) given 
-that it is the largest of the four pre-classified subgroups~\cite{Yang2014}.
-We reduced the original dataset to this subset to avoid inflated predictive 
-abilities due to large differences in the average performance of the different
-subpopulations~\cite{Windhausen2012}.
+To address objectives 2 and 3 a maize inbred line data set, subsequently
+referred to as \textit{Exp2}, was used~\cite{Yang2014}.
+Exp2 originally comprised 513 maize lines representing the global maize
+diversity  and was reduced to the set of tropical and subtropical lines
+($n = 211$) given that it is the largest of the four pre-classified
+subgroups~\cite{Yang2014}.
 All inbred lines were evaluated in five Chinese environments ranging from
@@
 
-This dataset was included to address objectives 2 and 3.
-
 % Please add the following required packages to your document preamble:
@@
 \toprule
-Trait              & Acronym & Data Set  \\ \midrule
+Trait              & Abbreviation & Data Set  \\ \midrule
 Cob weight         & CW           & Diversity \\
@@
 \Genetics2level{Population structure}
-Separately for each data set, principal component analyses (PCA) were run on 
-scaled and centered predictor matrices using the \emph{pca} function from the 
-R-package \textit{LEA}~\cite{Frichot2015}.
+Population structure in Exp2 was examined by running a STRUCTURE
+\cite{Pritchard2000} analysis with the number of putative ancestral
+populations $K$ ranging from 2 to 9.
+Based on the estimated probability of the data for each $K$, we chose $K=4$
+and ran a principal component analysis (PCA) on the scaled and centered
+genomic feature matrix using the \emph{pca} function from the R-pacakge LEA
+\cite{Frichot2015}.
 Least-squares estimates of ancestry proportions were estimated based on sparse 
@@
 \emph{snmf} function in \textit{LEA}.
-Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
-repetitions for each value of $K$. 
-For each $K$, only the repetition with the lowest cross-entropy was kept for 
-further analyses.
-
-\Genetics2level{Core sampling}
-In order to evaluate the influence of the genetic constitution of the set of
-genotypes that has only data on one out of two predictors, we generated core
-samples based on two different approaches.
-All core sets were assembled from the 149 maize inbred lines from the diversity
-panel which had gene expression as well as genomic information.
-We defined core samples as a subset of genotypes that is covered by two
-predictors whereas all other genotypes (\textit{i.e.}, the complement) are only
-covered by a single predictor.
-The first application involved assembling core sets of varying sizes by 
-maximizing the average genetic distance among core set members using the 
-Modified Rogers (MR) distance as the criterion~\cite{Thachuk2009}.
-The size of the core set was varied in increments of ten percentage points and 
-ranged from 10\% to 90\% of all 149 inbred lines.
-For the second application, we first ran an admixture analysis in combination
-with a PCA and kept $K=3$ as the number of putative ancestral population, which 
-seemed to describe the population structure best.
-Genotypes were then assigned to one of the three ancestral populations based on
-whether any of the three relationship coefficients (A1-A3) was equal to or
-greater than 0.5.
-For example, a genotype with an admixture coefficient for A2 $\geq 0.5$ was
-assigned to the second ancestral population.
-Thereby, we obtained $n_{A1} = 17$, $n_{A2} = 109$, $n_{A3} = 19$ and four 
-genotypes that could not be unambiguously assigned to any putative ancestral 
-population based on this criterion.
-To ensure commensurability of the three subpopulations within the set of 149
-genotypes, the subset of 109 genotypes (\textit{i.e.}, members of A2) was
-reduced to $n_{A2} = 19$ genotypes, again using the method of maximizing the
-average genetic distance among selected genotypes described above.
+Genotypes in the PCA plot (Fig.~\ref{fig:pca}) were colorized based on their
+primary assignment to any of the four putative ancestral populations.
+To avoid inflated predictive abilities, which can arise from large differences
+in the average performance of the different subpopulations
+\cite{Windhausen2012}, we reduced the panel of inbred lines to those primarily
+belonging to either cluster 1 or cluster 4 in Fig.~\ref{fig:pca}.
 
@@
 hybrids).
+
+\Genetics2level{Core sampling}
+In order to evaluate the influence of the fraction of genotypes that is not
+covered by the incomplete predictor on predictive ability we generated core
+sets of nine different sizes from Exp2 (Fig.~\ref{fig:compound-plot}~a).
+The base of each of these core sets was the superset of genotypes that were
+covered by both, genomic and trancriptomic data.
+A core set was created by artificially declaring a pre-specified fraction 
+$x \in {10\%, 20\%, \dots, 90\%}$ of genotypes as lacking of transcriptomic 
+information.
+For each value of $x$ we created 100 core sets $s \in {1, 2, 3, \dots, 100}$ by
+repeatedly sampling genotypes, whose trancsriptomic information was then 
+declared as missing, at random (Fig.~\ref{fig:compound-plot}~b).
+The artificially removed transriptomic information on these genotypes was then
+imputed.
 

### 2017-09-06 paper/manuscript.tex
paper, improve t0 trn description

@@
 between $D$ and $F$.
-The training set ($H_{TRN}$) for the hybrid to be predicted ($H_{ij}$) --- 
-where $i \in D_{TRN}$ and $j \in F_{TRN}$ --- was assembled as 
-$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$ with
-$D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.
+Let $H_{i \times j}$ denote the test set hybrid to be predicted.
+We can then define the parents of the training set hybrids as
+$D_{i} = D \setminus \{i\}$ and $F_{j} = F \setminus \{j\}$, respectively.
+The training set is then defined as 
+$H_{TRN}(x \times j) = H \cap (D_{i} \times F_{j})$.
 
@@
 $\mathbf{\hat{y}}$ is the vector of predicted values from each LOOCV run.
-Our confidence in the performance of each model was evalutaed by computing the
-coefficient of variation for the predicted values:
-
-\begin{equation} \label{eq:CV}
-CV = \sqrt{\frac{\sum_{i = 1}^{n} \sigma^{2}_{\hat{y}(i)} / n}{\sum_{i=1}^{n}\hat{y}_{i} / n}},
-\end{equation}
-
-where $n$ denotes the number of genotypes (\textit{i.e.}, inbred lines or 
-hybrids).
 

### 2017-09-06 analysis/bootstrap_predictions.R
adjust bootstrap to single lvl sampling

@@ -11,7 +11,7 @@ pacman::p_load(
 # Set the number of cores to use in parallel.
 if (interactive()) {
 
-  use_cores <- 4L
+  use_cores <- 3L
 
 } else {
 
@@ -46,8 +46,7 @@ boot_group_nms <- dat %>%
     Trait,
     Core_Fraction,
     Predictor,
-    Rnd_Level1,
-    Rnd_Level2
+    Rnd_Level1
   ) %>%
   dplyr::mutate(
     Core_Fraction = dplyr::if_else(
@@ -70,7 +69,6 @@ boot_group_nms <- dat %>%
     Core_Fraction,
     Predictor,
     Rnd_Level1,
-    Rnd_Level2,
     sep = "-"
   ) %>%
   dplyr::pull(Boot_Group)
@@ -81,8 +79,7 @@ boot_group_vars <- c(
   "Trait",
   "Core_Fraction",
   "Predictor",
-  "Rnd_Level1",
-  "Rnd_Level2"
+  "Rnd_Level1"
   )
 
 # Run the bootstrap.

### 2017-09-06 paper/manuscript.tex
paper, describe bootstrap
paper, captions for new figures

@@
 \emph{snmf} function in \textit{LEA}.
-Genotypes in the PCA plot (Fig.~\ref{fig:pca}) were colorized based on their
+Genotypes in the PCA plot (Fig.~\ref{fig:inbred-pca}) were colorized based on their
 primary assignment to any of the four putative ancestral populations.
@@
 \cite{Windhausen2012}, we reduced the panel of inbred lines to those primarily
-belonging to either cluster 1 or cluster 4 in Fig.~\ref{fig:pca}.
+belonging to either cluster 1 or cluster 4 in Fig.~\ref{fig:inbred-pca}.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics{./tables_figures/inbred_pca.pdf}
+  \caption{
+  Principal component analysis of 211 genotypes from the set
+  `tropical/subtropical' of the maize diversity panel inbred lines.
+  The four clusters are based on $K=4$ putative ancestral populations in a
+  STRUCTURE analysis of the corresponding genomic feature matrix.
+  Genotypes are colored based on their primary assignment to any of these four
+  clusters.
+}
+\label{fig:inbred-pca}
+\end{figure}
 
@@
 
+Standard errors for the predictive abilities were calculated by bootstrapping
+predictive and observed values 1,000 times using the \emph{R} package 
+\emph{boot}~\cite{Canty2017}.
+
 \Genetics2level{Core sampling}
@@
 imputed.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=1\textwidth]{./tables_figures/compound_plot.pdf}
+  \caption{
+  (\textbf{a}) Scheme for the amount of genotypes covered by predictors.
+  The two extended bars represent the entire set of genotypes in either Exp1 or
+  Exp2.
+  All genotypes are covered by at least one predictor, which is called the
+  `complete' predictor and depicted as a white bar.
+  The second bar depicts that only a fraction of genotypes is covered by both,
+  the `complete' and the `incomplete' predictor (dark gray) and that the
+  difference between the predictors will be imputed.
+  The four bars to the right outline the principle of generating core sets for
+  Exp2.
+  For nine different values of $x$, which represents the fraction of genotypes
+  covered by the `incomplete' predictor, core sets are build.
+  (\textbf{b}) Subsampling within core sets.
+  Each of the black symbols represents a different genotype.
+  For each value of $x$, 100 subsets $s$ are generated where $x\%$ of genotypes
+  are randomly declared to be covered by both, the `complete' and the
+  `incomplete' predictor whereas the remaining $100\% - x\%$ of genotypes
+  declared to be covered only by the `complete' predictor.
+  }
+\label{fig:compound-plot}
+\end{figure}
 
@@
 
-% Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
 \begin{figure}[H]
   \centering
-  \includegraphics{./tables_figures/inbred_line_combi_plot.pdf}
+  \includegraphics{./tables_figures/pred_ability_inbred.pdf}
   \caption{
-  Predictive abilities and average coefficient of variation for
-  tropical/subtropical inbred lines from the maize diversity panel and six 
-  agronomic traits.
+  Predictive abilities and SE for tropical/subtropical inbred lines from the 
+  maize diversity panel and six agronomic traits.
   As predictors, genomic (G) data, transcriptomic (T) data and their 
   combination (GT) were used.
-  (\textbf{A}) The `Core' panel includes a subset of 149 lines covered by
-  genomic and transcriptomic information whereas the `Full' panel comprises all
-  211 genotypes. In the `Full' set of genotypes, transcriptomic records were
-  imputed for 62 lines.
-  (\textbf{B}) Influence of structural missingness in the coverage of genotypes 
-  with transcriptomic information on single step-based predictive abilities 
-  using a `Core' set of 149 inbred lines.
-  Transcriptomic records were deleted for 17 to 19 genotypes that were
-  primarily associated with a particular subgroup in the material, as defined
-  by an admixture analysis and a PCA (Fig.~\ref{fig:PopStructure}).
-  (\textbf{C}) Influence of core set sampling on predictive abilities using
-  nine different sub set sizes where `Core Fraction' indicates the fraction
-  of 149 inbred lines that is covered by both, genomic and transcriptomic
-  data.
+  The `Core' panel includes a subset of 110 lines covered by genomic and
+  transcriptomic information whereas the `Full' panel comprises all 164
+  genotypes.
+  In the `Full' set of genotypes, transcriptomic records were imputed for 54
+  lines.
   }
@@
 
-% Script: ./analysis/uhoh_plot_predictive_abilities.R
 \begin{figure}[H]
 \centering
-  \includegraphics{./tables_figures/hybrid_predictive_ability.pdf}
+  \includegraphics{./tables_figures/pred_ability_hybrid.pdf}
   \caption{
-    Predictive abilities and average coefficients of variation for the set of 
-    maize hybrids and six agronomic traits.
+    Predictive abilities and average, bootstrapped standard errors for the set
+    of maize hybrids and six agronomic traits.
     As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
@@
 ranging from only 10\% to 90\% of all genotypes covered by the incomplete
-predictor (Fig.~\ref{fig:CoreSetPCA}).
+predictor (Fig.~\ref{fig:inbred-pca}).
 
@@
   \centering
-\includegraphics{./tables_figures/maizego_core_sampling_pca.pdf}
+  \includegraphics{./tables_figures/pred_ability_inbred_core_fraction.pdf}
   \caption{
-  Coverage of the genetic space by nine different core sets. 
-  Genotypes are distributed according to their scores based on the first two
-  principal components (PC).
-  The total genetic space was spanned by 149 genotypes from the maize diversity 
-  panel for which both, genomic and transcriptomic data, were available.
-  Genotypes that were members of the core set (orange triangles) were declared 
-  as being covered by both predictors whereas members of the complement (blue,
-  filled circles) were declared as only being covered by genomic information.
-  Assignments of genotypes to the core set was done by selecting a subset of
-  genotypes that maximized the average genetic distance in the core set based 
-  on the Modified Rogers distance between all genotypes.
-  Core set sizes ranged from 10\% to 90\% of the 149 available genotypes and
-  are given as decimals above each facet of the plot.
+  Influence of the fraction of genotypes covered by the incomplete predictor
+  on predictive abilities.
+  Predictive abilities are based on transcriptomic effects derived from a
+  single step prediction model involving transcriptomic (T) and genomic (G)
+  data, respectively.
+  Results are shown for the core set of 110 inbred lines from Exp2 for which
+  both, G and T are available for any genotype and six agronomic traits.
+  The fraction of genotypes (\textit{i.e.} `Core Fraction') for which
+  T was available is given on the x-axis.
+  Dotted lines indicate that $T > G$, dashed lines indicate that $T \approx G$
+  and solid lines indicate that $G > T$ when running predictions using single
+  predictors on the same data set.
   }
-\label{fig:CoreSetPCA}
+\label{fig:InbredResults}
 \end{figure}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/maizego_admixture_k3_pca.pdf}
-\caption{
-  Population structure of 149 maize inbred lines from the maize diversity panel
-  covered by genomic and transcriptomic data.
-  (\textbf{A}) Ancestry coefficients from an admixture analysis when 
-  considering $K=3$ ancestral populations.
-  (\textbf{B}) Principal component analysis with genotypes colored according to 
-  the most likely ancestral population (\textit{i.e.}, share $\geq 0.5$) as 
-  indicated by the admixture analysis.
-  Genotypes for which none of the three ancestry coefficients reached a value
-  of 0.5 were classified as ``mixed''.
-}
-\label{fig:PopStructure}
-\end{figure}
-
 Of the 149 genotypes, 17 belonged to more than 50\% to the first ancestral
@@
 (Fig.~\ref{fig:PCA}).
-
-\begin{figure}[H]
-\centering
-\includegraphics{./tables_figures/diversity_dent_flint_pca.pdf}
-\caption{
-  Principal component (PC) analysis i) of the 211 tropical/subtropical lines 
-  from the maize diversity panel, ii) of the 142 Dent parent lines of the 1,521
-  maize hybrids and iii) of the 103 Flint parent lines of the maize hybrids.
-  All inbred lines, which are covered by genomic and transcriptomic features
-  are depicted as orange triangles whereas genotypes for which only genomic
-  information are available are depicted as blue dots.
-}
-\label{fig:PCA}
-\end{figure}
 

### 2017-09-06 analysis/visualize_predictions.R
visualize scenario a only

@@ -134,10 +134,10 @@ a_colors <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6) %>%
 inbred_plot <- inbred_df %>%
   dplyr::filter(Core_Fraction == "1") %>%
   dplyr::rename(`Predictive Ability` = "avg_r") %>%
-  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
+  ggplot(aes(x = Combi, y = `Predictive Ability`, fill = Predictor)) +
   geom_bar(stat = "identity", position = dodge) +
   geom_errorbar(limits, position = dodge, width = 0.25) +
-  facet_grid(. ~ Combi) +
+  facet_grid(Trait ~ .) +
   scale_fill_manual(values = a_colors) +
   theme_pander(base_size = 10)
 
@@ -154,9 +154,35 @@ ggsave(
 
 
 ## -- INBREDS CORE_FRACTION != 1 ----------------------------------------------
-core_plot <- inbred_df %>%
-  dplyr::filter(Predictor == "GT", Combi %in% c("CIA", "CIB", "CIC")) %>%
+line_type_df <- inbred_df %>%
+  dplyr::filter(Predictor != "GT", Combi == "CIA") %>%
   dplyr::mutate_at(vars(Predictor), funs(as.factor)) %>%
+  dplyr::select(-se) %>%
+  tidyr::spread(key = Predictor, value = avg_r) %>%
+  dplyr::rename(Transcriptome = "T") %>%
+  dplyr::rename(Genome = "G") %>%
+  dplyr::mutate(Diff = Genome - Transcriptome) %>%
+  dplyr::mutate(BetterPredictor = dplyr::case_when(
+    Diff > 0.01 ~ "G",
+    Diff < -0.01 ~ "T",
+    TRUE ~ "None"
+    )
+  ) %>%
+  dplyr::mutate(LineType = dplyr::case_when(
+    BetterPredictor == "G" ~ "a",
+    BetterPredictor == "T" ~ "e",
+    BetterPredictor == "None" ~ "k"
+    )
+  ) %>%
+  dplyr::select(-Combi, -Core_Fraction)
+
+
+core_df <- inbred_df %>%
+  dplyr::filter(Predictor == "GT", Combi == "CIA") %>%
+  dplyr::mutate_at(vars(Predictor), funs(as.factor)) %>%
+  dplyr::inner_join(y = line_type_df, by = "Trait")
+
+core_plot <- core_df %>%
[…]

### 2017-09-06 paper/manuscript.tex
paper, minor methods fixes

@@
 pedigree relationships and the phenotypes of relatives.
-Individuals with accurate breeding values out of the first step (that is a
+Individuals with accurate breeding values out of the first step (that is, a
 high number of offspring), enter the second step in which conventional breeding
@@
 For instance, the range of the prediction errors will no longer be identically
-distributed, that is they should be weighted differently or estimated 
+distributed, \textit{i.e.}, they should be weighted differently or estimated 
 individually~\cite{Aguilar2010}.
@@
 \Genetics2level{Maize hybrids}
-To address the objectives 1 and 2, a maize hybrid data set, subsequently
+To address objectives 1 and 2, a maize hybrid data set, subsequently
 denoted as \textit{Exp1}, was used.
@@
 \Genetics2level{Maize diversity panel}
-To address objectives 2 and 3 a maize inbred line data set, subsequently
+To address objectives 2 and 3, a maize inbred line data set, subsequently
 referred to as \textit{Exp2}, was used~\cite{Yang2014}.
-Exp2 originally comprised 513 maize lines representing the global maize
+\textit{Exp2} originally comprised 513 maize lines representing the global maize
 diversity  and was reduced to the set of tropical and subtropical lines
@@
 \Genetics2level{Population structure}
-Population structure in Exp2 was examined by running a STRUCTURE
+Population structure in \textit{Exp2} was examined by running a STRUCTURE
 \cite{Pritchard2000} analysis with the number of putative ancestral
@@
 belonging to either cluster 1 or cluster 4 in Fig.~\ref{fig:inbred-pca}.
+This reduction yielded 164 genotypes, 110 of which were covered by both, genomic
+and transcriptomic data.
 
@@
 For the maize diversity panel inbred lines from \textit{Exp2}, we fit random 
-genotype effects $\mathbf{g}_{l}$ (\textit{i.e.}, breeding values) of the 
-lines for the $o$-th predictor.
+genotype effects $\mathbf{g}_{l}$ (\textit{i.e.}, breeding values).
 For both experiments, genetic effects $\mathbf{g}_{l}$ are associated with 
@@
 The structure of the residual imputation/prediction error is known and can 
-therefore be modelled.
+therefore be modeled.
 The covariates not covered by the incomplete predictor can be predicted using 
@@
 covered by the incomplete predictor on predictive ability we generated core
-sets of nine different sizes from Exp2 (Fig.~\ref{fig:compound-plot}~a).
+sets of nine different sizes from \textit{Exp2}
+(Fig.~\ref{fig:compound-plot}~a).
 The base of each of these core sets was the superset of genotypes that were
@@
 A core set was created by artificially declaring a pre-specified fraction 
-$x \in {10\%, 20\%, \dots, 90\%}$ of genotypes as lacking of transcriptomic 
+$x = {10\%, 20\%, \dots, 90\%}$ of genotypes as lacking of transcriptomic 
 information.
 For each value of $x$ we created 100 core sets $s \in {1, 2, 3, \dots, 100}$ by
-repeatedly sampling genotypes, whose trancsriptomic information was then 
-declared as missing, at random (Fig.~\ref{fig:compound-plot}~b).
+repeatedly sampling genotypes, whose trancsriptomic information was declared
+as missing, at random (Fig.~\ref{fig:compound-plot}~b).
 The artificially removed transriptomic information on these genotypes was then
@@
   (\textbf{a}) Scheme for the amount of genotypes covered by predictors.
-  The two extended bars represent the entire set of genotypes in either Exp1 or
-  Exp2.
+  The two extended bars represent the entire set of genotypes in either
+  \textit{Exp1} or \textit{Exp2}.
   All genotypes are covered by at least one predictor, which is called the
@@
   The four bars to the right outline the principle of generating core sets for
-  Exp2.
+  \textit{Exp2}.
   For nine different values of $x$, which represents the fraction of genotypes
@@
   data, respectively.
-  Results are shown for the core set of 110 inbred lines from Exp2 for which
-  both, G and T are available for any genotype and six agronomic traits.
+  Results are shown for the core set of 110 inbred lines from \textit{Exp2} for
+  which both, G and T are available for any genotype and six agronomic traits.
   The fraction of genotypes (\textit{i.e.} `Core Fraction') for which

### 2017-09-06 paper/manuscript.tex
paper, update results

@@
 \Genetics2level{Inbred lines}
-At the level of the 149 genotypes for which both, genomic (G) and 
+At the level of the 110 genotypes for which both, genomic (G) and 
 transcriptomic (T) information, were available, the use of T yielded higher
-predictive abilities than the use of G for four out of six traits
+predictive abilities than the use of G for two out of six traits
 (Fig.~\ref{fig:InbredResults}A).
 In terms of predictive ability, the absolute advantage of T over G was as large 
-as 0.04 as observed for the traits DS, ED, GW and KW\@.
-Except for KW, the predictive ability for predictions based on G were always
-higher for the full set of inbred lines ($n = 211$) than for the reduced set of 
-inbred lines ($n = 149$).
+as 0.15 as observed for the trait ED\@.
+The predictive ability for predictions based on G were always higher for the
+full set of inbred lines ($n = 164$) than for the reduced set of inbred lines
+($n = 110$).
 The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
-prediction of all 211 lines, performed at least as well as G alone when T was 
-superior over G for the reduced set of 149 lines.
-In the case of GW and ED, the use of GT yielded a slight improvement in predictive 
+prediction of all 164 lines, performed at least as well as G alone when T was 
+superior over G for the reduced set of 110 lines.
+In the case of KW and ED, the use of GT yielded a slight improvement in predictive 
 ability over G.
-Where T was a worse predictor for the reduced set of 149 lines than G,
-predictive abilities of GT for the full set of 211 lines were also worse than 
+Where T was a worse predictor for the reduced set of 110 lines than G,
+predictive abilities of GT for the full set of 164 lines were also worse than 
 those based on G alone.
@@
 \Genetics2level{Hybrids}
-The maize hybrid data set contains information on six agronomic traits and
+The maize hybrid data set contains information on seven agronomic traits and
 three predictors, namely pedigree data (P), genomic information (G), 
@@
 we denote as `Core'.
-For DMY, T was the best predictor in the core set of 685 hybrids with a 
-predictive ability that was 0.08 points higher than for genomic data, as the 
-second best predictor for this trait (Fig.~\ref{fig:HybridResults}).
+In the `core' set, transcriptomic data were the best single predictor for ADL,
+PRO and particulaly for DMY where the predictive ability obtained with T
+was 0.08 points higher than for G as the second best predictor for this trait
+(Fig.~\ref{fig:HybridResults}).
 In the core set, P had the highest predictive ability for STA and DMC\@.
-For the other three traits, G was the superior predictor inside the core set.
+For the FAT and SUG, G was the superior predictor inside the core set.
 In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
@@
 information were better than P alone for all traits and PT improved upon P in
-the core set for four traits.
+the `full' set for four traits.
 
@@
 
-\Genetics2level{Coverage of the genetic space}
+\Genetics2level{Impact of coverage by the incomplete predictor}
 Single-step approaches are based on the use of one complete predictor that
@@
 predictor on the predictive abilities obtained via single-step approaches by
-focusing exclusively on the 149 inbred lines from the maize diversity panel
+focusing exclusively on the 110 inbred lines from the maize diversity panel
 that are covered by both, genomic and transcriptomic information.
-First, we assembled nine different core sets, which we defined as subsets
-of genotypes that maximize the average genetic distance between genotypes of
-the full set.
+First, we assembled nine different core sets, each with 100 nested subsamples,
+which were generated by randomly removing transcriptomic information from
+some genotypes (Fig.~\ref{fig:compound-plot}).
 The nine core sets differ in that each covers a different number of genotypes,
@@
 Predictive abilities obtained with the different core sets varied widely within
-traits (Fig.~\ref{fig:InbredResults}B).
-For the traits CW and PH, which were better predicted via G than via T in the 
-reduced set of lines, predictive abilities were higher when only a small number 
-of genotypes was included in the core set.
+most traits (Fig.~\ref{fig:InbredResults}).
+For most traits, an initial increase in predictive abilities when moving
+from a core set size of 0.1 to 0.2/0.3 was observed, regardless of whether G
+or T individually was the superior predictor.
+For the traits CW, GW and PH, which were better predicted via G than via T in
+the reduced set of lines, predictive abilities were higher when only a small
+number of genotypes was included in the core set.
 This corresponds to a situation where a large number of genotypes is being 
 imputed and the majority of the information is derived from genomic information.
-For most other traits, an initial increase in predictive abilities when moving
-from a core set size of 0.1 to 0.5/0.6 was observed, followed by a drop around
-0.6/0.7 and a subsequent hike up to a core set size of 0.9.
-A principal component analysis depicts the coverage of the genetic space by
-genotypes (Fig.~\ref{fig:PopStructure}).
-With a core set size of 50--60\% of the 149 genotypes, the congruence between
-core set and complement was high whereas genotypes in the south-west direction
-of the PCA plot were overrepresented compared to genotypes in the east 
-direction at a core set size of 0.7.
-
-As an alternative to exploring the impact of genotypes sampled based on the
-maximization of genetic distance within the core set, we wanted to explore a
-more extreme scenario where we select core sets based on population structure.
-Hereto, we ran an admixture analysis on the same set of 149 inbred lines 
-assuming $K=3$ ancestral populations (Fig.~\ref{fig:PopStructure}).
-
-Of the 149 genotypes, 17 belonged to more than 50\% to the first ancestral
-population (A1), 109 belonged to more than 50\% to the second ancestral
-population (A2) and 19 belonged to more than 50\% to the third ancestral
-population while four genotypes could not be unequivocally assigned to either
-ancestral population (mixed).
-To reconcile the large size of the A2 set compared to A1 and A3, we selected 19 
-genotypes from A2 using previously described core sampling.
-For single-step prediction, each of the three genotype sets were then declared
-as having genomic and transcriptomic information whereas their complements were
-covered by transcriptomic data only.
-Genotypes from A1 and A3, which were at the far ends in the PCA plot 
-(Fig.~\ref{fig:PopStructure}B), were estimated with less precision than genotypes
-from the central cluster of genotypes belonging largely to the A2 population
-(Fig.~\ref{fig:PopStructure}A, Fig.~\ref{fig:InbredResults}C).
-
-Regarding the hybrid material, a PCA carried out separately for the two groups
-of parental inbred lines showed that the entire genetic space in each group 
-was covered well by genotypes that had both, transcriptomic and genomic data 
-(Fig.~\ref{fig:PCA}).
+A hike in predictive abilities, when moving from a core fraction of 10\% to 
+90\%, was observed for ED and KW, which were predicted better when using T
+individually compared to using G individually.
+The only trait that was barely affected by a change in the core fraction was DS
+for which the predictive abilities using either G or T individually were almost
+identical (Fig.~\ref{fig:InbredResults}).
 

### 2017-09-07 paper/manuscript.tex
paper, update discussion to scenario a

@@
 \Genetics2level{Superiority of single step models}
-\subsection{Inbred lines}
+\subsection{Exp2}
 When using a single step approach based on complete genomic and incomplete
@@
 Another possibility is that the number of additional phenotypes in the full
-set compared to the core set ($\Delta(n) = 62$) was not large enough to provide
+set compared to the core set ($\Delta(n) = 54$) was not large enough to provide
 considerably more useful information.
@@
 possible reasons:
-1) We reduced the dataset to the 211 tropical/subtropical lines and excluded 
-all other genotypes from our analyses whereas \citeNP{Guo2016} used genotypes 
-from four different subpopulations, modeled as a fixed effect, and used only a 
-subset of 157 genotypes from the set of tropical/subtropical lines.
+1) We reduced the dataset to 164 tropical/subtropical lines with little evident
+population structure and excluded all other genotypes from our analyses whereas
+\citeNP{Guo2016} used genotypes from four different subpopulations, modeled as a
+fixed effect.
 2) We applied quality checks for the predictor data after generating the subset
-of 211 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
+of 164 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
 genotypes.
@@
 
-\subsection{Maize hybrids}
+\subsection{Exp1}
 Hybrid breeding --- with noteworthy commercial applications in pig
@@
 Thus, even the minor gains when using transcriptomic and genomic data in a
-single step might justify additional investments in RNA-seq at least of a
+single step might justify additional investments in RNA-seq at least for a
 subset of genotypes that represents the genetic space of the breeding 
@@
 either transcriptomic or genomic information.
-Evaluations of core sets for the parental inbred lines of the maize hybrids were 
-outside the scope of this study, but results on the inbred lines have shown that 
-homogeneous coverage of the genetic space by the incomplete predictor can 
-considerably leverage the predictive ability in single step prediction compared 
-to using only the complete predictor.
 
@@
 complemented with information on another predictor.
-When using core sets of varying sizes, which were established with the
-constraint of maximizing the average genetic distance among their members, we
-noticed strong fluctuations in predictive abilities when moving from small core
-sets to large core sets.
-For traits that were better predicted by transcriptomic than by genomic
-information in the core set, predictive abilities typically peaked around
-core set size of 0.5/0.6, then dropped sharply at 0.7 and started to increase
-again with a size of 0.9.
-A principal component analysis showed that, at a core set size of 0.7, 
-genotypes assigned to the A3 ancestral population (Fig.~\ref{fig:PopStructure})
-were largely elements of the core set whereas genotypes with a large share at 
-the A1 ancestral population were predominantly excluded from the core set.
-At a core set size of 90\% of the original size, the majority of the A1
-genotypes were finally covered by the core set as well.
-This result suggests a negative bias in the imputation of genotypes from the A1
-population and could explain the drop in predictive ability at a core set size
-of 0.7, which is in agreement with the reduced predictive abilities observed
-when stripping A1 genotypes from their transcriptomic information
-(Fig.~\ref{fig:InbredResults}B).
+By randomly and repeatedly declaring trancsriptomic information missing from
+genotypes in \textit{Exp2} we examined how reliably single step prediction works
+depending on the fraction of genotypes covered by both, the complete and the
+incomplete predictor.
+We observed that about $20--30\%$ of genotypes needed to be covered by both
+predictors to match the performance achieved with the complete predictor
+individually.
+Beyond this fraction of covered genotypes the trend in predictive ability
+changes was stable.
 
@@
 By declaring different subsets of individuals as covered by one or two 
-predictors, respectively, we could show the importance of homogeneous coverage
-of the genetic space by genotypes covered by both predictors on predictive
-ability.
+predictors, respectively, we could elucidate the influence of differences
+between two predictors on predictive ability when used in a single step
+approach.
+When the discrepancy in predictive ability between the complete and the
+incomplete predictor was skewed profoundly in favor of the incomplete predictor,
+such as for pedigree and transcriptomic data, single step prediction was shown
+be highly successful.
 Extensions to more than two predictors were outside the scope of this study

### 2017-09-07 paper/manuscript.tex
paper, inbreds as Exp1, shift order of Exp1,Exp2

@@
 \section{Materials and Methods}
-\Genetics2level{Maize hybrids}
-To address objectives 1 and 2, a maize hybrid data set, subsequently
-denoted as \textit{Exp1}, was used.
-The material comprised 1,521 hybrids produced in 16 factorial mating designs
-between 142 Dent and 103 Flint parent lines~\cite{Westhues2017}. 
-Best linear unbiased estimates (BLUEs) were computed across all factorials and 
-three or more agro-ecologically diverse locations across Germany for six 
-agronomically important traits by accounting for year, location, field 
-replication, block and genotype effects as well as their interactions
-(Table~\ref{table:TraitKey}).
-All 245 parent lines have pedigree records back to the generation of their 
-grandparents or deeper~\cite{Westhues2017} and were genotyped using the Illumina 
-BeadChip MaizeSNP50~\cite{Ganal2011}.
-After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
-($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed 
-using the Beagle software~\cite{Browning2009}.
-The final number of polymorphic marker loci was 7,013 for the Dent and 6,212 for 
-the Flint lines, respectively.
-Gene expression data for seedlings seven days after sowing were obtained for a 
-subset of 60 parental Dent and 43 parental Flint lines, that are parents of 685
-hybrid progeny.
-These data were obtained through two-color hybridizations using a custom 2K 
-microarray (GPL22267)~\cite{Westhues2017}.
-After applying established normalization procedures~\cite{Smyth2003,Ritchie2007},
-BLUEs for gene expression were computed for 1,323 transcripts while adjusting
-for putative effects arising from the presence of multiple factorials
-(\textit{cf.} \citeNP{Westhues2017}) using the R-package
-\textit{limma}~\cite{Ritchie2015a}.
-
-\Genetics2level{Maize diversity panel}
+\Genetics2level{Exp1}
 To address objectives 2 and 3, a maize inbred line data set, subsequently
-referred to as \textit{Exp2}, was used~\cite{Yang2014}.
-\textit{Exp2} originally comprised 513 maize lines representing the global maize
+referred to as \textit{Exp1}, was used~\cite{Yang2014}.
+\textit{Exp1} originally comprised 513 maize lines representing the global maize
 diversity  and was reduced to the set of tropical and subtropical lines
@@
 
+\Genetics2level{Exp2}
+To address objectives 1 and 2, a maize hybrid data set, subsequently
+denoted as \textit{Exp2}, was used.
+The material comprised 1,521 hybrids produced in 16 factorial mating designs
+between 142 Dent and 103 Flint parent lines~\cite{Westhues2017}. 
+Best linear unbiased estimates (BLUEs) were computed across all factorials and 
+three or more agro-ecologically diverse locations across Germany for six 
+agronomically important traits by accounting for year, location, field 
+replication, block and genotype effects as well as their interactions
+(Table~\ref{table:TraitKey}).
+All 245 parent lines have pedigree records back to the generation of their 
+grandparents or deeper~\cite{Westhues2017} and were genotyped using the Illumina 
+BeadChip MaizeSNP50~\cite{Ganal2011}.
+After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
+($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed 
+using the Beagle software~\cite{Browning2009}.
+The final number of polymorphic marker loci was 7,013 for the Dent and 6,212 for 
+the Flint lines, respectively.
+Gene expression data for seedlings seven days after sowing were obtained for a 
+subset of 60 parental Dent and 43 parental Flint lines, that are parents of 685
+hybrid progeny.
+These data were obtained through two-color hybridizations using a custom 2K 
+microarray (GPL22267)~\cite{Westhues2017}.
+After applying established normalization procedures~\cite{Smyth2003,Ritchie2007},
+BLUEs for gene expression were computed for 1,323 transcripts while adjusting
+for putative effects arising from the presence of multiple factorials
+(\textit{cf.} \citeNP{Westhues2017}) using the R-package
+\textit{limma}~\cite{Ritchie2015a}.
+
 % Please add the following required packages to your document preamble:
@@
 \Genetics2level{Population structure}
-Population structure in \textit{Exp2} was examined by running a STRUCTURE
+Population structure in \textit{Exp1} was examined by running a STRUCTURE
 \cite{Pritchard2000} analysis with the number of putative ancestral
@@
 features.
-Note that $l$ is only defined for \textit{Exp1} where a distinction between 
+Note that $l$ is only defined for \textit{Exp2} where a distinction between 
 the two heterotic groups is necessary whereas $l$ can be ignored for all 
 models pertaining to the inbred lines of the maize diversity panel in 
-\textit{Exp2}.
+\textit{Exp1}.
 To avoid mentioning special cases, we will use the general notation with 
@@
 residual error.
-For the hybrid data from \textit{Exp1}, we fit separate additive effects for 
+For the hybrid data from \textit{Exp2}, we fit separate additive effects for 
 each of the $L$ heterotic groups and $\mathbf{g}_{l}$ now pertains to the 
@@
 Flint, as expressed by $l$).
-For the maize diversity panel inbred lines from \textit{Exp2}, we fit random 
+For the maize diversity panel inbred lines from \textit{Exp1}, we fit random 
 genotype effects $\mathbf{g}_{l}$ (\textit{i.e.}, breeding values).
@@
 covered by the incomplete predictor on predictive ability we generated core
-sets of nine different sizes from \textit{Exp2}
+sets of nine different sizes from \textit{Exp1}
 (Fig.~\ref{fig:compound-plot}~a).
@@
   The two extended bars represent the entire set of genotypes in either
-  \textit{Exp1} or \textit{Exp2}.
+  \textit{Exp2} or \textit{Exp1}.
   All genotypes are covered by at least one predictor, which is called the
@@
   The four bars to the right outline the principle of generating core sets for
-  \textit{Exp2}.
+  \textit{Exp1}.
   For nine different values of $x$, which represents the fraction of genotypes
@@
 \section*{Results}
-\Genetics2level{Inbred lines}
+\Genetics2level{Exp1}
 At the level of the 110 genotypes for which both, genomic (G) and 
@@
 
-\Genetics2level{Hybrids}
+\Genetics2level{Exp2}
 The maize hybrid data set contains information on seven agronomic traits and
@@
   data, respectively.
-  Results are shown for the core set of 110 inbred lines from \textit{Exp2} for
+  Results are shown for the core set of 110 inbred lines from \textit{Exp1} for
   which both, G and T are available for any genotype and six agronomic traits.
@@
 \Genetics2level{Superiority of single step models}
-\subsection{Exp2}
+\subsection{Exp1}
 When using a single step approach based on complete genomic and incomplete
@@
 
-\subsection{Exp1}
+\subsection{Exp2}
 Hybrid breeding --- with noteworthy commercial applications in pig
@@
 By randomly and repeatedly declaring trancsriptomic information missing from
-genotypes in \textit{Exp2} we examined how reliably single step prediction works
+genotypes in \textit{Exp1} we examined how reliably single step prediction works
 depending on the fraction of genotypes covered by both, the complete and the

### 2017-09-12 paper/manuscript.tex
paper, change 'Core' to 'Reduced'

@@
 days after sowing.
-By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
+By exploiting identity by descent (IBD) between 49,728 SNPs on the BeadChip 
 overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
@@
 Gene expression data were normalized using a normal quantile transformation to
-satisfy modelling assumptions~\cite{Fu2013}.
+satisfy modeling assumptions~\cite{Fu2013}.
 A subset of 28,850 annotated genes was available for 149 out of the 211
@@
 Gene expression data for seedlings seven days after sowing were obtained for a 
-subset of 60 parental Dent and 43 parental Flint lines, that are parents of 685
-hybrid progeny.
+subset of 60 Dent and 43 Flint lines, that are parents of 685 hybrid progeny.
 These data were obtained through two-color hybridizations using a custom 2K 
@@
 and ran a principal component analysis (PCA) on the scaled and centered
-genomic feature matrix using the \emph{pca} function from the R-pacakge LEA
-\cite{Frichot2015}.
+genomic feature matrix using the \emph{pca} function from the \emph{R}-package
+\emph{LEA} \cite{Frichot2015}.
 Least-squares estimates of ancestry proportions were estimated based on sparse 
 nonnegative matrix factorization algorithms~\cite{Frichot2014} using the 
-\emph{snmf} function in \textit{LEA}.
+\emph{snmf} function in \emph{LEA}.
 Genotypes in the PCA plot (Fig.~\ref{fig:inbred-pca}) were colorized based on their
@@
 \textit{Exp1}.
-To avoid mentioning special cases, we will use the general notation with 
-the index $l$ throughout this section, though.
 For the diversity panel maize lines, $n_{l}$ corresponds to the number of 
@@
 Following the notation in \citeNP{Fernando2014} the phenotypes can be untangled 
-like this:
+as:
 
@@
 In the case of the diversity panel maize inbred lines, LOOCV was performed by 
-using a single genotype as a hold-out sample, which will subsequently be 
-predicted, while using all other inbred lines for model training.
+using a single genotype as a hold-out sample, which will be predicted by using
+all other inbred lines for model training.
 This process is repeated until all inbred lines have been used once for testing
@@
   combination (GT) were used.
-  The `Core' panel includes a subset of 110 lines covered by genomic and
+  The `Reduced' panel includes a subset of 110 lines covered by genomic and
   transcriptomic information whereas the `Full' panel comprises all 164
@@
 transcriptomic data were available for 103 parent lines of 685 hybrids, which
-we denote as `Core'.
+we denote as `Reduced'.
 In the `core' set, transcriptomic data were the best single predictor for ADL,
@@
     The `Full' set of genotypes comprises 1,521 hybrid progeny based on 245
-    parental inbred lines, whereas the `Core' set comprises 685 hybrid progeny
-    based on 103 parental inbred lines.
+    parental inbred lines, whereas the `Reduced' set comprises 685 hybrid
+    progeny based on 103 parental inbred lines.
   }

### 2017-09-12 analysis/visualize_predictions.R

@@ -72,6 +72,10 @@ limits <- aes(
 
 hybrid_plot <- hyb_df %>%
   dplyr::rename(`Predictive Ability` = "avg_r") %>%
+  dplyr::mutate(Combi = forcats::fct_recode(
+    Combi,
+    "Reduced" = "Core"
+  )) %>%
   ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
   geom_bar(stat = "identity", position = dodge, color = "black") +
   geom_errorbar(limits, position = dodge, width = 0.25) +

### 2017-09-12 analysis/visualize_predictions.R
paper, 'core' to 'reduced' for inbreds

@@ -138,6 +138,11 @@ a_colors <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6) %>%
 inbred_plot <- inbred_df %>%
   dplyr::filter(Core_Fraction == "1") %>%
   dplyr::rename(`Predictive Ability` = "avg_r") %>%
+  dplyr::mutate(Combi = forcats::fct_recode(
+    Combi,
+    "Reduced" = "CIA",
+    "Full" = "FIA"
+  )) %>%
   ggplot(aes(x = Combi, y = `Predictive Ability`, fill = Predictor)) +
   geom_bar(stat = "identity", position = dodge) +
   geom_errorbar(limits, position = dodge, width = 0.25) +

### 2017-09-12 paper/manuscript.tex
paper, cosmetic changes

@@
 Best linear unbiased estimates (BLUEs) were computed across all factorials and 
-three or more agro-ecologically diverse locations across Germany for six 
+three or more agro-ecologically diverse locations across Germany for seven
 agronomically important traits by accounting for year, location, field 
@@
 Dry matter content & DMC          & Hybrid    \\
+Lignin             & ADL          & Hybrid    \\
 Fat                & FAT          & Hybrid    \\
@@
   `tropical/subtropical' of the maize diversity panel inbred lines.
-  The four clusters are based on $K=4$ putative ancestral populations in a
+  The four clusters are based on $K=4$ putative ancestral populations from a
   STRUCTURE analysis of the corresponding genomic feature matrix.
@@
   `complete' predictor and depicted as a white bar.
-  The second bar depicts that only a fraction of genotypes is covered by both,
+  The second bar shows that only a fraction of genotypes is covered by both,
   the `complete' and the `incomplete' predictor (dark gray) and that the
-  difference between the predictors will be imputed.
+  difference between the predictors will be imputed (light gray fraction).
   The four bars to the right outline the principle of generating core sets for
@@
   `incomplete' predictor whereas the remaining $100\% - x\%$ of genotypes
-  declared to be covered only by the `complete' predictor.
+  are declared to be covered only by the `complete' predictor.
   }
@@
   \caption{
-  Predictive abilities and SE for tropical/subtropical inbred lines from the 
-  maize diversity panel and six agronomic traits.
+  Predictive abilities and average, bootstrapped standard errors for
+  tropical/subtropical inbred lines from the maize diversity panel and six
+  agronomic traits.
   As predictors, genomic (G) data, transcriptomic (T) data and their 
@@
 we denote as `Reduced'.
-In the `core' set, transcriptomic data were the best single predictor for ADL,
-PRO and particulaly for DMY where the predictive ability obtained with T
+In the `Reduced' set, transcriptomic data were the best single predictor for
+ADL, PRO and particulaly for DMY where the predictive ability obtained with T
 was 0.08 points higher than for G as the second best predictor for this trait
 (Fig.~\ref{fig:HybridResults}).
-In the core set, P had the highest predictive ability for STA and DMC\@.
-For the FAT and SUG, G was the superior predictor inside the core set.
+In the reduced set, P had the highest predictive ability for STA and DMC\@.
+For FAT and SUG, G was the superior predictor inside the reduced set.
 In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
@@
 Except for PRO, compared to using only G, trait prediction did not benefit from 
-the single step approach when T was not superior to G in the core set.
+the single step approach when T was not superior to G in the reduced set.
 Predictive abilities resulting from combinations of pedigree with genomic (PG)
@@
     Predictive abilities and average, bootstrapped standard errors for the set
-    of maize hybrids and six agronomic traits.
+    of maize hybrids and seven agronomic traits.
     As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
@@
 For most traits, an initial increase in predictive abilities when moving
-from a core set size of 0.1 to 0.2/0.3 was observed, regardless of whether G
+from a core set size of 10\% to 20\%/30\% was observed, regardless of whether G
 or T individually was the superior predictor.
@@
 Another possibility is that the number of additional phenotypes in the full
-set compared to the core set ($\Delta(n) = 54$) was not large enough to provide
-considerably more useful information.
+set compared to the reduced set ($\Delta(n) = 54$) was not large enough to
+provide considerably more useful information.
 The latter was the case for \citeNP{Ashraf2016}, who imputed genotypes for about 
@@
 predicted candidates is a strongly convex function of the predictive ability.
-Thus, even the minor gains when using transcriptomic and genomic data in a
-single step might justify additional investments in RNA-seq at least for a
+Thus, even minor gains, when using transcriptomic and genomic data in a
+single step, might justify additional investments in RNA-seq at least for a
 subset of genotypes that represents the genetic space of the breeding 
@@
 Hence, the inclusion of trancsriptomic data could not further improve upon
-this result.
+genomic information alone.
 Predictive abilities obtained when using pedigree data could be 
@@
 either transcriptomic or genomic information.
+This suggests that single step prediction should always be considered when the
+incomplete predictor offers an appreciably higher predictive ability compared
+to the complete predictor.
 
@@
 predictor on only a subset of genotypes, thereby reducing predictor costs.
-Hence, we were interested in determining which individuals should be 
+Hence, we were interested in determining to what extent individuals should be 
 complemented with information on another predictor.
@@
 incomplete predictor.
-We observed that about $20--30\%$ of genotypes needed to be covered by both
+We observed that about $20-30\%$ of genotypes needed to be covered by both
 predictors to match the performance achieved with the complete predictor
@@
 \section*{Conclusions}
-We have successfully applied single step prediction to a hybrid plant data set 
+We successfully applied single step prediction to a hybrid plant data set 
 while imputing with a quantitative downstream `omics' predictor.
@@
 such as for pedigree and transcriptomic data, single step prediction was shown
-be highly successful.
+to be highly beneficial.
 Extensions to more than two predictors were outside the scope of this study

### 2017-09-13 paper/manuscript.tex
paper, add claas' remarks 2017-09-12

@@
 Consider now the situation in which one predictor is complete in the sense that
-it contains information on the whole population whereas a second predictor is
-incomplete and covers only a subset of the population.
+the whole population is equipped with features for that predictor whereas only
+a subset of that population has information on a second predictor.
 We can take a similar route as in \citeNP{Fernando2014} and impute covariates 
@@
 values of animals covered by both predictors).
-The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
-expectation given $\mathbf{g}_2$ and a residual error term:
+Under multivariate normality, the vector $\mathbf{g}_1$ can be written as the
+sum of the conditional expectation given $\mathbf{g}_2$ and a residual error
+term:
 \begin{align} \label{eq:mrna1}
@@
 therefore be modeled.
-The covariates not covered by the incomplete predictor can be predicted using 
-the expectation of a multivariate normal random vector given correlated 
-observations, which are related to the Best Linear Predictor.
 
@@
 
-Breeding values can be estimated as:
+Breeding values or general combining abilities can be estimated as:
 

### 2017-09-14 paper/manuscript.tex
paper, hybrid and quant pred rationale in intro

@@
 animal and plant breeding by a more efficient selection of promising candidates 
-without phenotypes~\cite{DeLosCampos2013,Garcia-Ruiz2016} and recently emerged
-as an important tool in human diagnostics~\cite{DeLosCampos2010,Vazquez2016}.
+without phenotypes~\cite{DeLosCampos2013,Garcia-Ruiz2016}.
 Its use is largely motivated by the high costs and oftentimes long duration of 
-phenotyping, which can be reduced when covering all genotypes with a predictor 
+phenotyping, which can be reduced when all genotypes are equipped with a predictor 
 while limiting phenotyping to a few genotypes with high precision
@@
 The performance of all other genotypes is then forecast using model parameters
-obtained from thorough resampling routines applied to the subset of individuals 
-with full information~\cite{Hastie2009}.
+estimated in the subset of individuals with full information.
 In animal breeding, genomic selection has increased the annual selection gain 
@@
 \citeNP{Christensen2010}, which does not require weights for $\mathbf{A}$ and 
-$\mathbf{G}$ and allows implicit modeling of the imputation error's structure.
+$\mathbf{G}$ and allows implicit modeling of the imputation error structure.
 
-% -   Mention scarcity of data on single step prediction in plant populations.
+% -   Objectives (Hybrids)
 In animal breeding, single step prediction, particularly H-BLUP, has already 
@@
 white spruce lines, respectively~\cite{Ratcliffe2017}.
-Our objectives were threefold: (1) Examine whether the single step prediction
-framework can be transfered to plant hybrids, (2) explore the utility of single 
-step predition with a quantitative predictor and (3) evaluate the impact of the
-proporition of genotypes not covered by the incomplete predictor on predictive
-ability.
+While single step prediction has been applied to pig hybrids~\cite{Xiang2016}
+this application is not equivalent to hybrid breeding in plants.
+There, the maternal and the paternal crossing partner can be selected without
+constraints whereas animal breeders select only males because of the limited
+fertility of the female animals.
+Technological advances in plant hybrid breeding such as the rapid generation
+of completely homozygous lines~\cite{Wedzony2009} now enable breeders to
+produce about $n = 1,000$ individuals per generation and sex.
+Assuming that about 90\% of these lines were never before tested in any hybrid
+combination about $n^{2} * 0.9^{2} = 810,000$ putative hybrids per generation
+would not have a single parent with phenotypically tested progeny
+\cite{Westhues2017}.
+Predicting these hybrid candidates with low relatedness to tested hybrids using
+only information from their parents is currently a hot topic in plant breeding
+\cite{Technow2014,Kadam2016,Westhues2017}.
+Hence, our first objective was to examine whether the single step prediction 
+framework can be transfered to hybrids in plants.
+
+% -   Objectives (quantitative predictores)
+Given that the availability of pedigree and genomic data is the most prevalent
+situation in animal breeding, previous studies have focused exclusively on the
+application of single step prediction to these predictors.
+Recent investigations into the utility of `omics' data for boosting predictive
+abilities in plants have suggested that metabolites and transcriptomic data can
+improve upon predictive abilities based on pedigree or genomic data
+\cite{Guo2016,Xu2016,Dan2016,Zenke-Philippi2017,Westhues2017}.
+This situation motivated our second objective, namely exploring the utility of
+single step predition with a quantitative predictor.
+Yet, these technologies are not fully mature and therefore incur higher costs
+to breeders than pedigree and genomic data, putting their utility into question
+despite their proven usefulness.
+Our third objective is therefore to evaluate the impact of equipping only a
+subset of all genotypes with both, the complete and the incomplete predictor on
+predictive ability.
 

### 2017-09-14 paper/manuscript.tex
paper-intro, update Exp1 and Exp2

@@
 To address objectives 2 and 3, a maize inbred line data set, subsequently
-referred to as \textit{Exp1}, was used~\cite{Yang2014}.
-\textit{Exp1} originally comprised 513 maize lines representing the global maize
+referred to as Exp1, was used~\cite{Yang2014}.
+Exp1 originally comprised 513 maize lines representing the global maize
 diversity  and was reduced to the set of tropical and subtropical lines
@@
 subgroups~\cite{Yang2014}.
-All inbred lines were evaluated in five Chinese environments ranging from
-$18$ to $30^{\circ}$N and from $102$ to $110^{\circ}$E~\cite{Yang2014}.
+All inbred lines were evaluated in five Chinese environments described in detail
+by~\citeNP{Yang2014}.
 Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
@@
 by genomic data.
-MaizeSNP50 BeadChip-data were available for the entire set of lines from the
+MaizeSNP50 BeadChip-data were available for the entire set of 513 lines from the
 diversity panel.
@@
 set of markers~\cite{Yang2014}.
-The same SNP quality checks as for the parental inbred lines of the hybrids
-were applied to the 211 inbred lines from the tropical/subtropical subset,
-yielding 37,760 SNPs.
+After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
+($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed 
+for 37,760 SNPs using the Beagle software~\cite{Browning2009}.
 Gene expression data were normalized using a normal quantile transformation to
@@
 To address objectives 1 and 2, a maize hybrid data set, subsequently
-denoted as \textit{Exp2}, was used.
+denoted as Exp2, was used.
 The material comprised 1,521 hybrids produced in 16 factorial mating designs
-between 142 Dent and 103 Flint parent lines~\cite{Westhues2017}. 
+between 142 Dent and 103 Flint parent lines (described in details by
+\citeNP{Westhues2017}).
 Best linear unbiased estimates (BLUEs) were computed across all factorials and 
-three or more agro-ecologically diverse locations across Germany for seven
-agronomically important traits by accounting for year, location, field 
-replication, block and genotype effects as well as their interactions
+three or more agro-ecologically diverse environments across Germany for seven
+agronomically important traits in silage maize by accounting for year, location,
+field replication, block and genotype effects as well as their interactions
 (Table~\ref{table:TraitKey}).
@@
 BeadChip MaizeSNP50~\cite{Ganal2011}.
-After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
-($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed 
-using the Beagle software~\cite{Browning2009}.
-The final number of polymorphic marker loci was 7,013 for the Dent and 6,212 for 
-the Flint lines, respectively.
+The same SNP quality checks as for the material in Exp1 were applied here,
+yielding 7,013 polymorphic marker loci for the Dent and 6,212 for the Flint
+lines, respectively.
 Gene expression data for seedlings seven days after sowing were obtained for a 
@@
 After applying established normalization procedures~\cite{Smyth2003,Ritchie2007},
-BLUEs for gene expression were computed for 1,323 transcripts while adjusting
-for putative effects arising from the presence of multiple factorials
-(\textit{cf.} \citeNP{Westhues2017}) using the R-package
-\textit{limma}~\cite{Ritchie2015a}.
+BLUEs for gene expression were computed for 1,323 transcripts
+\cite{Westhues2017} using the R-package \emph{limma}~\cite{Ritchie2015a}.
 
@@
 \Genetics2level{Population structure}
-Population structure in \textit{Exp1} was examined by running a STRUCTURE
+Population structure in Exp1 was examined by running a STRUCTURE
 \cite{Pritchard2000} analysis with the number of putative ancestral
@@
 features.
-Note that $l$ is only defined for \textit{Exp2} where a distinction between 
+Note that $l$ is only defined for Exp2 where a distinction between 
 the two heterotic groups is necessary whereas $l$ can be ignored for all 
 models pertaining to the inbred lines of the maize diversity panel in 
-\textit{Exp1}.
+Exp1.
 For the diversity panel maize lines, $n_{l}$ corresponds to the number of 
@@
 residual error.
-For the hybrid data from \textit{Exp2}, we fit separate additive effects for 
+For the hybrid data from Exp2, we fit separate additive effects for 
 each of the $L$ heterotic groups and $\mathbf{g}_{l}$ now pertains to the 
@@
 Flint, as expressed by $l$).
-For the maize diversity panel inbred lines from \textit{Exp1}, we fit random 
+For the maize diversity panel inbred lines from Exp1, we fit random 
 genotype effects $\mathbf{g}_{l}$ (\textit{i.e.}, breeding values).
@@
 covered by the incomplete predictor on predictive ability we generated core
-sets of nine different sizes from \textit{Exp1}
-(Fig.~\ref{fig:compound-plot}~a).
+sets of nine different sizes from Exp1 (Fig.~\ref{fig:compound-plot}~a).
 The base of each of these core sets was the superset of genotypes that were
@@
   The two extended bars represent the entire set of genotypes in either
-  \textit{Exp2} or \textit{Exp1}.
+  Exp2 or Exp1.
   All genotypes are covered by at least one predictor, which is called the
@@
   The four bars to the right outline the principle of generating core sets for
-  \textit{Exp1}.
+  Exp1.
   For nine different values of $x$, which represents the fraction of genotypes
@@
   data, respectively.
-  Results are shown for the core set of 110 inbred lines from \textit{Exp1} for
+  Results are shown for the core set of 110 inbred lines from Exp1 for
   which both, G and T are available for any genotype and six agronomic traits.
@@
 By randomly and repeatedly declaring trancsriptomic information missing from
-genotypes in \textit{Exp1} we examined how reliably single step prediction works
+genotypes in Exp1 we examined how reliably single step prediction works
 depending on the fraction of genotypes covered by both, the complete and the

### 2017-09-14 paper/manuscript.tex
paper, add experiment id to trait table

@@
 \toprule
-Trait              & Abbreviation & Data Set  \\ \midrule
-Cob weight         & CW           & Diversity \\
-Days to silking    & DS           & Diversity \\
-Ear diameter       & ED           & Diversity \\
-100-grain weight   & GW           & Diversity \\
-Kernel width       & KW           & Diversity \\
-Plant height       & PH           & Diversity \\
-Dry matter yield   & DMY          & Hybrid    \\
-Dry matter content & DMC          & Hybrid    \\
-Lignin             & ADL          & Hybrid    \\
-Fat                & FAT          & Hybrid    \\
-Protein            & PRO          & Hybrid    \\
-Starch             & STA          & Hybrid    \\
-Sugar              & SUG          & Hybrid    \\ \bottomrule
+Trait              & Abbreviation & Data Set  & Experiment\\ \midrule
+Cob weight         & CW           & Diversity & Exp1 \\
+Days to silking    & DS           & Diversity & Exp1 \\
+Ear diameter       & ED           & Diversity & Exp1 \\
+100-grain weight   & GW           & Diversity & Exp1 \\
+Kernel width       & KW           & Diversity & Exp1 \\
+Plant height       & PH           & Diversity & Exp1 \\
+Dry matter yield   & DMY          & Hybrid    & Exp2 \\
+Dry matter content & DMC          & Hybrid    & Exp2 \\
+Lignin             & ADL          & Hybrid    & Exp2 \\
+Fat                & FAT          & Hybrid    & Exp2 \\
+Protein            & PRO          & Hybrid    & Exp2 \\
+Starch             & STA          & Hybrid    & Exp2 \\
+Sugar              & SUG          & Hybrid    & Exp2 \\ \bottomrule
 \end{tabular}

### 2017-09-14 paper/manuscript.tex
paper-intro, update population structure text

@@
 \Genetics2level{Population structure}
-Population structure in Exp1 was examined by running a STRUCTURE
+Population structure in Exp1 was examined through a STRUCTURE
 \cite{Pritchard2000} analysis with the number of putative ancestral
@@
 genomic feature matrix using the \emph{pca} function from the \emph{R}-package
-\emph{LEA} \cite{Frichot2015}.
+\emph{LEA}~\cite{Frichot2015}.
 Least-squares estimates of ancestry proportions were estimated based on sparse 
@@
 primary assignment to any of the four putative ancestral populations.
-To avoid inflated predictive abilities, which can arise from large differences
-in the average performance of the different subpopulations
-\cite{Windhausen2012}, we reduced the panel of inbred lines to those primarily
-belonging to either cluster 1 or cluster 4 in Fig.~\ref{fig:inbred-pca}.
+To mitigate the possibility of inflated predictive abilities, which can arise
+from large differences in the average performance of the different
+subpopulations~\cite{Windhausen2012}, we reduced the panel of inbred lines to
+those primarily belonging to either cluster 1 or cluster 4 in 
+Fig.~\ref{fig:inbred-pca}.
 This reduction yielded 164 genotypes, 110 of which were covered by both, genomic

### 2017-09-14 paper/manuscript.tex
paper, clarify predictions

@@
 \label{table:TraitKey}
-\begin{tabular}{@{}lll@{}}
+\begin{tabular}{@{}llll@{}}
 \toprule
-Trait              & Abbreviation & Data Set  & Experiment\\ \midrule
+Trait              & Abbreviation & Data Set  & Experiment \\ \midrule
 Cob weight         & CW           & Diversity & Exp1       \\
@@
 transcriptompic data (T).
-The corresponding feature matrix of the $l$-th group of inbred lines 
-($\mathbf{W}_{l}$) has dimensions $n_{l} \times p$ where $n_{l}$ pertains to 
-the number of genotypes in the $l$-th group and $p$ pertains to the number of 
-features.
-Note that $l$ is only defined for Exp2 where a distinction between 
-the two heterotic groups is necessary whereas $l$ can be ignored for all 
-models pertaining to the inbred lines of the maize diversity panel in 
-Exp1.
-For the diversity panel maize lines, $n_{l}$ corresponds to the number of 
-genotypes whereas, in the case of the hybrid data, $n_{l}$ corresponds to the 
-number of parent lines in the $l$-th heterotic group.
-In generating kernels for each predictor, all features in $\mathbf{W}_{l}$ were
+The corresponding feature matrix of the inbred lines ($\mathbf{M}$) has
+dimensions $n \times p$ where $n$ pertains to the number of genotypes.
+Note that, for the sake of clarity, we do not distinguish between models for
+Exp1 and Exp2 at this point.
+In the case of Exp2 the term `genotype' refers strictly to the parent lines.
+In generating kernels for each predictor, all features in $\mathbf{M}$ were
 centered and standardized to unit variance.
@@
 \begin{equation} \label{eq:GenomicRelationship}
-  \mathbf{K}_{l} = \frac{1}{W_{l}} \mathbf{W}_{l} \mathbf{W}_{l}^{\top},
+  \mathbf{K} = \frac{1}{M} \mathbf{M} \mathbf{M}^{\top},
 \end{equation}
 
-where $W_{l}$ denotes the number of features~\cite{VanRaden2008}.
+where $M$ denotes the number of features~\cite{VanRaden2008}.
 In the case of pedigree data (P), coancestry coefficients were used directly
-for $\mathbf{K}_{l}$.
+for $\mathbf{K}$.
 
-\subsection{Breeding value model}
-The general model for breeding values was:
+\subsection{Genetic model}
+The general model for genetic values was:
 
 \begin{equation} \label{eq:KBLUPModel}
-  \mathbf{y} = \mu + 
-  \sum_{l=1}^{L} \mathbf{Z}_{l} \mathbf{g}_{l} + \mathbf{e},
+  \mathbf{y} = \mu + \mathbf{Z} \mathbf{g} + \mathbf{e},
 \end{equation}
@@
 residual error.
-For the hybrid data from Exp2, we fit separate additive effects for 
-each of the $L$ heterotic groups and $\mathbf{g}_{l}$ now pertains to the 
-general combining ability effects of the parental inbred lines (either Dent or 
-Flint, as expressed by $l$).
-For the maize diversity panel inbred lines from Exp1, we fit random 
-genotype effects $\mathbf{g}_{l}$ (\textit{i.e.}, breeding values).
-For both experiments, genetic effects $\mathbf{g}_{l}$ are associated with 
-the vector of phenotypic performance $\mathbf{y}$ through a corresponding 
-design matrix $\mathbf{Z}_{l}$.
-The random effects ($\mathbf{g}_{l}$) have expectation zero and variance equal 
-to $\mathbf{K}_{l} \sigma^{2}_{{g}^{l}}$ and $\mathbf{I} \sigma^2_{e}$ 
-for the residual error.
+For the inbred lines we fit random genotype effects $\mathbf{g}$.
+For both experiments, genetic effects $\mathbf{g}$ are associated with the
+vector of phenotypic performance $\mathbf{y}$ through a corresponding design
+matrix $\mathbf{Z}$.
+The random effects ($\mathbf{g}$) have expectation zero and variance equal to
+$\mathbf{K} \sigma^{2}_{g}$ and $\mathbf{I} \sigma^2_{e}$ for the residual
+error.
 Here, we did not consider the inclusion of dominance effects for the hybrid
 material after a previous investigation into the predictive ability of various
-``omics'' predictors in the same material concluded no considerably gains 
+``omics'' predictors in the same material concluded no considerable gains 
 \cite{Westhues2017}.
@@
 \subsection{Single step prediction}
-For simplicity, let us assume throughout this section that we have $L=1$.
-If the breeding values are fully captured by the features in $\mathbf{W}$,
+If the genetic values are fully captured by the features in $\mathbf{M}$,
 Eq.~\ref{eq:KBLUPModel} can then be expressed as
@@
 	\mathbf{y} &= \mu + \mathbf{Z} \mathbf{g} + \boldsymbol{e} \\
-	&= \mu + \mathbf{Z} \mathbf{W} \boldsymbol{\alpha} + \boldsymbol{e}
+	&= \mu + \mathbf{Z} \mathbf{M} \boldsymbol{\alpha} + \boldsymbol{e},
 \end{align}
-\cite{Fernando2014} since the ridge regression model is equivalent to the BLUP
-model~\cite{Ruppert2003}.
+
+\cite{Fernando2014} because the ridge regression model is equivalent to the
+BLUP model~\cite{Ruppert2003}.
+In Eq.~\ref{eq:KBLUPModelSimplified}, $\mathbf{M}$ is a matrix of feature
+covariates and $\boldsymbol{\alpha}$ is the vector of random, partial-regression
+coefficients of the feature covariates.
 
@@
 \begin{equation} \label{eq:mrnaebv}
-	\mathbf{\hat{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
+	\mathbf{\hat{g}} = \mathbf{M}\boldsymbol{\hat{\alpha}},
 \end{equation}
 
-where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model.
+where $\boldsymbol{\hat{\alpha}}$ is the solution to the ridge regression model
+(Eq.~\ref{eq:KBLUPModel}).
 
@@
 incomplete predictor, are indicated by the subscript $2$.
-Further, the covariates in $\mathbf{W}$ are centered.
-In this case we need to partition the vector of breeding values $\mathbf{g}$
-into a component $\mathbf{g}_{1}$ (\textit{i.e.,} breeding values of animals 
+Further, we center the covariates of the incomplete predictor and denote it as
+$\mathbf{W}$.
+Correspondingly, $\mathbf{K}$ corresponds to the kernel for the complete
+predictor.
+In this case we need to partition the vector of genetic values $\mathbf{g}$
+into a component $\mathbf{g}_{1}$ (\textit{i.e.,} genetic values of genotypes
 covered only by the complete predictor) and a component
-$\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} breeding 
-values of animals covered by both predictors).
+$\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} genetic
+values of genotypes covered by both predictors, where $\boldsymbol{\alpha}$
+pertains to the random feature effects of the incomplete predictor).
 Under multivariate normality, the vector $\mathbf{g}_1$ can be written as the
@@
 	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
-	&= \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K_{12}}\mathbf{K_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
+	&= \mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{W}_{2}
+	\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K}_{12}
+	\mathbf{K}^{-1}_{2}
+	\boldsymbol{\hat{\alpha}}) \\
 	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
@@
 where $\mathbf{K}_{ij}$ are partitions of $\mathbf{K}$ that correspond to 
-$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014}.
+$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and
+where $\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values in the
+incomplete predictor.
 
-The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{K}_{11} - \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{K}_{21} = (\mathbf{K}^{11})^{-1}$
+The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{K}_{11} -
+\mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{K}_{21} =
+(\mathbf{K}^{11})^{-1}$
 \cite{Legarra2009}.
@@
 \begin{bmatrix}
-  X_1 \\
-  X_2 
+  \mathbf{X}_1 \\
+  \mathbf{X}_2
  \end{bmatrix},
@@
 \begin{bmatrix}
-  Z_1\hat{\mathbf{W}_1} \\
+  \mathbf{Z}_1\mathbf{\hat{W}}_{1} \\
   \mathbf{W}_2
@@
 \begin{bmatrix}
-  Z_1 \\
+  \mathbf{Z}_1 \\
   0
@@
 
-where $\hat{\mathbf{W}}_{1} = \mathbf{K}_{12} \mathbf{K}_{22}^{-1} \mathbf{W}_{2}$.
+where $\mathbf{\hat{W}}_{1} = \mathbf{K}_{12} 
+\mathbf{K}^{-1}_{22} \mathbf{W}_{2}$.
 
@@
 \begin{bmatrix}
-  y_1 \\
-  y_2 
+  \mathbf{y}_1 \\
+  \mathbf{y}_2
  \end{bmatrix}
@@
  \begin{bmatrix}
-  X_1 \\
-  X_2 
+  \mathbf{X}_1 \\
+  \mathbf{X}_2
  \end{bmatrix}
@@
  \begin{bmatrix}
-  Z_1 & 0 \\
-  0 & Z_2 
+  \mathbf{Z}_1 & \boldsymbol{0} \\
+  \boldsymbol{0} & \mathbf{Z}_2
  \end{bmatrix}
 \begin{bmatrix}
-  g_1 \\
-  g_2 
+  \mathbf{g}_1 \\
+  \mathbf{g}_2
  \end{bmatrix}
@@
  \begin{bmatrix}
-  X_1 \\
-  X_2 
+ \mathbf{X}_1 \\
+ \mathbf{X}_2
  \end{bmatrix}
@@
  \begin{bmatrix}
-  Z_1 & 0 \\
-  0 & Z_2 
+  \mathbf{Z}_1 & \boldsymbol{0} \\
+  \boldsymbol{0} & \mathbf{Z}_2
  \end{bmatrix}
 \begin{bmatrix}
-  \mathbf{K}_{12}\mathbf{K}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}  \\
+  \mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{W}_2
+  \boldsymbol{\alpha} + \boldsymbol{\epsilon}  \\
   \mathbf{W}_2\boldsymbol{\alpha} \\
@@
  \begin{bmatrix}
-  Z_1 & 0 \\
-  0 & Z_2 
+  \mathbf{Z}_1 & \boldsymbol{0} \\
+  \boldsymbol{0} & \mathbf{Z}_2
  \end{bmatrix}
 \begin{bmatrix}
-  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon} \\
+  \mathbf{\hat{W}}_{1}\boldsymbol{\alpha} + \boldsymbol{\epsilon} \\
   \mathbf{W}_2\boldsymbol{\alpha} \\
@@
 \mathbf{y} &= \mathbf{Xb} +
-\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
+\mathbf{Z}\mathbf{W} \boldsymbol{\alpha}_{l} + 
 \sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l} \boldsymbol{\epsilon}_{l} +

### 2017-09-14 paper/manuscript.tex
paper, explain partitions of W matrix

@@
 dimensions $n \times p$ where $n$ pertains to the number of genotypes.
-Note that, for the sake of clarity, we do not distinguish between models for
-Exp1 and Exp2 at this point.
-In the case of Exp2 the term `genotype' refers strictly to the parent lines.
-In generating kernels for each predictor, all features in $\mathbf{M}$ were
-centered and standardized to unit variance.
-Then, each kernel can be defined as
+For the sake of clarity, we do not distinguish between models for Exp1 and Exp2
+at this point.
+It should be noted, however, that in the case of Exp2 the term `genotype' refers
+strictly to the parent lines.
+For the complete predictor we generated kernels by centering and standardizing
+all features in $\mathbf{M}$ to unit variance:
 
@@
 If the genetic values are fully captured by the features in $\mathbf{M}$,
-Eq.~\ref{eq:KBLUPModel} can then be expressed as
+Eq.~\ref{eq:KBLUPModel} can be expressed as
 
@@
 BLUP model~\cite{Ruppert2003}.
-In Eq.~\ref{eq:KBLUPModelSimplified}, $\mathbf{M}$ is a matrix of feature
-covariates and $\boldsymbol{\alpha}$ is the vector of random, partial-regression
-coefficients of the feature covariates.
+In the previous equation, $\mathbf{M}$ is a matrix of feature covariates and
+$\boldsymbol{\alpha}$ is the vector of random, partial-regression coefficients
+of the feature covariates.
 
@@
 
-where $\boldsymbol{\hat{\alpha}}$ is the solution to the ridge regression model
-(Eq.~\ref{eq:KBLUPModel}).
+where $\boldsymbol{\hat{\alpha}}$ is the solution to the ridge regression model.
 
@@
 incomplete predictor, are indicated by the subscript $2$.
-Further, we center the covariates of the incomplete predictor and denote it as
-$\mathbf{W}$.
-Correspondingly, $\mathbf{K}$ corresponds to the kernel for the complete
-predictor.
+Let us further denote the matrix of centered covariates of the incomplete
+predictor as
+\begin{equation} \label{eq:incomplete-covariates}
+\mathbf{W} =
+\begin{bmatrix}
+  \mathbf{\hat{W}}_{1} \\
+  \mathbf{W}_{2}
+ \end{bmatrix},
+\end{equation}
+where observed feature covariates are stored in $\mathbf{W}_{2}$ while marker
+covariates for genotypes not covered by the incomplete predictor
+($\mathbf{\hat{W}}_{1}$) need to be imputed.
 In this case we need to partition the vector of genetic values $\mathbf{g}$
@@
 $\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} genetic
-values of genotypes covered by both predictors, where $\boldsymbol{\alpha}$
+values of genotypes covered by both predictors where $\boldsymbol{\alpha}$
 pertains to the random feature effects of the incomplete predictor).
@@
 $\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and
-where $\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values in the
-incomplete predictor.
+where $\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values of genotypes
+not equipped with the incomplete predictor.
 

### 2017-09-14 paper/manuscript.tex
paper-prediction, extend equations to hybrids

@@
 trait predictions, namely pedigree data (P), genomic data (G) and
-transcriptompic data (T).
+transcriptomic data (T).
 The corresponding feature matrix of the inbred lines ($\mathbf{M}$) has
@@
 \begin{bmatrix}
-  \mathbf{\hat{W}}_{1} \\
+  \mathbf{Z}_{1}\mathbf{\hat{W}}_{1} \\
   \mathbf{W}_{2}
@@
 covariates for genotypes not covered by the incomplete predictor
-($\mathbf{\hat{W}}_{1}$) need to be imputed.
+($\mathbf{\hat{W}}_{1} = \mathbf{K}_{12} \mathbf{K}^{-1}_{22} \mathbf{W}_{2}$)
+need to be imputed.
 In this case we need to partition the vector of genetic values $\mathbf{g}$
@@
 \begin{equation} \label{eq:single-step-model}
-\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U} \boldsymbol{\epsilon} + \mathbf{e},
+\mathbf{y} =
+\mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U} \boldsymbol{\epsilon} + \mathbf{e},
 \end{equation}
@@
  \end{bmatrix},
- \mathbf{W} =
-\begin{bmatrix}
-  \mathbf{Z}_1\mathbf{\hat{W}}_{1} \\
-  \mathbf{W}_2
- \end{bmatrix},
  \mathbf{U} =
@@
   0
- \end{bmatrix},
+ \end{bmatrix}.
 \end{equation}
-
-where $\mathbf{\hat{W}}_{1} = \mathbf{K}_{12} 
-\mathbf{K}^{-1}_{22} \mathbf{W}_{2}$.
 
@@
 
-Our final model looks like this:
+The above equation implies that ultimately single step prediction uses only
+one random effect vector $\mathbf{g}$, which entails the observed effects of
+the incomplete predictor $\mathbf{g}_{2}$ and the imputed effects
+$\mathbf{\hat{g}}_{1}$ for genotypes that are not equipped with the incomplete
+predictor plus the residual imputation error $\boldsymbol{\epsilon}$.
+
+Our final model is now:
 
@@
 \mathbf{y} &= \mathbf{Xb} +
-\mathbf{Z}\mathbf{W} \boldsymbol{\alpha}_{l} + 
-\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l} \boldsymbol{\epsilon}_{l} +
+\mathbf{Z}\mathbf{W} \boldsymbol{\alpha} +
+\mathbf{Z}\mathbf{U} \boldsymbol{\epsilon} +
 \mathbf{e}.
@@
 
-Breeding values or general combining abilities can be estimated as:
+Extending this model to hybrids with parent lines from two heterotic groups
+simply requires us to separate the feature matrices of the two predictors
+into a paternal and a maternal component, respectively.
+Using the index $D$ for the Dent and the index $F$ for the Flint lines in Exp2
+we then have a feature matrix $\mathbf{M}_{D}$ and $\mathbf{M}_{F}$ for each
+predictor.
+For the hybrids the final model is therefore simply
+\begin{align} \label{eq:final-model}
+\mathbf{y} &= \mathbf{Xb} +
+\mathbf{Z}_{D}\mathbf{W}_{D} \boldsymbol{\alpha}_{D} + 
+\mathbf{Z}_{D}\mathbf{U}_{D} \boldsymbol{\epsilon}_{D} +
+\mathbf{Z}_{F}\mathbf{W}_{F} \boldsymbol{\alpha}_{F} + 
+\mathbf{Z}_{F}\mathbf{U}_{F} \boldsymbol{\epsilon}_{F} +
+\mathbf{e}.
+\end{align}
+
+Breeding values or, separately for each heterotic group, general combining
+abilities can be estimated as:
 
@@
 half their breeding values.
-The predicted agronomic performance is:
+The predicted agronomic performance of the inbred lines is
 
-\begin{equation} \label{eq:predicted-performance}
-\hat{\mathbf{y}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\hat{\mathbf{g}}_{l} 
+\begin{equation} \label{eq:inbred-predicted-performance}
+\hat{\mathbf{y}} = \mathbf{Z}\mathbf{\hat{g}} 
+\end{equation}
+
+and that of the hybrids is equivalently
+
+\begin{equation} \label{eq:inbred-predicted-performance}
+\hat{\mathbf{y}} = \mathbf{Z}_{D}\mathbf{\hat{g}}_{D} + \mathbf{Z}_{F}\mathbf{\hat{g}}_{F}.
 \end{equation}

### 2017-09-14 analysis/imputed_genotype_predictive_ability.R
compare r for imputed genos only

@@ -1,0 +1,85 @@
+# Goal: Compare single step prediction with individual predictor results.
+
+## -- PACKAGES ----------------------------------------------------------------
+if (!require("pacman")) install.packages("pacman")
+pacman::p_load("tidyverse", "dtplyr", "data.table", "broom")
+
+
+## -- DATA --------------------------------------------------------------------
+dat <- "./data/processed/predictions/predictions.RDS" %>%
+  readRDS() %>%
+  dplyr::mutate(Combi = substr(Interval, start = 1, stop = 3)) %>%
+  data.table::as.data.table()
+
+
+# Find the complement of the full set and the reduced set.
+geno_df <- "./data/derived/predictor_subsets/geno_df.RDS" %>%
+  readRDS()
+
+hybrid_imp_genos <- geno_df %>%
+  dplyr::filter(Material == "Hybrid") %>%
+  base::split(x = ., f = list(.$Extent)) %>%
+  purrr::map(~dplyr::pull(., G)) %>%
+  .[c("Full", "Core")] %>%
+  purrr::reduce(setdiff)
+
+# Specify a unified predictor order.
+pred_order <- c("P", "G", "T", "PG", "PT", "GT")
+
+# Specify a unique trait order.
+trait_order <- c("DMY", "DMC", "ADL", "FAT", "PRO", "STA", "SUG")
+
+# Predefine the color scheme.
+mycols <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6)
+names(mycols) <- c("P", "G", "T", "PG", "PT", "GT")
+
+imputed_hybrid_plot <- dat %>%
+  tibble::as_data_frame() %>%
+  dplyr::filter(Combi == "FHN", TST_Geno %in% hybrid_imp_genos) %>%
+  dplyr::group_by(Trait, Predictor) %>%
+  dplyr::summarize(r = cor(y, yhat)) %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(Trait = forcats::fct_recode(
+    Trait,
+    "DMY" = "GTM",
+    "DMC" = "GTS",
+    "FAT" = "FETT",
+    "PRO" = "RPR",
+    "SUG" = "XZ"
+  )) %>%
+  dplyr::mutate(Trait = forcats::fct_relevel(Trait, trait_order)) %>%
+  dplyr::mutate(
+    Predictor = gsubfn::gsubfn(
+      pattern = "\\S+",
[…]

### 2017-09-14 paper/manuscript.tex
paper-loocv, update loocv

@@
 between $D$ and $F$.
-Let $H_{i \times j}$ denote the test set hybrid to be predicted.
+Note that $H$ does not comprise all possible crosses between $D$ and $F$,
+which would have been $142 \times 103 = 14,626$ hybrids, but a subset of 1,521
+hybrids.
+Let $H_{i \times j}$ denote the hybrid to be predicted.
 We can then define the parents of the training set hybrids as
@@
 The training set is then defined as 
-$H_{TRN}(x \times j) = H \cap (D_{i} \times F_{j})$.
+$H_{TRN}(i \times j) = H \cap (D_{i} \times F_{j})$.
 
@@
 which is calculated as $\rho(\mathbf{y}, \mathbf{\hat{y}})$, where 
-$\mathbf{\hat{y}}$ is the vector of predicted values from each LOOCV run.
+$\mathbf{\hat{y}}$ is the vector of predicted values for all genotypes.
 
 Standard errors for the predictive abilities were calculated by bootstrapping
-predictive and observed values 1,000 times using the \emph{R} package 
+the predicted and observed values 1,000 times using the \emph{R} package 
 \emph{boot}~\cite{Canty2017}.

### 2017-09-14 paper/manuscript.tex
paper-results, update results

@@
 \Genetics2level{Exp2}
-The maize hybrid data set contains information on seven agronomic traits and
-three predictors, namely pedigree data (P), genomic information (G), 
-transcriptomic information (T), and combinations thereof.
+The maize hybrid data set encompasses seven agronomic traits and three
+predictors, namely pedigree data (P), genomic information (G), transcriptomic
+information (T), and combinations thereof.
 Pedigree records and genomic information were available for the 245 parent 
@@
 For FAT and SUG, G was the superior predictor inside the reduced set.
-In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
+In the full set of 1,521 hybrids, predictive abilities obtained with G increased
 markedly for all traits with the exception of FAT\@.
@@
   }
-\label{fig:InbredResults}
+\label{fig:core-results}
 \end{figure}
@@
 Predictive abilities obtained with the different core sets varied widely within
-most traits (Fig.~\ref{fig:InbredResults}).
+most traits (Fig.~\ref{fig:core-results}).
 For most traits, an initial increase in predictive abilities when moving
@@
 genotypes is smaller than the number of features of the incomplete predictor.
-Nevertheless, SS-BLUP suffers from a variety of issues, which we will present
-briefly:
+Nevertheless, SS-BLUP suffers from a variety of issues.
 First, the combination of the numerator with the genomic relationship matrix

### 2017-09-14 analysis/imputed_genotype_predictive_ability.R

@@ -76,7 +76,7 @@ imputed_hybrid_plot <- dat %>%
 
 ggsave(
   plot = imputed_hybrid_plot,
-  filename = "./tabs_figs/pred_ability_hybrid.pdf",
+  filename = "./tabs_figs/imputed_pred_ability_hybrid.pdf",
   width = 7,
   height = 4,
   units = "in"

### 2017-09-14 paper/manuscript.tex
paper-discussion, update

@@
 provide considerably more useful information.
-The latter was the case for \citeNP{Ashraf2016}, who imputed genotypes for about 
-10,000 wheat lines using pedigree information and observed greater accuracy of
-the single step method over genomic BLUP for all four evaluated traits.
+A larger dataset was available to~\citeNP{Ashraf2016}, who imputed genotypes
+for about 10,000 wheat lines using pedigree information and observed greater
+prediction accuracy of the single step method over genomic BLUP for all four
+evaluated traits.
 Predictive abilities reported for the maize diversity panel in this study were
@@
 population structure and excluded all other genotypes from our analyses whereas
-\citeNP{Guo2016} used genotypes from four different subpopulations, modeled as a
-fixed effect.
+\citeNP{Guo2016} used genotypes from four different subpopulations with a model
+using a fixed effect to account for the differences between subpopulations.
 2) We applied quality checks for the predictor data after generating the subset
@@
 For the hybrid data, the predictive abilities of genomic information
-for the whole set of available parent lines rose considerably for several
-agronomic traits.
+increased considerably in the full set of genotypes compared to the reduced
+set for several agronomic traits.
 Hence, the inclusion of trancsriptomic data could not further improve upon
@@
 Beyond this fraction of covered genotypes the trend in predictive ability
-changes was stable.
+changes was stable in either direction.
+Predictive abilities increased simultaneously with $x$ when the incomplete
+predictor outperformed the complete predictor in the reduced set of genotypes
+and decreased with increasing $x$ when the complete predictor was the better
+predictor.
+This pattern clearly elucidates the influence on the difference between the
+two predictors on predictive ability in a single step approach.
 
 \section*{Conclusions}
-We successfully applied single step prediction to a hybrid plant data set 
-while imputing with a quantitative downstream `omics' predictor.
+We successfully applied single step prediction to an inbred and a hybrid plant
+data sets while imputing with a quantitative downstream `omics' predictor.
 By declaring different subsets of individuals as covered by one or two 

### 2017-09-14 paper/manuscript.tex
paper, fix typos

@@
 The material comprised 1,521 hybrids produced in 16 factorial mating designs
-between 142 Dent and 103 Flint parent lines (described in details by
+between 142 Dent and 103 Flint parent lines (described in detail by
 \citeNP{Westhues2017}).
@@
 \end{equation}
-where observed feature covariates are stored in $\mathbf{W}_{2}$ while marker
+where observed feature covariates are stored in $\mathbf{W}_{2}$ while feature
 covariates for genotypes not covered by the incomplete predictor
@@
 Extending this model to hybrids with parent lines from two heterotic groups
-simply requires us to separate the feature matrices of the two predictors
-into a paternal and a maternal component, respectively.
+simply requires separating the feature matrices of the two predictors into a
+paternal and a maternal component, respectively.
 Using the index $D$ for the Dent and the index $F$ for the Flint lines in Exp2

### 2017-09-14 paper/manuscript.tex
paper-methods, tensor matrices

@@
 \usepackage{hyperref} % web links
+\usepackage{tensor} % tensor matrices
 
@@
 incomplete predictor, are indicated by the subscript $2$.
+Let the index $C$ denote matrices and vectors pertaining to the complete
+predictor and let the index $I$ denote matrices and vectors pertaining to the
+incomplete predictor, respectively.
+
 Let us further denote the matrix of centered covariates of the incomplete
@@
 \begin{equation} \label{eq:incomplete-covariates}
-\mathbf{W} =
+\tensor[_{I}]{\mathbf{W}}{} =
 \begin{bmatrix}
-  \mathbf{Z}_{1}\mathbf{\hat{W}}_{1} \\
-  \mathbf{W}_{2}
+  \mathbf{Z}_{1} \; \tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} \\
+  \tensor[_I]{\mathbf{W}}{_{2}}
  \end{bmatrix},
 \end{equation}
-where observed feature covariates are stored in $\mathbf{W}_{2}$ while feature
+where observed feature covariates are stored in $\tensor[_I]{\mathbf{W}}{_{2}}$ while feature
 covariates for genotypes not covered by the incomplete predictor
-($\mathbf{\hat{W}}_{1} = \mathbf{K}_{12} \mathbf{K}^{-1}_{22} \mathbf{W}_{2}$)
+($\tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} = \tensor[_C]{\mathbf{K}}{_{12}} \; \tensor[_C]{\mathbf{K}}{_{22}^{-1}} \; \tensor[_I]{\mathbf{W}}{_{2}}$)
 need to be imputed.
@@
 covered only by the complete predictor) and a component
-$\mathbf{g}_{2} = \mathbf{W}_{2}\boldsymbol{\alpha}$ (\textit{i.e.,} genetic
-values of genotypes covered by both predictors where $\boldsymbol{\alpha}$
+$\mathbf{g}_{2} = \tensor[_{I}]{\mathbf{W}}{_{2}} \; 
+\tensor[_{I}]{\boldsymbol{\alpha}}{}$ (\textit{i.e.,} genetic
+values of genotypes covered by both predictors where $\tensor[_{I}]{\boldsymbol{\alpha}}{}$
 pertains to the random feature effects of the incomplete predictor).
@@
 	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
-	&= \mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{W}_{2}
-	\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{K}_{12}
-	\mathbf{K}^{-1}_{2}
-	\boldsymbol{\hat{\alpha}}) \\
+	&= \tensor[_{C}]{\mathbf{K}}{_{12}} \; \tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \; \tensor[_{I}]{\mathbf{W}}{_{2}} \;
+	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}} + (\mathbf{g_1} - \tensor[_{C}]{\mathbf{K}}{_{12}} \;
+	\tensor[_{C}]{\mathbf{K}}{_{2}^{-1}} \;
+	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}}{}) \\
 	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
@@
 
-where $\mathbf{K}_{ij}$ are partitions of $\mathbf{K}$ that correspond to 
-$\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and
-where $\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values of genotypes
+where $\tensor[_{C}]{\mathbf{K}}{_{ij}}$ are partitions of
+$\tensor[_{C}]{\mathbf{K}}{}$ that correspond to $\mathbf{g}_{1}$ and
+$\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and where
+$\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values of genotypes
 not equipped with the incomplete predictor.
 
-The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{K}_{11} -
-\mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{K}_{21} =
-(\mathbf{K}^{11})^{-1}$
-\cite{Legarra2009}.
+The covariance matrix of $\boldsymbol{\epsilon}$ is
+$\tensor[_{C}]{\mathbf{K}}{_{11}} -
+\tensor[_{C}]{\mathbf{K}}{_{12}} \;
+\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
+\tensor[_{C}]{\mathbf{K}}{_{21}} =
+(\tensor[_{C}]{\mathbf{K}}{^{11}})^{-1}$
+\cite{Legarra2009}
+
+and the covariance of $\mathbf{\hat{g}}_{1}$ is
+$\tensor[_{C}]{\mathbf{K}}{_{12}} \;
+\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
+\tensor[_{I}]{\mathbf{W}}{_{2}} \;
+\tensor[_{I}]{\mathbf{W}}{_{2}^{\top}} \;
+\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
+\tensor[_{C}]{\mathbf{K}}{_{21}}\sigma^{2}_{\alpha}$
+\cite{Fernando2014}.
 
@@
 \mathbf{y} =
-\mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U} \boldsymbol{\epsilon} + \mathbf{e},
+\mathbf{Xb} + \tensor[_{I}]{\mathbf{W}}{} \; \tensor[_{I}]{\boldsymbol{\alpha}} + \mathbf{U} \boldsymbol{\epsilon} + \mathbf{e},
 \end{equation}
@@
 \begin{bmatrix}
-  \mathbf{K}_{12}\mathbf{K}^{-1}_{22}\mathbf{W}_2
-  \boldsymbol{\alpha} + \boldsymbol{\epsilon}  \\
-  \mathbf{W}_2\boldsymbol{\alpha} \\
+  \tensor[_{C}]{\mathbf{K}}{_{12}} \; \tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
+  \tensor[_{I}]{\mathbf{W}}{_{2}} \;
+  \tensor[_{I}]{\boldsymbol{\alpha}} + \boldsymbol{\epsilon}  \\
+  \tensor[_{I}]{\mathbf{W}}{_{2}} \; \tensor[_{I}]{\boldsymbol{\alpha}} \\
  \end{bmatrix}
@@
 \begin{bmatrix}
-  \mathbf{\hat{W}}_{1}\boldsymbol{\alpha} + \boldsymbol{\epsilon} \\
-  \mathbf{W}_2\boldsymbol{\alpha} \\
+  \tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} \; \tensor[_{I}]{\boldsymbol{\alpha}} + \boldsymbol{\epsilon} \\
+  \tensor[_{I}]{\mathbf{W}}{_{2}} \; \tensor[_{I}]{\boldsymbol{\alpha}} \\
  \end{bmatrix}
@@
 \mathbf{y} &= \mathbf{Xb} +
-\mathbf{Z}\mathbf{W} \boldsymbol{\alpha} +
+\mathbf{Z} \; \tensor[_{I}]{\mathbf{W}}{} \;
+\tensor[_{I}]{\boldsymbol{\alpha}} +
 \mathbf{Z}\mathbf{U} \boldsymbol{\epsilon} +
@@
 \mathbf{y} &= \mathbf{Xb} +
-\mathbf{Z}_{D}\mathbf{W}_{D} \boldsymbol{\alpha}_{D} + 
-\mathbf{Z}_{D}\mathbf{U}_{D} \boldsymbol{\epsilon}_{D} +
-\mathbf{Z}_{F}\mathbf{W}_{F} \boldsymbol{\alpha}_{F} + 
+\mathbf{Z}_{D} \;
+\tensor[_{I}]{\mathbf{W}}{_{D}} \; \tensor[_{I}]{\boldsymbol{\alpha}}{_{D}} +
+\mathbf{Z}_{D} \mathbf{U}_{D} \boldsymbol{\epsilon}_{D} +
+\mathbf{Z}_{F} \;
+\tensor[_{I}]{\mathbf{W}}{_{F}} \;
+\tensor[_{I}]{\boldsymbol{\alpha}}{_{F}} +
 \mathbf{Z}_{F}\mathbf{U}_{F} \boldsymbol{\epsilon}_{F} +
@@
  \begin{bmatrix}
-  \hat{\mathbf{W}_1} \\
-  \mathbf{W}_2 
+  \tensor[_{I}]{\hat{\mathbf{W}}}{_{1}} \\
+  \tensor[_{I}]{\mathbf{W}}{_{2}}
  \end{bmatrix}
- \hat{\boldsymbol{\alpha}}
+ \tensor[_{I}]{\hat{\boldsymbol{\alpha}}}{}
  +

### 2017-09-15 analysis/visualize_predictions.R
add G and T to core-r plot

@@ -188,7 +188,13 @@ line_type_df <- inbred_df %>%
 
 
 core_df <- inbred_df %>%
-  dplyr::filter(Predictor == "GT", Combi == "CIA") %>%
+  dplyr::filter(Combi == "CIA") %>%
+  dplyr::mutate(Core_Fraction = dplyr::case_when(
+    Predictor == "T" ~ "1",
+    Predictor == "G" ~ "0",
+    Predictor == "GT" ~ Core_Fraction
+    )
+  ) %>%
   dplyr::mutate_at(vars(Predictor), funs(as.factor)) %>%
   dplyr::inner_join(y = line_type_df, by = "Trait")
 
@@ -203,11 +209,13 @@ core_plot <- core_df %>%
     )
   ) +
   geom_line(aes(linetype = LineType), size = 1.2) +
+  geom_point(size = 3, shape = 4, color = "black") +
   ggthemes::scale_color_tableau() +
   ggthemes::theme_pander(base_size = 10) +
   theme(legend.position = "top") +
   guides(color = guide_legend(override.aes = list(size = 3))) +
-  scale_linetype(guide = "none")
+  scale_linetype(guide = "none") +
+  scale_shape(guide = "none")
 
 ggsave(
   plot = core_plot,

### 2017-09-15 paper/manuscript.tex
update figure captions

@@
 \centering
-\caption{Acronyms for the trait names}
+\caption{Abbreviations for the trait names}
 \label{table:TraitKey}
@@
   \caption{
-  Principal component analysis of 211 genotypes from the set
+  Principal component analysis (PCA) of 211 genotypes from the set
   `tropical/subtropical' of the maize diversity panel inbred lines.
@@
   clusters.
+  The explained variance for each of the two PCs is given in the corresponding
+  axis labels.
 }
@@
   The two extended bars represent the entire set of genotypes in either
-  Exp2 or Exp1.
+  Exp1 or Exp2.
   All genotypes are covered by at least one predictor, which is called the
@@
   (\textbf{b}) Subsampling within core sets.
-  Each of the black symbols represents a different genotype.
+  Each of the white symbols represents a different genotype.
   For each value of $x$, 100 subsets $s$ are generated where $x\%$ of genotypes
@@
   on predictive abilities.
-  Predictive abilities are based on transcriptomic effects derived from a
-  single step prediction model involving transcriptomic (T) and genomic (G)
-  data, respectively.
   Results are shown for the core set of 110 inbred lines from Exp1 for
@@
   T was available is given on the x-axis.
+  Predictive abilities for core fractions between 0.1 and 0.9 are based on
+  transcriptomic effects derived from a single step prediction model involving
+  transcriptomic (T) and genomic (G) data, respectively.
+  The predictive ability at a core fraction of 0 is based only on G and at a
+  core fraction of 1 it is based only on T.
   Dotted lines indicate that $T > G$, dashed lines indicate that $T \approx G$

### 2017-09-15 paper/manuscript.tex
paper, init abstract

@@
 \begin{abstract}
+Predicting the performance of candidates is indispensable for selection
+decisions in animal and plant breeding.
+Novel `omics' predictors have shown to improve upon genomics-based predictions
+of some complex traits but are yet expensive to sample.
+Here we show that single step prediction can be applied to a quantitative
+predictor.
 
-  $\dots$
+Using two maize datasets of inbred lines and hybrids we show that transcriptomic
+data can successfully be imputed, boosting forecasts when their predictive
+ability exceeds that of pedigree or genomic data.
 
+Equipping only a subset of genotypes with `omics' predictors and imputing all
+others using pedigree or genomic data could enable breeders to improve trait
+predictions while keeping costs manageable.
+Employing `omics' predictors could particularly benefit candidate selection in
+hybrid breeding.
 \end{abstract}

### 2017-09-15 paper/manuscript.tex
AEM suggestions up to methods-prediction

@@
 Breeding programs --- particularly in animal breeding --- are characterized by
-the presence of one predictor type, which is available for all recorded
+the availability of one predictor type, which is available for all recorded
 genotypes while other predictors are incomplete in that they cover only a 
@@
 To utilize all available phenotypic information for model training, methods were 
-developed that allowed for a combined use of genomic and pedigree information 
+developed that allow the combined use of genomic and pedigree information 
 \cite{Hayes2009a,VanRaden2009}.
@@
 pedigree relationships and the phenotypes of relatives.
-Individuals with accurate breeding values out of the first step (that is, a
-high number of offspring), enter the second step in which conventional breeding
+Individuals with accurate breeding values out of the first step, which have a
+high number of offspring, enter the second step in which conventional breeding
 values are regressed onto SNP genotypes.
-The estimated model coefficients can then be used to predict genomic breeding 
-values for solely genotyped and potentially very young individuals.
+The estimated model coefficients can then be used to predict breeding values
+for solely genotyped and potentially very young individuals.
 However, as the breeding values from step one enter the genomic prediction
@@
 For instance, the range of the prediction errors will no longer be identically
-distributed, \textit{i.e.}, they should be weighted differently or estimated 
+distributed and thus, they should be weighted differently or estimated 
 individually~\cite{Aguilar2010}.
@@
 trait-independent~\cite{Vitezica2011,Ashraf2016}.
-\citeNP{Fernando2014} derived the equivalent single step marker-effect-model to 
+\citeN{Fernando2014} derived the equivalent single step marker-effect-model to 
 the single step breeding value model of \citeNP{Legarra2009} and 
@@
 Hence, our first objective was to examine whether the single step prediction 
-framework can be transfered to hybrids in plants.
+framework can be transfered to a situation where a very large number of hybrids
+($\approx$ 1 million) needs to be predicted.
 
@@
 \cite{Guo2016,Xu2016,Dan2016,Zenke-Philippi2017,Westhues2017}.
-This situation motivated our second objective, namely exploring the utility of
-single step predition with a quantitative predictor.
 Yet, these technologies are not fully mature and therefore incur higher costs
@@
 despite their proven usefulness.
-Our third objective is therefore to evaluate the impact of equipping only a
+Our second objective is therefore to evaluate the impact of covering only a
 subset of all genotypes with both, the complete and the incomplete predictor on
-predictive ability.
+predictive ability, where for the remaining individuals only genomic data are
+available.
 
@@
 \Genetics2level{Exp1}
-To address objectives 2 and 3, a maize inbred line data set, subsequently
+To address the second objective a maize inbred line data set, subsequently
 referred to as Exp1, was used~\cite{Yang2014}.
@@
 by~\citeNP{Yang2014}.
-Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
-six are presented in this study (Table~\ref{table:TraitKey}).
-The selection of these traits was done so that four traits were better 
-predicted by transcriptomic data and two further traits were better predicted 
-by genomic data.
+Best linear unbiased predictors (BLUPs) were calculated for 17
+traits~\cite{Guo2016} of which six are presented in this study
+(Table~\ref{table:TraitKey}).
+The selection of these traits was done so that four traits were predicted with
+greater precision by transcriptomic data and two further traits were predicted
+with greater precision by genomic data.
 MaizeSNP50 BeadChip-data were available for the entire set of 513 lines from the
@@
 A subset of 28,850 annotated genes was available for 149 out of the 211
-genotypes and was kept for further analyses~\cite{Li2013}.
+genotypes~\cite{Fu2013} and provided the basis for further analyses.
 
@@
 The material comprised 1,521 hybrids produced in 16 factorial mating designs
-between 142 Dent and 103 Flint parent lines (described in detail by
-\citeNP{Westhues2017}).
+between 142 Dent and 103 Flint parent lines described in detail by
+\citeN{Westhues2017}.
 Best linear unbiased estimates (BLUEs) were computed across all factorials and 
@@
 Based on the estimated probability of the data for each $K$, we chose $K=4$
-and ran a principal component analysis (PCA) on the scaled and centered
+and carried out a principal component analysis (PCA) on the scaled and centered
 genomic feature matrix using the \emph{pca} function from the \emph{R}-package
@@
 subpopulations~\cite{Windhausen2012}, we reduced the panel of inbred lines to
-those primarily belonging to either cluster 1 or cluster 4 in 
-Fig.~\ref{fig:inbred-pca}.
+those assigned to either cluster 1 or cluster 4 in Fig.~\ref{fig:inbred-pca}.
 This reduction yielded 164 genotypes, 110 of which were covered by both, genomic

### 2017-09-15 paper/manuscript.tex
updates after aem confus

@@
 
-% -   Objectives (Hybrids)
+% -   Hybrids
 In animal breeding, single step prediction, particularly H-BLUP, has already 
@@
 white spruce lines, respectively~\cite{Ratcliffe2017}.
-While single step prediction has been applied to pig hybrids~\cite{Xiang2016}
-this application is not equivalent to hybrid breeding in plants.
-There, the maternal and the paternal crossing partner can be selected without
-constraints whereas animal breeders select only males because of the limited
-fertility of the female animals.
+Moreover, while an application of the single step framework to animals exists
+for pigs~\cite{Xiang2016}, it is yet lacking for plant hybrids.
 Technological advances in plant hybrid breeding such as the rapid generation
@@
 \cite{Technow2014,Kadam2016,Westhues2017}.
-Hence, our first objective was to examine whether the single step prediction 
-framework can be transfered to a situation where a very large number of hybrids
-($\approx$ 1 million) needs to be predicted.
 
-% -   Objectives (quantitative predictores)
+% -   quantitative predictores
 Given that the availability of pedigree and genomic data is the most prevalent
@@
 despite their proven usefulness.
-Our second objective is therefore to evaluate the impact of covering only a
-subset of all genotypes with both, the complete and the incomplete predictor on
-predictive ability, where for the remaining individuals only genomic data are
-available.
+
+% - Objectives
+Hence, our objectives were to i) examine whether the single step prediction 
+framework can be transfered to a situation where a very large number of hybrids
+($\approx$ 1 million) needs to be predicted and ii) evaluate the impact of
+covering only a subset of all genotypes with both, the complete and the
+incomplete predictor on predictive ability, where for the remaining individuals
+only genomic data are available.
 
@@
 diversity  and was reduced to the set of tropical and subtropical lines
-($n = 211$) given that it is the largest of the four pre-classified
-subgroups~\cite{Yang2014}.
+($n = 211$) given that it is the largest of the four subgroups classified by
+\citeN{Yang2014}.
 All inbred lines were evaluated in five Chinese environments described in detail
 by~\citeNP{Yang2014}.
-Best linear unbiased predictors (BLUPs) were calculated for 17
-traits~\cite{Guo2016} of which six are presented in this study
+Best linear unbiased predictors (BLUPs) were calculated by~\citeN{Yang2014} for
+17 traits~\cite{Guo2016} of which six are presented in this study
 (Table~\ref{table:TraitKey}).
@@
 subpopulations~\cite{Windhausen2012}, we reduced the panel of inbred lines to
-those assigned to either cluster 1 or cluster 4 in Fig.~\ref{fig:inbred-pca}.
+those assigned to the largely overlapping clusters 1 and 4 in
+Fig.~\ref{fig:inbred-pca}.
 This reduction yielded 164 genotypes, 110 of which were covered by both, genomic
@@
 \subsection{Kernels}
-BLUP models are computationally attractive when the number of features $p$
-exceeds the number of genotypes $n$ and are equivalent to a selection index
-when fixed effects have been accomodated by the dependent variable 
-\cite{Mrode2014}.
 Depending on the data set, up to three predictors were available for agronomic
@@
 The corresponding feature matrix of the inbred lines ($\mathbf{M}$) has
-dimensions $n \times p$ where $n$ pertains to the number of genotypes.
-For the sake of clarity, we do not distinguish between models for Exp1 and Exp2
-at this point.
-It should be noted, however, that in the case of Exp2 the term `genotype' refers
-strictly to the parent lines.
+dimensions $n \times p$ where $n$ pertains to the number of genotypes and $p$
+pertains to the number of features; that is transcripts for T and SNPs for G.
+For the sake of simplicity the derivations are detailed first for predicting
+breeding values of inbred lines.
+Later, we briefly describe the extension to models required for predicting the
+genetic value of hybrids.
 For the complete predictor we generated kernels by centering and standardizing
@@
 where $M$ denotes the number of features~\cite{VanRaden2008}.
-In the case of pedigree data (P), coancestry coefficients were used directly
-for $\mathbf{K}$.
+In the case of pedigree data (P), the numerator relationship matrix was used
+directly for $\mathbf{K}$.
 
 \subsection{Genetic model}
-The general model for genetic values was:
+The general model for breeding values of the lines was:
 
@@
 \subsection{Single step prediction}
-If the genetic values are fully captured by the features in $\mathbf{M}$,
+If the breeding values are fully captured by the features in $\mathbf{M}$,
 Eq.~\ref{eq:KBLUPModel} can be expressed as
@@
 \begin{align} \label{eq:KBLUPModelSimplified}
-	\mathbf{y} &= \mu + \mathbf{Z} \mathbf{g} + \boldsymbol{e} \\
 	&= \mu + \mathbf{Z} \mathbf{M} \boldsymbol{\alpha} + \boldsymbol{e},
@@
 \end{equation}
-where observed feature covariates are stored in $\tensor[_I]{\mathbf{W}}{_{2}}$ while feature
-covariates for genotypes not covered by the incomplete predictor
-($\tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} = \tensor[_C]{\mathbf{K}}{_{12}} \; \tensor[_C]{\mathbf{K}}{_{22}^{-1}} \; \tensor[_I]{\mathbf{W}}{_{2}}$)
-need to be imputed.
+where observed feature covariates are stored in $\tensor[_I]{\mathbf{W}}{_{2}}$
+while the feature covariates for genotypes not covered by the incomplete
+predictor can be estimated/imputed with the aid of the covariance matrix of the
+complete predictor $\tensor[_{C}]{\mathbf{K}}{}$ and the feature matrix of the
+incomplete predictor $\tensor[_{I}]{\mathbf{W}}{_{22}}$ by calculating
+($\tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} = \tensor[_C]{\mathbf{K}}{_{12}} \;
+\tensor[_C]{\mathbf{K}}{_{22}^{-1}} \; \tensor[_I]{\mathbf{W}}{_{2}}$).
 In this case we need to partition the vector of genetic values $\mathbf{g}$
@@
 \begin{align} \label{eq:mrna1}
-	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
-	&= \tensor[_{C}]{\mathbf{K}}{_{12}} \; \tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \; \tensor[_{I}]{\mathbf{W}}{_{2}} \;
-	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}} + (\mathbf{g_1} - \tensor[_{C}]{\mathbf{K}}{_{12}} \;
-	\tensor[_{C}]{\mathbf{K}}{_{2}^{-1}} \;
+	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) +
+	\boldsymbol{\epsilon} \\
+	&= \tensor[_{C}]{\mathbf{K}}{_{12}} \;
+       	\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
+       	\tensor[_{I}]{\mathbf{W}}{_{2}} \;
+	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}}{} +
+       	(\mathbf{g_1} - \tensor[_{C}]{\mathbf{K}}{_{12}} \;
+	\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
+       	\tensor[_{I}]{\mathbf{W}}{_{2}} \;
 	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}}{}) \\
@@
 
-where $\tensor[_{C}]{\mathbf{K}}{_{ij}}$ are partitions of
+where $\tensor[_{C}]{\mathbf{K}}{_{ij}}$ are partitions of the matrix
 $\tensor[_{C}]{\mathbf{K}}{}$ that correspond to $\mathbf{g}_{1}$ and
@@
 \tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-\tensor[_{C}]{\mathbf{K}}{_{21}}\sigma^{2}_{\alpha}$
+\tensor[_{C}]{\mathbf{K}}{_{21}} \;
+\tensor[_{I}]{\sigma^{2}_{\alpha}}{}$
 \cite{Fernando2014}.
 
-The structure of the residual imputation/prediction error is known and can 
-therefore be modeled.
+The structure of the residual imputation/prediction error is known and,
+therefore, can modeled.
 
@@
 \mathbf{y} =
-\mathbf{Xb} + \tensor[_{I}]{\mathbf{W}}{} \; \tensor[_{I}]{\boldsymbol{\alpha}} + \mathbf{U} \boldsymbol{\epsilon} + \mathbf{e},
+\mathbf{Xb} +
+\tensor[_{I}]{\mathbf{W}}{} \;
+\tensor[_{I}]{\boldsymbol{\alpha}}{} +
+\mathbf{U} \boldsymbol{\epsilon} +
+\mathbf{e},
 \end{equation}
@@
 
-Following the notation in \citeNP{Fernando2014} the phenotypes can be untangled 
-as:
+According to the notation in \citeNP{Fernando2014} the phenotypes can be
+untangled as:
 
@@
   \tensor[_{I}]{\mathbf{W}}{_{2}} \;
-  \tensor[_{I}]{\boldsymbol{\alpha}} + \boldsymbol{\epsilon}  \\
+  \tensor[_{I}]{\boldsymbol{\alpha}}{} + \boldsymbol{\epsilon}  \\
   \tensor[_{I}]{\mathbf{W}}{_{2}} \; \tensor[_{I}]{\boldsymbol{\alpha}} \\
@@
 \begin{bmatrix}
-  \tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} \; \tensor[_{I}]{\boldsymbol{\alpha}} + \boldsymbol{\epsilon} \\
+  \tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} \;
+  \tensor[_{I}]{\boldsymbol{\alpha}}{} + \boldsymbol{\epsilon} \\
   \tensor[_{I}]{\mathbf{W}}{_{2}} \; \tensor[_{I}]{\boldsymbol{\alpha}} \\
@@
 
-The above equation implies that ultimately single step prediction uses only
-one random effect vector $\mathbf{g}$, which entails the observed effects of
-the incomplete predictor $\mathbf{g}_{2}$ and the imputed effects
-$\mathbf{\hat{g}}_{1}$ for genotypes that are not equipped with the incomplete
-predictor plus the residual imputation error $\boldsymbol{\epsilon}$.
+The model above indicates that single step prediction models ultimately use
+only the matrix of known and imputed feature covariates $\mathbf{W}$ for the
+estimation of breeding values while accounting for the imputation error
+$\boldsymbol{\epsilon}$.
 
@@
 \mathbf{y} &= \mathbf{Xb} +
-\mathbf{Z} \; \tensor[_{I}]{\mathbf{W}}{} \;
-\tensor[_{I}]{\boldsymbol{\alpha}} +
+\mathbf{Z} \;
+\tensor[_{I}]{\mathbf{W}}{} \;
+\tensor[_{I}]{\boldsymbol{\alpha}}{} +
 \mathbf{Z}\mathbf{U} \boldsymbol{\epsilon} +
@@
 
-Extending this model to hybrids with parent lines from two heterotic groups
-simply requires separating the feature matrices of the two predictors into a
-paternal and a maternal component, respectively.
-Using the index $D$ for the Dent and the index $F$ for the Flint lines in Exp2
-we then have a feature matrix $\mathbf{M}_{D}$ and $\mathbf{M}_{F}$ for each
-predictor.
-For the hybrids the final model is therefore simply
-\begin{align} \label{eq:final-model}
-\mathbf{y} &= \mathbf{Xb} +
-\mathbf{Z}_{D} \;
-\tensor[_{I}]{\mathbf{W}}{_{D}} \; \tensor[_{I}]{\boldsymbol{\alpha}}{_{D}} +
-\mathbf{Z}_{D} \mathbf{U}_{D} \boldsymbol{\epsilon}_{D} +
-\mathbf{Z}_{F} \;
-\tensor[_{I}]{\mathbf{W}}{_{F}} \;
-\tensor[_{I}]{\boldsymbol{\alpha}}{_{F}} +
-\mathbf{Z}_{F}\mathbf{U}_{F} \boldsymbol{\epsilon}_{F} +
-\mathbf{e}.
-\end{align}
-
-Breeding values or, separately for each heterotic group, general combining
-abilities can be estimated as:
+Breeding values can be estimated as:
 
@@
 
-For the hybrid data, the obtained GCA effects for the inbred lines represent 
-half their breeding values.
 The predicted agronomic performance of the inbred lines is
@@
 
-and that of the hybrids is equivalently
+The previous models can simply be extended to hybrids by defining separate
+feature matrices for each predictor and each heterotic group; here
+$\mathbf{M}_{D}$ for Dent and $\mathbf{M}_{F}$ for Flint parent lines in Exp2.
+The general model for the calculation of genetic values in Exp2 is then
 
-\begin{equation} \label{eq:inbred-predicted-performance}
-\hat{\mathbf{y}} = \mathbf{Z}_{D}\mathbf{\hat{g}}_{D} + \mathbf{Z}_{F}\mathbf{\hat{g}}_{F}.
+\begin{equation} \label{eq:hybrid-genetic-value-model}
+  \mathbf{y} = \mu + \mathbf{Z}_{D} \mathbf{g}_{D} +
+  \mathbf{Z}_{F} \mathbf{g}_{F} +
+  \mathbf{e},
 \end{equation}
+
+corresponding to Eq.~\ref{eq:KBLUPModel}.
 
@@
 Flint group, respectively.
-Further, let $H \cap [D \times F]$ denote the set of hybrids from crosses 
-between $D$ and $F$.
-Note that $H$ does not comprise all possible crosses between $D$ and $F$,
-which would have been $142 \times 103 = 14,626$ hybrids, but a subset of 1,521
+The set $H$ is the subset of 1,521 phenotyped hybrids within the set of all
+possible hybrids $D \times F$, which would comprise $14,626 = 142 * 103$
 hybrids.
-Let $H_{i \times j}$ denote the hybrid to be predicted.
+Let $i \times j$ denote the hybrid to be predicted.
 We can then define the parents of the training set hybrids as
@@
 We judged the performance of each model by looking at its predictive ability,
-which is calculated as $\rho(\mathbf{y}, \mathbf{\hat{y}})$, where 
-$\mathbf{\hat{y}}$ is the vector of predicted values for all genotypes.
+which was calculated as the correlation $\rho(\mathbf{y}, \mathbf{\hat{y}})$,
+where $\mathbf{\hat{y}}$ is the vector of predicted values for all genotypes.
 
@@
 sets of nine different sizes from Exp1 (Fig.~\ref{fig:compound-plot}~a).
-The base of each of these core sets was the superset of genotypes that were
+The base of each of these core sets was the set of 110 genotypes that were
 covered by both, genomic and trancriptomic data.
@@
 imputed.
+For the each core set, comprising 110 genotypes, predictive abilities were again
+computed as described above, using LOOCV and bootstrapping.
 
@@
 At the level of the 110 genotypes for which both, genomic (G) and 
-transcriptomic (T) information, were available, the use of T yielded higher
+transcriptomic (T) information were available, the use of T yielded higher
 predictive abilities than the use of G for two out of six traits
@@
 ($n = 110$).
-The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
+The combination of G and T (\textit{i.e.}, $\tensor[_{C}]{G}{}
+\tensor[_{I}]{T}{}$, where the subscripts correspond to the complete ($C$) and
+the incomplete ($I$) predictor, respectively) in a single-step approach for the 
 prediction of all 164 lines, performed at least as well as G alone when T was 
 superior over G for the reduced set of 110 lines.
-In the case of KW and ED, the use of GT yielded a slight improvement in predictive 
-ability over G.
+In the case of KW and ED, the use of $\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$
+yielded a slight improvement in predictive ability over G.
 Where T was a worse predictor for the reduced set of 110 lines than G,
-predictive abilities of GT for the full set of 164 lines were also worse than 
-those based on G alone.
+predictive abilities of $\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$ for the full set
+of 164 lines were also worse than those based on G alone.
 
@@
   As predictors, genomic (G) data, transcriptomic (T) data and their 
-  combination (GT) were used.
+  combination ($\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$) were used.
   The `Reduced' panel includes a subset of 110 lines covered by genomic and
@@
 markedly for all traits with the exception of FAT\@.
-The combination of genomic and transriptomic information (GT) in a single-step
-model yielded a small improvement over G alone for the traits DMY and PRO\@.
+The combination of genomic and transriptomic information
+($\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$) in a single-step model yielded a small
+improvement over G alone for the traits DMY and PRO\@.
 Except for PRO, compared to using only G, trait prediction did not benefit from 
 the single step approach when T was not superior to G in the reduced set.
-Predictive abilities resulting from combinations of pedigree with genomic (PG)
-information were better than P alone for all traits and PT improved upon P in
+Predictive abilities resulting from combinations of pedigree with genomic
+($\tensor[_{C}]{P}{}\tensor[_{I}]{G}{}$) information were better than P alone
+for all traits and $\tensor[_{C}]{P}{}\tensor[_{I}]{T}{}$ improved upon P in
 the `full' set for four traits.
+
+A previous study using the same data as in Exp2 concluded that modeling of
+specific combining ability (SCA) effects did not appreciably improve predictive
+abilities over that of a purely additive model for any of the investigated
+traits.
+Nevertheless, the inclusion of SCA effects in a single step prediction model
+would be as straightforward as computing the Kronecker product between extension of single step prediction
 
@@
     As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
-    combinations thereof we used.
+    combinations thereof were used.
     The `Full' set of genotypes comprises 1,521 hybrid progeny based on 245
@@
 \Genetics2level{Impact of coverage by the incomplete predictor}
-Single-step approaches are based on the use of one complete predictor that
-covers all genotypes and one incomplete predictor that covers only a subset of
-all available genotypes.
-Here, we explore the impact of which data are covered by the incomplete 
-predictor on the predictive abilities obtained via single-step approaches by
-focusing exclusively on the 110 inbred lines from the maize diversity panel
-that are covered by both, genomic and transcriptomic information.
-First, we assembled nine different core sets, each with 100 nested subsamples,
-which were generated by randomly removing transcriptomic information from
-some genotypes (Fig.~\ref{fig:compound-plot}).
-The nine core sets differ in that each covers a different number of genotypes,
-ranging from only 10\% to 90\% of all genotypes covered by the incomplete
-predictor (Fig.~\ref{fig:inbred-pca}).
+Predictive abilities for the six traits in the the different core sets of Exp1
+varied widely within most traits (Fig.~\ref{fig:core-results}).
+For most traits, an initial increase in predictive abilities when moving
+from a core set size of 10\% to 20\%/30\% was observed, regardless of whether G
+or T individually was the superior predictor.
+For the traits CW, GW and PH, which were better predicted via G than via T in
+the reduced set of lines, predictive abilities were higher when only a small
+number of genotypes was included in the core set.
+This corresponds to a situation where a large number of genotypes is being 
+imputed and the majority of the information is derived from genomic information.
+A hike in predictive abilities, when moving from a core fraction of 10\% to 
+90\%, was observed for ED and KW, which were predicted better when using T
+individually compared to using G individually.
+The only trait that was barely affected by a change in the core fraction was DS
+for which the predictive abilities using either G or T individually were almost
+identical (Fig.~\ref{fig:InbredResults}).
 
@@
 \end{figure}
-
-Predictive abilities obtained with the different core sets varied widely within
-most traits (Fig.~\ref{fig:core-results}).
-For most traits, an initial increase in predictive abilities when moving
-from a core set size of 10\% to 20\%/30\% was observed, regardless of whether G
-or T individually was the superior predictor.
-For the traits CW, GW and PH, which were better predicted via G than via T in
-the reduced set of lines, predictive abilities were higher when only a small
-number of genotypes was included in the core set.
-This corresponds to a situation where a large number of genotypes is being 
-imputed and the majority of the information is derived from genomic information.
-A hike in predictive abilities, when moving from a core fraction of 10\% to 
-90\%, was observed for ED and KW, which were predicted better when using T
-individually compared to using G individually.
-The only trait that was barely affected by a change in the core fraction was DS
-for which the predictive abilities using either G or T individually were almost
-identical (Fig.~\ref{fig:InbredResults}).
 
@@
 metabolomic data.
+While the heritability of relationship coefficients and genomic markers can be
+assumed to be equal to 1, this is not the case for the aforementioned
+quantitative `omics' predictors.
 Here, we considered single step approaches for a maize inbred line diversity
@@
 provide considerably more useful information.
-A larger dataset was available to~\citeNP{Ashraf2016}, who imputed genotypes
+A larger dataset was available to~\citeN{Ashraf2016}, who imputed genotypes
 for about 10,000 wheat lines using pedigree information and observed greater
@@
 Predictive abilities reported for the maize diversity panel in this study were
-slightly different from those reported by \citeNP{Guo2016}, for three
+slightly different from those reported by~\citeN{Guo2016}, for three
 possible reasons:
@@
 population structure and excluded all other genotypes from our analyses whereas
-\citeNP{Guo2016} used genotypes from four different subpopulations with a model
+\citeN{Guo2016} used genotypes from four different subpopulations with a model
 using a fixed effect to account for the differences between subpopulations.
 2) We applied quality checks for the predictor data after generating the subset
-of 164 inbred lines, whereas \citeNP{Guo2016} applied the quality checks to 368
+of 164 inbred lines, whereas~\citeN{Guo2016} applied the quality checks to 368
 genotypes.
-3) While \citeNP{Guo2016} used five-fold cross-validation with 500 repetitions,
+3) While~\citeN{Guo2016} used five-fold cross-validation with 500 repetitions,
 we employed a leave-one-out cross-validation scheme (LOOCV).
@@
 progeny.
-\citeNP{Westhues2017} have shown that, in hybrid breeding, the probability 
+\citeN{Westhues2017} have shown that, in hybrid breeding, the probability 
 of successfully selecting the best observed genotypes based on the best 
@@
 incomplete predictor.
-We observed that about $20-30\%$ of genotypes needed to be covered by both
+We observed that about $x=20-30\%$ of genotypes needed to be covered by both
 predictors to match the performance achieved with the complete predictor

### 2017-09-17 paper/manuscript.tex
paper, first full abstract draft

@@
 \begin{abstract}
-Predicting the performance of candidates is indispensable for selection
-decisions in animal and plant breeding.
-Novel `omics' predictors have shown to improve upon genomics-based predictions
-of some complex traits but are yet expensive to sample.
-Here we show that single step prediction can be applied to a quantitative
-predictor.
-
+Predicting genotypic values with great precision is pivotal for effective
+candidate selections in animal and plant breeding.
+Novel `omics' predictors have shown to improve upon established genome-based
+predictions of some complex traits but are yet expensive to sample.
+In animal breeding, single step prediction was introduced to efficiently use
+pedigree information, collected on a large number of animals, and genomic
+information, collected on a smaller subset of animals, simultaneously without
+bias.
+Single step prediction has not been attempted with a predictor whose features
+display heritabilities smaller than one, yet.
 Using two maize datasets of inbred lines and hybrids we show that transcriptomic
-data can successfully be imputed, boosting forecasts when their predictive
+data can be imputed successfully, boosting forecasts when their predictive
 ability exceeds that of pedigree or genomic data.
-
-Equipping only a subset of genotypes with `omics' predictors and imputing all
+Covering only a subset of genotypes with `omics' predictors and imputing all
 others using pedigree or genomic data could enable breeders to improve trait
@@
 Employing `omics' predictors could particularly benefit candidate selection in
-hybrid breeding.
+hybrid breeding where the success of forecasts is a strongly convex function of
+predictive ability.
 \end{abstract}
-
-% Please add here a significance statement to explain the relevance of your work
-\section{Article Summary}
-$\dots$
 
@@
   are randomly declared to be covered by both, the `complete' and the
-  `incomplete' predictor whereas the remaining $100\% - x\%$ of genotypes
-  are declared to be covered only by the `complete' predictor.
+  `incomplete' predictor (dark gray) whereas the remaining $100\% - x\%$ of
+  genotypes are declared to be covered only by the `complete' predictor (light
+  gray).
   }
@@
 the `full' set for four traits.
-
 A previous study using the same data as in Exp2 concluded that modeling of
@@
 Nevertheless, the inclusion of SCA effects in a single step prediction model
-would be as straightforward as computing the Kronecker product between extension of single step prediction
+would be as straightforward as computing the Kronecker product between the
+feature matrices $\mathbf{M}_{D}$ and $\mathbf{M}_{F}$ of the two corresponding
+heterotic groups~\cite{Martini2016}.
 

### 2017-09-17 paper/manuscript.tex
paper, init cover letter

@@
 \end{abstract}
+
+% Cover Letter
+\section{Cover Letter}
+Dear Managing editors of TAG,
+
+We wish to submit the enclosed manuscript "Omics-based Single Step Prediction"
+by Westhues et al.\@ for publication in TAG.
+
+Recent studies in maize inbred lines~\cite{Guo2016}, maize hybrids
+\cite{Zenke-Philippi2017,Westhues2017}, rice hybrids~\cite{Dan2016,Xu2016} and
+humans~\cite{Vazquez2016} have suggested that `omics' predictors, such as the
+metabolome and the transcriptome, can improve the prediction of traits with
+complex genetic architectures over that achieved with pedigree or genomic data.
+Particularly hybrid breeding, where the success of forecasts increases
+exponentionally with predictive abilities, could benefit from the use of
+predictors that yield more precise estimates of candidate performance.
+Currently, the adoption of these novel predictors by breeders might be impeded
+by their higher sampling costs compared to pedigree and genomic data.
+Here, we present single step prediction of complex traits as a framework for
+combining pedigree and/or genomic data with `omics' predictors collected on a
+subset of all available genotypes to boost predictive abilities while managing
+costs.
+
+Using two extensive phenotypic data sets from a public hybrid silage maize
+breeding program and a maize diversity panel, we a) investigate the success of
+imputing a quantitative predictor with feature heritabilities less than one and
+b) the influence of the fraction of genotypes covered by two predictors on
+predictive abilities in single step prediction.
+
+In our current study we obtained the following results:
+
+\begin{enumerate}
+\item In both data sets, single step prediction improved predictive abilities
+	over those of an individual predictor when the discrepancy between the
+	two predictors in terms of predictive ability were moderate to large in
+	favor of the incomplete predictor.
+\item When 20\%-30\% ($n \approx 25$) of the genotypes were covered by both
+	predictors, single step-based predictive abilities converged to those
+	obtained with the incomplete predictor.
+\end{enumerate}
+
+In conclusion our results demonstrate that single step prediction, using
+an established predictor such as pedigree information, can benefit from sampling
+a highly performant predictor, such as transcriptomic data, on a subset of all
+genotypes.
+The framework presented herein should help promoting the use of attractive
+`omics' predictors by balancing their advantages with their currently higher
+sampling costs.
 

### 2017-09-17 paper/manuscript.tex
paper, fix typos

@@
 
-We wish to submit the enclosed manuscript "Omics-based Single Step Prediction"
+We wish to submit the enclosed manuscript ``Omics-based Single Step Prediction''
 by Westhues et al.\@ for publication in TAG.
@@
 genotypes.
-The framework presented herein should help promoting the use of attractive
+The framework presented herein should help promote the use of attractive
 `omics' predictors by balancing their advantages with their currently higher
@@
 ($n = 211$) given that it is the largest of the four subgroups classified by
-\citeN{Yang2014}.
+\citeN{Guo2016}.
 All inbred lines were evaluated in five Chinese environments described in detail
@@
 Best linear unbiased predictors (BLUPs) were calculated by~\citeN{Yang2014} for
-17 traits~\cite{Guo2016} of which six are presented in this study
-(Table~\ref{table:TraitKey}).
-The selection of these traits was done so that four traits were predicted with
-greater precision by transcriptomic data and two further traits were predicted
+17 traits of which six are presented in this study (Table~\ref{table:TraitKey}).
+The selection of these traits was done so that two traits were predicted with
+greater precision by transcriptomic data and four further traits were predicted
 with greater precision by genomic data.
@@
 complete predictor $\tensor[_{C}]{\mathbf{K}}{}$ and the feature matrix of the
-incomplete predictor $\tensor[_{I}]{\mathbf{W}}{_{22}}$ by calculating
+incomplete predictor $\tensor[_{I}]{\mathbf{W}}{_{2}}$ by calculating
 ($\tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} = \tensor[_C]{\mathbf{K}}{_{12}} \;
@@
 \Genetics2level{Impact of coverage by the incomplete predictor}
-Predictive abilities for the six traits in the the different core sets of Exp1
+Predictive abilities for the six traits in the different core sets of Exp1
 varied widely within most traits (Fig.~\ref{fig:core-results}).
@@
 or T individually was the superior predictor.
-For the traits CW, GW and PH, which were better predicted via G than via T in
-the reduced set of lines, predictive abilities were higher when only a small
-number of genotypes was included in the core set.
+For the traits CW, GW and PH, which were predicted with higher precision via G
+than via T in the reduced set of lines, predictive abilities were higher when
+only a small number of genotypes was included in the core set.
 This corresponds to a situation where a large number of genotypes is being 
@@
 A hike in predictive abilities, when moving from a core fraction of 10\% to 
-90\%, was observed for ED and KW, which were predicted better when using T
-individually compared to using G individually.
+90\%, was observed for ED and KW, which were predicted with greater precision
+when using T individually compared to using G individually.
 The only trait that was barely affected by a change in the core fraction was DS

### 2017-10-08 paper/manuscript.tex
paper, add AEM review from 2017-09-20

@@
 Novel `omics' predictors have shown to improve upon established genome-based
-predictions of some complex traits but are yet expensive to sample.
-In animal breeding, single step prediction was introduced to efficiently use
+predictions of some complex traits but yet require laborious and expensive
+assays.
+In animal breeding, single-step prediction was introduced to efficiently use
 pedigree information, collected on a large number of animals, and genomic
-information, collected on a smaller subset of animals, simultaneously without
-bias.
-Single step prediction has not been attempted with a predictor whose features
-display heritabilities smaller than one, yet.
+information, collected on a smaller subset of animals, simultaneously for breeding
+value estimation without bias.
+Single-step prediction has not yet been attempted with predictors calculated from
+quantitative features displaying heritabilities smaller than one.
 Using two maize datasets of inbred lines and hybrids we show that transcriptomic
@@
 others using pedigree or genomic data could enable breeders to improve trait
-predictions while keeping costs manageable.
-Employing `omics' predictors could particularly benefit candidate selection in
+predictions while keeping costs under control.
+Employing `omics' predictors could particularly improve candidate selection in
 hybrid breeding where the success of forecasts is a strongly convex function of
@@
 predictors that yield more precise estimates of candidate performance.
-Currently, the adoption of these novel predictors by breeders might be impeded
+Currently, the adoption of these novel predictors by breeders is impeded
 by their higher sampling costs compared to pedigree and genomic data.
-Here, we present single step prediction of complex traits as a framework for
-combining pedigree and/or genomic data with `omics' predictors collected on a
-subset of all available genotypes to boost predictive abilities while managing
-costs.
+Here, we present single-step prediction of complex traits as a framework for
+combining pedigree and/or genomic data with other `omics' predictors collected
+on a subset of all available genotypes to boost predictive abilities while
+managing costs.
 
@@
 \begin{enumerate}
-\item In both data sets, single step prediction improved predictive abilities
-	over those of an individual predictor when the discrepancy between the
-	two predictors in terms of predictive ability were moderate to large in
-	favor of the incomplete predictor.
+\item In both data sets, single-step prediction improved predictive abilities
+	over those of the predictor covering all genotypes when the discrepancy 
+	between the	two predictors in terms of predictive ability were moderate to
+	large in favor of the incomplete predictor.
 \item When 20\%-30\% ($n \approx 25$) of the genotypes were covered by both
@@
 
-In conclusion our results demonstrate that single step prediction, using
+In conclusion our results demonstrate that single-step prediction, using
 an established predictor such as pedigree information, can benefit from sampling
@@
 `omics' predictors by balancing their advantages with their currently higher
-sampling costs.
+costs.
 
@@
 animal and plant breeding by a more efficient selection of promising candidates 
-without phenotypes~\cite{DeLosCampos2013,Garcia-Ruiz2016}.
-Its use is largely motivated by the high costs and oftentimes long duration of 
-phenotyping, which can be reduced when all genotypes are equipped with a predictor 
+without phenotypic records~\cite{DeLosCampos2013,Garcia-Ruiz2016}.
+Its use is mainly motivated by the high costs and oftentimes time-consuming
+phenotyping, which can be reduced when all genotypes are covered by a predictor 
 while limiting phenotyping to a few genotypes with high precision
@@
 the availability of one predictor type, which is available for all recorded
-genotypes while other predictors are incomplete in that they cover only a 
-subset of all genotypes~\cite{Fragomeni2015}.
+genotypes with phenotypic information while other predictors are incomplete in
+that they cover only a subset of all genotypes~\cite{Fragomeni2015}.
 Usually, the complete predictor is pedigree information while genomic data are
@@
 The first step consists of conventionally estimating breeding values based on
-pedigree relationships and the phenotypes of relatives.
-Individuals with accurate breeding values out of the first step, which have a
-high number of offspring, enter the second step in which conventional breeding
-values are regressed onto SNP genotypes.
+pedigree relationships among the candidates and phenotypes of their relatives.
+Individuals with higly reliable breeding values out of the first step,
+(\textit{i.e.} those having a high number of offspring) enter the second step in
+which conventional breeding values are regressed onto SNP genotypes.
 The estimated model coefficients can then be used to predict breeding values
@@
 individually~\cite{Aguilar2010}.
-One popular approach to account for this problem is deregressing the breeding
+One popular approach to tackle this problem is deregressing the breeding
 values and weighting the residuals according to the prediction error variances
@@
 % Workings and advantages of the single step procedure.
-Independently from each other,~\citeNP{Legarra2009} and
-\citeNP{Christensen2010} developed a single step BLUP method that blends the 
-numerator relationship matrix $\mathbf{A}$ with the genomic relationship matrix
-$\mathbf{G}$ in a mutual matrix $\mathbf{H}$ to use all predictor information
-simultaneously.
+\citeNP{Legarra2009} and \citeNP{Christensen2010} independently from each other
+developed a single-step BLUP method that blends the numerator relationship matrix
+$\mathbf{A}$ with the genomic relationship matrix $\mathbf{G}$ in a mutual matrix
+$\mathbf{H}$ to use all predictor information simultaneously.
 Issues with this H-BLUP approach were the compatability of $\mathbf{A}$ with
@@
 trait-independent~\cite{Vitezica2011,Ashraf2016}.
-\citeN{Fernando2014} derived the equivalent single step marker-effect-model to 
-the single step breeding value model of \citeNP{Legarra2009} and 
+\citeN{Fernando2014} derived the equivalent single-step marker-effect-model to 
+the single-step breeding value model of \citeNP{Legarra2009} and 
 \citeNP{Christensen2010}, which does not require weights for $\mathbf{A}$ and 
@@
 % -   Hybrids
-In animal breeding, single step prediction, particularly H-BLUP, has already 
-become a routine procedure~\cite{Legarra2014} but, to the best of our knowledge, 
-has only been considered in plants in a study on wheat~\cite{Ashraf2016} and
-white spruce lines, respectively~\cite{Ratcliffe2017}.
-Moreover, while an application of the single step framework to animals exists
-for pigs~\cite{Xiang2016}, it is yet lacking for plant hybrids.
-Technological advances in plant hybrid breeding such as the rapid generation
-of completely homozygous lines~\cite{Wedzony2009} now enable breeders to
-produce about $n = 1,000$ individuals per generation and sex.
+In animal breeding, single-step prediction, particularly H-BLUP, has already 
+become a routine procedure~\cite{Legarra2014} for predicting breeding values
+but, to the best of our knowledge, has only been considered in plants in a study
+on wheat~\cite{Ashraf2016} and white spruce lines, respectively
+\cite{Ratcliffe2017}.
+Moreover, while an extension of the single-step framework to hybrid genotypes
+was developed for pigs~\cite{Xiang2016}, it is yet lacking for the prediction
+of single-cross hybrids between purely inbred lines.
+Technological advances in hybrid breeding of several crop species, such as the
+rapid generation of completely homozygous lines~\cite{Wedzony2009}, now enable
+plant breeders to produce about $n = 1,000$ individuals per generation and sex.
 Assuming that about 90\% of these lines were never before tested in any hybrid
@@
 \cite{Westhues2017}.
-Predicting these hybrid candidates with low relatedness to tested hybrids using
-only information from their parents is currently a hot topic in plant breeding
-\cite{Technow2014,Kadam2016,Westhues2017}.
+Predicting the performance of these hybrid candidates with low relatedness to 
+tested hybrids using only information from their parents is currently a hot topic
+in plant breeding~\cite{Technow2014,Kadam2016,Westhues2017}.
 
@@
 situation in animal breeding, previous studies have focused exclusively on the
-application of single step prediction to these predictors.
+application of single-step prediction to these predictors.
 Recent investigations into the utility of `omics' data for boosting predictive
-abilities in plants have suggested that metabolites and transcriptomic data can
+abilities in plants have suggested that metabolic and transcriptomic data can
 improve upon predictive abilities based on pedigree or genomic data
 \cite{Guo2016,Xu2016,Dan2016,Zenke-Philippi2017,Westhues2017}.
-Yet, these technologies are not fully mature and therefore incur higher costs
-to breeders than pedigree and genomic data, putting their utility into question
-despite their proven usefulness.
+Yet, these two `omics' technologies are not fully mature and therefore incur
+higher costs to breeders than pedigree and genomic data, putting their utility
+into question despite their proven usefulness.
 
 % - Objectives
-Hence, our objectives were to i) examine whether the single step prediction 
-framework can be transfered to a situation where a very large number of hybrids
-($\approx$ 1 million) needs to be predicted and ii) evaluate the impact of
-covering only a subset of all genotypes with both, the complete and the
-incomplete predictor on predictive ability, where for the remaining individuals
-only genomic data are available.
+Hence, our objectives were to (i) extend the existing theoretical framework for
+single-step prediction to the imputation of quantitative `omics' features, here
+exemplified by transcriptomic data, (ii) evaluate the impact of the proportion of
+imputed genotypes in the training set on the predictive ability and (iii) transfer
+the single-step framework for breeding value prediction to the prediction of
+hybrids.
 
@@
 \Genetics2level{Exp1}
-To address the second objective a maize inbred line data set, subsequently
-referred to as Exp1, was used~\cite{Yang2014}.
-Exp1 originally comprised 513 maize lines representing the global maize
+To address objectives (i) and (ii) a published data set of maize inbred lines
+\cite{Yang2014} was used.
+It originally comprised 513 maize lines representing the global maize
 diversity  and was reduced to the set of tropical and subtropical lines
-($n = 211$) given that it is the largest of the four subgroups classified by
-\citeN{Guo2016}.
-All inbred lines were evaluated in five Chinese environments described in detail
-by~\citeNP{Yang2014}.
+($n = 211$), subsequently referred to as Exp1, which is the largest of the four
+subgroups classified by~\citeN{Guo2016}.
+All inbred lines were evaluated in five different environments in China
+described in detail by~\citeNP{Yang2014}.
 Best linear unbiased predictors (BLUPs) were calculated by~\citeN{Yang2014} for
-17 traits of which six are presented in this study (Table~\ref{table:TraitKey}).
-The selection of these traits was done so that two traits were predicted with
-greater precision by transcriptomic data and four further traits were predicted
-with greater precision by genomic data.
-MaizeSNP50 BeadChip-data were available for the entire set of 513 lines from the
-diversity panel.
-Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
-days after sowing.
-By exploiting identity by descent (IBD) between 49,728 SNPs on the BeadChip 
-overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
-for these lines~\cite{Fu2013,Li2013}.
-For the remaining 145 maize lines without RNA-seq data, high density markers
-were inferred via projection of IBD regions onto the BeadChip, using this core 
-set of markers~\cite{Yang2014}.
+all 211 inbred lines and for 17 traits of which six are presented in this study 
+(Table~\ref{table:TraitKey}).
+The selection of these traits was done based on our prediction results so that
+two traits were predicted with greater precision by transcriptomic data and four
+further traits were predicted with greater precision by genomic data.
 After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
 ($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed
-for 37,760 SNPs using the Beagle software~\cite{Browning2009}.
-Gene expression data were normalized using a normal quantile transformation to
-satisfy modeling assumptions~\cite{Fu2013}.
+using the Beagle software~\cite{Browning2009} resulting in 37,760 SNPs for these
+211 genotypes.
 A subset of 28,850 annotated genes was available for 149 out of the 211
@@
 
+\subsection{Population structure}
+Population structure in Exp1 was examined through a STRUCTURE
+\cite{Pritchard2000} analysis with the number of putative ancestral
+populations $K$ ranging from 2 to 9.
+Based on the estimated likelihood of the data for each $K$, we chose $K=4$
+and carried out a principal component analysis (PCA) on the scaled and centered
+genomic feature matrix using the \emph{pca} function from the \emph{R}-package
+\emph{LEA}~\cite{Frichot2015}.
+Least-squares estimates of ancestry proportions were estimated based on sparse 
+nonnegative matrix factorization algorithms~\cite{Frichot2014} using the 
+\emph{snmf} function in \emph{LEA}.
+Genotypes in the PCA plot (Fig.~\ref{fig:inbred-pca}) were colorized based on their
+primary assignment to any of the four putative ancestral populations.
+To mitigate the possibility of inflated predictive abilities, which can arise
+from large differences in the average performance of the different
+subpopulations~\cite{Windhausen2012}, we reduced the panel of inbred lines to
+164 genotypes, which were assigned to the largely overlapping clusters 1 and 4 in
+Fig.~\ref{fig:inbred-pca}.
+In summary, genomic and phenotypic information were available for 164 inbred
+lines, which we denote as the `Full' set whereas transcriptomic data were available
+for a subset of 110 inbred lines, which we denote as `Reduced'.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics{./tables_figures/inbred_pca.pdf}
+  \caption{
+  Principal component analysis (PCA) of 211 genotypes from the set
+  `tropical/subtropical' of the maize diversity panel inbred lines.
+  The four clusters are based on $K=4$ putative ancestral populations from a
+  STRUCTURE analysis of the corresponding genomic feature matrix.
+  Genotypes are colored based on their primary assignment to any of these four
+  clusters.
+  The explained variance for each of the two PCs is given in the corresponding
+  axis labels.
+}
+\label{fig:inbred-pca}
+\end{figure}
+
 \Genetics2level{Exp2}
-To address objectives 1 and 2, a maize hybrid data set, subsequently
+To address objectives (i) and (iii), a maize hybrid data set, subsequently
 denoted as Exp2, was used.
@@
 
-\Genetics2level{Population structure}
-Population structure in Exp1 was examined through a STRUCTURE
-\cite{Pritchard2000} analysis with the number of putative ancestral
-populations $K$ ranging from 2 to 9.
-Based on the estimated probability of the data for each $K$, we chose $K=4$
-and carried out a principal component analysis (PCA) on the scaled and centered
-genomic feature matrix using the \emph{pca} function from the \emph{R}-package
-\emph{LEA}~\cite{Frichot2015}.
-Least-squares estimates of ancestry proportions were estimated based on sparse 
-nonnegative matrix factorization algorithms~\cite{Frichot2014} using the 
-\emph{snmf} function in \emph{LEA}.
-Genotypes in the PCA plot (Fig.~\ref{fig:inbred-pca}) were colorized based on their
-primary assignment to any of the four putative ancestral populations.
-To mitigate the possibility of inflated predictive abilities, which can arise
-from large differences in the average performance of the different
-subpopulations~\cite{Windhausen2012}, we reduced the panel of inbred lines to
-those assigned to the largely overlapping clusters 1 and 4 in
-Fig.~\ref{fig:inbred-pca}.
-This reduction yielded 164 genotypes, 110 of which were covered by both, genomic
-and transcriptomic data.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics{./tables_figures/inbred_pca.pdf}
-  \caption{
-  Principal component analysis (PCA) of 211 genotypes from the set
-  `tropical/subtropical' of the maize diversity panel inbred lines.
-  The four clusters are based on $K=4$ putative ancestral populations from a
-  STRUCTURE analysis of the corresponding genomic feature matrix.
-  Genotypes are colored based on their primary assignment to any of these four
-  clusters.
-  The explained variance for each of the two PCs is given in the corresponding
-  axis labels.
-}
-\label{fig:inbred-pca}
-\end{figure}
-
 \Genetics2level{Prediction}
@@
 
-where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
-respectively, $\mu$ is the fixed model intercept, and $\mathbf{e}$ is the 
-residual error.
-For the inbred lines we fit random genotype effects $\mathbf{g}$.
-For both experiments, genetic effects $\mathbf{g}$ are associated with the
-vector of phenotypic performance $\mathbf{y}$ through a corresponding design
-matrix $\mathbf{Z}$.
+where $\mathbf{y}$ is the vector of observed inbred line performance $\mu$ is
+the fixed model intercept, and $\mathbf{e}$ is the residual error.
+For the inbred lines we fit random genotype effects $\mathbf{g}$, which are
+associated with the vector of phenotypic performance $\mathbf{y}$ through a
+corresponding design matrix $\mathbf{Z}$.
 The random effects ($\mathbf{g}$) have expectation zero and variance equal to
 $\mathbf{K} \sigma^{2}_{g}$ and $\mathbf{I} \sigma^2_{e}$ for the residual
-error.
+error, respectively.
 Here, we did not consider the inclusion of dominance effects for the hybrid
@@
 
-\subsection{Single step prediction}
+\subsection{Single-step prediction}
 If the breeding values are fully captured by the features in $\mathbf{M}$,
-Eq.~\ref{eq:KBLUPModel} can be expressed as
+Eq.~\ref{eq:KBLUPModel} can be expressed as~\cite{Fernando2014}
 
@@
 
-\cite{Fernando2014} because the ridge regression model is equivalent to the
-BLUP model~\cite{Ruppert2003}.
-In the previous equation, $\mathbf{M}$ is a matrix of feature covariates and
-$\boldsymbol{\alpha}$ is the vector of random, partial-regression coefficients
-of the feature covariates.
-
-Hence, the breeding values in a prediction model are given by:
+with $\boldsymbol{\alpha}$ being the vector of random, partial-regression
+coefficients of the feature covariates, because the ridge regression model is
+equivalent to the BLUP model \cite{Ruppert2003}.
+
+Hence, the breeding values from the prediction model are given by:
 \begin{equation} \label{eq:mrnaebv}
@@
 Consider now the situation in which one predictor is complete in the sense that
-the whole population is equipped with features for that predictor whereas only
-a subset of that population has information on a second predictor.
+the whole population is covered by features for that predictor but only a subset
+of that population has information for a second predictor.
 We can take a similar route as in \citeNP{Fernando2014} and impute covariates 
-of the incomplete predictor by using covariates of the complete predictor.
-Let the subscript $1$ denote individuals covered only by the complete
-predictor whereas individuals covered by both, the complete and the
-incomplete predictor, are indicated by the subscript $2$.
-Let the index $C$ denote matrices and vectors pertaining to the complete
-predictor and let the index $I$ denote matrices and vectors pertaining to the
-incomplete predictor, respectively.
-
-Let us further denote the matrix of centered covariates of the incomplete
-predictor as
+of the incomplete predictor, denoted as `predictor' and indicated through the
+asterisk symbol `*', by using covariates of the complete predictor, denoted as
+`imputor'.
+Let the subscript $1$ denote individuals covered only by the predictor whereas
+individuals covered by both, the imputor and the predictor, are indicated by the subscript $2$.
+
+Let us further denote the matrix of centered covariates of the predictor as
 \begin{equation} \label{eq:incomplete-covariates}
-\tensor[_{I}]{\mathbf{W}}{} =
-\begin{bmatrix}
-  \mathbf{Z}_{1} \; \tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} \\
-  \tensor[_I]{\mathbf{W}}{_{2}}
+\tensor[]{\mathbf{W}}{^{*}} =
+\begin{bmatrix}
+  \mathbf{Z}_{1} \; \tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} \\
+  \tensor[]{\mathbf{W}}{^{*}_{2}}
  \end{bmatrix},
 \end{equation}
-where observed feature covariates are stored in $\tensor[_I]{\mathbf{W}}{_{2}}$
-while the feature covariates for genotypes not covered by the incomplete
-predictor can be estimated/imputed with the aid of the covariance matrix of the
-complete predictor $\tensor[_{C}]{\mathbf{K}}{}$ and the feature matrix of the
-incomplete predictor $\tensor[_{I}]{\mathbf{W}}{_{2}}$ by calculating
-($\tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} = \tensor[_C]{\mathbf{K}}{_{12}} \;
-\tensor[_C]{\mathbf{K}}{_{22}^{-1}} \; \tensor[_I]{\mathbf{W}}{_{2}}$).
-In this case we need to partition the vector of genetic values $\mathbf{g}$
-into a component $\mathbf{g}_{1}$ (\textit{i.e.,} genetic values of genotypes
-covered only by the complete predictor) and a component
-$\mathbf{g}_{2} = \tensor[_{I}]{\mathbf{W}}{_{2}} \; 
-\tensor[_{I}]{\boldsymbol{\alpha}}{}$ (\textit{i.e.,} genetic
-values of genotypes covered by both predictors where $\tensor[_{I}]{\boldsymbol{\alpha}}{}$
-pertains to the random feature effects of the incomplete predictor).
-Under multivariate normality, the vector $\mathbf{g}_1$ can be written as the
-sum of the conditional expectation given $\mathbf{g}_2$ and a residual error
-term:
+where observed feature covariates are stored in $\tensor[]{\mathbf{W}}{^{*}_{2}}$
+while the feature covariates for genotypes not covered by the predictor can be estimated/imputed with the aid of the covariance matrix of the
+imputor $\mathbf{K}$ and the feature matrix of the predictor 
+$\tensor[]{\mathbf{W}}{^{*}_{2}}$ by calculating
+($\tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} = \tensor[]{\mathbf{K}}{_{12}} \;
+\tensor[]{\mathbf{K}}{_{22}^{-1}} \; \tensor[]{\mathbf{W}}{^{*}_{2}}$).
+In this case we need to partition the vector of breeding values $\mathbf{g}$
+into a component $\mathbf{g}_{1}$ (\textit{i.e.,} breeding values of genotypes
+covered only by the imputor) and a component
+$\mathbf{g}_{2} = \tensor[]{\mathbf{W}}{^{*}_{2}} \; 
+\tensor[]{\boldsymbol{\alpha}}{^{*}}$ (\textit{i.e.,} genetic
+values of genotypes covered by both predictors, where
+$\tensor[]{\boldsymbol{\alpha}}{^{*}}$ pertains to the random feature effects of
+the predictor).
+Features are distributed according to a multivariate normal distribution:
 \begin{align} \label{eq:mrna1}
@@
 	\boldsymbol{\epsilon} \\
-	&= \tensor[_{C}]{\mathbf{K}}{_{12}} \;
-       	\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-       	\tensor[_{I}]{\mathbf{W}}{_{2}} \;
-	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}}{} +
-       	(\mathbf{g_1} - \tensor[_{C}]{\mathbf{K}}{_{12}} \;
-	\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-       	\tensor[_{I}]{\mathbf{W}}{_{2}} \;
-	\tensor[_{I}]{\boldsymbol{\hat{\alpha}}}{}) \\
+	&= \tensor[]{\mathbf{K}}{_{12}} \;
+       	\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
+       	\tensor[]{\mathbf{W}}{^{*}_{2}} \;
+	\tensor[]{\boldsymbol{\hat{\alpha}}}{^{*}} +
+       	(\mathbf{g_1} - \tensor[]{\mathbf{K}}{_{12}} \;
+	\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
+       	\tensor[]{\mathbf{W}}{^{*}_{2}} \;
+	\tensor[]{\boldsymbol{\hat{\alpha}}}{^{*}}) \\
 	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
@@
 
-where $\tensor[_{C}]{\mathbf{K}}{_{ij}}$ are partitions of the matrix
-$\tensor[_{C}]{\mathbf{K}}{}$ that correspond to $\mathbf{g}_{1}$ and
-$\mathbf{g}_{2}$, respectively~\cite{Fernando2014} and where
-$\mathbf{\hat{g}}_{1}$ pertains to the imputed genetic values of genotypes
-not equipped with the incomplete predictor.
+where the first summand pertains to $\mathbf{\hat{g}}_{1}$, the second summand
+pertains to $\boldsymbol{\epsilon}$ and where the partitions of the matrix
+\begin{equation} \label{eq:imputor}
+\tensor[]{\mathbf{K}}{} =
+\begin{bmatrix}
+  \mathbf{K}_{11} & \mathbf{K}_{12} \\
+  \mathbf{K}_{21} & \tensor[]{\mathbf{K}_{22}}{^{*}}
+ \end{bmatrix},
+\end{equation}
+
+correspond to $\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively
+\cite{Fernando2014} and $\mathbf{\hat{g}}_{1}$ pertains to breeding values
+estimated by imputing the predictor.
 
 The covariance matrix of $\boldsymbol{\epsilon}$ is
-$\tensor[_{C}]{\mathbf{K}}{_{11}} -
-\tensor[_{C}]{\mathbf{K}}{_{12}} \;
-\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-\tensor[_{C}]{\mathbf{K}}{_{21}} =
-(\tensor[_{C}]{\mathbf{K}}{^{11}})^{-1}$
+$\tensor[]{\mathbf{K}}{_{11}} -
+\tensor[]{\mathbf{K}}{_{12}} \;
+\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
+\tensor[]{\mathbf{K}}{_{21}} =
+(\tensor[]{\mathbf{K}}{^{11}})^{-1}$
 \cite{Legarra2009}
@@
 and the covariance of $\mathbf{\hat{g}}_{1}$ is
-$\tensor[_{C}]{\mathbf{K}}{_{12}} \;
-\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-\tensor[_{I}]{\mathbf{W}}{_{2}} \;
-\tensor[_{I}]{\mathbf{W}}{_{2}^{\top}} \;
-\tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-\tensor[_{C}]{\mathbf{K}}{_{21}} \;
-\tensor[_{I}]{\sigma^{2}_{\alpha}}{}$
+$\tensor[]{\mathbf{K}}{_{12}} \;
+\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
+\tensor[]{\mathbf{W}}{^{*}_{2}} \;
+\tensor[]{\mathbf{W}}{^{*}_{2}^{\top}} \;
+\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
+\tensor[]{\mathbf{K}}{_{21}} \;
+\tensor[]{\sigma^{2}_{\alpha}}{^{*}}$
 \cite{Fernando2014}.
 
-The structure of the residual imputation/prediction error is known and,
-therefore, can modeled.
-
-The general model for the single step procedure can then be written as
+The structure of the residual imputation/prediction error is known, and
+therefore, can be modeled.
+
+The general model for the single-step procedure can then be written as
 
@@
 \mathbf{Xb} +
-\tensor[_{I}]{\mathbf{W}}{} \;
-\tensor[_{I}]{\boldsymbol{\alpha}}{} +
+\tensor[]{\mathbf{W}}{^{*}} \;
+\tensor[]{\boldsymbol{\alpha}}{^{*}} +
 \mathbf{U} \boldsymbol{\epsilon} +
@@
 \begin{bmatrix}
-  \tensor[_{C}]{\mathbf{K}}{_{12}} \; \tensor[_{C}]{\mathbf{K}}{_{22}^{-1}} \;
-  \tensor[_{I}]{\mathbf{W}}{_{2}} \;
-  \tensor[_{I}]{\boldsymbol{\alpha}}{} + \boldsymbol{\epsilon}  \\
-  \tensor[_{I}]{\mathbf{W}}{_{2}} \; \tensor[_{I}]{\boldsymbol{\alpha}} \\
+  \tensor[]{\mathbf{K}}{_{12}} \; \tensor[]{\mathbf{K}}{_{22}^{-1}} \;
+  \tensor[]{\mathbf{W}}{^{*}_{2}} \;
+  \tensor[]{\boldsymbol{\alpha}}{^{*}} + \boldsymbol{\epsilon}  \\
+  \tensor[]{\mathbf{W}}{^{*}_{2}} \; \tensor[]{\boldsymbol{\alpha}}{^{*}} \\
  \end{bmatrix}
@@
 \begin{bmatrix}
-  \tensor[_{I}]{\mathbf{\hat{W}}}{_{1}} \;
-  \tensor[_{I}]{\boldsymbol{\alpha}}{} + \boldsymbol{\epsilon} \\
-  \tensor[_{I}]{\mathbf{W}}{_{2}} \; \tensor[_{I}]{\boldsymbol{\alpha}} \\
+  \tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} \;
+  \tensor[]{\boldsymbol{\alpha}}{^{*}} + \boldsymbol{\epsilon} \\
+  \tensor[]{\mathbf{W}}{^{*}_{2}} \; \tensor[]{\boldsymbol{\alpha}}{^{*}} \\
  \end{bmatrix}
@@
 
-The model above indicates that single step prediction models ultimately use
+The model above indicates that single-step prediction models ultimately use
 only the matrix of known and imputed feature covariates $\mathbf{W}$ for the
@@
 
-Our final model is now:
+Our final model for the predictor is now:
 
@@
 \mathbf{Z} \;
-\tensor[_{I}]{\mathbf{W}}{} \;
-\tensor[_{I}]{\boldsymbol{\alpha}}{} +
+\tensor[]{\mathbf{W}}{^{*}} \;
+\tensor[]{\boldsymbol{\alpha}}{^{*}} +
 \mathbf{Z}\mathbf{U} \boldsymbol{\epsilon} +
@@
  \begin{bmatrix}
-  \tensor[_{I}]{\hat{\mathbf{W}}}{_{1}} \\
-  \tensor[_{I}]{\mathbf{W}}{_{2}}
- \end{bmatrix}
- \tensor[_{I}]{\hat{\boldsymbol{\alpha}}}{}
+  \tensor[]{\hat{\mathbf{W}}}{^{*}_{1}} \\
+  \tensor[]{\mathbf{W}}{^{*}_{2}}
+ \end{bmatrix}
+ \tensor[]{\hat{\boldsymbol{\alpha}}}{^{*}}
  +
@@
 $\mathbf{M}_{D}$ for Dent and $\mathbf{M}_{F}$ for Flint parent lines in Exp2.
-The general model for the calculation of genetic values in Exp2 is then
+The general model for the calculation of genetic values of hybrids in Exp2 is
+then
 
@@
 using a single genotype as a hold-out sample, which will be predicted by using
-all other inbred lines for model training.
-This process is repeated until all inbred lines have been used once for testing
-and $n - 1$ times for model training.
+all other $164 - 1$ inbred lines for model training.
+This process is repeated until all $164$ inbred lines have been used once for
+testing and $164 - 1$ times for model training.
+For the single-step prediction models, imputation was performed on the
+transcriptomic data.
 
@@
 $H_{TRN}(i \times j) = H \cap (D_{i} \times F_{j})$.
+For the single-step prediction models, imputation was performed on either the
+genomic or the transcriptomic data.
 
@@
 The base of each of these core sets was the set of 110 genotypes that were
-covered by both, genomic and trancriptomic data.
+covered by both, genomic and transcriptomic data.
 A core set was created by artificially declaring a pre-specified fraction 
@@
 In terms of predictive ability, the absolute advantage of T over G was as large 
-as 0.15 as observed for the trait ED\@.
+as 0.15, observed for the trait ED\@.
 The predictive ability for predictions based on G were always higher for the
@@
 ($n = 110$).
-The combination of G and T (\textit{i.e.}, $\tensor[_{C}]{G}{}
-\tensor[_{I}]{T}{}$, where the subscripts correspond to the complete ($C$) and
-the incomplete ($I$) predictor, respectively) in a single-step approach for the 
-prediction of all 164 lines, performed at least as well as G alone when T was 
-superior over G for the reduced set of 110 lines.
-In the case of KW and ED, the use of $\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$
-yielded a slight improvement in predictive ability over G.
+The combination of G and T in a single step (\textit{i.e.}, $G|T^{*}$, where
+$G$ is the imputor and $T^{*}$ is the predictor, for the prediction of all 164
+lines, performed at least as well as G alone when T was superior over G for the
+reduced set of 110 lines.
+In the case of KW and ED, the use of $G|T^{*}$ yielded a slight improvement in
+predictive ability over G.
 Where T was a worse predictor for the reduced set of 110 lines than G,
-predictive abilities of $\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$ for the full set
-of 164 lines were also worse than those based on G alone.
+predictive abilities of $G|T^{*}$ for the full set of 164 lines were also worse
+than those based on G alone.
 
@@
 \Genetics2level{Exp2}
-The maize hybrid data set encompasses seven agronomic traits and three
-predictors, namely pedigree data (P), genomic information (G), transcriptomic
-information (T), and combinations thereof.
-Pedigree records and genomic information were available for the 245 parent 
-lines of all 1,521 hybrids and we denote these data as the `Full' set whereas
-transcriptomic data were available for 103 parent lines of 685 hybrids, which
-we denote as `Reduced'.
-In the `Reduced' set, transcriptomic data were the best single predictor for
-ADL, PRO and particulaly for DMY where the predictive ability obtained with T
-was 0.08 points higher than for G as the second best predictor for this trait
-(Fig.~\ref{fig:HybridResults}).
+In the `Reduced' set of Exp2, transcriptomic data were the best single predictor
+for the traits ADL, PRO and particulaly for DMY where the predictive ability
+obtained with T was 0.08 points higher than for G as the second best predictor for
+this trait (Fig.~\ref{fig:HybridResults}).
 In the reduced set, P had the highest predictive ability for STA and DMC\@.
@@
 markedly for all traits with the exception of FAT\@.
-The combination of genomic and transriptomic information
-($\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$) in a single-step model yielded a small
-improvement over G alone for the traits DMY and PRO\@.
+The single-step combination of genomic and transriptomic information
+($G|T^{*}$) yielded a small improvement over G alone for the traits DMY and PRO\@.
 Except for PRO, compared to using only G, trait prediction did not benefit from 
-the single step approach when T was not superior to G in the reduced set.
+the single-step approach when T was not superior to G in the reduced set.
 Predictive abilities resulting from combinations of pedigree with genomic
-($\tensor[_{C}]{P}{}\tensor[_{I}]{G}{}$) information were better than P alone
-for all traits and $\tensor[_{C}]{P}{}\tensor[_{I}]{T}{}$ improved upon P in
-the `full' set for four traits.
-A previous study using the same data as in Exp2 concluded that modeling of
-specific combining ability (SCA) effects did not appreciably improve predictive
-abilities over that of a purely additive model for any of the investigated
-traits.
-Nevertheless, the inclusion of SCA effects in a single step prediction model
-would be as straightforward as computing the Kronecker product between the
-feature matrices $\mathbf{M}_{D}$ and $\mathbf{M}_{F}$ of the two corresponding
-heterotic groups~\cite{Martini2016}.
+($P|G^{*}$) information were better than P alone for all traits and $P|T^{*}$
+improved upon P in the `full' set for four traits.
 
@@
 varied widely within most traits (Fig.~\ref{fig:core-results}).
-For most traits, an initial increase in predictive abilities when moving
-from a core set size of 10\% to 20\%/30\% was observed, regardless of whether G
-or T individually was the superior predictor.
+For the majority of traits, an initial increase in predictive abilities when
+moving from a core set size of 10\% to 20\%/30\% was observed, regardless of
+whether G or T individually was the superior predictor.
 For the traits CW, GW and PH, which were predicted with higher precision via G
@@
   Predictive abilities for core fractions between 0.1 and 0.9 are based on
-  transcriptomic effects derived from a single step prediction model involving
+  transcriptomic effects derived from a single-step prediction model involving
   transcriptomic (T) and genomic (G) data, respectively.
@@
 \Genetics2level{Advantages of marker effect models}
-Single step approaches in hybrid populations have so far been limited to pig
+Single-step approaches in hybrid populations have so far been limited to pig
 breeding~\cite{Xiang2015,Xiang2016,Tusell2016}.
-All of these studies used breeding value-based single step BLUP methods
+All of these studies used breeding value-based single-step BLUP methods
 (SS-BLUP) conceived by \citeNP{Legarra2009} and \citeNP{Christensen2010} with
 extensions to hybrid populations~\cite{Christensen2014,Christensen2015}.
-An attactive property of SS-BLUP models is their computational advantage over 
-marker-effect single step Bayesian regression models (SSBR) when the number of 
+An attractive property of SS-BLUP models is their computational advantage over 
+marker-effect single-step Bayesian regression models (SSBR) when the number of 
 genotypes is smaller than the number of features of the incomplete predictor.
@@
 Advantages of the SSBR algorithm, conceived by \citeNP{Fernando2014}, include
-that i) it does not require commensurability of $\mathbf{A}$ and $\mathbf{G}$, 
-ii) it models the bias incurred by the incomplete predictor explicitly as 
-$\mathbf{\epsilon}$ and iii) its speed depends largely on the number of 
+that (i) it does not require commensurability of $\mathbf{A}$ and $\mathbf{G}$, 
+(ii) it models the bias incurred by the incomplete predictor explicitly as 
+$\mathbf{\epsilon}$ and (iii) its speed depends largely on the number of 
 features.
@@
 Compared to pedigree information, which represent the expected relationship
-among individuals, genomic information capture Mendelian sampling and thereby 
-provide an improved proxy of the realized relationship among individuals.
+among individuals, genomic information capture Mendelian sampling as well as the
+effects of selection and thereby provide an improved proxy of the realized
+relationship among individuals.
 Nevertheless, genomic data do not exhaustively capture physiological epistasis 
@@
 maize inbred lines~\cite{Guo2016} and maize hybrids~\cite{Westhues2017}.
-While single step prediction has become an established method when complete
+While single-step prediction has become an established method when complete
 pedigree but only incomplete genomic records are available, its application has
-not yet been considered for the imputation of a complete predictor based on 
-incomplete downstream `omics' predictors such as transcriptomic, proteomic or 
-metabolomic data.
-While the heritability of relationship coefficients and genomic markers can be
-assumed to be equal to 1, this is not the case for the aforementioned
-quantitative `omics' predictors.
-Here, we considered single step approaches for a maize inbred line diversity
+not yet been considered for the imputation of downstream `omics' predictors
+such as transcriptomic, proteomic or metabolomic data.
+Here, we considered single-step approaches for a maize inbred line diversity
 panel as well as for a collection of maize hybrids using pedigree, genomic and
 transcriptomic data as predictors.
-
-\Genetics2level{Superiority of single step models}
+While the repeatability of relationship coefficients and genomic markers can be
+assumed to be equal to 1, this is not the case for the aforementioned
+quantitative `omics' predictors, which would negatively affect the precision of
+the imputation and was not considered here.
+
+\Genetics2level{Superiority of single-step models}
 \subsection{Exp1}
-When using a single step approach based on complete genomic and incomplete
+When using a single-step approach based on complete genomic and incomplete
 transcriptomic information, gains in predictive ability for the inbred lines
@@
 for about 10,000 wheat lines using pedigree information and observed greater
-prediction accuracy of the single step method over genomic BLUP for all four
+prediction accuracy of the single-step method over genomic BLUP for all four
 evaluated traits.
@@
 possible reasons:
-1) We reduced the dataset to 164 tropical/subtropical lines with little evident
-population structure and excluded all other genotypes from our analyses whereas
-\citeN{Guo2016} used genotypes from four different subpopulations with a model
-using a fixed effect to account for the differences between subpopulations.
-2) We applied quality checks for the predictor data after generating the subset
-of 164 inbred lines, whereas~\citeN{Guo2016} applied the quality checks to 368
-genotypes.
-3) While~\citeN{Guo2016} used five-fold cross-validation with 500 repetitions,
+First, we reduced the dataset to 164 tropical/subtropical lines with little
+evident population structure and excluded all other genotypes from our analyses
+whereas \citeN{Guo2016} used genotypes from four different subpopulations with a 
+model using a fixed effect to account for the differences between subpopulations.
+Second, we applied quality checks for the predictor data after generating the
+subset of 164 inbred lines, whereas~\citeN{Guo2016} applied the quality checks to
+368 genotypes.
+Third, while~\citeN{Guo2016} used five-fold cross-validation with 500 repetitions,
 we employed a leave-one-out cross-validation scheme (LOOCV).
@@
 \subsection{Exp2}
-Hybrid breeding --- with noteworthy commercial applications in pig
-\cite{Xiang2016,Tusell2016} and maize breeding~\cite{TheRoyalSociety2009} --- 
-is a particularly challenging application field for prediction tools 
-\cite{Kadam2016}.
+Hybrid breeding is a particularly challenging application field for prediction
+tools \cite{Kadam2016}.
 Here, $2n$ parent individuals from two genetically distinct heterotic groups are
@@
 material well.
-For the hybrid data, the predictive abilities of genomic information
-increased considerably in the full set of genotypes compared to the reduced
-set for several agronomic traits.
+For the hybrid data, the predictive abilities of G increased considerably in the
+full set of genotypes compared to the reduced set for several agronomic traits.
 Hence, the inclusion of trancsriptomic data could not further improve upon
@@
 either transcriptomic or genomic information.
-This suggests that single step prediction should always be considered when the
+This suggests that single-step prediction should always be considered when the
 incomplete predictor offers an appreciably higher predictive ability compared
 to the complete predictor.
+A previous study using the same data as in Exp2 concluded that modeling of
+specific combining ability (SCA) effects did not appreciably improve predictive
+abilities over that of a purely additive model for any of the investigated
+traits~\cite{Westhues2017}.
+Nevertheless, the inclusion of SCA effects in a single-step prediction model
+would be as straightforward as computing the element-wise product between the
+feature matrices $\mathbf{M}_{D}$ and $\mathbf{M}_{F}$ of the two corresponding
+heterotic groups~\cite{Martini2016}.
 
@@
 be done at random once more than 10,000 animals have pedigree records 
-\cite{Fragomeni2015,Lourenco2015,Masuda2016}, pedigrees in plant breeding 
+\cite{Fragomeni2015,Lourenco2015,Masuda2016}, population sizes in plant breeding 
 programs are typically considerably smaller.
-Single step prediction offers the promise of leveraging the predictive ability
+Single-step prediction offers the promise of leveraging the predictive ability
 for all genotypes by borrowing the superior performance of an auxiliary
@@
 complemented with information on another predictor.
-By randomly and repeatedly declaring trancsriptomic information missing from
-genotypes in Exp1 we examined how reliably single step prediction works
+By randomly and repeatedly declaring transcriptomic information missing from
+genotypes in Exp1 we examined how reliably single-step prediction works
 depending on the fraction of genotypes covered by both, the complete and the
 incomplete predictor.
-We observed that about $x=20-30\%$ of genotypes needed to be covered by both
+We observed that about $x=20-30\%$ of genotypes need to be covered by both
 predictors to match the performance achieved with the complete predictor
@@
 predictor.
-This pattern clearly elucidates the influence on the difference between the
-two predictors on predictive ability in a single step approach.
+This pattern clearly elucidates the influence of the difference between the
+two predictors on predictive ability in a single-step approach.
 
 \section*{Conclusions}
-We successfully applied single step prediction to an inbred and a hybrid plant
-data sets while imputing with a quantitative downstream `omics' predictor.
+We successfully applied single-step prediction to inbred and hybrid data sets
+while imputing with a quantitative downstream `omics' predictor.
 By declaring different subsets of individuals as covered by one or two 
 predictors, respectively, we could elucidate the influence of differences
-between two predictors on predictive ability when used in a single step
+between two predictors on predictive ability when using a single-step
 approach.
@@
 incomplete predictor was skewed profoundly in favor of the incomplete predictor,
-such as for pedigree and transcriptomic data, single step prediction was shown
+such as for pedigree and transcriptomic data, single-step prediction was shown
 to be highly beneficial.

### 2017-10-16 paper/manuscript.tex
paper, edits from meeting on 2017_10_12

@@
 %% THE MANUSCRIPT TITLE
-\title{Omics-based Single Step Trait Prediction}
+\title{Efficient Phenotype Prediction Using Incomplete Omics Data}
 
@@
 
+\section{Key message}
+Borrowing information from a highly performant `omics' predictor, sampled on a
+subset of inbred lines, can efficiently improve the prediction of complex
+traits.
+
 \begin{abstract}
 Predicting genotypic values with great precision is pivotal for effective
-candidate selections in animal and plant breeding.
+selection of candidates in animal and plant breeding.
 Novel `omics' predictors have shown to improve upon established genome-based
@@
 Single-step prediction has not yet been attempted with predictors calculated from
-quantitative features displaying heritabilities smaller than one.
+quantitative features displaying repeatabilities smaller than one.
 Using two maize datasets of inbred lines and hybrids we show that transcriptomic
@@
 ability exceeds that of pedigree or genomic data.
-Covering only a subset of genotypes with `omics' predictors and imputing all
-others using pedigree or genomic data could enable breeders to improve trait
-predictions while keeping costs under control.
+Our results suggest that covering only a subset of inbred lines with `omics'
+predictors and imputing all others using pedigree or genomic data could enable
+breeders to improve trait predictions while keeping costs under control.
 Employing `omics' predictors could particularly improve candidate selection in
@@
 
-We wish to submit the enclosed manuscript ``Omics-based Single Step Prediction''
-by Westhues et al.\@ for publication in TAG.
+We wish to submit the enclosed manuscript 
+``Efficient Phenotype Prediction Using Incomplete Omics Data'' by Westhues et 
+al.\@ for publication in TAG.
 
 Recent studies in maize inbred lines~\cite{Guo2016}, maize hybrids
-\cite{Zenke-Philippi2017,Westhues2017}, rice hybrids~\cite{Dan2016,Xu2016} and
-humans~\cite{Vazquez2016} have suggested that `omics' predictors, such as the
-metabolome and the transcriptome, can improve the prediction of traits with
-complex genetic architectures over that achieved with pedigree or genomic data.
+\cite{Zenke-Philippi2017,Westhues2017} and rice hybrids~\cite{Dan2016,Xu2016}
+have suggested that `omics' predictors, such as the metabolome and the
+transcriptome, can improve the prediction of traits with complex genetic 
+architectures over that achieved with pedigree or genomic data.
 Particularly hybrid breeding, where the success of forecasts increases
@@
 Currently, the adoption of these novel predictors by breeders is impeded
-by their higher sampling costs compared to pedigree and genomic data.
+by their higher costs compared to pedigree and genomic data.
 Here, we present single-step prediction of complex traits as a framework for
 combining pedigree and/or genomic data with other `omics' predictors collected
-on a subset of all available genotypes to boost predictive abilities while
-managing costs.
+on a subset of all available inbred lines to boost predictive abilities while
+controlling costs.
 
 Using two extensive phenotypic data sets from a public hybrid silage maize
-breeding program and a maize diversity panel, we a) investigate the success of
-imputing a quantitative predictor with feature heritabilities less than one and
-b) the influence of the fraction of genotypes covered by two predictors on
-predictive abilities in single step prediction.
+breeding program and a maize diversity panel published in TAG~\cite{Westhues2017,
+Guo2016}, we a) investigate the success of imputing a quantitative predictor
+with feature repeatabilities less than one and b) the influence of the fraction
+of inbred lines covered by two predictors on predictive abilities in prediction
+based on incomplete `omics' data.
 
@@
 \item In both data sets, single-step prediction improved predictive abilities
-	over those of the predictor covering all genotypes when the discrepancy 
-	between the	two predictors in terms of predictive ability were moderate to
-	large in favor of the incomplete predictor.
-\item When 20\%-30\% ($n \approx 25$) of the genotypes were covered by both
+	over those of the predictor covering all inbred lines when the
+	discrepancy between the	two predictors in terms of predictive ability
+	were moderate to large in favor of the incomplete predictor.
+\item When 20\%-30\% ($n \approx 25$) of the inbred lines were covered by both
 	predictors, single step-based predictive abilities converged to those
@@
 a highly performant predictor, such as transcriptomic data, on a subset of all
-genotypes.
+inbred lines.
 The framework presented herein should help promote the use of attractive
@@
 Its use is mainly motivated by the high costs and oftentimes time-consuming
-phenotyping, which can be reduced when all genotypes are covered by a predictor 
-while limiting phenotyping to a few genotypes with high precision
-\cite{Kadam2016}.
+phenotyping.
+Costs can be reduced when all genotypes, equivalent to inbred lines in plant
+breeding and animals in animal breeding, are covered by a predictor while
+limiting phenotyping to a few genotypes with high precision \cite{Kadam2016}.
 The performance of all other genotypes is then forecast using model parameters
@@
 \section{Materials and Methods}
-\Genetics2level{Exp1}
+\Genetics2level{Experiment 1}
 To address objectives (i) and (ii) a published data set of maize inbred lines
@@
 Best linear unbiased predictors (BLUPs) were calculated by~\citeN{Yang2014} for
-all 211 inbred lines and for 17 traits of which six are presented in this study 
-(Table~\ref{table:TraitKey}).
+all 211 inbred lines and for 17 traits of which six are presented in this study.
 The selection of these traits was done based on our prediction results so that
@@
 In summary, genomic and phenotypic information were available for 164 inbred
-lines, which we denote as the `Full' set whereas transcriptomic data were available
-for a subset of 110 inbred lines, which we denote as `Reduced'.
+lines, which we denote as the `Full' set, whereas transcriptomic data were
+available for a subset of 110 inbred lines, which we denote as `Reduced'.
 
@@
 
-\Genetics2level{Exp2}
+\Genetics2level{Experiment 2}
 To address objectives (i) and (iii), a maize hybrid data set, subsequently
@@
 agronomically important traits in silage maize by accounting for year, location,
-field replication, block and genotype effects as well as their interactions
-(Table~\ref{table:TraitKey}).
+field replication, block and genotype effects as well as their interactions.
 All 245 parent lines have pedigree records back to the generation of their 
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\centering
-\caption{Abbreviations for the trait names}
-\label{table:TraitKey}
-\begin{tabular}{@{}llll@{}}
-\toprule
-Trait              & Abbreviation & Data Set  & Experiment \\ \midrule
-Cob weight         & CW           & Diversity & Exp1       \\
-Days to silking    & DS           & Diversity & Exp1       \\
-Ear diameter       & ED           & Diversity & Exp1       \\
-100-grain weight   & GW           & Diversity & Exp1       \\
-Kernel width       & KW           & Diversity & Exp1       \\
-Plant height       & PH           & Diversity & Exp1       \\
-Dry matter yield   & DMY          & Hybrid    & Exp2       \\
-Dry matter content & DMC          & Hybrid    & Exp2       \\
-Lignin             & ADL          & Hybrid    & Exp2       \\
-Fat                & FAT          & Hybrid    & Exp2       \\
-Protein            & PRO          & Hybrid    & Exp2       \\
-Starch             & STA          & Hybrid    & Exp2       \\
-Sugar              & SUG          & Hybrid    & Exp2       \\ \bottomrule
-\end{tabular}
-\end{table}
-
 \Genetics2level{Prediction}
@@
 For the complete predictor we generated kernels by centering and standardizing
-all features in $\mathbf{M}$ to unit variance:
+all features in $\mathbf{M}$ to unit variance~\cite{VanRaden2008}:
 
 \begin{equation} \label{eq:GenomicRelationship}
-  \mathbf{K} = \frac{1}{M} \mathbf{M} \mathbf{M}^{\top},
+  \mathbf{K} = \frac{1}{p} \mathbf{M} \mathbf{M}^{\top}.
 \end{equation}
 
-where $M$ denotes the number of features~\cite{VanRaden2008}.
 In the case of pedigree data (P), the numerator relationship matrix was used
@@
 error, respectively.
-Here, we did not consider the inclusion of dominance effects for the hybrid
-material after a previous investigation into the predictive ability of various
-``omics'' predictors in the same material concluded no considerable gains 
-\cite{Westhues2017}.
 All prediction models were implemented using the R-packages \textit{BGLR}
@@
 If the breeding values are fully captured by the features in $\mathbf{M}$,
-Eq.~\ref{eq:KBLUPModel} can be expressed as~\cite{Fernando2014}
+Eq.(~\ref{eq:KBLUPModel}) can be expressed as~\cite{Fernando2014}
 
@@
 of that population has information for a second predictor.
-We can take a similar route as in \citeNP{Fernando2014} and impute covariates 
+We can take a similar route as in~\citeNP{Fernando2014} and impute covariates 
 of the incomplete predictor, denoted as `predictor' and indicated through the
@@
 where observed feature covariates are stored in $\tensor[]{\mathbf{W}}{^{*}_{2}}$
-while the feature covariates for genotypes not covered by the predictor can be estimated/imputed with the aid of the covariance matrix of the
-imputor $\mathbf{K}$ and the feature matrix of the predictor 
-$\tensor[]{\mathbf{W}}{^{*}_{2}}$ by calculating
+while the feature covariates for genotypes not covered by the predictor can be
+imputed with the aid of the covariance matrix of the imputor $\mathbf{K}$ and
+the feature matrix of the predictor $\tensor[]{\mathbf{W}}{^{*}_{2}}$ by
+calculating
 ($\tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} = \tensor[]{\mathbf{K}}{_{12}} \;
@@
 
-The structure of the residual imputation/prediction error is known, and
-therefore, can be modeled.
+The structure of the residual imputation error is known, and therefore, can be
+modeled.
 
@@
 The model above indicates that single-step prediction models ultimately use
-only the matrix of known and imputed feature covariates $\mathbf{W}$ for the
+only the matrix of known and imputed feature covariates $\mathbf{W}^{*}$ for the
 estimation of breeding values while accounting for the imputation error
@@
 \Genetics2level{Core sampling}
-In order to evaluate the influence of the fraction of genotypes that is not
-covered by the incomplete predictor on predictive ability we generated core
-sets of nine different sizes from Exp1 (Fig.~\ref{fig:compound-plot}~a).
+In order to evaluate how prediction accuracy is influenced by the fraction of
+genotypes not covered by the incomplete predictor on predictive ability we
+generated core sets of nine different sizes from Exp1
+(Fig.~\ref{fig:compound-plot}~a).
 The base of each of these core sets was the set of 110 genotypes that were
@@
 
+\Genetics2level{Data availability}
+The code for the analyses can be accessed from 
+\url{https:://github.com/mwesthues/single\_step} and
+\url{https://github.com/mwesthues/sspredr}.
+All data from Exp1 have been published before and are publicly available.
+Data for Exp2 can be made available upon request.
+
 \section*{Results}
-\Genetics2level{Exp1}
+\Genetics2level{Experiment 1}
 At the level of the 110 genotypes for which both, genomic (G) and 
@@
 In terms of predictive ability, the absolute advantage of T over G was as large 
-as 0.15, observed for the trait ED\@.
+as 0.15, observed for the trait ear diameter\@.
 The predictive ability for predictions based on G were always higher for the
@@
 ($n = 110$).
-The combination of G and T in a single step (\textit{i.e.}, $G|T^{*}$, where
+The combination of G and T in a single step (\textit{i.e.}, $T^{*}|G$, where
 $G$ is the imputor and $T^{*}$ is the predictor, for the prediction of all 164
@@
 reduced set of 110 lines.
-In the case of KW and ED, the use of $G|T^{*}$ yielded a slight improvement in
-predictive ability over G.
-Where T was a worse predictor for the reduced set of 110 lines than G,
-predictive abilities of $G|T^{*}$ for the full set of 164 lines were also worse
-than those based on G alone.
+In the case of kernel width and ear diameter, the use of $T^{*}|G$ yielded a slight 
+improvement in predictive ability over G.
+For the other four traits, where T was a worse predictor for the reduced set of
+110 lines than G, predictive abilities of $T^{*}|G$ for the full set of 164
+lines were also worse than those based on G alone.
 
@@
   As predictors, genomic (G) data, transcriptomic (T) data and their 
-  combination ($\tensor[_{C}]{G}{}\tensor[_{I}]{T}{}$) were used.
+  combination ($T^{*}|G$) were used.
   The `Reduced' panel includes a subset of 110 lines covered by genomic and
@@
 
-\Genetics2level{Exp2}
+\Genetics2level{Experiment 2}
 In the `Reduced' set of Exp2, transcriptomic data were the best single predictor
-for the traits ADL, PRO and particulaly for DMY where the predictive ability
-obtained with T was 0.08 points higher than for G as the second best predictor for
-this trait (Fig.~\ref{fig:HybridResults}).
-In the reduced set, P had the highest predictive ability for STA and DMC\@.
-For FAT and SUG, G was the superior predictor inside the reduced set.
+for the traits lignin, protein and particulaly for dry matter yield where the
+predictive ability obtained with T was 0.08 points higher than for G as the
+second best predictor for this trait (Fig.~\ref{fig:HybridResults}).
+In the reduced set, P had the highest predictive ability for starch and dry
+matter content\@.
+For fat and sugar, G was the superior predictor inside the reduced set.
 In the full set of 1,521 hybrids, predictive abilities obtained with G increased
-markedly for all traits with the exception of FAT\@.
+markedly for all traits with the exception of fat\@.
 The single-step combination of genomic and transriptomic information
-($G|T^{*}$) yielded a small improvement over G alone for the traits DMY and PRO\@.
-Except for PRO, compared to using only G, trait prediction did not benefit from 
-the single-step approach when T was not superior to G in the reduced set.
+($T^{*}|G$) yielded a small improvement over G alone for the traits dry matter
+yield and protein.\@.
+Compared to using only G, trait prediction did not benefit from the single-step
+approach when T was not superior to G in the reduced set.
+The only exception was protein.
 Predictive abilities resulting from combinations of pedigree with genomic
-($P|G^{*}$) information were better than P alone for all traits and $P|T^{*}$
+($G^{*}|P$) information were higher than P alone for all traits and $T^{*}|P$
 improved upon P in the `full' set for four traits.
+Here, we did not consider the inclusion of dominance effects for the hybrid
+material after a previous investigation into the predictive ability of various
+``omics'' predictors in the same material concluded no considerable gains 
+\cite{Westhues2017}.
 
@@
 varied widely within most traits (Fig.~\ref{fig:core-results}).
-For the majority of traits, an initial increase in predictive abilities when
-moving from a core set size of 10\% to 20\%/30\% was observed, regardless of
+For the majority of traits, an initial increase in predictive abilities was
+observed when moving from a core set size of 10\% to 20\%/30\%, regardless of
 whether G or T individually was the superior predictor.
-For the traits CW, GW and PH, which were predicted with higher precision via G
-than via T in the reduced set of lines, predictive abilities were higher when
-only a small number of genotypes was included in the core set.
+For the traits cob weight, grain weight and plant height, which were predicted
+with higher precision via G than via T in the reduced set of lines, predictive
+abilities were higher when only a small number of genotypes was included in the
+core set.
 This corresponds to a situation where a large number of genotypes is being 
@@
 A hike in predictive abilities, when moving from a core fraction of 10\% to 
-90\%, was observed for ED and KW, which were predicted with greater precision
-when using T individually compared to using G individually.
-The only trait that was barely affected by a change in the core fraction was DS
-for which the predictive abilities using either G or T individually were almost
-identical (Fig.~\ref{fig:InbredResults}).
+90\%, was observed for ear diameter and kernel width, which were predicted with
+greater precision when using T individually compared to using G individually.
+The only trait that was barely affected by a change in the core fraction was
+days until silking for which the predictive abilities using either G or T
+individually were almost identical (Fig.~\ref{fig:InbredResults}).
 
@@
 utility of downstream `omics' predictors for integrating such interactions.
-Recently, encouraging results have been found for humans~\cite{Vazquez2016}, 
-maize inbred lines~\cite{Guo2016} and maize hybrids~\cite{Westhues2017}.
+Recently, encouraging results with regard to predicting complex traits using
+`omics` data have been found for humans~\cite{Vazquez2016}, maize inbred lines
+\cite{Guo2016} and maize hybrids~\cite{Westhues2017}.
 While single-step prediction has become an established method when complete
@@
 such as transcriptomic, proteomic or metabolomic data.
+High correlations between the off-diagonal elements of the kernels for G and T
+and abundant interactions between distal SNPs and transcripts in hybrid maize
+\cite{Westhues2017} suggested that the substitution of G with T as a single-step
+predictor would be a logical next step.
 Here, we considered single-step approaches for a maize inbred line diversity
@@
 While the repeatability of relationship coefficients and genomic markers can be
-assumed to be equal to 1, this is not the case for the aforementioned
-quantitative `omics' predictors, which would negatively affect the precision of
-the imputation and was not considered here.
-
-\Genetics2level{Superiority of single-step models}
-\subsection{Exp1}
+assumed to be equal to 1, this would not be the case for the aforementioned
+quantitative `omics' predictors. 
+This would negatively affect the precision of the imputation when using them as
+imputors, which was not considered here.
+
+\Genetics2level{Superiority of single-step models in line prediction}
 When using a single-step approach based on complete genomic and incomplete
@@
 First, we reduced the dataset to 164 tropical/subtropical lines with little
-evident population structure and excluded all other genotypes from our analyses
-whereas \citeN{Guo2016} used genotypes from four different subpopulations with a 
-model using a fixed effect to account for the differences between subpopulations.
+evident population structure and excluded all other genotypes from our analyses.
+In contrast~\citeN{Guo2016} used genotypes from four different subpopulations
+with a model using a fixed effect to account for the differences between
+subpopulations.
 Second, we applied quality checks for the predictor data after generating the
@@
 
-\subsection{Exp2}
+\Genetics2level{Superiority of single-step models in hybrid prediction}
 Hybrid breeding is a particularly challenging application field for prediction
@@
 full set of genotypes compared to the reduced set for several agronomic traits.
-Hence, the inclusion of trancsriptomic data could not further improve upon
+Hence, the inclusion of transcriptomic data could not further improve upon
 genomic information alone.
@@
 heterotic groups~\cite{Martini2016}.
+While different studies have found that the low magnitude of SCA in hybrid maize
+breeding programs typically does not warrant its inclusion, this situation
+might be different in crops where heterotic patterns are not yet clearly
+defined~\cite{Zhao2015}.
 
@@
 of a core set of individuals with records on multiple predictors can effectively 
-be done at random once more than 10,000 animals have pedigree records 
+be done at random once more than 10,000 animals have predictor information
 \cite{Fragomeni2015,Lourenco2015,Masuda2016}, population sizes in plant breeding 
